---
title: "Transit Analysis"
author: "Kristin Mitchell"
format: html
editor: visual
toc: true
---

```{r message=FALSE, warning=FALSE}

library(psrc.travelsurvey)
library(magrittr)
library(data.table)
library(dplyr)
library(stringr)
library(psrcplot)
library(tidyverse)
library(openxlsx)

```

```{r}
#| echo: false

# Specify which variables to retrieve
vars <- c("mode_class", "mode_class_5", "transit_freq", "mode_acc", "commute_mode", "dest_purpose", "home_county", "home_rgcname", "home_jurisdiction", "dest_purpose_cat", "dest_purpose_cat_5", "hhincome_broad", "hhsize", "transportation_statement_transit", "commute_subsidy_transit", "commute_subsidy_parking", "commute_subsidy_1", "disability_person", "race_category", "mode_1", "distance_miles", "work_from_home", "age", "gender", "distance_miles", "vehicle_count")

survey_year <- c(
  "2017", "2019", "2021", "2023", 
  "2025")

# Retrieve the data
hts_data <- get_psrc_hts(survey_vars = vars, survey_year=survey_year)  # default includes all survey_years

```

```{r}
#| echo: false
## add additional variables

#add bucket age
hts_data$person <- mutate(
  hts_data$person,
  age_bucket=case_when(
    age%in%c('18-24 years', '25-34 years', '35-44 years', '45-54 years', '55-64 years') ~ '18-64 years',
    age%in%c('65-74 years', '75-84 years', '85 years or older') ~ '65 years and older'
       ))

#add bucket vehicles
hts_data$hh <- mutate(
  hts_data$hh,
  vehicles=case_when(
    vehicle_count%in%c('0 (no vehicles)') ~ '0 (no vehicles)',
    vehicle_count%in%c('1 vehicle') ~ '1 vehicle',
    vehicle_count%in%c('2 vehicles') ~ '2 vehicles',
    vehicle_count%in%c('3 vehicles', '4 vehicles', '5 vehicles', '6 vehicles', '7 vehicles', '8 vehicles', '9 vehicles', '10 or more vehicles') ~ '3 or more vehicles'
       ))

#add bucket gender
hts_data$person <- mutate(
  hts_data$person,
  gender_bucket=case_when(
    gender%in%c('Boy/Man (cisgender or transgender)', 'Male') ~ 'Boy/Man (cisgender or transgender)',
    gender%in%c('Girl/Woman (cisgender or transgender)', 'Female') ~ 'Girl/Woman (cisgender or transgender)',
    # gender%in%c('Non-Binary', 'Non-binary/Something else fits better') ~ 'Non-Binary',
    gender%in%c('Non-Binary', 'Non-binary/Something else fits better', 'Not listed here / prefer not to answer', 'Another') ~ 'Non-Binary/Another/Prefer Not to Answer'
       ))

#add bucket gender
hts_data$person <- mutate(
  hts_data$person,
  transit_statement_bucket=case_when(
    transportation_statement_transit%in%c('Strongly disagree', 'Disagree') ~ 'Disagree/Strongly Disagree',
    transportation_statement_transit%in%'Neither agree nor disagree' ~ 'Neither agree nor disagree',
    transportation_statement_transit%in%c('Strongly Agree', 'Agree') ~ 'Agree/Strongly Agree'
       ))

#add bucket frequency
hts_data$person <- mutate(
  hts_data$person,
  transit_freq_bucket=case_when(
    transit_freq %in%'6-7 days a week' ~ '6-7 days a week',
    transit_freq %in% '5 days a week' ~ '5 days a week',
    transit_freq%in%c('2 days a week', '3 days a week', '4 days a week', '2-4 days a week') 
    ~ '2-4 days a week',
    transit_freq%in%'1 day a week'~ '1 day a week',
    transit_freq%in%'1-3 days in the past month'~ '1-3 days in the past month',
    transit_freq%in%c('Never in the past 30 days', NA)~ 'Never in the past 30 days'
    #transit_freq%in%NA~ 'NA'
    ))

#add homegeog
hts_data$hh <- mutate(
  hts_data$hh,
  homegeog=case_when(
    home_rgcname=='Seattle Downtown'~ 'Seattle Downtown',
    (home_rgcname!='Seattle Downtown' & home_jurisdiction=='Seattle') ~ 'Seattle Outside Downtown',
    home_jurisdiction=='Bellevue' ~ 'Bellevue',
    ((home_county=='King' &!home_jurisdiction %in% c('Seattle', 'Bellevue'))|home_county=='Snohomish')~'Snohomish-King Suburban',
    .default=home_county))

# add home rgc variable
hts_data$hh <- mutate(
  hts_data$hh,
  rgchome=if_else((!is.na(home_rgcname) & home_rgcname!='Not RGC'),"Regional Growth Center", "Not Regional Growth Center"))

# add hyrbid variable

#add bucket frequency
hts_data$person <- mutate(
  hts_data$person,
  wfh_hybrid=case_when(
    work_from_home %in%'Yes, all of the time (100% of the time)' ~ 'Yes, all of the time (100% of the time)',
    work_from_home%in%c('Yes, most of the time (at least than 50% but less than 100% of the time)', 'Yes, some of the time (less than 50% of the time)') 
    ~ 'Hybird',
    work_from_home%in%'No'~ 'No'
    ))

## add two and three income buckets

hts_data$hh <- mutate(
  hts_data$hh,
  income_3group=factor(case_when(
        hhincome_broad %in% c("Under $25,000","$25,000-$49,999")~'Under $50,000',
        hhincome_broad %in% c("$50,000-$74,999", "$75,000-$99,999")~"$50,000-$99,999",
        hhincome_broad %in% c("$100,000 or more","$100,000-$199,999", "$200,000 or more")~"$100,000 or more",
        TRUE~hhincome_broad
                ),
      levels = c('Under $50,000',"$50,000-$99,999","$100,000 or more","Prefer not to answer")
      ))

hts_data$hh <- mutate(
  hts_data$hh,
 income_2group = factor(
      case_when(
        hhincome_broad %in% c("Under $25,000","$25,000-$49,999","$50,000-$74,999")~ "Under $75,000",
        hhincome_broad %in% c("$75,000-$99,999","$100,000 or more","$100,000-$199,999","$200,000 or more")~ "$75,000 or more",
        hhincome_broad == "Prefer not to answer"~ NA
                ),
      levels= c("Under $75,000","$75,000 or more")
      ))

hts_data$person <- mutate(
  hts_data$person,
  re_4group=factor(case_when(
        race_category %in% c("AANHPI non-Hispanic")~"AANHPI",
        race_category %in% c("Black or African American non-Hispanic")~"Black or African American",
        race_category %in% c("Some Other Race non-Hispanic", "Hispanic", "Two or More Races non-Hispanic")~"Another Person of Color",
        race_category %in% c("White non-Hispanic")~"White",
        TRUE~race_category
                ),
      levels = c("AANHPI","Black or African American", "Another Person of Color", "White")
      ))

hts_data$person <- mutate(
  hts_data$person,
  re_2group=factor(case_when(
        race_category %in% c("Black or African American non-Hispanic", "Hispanic", "Some Other Race non-Hispanic", "AANHPI non-Hispanic", "Two or More Races non-Hispanic")~"People of Color",
        race_category %in% c("White non-Hispanic")~"White",
        TRUE~race_category
                ),
      levels = c("People of Color", "White")
      ))

# set years as characters
hts_data$day$survey_year<-as.character(hts_data$day$survey_year)
hts_data$hh$survey_year<-as.character(hts_data$hh$survey_year)
hts_data$person$survey_year<-as.character(hts_data$person$survey_year)
hts_data$trip$survey_year<-as.character(hts_data$trip$survey_year)
hts_data$vehicle$survey_year<-as.character(hts_data$vehicle$survey_year)

```

## Overview

-   The Household Travel Survey asks various questions about transit use and transit access. This information can then be paired with other information including if someone lives in a regional growth center.

## Total trips

-   Total trips in the region have exceeded pre-pandemic numbers.

-   Walk trips peaked in 2021.

```{r}
#| echo: false

# Calculate a categorical summary, i.e. count/share
hts_data_mode <- hts_bin_mode(hts_data)

df <- psrc_hts_stat(hts_data, "trip", "mode_class_5", incl_na=FALSE) 

df$mode_class_5 <- factor(df$mode_class_5,
         levels = c("Bike/Micromobility", "Drive", "Other",
                    "Transit", "Walk"))

```

```{r}
#| echo: false

# Trips Count
df_count <- psrc_hts_stat(hts_data_mode, "trip", "mode_class_5", incl_na=FALSE) %>% group_by(survey_year) %>% summarise(total = sum(est))

trips_chart <- static_column_chart(df_count, x='survey_year', y='total', fill='survey_year',
                                     ylabel= 'Count',
                                     #xlabel='Mode',
                                     #moe='prop_moe',
                                     title='Total Trips in Region'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14), 
          axis.title.y=element_text(size=20), 
          axis.title.x=element_text(size=20)) 

trips_chart

```

```{r}
#| echo: false

mode_chart <- static_column_chart(df, x='mode_class_5', y='prop', fill='survey_year',
                                     ylabel= 'Share (%)',
                                     #xlabel='Mode',
                                     moe='prop_moe',
                                     title='Share Trips in Region'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14), 
          axis.title.y=element_text(size=20), 
          axis.title.x=element_text(size=20)) 

mode_chart
```

```{r}
#| echo: false
df_year <- df %>% filter(survey_year==2025)

mode_chart_year <- static_column_chart(df_year, x='mode_class_5', y='prop', fill='mode_class_5',
                                     ylabel= 'Share (%)',
                                     #xlabel='Mode',
                                     moe='prop_moe',
                                     title='Trips by Mode - Share (2025)'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14), 
          axis.title.y=element_text(size=20), 
          axis.title.x=element_text(size=20)) 

mode_chart_year

```

```{r}
df
```

## Transit trips

-   Transit trips have continued to increase post-pandemic, but have not reached pre-pandemic levels.

-   In 2025, there was a significant increase in transit trips by count and share compared to 2021 and 2023.

```{r}
#| echo: false
## transit trips count
df2 <- df %>% filter(mode_class_5=="Transit")

transit_chart_count <- static_column_chart(df2, x='survey_year', y='est', fill='survey_year',
                                     ylabel= 'Count',
                                     #xlabel='Mode',
                                     #moe='est_moe',
                                     title='Transit Trips - Count'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14), 
          axis.title.y=element_text(size=20), 
          axis.title.x=element_text(size=20)) 

transit_chart_count
```

```{r}
#| echo: false

transit_chart_count_err <- static_column_chart(df2, x='survey_year', y='est', fill='survey_year',
                                     ylabel= 'Count',
                                     #xlabel='Mode',
                                     moe='est_moe',
                                     title='Transit Trips - Count'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14), 
          axis.title.y=element_text(size=20), 
          axis.title.x=element_text(size=20)) 

transit_chart_count_err
```

```{r}
#| echo: false
## transit trips share
transit_chart <- static_column_chart(df2, x='survey_year', y='prop', fill='survey_year',
                                     ylabel= 'Share (%)',
                                     #xlabel='Mode',
                                     moe='prop_moe',
                                     title='Transit Trips - Share'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14), 
          axis.title.y=element_text(size=20), 
          axis.title.x=element_text(size=20)) 

transit_chart
```

```{r}
df2
```

```{r}
# #| echo: false
# ## walk trips count
# df3 <- df %>% filter(mode_class_5=="Walk")
# 
# walk_chart_count <- static_column_chart(df3, x='survey_year', y='est', fill='survey_year',
#                                      ylabel= 'Count',
#                                      #xlabel='Mode',
#                                      moe='est_moe',
#                                      title='Walk Trips - Count'
# ) + theme(axis.text.x=element_text(size=14),
#           axis.text.y=element_text(size=14),
#           legend.text = element_text(size=14), 
#           axis.title.y=element_text(size=20), 
#           axis.title.x=element_text(size=20)) 
# 
# walk_chart_count
```

```{r}
#| echo: false
# ## walk trips share
# walk_chart <- static_column_chart(df3, x='survey_year', y='prop', fill='survey_year',
#                                      ylabel= 'Share (%)',
#                                      #xlabel='Mode',
#                                      moe='prop_moe',
#                                      title='Walk Trips - Share'
# ) + theme(axis.text.x=element_text(size=14),
#           axis.text.y=element_text(size=14),
#           legend.text = element_text(size=14), 
#           axis.title.y=element_text(size=20), 
#           axis.title.x=element_text(size=20)) 
# 
# walk_chart
```

```{r}
# df3
```

```{r}
#| echo: false
## bike/micro trips count
# df4 <- df %>% filter(mode_class_5=="Bike/Micromobility")
# 
# bike_chart_count <- static_column_chart(df4, x='survey_year', y='est', fill='survey_year',
#                                      ylabel= 'Count',
#                                      #xlabel='Mode',
#                                      moe='est_moe',
#                                      title='Bike/Micromobility Trips - Count'
# ) + theme(axis.text.x=element_text(size=14),
#           axis.text.y=element_text(size=14),
#           legend.text = element_text(size=14), 
#           axis.title.y=element_text(size=20), 
#           axis.title.x=element_text(size=20)) 
# 
# bike_chart_count
```

```{r}
#| echo: false
## bike/micro trips share
# bike_chart <- static_column_chart(df4, x='survey_year', y='prop', fill='survey_year',
#                                      ylabel= 'Share (%)',
#                                      #xlabel='Mode',
#                                      moe='prop_moe',
#                                      title='Bike/Micromobility Trips - Share'
# ) + theme(axis.text.x=element_text(size=14),
#           axis.text.y=element_text(size=14),
#           legend.text = element_text(size=14), 
#           axis.title.y=element_text(size=20), 
#           axis.title.x=element_text(size=20)) 
# 
# bike_chart
```

```{r}
# df4
```

## Transit frequency

-   In 2025, there was a significant increase in transit use 2-4 days a week.

-   In 2025, there was a significant increase in transit use 1-3 days a week.

-   Transit frequency of 5 days a week has not returned to pre-pandemic levels.

-   In 2025, there was a significant decrease in the share of riders that answered they never in the past 30 days used transit. Historically, 75% to 84% of respondents said they have not used transit in the past 30 days. That dropped to 67% in 2025.

```{r}
#| echo: false

df5 <- psrc_hts_stat(hts_data, "trip", c("transit_freq_bucket"), incl_na=TRUE)

# df5.1 <- psrc_hts_stat(hts_data, "trip", c("transit_freq"), incl_na=TRUE)

df5$transit_freq_bucket <-
  factor(df5$transit_freq_bucket,
         levels = c("6-7 days a week", "5 days a week", "2-4 days a week", "1 day a week", "1-3 days in the past month", 
        "Never in the past 30 days"))

transit_freq_chart <- static_column_chart(df5, x='transit_freq_bucket', y='prop', fill='survey_year',ylabel= 'Share (%)',
                                  #xlabel='',
                                  title='Transit Use Frequency',
                                  #moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
        axis.text.y=element_text(size=14),
        legend.text = element_text(size=14), 
        axis.title.y=element_text(size=20), 
        axis.title.x=element_text(size=20))

transit_freq_chart

transit_freq_chart_err <- static_column_chart(df5, x='transit_freq_bucket', y='prop', fill='survey_year',ylabel= 'Share',
                                  #xlabel='',
                                  title='Transit Use Frequency',
                                  moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
        axis.text.y=element_text(size=14),
        legend.text = element_text(size=14), 
        axis.title.y=element_text(size=20), 
        axis.title.x=element_text(size=20))

transit_freq_chart_err

transit_freq_chart_count <- static_column_chart(df5, x='transit_freq_bucket', y='est', fill='survey_year',ylabel= 'Count',
                                  #xlabel='',
                                  title='Transit Use Frequency',
                                  #moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
        axis.text.y=element_text(size=14),
        legend.text = element_text(size=14),
        axis.title.y=element_text(size=20),
        axis.title.x=element_text(size=20))

transit_freq_chart_count
```

```{r}
df5
```

## Access to transit trips

-   Walked or jogged is by far the most dominant way people access transit.

-   The share of walked or jogged has significantly increased in 2023 and 2025 compared to 2019 and 2021. Accessing transit by vehicle has significantly decreased over the same time period.

```{r}
#| echo: false

# mode_acc to transit
hts_data2 <-  
  hts_bin_transit_mode_acc(hts_data)

df6 <- psrc_hts_stat(hts_data2, "trip", c("transit_mode_acc"), incl_na=FALSE) 

transit_acc_chart <- static_column_chart(df6, x='survey_year', y='prop', fill='transit_mode_acc',ylabel= 'Share (%)',
                                         title='Access to Transit Trips - Share', 
                                         moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
        axis.text.y=element_text(size=14),
        legend.text = element_text(size=14), 
        axis.title.y=element_text(size=20), 
        axis.title.x=element_text(size=20))

transit_acc_chart
```

```{r}
#| echo: false

df6.2025 <- df6 %>% filter(survey_year==2025)
transit_acc_chart_2025 <- static_column_chart(df6.2025, x='survey_year', y='prop', fill='transit_mode_acc',ylabel= 'Share (%)',
                                         title='Access to Transit Trips - Share (2025)'
) + theme(axis.text.x=element_text(size=14),
        axis.text.y=element_text(size=14),
        legend.text = element_text(size=14), 
        axis.title.y=element_text(size=20), 
        axis.title.x=element_text(size=20))

transit_acc_chart_2025
```

```{r}
#| echo: false
transit_acc_chart_count <- static_column_chart(df6, x='survey_year', y='est', fill='transit_mode_acc',ylabel= 'Count',
                                         title='Access to Transit Trips - Count',
                                  #moe='est_moe'
) + theme(axis.text.x=element_text(size=14),
        axis.text.y=element_text(size=14),
        legend.text = element_text(size=14), 
        axis.title.y=element_text(size=20), 
        axis.title.x=element_text(size=20))

transit_acc_chart_count
```

```{r}
#| echo: false
transit_acc_chart_count_err <- static_column_chart(df6, x='survey_year', y='est', fill='transit_mode_acc',ylabel= 'Count',
                    title='Access to Transit Trips - Count',
                                  moe='est_moe'
) + theme(axis.text.x=element_text(size=14),
        axis.text.y=element_text(size=14),
        legend.text = element_text(size=14), 
        axis.title.y=element_text(size=20), 
        axis.title.x=element_text(size=20))

transit_acc_chart_count_err
```

```{r}
df6
```

### Access to transit not including walk

-   Taking a closer look at accessing transit by bike/micromobility or a vehicle shows the significant decrease in accessing transit by vehicle in 2023 and 2025.

```{r}
#| echo: false

df6.1 <- df6 %>% filter(transit_mode_acc!="Walked or jogged")

transit_acc_chart_nowalk <- static_column_chart(df6.1, x='survey_year', y='prop', fill='transit_mode_acc',ylabel= 'Share (%)',
                                         title='Access to Transit Trips (No Walk) - Share', 
                                         moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14), 
          axis.title.y=element_text(size=20), 
          axis.title.x=element_text(size=20))

transit_acc_chart_nowalk

```

## Access to transit by destination purpose 

-   Destination purpose is only significantly different for work/school.

-   Share is higher for vehicular access to transit for work/school destination compared to walk/jog access for other destinations.

```{r}
#| echo: false

# mode_acc to transit
hts_data3 <-  
  hts_bin_transit_mode_acc(hts_data)

df7 <- psrc_hts_stat(hts_data3, "trip", c("dest_purpose_cat_5", "transit_mode_acc"), incl_na=FALSE) %>% 
#  filter(dest_purpose_cat_5=="Social/Recreation") %>%
  filter(transit_mode_acc=="Walked or jogged") %>% 
  filter(survey_year=="2025") 

dest_walk_chart <- static_column_chart(df7, x='survey_year', y='prop', fill='dest_purpose_cat_5',ylabel= 'Share (%)',
                                         title='Walk Access to Transit by Destination Purpose - Share', 
                                         moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14), 
          axis.title.y=element_text(size=20), 
          axis.title.x=element_text(size=20))

dest_walk_chart

```

```{r}
#| echo: false

df7.1 <- psrc_hts_stat(hts_data3, "trip", c("dest_purpose_cat_5", "transit_mode_acc"), incl_na=FALSE) %>% 
#  filter(dest_purpose_cat_5=="Social/Recreation") %>%
  filter(transit_mode_acc=="Vehicular") %>% 
  filter(survey_year=="2025") 

dest_veh_chart <- static_column_chart(df7.1, x='survey_year', y='prop', fill='dest_purpose_cat_5',ylabel= 'Share (%)',
                                         title='Vehicular Access to Transit by Destination Purpose - Share', 
                                         moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14), 
          axis.title.y=element_text(size=20), 
          axis.title.x=element_text(size=20))

dest_veh_chart
```

```{r}
df7.1
```

## Transit use by income

-   In 2025, members of households with an income under \$50,000 are significantly more likely to use transit compared to people in other income brackets.

-   In 2025, there was a significant increase from 2021 and 2023 in share of people in households with an income of \$50,000 to \$99,999 and \$100,000 or more using transit. 2025 was comparable to pre-pandemic transit use by people in these two income brackets.

```{r}
#| echo: false

df8 <- psrc_hts_stat(hts_data_mode, "trip", c("income_3group", "mode_class_5"), incl_na=FALSE) %>%
  filter(mode_class_5=="Transit") %>% 
  filter(survey_year=="2025") %>%
  filter(income_3group!="Prefer not to answer")

transit_income_chart <- static_bar_chart(df8, x='prop', y='income_3group', fill='income_3group',ylabel= 'Income',
                                         xlabel='Share (%)',
                    title='Transit Mode by Income - Share (2025)', 
                                         moe='prop_moe',
                                         color='pgnobgy_10'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14), 
          axis.title.y=element_text(size=20), 
          axis.title.x=element_text(size=20))

transit_income_chart
```

```{r}
df8
```

```{r}
#| echo: false

df9 <- psrc_hts_stat(hts_data_mode, "trip", c("income_3group", "mode_class_5"), incl_na=FALSE) 

df9.1 <- df9 %>%
  filter(mode_class_5=="Transit") %>% 
  filter(income_3group!="Prefer not to answer")

transit_income_chart_years <- static_column_chart(df9.1, x='survey_year', y='prop', fill='income_3group',ylabel= 'Share (%)',
                    title='Transit Mode by Income - Share', 
                                         moe='prop_moe',
                                         color='pgnobgy_10'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14), 
          axis.title.y=element_text(size=20), 
          axis.title.x=element_text(size=20))

transit_income_chart_years
```

```{r}
df9.1
```

## How people feel about being a transit rider

-   The 2025 Household Travel Survey included a new question about how people feel about being a transit rider.

-   Across all income brackets, people overwhelmingly agree or strongly agree with the statement, "I like the idea of public transit as a means of travel for me personally."

```{r}
#| echo: false
df10 <- psrc_hts_stat(hts_data_mode, "trip", c("income_3group", "mode_class_5", "transportation_statement_transit"), incl_na=FALSE) 

df10.2 <- df10 %>%
  filter(mode_class_5=="Transit") %>% 
  filter(income_3group!="Prefer not to answer") 
#%>%
  #filter(transportation_statement_transit=="2-Disagree") %>% 
  #filter(hhincome_broad=="Under $25,000")

income_transit_statement_chart <- static_column_chart(df10.2, x='transportation_statement_transit', y='prop', fill='income_3group',ylabel= 'Share (%)',
                    title='Level of Agreement with:',
                    subtitle='"I like the idea of public transit as a means of travel for me personally." by Income - Share (2025)',
                                         moe='prop_moe',
                                         color='pgnobgy_10'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14), 
          axis.title.y=element_text(size=20), 
          axis.title.x=element_text(size=20))

income_transit_statement_chart
```

```{r}
#| echo: false

df10.1 <- df10 %>%
  filter(mode_class_5=="Transit") %>% 
  filter(income_3group!="Prefer not to answer") %>%
  filter(transportation_statement_transit=="5-Strongly Agree") 
#%>% filter(hhincome_broad=="Under $25,000")

transit_statement_chart <- static_column_chart(df10.1, x='income_3group', y='prop', fill='income_3group',ylabel= 'Share (%)',
                    title='Strongly Agree with:',
                    subtitle='"I like the idea of public transit as a means of travel for me personally." by Income - Share (2025)',
                                         moe='prop_moe',
                                         color='pgnobgy_10'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14), 
          axis.title.y=element_text(size=20), 
          axis.title.x=element_text(size=20))

transit_statement_chart
```

```{r}
df10
```

## Employer offered transit passes

-   Over the years, people in households with an income \$100,000 or more have consistently been less likely to have an employer offer a transit pass.

```{r}
#| echo: false

# transit pass and parking analysis: "commute_subsidy_parking", "commute_subsidy_1", "commute_subsidy_transit"
df11 <- psrc_hts_stat(hts_data, "trip", c("income_3group", "commute_subsidy_transit"), incl_na=FALSE) 

df11.1 <- df11 %>%
  filter(income_3group!="Prefer not to answer") %>%
  filter(commute_subsidy_transit=="Not offered")
#%>% filter(hhincome_broad=="Under $25,000")
 
transit_pass_chart <- static_column_chart(df11.1, x='survey_year', y='prop', fill='income_3group',ylabel= 'Share (%)',
                    title='Transit Pass Not Offered - Share',
                                         moe='prop_moe',
                                         color='pgnobgy_10'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14), 
          axis.title.y=element_text(size=20), 
          axis.title.x=element_text(size=20))

transit_pass_chart
```

```{r}
df11
```

## Transit trips by race/ethnicity

-   People of color are more likely to use transit for trips.

-   Black or African American residents are more likely to use transit.

```{r}
#| echo: false

df12 <- psrc_hts_stat(hts_data_mode, "trip", c("re_2group", "mode_class_5"), incl_na=FALSE) %>% 
  filter(mode_class_5=="Transit") %>% 
  filter(survey_year=="2025")

transit_race_chart <- static_bar_chart(df12, x='prop', y='re_2group', fill='mode_class_5',xlabel= 'Share (%)',
                                         title='Transit Trips by R/E - 2025',
                                         moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14),
          axis.title.y=element_text(size=20),
          axis.title.x=element_text(size=20))

transit_race_chart

```

```{r}
df12.1 <- psrc_hts_stat(hts_data_mode, "trip", c("re_4group", "mode_class_5"), incl_na=FALSE) %>% 
  filter(mode_class_5=="Transit") %>% 
  filter(survey_year=="2025")

transit_race_chart_detailed <- static_bar_chart(df12.1, x='prop', y='re_4group', fill='mode_class_5',xlabel= 'Share (%)',
                                         title='Transit Trips by R/E - Share',
                                         moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14),
          axis.title.y=element_text(size=20),
          axis.title.x=element_text(size=20))

transit_race_chart_detailed

```

```{r}
df12.1
```

```{r}
#| echo: false

df12.2 <- psrc_hts_stat(hts_data_mode, "trip", c("re_4group", "mode_class_5"), incl_na=FALSE) %>% 
  filter(mode_class_5=="Transit") 

transit_race_chart_years <- static_column_chart(df12.2, x='survey_year', y='prop', fill='re_4group',ylabel= 'Share (%)',
                                         title='Transit Trips by R/E - Share',
                                         moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14),
          axis.title.y=element_text(size=20),
          axis.title.x=element_text(size=20))

transit_race_chart_years

transit_race_count_years_count <- static_column_chart(df12.1, x='survey_year', y='est', fill='re_4group',ylabel= 'Count',
                                         title='Transit Trips by R/E',
                                         moe='est_moe'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14),
          axis.title.y=element_text(size=20),
          axis.title.x=element_text(size=20))

transit_race_count_years_count
```

```{r}
df12.2
```

## Transit/mode share by disability status

-   The question about disability status was added to the Household Travel Survey in 2023.

-   People who do not have a disability increased their share of transit ridership from 2.4% to 3.7% from 2023 to 2025. This was a significant increase.

-   People with a disability use of the Other mode share jumped significantly from 2023 to 2025 from 1.6% to 7%. The Other mode includes paratransit and ride hail services such as Uber/Lyft, taxi, or car service.

-   Driving is the dominant mode for people with a disability.

```{r}
#| echo: false
hts_data5 <-
  hts_bin_mode(hts_data)

df13 <- psrc_hts_stat(hts_data5, "trip", c("disability_person", "mode_class_5"), incl_na=FALSE) %>% filter(mode_class_5=="Transit")
#%>% filter(survey_year=="2025")

transit_dis_chart <- static_column_chart(df13, x='survey_year', y='prop', fill='disability_person',xlabel= 'Share (%)',
                                         title='Transit Trips by Disability Status - Share',
                                         moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14),
          axis.title.y=element_text(size=20),
          axis.title.x=element_text(size=20))

transit_dis_chart

df13.1 <- psrc_hts_stat(hts_data5, "trip", c("disability_person", "mode_class_5"), incl_na=FALSE) %>% filter(mode_class_5=="Transit")

transit_dis_count <- static_column_chart(df13.1, x='survey_year', y='est', fill='disability_person',ylabel= 'Count',
                                         title='Transit Trips by Disability Status - Count',
                                         moe='est_moe'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14),
          axis.title.y=element_text(size=20),
          axis.title.x=element_text(size=20))

transit_dis_count

```

```{r}
#| echo: false

df14 <- psrc_hts_stat(hts_data_mode, "trip", c("disability_person", "mode_class_5"), incl_na=FALSE) %>% 
  filter(mode_class_5=="Other") %>% 
  #filter(survey_year=="2025") 
#%>%  
  filter(disability_person=="Yes")

other_dis_chart <- static_column_chart(df14, x='survey_year', y='prop', fill='mode_class_5',xlabel= 'Share (%)',
                                         title='Other Trips for People with a Disability - Share',
                                         moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14),
          axis.title.y=element_text(size=20),
          axis.title.x=element_text(size=20))

other_dis_chart

```

```{r}
df13
```

```{r}
#| echo: false

df14.1 <- psrc_hts_stat(hts_data5, "trip", c("disability_person", "mode_class_5"), incl_na=FALSE) %>% 
  #filter(survey_year=="2025") 
#%>%  
  filter(disability_person=="Yes")

mode_dis_chart <- static_column_chart(df14.1, x='survey_year', y='prop', fill='mode_class_5',xlabel= 'Share',
                                         title='Mode by Disability Status - Share (%)',
                                         moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14),
          axis.title.y=element_text(size=20),
          axis.title.x=element_text(size=20))

mode_dis_chart
```

```{r}
df14.1
```

## Access to transit by home geography

-   Walking to access transit does vary based on home geography, but what would be expected across the region. People in downtown Seattle are most likely to walk to access transit at 100%.

-   People living in Seattle but outside of downtown (94%) are comparable to Bellevue (91%) in walking to access transit.

```{r}
#| echo: false

#home geography

hts_data6 <-
  hts_bin_mode(hts_data)

df15 <- psrc_hts_stat(hts_data6, "trip", c("homegeog", "mode_acc"), incl_na=FALSE) %>% filter(survey_year=="2025") %>% filter(mode_acc=="Walked or jogged")

transit_acc_chart_detail <- static_bar_chart(df15, x='prop', y='homegeog', fill='survey_year',ylabel= 'Share (%)',
                                         #xlabel='Mode',
                                         moe='prop_moe',
                                         color='pgnobgy_10',
                                         title='Walked or Jogged Access to Transit by Home Geograhpy - Share'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14),
          axis.title.y=element_text(size=20),
          axis.title.x=element_text(size=20))

transit_acc_chart_detail
```

```{r}
df15
```

## Transit trips based on if home is in a regional growth center

-   In 2025, people who live in a regional growth center were significantly more likely to take transit (13.6%) compared to people who do not (2.8%).

-   In 2025, people who live in a regional growth center were significantly more likely to walk or jog to access transit (93.7%) compared to people who do not (81.4%).

```{r}
#| echo: false
# mode based on home rgc/not rgc

hts_data7 <-
  hts_bin_mode(hts_data)

df16 <- psrc_hts_stat(hts_data7, "trip", c("rgchome", "mode_class_5"), incl_na=FALSE) %>% filter(mode_class_5=="Transit")

transit_rgchome <- static_column_chart(df16, x='survey_year', y='prop', fill='rgchome',ylabel= 'Share (%)',
                                         #xlabel='Mode',
                                         moe='prop_moe',
                                         color='pgnobgy_10',
                                         title='Transit Trips by Home in RGCs - 2025'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14),
          axis.title.y=element_text(size=20),
          axis.title.x=element_text(size=20))

transit_rgchome
```

```{r}
df16
```

```{r}
#| echo: false
 
df17 <- psrc_hts_stat(hts_data, "trip", c("mode_class_5", "rgchome", "mode_acc"), incl_na=FALSE) %>% 
  #filter(survey_year=="2025") %>% 
  filter(mode_acc=="Walked or jogged") %>%
  filter(mode_class_5=="Transit")

transit_acc_rgchome <- static_column_chart(df17, x='survey_year', y='prop', fill='rgchome',ylabel= 'Share (%)',
                                         #xlabel='Mode',
                                         moe='prop_moe',
                                         color='pgnobgy_10',
                                         title='Access to Transit by Home in RGCs - 2025'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14),
          axis.title.y=element_text(size=20),
          axis.title.x=element_text(size=20))

transit_acc_rgchome
```

```{r}
df17
```

```{r}
#| echo: false
## access to transit trips without walk
df18 <- psrc_hts_stat(hts_data7, "trip", c("rgchome", "mode_acc"), incl_na=FALSE) %>% filter(survey_year=="2025") %>% filter(mode_acc!="Walked or jogged")

transit_acc_chart_wo_walk <- static_column_chart(df18, x='mode_acc', y='prop', fill='rgchome',ylabel= 'Share (%)',
                                         title='Access to Transit Trips - Share (no walk)',
                                         moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14),
          axis.title.y=element_text(size=20),
          axis.title.x=element_text(size=20))

transit_acc_chart_wo_walk

```

## Transit trips by county

-   In 2023, Pierce County saw a rebound in share of transit trips comparable to the 2017 share.

-   In 2025, King County saw a rebound in share of transit trips to pre-pandemic levels.

-   When factoring in the margin of error, trends are not seen for Kitsap or Snohomish counties.

```{r}
#| echo: false

# trip mode Share (%) by county

df19 <- psrc_hts_stat(hts_data7, "trip", c("home_county", "mode_class_5"), incl_na=FALSE) %>% filter(mode_class_5=="Transit")

mode_county <- static_column_chart(df19, x='survey_year', y='prop', fill='home_county',ylabel= 'Share (%)',
                                         title='Transit Trips by County - Share',
                                         moe='prop_moe'
) + theme(axis.text.x=element_text(size=14),
          axis.text.y=element_text(size=14),
          legend.text = element_text(size=14),
          axis.title.y=element_text(size=20),
          axis.title.x=element_text(size=20))

mode_county
```

```{r}
df19
```

## Transit trips by gender

-   In 2025, men/boys and women/girls were comparable on the share of trips they chose to use transit for, but non-binary/another/prefer not to answer was significantly more likely to choose transit for a trip.

```{r}
#| echo: false

df20 <- psrc_hts_stat(hts_data_mode, "trip", c("gender_bucket", "mode_class_5"), incl_na=FALSE) %>% filter(mode_class_5=="Transit") 

transit_gender_chart <- static_column_chart(df20, x='survey_year', y='prop', fill='gender_bucket',ylabel= 'Share', xlabel='Gender', title='Transit by Gender - Share', moe='prop_moe', color='pgnobgy_10' ) + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

transit_gender_chart

```

```{r}
df20
```

## Transit trips and trip miles

-   Due to lower counts for higher mile trips, limit takeaways to looking at zero to six miles and six to 12 miles compared to everything else.

-   There are meaningful trends to take away in general.

-   In 2023, when transit is the primary mode there was a decrease in length of trips from zero to six miles and increase of trips six to 12 miles.

-   Other than in 2023, when transit is the primary mode over 50% of transit trips are zero to six miles.

```{r}
#| echo: false

df21 <- psrc_hts_stat(hts_data_mode, "trip", c("mode_class_5", "distance_miles"), incl_na=FALSE) %>% filter(mode_class_5=="Transit") %>% #filter(survey_year=="2025") %>% 
  filter(distance_miles!='Exactly 0') %>% filter(distance_miles!='Missing')

transit_miles_chart <- static_column_chart(df21, x='distance_miles', y='prop', fill='survey_year',ylabel= 'Share', xlabel='Miles', title='Trip Length when Transit is Primary Mode - Share', moe='prop_moe', color='pgnobgy_10' ) + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

transit_miles_chart
```

```{r}
#| echo: false

df21.1 <- psrc_hts_stat(hts_data_mode, "trip", c("mode_class_5", "distance_miles"), incl_na=FALSE) %>% filter(mode_class_5=="Transit") %>% filter(survey_year=="2025") %>% filter(distance_miles!='Exactly 0') %>% filter(distance_miles!='Missing')

mode_miles_chart <- static_column_chart(df21.1, x='distance_miles', y='prop', 
                                        fill='survey_year',ylabel= 'Share', xlabel='Miles',                                           title='Trip Length when Transit is Primary Mode - Share (2025)', #moe='prop_moe', 
color='pgnobgy_10' ) + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20)) + facet_wrap(~ mode_class_5)

mode_miles_chart

```

```{r}
#| echo: false

hts_data8 <-hts_bin_transit_mode_acc(hts_data)

df21.2 <- psrc_hts_stat(hts_data8, "trip", c("transit_mode_acc", "distance_miles"), incl_na=FALSE) %>% 
  filter(survey_year=="2025") %>%
  filter(transit_mode_acc=="Walked or jogged") %>%
  filter(distance_miles!='Exactly 0') %>%
  filter(distance_miles!='Missing')

acc_miles_chart <- static_column_chart(df21.2, x='distance_miles', y='prop', fill='transit_mode_acc',ylabel= 'Share', xlabel='Miles', title='Mode Access to Transit by Miles - Share', moe='prop_moe', color='pgnobgy_10' ) + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20)) + facet_wrap(~ survey_year)

acc_miles_chart

df21.3 <- psrc_hts_stat(hts_data8, "trip", c("transit_mode_acc", "distance_miles", "rgchome"), incl_na=FALSE) %>% filter(transit_mode_acc=="Walked or jogged") %>% filter(survey_year=="2025") %>% filter(distance_miles!='Exactly 0') %>% filter(distance_miles!='Missing')

rgc_miles_walktotransit_chart <- static_column_chart(df21.3, x='distance_miles', y='prop', fill='rgchome',ylabel= 'Share', xlabel='Miles', title='Transit Access Mode Walk/Jog by Miles - Share', moe='prop_moe', color='pgnobgy_10' ) + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20)) + facet_wrap(~ survey_year)

rgc_miles_walktotransit_chart

```

```{r}
df21
```

```{r}
df21.2
```

## Trip miles by regional growth centers

-   People who live in a regional growth center take shorter trips (all modes) from zero to six miles compared to those who live outside of regional growth centers.

```{r}
#| echo: false

df21.4 <- psrc_hts_stat(hts_data8, "trip", c("rgchome", "distance_miles"), incl_na=FALSE) %>% #filter(survey_year=="2025") %>% 
  filter(distance_miles!='Exactly 0') %>% filter(distance_miles!='Missing')

rgc_miles_chart <- static_column_chart(df21.4, x='distance_miles', y='prop', fill='rgchome',ylabel= 'Share', xlabel='Miles', title='Distance in Miles per Trip by RGC - Share', moe='prop_moe', color='pgnobgy_10' ) + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20)) + facet_wrap(~ survey_year)

rgc_miles_chart

```

```{r}
df21.4
```

## Vehicle ownership and regional growth centers

-   People who live in a regional growth center are significantly more likely to live in a household without a vehicle.

```{r}
#| echo: false

 df22 <- psrc_hts_stat(hts_data, "hh", c("rgchome", "vehicles"), incl_na=FALSE) #%>% # filter(survey_year=="2025")

rgc_vehicle_chart <- static_column_chart(df22, x='vehicles', y='prop', fill='rgchome',ylabel= 'Share', xlabel='# of Vehicles', title='Vehicle Ownership by RGC - Share', moe='prop_moe', color='pgnobgy_10' ) + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20)) + facet_wrap(~ survey_year)

rgc_vehicle_chart

```

```{r}
#| echo: false

df22.1 <- psrc_hts_stat(hts_data, "hh", c("rgchome", "vehicles"), incl_na=FALSE) %>% filter(vehicles=="0 (no vehicles)")

rgc_novehicle_chart <- static_column_chart(df22.1, x='survey_year', y='prop', fill='rgchome',ylabel= 'Share', xlabel='Years', title='No Vehicle by RGC - Share', moe='prop_moe', color='pgnobgy_10' ) + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

rgc_novehicle_chart

```

```{r}
#| echo: false

df22.2 <- psrc_hts_stat(hts_data, "hh", c("rgchome", "vehicles"), incl_na=FALSE) %>% filter(vehicles=="1 vehicle")

rgc_onevehicle_chart <- static_column_chart(df22.2, x='survey_year', y='prop', fill='rgchome',ylabel= 'Share', xlabel='Years', title='One Vehicle by RGC - Share', moe='prop_moe', color='pgnobgy_10' ) + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

rgc_onevehicle_chart

```

```{r}
df22
```

## Zero vehicle and modes

-   People who live in a zero vehicle households walk as a primary mode of travel followed by transit.

```{r}
#| echo: false

df23 <- psrc_hts_stat(hts_data_mode, "trip", c("vehicles", "mode_class_5"), incl_na=FALSE) %>% filter(vehicles=="0 (no vehicles)")

transit_novehicle_chart <- static_column_chart(df23, x='survey_year', y='prop', fill='mode_class_5',ylabel= 'Share (%)', xlabel='Years', title='No Vehicle and Modes - Share', 
                                               #moe='prop_moe', 
                                               color='pgnobgy_10' ) + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

transit_novehicle_chart

transit_novehicle_chart_err <- static_column_chart(df23, x='survey_year', y='prop', fill='mode_class_5',ylabel= 'Share (%)', xlabel='Years', title='No Vehicle and Modes - Share', moe='prop_moe', color='pgnobgy_10' ) + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

transit_novehicle_chart_err

```

```{r}
df23
```

## Number of vehicles and transit riders

-   During the pandemic in 2021, 61% of transit riders live in a zero vehicle household.

-   In 2023, transit riders living in a household with one vehicle came back to riding transit and ridership was comparable to riders living in zero vehicle households.

-   In 2025, transit riders living in a household with two vehicles came back to riding transit comparable to riders living in zero or one vehicle households.

```{r}
#| echo: false

df23.1 <- psrc_hts_stat(hts_data_mode, "trip", c("mode_class_5", "vehicles"), incl_na=FALSE) %>% filter(mode_class_5=="Transit")

transit_vehicle_chart <- static_column_chart(df23.1, x='survey_year', y='prop', fill='vehicles',ylabel= 'Share', xlabel='Years', title='# of Vehicles for Transit Riders - Share', #moe='prop_moe',
        color='pgnobgy_10' ) + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

transit_vehicle_chart

df23.1 <- psrc_hts_stat(hts_data_mode, "trip", c("mode_class_5", "vehicles"), incl_na=FALSE) %>% filter(mode_class_5=="Transit")

transit_vehicle_chart_err <- static_column_chart(df23.1, x='survey_year', y='prop', fill='vehicles',ylabel= 'Share', xlabel='Years', title='# of Vehicles for Transit Riders - Share', moe='prop_moe', color='pgnobgy_10' ) + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

transit_vehicle_chart_err

transit_vehicle_chart_count <- static_column_chart(df23.1, x='survey_year', y='est', fill='vehicles',ylabel= 'Estimate', xlabel='Years', title='# of Vehicles for Transit Riders - Estimate', #moe='prop_moe', 
      color='pgnobgy_10' ) + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

transit_vehicle_chart_count

```

```{r}
df23.1
```
