---
title: "Gender by Telecommute Frequency"
author: "Christy"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r, message=FALSE, warning=FALSE}
source('gender_telecommute.R')
```

## About

**gender**:

- persons table
  
**telecommute_freq**: 

- persons table
- if workplace = fixed, varied, telecommutes some days


Not quite apples-to-apples comparison because the question for `telecommute_freq` was different across years.

- 2019: How frequently do you telecommute?
  - includes 'per month' answers
- 2021: How frequently did you telecommute in the past week?

Values for `Gender` are also different across years. 2021 introduced 'Non-Binary' and combined 'Not listed here / prefer not to answer'.

- 2019: Female, Male, Prefer not to answer, Another
- 2021: Female, Male, Not listed here / prefer not to answer, Non-Binary

Exclude 'per month' responses from 2019 survey because of the small sample size.

```{r data summaries by gender, eval=TRUE, include=FALSE}

vars_02 <- c('workplace', 'age_category', 'gender', 'worker', 'telecommute_freq')

# 2017/2019: How frequently do you telecommute? ----
wrkr_1719 <- gender_tcf_get_hhts(survey_a$survey, 'p', vars_02)

# summarise 2017/2019 and add telecommute frequency lookup
gen_tc_freq_1719 <- hhts_count(wrkr_1719, 
                               # spec_wgt = 'hh_weight_2017_2019_v2021_adult',
                               spec_wgt = 'hh_weight_2017_2019_adult',
                               group_vars = c('gender_group', 'tc_freq_group')) %>%
  mutate(period = survey_a$label)

gen_tc_freq_1719_simple <- hhts_count(wrkr_1719, 
                                      # spec_wgt = 'hh_weight_2017_2019_v2021_adult',
                                      spec_wgt = 'hh_weight_2017_2019_adult',
                                      group_vars = c('gender_group', 'tc_freq_group_simple')) %>%
  mutate(period = survey_a$label) %>% 
  rename(tc_freq_group = tc_freq_group_simple)

# 2021: How frequently did you telecommute in the past week? ----
wrkr_2021 <- gender_tcf_get_hhts(survey_b$survey, 'p', vars_02)
# browser()
# summarise 2021
gen_tc_freq_2021 <- hhts_count(wrkr_2021, 
                               group_vars = c('gender_group', 'telecommute_freq'), 
                               # spec_wgt = 'person_weight_2021_ABS_Panel_adult'
                               spec_wgt = 'person_adult_weight_2021'
                               ) %>% 
  mutate(period = survey_b$label) %>% 
  rename(tc_freq_group = telecommute_freq)

gen_tc_freq_2021_simple <- hhts_count(wrkr_2021, 
                                      # spec_wgt = 'person_weight_2021_ABS_Panel_adult',
                                      spec_wgt = 'person_adult_weight_2021',
                                      group_vars = c('gender_group', 'tc_freq_group_simple')) %>% 
  mutate(period = survey_b$label) %>% 
  rename(tc_freq_group = tc_freq_group_simple)

# 2017 single-year ----
wrkr_2017 <- gender_tcf_get_hhts(survey_c$survey, 'p', vars_02)

# summarise 2017 and add telecommute frequency lookup
gen_tc_freq_2017 <- hhts_count(wrkr_2017, 
                               # spec_wgt = 'hh_weight_2017_v2021',
                               spec_wgt = 'hh_weight_2017',
                               group_vars = c('gender_group', 'tc_freq_group')) %>%
  mutate(period = survey_c$label)

gen_tc_freq_2017_simple <- hhts_count(wrkr_2017, 
                                      # spec_wgt = 'hh_weight_2017_v2021',
                                      spec_wgt = 'hh_weight_2017',
                                      group_vars = c('gender_group', 'tc_freq_group_simple')) %>%
  mutate(period = survey_c$label) %>% 
  rename(tc_freq_group = tc_freq_group_simple)

# 2019 single-year ----
wrkr_2019 <- gender_tcf_get_hhts(survey_d$survey, 'p', vars_02)

# summarise 2019 and add telecommute frequency lookup
gen_tc_freq_2019 <- hhts_count(wrkr_2019, 
                               # spec_wgt = 'hh_weight_2019_v2021',
                               spec_wgt = 'hh_weight_2019',
                               group_vars = c('gender_group', 'tc_freq_group')) %>%
  mutate(period = survey_d$label)

gen_tc_freq_2019_simple <- hhts_count(wrkr_2019, 
                                      # spec_wgt = 'hh_weight_2019_v2021',
                                      spec_wgt = 'hh_weight_2019',
                                      group_vars = c('gender_group', 'tc_freq_group_simple')) %>%
  mutate(period = survey_d$label) %>% 
  rename(tc_freq_group = tc_freq_group_simple)


# merge dfs ----
gen_tc_freq <- bind_rows(gen_tc_freq_1719, gen_tc_freq_2021, gen_tc_freq_2017, gen_tc_freq_2019) %>% 
  mutate(period = as.factor(period))

gen_tc_freq_simple <- bind_rows(gen_tc_freq_1719_simple, gen_tc_freq_2021_simple, gen_tc_freq_2017_simple, gen_tc_freq_2019_simple) %>% 
  mutate(period = as.factor(period))

## plot ----

plot.title <- 'Gender and Telecommute Frequency'
plot.title.01 <- 'Telecommute Frequency'
plot.subtitle <- 'Times per week'
plot.caption.01 <- "\n*Includes 'Non-Binary' responses but unable to report separately because of small sample size"

plot.facet.labels <- c('Female' = 'Female', 'Male' = 'Male', 'Prefer not to answer / Another' = 'Prefer not to answer / Another*')                       

df <- gen_tc_freq %>% 
  filter(tc_freq_group != 'Total') %>% 
  mutate(upper = share + share_moe,
         lower = share - share_moe)

df_simple <- gen_tc_freq_simple %>% 
  filter(tc_freq_group != 'Total') %>% 
  mutate(upper = share + share_moe,
         lower = share - share_moe,
         tc_freq_group = factor(tc_freq_group, levels = c('Telecommuted at least once per week', 'Never / None')))


```

```{r data summaries combined gender, include=FALSE}

# 2017/2019 All Genders ----

gen_tc_freq_1719_tot <- hhts_count(wrkr_1719,
                               # spec_wgt = 'hh_weight_2017_2019_v2021_adult',
                               spec_wgt = 'hh_weight_2017_2019_adult',
                               group_vars = 'tc_freq_group') %>%
  mutate(period = survey_a$label)

gen_tc_freq_1719_simple_tot <- hhts_count(wrkr_1719,
                                      # spec_wgt = 'hh_weight_2017_2019_v2021_adult',
                                      spec_wgt = 'hh_weight_2017_2019_adult',
                                      group_vars = 'tc_freq_group_simple') %>%
  mutate(period = survey_a$label) %>%
  rename(tc_freq_group = tc_freq_group_simple)

# 2017 All Genders ----

gen_tc_freq_2017_tot <- hhts_count(wrkr_2017, 
                               # spec_wgt = 'hh_weight_2017_v2021',
                               spec_wgt = 'hh_weight_2017',
                               group_vars = 'tc_freq_group') %>%
  mutate(period = survey_c$label)

gen_tc_freq_2017_simple_tot <- hhts_count(wrkr_2017, 
                                      # spec_wgt = 'hh_weight_2017_v2021',
                                      spec_wgt = 'hh_weight_2017',
                                      group_vars = 'tc_freq_group_simple') %>%
  mutate(period = survey_c$label) %>% 
  rename(tc_freq_group = tc_freq_group_simple)

# 2019 All Genders ----

gen_tc_freq_2019_tot <- hhts_count(wrkr_2019, 
                               # spec_wgt = 'hh_weight_2019_v2021',
                               spec_wgt = 'hh_weight_2019',
                               group_vars = 'tc_freq_group') %>%
  mutate(period = survey_d$label)

gen_tc_freq_2019_simple_tot <- hhts_count(wrkr_2019, 
                                      # spec_wgt = 'hh_weight_2019_v2021',
                                      spec_wgt = 'hh_weight_2019',
                                      group_vars = 'tc_freq_group_simple') %>%
  mutate(period = survey_d$label) %>% 
  rename(tc_freq_group = tc_freq_group_simple)

# 2021 All Genders ----

gen_tc_freq_2021_tot <- hhts_count(wrkr_2021, 
                                   group_vars = 'telecommute_freq', 
                                   # spec_wgt = 'person_weight_2021_ABS_Panel_adult'
                                   spec_wgt = 'person_adult_weight_2021'
                                   ) %>% 
  mutate(period = survey_b$label) %>% 
  rename(tc_freq_group = telecommute_freq)

gen_tc_freq_2021_simple_tot <- hhts_count(wrkr_2021, 
                                      # spec_wgt = 'person_weight_2021_ABS_Panel_adult',
                                      spec_wgt = 'person_adult_weight_2021',
                                      group_vars = 'tc_freq_group_simple') %>% 
  mutate(period = survey_b$label) %>% 
  rename(tc_freq_group = tc_freq_group_simple)

gen_tc_freq_tot <- bind_rows(gen_tc_freq_1719_tot, gen_tc_freq_2021_tot, gen_tc_freq_2017_tot, gen_tc_freq_2019_tot) %>% 
  mutate(period = as.factor(period))

gen_tc_freq_simple_tot <- bind_rows(gen_tc_freq_1719_simple_tot, gen_tc_freq_2021_simple_tot, gen_tc_freq_2017_simple_tot, gen_tc_freq_2019_simple_tot) %>% 
  mutate(period = as.factor(period))

plot.title.tot <- 'Gender and Telecommute Frequency'
plot.subtitle.tot <- 'Times per week'
plot.caption.tot <- "\n*Includes 'Non-Binary' responses but unable to report separately because of small sample size"

df_tot <- gen_tc_freq_tot %>% 
  filter(tc_freq_group != 'Total') %>% 
  mutate(upper = share + share_moe,
         lower = share - share_moe)

df_simple_tot <- gen_tc_freq_simple_tot %>% 
  filter(tc_freq_group != 'Total') %>% 
  mutate(upper = share + share_moe,
         lower = share - share_moe,
         tc_freq_group = factor(tc_freq_group, levels = c('Telecommuted at least once per week', 'Never / None')))

# root <- "T:\\2022July\\christy"
# write.csv(df_simple_tot, file.path(root, "gtf_simple_tot.csv"))
```

## Plots

```{r plot simple 2017-19 2021 a, echo=FALSE}
df_simple %>%
  filter(survey %in% c('2017_2019', '2021')) %>%
  filter(tc_freq_group != 'Never / None') %>%
gender_tcf(plot.source.year = '2017/19, 2021', plot.caption = plot.caption.01, plot.title = plot.title, palette = "Paired", hjust = .5, vjust = 0, angle = 0, facet = TRUE)
```

There has been an increase of females telecommuting at least once per week in 2021 compared to pre-COVID years.

```{r plot 2017-19 2021 a, echo=FALSE}
df %>%
  filter(survey %in% c('2017_2019', '2021')) %>%
  filter(!(tc_freq_group %in% c('Never / None', 'A few times per month', 'Less than monthly'))) %>%
gender_tcf(plot.source.year = '2017/19, 2021', plot.caption = plot.caption.01, plot.title = plot.title, palette = "Paired", hjust = .5, vjust = 0, angle = 0, facet = TRUE)
```

While distribution of telecommuters are somewhat fairly distributed across the bins in 2021, there appears to be slightly more 1-2 day female telecommuters than before while for males the data skews towards more 5+ day telecommuting.

```{r plot simple 2017-19 2021, include=FALSE}
# filter df for just 2017/19 2021
df_simple %>% filter(survey %in% c('2017_2019', '2021')) %>% 
gender_tcf(plot.source.year = '2017/19, 2021', plot.caption = plot.caption.01, plot.title = plot.title, palette = "Paired", hjust = .5, vjust = 0, angle = 0, facet = TRUE)
```

```{r plot 2017-19 2021, include=FALSE}
# filter df for just 2017/19 2021
df %>% filter(survey %in% c('2017_2019', '2021')) %>% 
gender_tcf(plot.source.year = '2017/19, 2021', plot.caption = plot.caption.01, plot.title = plot.title, palette = "Paired", hjust = .5, vjust = 0, angle = 0, facet = TRUE)
```


### All Genders

```{r, echo=FALSE}
df_simple_tot %>% 
  filter(tc_freq_group != 'Never / None') %>%
  gender_tcf(plot.source.year = '2017, 2019, 2017/19, 2021', plot.caption = '', plot.title = plot.title.01, palette = "Set3", hjust = .5, vjust = 0, angle = 0, facet = FALSE)
```

For all workers, between 2017 and 2021 there is an increase in people telecommuting at least once per week.

```{r, echo=FALSE}
df_tot %>% 
  filter(!(tc_freq_group %in% c('Never / None', 'A few times per month', 'Less than monthly'))) %>%
  gender_tcf(plot.source.year = '2017, 2019, 2017/19, 2021', plot.caption = '', plot.title = plot.title.01, palette = "Set3", hjust = .5, vjust = 0, angle = 0, facet = FALSE)
```

For all workers, there is a definite increase in 3-4 day telecommuting in 2021 compared to pre-COVID years, but there appears to have been a larger proportion that has been telecommuting for 1-2 days.

```{r plot all years, echo=FALSE}
df %>% 
  filter(!(tc_freq_group %in% c('Never / None', 'A few times per month', 'Less than monthly'))) %>% 
  filter(gender_group != 'Prefer not to answer / Another') %>% 
  gender_tcf(plot.source.year = '2017, 2019, 2017/19, 2021', plot.caption = plot.caption.01, plot.title = plot.title.01, palette = "Set3", hjust = .5, vjust = 0, angle = 0, facet = TRUE)

```

Between 2017 and 2021, there was a sharp increase in females telecommuting 1-2 days and an increase of males telecommuting 5+ days in 2021.While there is a significant increase in females telecommuting 3-4 days, a larger proportion of females are commuting 1-2 days. 


## Literature

- https://www.intereconomics.eu/contents/year/2021/number/5/article/teleworking-a-curse-or-a-blessing-for-gender-equality-and-work-life-balance.html
  - the pandemic has hit women harder than men, and the risk of the reversal of women’s hard-won gains is real, the pandemic has also opened up new opportunities to correct structural gender inequalities
- https://www.theurbanist.org/2020/03/11/telecommuting-surged-transit-use-lagged-among-downtown-seattle-workers-in-2019/
