---
title: "Race by Simple Mode"
author: "Christy"
date: '2022-06-06'
output: html_document
---

```{r}
library(psrc.travelsurvey)
library(tidyverse)
library(ggiraph)

# global variables
survey_a <- list(survey = '2017_2019', label = '2017/2019')
survey_b <- list(survey = '2021', label = '2021')
survey_c <- list(survey = '2017', label = '2017')
survey_d <- list(survey = '2019', label = '2019')


gridline.color <- '#ededed'
background.color <- 'white'
plot.source <- ""
```

## Race Categories by Mode

```{r cleaning}
vars <- c('race_eth_broad', 'mode_simple')

rm_get_hhts <- function(survey, level, vars) {

  df <- get_hhts(survey = survey,
           level = level,
           vars = vars) %>% 
    mutate(race_eth_broad_group = case_when(!(race_eth_broad %in% c('White only, non-Hispanic/Latinx', 'No Answer', 'Under 18 years, not asked race')) ~ 'Persons of Color',
                                            !is.na(race_eth_broad) ~ race_eth_broad),
           race_eth_broad_group2 = case_when(!(race_eth_broad %in% c("White only, non-Hispanic/Latinx", "Asian only, non-Hispanic/Latinx")) ~ 'Other', 
                                               !is.na(race_eth_broad) ~ race_eth_broad),
           mode_simple_group = case_when((mode_simple %in% c('Bike', 'Walk', 'Other')) ~ 'Other',
                                         !is.na(mode_simple) ~ mode_simple)
           )
  
  df_sum <- hhts_count(df, 
                       group_vars = vars) %>% 
    mutate(period = survey)

  df_simple <- hhts_count(df,
                          group_vars = c('race_eth_broad_group', 'mode_simple')) %>%
    mutate(period = survey) %>%
    rename(race_eth_broad = race_eth_broad_group)

  df_nodrive <- df %>%
    filter(is.na(mode_simple)|mode_simple != 'Drive') %>%
    hhts_count(group_vars = vars) %>%
    mutate(period = survey)

  return(list(df=df, sum=df_sum, simple=df_simple, nodrive=df_nodrive))
}

```

```{r race, eval=TRUE}

# hh_race_category # combined_adult_weight
# mode_1 # Primary mode, combined_adult_trip_weight
# mode_simple

r_vars <- c('race_eth_broad', 'mode_simple')

# summarise 2019
rm1719 <- rm_get_hhts(survey_a$survey, 't', vars)

# summarise 2021
rm2021 <- rm_get_hhts(survey_b$survey, 't', vars)


## plot ----

plot.title.rm <- 'Race/Ethnicity by Primary Mode of Transportation'

rm_df <- bind_rows(rm1719$sum, rm2021$sum) %>%
  filter(is.na(mode_simple)|mode_simple != 'Total') %>%
  mutate(upper = share + share_moe,
         lower = share - share_moe)

rm_df_simple <- bind_rows(rm1719$simple, rm2021$simple) %>%
  filter(is.na(mode_simple)|mode_simple != 'Total') %>%
  mutate(upper = share + share_moe,
         lower = share - share_moe,
         race_eth_broad = factor(race_eth_broad, levels = c('Persons of Color', 'White only, non-Hispanic/Latinx', 'No Answer')))

rm_df_nodrive <- bind_rows(rm1719$nodrive, rm2021$nodrive) %>%
  filter(is.na(mode_simple)|mode_simple != 'Total') %>%
  mutate(upper = share + share_moe,
         lower = share - share_moe)
```

```{r race plot function}
my_theme <- function(hjust = .5, vjust = 1.2, angle = 45) {
    theme(axis.text.x = element_text(hjust = hjust,
                                     vjust = vjust,
                                     angle = angle,
                                     size = 6),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = 'bottom',
          legend.title = element_blank(),
          panel.background = element_rect(fill = background.color),
          panel.grid.major.y = element_line(color = gridline.color),
          panel.grid.minor.y = element_line(color = gridline.color),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          plot.caption = element_text(size=5),
          strip.text = element_text(size = 7)
          )
}

race_mode <- function(table, hjust = .5, vjust = 1.2, angle = 45, subtitle = NULL) {
  r <- ggplot(table,
              aes(x = mode_simple,
                  y = share,
                  fill = period,
                  tooltip = paste0(mode_simple, ' ', period, ': ', round(share*100, 0), '%')
              )) +
    geom_col_interactive(position = position_dodge(width=.9)) +
    geom_linerange(aes(ymin = lower,
                       ymax = upper),
                   position = position_dodge(width=.9)) +
    scale_fill_brewer(palette = "Paired") +
    scale_y_continuous(labels = scales::label_percent()) +
    scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 20)) +
    my_theme(hjust = hjust, vjust = vjust, angle = angle) +
    labs(x = NULL,
         y = NULL,
         subtitle = subtitle,
         title = plot.title.rm,
         caption = plot.source
         ) +
  facet_wrap(vars(race_eth_broad), labeller = labeller(race_eth_broad = label_wrap_gen(20)))
   
  girafe(ggobj = r)
}
```

```{r race plot all}
race_mode(rm_df, hjust = .5, vjust = .5, angle = 45)
```

```{r race plot simple}
race_mode(rm_df_simple)
```

```{r race plot no drive trips}
race_mode(rm_df_nodrive, hjust = .5, vjust = .5, angle = 45, subtitle = 'Excluding Driving Trips')
```

- Driving is the predominant primary mode of transportation across race and ethnicity.
- POC saw slightly more walking trips
- Black/African-American (non-Hispanic/Latinx) & Hispanic or Latinx had more driving trips while other groups saw decreases in driving between the two time periods.
- Transit trips for Asian (non-Hispanic/Latinx) dropped slightly while Black or African American (non-Hispanic/Latinx) saw a larger drop in transit trips.
