---
title: "Summaries 2019 v 2021"
author: "Christy"
date: "4/21/2022"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r global, warning=FALSE, message=FALSE, echo=FALSE}
library(psrc.travelsurvey)
library(tidyverse)
library(ggiraph)

# global variables
survey_a <- list(survey = '2017_2019', label = '2017/2019')
survey_b <- list(survey = '2021', label = '2021')
survey_c <- list(survey = '2017', label = '2017')
survey_d <- list(survey = '2019', label = '2019')


gridline.color <- '#ededed'
background.color <- 'white'
```


```{r, child='gender_telecommute.Rmd'}
# gender by telecommute frequency analysis
```


## Race Categories by Mode

```{r race, eval=TRUE}

# hh_race_category # combined_adult_weight
# mode_1 # Primary mode, combined_adult_trip_weight
# mode_simple

r_vars <- c('race_eth_broad', 'mode_simple')

# summarise 2019
r_m1_2019 <- get_hhts(survey = survey_a$survey,
                      level = 't',
                      vars = r_vars) %>% 
   mutate(race_eth_broad_group = case_when(!(race_eth_broad %in% c('White only, non-Hispanic/Latinx', 'No Answer', 'Under 18 years, not asked race')) ~ 'Persons of Color',
                                          !is.na(race_eth_broad) ~ race_eth_broad))

r_m1_sum_2019 <- hhts_count(r_m1_2019, 
                            group_vars = r_vars) %>% 
  mutate(period = survey_a$label)

r_m1_sum_2019_simple <- hhts_count(r_m1_2019, 
                            group_vars = c('race_eth_broad_group', 'mode_simple')) %>% 
  mutate(period = survey_a$label) %>% 
  rename(race_eth_broad = race_eth_broad_group)

r_m1_sum_2019_nodrive <- r_m1_2019 %>% 
  filter(is.na(mode_simple)|mode_simple != 'Drive') %>% 
  hhts_count(group_vars = r_vars) %>% 
  mutate(period = survey_a$label)

# summarise 2021
r_m1_2021 <- get_hhts(survey = survey_b$survey,
                      level = 't',
                      vars = r_vars) %>% 
   mutate(race_eth_broad_group = case_when(!(race_eth_broad %in% c('White only, non-Hispanic/Latinx', 'No Answer', 'Under 18 years, not asked race')) ~ 'Persons of Color',
                                          !is.na(race_eth_broad) ~ race_eth_broad))

r_m1_sum_2021 <- hhts_count(r_m1_2021, 
                           group_vars = r_vars) %>% 
  mutate(period = survey_b$label)

r_m1_sum_2021_simple <- hhts_count(r_m1_2021, 
                           group_vars = c('race_eth_broad_group', 'mode_simple')) %>% 
  mutate(period = survey_b$label) %>% 
  rename(race_eth_broad = race_eth_broad_group)

r_m1_sum_2021_nodrive <- r_m1_2021 %>% 
  filter(is.na(mode_simple)|mode_simple != 'Drive') %>% 
  hhts_count(r_m1_2021, 
             group_vars = r_vars) %>% 
  mutate(period = survey_b$label)

## plot ----

plot.title.rm <- 'Race/Ethnicity by Primary Mode of Transportation'

rm_df <- bind_rows(r_m1_sum_2019, r_m1_sum_2021) %>% 
  filter(is.na(mode_simple)|mode_simple != 'Total') %>% 
  mutate(upper = share + share_moe,
         lower = share - share_moe)

rm_df_simple <- bind_rows(r_m1_sum_2019_simple, r_m1_sum_2021_simple) %>% 
  filter(is.na(mode_simple)|mode_simple != 'Total') %>% 
  mutate(upper = share + share_moe,
         lower = share - share_moe,
         race_eth_broad = factor(race_eth_broad, levels = c('Persons of Color', 'White only, non-Hispanic/Latinx', 'No Answer')))

rm_df_nodrive <- bind_rows(r_m1_sum_2019_nodrive, r_m1_sum_2021_nodrive) %>% 
  filter(is.na(mode_simple)|mode_simple != 'Total') %>% 
  mutate(upper = share + share_moe,
         lower = share - share_moe)
```

```{r race plot function}
race_mode <- function(table, hjust = .5, vjust = 1.2, angle = 45, subtitle = NULL) {
  r <- ggplot(table,
              aes(x = mode_simple,
                  y = share,
                  fill = period,
                  tooltip = paste0(mode_simple, ' ', period, ': ', round(share*100, 0), '%')
              )) +
    geom_col_interactive(position = position_dodge(width=.9)) +
    geom_linerange(aes(ymin = lower,
                       ymax = upper),
                   position = position_dodge(width=.9)) +
    scale_fill_brewer(palette = "Paired") +
    scale_y_continuous(labels = scales::label_percent()) +
    scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 20)) +
    theme(axis.text.x = element_text(hjust = hjust,
                                     vjust = vjust,
                                     angle = angle,
                                     size = 6),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = 'bottom',
          legend.title = element_blank(),
          panel.background = element_rect(fill = background.color),
          panel.grid.major.y = element_line(color = gridline.color),
          panel.grid.minor.y = element_line(color = gridline.color),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          plot.caption = element_text(size=5),
          strip.text = element_text(size = 7)
          ) +
    labs(x = NULL,
         y = NULL,
         subtitle = subtitle,
         title = plot.title.rm,
         caption = plot.source
         ) +
  facet_wrap(vars(race_eth_broad), labeller = labeller(race_eth_broad = label_wrap_gen(20)))
   
  girafe(ggobj = r)
}
```

```{r race plot all}
race_mode(rm_df, hjust = .5, vjust = .5, angle = 45)
```

```{r race plot simple}
race_mode(rm_df_simple)
```

```{r race plot no drive trips}
race_mode(rm_df_nodrive, hjust = .5, vjust = .5, angle = 45, subtitle = 'Excluding Driving Trips')
```

- Driving is the predominant primary mode of transportation across race and ethnicity.
- POC saw slightly more walking trips
- Black/African-American (non-Hispanic/Latinx) & Hispanic or Latinx had more driving trips while other groups saw decreases in driving between the two time periods.
- Transit trips for Asian (non-Hispanic/Latinx) dropped slightly while Black or African American (non-Hispanic/Latinx) saw a larger drop in transit trips.


## Detailed Household Income by Trip Path Distance 

```{r income process, eval=FALSE}
t_vars <- c('mode_1', 'trip_path_distance')
inc_dist_2019 <- get_hhts(dyear = 2019,
         level = 't',
         vars = t_vars)

t <- inc_dist_2019 %>% 
  filter(trip_path_distance > 200)

t2 <- inc_dist_2019 %>% 
  filter(mode_1 == "Airplane or helicopter")

```

```{r income, eval=TRUE}
# hhincome_detailed # abs_hh_weight
# trip_path_distance # Bing-estimated trip distance (miles), combined_adult_trip_weight

stat <- 'median'
tpd <- 100

i_vars <- c('mode_1', 'hhincome_detailed', 'trip_path_distance')
inc_levels <- c("Under $10,000" ,"$10,000-$24,999","$25,000-$34,999","$35,000-$49,999",
                "$50,000-$74,999", "$75,000-$99,999", "$100,000-$149,999", "$150,000-$199,999",
                "$200,000-$249,999", "$250,000 or more", "Prefer not to answer", "Total")

inc_dist_2019 <- get_hhts(survey = survey_a$survey,
         level = 't',
         vars = i_vars) %>% 
  filter(mode_1 != 'Airplane or helicopter') %>% 
  filter(trip_path_distance < tpd)

inc_dist_sum_2019 <- hhts_median(inc_dist_2019,
                                stat_var = 'trip_path_distance', # numeric variable (distance)
                                spec_wgt = 'trip_weight_2017_2019_v2021_adult',
                                group_vars = c('hhincome_detailed')) # category (income, race, gender)

id_19 <- inc_dist_sum_2019 %>% 
  rename(m = ends_with(stat), moe = ends_with('moe')) %>% 
  mutate(hhincome_detailed = factor(hhincome_detailed, levels = inc_levels),
         lower = m - moe,
         upper = m + moe,
         year = survey_a$label)

inc_dist_2021 <- get_hhts(survey = survey_b$survey,
         level = 't',
         vars = i_vars) %>% 
  filter(mode_1 != 'Airplane or helicopter') %>% 
  filter(trip_path_distance < tpd)

inc_dist_sum_2021 <- hhts_median(inc_dist_2021,
                                stat_var = 'trip_path_distance',
                                spec_wgt = 'trip_weight_2021_ABS_Panel_adult',
                                group_vars = c('hhincome_detailed'))

id_21 <- inc_dist_sum_2021 %>% 
  rename(m = ends_with(stat), moe = ends_with('moe')) %>% 
  mutate(hhincome_detailed = factor(hhincome_detailed, levels = inc_levels),
         lower = m - moe,
         upper = m + moe,
         year = survey_b$label)

## plot ----

plot.title.itpd <- paste('Detailed Household Income by', str_to_title(stat), 'Trip Path Distance')
plot.subtitle.itpd <- paste('Trips less than', tpd, 'miles')

id_1921 <- bind_rows(id_19, id_21) %>% 
  mutate(year = as.factor(year))



i <- ggplot(id_1921, aes(x = hhincome_detailed, 
                         y = m, 
                         fill = year,
                         tooltip = paste0(year, ": ", round(m, 1), ' miles'))) +
  geom_col_interactive(position = position_dodge(width=.9)) +
  geom_linerange(aes(ymin = lower,
                    ymax = upper),
                 position = position_dodge(width=.9)) +
  scale_fill_brewer(palette = "Paired") +
  scale_y_continuous(labels = scales::label_comma()) +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1,
                                   size = 6),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        legend.title = element_blank(),
        legend.position = 'bottom',
        panel.background = element_rect(fill = background.color),
        panel.grid.major.y = element_line(color = gridline.color),
        panel.grid.minor.y = element_line(color = gridline.color),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.caption = element_text(size=5)) +
  labs(x = NULL,
       y = 'Miles',
       fill = NULL,
       title = plot.title.itpd,
       caption = plot.source,
       subtitle = plot.subtitle.itpd)

girafe(ggobj = i)

```

- Lower Household earnings ($34k and less) had trips with higher distances in 2021.
- Households with earnings between $35k and $150k or $200k+ saw a decrease in median trip distance.
