---
title: "telework_for_presentation"
author: "suzanne"
date: "2024-04-08"
output: html_document
---


```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
library(forcats)
source('../survey-23-preprocess.R')
```


```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad,home_rgcname,survey_year, hh_weight
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category,work_rgcname,
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, industry,office_available, commute_freq 
                from HHSurvey.v_persons_labels")
                

setDT(hh)
setDT(person)
```
#race_category needs to be added to the codebook

# to do: add Seattle, Bellevue geographies

## Read in Codebook
```{r pressure, echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```

## Set IDs as characters
TO DO: functionalize convert all ids to characters, or store as characters upfront
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]

person[, person_id := as.character(person_id)]




hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))

```

Variables that need grouping:
telework_time: use group_2, group_2_value: telework_time_broad
gender: gender_grp group_1
telecommute_freq: telecommute_freq_simple: group_1-- this variable is confusing right now, need some digging
dest_purpose: dest_purpose_simple, group_1
mode-> mode_simple, group_1

```{r}
center.lyr <- st_read_elmergeo('URBAN_CENTERS')
```




```{r}

variable_list<-rbind(
    variable_list,
    data.table(
      variable = c("mode_simple", 'dest_purpose_simple', 'telework_time_broad',  'telework_time_hr', 'gender_grp','commute_freq_simple', 'commute_freq_all','telecommute_freq_simple','telecommute_status', 'telecommute_status_broad', 'telecommute_freq_cond','seattle_bellevue'),
      is_checkbox = c(0,0,0,0,0,0,0,0,0,0,0,0,0),
      hh = c(0,0,0,0,0,0,0,0,0,0,0,1),
      person = c(0,0,0,0,1,1,1, 1,1,1,1,1,0),
      day = c(0,0,1,1,0,0,0,0, 0,0,0,0,1),
      trip = c(1,1, 0,0, 0,0, 0,0,0,0,0,0,1),
      vehicle = c(0,0,0,0,0,0,0,0,0,0,0,0,0),
      location = c(0,0,0,0,0,0,0,0,0,0,0,0,0),
      description = c("Mode Group", "Trip Purpose", "Telework Hours", "Telework Time Group", "Gender", 'Commute Frequency', 'Telecommute Frequency', 'Telecommute Status', 'Telecommute Status Broad','Telecommute Detail', 'Telecommute or not', 'Home City'),
      logic = c('mode aggregation', 'destination aggregation',"telework time aggregation","telework time aggregation", "gender group", 'commute frequency aggregation', 'commute frequency aggregation','telecommute frequency group', "telecommute status","telecommute status broad", 'telecommute frequency group detail', 'telework or not',  'home in Seattle or Bellevue'),
      data_type = c("integer/categorical", "integer/categorical","integer/categorical", "integer/categorical", "integer/categorical", "integer/categorical","integer/categorical", "integer/categorical","integer/categorical","integer/categorical", "integer/categorical","integer/categorical","integer/categorical"),
      shared_name = c("mode_simple", 'dest_purpose_simple', 'telework_time_broad', 'telework_time_hr','gender_grp', 'commute_freq_simple', 'commute_freq_all','telecommute_freq_simple', 'telecommute_status', 'telecommute_status_broad','telecommute_status_detail', 'telecommute_freq_cond', 'seattle_bellevue')
    )
  )

```
Preprocessing to handle new variables
```{r}
group_labels<-get_grouped_labels(group_id='group_2', group_name='telecommute_freq_cond')
value_labels<-add_values_code(group_name='telecommute_freq_cond')
person<-grp_to_tbl(tbl=person, ungrouped_name='telecommute_freq', grouped_name='telecommute_freq_cond')
```

```{r}
group_labels<-get_grouped_labels(group_id='group_3', group_name='telecommute_freq_broad')
value_labels<-add_values_code(group_name='telecommute_freq_broad')
person<-grp_to_tbl(tbl=person, ungrouped_name='telecommute_freq', grouped_name='telecommute_freq_broad')
```

```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='commute_freq_simple')
value_labels<-add_values_code(group_name='commute_freq_simple')
person<-grp_to_tbl(tbl=person, ungrouped_name='commute_freq', grouped_name='commute_freq_simple')
```


```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='telecommute_freq_simple')
value_labels<-add_values_code(group_name='telecommute_freq_simple')
person<-grp_to_tbl(tbl=person, ungrouped_name='telecommute_freq', grouped_name='telecommute_freq_simple')%>%
  mutate(telecommute_freq_simple=gsub("[\r\n]","", telecommute_freq_simple))# there's some weird characters in there i can't figure out how to remove
```



Adding more complex value definitions that require multiple fields

```{r}

value<- c(0,1,2,3)
label=c('Not Worker', 'Remote', 'Hybrid', 'Onsite')
tele_val_labels<-data.frame(value, label)
tele_val_labels<-tele_val_labels%>%mutate(variable='telecommute_status')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, tele_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  


```

#make this a function
```{r}

value<- c(0,1,2,3,4)
label=c('Not Worker', 'Remote', 'Mostly In Person', 'Mostly Remote', 'In Person')
tele_val_labels<-data.frame(value, label)
tele_val_labels<-tele_val_labels%>%mutate(variable='telecommute_status_broad')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, tele_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  


```

```{r}

value<- c(0,1,2,3, 4, 5)
label=c('Never or less than weekly', '1 days', '2 days', '3 days', '4 days', '5+ days')
val_labels<-data.frame(value, label)
val_labels<-tele_val_labels%>%mutate(variable='commute_freq_all')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, tele_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  

```


More detailed variable definitions
```{r}
person<-person%>%mutate(telecommute_status=
        case_when(workplace %in%c('Missing: Skip logic', "Missing: Skip Logic") 
                  ~ 'Not Worker',
                  (is.na(workplace)) ~ 'Not Worker',
                  (workplace=='At home (telecommute or self-employed with home office)' )
                  ~ 'Fully At Home',
                  (telecommute_freq_cond == '1+ days per week')
                  ~'Hybrid',
                  .default = 'Fully In Person'
        )
)

```




```{r}
person<-person%>%mutate(telecommute_status_broad=
        case_when(
                  (telecommute_freq_broad == '1-2 days')
                  ~'Mostly In Person',
                  (telecommute_freq_broad == '3+ days')
                  ~'Mostly At Home',
                  .default = telecommute_status
        )
)

```

How much are people commuting?
Combine At home with other workers

```{r}
person<-person%>%mutate(commute_freq_all=
                           if_else(workplace== "At home (telecommute or self-employed with home office)", 
                                   'Never or less than weekly', commute_freq))



person<- person%>%mutate(commute_freq_all=case_match(commute_freq_all,
                               "1-3 days a month" ~ "Never or less than weekly",
                               'Never' ~ "Never or less than weekly",
                                "A few times per month"   ~  "Never or less than weekly",  
                                "Less than monthly"   ~  "Never or less than weekly",  
                               "None, did not commute to workplace last week" ~"Never or less than weekly",
                                '5 days a week' ~ '5+ days',
                                '6-7 days a week' ~ '5+ days',
                               .default= commute_freq_all
                               ))


person<- person%>%mutate(commute_freq_all=str_replace(commute_freq_all, ' a week', ''))
                               
person<- person%>%mutate(commute_freq_all=case_when(
  commute_freq_all %in% c('1 day', '2 days') ~ '1-2 days',
  commute_freq_all %in% c('3 days', '4 days') ~ '3-4 days',
  .default = commute_freq_all
  
))  

person<- person%>%mutate(commute_freq_all=factor(commute_freq_all, levels=c("Never or less than weekly", '1-2 days', '3-4 days', '5+ days')))


```


```{r}
workers<-person%>%filter(!workplace %in% c('Missing: Skip logic', "Missing: Skip Logic"))%>%drop_na(workplace)


```



```{r}
variable_list<-add_variable(variable_list, 'home_rgc_cat','hh')
```

```{r}
variable_list<-add_variable(variable_list, 'work_rgc_cat','person')
```


```{r}

value<- c(0,1,2)
label=c('Urban', 'Metro','Not RGC')
val_labels<-data.frame(value, label)
val_labels<-val_labels%>%mutate(variable='home_rgc_cat')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  
```
```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```


```{r}
hh<-left_join(hh, center.lyr, by=join_by(home_rgcname==name))

```


we know the current home of people whether they moved in the past five years or no, gives us a bigger sample
```{r}
hh<-hh%>%mutate(home_rgc_cat=case_when(
  is.na(category) ~'Not RGC',
  category=='Not RGC' ~ 'Not RGC',
  TRUE ~ category))



```



```{r}

value<- c(0,1,2)
label=c('Urban', 'Metro','Not RGC')
val_labels<-data.frame(value, label)
val_labels<-val_labels%>%mutate(variable='work_rgc_cat')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  
```

```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```


```{r}
worker<-left_join(workers, center.lyr, by=join_by(work_rgcname==name))

```

```{r}
worker_hh<-left_join(worker, hh, by='hh_id', suffix=c('worker', ''))
```

we know the current home of people whether they moved in the past five years or no, gives us a bigger sample
```{r}
worker<-worker_hh%>%mutate(work_rgc_cat=case_when(
  telecommute_status=='Fully At Home' ~ home_rgc_cat,
  is.na(work_rgcname) ~'Not RGC',
  work_rgcname=='Not RGC' ~ 'Not RGC',
  TRUE ~ categoryworker))



```


```{r}
hts_data<-list( hh=hh, person=worker)

id_cols=c('hh_id', 'person_id')
wt_cols=c('hh_weight', 'person_weight')
```


```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```

```{r}

```

```{r}
hts_data<-list( hh=hh, person=worker)

id_cols=c('hh_id', 'person_id')
wt_cols=c('hh_weight', 'person_weight')
```


```{r}

telecommute_status <- summarize_weighted(hts_data=hts_data,
                               summarize_var = 'telecommute_status',
                               summarize_by = c('survey_year', 'work_rgc_cat'),
                               id_cols=id_cols,
                               wt_cols=wt_cols,
                               wtname= 'person_weight'
                               )

```

```{r}

telecommute_summary<-telecommute_status$summary$wtd%>%mutate(telecommute_status=forcats::fct_reorder(telecommute_status, -prop))
static<-static_facet_column_chart(telecommute_summary, y='prop', x='telecommute_status', fill='survey_year', facet='work_rgc_cat', xlabel='Work Location Status', ylabel='Worker Share', scale='fixed') 
static

write.csv(telecommute_summary,'telecommute_summary_trend_rgcwork.csv')
```
```{r}

telecommute_status_home <- summarize_weighted(hts_data=hts_data,
                               summarize_var = 'telecommute_status',
                               summarize_by = c('survey_year', 'home_rgc_cat'),
                               id_cols=id_cols,
                               wt_cols=wt_cols,
                               wtname= 'person_weight'
                               )

```

```{r}

telecommute_summary_home<-telecommute_status_home$summary$wtd%>%mutate(telecommute_status=forcats::fct_reorder(telecommute_status, -prop))
static<-static_facet_column_chart(telecommute_summary_home, y='prop', x='telecommute_status', fill='survey_year', facet='home_rgc_cat', xlabel='Work Location Status', ylabel='Worker Share', scale='fixed') 

static

write.csv(telecommute_summary_home,'telecommute_summary_trend_rgc_home.csv')
```


                                                                                                                                  






