---
title: "telework_for_presentation"
author: "suzanne"
date: "2024-04-08"
output: html_document
---


```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
library(forcats)
source('../survey-23-preprocess.R')
```


```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad,home_jurisdiction,survey_year, hh_weight
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category,
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, industry,office_available, commute_freq 
                from HHSurvey.v_persons_labels")
                
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,
person_id, mode_1, dest_purpose, survey_year, trip_weight 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)
setDT(day)
setDT(trip)
```
#race_category needs to be added to the codebook

# to do: add Seattle, Bellevue geographies

## Read in Codebook
```{r pressure, echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```

## Set IDs as characters
TO DO: functionalize convert all ids to characters, or store as characters upfront
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
day[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
day[, person_id := as.character(person_id)]
trip[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
trip[, day_id := as.character(day_id)]

trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
day <- day%>%mutate(survey_year=as.character(survey_year))
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```

Variables that need grouping:
telework_time: use group_2, group_2_value: telework_time_broad
gender: gender_grp group_1
telecommute_freq: telecommute_freq_simple: group_1-- this variable is confusing right now, need some digging
dest_purpose: dest_purpose_simple, group_1
mode-> mode_simple, group_1


```{r}
hh_person_seattle<-merge(hh,person, by='hh_id')%>%filter(home_jurisdiction=='Seattle')
person<-hh_person_seattle
```


```{r}

variable_list<-rbind(
    variable_list,
    data.table(
      variable = c("mode_simple", 'dest_purpose_simple', 'telework_time_broad',  'telework_time_hr', 'gender_grp','commute_freq_simple', 'commute_freq_all','telecommute_freq_simple','telecommute_status', 'telecommute_status_broad', 'telecommute_freq_cond','seattle_bellevue'),
      is_checkbox = c(0,0,0,0,0,0,0,0,0,0,0,0,0),
      hh = c(0,0,0,0,0,0,0,0,0,0,0,1),
      person = c(0,0,0,0,1,1,1, 1,1,1,1,1,0),
      day = c(0,0,1,1,0,0,0,0, 0,0,0,0,1),
      trip = c(1,1, 0,0, 0,0, 0,0,0,0,0,0,1),
      vehicle = c(0,0,0,0,0,0,0,0,0,0,0,0,0),
      location = c(0,0,0,0,0,0,0,0,0,0,0,0,0),
      description = c("Mode Group", "Trip Purpose", "Telework Hours", "Telework Time Group", "Gender", 'Commute Frequency', 'Telecommute Frequency', 'Telecommute Status', 'Telecommute Status Broad','Telecommute Detail', 'Telecommute or not', 'Home City'),
      logic = c('mode aggregation', 'destination aggregation',"telework time aggregation","telework time aggregation", "gender group", 'commute frequency aggregation', 'commute frequency aggregation','telecommute frequency group', "telecommute status","telecommute status broad", 'telecommute frequency group detail', 'telework or not',  'home in Seattle or Bellevue'),
      data_type = c("integer/categorical", "integer/categorical","integer/categorical", "integer/categorical", "integer/categorical", "integer/categorical","integer/categorical", "integer/categorical","integer/categorical","integer/categorical", "integer/categorical","integer/categorical","integer/categorical"),
      shared_name = c("mode_simple", 'dest_purpose_simple', 'telework_time_broad', 'telework_time_hr','gender_grp', 'commute_freq_simple', 'commute_freq_all','telecommute_freq_simple', 'telecommute_status', 'telecommute_status_broad','telecommute_status_detail', 'telecommute_freq_cond', 'seattle_bellevue')
    )
  )

```
Preprocessing to handle new variables
```{r}
group_labels<-get_grouped_labels(group_id='group_2', group_name='telecommute_freq_cond')
value_labels<-add_values_code(group_name='telecommute_freq_cond')
person<-grp_to_tbl(tbl=person, ungrouped_name='telecommute_freq', grouped_name='telecommute_freq_cond')
```

```{r}
group_labels<-get_grouped_labels(group_id='group_3', group_name='telecommute_freq_broad')
value_labels<-add_values_code(group_name='telecommute_freq_broad')
person<-grp_to_tbl(tbl=person, ungrouped_name='telecommute_freq', grouped_name='telecommute_freq_broad')
```

```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='commute_freq_simple')
value_labels<-add_values_code(group_name='commute_freq_simple')
person<-grp_to_tbl(tbl=person, ungrouped_name='commute_freq', grouped_name='commute_freq_simple')
```


```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='telecommute_freq_simple')
value_labels<-add_values_code(group_name='telecommute_freq_simple')
person<-grp_to_tbl(tbl=person, ungrouped_name='telecommute_freq', grouped_name='telecommute_freq_simple')%>%
  mutate(telecommute_freq_simple=gsub("[\r\n]","", telecommute_freq_simple))# there's some weird characters in there i can't figure out how to remove
```

```{r}
group_labels<-get_grouped_labels(group_id='group_2', group_name='telework_time_broad')
value_labels<-add_values_code(group_name='telework_time_broad')
day<-grp_to_tbl(tbl=day, ungrouped_name='telework_time', grouped_name='telework_time_broad')
```

```{r}
group_labels<-get_grouped_labels(group_id='group_3', group_name='telework_time_23')
value_labels<-add_values_code(group_name='telework_time_23')
day<-grp_to_tbl(tbl=day, ungrouped_name='telework_time', grouped_name='telework_time_23')
```


```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='telework_time_hr')
value_labels<-add_values_code(group_name='telework_time_hr')
day<-grp_to_tbl(tbl=day, ungrouped_name='telework_time', grouped_name='telework_time_hr')
```

Adding more complex value definitions that require multiple fields

```{r}

value<- c(0,1,2,3)
label=c('Not Worker', 'Remote', 'Hybrid', 'Onsite')
tele_val_labels<-data.frame(value, label)
tele_val_labels<-tele_val_labels%>%mutate(variable='telecommute_status')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, tele_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  


```

#make this a function
```{r}

value<- c(0,1,2,3,4)
label=c('Not Worker', 'Remote', 'Mostly In Person', 'Mostly Remote', 'In Person')
tele_val_labels<-data.frame(value, label)
tele_val_labels<-tele_val_labels%>%mutate(variable='telecommute_status_broad')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, tele_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  


```

```{r}

value<- c(0,1,2,3, 4, 5)
label=c('Never or less than weekly', '1 days', '2 days', '3 days', '4 days', '5+ days')
val_labels<-data.frame(value, label)
val_labels<-tele_val_labels%>%mutate(variable='commute_freq_all')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, tele_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  

```


More detailed variable definitions
```{r}
person<-person%>%mutate(telecommute_status=
        case_when(workplace %in%c('Missing: Skip logic', "Missing: Skip Logic") 
                  ~ 'Not Worker',
                  (is.na(workplace)) ~ 'Not Worker',
                  (workplace=='At home (telecommute or self-employed with home office)' )
                  ~ 'Fully At Home',
                  (telecommute_freq_cond == '1+ days per week')
                  ~'Hybrid',
                  .default = 'Fully In Person'
        )
)

```




```{r}
person<-person%>%mutate(telecommute_status_broad=
        case_when(
                  (telecommute_freq_broad == '1-2 days')
                  ~'Mostly In Person',
                  (telecommute_freq_broad == '3+ days')
                  ~'Mostly At Home',
                  .default = telecommute_status
        )
)

```

How much are people commuting?
Combine At home with other workers

```{r}
person<-person%>%mutate(commute_freq_all=
                           if_else(workplace== "At home (telecommute or self-employed with home office)", 
                                   'Never or less than weekly', commute_freq))



person<- person%>%mutate(commute_freq_all=case_match(commute_freq_all,
                               "1-3 days a month" ~ "Never or less than weekly",
                               'Never' ~ "Never or less than weekly",
                                "A few times per month"   ~  "Never or less than weekly",  
                                "Less than monthly"   ~  "Never or less than weekly",  
                               "None, did not commute to workplace last week" ~"Never or less than weekly",
                                '5 days a week' ~ '5+ days',
                                '6-7 days a week' ~ '5+ days',
                               .default= commute_freq_all
                               ))


person<- person%>%mutate(commute_freq_all=str_replace(commute_freq_all, ' a week', ''))
                               
person<- person%>%mutate(commute_freq_all=case_when(
  commute_freq_all %in% c('1 day', '2 days') ~ '1-2 days',
  commute_freq_all %in% c('3 days', '4 days') ~ '3-4 days',
  .default = commute_freq_all
  
))  

#person<- person%>%mutate(commute_freq_all=factor(commute_freq_all, levels=c("Never or less than #weekly", '1-2 days', '3-4 days', '5+ days')))


```


```{r}
workers<-person%>%filter(!workplace %in% c('Missing: Skip logic', "Missing: Skip Logic"))%>%drop_na(workplace)%>%rename(survey_year=survey_year.x)
worker_list<-list(person=workers)

```





```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```



```{r}

telecommute_status <- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_status',
                               summarize_by = 'survey_year',
                               id_cols=c('person_id'),
                               wt_cols=c('person_weight'),
                               wtname= 'person_weight'
                               )

```

```{r}

telecommute_summary<-telecommute_status$summary$wtd%>%mutate(telecommute_status=forcats::fct_reorder(telecommute_status, -prop))
static<-static_column_chart(telecommute_summary, y='prop', x='telecommute_status', fill='survey_year', xlabel='Work Location Status', ylabel='Worker Share') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

inter<-interactive_column_chart(telecommute_summary, y='prop', x='telecommute_status', fill='survey_year', xlabel='Work Location Status', ylabel='Worker Share')
static
inter
write.csv(telecommute_summary,'telecommute_summary_trend.csv')
```

```{r}

telecommute_status_2 <- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_status_broad',
                               summarize_by = 'survey_year',
                               id_cols=c('person_id'),
                               wt_cols=c('person_weight'),
                               wtname= 'person_weight'
                               )

```

```{r}

telecommute_summary_2<-telecommute_status_2$summary$wtd%>%mutate(telecommute_status_broad=factor(telecommute_status_broad, levels=c('Fully In Person', 'Mostly In Person', 'Mostly At Home', 'Fully At Home')))
static_column_chart(telecommute_summary_2, y='prop', x='telecommute_status_broad', fill='survey_year', xlabel='Work Location Status', ylabel='Worker Share') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

```






```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```

Summarize Workplace variable
```{r}
workers<-person%>%filter(!workplace %in% c('Missing: Skip logic', "Missing: Skip Logic"))%>%drop_na(workplace)%>%rename(survey_year=survey_year.x)
worker_list<-list(person=workers)

```



# regional
```{r}

workplace <- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'workplace',
                               summarize_by = 'survey_year',
                               id_cols=c('person_id'),
                               wt_cols=c('person_weight'),
                               wtname= 'person_weight'
                               )

```
## Visualize


```{r}
workplace_summary_23<-workplace$summary$wtd%>%filter(survey_year %in% c( '2023'))

workplace_summary_23<-workplace_summary_23%>%
  mutate(workplace=case_match(workplace,  
         "Usually the same location (outside home)"  ~ "One outside location",
         "Telework some days and travel to a work location some days" ~ "Hybrid telework",
        "Workplace regularly varies (different offices or jobsites)" ~ "Variable outside locations",
         "At home (telecommute or self-employed with home office)" ~ "At home",
          "Drives for a living (e.g., bus driver, salesperson)"   ~ "Drives for work"
         
         ))%>%
  mutate(workplace=forcats::fct_reorder(workplace, -prop))

static_bar_chart(workplace_summary_23, x='prop', y='workplace', fill='survey_year')+ theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=20))

```

```{r}
workplace_summary_21_23<-workplace$summary$wtd%>%filter(survey_year %in% c('2021', '2023'))

workplace_summary_21_23<-workplace_summary_21_23%>%
  mutate(workplace=forcats::fct_reorder(workplace, -prop))

static_bar_chart(workplace_summary_21_23, x='prop', y='workplace', fill='survey_year')+ theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=20))

```



```{r}

workplace_summary<-workplace$summary$wtd%>%filter(workplace== 'At home (telecommute or self-employed with home office)')
static_column_chart(workplace_summary, y='prop', x='workplace', fill='survey_year' )+ theme(axis.text.x=element_text(size=20), legend.text = element_text(size=20))

```


```{r}
workers<-person%>%filter(!workplace %in% c('Missing: Skip logic', "Missing: Skip Logic"))%>%drop_na(workplace)%>%filter(commute_freq_all!= 'Missing: Skip Logic')%>%drop_na(commute_freq_all)%>%rename(survey_year=survey_year.x)
worker_list<-list(person=workers)

```



```{r}

commute_freq_all <- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'commute_freq_all',
                               summarize_by = 'survey_year',
                               id_cols=c('person_id'),
                               wt_cols=c('person_weight'),
                               wtname= 'person_weight'
                               )

```

```{r}

commute_summary<-commute_freq_all$summary$wtd
static<-static_column_chart(commute_summary, y='prop', x='commute_freq_all', fill='survey_year', ylabel= 'Worker Share', xlabel='Days Commuting per Week') + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=16), axis.title.x=element_text(size=16))

inter<-interactive_column_chart(commute_summary, y='prop', x='commute_freq_all', fill='survey_year', xlabel='Days Commuting per Week', ylabel='Worker Share')
   
  

static
inter
write.csv(commute_summary,'commute_summary_trend.csv')


```


```{r}
workers<-person%>%filter(!telecommute_freq %in% c('Missing: Skip logic', "Missing: Skip Logic"))%>%drop_na(telecommute_freq)%>%filter(telecommute_freq!='NA')%>% mutate(telecommute_freq_simple=factor(telecommute_freq_simple, levels=c("Never or less than weekly", '1-2 days', '3-4 days', '5+ days')))%>%rename(survey_year=survey_year.x)
worker_list<-list(person=workers)

```

```{r}
telecommute_freq_simple <- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_freq_simple',
                               summarize_by = 'survey_year',
                               id_cols=c('person_id'),
                               wt_cols=c('person_weight'),
                               wtname= 'person_weight'
                               )
```



This one is confusing because it does not include the always work at home people.
We do NOT know how much people who always work at home telework.
```{r}


telecommute_freq_simple_summary<-telecommute_freq_simple$summary$wtd
static_column_chart(telecommute_freq_simple_summary, y='prop', x='telecommute_freq_simple', fill='survey_year', ylabel= 'Worker Share, of those who do not always work at home', xlabel='Telecommute Frequency') + theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=20),legend.text = element_text(size=20))

```

```{r}
workers_23<-person%>%filter(!workplace %in% c('Missing: Skip logic', "Missing: Skip Logic"))%>%drop_na(workplace)%>%filter(survey_year.x=='2023')%>%rename(survey_year=survey_year.x)
worker_list_23<-list(person=workers_23)
```


```{r}

office_available <- summarize_weighted(hts_data= worker_list_23,
                               summarize_var = 'office_available',
                               summarize_by = 'workplace',
                               id_cols=c('person_id'),
                               wt_cols=c('person_weight'),
                               wtname="person_weight"
                               )

```

## Visualize
```{r}

office_avail_summary<-office_available$summary$wtd%>%
  mutate(office_available=
    case_when(
    office_available=='No, I do not have a dedicated and private space available outside the home' ~ 'No',
    office_available=='Yes, I have a dedicated and private space available outside the home that I do not use.' ~ 'Yes',
    .default= 'Missing')
  )%>%mutate(office_available=factor(office_available, levels=c("Yes", 'No')))
static_column_chart(office_avail_summary, x='office_available', y='prop', fill='workplace')+ theme(axis.text.x=element_text(size=20), axis.text.y=element_text(size=20),legend.text = element_text(size=20))

```

                                                                                                                                  






