---
title: "telework_for_presentation"
author: "suzanne"
date: "2024-04-08"
output: html_document
---


```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
library(forcats)
source('survey-23-preprocess.R')
```


```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad,home_jurisdiction, home_county, hh_weight,vehicle_count, home_rgcname, home_lat, home_lng, survey_year
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category, race_category_2023,
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, sexuality, industry,office_available, commute_freq, education, work_lat, work_lng , work_rgcname, work_jurisdiction, work_county
                from HHSurvey.v_persons_labels")
                
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,
person_id, mode_1, dest_purpose, survey_year, trip_weight, distance_miles, mode_characterization, travelers_total 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)
setDT(day)
setDT(trip)
```
#race_category needs to be added to the codebook

# to do: add Seattle, Bellevue geographies

## Read in Codebook
```{r pressure, echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```

## Set IDs as characters
TO DO: functionalize convert all ids to characters, or store as characters upfront
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
day[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
day[, person_id := as.character(person_id)]
trip[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
trip[, day_id := as.character(day_id)]

trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
day <- day%>%mutate(survey_year=as.character(survey_year))
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```







```{r}
variable_list<-add_variable(variable_list, 'telecommute_freq_cond','person')
group_labels<-get_grouped_labels(group_id='group_2', group_name='telecommute_freq_cond')
value_labels<-add_values_code(group_name='telecommute_freq_cond')
person<-grp_to_tbl(tbl=person, ungrouped_name='telecommute_freq', grouped_name='telecommute_freq_cond')
```

```{r}
variable_list<-add_variable(variable_list, 'telecommute_freq_broad','person')
group_labels<-get_grouped_labels(group_id='group_3', group_name='telecommute_freq_broad')
value_labels<-add_values_code(group_name='telecommute_freq_broad')
person<-grp_to_tbl(tbl=person, ungrouped_name='telecommute_freq', grouped_name='telecommute_freq_broad')
```

```{r}
variable_list<-add_variable(variable_list, 'commute_freq_simple','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='commute_freq_simple')
value_labels<-add_values_code(group_name='commute_freq_simple')
person<-grp_to_tbl(tbl=person, ungrouped_name='commute_freq', grouped_name='commute_freq_simple')
```
```{r}
variable_list<-add_variable(variable_list, 'industry_cond','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='industry_cond')
value_labels<-add_values_code(group_name='industry_cond')
person<-grp_to_tbl(tbl=person, ungrouped_name='industry', grouped_name='industry_cond')
```
```{r}
variable_list<-add_variable(variable_list, 'industry_usim','person')
group_labels<-get_grouped_labels(group_id='group_2', group_name='industry_usim')
value_labels<-add_values_code(group_name='industry_usim')
person<-grp_to_tbl(tbl=person, ungrouped_name='industry', grouped_name='industry_usim')
```


```{r}
variable_list<-add_variable(variable_list, 'age_grp_2','person')
group_labels<-get_grouped_labels(group_id='group_3', group_name='age_grp_2')
value_labels<-add_values_code(group_name='age_grp_2')
person<-grp_to_tbl(tbl=person, ungrouped_name='age', grouped_name='age_grp_2')
```

```{r}
variable_list<-add_variable(variable_list, 'telecommute_freq_simple','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='telecommute_freq_simple')
value_labels<-add_values_code(group_name='telecommute_freq_simple')
person<-grp_to_tbl(tbl=person, ungrouped_name='telecommute_freq', grouped_name='telecommute_freq_simple')%>%
  mutate(telecommute_freq_simple=gsub("[\r\n]","", telecommute_freq_simple))# there's some weird characters in there i can't figure out how to remove
```
```{r}
variable_list<-add_variable(variable_list, 'edu_simple','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='edu_simple')
value_labels<-add_values_code(group_name='edu_simple')
person<-grp_to_tbl(tbl=person, ungrouped_name='education', grouped_name='edu_simple')
```

```{r}
variable_list<-add_variable(variable_list, 'sexuality_broad','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='sexuality_broad')
value_labels<-add_values_code(group_name='sexuality_broad')
person<-grp_to_tbl(tbl=person, ungrouped_name='sexuality', grouped_name='sexuality_broad')
```
```{r}
variable_list<-add_variable(variable_list, 'vehicle_count_simple','hh')
group_labels<-get_grouped_labels(group_id='group_1', group_name='vehicle_count_simple')
value_labels<-add_values_code(group_name='vehicle_count_simple')
hh<-grp_to_tbl(tbl=hh, ungrouped_name='vehicle_count', grouped_name='vehicle_count_simple')
```

```{r}
variable_list<-add_variable(variable_list, 'telecommute_status', 'person')
variable_list<-add_variable(variable_list, 'telecommute_broad', 'person')
variable_list<-add_variable(variable_list, 'telecommute_status_broad', 'person')
variable_list<-add_variable(variable_list, 'commute_freq_all', 'person')

```

# add necessary variables to the persons table
# update telecommute_freq_cond to also include at home people




```{r}
hh<-hh%>%mutate(hhincome_broad=if_else(hhincome_broad %in% 
c('Under $25,000', '$25,000-$49,999', '$50,000-$74,999'), 'Under $75,000', hhincome_broad ))%>%
 mutate(hhincome_broad=if_else(hhincome_broad %in% 
c('$75,000-$99,999', '$100,000-$199,000'), '$75,000-$199,999', hhincome_broad )) 
```


```{r}
value<- c(0,3)
label=c('Under $75,000', '$75,000-$199,999')
inc_val_labels<-data.frame(value, label)
inc_val_labels<-inc_val_labels%>%mutate(variable='hhincome_broad')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, inc_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels 
hh<- order_factors(hh, 'hhincome_broad', value_labels)
```

```{r}
person<- order_factors(person, 'age', value_labels)
person<-person%>%mutate(telecommute_freq_cond=case_when(
                  workplace=='At home (telecommute or self-employed with home office)'
                  ~ '1+ days per week',
                  .default=telecommute_freq_cond))

```



```{r}

value<- c(0,1,2,3)
label=c('Not Worker', 'Fully at Home', 'Hybrid', 'Fully in Person')
tele_val_labels<-data.frame(value, label)
tele_val_labels<-tele_val_labels%>%mutate(variable='telecommute_status')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, tele_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  


```


```{r}

value<- c(0,1,2,3,4)
label=c('Not Worker', 'Fully at Home', 'Mostly In Person', 'Mostly at Home', 'Fully in Person')
tele_val_labels<-data.frame(value, label)
tele_val_labels<-tele_val_labels%>%mutate(variable='telecommute_status_broad')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, tele_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  


```

```{r}

value<- c(0,1,2,3, 4, 5)
label=c('Never or less than weekly', '1 days', '2 days', '3 days', '4 days', '5+ days')
val_labels<-data.frame(value, label)
val_labels<-tele_val_labels%>%mutate(variable='commute_freq_all')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, tele_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  

```


More detailed variable definitions
```{r}
person<-person%>%mutate(telecommute_status=
        case_when(workplace %in%c('Missing: Skip logic', "Missing: Skip Logic") 
                  ~ 'Not Worker',
                  (is.na(workplace)) ~ 'Not Worker',
                  (workplace=='At home (telecommute or self-employed with home office)' )
                  ~ 'Fully At Home',
                  (telecommute_freq_cond == '1+ days per week')
                  ~'Hybrid',
                  .default = 'Fully In Person'
        )
)

```




```{r}
person<-person%>%mutate(telecommute_status_broad=
        case_when(
                  (telecommute_freq_broad == '1-2 days')
                  ~'Mostly In Person',
                  (telecommute_freq_broad == '3+ days')
                  ~'Mostly At Home',
                  .default = telecommute_status
        )
)

```

How much are people commuting?
Combine At home with other workers

```{r}
person<-person%>%mutate(commute_freq_all=
                           if_else(workplace== "At home (telecommute or self-employed with home office)", 
                                   'Never or less than weekly', commute_freq))



person<- person%>%mutate(commute_freq_all=case_match(commute_freq_all,
                               "1-3 days a month" ~ "Never or less than weekly",
                               'Never' ~ "Never or less than weekly",
                                "A few times per month"   ~  "Never or less than weekly",  
                                "Less than monthly"   ~  "Never or less than weekly",  
                               "None, did not commute to workplace last week" ~"Never or less than weekly",
                                '5 days a week' ~ '5+ days',
                                '6-7 days a week' ~ '5+ days',
                               .default= commute_freq_all
                               ))


person<- person%>%mutate(commute_freq_all=str_replace(commute_freq_all, ' a week', ''))
                               
person<- person%>%mutate(commute_freq_all=case_when(
  commute_freq_all %in% c('1 day', '2 days') ~ '1-2 days',
  commute_freq_all %in% c('3 days', '4 days') ~ '3-4 days',
  .default = commute_freq_all
  
))  

#person<- person%>%mutate(commute_freq_all=factor(commute_freq_all, levels=c("Never or less than #weekly", '1-2 days', '3-4 days', '5+ days')))


```


```{r}
person<-person%>%mutate(telecommute_status=
        case_when(workplace %in%c('Missing: Skip logic', "Missing: Skip Logic") 
                  ~ 'Not Worker',
                  (is.na(workplace)) ~ 'Not Worker',
                  (workplace=='At home (telecommute or self-employed with home office)' )
                  ~ 'Fully At Home',
                  (telecommute_freq_cond == '1+ days per week')
                  ~'Hybrid',
                  .default = 'Fully In Person'
        )
)

```




```{r}
person<-person%>%mutate(telecommute_status_broad=
        case_when(
                  (telecommute_freq_broad == '1-2 days')
                  ~'1-2 days',
                  (telecommute_freq_broad == '3+ days')
                  ~'3-4 days',
                  .default = telecommute_status
        )
)

```





```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```





```{r}
workers<-person%>%filter(!workplace %in% c('Missing: Skip logic', "Missing: Skip Logic"))%>%drop_na(workplace)
worker_list<-list(hh=hh, person=workers)

```



People having location available outside home

Ability to Telework vs Telework

Vehicle Miles Traveled per Capita by Telecommute Amount

-	By Home Location
-	By Work Location







