---
title: "Displacement Analysis"
format: html
---

```{r}
#| include: false
# YOU MIGHT need to download some special packages on github: travelSurveyTools, psrcelmer, and psrcplot. IF you need to download them, here's how:
# library(remotes)
 
# remotes::install_github('RSGInc/travelSurveyTools')
# remotes::install_github('psrc/psrcelmer')
# remotes::install_github('psrc/psrcplot')
# using this package
# remotes::install_github('psrc/psrc.travelsurvey', force=TRUE)

library(data.table)
library(psrcelmer)
library(psrcplot)
library(tidyverse)
library(table1)
library(psrc.travelsurvey)
library(scales)

library(sf) #for regional geographies
sf_use_s2(FALSE)
library(corrplot) #for correlation analysis

# source('../survey-23-preprocess.R')
```

<!-- ## Read in Codebook -->

```{r}
# cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")
# 
# variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
# value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')
# 
# setDT(variable_list)
# setDT(value_labels)
```

<!-- ## Read in Data from Elmer -->

```{r}
# Specify which variables to retrieve
# View(psrc.travelsurvey:::init_variable_list)

vars <- c(
  #household-level variables
  'household_id', 'res_dur', 'prev_home_wa', 'home_rgcname', 
  'prev_home_rgcname', 'home_county', 'prev_home_county',
  'home_jurisdiction', 'prev_rent_own', 'rent_own', 
  'prev_res_type', 'res_type',
  'hhsize', 'numadults', 'numchildren', 'numworkers', 'vehicle_count', 
  'hhincome_detailed', 'hhincome_broad',
  'prev_res_factors_amenities', 'prev_res_factors_community_change',
  'prev_res_factors_crime', 'prev_res_factors_employment', 
  'prev_res_factors_forced', 'prev_res_factors_hh_size', 
  'prev_res_factors_housing_cost', 'prev_res_factors_income_change', 
  'prev_res_factors_less_space', 'prev_res_factors_more_space', 
  'prev_res_factors_no_answer', 'prev_res_factors_other', 
  'prev_res_factors_quality', 'prev_res_factors_school', 
  'prev_res_factors_specify', 'prev_res_factors_telework', 
  'hh_race_category',
  "home_jurisdiction","home_lat","home_lng", 'displaced',"hhincome_bin5",
  
  #person-level variables
  'race_category', 'race_afam', 'race_aiak', 'race_asian', 'race_hapi', 
  'race_noanswer', 'race_other', 'race_other_specify', 'race_white',
  'ethnicity_1', 'ethnicity_2', 'ethnicity_3', 'ethnicity_4', 
  'ethnicity_997', 'ethnicity_999', 'ethnicity_other',
  'person_id', 'gender', 'age',
  'education', 'employment')

# Retrieve the data
survey_years = c(2017, 2019, 2021, 2023, 2025)
hts_data <- get_psrc_hts(survey_years = survey_years, survey_vars = vars)  # default includes all survey_years
```

## Percent displaced by survey year

(number of people who lived in washington before and have moved in the past 5 years who moved for the four displacement reasons)/(number of people who lived in washington before and have moved in the past 5 years)

```{r}
hh_test <- hts_data$hh %>% 
  mutate(movers=case_when(res_dur=="Less than a year" |
                            res_dur=="Between 1 and 2 years" | 
                            res_dur=="Between 2 and 3 years" |
                            res_dur=="Between 3 and 5 years"
                          ~ 'Did move',
                          res_dur=="Missing Response" ~ 'No response',
                          TRUE ~ 'Did not move'),
         displaced=case_when(prev_res_factors_forced=='Selected'|
                               prev_res_factors_income_change=='Selected'|
                               prev_res_factors_housing_cost =='Selected'|
                               prev_res_factors_community_change=='Selected'~ 'Displaced',
                             TRUE~'Not displaced')
  ) %>%
    filter(
     prev_home_wa == "Yes, previous home was in Washington"
    )

#table(hh_test$movers, hh_test$displaced)


# There is an issue with the 2023 data where some people who hadn't moved within the past 5 years were able to select reasons for moving (prev_res_factors). This shouldn't have happened: all respondents were asked how long they had been in their current homes > if they chose <5 years they were asked if their previous residence had been in WA > if yes, they were asked to select why they moved. As a result, we will need to filter out the people who did not move.

hh_test_move <- hh_test %>%
  filter(movers == "Did move")
```

```{r}
hh_displaced <- hh_test_move %>%
  mutate(moved = case_when(res_dur %in% c('Less than a year', 'Between 1 and 2 years', 'Between 2 and 3 years', 'Between 3 and 5 years')~ "moved",
                           TRUE~"not moved"),
        displaced_no_res = case_when(is.na(prev_res_factors_forced) ~ 'No Response',
                                      TRUE~displaced)
         )
test <- hts_data
test$hh <- hh_displaced

res <- psrc_hts_stat(test, "hh", c("moved","displaced_no_res"), incl_na = TRUE)

res
```

```{r}
plot_data <- res %>%
  filter(displaced_no_res == "Displaced")

ggplot(plot_data,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1]) +
   geom_errorbar(
    aes(
      ymin = pmax(prop - prop_moe, 0),
      ymax = pmin(prop + prop_moe, 1)
    ),
    width = 0.15,
    linewidth = 0.7
  ) +
  geom_text(
    aes(
      y = pmin(prop + prop_moe + 0.02, 1),
      label = percent(prop, accuracy = 0.1)
    ),
    size = 5
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    title = "Percent Displaced by Survey Year",
    x = "Survey Year",
    y = "Percent Displaced"
  ) +
  theme_minimal()
```

## percent displaced by county

```{r}
# create new variables with exhaustive filters 
hts_data$hh <- hts_data$hh %>% 
  mutate(full_sample=case_when(!is.na(hh_weight)~"Full sample",
                               TRUE~"No"),
         movers=case_when(res_dur=="Less than a year" & full_sample=="Full sample"|
                            res_dur=="Between 1 and 2 years" & full_sample=="Full sample"| 
                            res_dur=="Between 2 and 3 years" & full_sample=="Full sample"|
                            res_dur=="Between 3 and 5 years" & full_sample=="Full sample"~ 'Did move',
                          TRUE ~ 'Did not move'),
         movers_wa=case_when(movers=="Did move" & grepl("Yes", prev_home_wa)|
                               movers=="Did move" & grepl("Yes", prev_home_wa)| 
                               movers=="Did move" & grepl("Yes", prev_home_wa)|
                               movers=="Did move" & grepl("Yes", prev_home_wa) ~ 'Did move in WA',
                             TRUE ~ 'Did not move in WA'),
         displaced=case_when(prev_res_factors_forced=='Selected' & movers_wa=="Did move in WA"|
                               prev_res_factors_income_change=='Selected' & movers_wa=="Did move in WA"|
                               prev_res_factors_housing_cost =='Selected' & movers_wa=="Did move in WA"|
                               prev_res_factors_community_change=='Selected' & movers_wa=="Did move in WA" ~ 'Displaced',
                             TRUE~'Not displaced')
  )

# generate counts/weighted estimates for subpopulations
total_hh <- psrc_hts_stat(hts_data, "hh", c("full_sample"), incl_na = TRUE)
moving_wa <- psrc_hts_stat(hts_data, "hh", c("full_sample", "movers_wa"), incl_na = TRUE)
displaced_moving_wa <- psrc_hts_stat(hts_data, "hh", c("movers_wa", "displaced"), incl_na = TRUE)

displaced_moving_wa_county <- psrc_hts_stat(hts_data, "hh", c("home_county","movers_wa", "displaced"), incl_na = TRUE)
print(displaced_moving_wa_county)
displaced_moving_wa_county_filter <- displaced_moving_wa_county %>%
  filter(movers_wa=="Did move in WA",
         displaced=="Displaced"
  )
print(displaced_moving_wa_county_filter)
```

## bar graph percent displaced by county

```{r}
ggplot(displaced_moving_wa_county_filter,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
  geom_errorbar(
    aes(
      ymin = pmax(prop - prop_moe, 0),
      ymax = pmin(prop + prop_moe, 1)
    ),
    width = 0.15,
    linewidth = 0.7
  ) +
  geom_text(
    aes(
      y = pmin(prop + prop_moe + 0.02, 1),
      label = percent(prop, accuracy = 0.1)
    ),
    size = 3.5
  ) +
  facet_wrap(~home_county) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.8),   # increased to leave room for MOE and labels
    breaks = seq(0, 0.8, by = 0.1)
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percent Displaced",
    title = "Percentage of WA Movers Who Were Displaced by County"
  ) +
  theme_minimal(base_size = 12)
```

## percent displaced by reason

```{r}
vars <- c(
  #household-level variables
  'household_id', 'res_dur', 'prev_home_wa', 'home_rgcname', 
  'prev_home_rgcname', 'home_county', 'prev_home_county',
  'prev_res_factors_amenities', 'prev_res_factors_community_change',
  'prev_res_factors_crime', 'prev_res_factors_employment', 
  'prev_res_factors_forced', 'prev_res_factors_hh_size', 
  'prev_res_factors_housing_cost', 'prev_res_factors_income_change', 
  'prev_res_factors_less_space', 'prev_res_factors_more_space', 
  'prev_res_factors_no_answer', 'prev_res_factors_other', 
  'prev_res_factors_quality', 'prev_res_factors_school', 
  'prev_res_factors_specify', 'prev_res_factors_telework', 
  'hh_race_category',
  "home_jurisdiction","home_lat","home_lng",
  "prev_home_lat", "prev_home_lng", 
  "home_tract20", "prev_home_tract20")

# this code is clean for reference purposes - it is the most restrictive in terms of filters so that any survey logic errors are accounted for and the sub-populations are those we are most interested in

# download data
hts_data <- get_psrc_hts(survey_vars = vars)  # default includes all survey_years

# create new variables with exhaustive filters 
hts_data$hh <- hts_data$hh %>% 
  mutate(full_sample=case_when(!is.na(hh_weight)~"Full sample",
                               TRUE~"No"),
         movers=case_when(res_dur=="Less than a year" & full_sample=="Full sample"|
                            res_dur=="Between 1 and 2 years" & full_sample=="Full sample"| 
                            res_dur=="Between 2 and 3 years" & full_sample=="Full sample"|
                            res_dur=="Between 3 and 5 years" & full_sample=="Full sample"~ 'Did move',
                          TRUE ~ 'Did not move'),
         movers_wa=case_when(movers=="Did move" & grepl("Yes", prev_home_wa)|
                               movers=="Did move" & grepl("Yes", prev_home_wa)| 
                               movers=="Did move" & grepl("Yes", prev_home_wa)|
                               movers=="Did move" & grepl("Yes", prev_home_wa) ~ 'Did move in WA',
                             TRUE ~ 'Did not move in WA'),
         displaced=case_when(prev_res_factors_forced=='Selected' & movers_wa=="Did move in WA"|
                               prev_res_factors_income_change=='Selected' & movers_wa=="Did move in WA"|
                               prev_res_factors_housing_cost =='Selected' & movers_wa=="Did move in WA"|
                               prev_res_factors_community_change=='Selected' & movers_wa=="Did move in WA" ~ 'Displaced',
                             TRUE~'Not displaced')
  )

# generate counts/weighted estimates for subpopulations
total_hh <- psrc_hts_stat(hts_data, "hh", c("full_sample"), incl_na = TRUE)
moving_wa <- psrc_hts_stat(hts_data, "hh", c("full_sample", "movers_wa"), incl_na = TRUE)
displaced_moving_wa <- psrc_hts_stat(hts_data, "hh", c("movers_wa", "displaced"), incl_na = TRUE)

displaced_moving_wa_county <- psrc_hts_stat(hts_data, "hh", c("home_county","movers_wa", "displaced"), incl_na = TRUE)

displaced_moving_wa_county_filter <- displaced_moving_wa_county %>%
  filter(movers_wa=="Did move in WA",
         displaced=="Displaced"
  )
```

### reason: forced

```{r}
# creating simplified forced variable
hts_data$hh <- hts_data$hh %>% 
  mutate(forced = case_when(prev_res_factors_forced=="Selected"~"Forced",
                           prev_res_factors_forced=="Not Selected"~"No"))

# calculating estimates/shares
forced_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "forced"), 
                              incl_na = TRUE)
print(forced_trial)

```

#### bar graph: forced

```{r}

library(dplyr)
library(ggplot2)
library(scales)

forced_plot <- forced_trial %>%
  filter(
    forced == "Forced",
    movers_wa == "Did move in WA"
  )

ggplot(forced_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
  geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.45),   # increase to accommodate MOE
    breaks = seq(0, 0.45, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Due to Forced"
  ) +
  theme_minimal(base_size = 12)
```

### reason: income change

```{r}
# income
# table(hts_data$hh$prev_res_factors_income_change)
hts_data$hh <- hts_data$hh %>% 
  mutate(income = case_when(prev_res_factors_income_change=="Selected"~"Income change",
                           prev_res_factors_income_change=="Not Selected"~"No"))

income_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "income"), 
                              incl_na = TRUE)


print(income_trial)
```

#### bar graph: income change

```{r}

library(dplyr)
library(ggplot2)
library(scales)

income_plot <- income_trial %>%
  filter(
    income == "Income change",
    movers_wa == "Did move in WA"
    
  )

ggplot(income_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
  geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.45),   # increase to accommodate MOE
    breaks = seq(0, 0.45, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Due to Income Change"
  ) +
  theme_minimal(base_size = 12)
```

### reason: housing cost

```{r}
# housing cost
# table(hts_data$hh$prev_res_factors_housing_cost)
hts_data$hh <- hts_data$hh %>% 
  mutate(housing = case_when(prev_res_factors_housing_cost=="Selected"~"Housing cost",
                           prev_res_factors_housing_cost=="Not Selected"~"No"))

housing_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "housing"), 
                              incl_na = TRUE)

print(housing_trial)
```

#### bar graph: housing cost

```{r}

library(dplyr)
library(ggplot2)
library(scales)

housing_plot <- housing_trial %>%
  filter(
    housing == "Housing cost",
    movers_wa == "Did move in WA"
  )

ggplot(housing_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
  geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.70),   # increase to accommodate MOE
    breaks = seq(0, 0.70, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Due to Income Change"
  ) +
  theme_minimal(base_size = 12)
```

### reason: community change

```{r}
# community change
# table(hts_data$hh$prev_res_factors_community_change)
hts_data$hh <- hts_data$hh %>% 
  mutate(community = case_when(prev_res_factors_community_change=="Selected"~"Community Change",
                           prev_res_factors_community_change=="Not Selected"~"No"))

community_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "community"), 
                              incl_na = TRUE)

print(community_trial)
```

#### bar graph: community change

```{r}
library(dplyr)
library(ggplot2)
library(scales)

community_plot <- community_trial %>%
  filter(
    community == "Community Change",
    movers_wa == "Did move in WA"
  )

ggplot(community_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
    geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.40),   # increase to accommodate MOE
    breaks = seq(0, 0.40, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Reporting Community Change as a Reason"
  ) +
  theme_classic(base_size = 12)
```

### reason: amenities

```{r}
# amenities
# table(hts_data$hh$prev_res_factors_amenities)
hts_data$hh <- hts_data$hh %>% 
  mutate(amenities = case_when(prev_res_factors_amenities=="Selected"~"Amenities",
                           prev_res_factors_amenities=="Not Selected"~"No"))

amenities_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "amenities"), 
                              incl_na = TRUE)

print(amenities_trial)
```

#### bar graph: amenities

```{r}
library(dplyr)
library(ggplot2)
library(scales)

amenities_plot <- amenities_trial %>%
  filter(
    amenities == "Amenities",
    movers_wa == "Did move in WA"
  )

ggplot(amenities_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
    geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.30),   # increase to accommodate MOE
    breaks = seq(0, 0.30, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Reporting Amenities as a Reason"
  ) +
  theme_classic(base_size = 12)
```

### reason: crime

```{r}
# crime
# table(hts_data$hh$prev_res_factors_crime)
hts_data$hh <- hts_data$hh %>% 
  mutate(crime = case_when(prev_res_factors_crime=="Selected"~"Crime",
                           prev_res_factors_crime=="Not Selected"~"No"))

crime_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "crime"), 
                              incl_na = TRUE)

print(crime_trial)
```

#### bar graph: crime

```{r}
library(dplyr)
library(ggplot2)
library(scales)

crime_plot <- crime_trial %>%
  filter(
    crime == "Crime",
    movers_wa == "Did move in WA"
  )

ggplot(crime_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
    geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.30),   # increase to accommodate MOE
    breaks = seq(0, 0.30, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Reporting Crime as a Reason"
  ) +
  theme_classic(base_size = 12)
```

### reason: employment

```{r}
# employment
# table(hts_data$hh$prev_res_factors_employment)
hts_data$hh <- hts_data$hh %>% 
  mutate(employment = case_when(prev_res_factors_employment=="Selected"~"Employment",
                           prev_res_factors_employment=="Not Selected"~"No"))

employment_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "employment"), 
                              incl_na = TRUE)

print(employment_trial)
```

#### bar graph: employment

```{r}
library(dplyr)
library(ggplot2)
library(scales)

employment_plot <- employment_trial %>%
  filter(
    employment == "Employment",
    movers_wa == "Did move in WA"
  )

ggplot(employment_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
  geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.35),   # increase to accommodate MOE
    breaks = seq(0, 0.35, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Reporting Employment as a Reason"
  ) +
  theme_classic(base_size = 12)
```

### reason: household size

```{r}
# household size
# table(hts_data$hh$prev_res_factors_hh_size)
hts_data$hh <- hts_data$hh %>% 
  mutate(hhsize = case_when(prev_res_factors_hh_size=="Selected"~"Household size",
                           prev_res_factors_hh_size=="Not Selected"~"No"))

hhsize_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "hhsize"), 
                              incl_na = TRUE)

print(hhsize_trial)
```

#### bar graph: household size

```{r}
library(dplyr)
library(ggplot2)
library(scales)

hhsize_plot <- hhsize_trial %>%
  filter(
    hhsize == "Household size",
    movers_wa == "Did move in WA"
  )

ggplot(hhsize_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
  geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.45),   # increase to accommodate MOE
    breaks = seq(0, 0.45, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Reporting Household Size as a Reason"
  ) +
  theme_classic(base_size = 12)
```

### reason: less space

```{r}
# less space
# table(hts_data$hh$prev_res_factors_less_space)
hts_data$hh <- hts_data$hh %>% 
  mutate(lessspace = case_when(prev_res_factors_less_space=="Selected"~"Less space",
                           prev_res_factors_less_space=="Not Selected"~"No"))

lessspace_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "lessspace"), 
                              incl_na = TRUE)

print(lessspace_trial)
```

#### bar graph: less space

```{r}
library(dplyr)
library(ggplot2)
library(scales)

lessspace_plot <- lessspace_trial %>%
  filter(
    lessspace == "Less space",
    movers_wa == "Did move in WA"
  )

ggplot(lessspace_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
  geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.10),   # increase to accommodate MOE
    breaks = seq(0, 0.10, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Reporting Less Space as a Reason"
  ) +
  theme_classic(base_size = 12)
```

### reason: more space

```{r}
# more space
# table(hts_data$hh$prev_res_factors_more_space)
hts_data$hh <- hts_data$hh %>% 
  mutate(morespace = case_when(prev_res_factors_more_space=="Selected"~"More space",
                           prev_res_factors_more_space=="Not Selected"~"No"))

morespace_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "morespace"), 
                              incl_na = TRUE)

print(morespace_trial)
```

#### bar graph: more space

```{r}
library(dplyr)
library(ggplot2)
library(scales)

morespace_plot <- morespace_trial %>%
  filter(
    morespace == "More space",
    movers_wa == "Did move in WA"
  )

ggplot(morespace_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
  geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.50),   # increase to accommodate MOE
    breaks = seq(0, 0.50, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Reporting More Space as a Reason"
  ) +
  theme_classic(base_size = 12)
```

### reason: no answer

```{r}
# no answer
# table(hts_data$hh$prev_res_factors_no_answer)
hts_data$hh <- hts_data$hh %>% 
  mutate(noanswer = case_when(prev_res_factors_no_answer=="Selected"~"No answer",
                           prev_res_factors_no_answer=="Not Selected"~"No"))

noanswer_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "noanswer"), 
                              incl_na = TRUE)

print(noanswer_trial)
```

#### bar graph: no answer

```{r}
library(dplyr)
library(ggplot2)
library(scales)

noanswer_plot <- noanswer_trial %>%
  filter(
    noanswer == "No answer",
    movers_wa == "Did move in WA"
  )

ggplot(noanswer_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
  geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.10),   # increase to accommodate MOE
    breaks = seq(0, 0.10, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Reporting No Answer as a Reason"
  ) +
  theme_classic(base_size = 12)
```

### reason: other

```{r}
# other
# table(hts_data$hh$prev_res_factors_other)
hts_data$hh <- hts_data$hh %>% 
  mutate(other = case_when(prev_res_factors_other=="Selected"~"Other",
                           prev_res_factors_other=="Not Selected"~"No"))

other_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "other"), 
                              incl_na = TRUE)

print(other_trial)
```

#### bar graph: other

```{r}
library(dplyr)
library(ggplot2)
library(scales)

other_plot <- other_trial %>%
  filter(
    other == "Other",
    movers_wa == "Did move in WA"
  )

ggplot(other_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
  geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.15),   # increase to accommodate MOE
    breaks = seq(0, 0.15, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Reporting Other as a Reason"
  ) +
  theme_classic(base_size = 12)
```

### reason: quality

```{r}
# quality
# table(hts_data$hh$prev_res_factors_quality)
hts_data$hh <- hts_data$hh %>% 
  mutate(quality = case_when(prev_res_factors_quality=="Selected"~"Quality",
                           prev_res_factors_quality=="Not Selected"~"No"))

quality_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "quality"), 
                              incl_na = TRUE)

print(quality_trial)
```

#### bar graph: quality

```{r}
library(dplyr)
library(ggplot2)
library(scales)

quality_plot <- quality_trial %>%
  filter(
    quality == "Quality",
    movers_wa == "Did move in WA"
  )

ggplot(quality_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
  geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.45),   # increase to accommodate MOE
    breaks = seq(0, 0.45, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Reporting Quality as a Reason"
  ) +
  theme_classic(base_size = 12)
```

### reason: school

```{r}
# school
# table(hts_data$hh$prev_res_factors_school)
hts_data$hh <- hts_data$hh %>% 
  mutate(school = case_when(prev_res_factors_school=="Selected"~"School",
                           prev_res_factors_school=="Not Selected"~"No"))

school_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "school"), 
                              incl_na = TRUE)

print(school_trial)
```

#### bar graph: school

```{r}
library(dplyr)
library(ggplot2)
library(scales)

school_plot <- school_trial %>%
  filter(
    school == "School",
    movers_wa == "Did move in WA"
  )

ggplot(school_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
  geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.15),   # increase to accommodate MOE
    breaks = seq(0, 0.15, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Reporting School as a Reason"
  ) +
  theme_classic(base_size = 12)
```

### reason: telework

```{r}
# telework
# table(hts_data$hh$prev_res_factors_telework)
hts_data$hh <- hts_data$hh %>% 
  mutate(telework = case_when(prev_res_factors_telework=="Selected"~"Telework",
                           prev_res_factors_telework=="Not Selected"~"No"))

telework_trial <- psrc_hts_stat(hts_data, "hh", 
                              c("movers_wa", "telework"), 
                              incl_na = TRUE)

print(telework_trial)
```

#### bar graph: telework

```{r}
library(dplyr)
library(ggplot2)
library(scales)

telework_plot <- telework_trial %>%
  filter(
    telework == "Telework",
    movers_wa == "Did move in WA"
  )

ggplot(telework_plot,
       aes(x = factor(survey_year), y = prop)) +
  geom_col(fill = psrc_colors$pgnobgy_5[1], width = 0.6) +
  geom_errorbar(
  aes(
    ymin = pmax(prop - prop_moe, 0),
    ymax = pmin(prop + prop_moe, 1)
  ),
  width = 0.15,
  linewidth = 0.8
  ) +
  geom_text(
    aes(label = percent(prop, accuracy = 0.1)),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.05),   # increase to accommodate MOE
    breaks = seq(0, 0.05, by = 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "Survey Year",
    y = "Weighted Percentage",
    title = "Displaced Movers Reporting Telework as a Reason"
  ) +
  theme_classic(base_size = 12)
```

```{r}
library(dplyr)

reasons <- bind_rows(

  forced_trial %>%
    filter(forced == "Forced", movers_wa == "Did move in WA") %>%
    mutate(reason = "Forced"),

  income_trial %>%
    filter(income == "Income change", movers_wa == "Did move in WA") %>%
    mutate(reason = "Income change"),

  housing_trial %>%
    filter(housing == "Housing cost", movers_wa == "Did move in WA") %>%
    mutate(reason = "Housing cost"),

  community_trial %>%
    filter(community == "Community Change", movers_wa == "Did move in WA") %>%
    mutate(reason = "Community change"),

  amenities_trial %>%
    filter(amenities == "Amenities", movers_wa == "Did move in WA") %>%
    mutate(reason = "Amenities"),

  crime_trial %>%
    filter(crime == "Crime", movers_wa == "Did move in WA") %>%
    mutate(reason = "Crime"),

  employment_trial %>%
    filter(employment == "Employment", movers_wa == "Did move in WA") %>%
    mutate(reason = "Employment"),

  hhsize_trial %>%
    filter(hhsize == "Household size", movers_wa == "Did move in WA") %>%
    mutate(reason = "Household size"),

  lessspace_trial %>%
    filter(lessspace == "Less space", movers_wa == "Did move in WA") %>%
    mutate(reason = "Less space"),

  morespace_trial %>%
    filter(morespace == "More space", movers_wa == "Did move in WA") %>%
    mutate(reason = "More space"),

  noanswer_trial %>%
    filter(noanswer == "No answer", movers_wa == "Did move in WA") %>%
    mutate(reason = "No answer"),

  other_trial %>%
    filter(other == "Other", movers_wa == "Did move in WA") %>%
    mutate(reason = "Other"),

  quality_trial %>%
    filter(quality == "Quality", movers_wa == "Did move in WA") %>%
    mutate(reason = "Quality"),

  school_trial %>%
    filter(school == "School", movers_wa == "Did move in WA") %>%
    mutate(reason = "School"),

  telework_trial %>%
    filter(telework == "Telework", movers_wa == "Did move in WA") %>%
    mutate(reason = "Telework")

) %>%
  select(survey_year, reason, prop, prop_moe)
```

## 2025 reasons plot

```{r}
library(dplyr)
library(ggplot2)
library(scales)

plot_2025 <- reasons %>%
  filter(survey_year == 2025) %>%
  arrange(desc(prop)) %>%
  mutate(
  reason = factor(reason, levels = unique(reason)),
  highlight = if_else(
    reason %in% c("Housing cost", "Income change", "Forced", "Community change"),
    "Highlight",
    "Other"
  )
)
ggplot(plot_2025,
       aes(x = reason, y = prop)) +
  geom_col(aes(fill = highlight), width = 0.7) +
  geom_errorbar(
    aes(
      ymin = pmax(prop - prop_moe, 0),
      ymax = pmin(prop + prop_moe, 1)
    ),
    width = 0.2,
    linewidth = 0.7
  ) +
  geom_text(
    aes(
      y = pmin(prop + prop_moe + 0.02, 1),
      label = percent(prop, accuracy = 0.1)
    ),
    size = 3.5
  ) +
  scale_fill_manual(
    values = c(
      "Highlight" = psrc_colors$pgnobgy_5[2],
      "Other" = psrc_colors$pgnobgy_5[1]
    ),
    guide = "none"
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.08))
  ) +
  labs(
    title = "Reasons for Displacement, 2025",
    x = NULL,
    y = "Weighted Percent"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## 2023 reasons plot

```{r}
library(dplyr)
library(ggplot2)
library(scales)

plot_2023 <- reasons %>%
  filter(survey_year == 2023) %>%
  arrange(desc(prop)) %>%
  mutate(
  reason = factor(reason, levels = unique(reason)),
  highlight = if_else(
    reason %in% c("Housing cost", "Income change", "Forced", "Community change"),
    "Highlight",
    "Other"
  )
)

ggplot(plot_2023,
       aes(x = reason, y = prop)) +
  geom_col(aes(fill = highlight), width = 0.7) +
  geom_errorbar(
    aes(
      ymin = pmax(prop - prop_moe, 0),
      ymax = pmin(prop + prop_moe, 1)
    ),
    width = 0.2,
    linewidth = 0.7
  ) +
  geom_text(
    aes(
      y = pmin(prop + prop_moe + 0.02, 1),
      label = percent(prop, accuracy = 0.1)
    ),
    size = 3.5
  ) +
  scale_fill_manual(
    values = c(
      "Highlight" = psrc_colors$pgnobgy_5[2],
      "Other" = psrc_colors$pgnobgy_5[1]
    ),
    guide = "none"
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.08))
  ) +
  labs(
    title = "Reasons for Displacement, 2023",
    x = NULL,
    y = "Weighted Percent"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## 2021 reasons plot

```{r}
library(dplyr)
library(ggplot2)
library(scales)

plot_2021 <- reasons %>%
  filter(survey_year == 2021) %>%
  arrange(desc(prop)) %>%
  mutate(
  reason = factor(reason, levels = unique(reason)),
  highlight = if_else(
    reason %in% c("Housing cost", "Income change", "Forced", "Community change"),
    "Highlight",
    "Other"
  )
)

ggplot(plot_2021,
       aes(x = reason, y = prop)) +
  geom_col(aes(fill = highlight), width = 0.7) +
  geom_errorbar(
    aes(
      ymin = pmax(prop - prop_moe, 0),
      ymax = pmin(prop + prop_moe, 1)
    ),
    width = 0.2,
    linewidth = 0.7
  ) +
  geom_text(
    aes(
      y = pmin(prop + prop_moe + 0.02, 1),
      label = percent(prop, accuracy = 0.1)
    ),
    size = 3.5
  ) +
  scale_fill_manual(
    values = c(
      "Highlight" = psrc_colors$pgnobgy_5[2],
      "Other" = psrc_colors$pgnobgy_5[1]
    ),
    guide = "none"
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.08))
  ) +
  labs(
    title = "Reasons for Displacement, 2021",
    x = NULL,
    y = "Weighted Percent"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## 2019 reasons plot

```{r}
library(dplyr)
library(ggplot2)
library(scales)

plot_2019 <- reasons %>%
  filter(survey_year == 2019) %>%
  arrange(desc(prop)) %>%
  mutate(
  reason = factor(reason, levels = unique(reason)),
  highlight = if_else(
    reason %in% c("Housing cost", "Income change", "Forced", "Community change"),
    "Highlight",
    "Other"
  )
)

ggplot(plot_2019,
       aes(x = reason, y = prop)) +
  geom_col(aes(fill = highlight), width = 0.7) +
  geom_errorbar(
    aes(
      ymin = pmax(prop - prop_moe, 0),
      ymax = pmin(prop + prop_moe, 1)
    ),
    width = 0.2,
    linewidth = 0.7
  ) +
  geom_text(
    aes(
      y = pmin(prop + prop_moe + 0.02, 1),
      label = percent(prop, accuracy = 0.1)
    ),
    size = 3.5
  ) +
  scale_fill_manual(
    values = c(
      "Highlight" = psrc_colors$pgnobgy_5[2],
      "Other" = psrc_colors$pgnobgy_5[1]
    ),
    guide = "none"
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.08))
  ) +
  labs(
    title = "Reasons for Displacement, 2019",
    x = NULL,
    y = "Weighted Percent"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
