---
title: "Telework Analysis"
date: today

params:
  errorbar: TRUE
---

```{r}
#| include: false

# get data and load standard data processing
source("telework_processing.R")  

plot_formatting <- function(colors = c("#4C4C4C", "#91268F", "#8CC63E", "#F05A28", "#00A7A0")){
  list(
    geom_hline(yintercept = 0, color = "#cbcbcb", linewidth = 0.3),
    scale_fill_manual(values = colors),
    scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, 0.1))),
    psrc_style(),
    theme(panel.grid.major.y = element_blank(),
          legend.text = element_text(size = 11),
          axis.title = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank())
  )
}

```

## Telework definition

- teleworkers are the workers who telework more than one day per week or work fully at home
- 2025 telework status assignments

```{r}
psrc_hts_stat(df_hts_data, "person", 
              c("employment_simple","work_from_home","workplace_simple","telecommute_freq_1plus","telecommute_status"),
              # exclude NAs in calculation: only considering workers
              incl_na=TRUE) %>%
  # function sourced from Analysis/util.R
  # get_table_format() %>%
  arrange(employment_simple,telecommute_status,work_from_home,workplace_simple,telecommute_freq_1plus) %>%
  filter(survey_year == 2025) %>%
  select(employment_simple,workplace_simple,telecommute_freq_1plus,telecommute_status,count)
```


## Telework trend

- trend remained similar since 2023

```{r}
#| label: fig-telework-trend
#| fig-cap: "Telework trend"
#| fig-asp: 0.75
#| fig-width: 6
#| fig-format: svg

df_plot <- psrc_hts_stat(df_hts_data, "person", 
                         c("telecommute_status"),
                         # exclude NAs in calculation: only considering workers
                         incl_na=FALSE) %>%
  group_by(survey_year) %>%
  mutate(
    prop_position = case_when(telecommute_status=="Fully At Home"~0,
                              telecommute_status=="Hybrid"~.data$prop[1],
                              telecommute_status=="Fully In Person"~sum(.data$prop[c(1,2)]))
  )

ggplot(df_plot, 
       aes(x=factor(survey_year), y=prop, fill=fct_rev(telecommute_status))) +
  geom_col(width = 0.7) +
  geom_text(aes(y= prop_position, label=scales::percent(prop, accuracy = 1),
                color=telecommute_status),
            vjust = -0.6) +
  geom_hline(yintercept = c(0,1), color = "#cbcbcb", linewidth = 0.3) +
  scale_fill_manual(values = c("#4C4C4C","#91268F","#F05A28"))+
  scale_color_manual(values = c("#4C4C4C","white","white"), guide = "none")+
  scale_y_continuous(labels = scales::percent) +
  psrc_style() +
  theme(axis.title = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.text.y = element_blank())
```



## Commute mode share

- Active Modes include walk, bike and micromobility

```{r}
#| label: fig-commute-mode-share
#| fig-cap: "Commute mode share"
#| fig-asp: 0.4
#| fig-width: 9
#| fig-format: svg

res <- psrc_hts_stat(df_hts_data, "person", 
                     c("work_mode_simple"),
                     incl_na=FALSE)

ggplot(res,
       aes(x=factor(survey_year), y=prop, fill=factor(survey_year))) +
  geom_col(width = 0.8) +
  geom_text(aes(label=scales::percent(prop, accuracy = 1)),
            vjust = -0.3) +
  errorbar_switch(geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                                width = 0.1), 
                  on=params$errorbar) +
  geom_hline(yintercept = 0, color = "#cbcbcb", linewidth = 0.3) +
  facet_wrap(~work_mode_simple, ncol=4, 
             # scale = "free"
             )+
  scale_fill_manual(values = psrc_colors$greens_dec)+
  scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, 0.1))) +
  psrc_style() +
  theme(legend.position = "none",
        strip.text = element_text(size = 11),
        panel.grid.major.y = element_blank(),
        panel.spacing.x = unit(0.5, "lines"),
        axis.title = element_blank(),
        axis.text.y = element_blank())
```

```{r}
res %>%
    # function sourced from Analysis/util.R
    get_table_format()
```


```{r}
  
res_region <- psrc_hts_stat(df_hts_data, "person", 
                            c("telecommute_freq_1plus"),
                            incl_na=FALSE) %>%
  filter(telecommute_freq_1plus == "telecommute at least 1 day/week") %>%
  mutate(group_var = "Region Total")

stat_telework_type <- function(group_var){
  
  res <- psrc_hts_stat(df_hts_data, "person", 
                       c(group_var,"telecommute_freq_1plus"),
                       incl_na=FALSE) %>%
    filter(telecommute_freq_1plus == "telecommute at least 1 day/week") %>%
    mutate(group_var = !!sym(group_var)) 
  
  res <- res %>% 
    select(names(res_region)) %>% 
    add_row(res_region %>% filter(survey_year %in% res$survey_year)) %>%
    filter(group_var != "Prefer not to answer") %>%
    mutate(group_var = ordered(group_var, levels = c("Region Total",levels(res$group_var))))

  return(res)
}

plot_telework_type <- function(group_var, type="both", 
                               color_year=c("#4C4C4C", "#91268F", "#8CC63E", "#F05A28", "#00A7A0"),
                               color_cat=psrc_colors$blues_dec){
  
  res <- stat_telework_type(group_var)
  
  if( type %in% c("year","both") ){
    
    # print your charts
    print(
      ggplot(res, 
             aes(x=telecommute_freq_1plus, y=prop, fill=group_var)) +
        geom_col(position = position_dodge(),width = 0.95) +
        geom_text(aes(label=scales::percent(prop, accuracy = 1)),
                      position = position_dodge(0.95),
                  vjust = -0.3, 
                  size = 10, size.unit="pt") +
        errorbar_switch(geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                                      position = position_dodge(0.95),
                                      width = 0.1), 
                        on=params$errorbar) +
        facet_wrap(~ survey_year, ncol=5)+
        geom_hline(yintercept = 0, color = "#cbcbcb", linewidth = 0.3) +
        scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, 0.1))) +
        scale_fill_manual(values = color_year)+
        psrc_style() +
        theme(panel.grid.major.y = element_blank(),
              legend.text = element_text(size = 11),
              axis.title = element_blank(),
              axis.text.x = element_blank(),
              axis.text.y = element_blank())
      )
  }
  
  if( type %in% c("cat","both") ){
    
    print(
      ggplot(res, 
             aes(x=factor(survey_year), y=prop, fill=factor(survey_year))) +
        geom_col(width = 0.8) +
        geom_text(aes(label=scales::percent(prop, accuracy = 1)),
                  vjust = -0.3, 
                  size = 10.5, size.unit="pt") +
        errorbar_switch(geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                                      width = 0.1), 
                        on=params$errorbar) +
        facet_wrap(~group_var, ncol=5)+
        geom_hline(yintercept = 0, color = "#cbcbcb", linewidth = 0.3) +
        scale_fill_manual(values = color_cat)+
        scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, 0.1))) +
        psrc_style() +
        theme(legend.position = "none",
              strip.text = element_text(size = 11),
              panel.grid.major.y = element_blank(),
              panel.spacing.x = unit(0.5, "lines"),
              axis.title = element_blank(),
              axis.text.x = element_text(size = 10.5),
              axis.text.y = element_blank())
    )    
    
  }
  
}

tbl_telework_type <- function(group_var){
  
  # result table with formatting
  stat_telework_type(group_var) %>%
    select("survey_year","group_var","telecommute_freq_1plus","count","prop","prop_moe","est","est_moe") %>%
    # function sourced from Analysis/util.R
    get_table_format()
  
}
```

## telework by demographic characteristics

- share of workers that are hybrid or fully at home
- strong correlation with income and working from home



### income (bins: 100K,200K)

- income with (100K/200K) bins have more even person counts than (50K/100K) bins

```{r}
#| label: fig-telework-income1
#| fig-cap: "[year] Telework by Income"
#| fig-asp: 0.618
#| fig-width: 7
#| fig-format: svg

plot_telework_type("income_100_200","year")
```

```{r}
#| label: fig-telework-income2
#| fig-cap: "[category] Telework by Income"
#| fig-asp: 0.618
#| fig-width: 6.5
#| fig-format: svg

plot_telework_type("income_100_200","cat", color_cat = psrc_colors$blues_dec[4:5])
```

```{r}
tbl_telework_type("income_100_200")
```

### income (bins: 50K,100K)

```{r}
#| label: fig-telework-income3
#| fig-cap: "[year] Telework by Income"
#| fig-asp: 0.5
#| fig-width: 7.5
#| fig-format: svg

plot_telework_type("income_50_100","year")
```

```{r}
#| label: fig-telework-income4
#| fig-cap: "[category] Telework by Income"
#| fig-asp: 0.4
#| fig-width: 8
#| fig-format: svg

plot_telework_type("income_50_100","cat")
```

```{r}
tbl_telework_type("income_50_100")
```


### industry

```{r}
#| label: fig-telework-industry1
#| fig-cap: "[year] Telework by Industry"
#| fig-asp: 0.3
#| fig-width: 11.5
#| fig-format: svg

plot_telework_type("industry_sector","year")
```

```{r}
#| label: fig-telework-industry2
#| fig-cap: "[category] Telework by Industry"
#| fig-asp: 0.3
#| fig-width: 12.5
#| fig-format: svg

plot_telework_type("industry_sector","cat", color_cat = psrc_colors$blues_dec[3:5])
```

```{r}
tbl_telework_type("industry_sector")
```



## telework by home/work location

- share of teleworkers: King > Pierce and Kitsap (significant)


### home location (incl. fully in person, hybrid and fully at home)


```{r}
#| label: fig-telework-homefull
#| fig-cap: "Worker Type by Home Location"
#| fig-asp: 0.7
#| fig-width: 6
#| fig-format: svg

df_plot <- psrc_hts_stat(df_hts_data, "person", 
                         c("home_county","telecommute_status"),
                         # exclude NAs in calculation: only considering workers
                         incl_na=FALSE) %>%
  filter(survey_year==2025) %>%
  group_by(home_county) %>%
  mutate(
    prop_position = case_when(telecommute_status=="Fully At Home"~0,
                              telecommute_status=="Hybrid"~.data$prop[1],
                              telecommute_status=="Fully In Person"~sum(.data$prop[c(1,2)]))
  )

ggplot(df_plot, 
       aes(x=home_county, y=prop, fill=fct_rev(telecommute_status))) +
  geom_col(width = 0.7) +
  geom_text(aes(y= prop_position, label=scales::percent(prop, accuracy = 1),
                color=telecommute_status),
            vjust = -0.6) +
  geom_hline(yintercept = c(0,1), color = "#cbcbcb", linewidth = 0.3) +
  scale_fill_manual(values = c("#4C4C4C","#91268F","#F05A28"))+
  scale_color_manual(values = c("#4C4C4C","white","white"), guide = "none")+
  scale_y_continuous(labels = scales::percent) +
  psrc_style() +
    theme(axis.title = element_blank(),
          panel.grid.major.y = element_blank(),
          axis.text.y = element_blank())
```

```{r}
df_plot %>%
    # function sourced from Analysis/util.R
    get_table_format() 
```


### home location (percentage of workers working at least one day)

```{r}
#| label: fig-telework-homeloc1
#| fig-cap: "[year] Telework by Home Location"
#| fig-asp: 0.4
#| fig-width: 9
#| fig-format: svg

plot_telework_type("home_county","year")
```

```{r}
#| label: fig-telework-homeloc2
#| fig-cap: "[category] Telework by Home Location"
#| fig-asp: 0.35
#| fig-width: 10.5
#| fig-format: svg

plot_telework_type("home_county","cat")
```

```{r}
tbl_telework_type("home_county")
```


### work location (incl. fully in person and hybrid)

-  workers working fully at home don't have work location outside home


```{r}
#| label: fig-telework-workfull
#| fig-cap: "Worker Type by Work Location"
#| fig-asp: 0.6
#| fig-width: 6
#| fig-format: svg

df_plot <- psrc_hts_stat(df_hts_data, "person", 
                         c("work_county","telecommute_status"),
                         # exclude NAs in calculation: only considering workers
                         incl_na=FALSE) %>%
  filter(survey_year==2025) %>%
  group_by(work_county) %>%
  mutate(
    prop_position = case_when(telecommute_status=="Hybrid"~0,
                              telecommute_status=="Fully In Person"~.data$prop[1])
  )

ggplot(df_plot, 
       aes(x=work_county, y=prop, fill=fct_rev(telecommute_status))) +
  geom_col(width = 0.7) +
  geom_text(aes(y= prop_position, label=scales::percent(prop, accuracy = 1),
                color=telecommute_status),
            vjust = -0.6) +
  geom_hline(yintercept = c(0,1), color = "#cbcbcb", linewidth = 0.3) +
  scale_fill_manual(values = c("#4C4C4C","#91268F","#F05A28"))+
  scale_color_manual(values = c("white","white"), guide = "none")+
  scale_y_continuous(labels = scales::percent) +
  psrc_style() +
    theme(axis.title = element_blank(),
          panel.grid.major.y = element_blank(),
          axis.text.y = element_blank())
```

### work location

-  workers working fully at home don't have work location outside home

```{r}
#| label: fig-telework-workloc1
#| fig-cap: "[year] Telework by Work Location"
#| fig-asp: 0.4
#| fig-width: 9
#| fig-format: svg

plot_telework_type("work_county","year")
```

```{r}
#| label: fig-telework-workloc2
#| fig-cap: "[category] Telework by Work Location (excluding always work at home)"
#| fig-asp: 0.35
#| fig-width: 10.5
#| fig-format: svg

plot_telework_type("work_county","cat")
```

```{r}
tbl_telework_type("work_county")
```





## Travel behaviors

### Workweek trips by telework status

- Fully in person workers make fewer "Social/Recreation/Meal" non-mandatory trips

```{r}
#| label: fig-telework-triprate
#| fig-cap: "Workweek trips by telework"
#| fig-asp: 0.4
#| fig-width: 8
#| fig-format: svg
# number of people by telecommute status
df_telecommute_status <- psrc_hts_stat(df_hts_data, "person",
                                       c("telecommute_status"),
                                       # exclude NAs in calculation: only considering workers
                                       incl_na=FALSE) %>%
  filter(survey_year == 2025) %>%
  rename(n_person = est) %>%
  select(telecommute_status, n_person)

# trip rate by person
res <- psrc_hts_stat(df_hts_data, "trip", 
                     c("telecommute_status", "dest_purpose_rate"),
                     # exclude NAs in calculation: only considering workers
                     incl_na=FALSE) %>%
  filter(survey_year == 2025) %>%
  rename(n_trip = est) %>%
  select(telecommute_status, dest_purpose_rate, n_trip) %>%
  left_join(df_telecommute_status, by = "telecommute_status") %>%
  mutate(trip_rate = (n_trip * 5) /n_person) %>%
  arrange(dest_purpose_rate) %>%
  group_by(telecommute_status) %>%
  mutate(
    trip_rate_position = case_when(dest_purpose_rate=="Errands/Shopping/School/Other"~.data$trip_rate[1],
                                   dest_purpose_rate=="Social/Recreation/Meal"~sum(.data$trip_rate[c(1,2)]),
                                   dest_purpose_rate=="Work"~sum(.data$trip_rate[c(1,2,3)]))
  )

ggplot(res, 
       aes(x=telecommute_status, y=trip_rate, fill=fct_rev(dest_purpose_rate))) +
  geom_col(width = 0.7) +
  geom_text(aes(y= trip_rate_position, label=scales::number(trip_rate, accuracy = 0.1),
                color=dest_purpose_rate),
            hjust = 1.1) +
  geom_hline(yintercept = 0, color = "#cbcbcb", linewidth = 0.3) +
  scale_fill_manual(values = c("#F05A28","#00A7A0","#8CC63E"))+
  scale_color_manual(values = c("#4C4C4C","#4C4C4C","#4C4C4C"), guide = "none")+
  coord_flip()+
  psrc_style() +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        panel.grid.major.y = element_blank())

res %>%
  mutate(trip_rate = scales::number(trip_rate, accuracy=0.01)) %>%
  pivot_wider(id_cols="telecommute_status", names_from="dest_purpose_rate", values_from="trip_rate")
```



### All trip mode share by telework status

- workers who work fully at home have significantly higher walk mode share
- hybrid workers have the highest transit mode share

```{r}
#| label: fig-telework-modeshare
#| fig-cap: "Trip mode share by worker type"
#| fig-asp: 0.4
#| fig-width: 9
#| fig-format: svg

res <- psrc_hts_stat(df_hts_data, "trip", 
                     c("telecommute_status", "mode_class_5"),
                     # exclude NAs in calculation: only considering workers
                     incl_na=FALSE) %>%
  filter(survey_year==2025) %>%
  group_by(telecommute_status) %>%
  mutate(
    prop_position = case_when(mode_class_5=="Drive"~0,
                              mode_class_5=="Transit"~.data$prop[1],
                              mode_class_5=="Walk"~sum(.data$prop[c(1,2)]),
                              mode_class_5=="Bike/Micromobility"~sum(.data$prop[c(1,2,3)]),
                              mode_class_5=="Other"~sum(.data$prop[c(1,2,3,4)]))
    ) #%>%
  # mutate(prop = scales::percent(prop, accuracy=0.01),
  #        mode_class_5 = factor(mode_class_5, levels=c("Drive", "Walk", "Transit", "Bike/Micromobility", "Other"))) %>%
  # arrange(mode_class_5) %>%
  # pivot_wider(id_cols = "telecommute_status3", names_from = "mode_class_5", values_from = "prop")


ggplot(res, 
       aes(x=telecommute_status, y=prop, fill=fct_rev(mode_class_5))) +
  geom_col(width = 0.7) +
  geom_text(aes(y= prop_position, label=scales::percent(prop, accuracy = 1),
                color=mode_class_5),
            hjust = -0.05) +
  geom_hline(yintercept = c(0,1), color = "#cbcbcb", linewidth = 0.3) +
  scale_fill_manual(values = c("#F05A28","#00A7A0","#91268F","#8CC63E","#4C4C4C"))+
  scale_color_manual(values = c("white","#4C4C4C","white","#4C4C4C","#4C4C4C"), guide = "none")+
  scale_y_continuous(labels = scales::percent) +
  coord_flip()+
  psrc_style() +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        panel.grid.major.y = element_blank())
```

### VMT per workweek by telework status

- VMT per workweek: Fully At Home < Hybrid < Fully In Person

```{r}
psrc_hts_vmtrate(df_hts_data, "telecommute_status", incl_na=FALSE) %>%
  mutate(workweek_vmt = mean * 5,
         workday_vmt = mean) %>%
  filter(survey_year == 2025) %>%
  select(survey_year,telecommute_status,workday_vmt,workweek_vmt)
```

