---
title: "Telework Analysis"
date: today
---

```{r}
#| include: false

# get data and load standard data processing
source("telework_processing.R")  

hts_source <- "Puget Sound Regional Council (PSRC) 2017-2025 Household Travel Survey"

plot_formatting <- function(colors, title){
  list(
    scale_fill_manual(values = colors),
    scale_y_continuous(labels = scales::percent),
    psrc_style(),
    theme(panel.grid.major.y = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_blank()),
    labs(caption = hts_source,
         title = title)
    )
}

```

## Telework definition

- teleworkers are the workers who telework more than one day per week or work fully at home
- 2025 telework status assignments

```{r}
psrc_hts_stat(df_hts_data, "person", 
              c("employment_simple","work_from_home","workplace_simple","telecommute_freq_1plus","telecommute_status"),
              # exclude NAs in calculation: only considering workers
              incl_na=TRUE) %>%
  # function sourced from Analysis/util.R
  get_table_format() %>%
  arrange(employment_simple,telecommute_status,work_from_home,workplace_simple,telecommute_freq_1plus) %>%
  filter(survey_year == 2025) %>%
  select(employment_simple,workplace_simple,telecommute_freq_1plus,telecommute_status,count)
```


## Telework trend

- trend remained similar since 2023

```{r}
#| out-width: "50%"

df_plot <- psrc_hts_stat(df_hts_data, "person", 
                         c("telecommute_status"),
                         # exclude NAs in calculation: only considering workers
                         incl_na=FALSE) %>%
  group_by(survey_year) %>%
  mutate(
    prop_position = case_when(telecommute_status=="Fully At Home"~0,
                              telecommute_status=="Hybrid"~.data$prop[1],
                              telecommute_status=="Fully In Person"~sum(.data$prop[c(1,2)]))
  )

ggplot(df_plot, 
       aes(x=factor(survey_year), y=prop, fill=fct_rev(telecommute_status))) +
  geom_col(width = 0.7) +
  geom_text(aes(y= prop_position, label=scales::percent(prop, accuracy = 1),
                color=telecommute_status),
            vjust = -0.6) +
  scale_fill_manual(values = c("#4C4C4C","#91268F","#F05A28"))+
  scale_color_manual(values = c("#4C4C4C","white","white"), guide = "none")+
  scale_y_continuous(labels = scales::percent) +
  psrc_style() +
  theme(axis.title = element_blank())+
  labs(caption = hts_source, title = "Telework trend")
```



## Commute mode share

- Active Modes include walk, bike and micromobility

```{r}
#| out-width: "50%"
#| fig-width: 5
res <- psrc_hts_stat(df_hts_data, "person", 
                     c("work_mode_simple"),
                     incl_na=FALSE)

ggplot(res, 
       aes(x=work_mode_simple, y=prop, fill=factor(survey_year))) +
  geom_col(position = position_dodge(),width = 0.8) +
  geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                position = position_dodge(0.8),
                width = 0.4) +
  facet_wrap(~work_mode_simple, ncol=5, scales = "free")+
  plot_formatting(psrc_colors$blues_inc[-1],"Commute mode trend: excluding work at home workers ")

res %>%
    # function sourced from Analysis/util.R
    get_table_format()
```



```{r}

stat_telework_type <- function(group_var, title){
  
  res <- psrc_hts_stat(df_hts_data, "person", 
                       c(group_var,"telecommute_freq_1plus"),
                       incl_na=FALSE) %>%
    filter(telecommute_freq_1plus == "telecommute at least 1 day/week") %>%
    mutate(group_var = !!sym(group_var))
  
  res2 <- psrc_hts_stat(df_hts_data, "person", 
                       c("telecommute_freq_1plus"),
                       incl_na=FALSE) %>%
    filter(telecommute_freq_1plus == "telecommute at least 1 day/week",
           survey_year %in% res$survey_year) %>%
    mutate(group_var = "Region")
  res <- res %>% 
    select(names(res2)) %>% 
    add_row(res2) %>%
    filter(group_var != "Prefer not to answer") %>%
    mutate(group_var = ordered(group_var, levels = c("Region",levels(res$group_var))))
    
  # result table with formatting
  df_plot <- res %>%
    # function sourced from Analysis/util.R
    get_table_format() %>%
    select(survey_year,group_var,count,prop,prop_moe,est)
  
  # print your charts
  print(
    ggplot(res, 
           aes(x=telecommute_freq_1plus, y=prop, fill=group_var)) +
      geom_col(position = position_dodge(),width = 0.8) +
      geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                    position = position_dodge(0.8),
                    width = 0.4) +
      facet_wrap(~ survey_year, ncol=5)+
      plot_formatting(c("#4C4C4C", "#91268F", "#8CC63E", "#F05A28", "#00A7A0", "#630460", "#588527", "#9f3913", "#00716c", "#3e4040"),
                      title)
  )
  
  print(
    ggplot(res, 
           aes(x=telecommute_freq_1plus, y=prop, fill=factor(survey_year))) +
      geom_col(position = position_dodge(),width = 0.8) +
      geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                    position = position_dodge(0.8),
                    width = 0.4) +
      facet_wrap(~group_var, ncol=4)+
      plot_formatting(psrc_colors$blues_inc[-1],title)
  )
  
  
  # print result table by calling df_plot
  return(df_plot)
}
```

## telework by demographic characteristics

- share of workers that are hybrid or fully at home
- strong correlation with income and working from home


:::{.panel-tabset}

### income (100K/200K)

- income with (100K/200K) bins have more even person counts than (50K/100K) bins

```{r}
#| layout: [[50, 50], [100]]
#| out-width: "50%"
stat_telework_type("income_100_200","telework by income")

```

### income (50K/100K)

```{r}
#| layout: [[50, 50], [100]]
#| out-width: "50%"
stat_telework_type("income_50_100","telework by income")

```

### industry

```{r}
#| fig-width: 6
#| layout: [[50, 50], [100]]
#| out-width: "50%"
stat_telework_type("industry_sector","telework by industry")
```

:::

## telework by home/work location

- share of teleworkers: King > Pierce > Snohomish > Kitsap

:::{.panel-tabset}

### home location (all types)


```{r}
#| out-width: "50%"
df_plot <- psrc_hts_stat(df_hts_data, "person", 
                         c("home_county","telecommute_status"),
                         # exclude NAs in calculation: only considering workers
                         incl_na=FALSE) %>%
  filter(survey_year==2025) %>%
  group_by(home_county) %>%
  mutate(
    prop_position = case_when(telecommute_status=="Fully At Home"~0,
                              telecommute_status=="Hybrid"~.data$prop[1],
                              telecommute_status=="Fully In Person"~sum(.data$prop[c(1,2)]))
  )

ggplot(df_plot, 
       aes(x=home_county, y=prop, fill=fct_rev(telecommute_status))) +
  geom_col(width = 0.7) +
  geom_text(aes(y= prop_position, label=scales::percent(prop, accuracy = 1),
                color=telecommute_status),
            vjust = -0.6) +
  scale_fill_manual(values = c("#4C4C4C","#91268F","#F05A28"))+
  scale_color_manual(values = c("#4C4C4C","white","white"), guide = "none")+
  scale_y_continuous(labels = scales::percent) +
  psrc_style() +
  theme(axis.title = element_blank())+
  labs(caption = hts_source, title = "Telework status by home location")

df_plot %>%
    # function sourced from Analysis/util.R
    get_table_format() %>%
    select(survey_year,home_county,telecommute_status,count,prop,prop_moe,est)
```

### home location (at least one day)

```{r}
#| layout: [[50, 50], [100]]
#| out-width: "50%"
#| fig-width: 5
stat_telework_type("home_county","telework by home location")

```

### work location

```{r}
#| layout: [[50, 50], [100]]
#| out-width: "50%"
#| fig-width: 5
stat_telework_type("work_county","telework by work location: excluding always work at home")

```

:::

```{r}
# work distance

# rs1 <- psrc_hts_stat(df_hts_data,
#                      analysis_unit = "trip",
#                      group_vars = c("telecommute_status","dest_purpose_cat", "distance_bins"),
#                      incl_na = FALSE) %>%
#   filter(telecommute_status != "Fully At Home",
#          dest_purpose_cat == "Work")
```





## Travel behaviors

### Workweek trips by telework status

- Fully in person workers make fewer "Social/Recreation/Meal" non-mandatory trips

```{r}
#| out-width: "50%"
# number of people by telecommute status
df_telecommute_status <- psrc_hts_stat(df_hts_data, "person",
                                       c("telecommute_status"),
                                       # exclude NAs in calculation: only considering workers
                                       incl_na=FALSE) %>%
  filter(survey_year == 2025) %>%
  rename(n_person = est) %>%
  select(telecommute_status, n_person)

# trip rate by person
res <- psrc_hts_stat(df_hts_data, "trip", 
                     c("telecommute_status", "dest_purpose_rate"),
                     # exclude NAs in calculation: only considering workers
                     incl_na=FALSE) %>%
  filter(survey_year == 2025) %>%
  rename(n_trip = est) %>%
  select(telecommute_status, dest_purpose_rate, n_trip) %>%
  left_join(df_telecommute_status, by = "telecommute_status") %>%
  mutate(trip_rate = (n_trip * 5) /n_person) %>%
  arrange(dest_purpose_rate) %>%
  group_by(telecommute_status) %>%
  mutate(
    trip_rate_position = case_when(dest_purpose_rate=="Errands/Shopping/School/Other"~.data$trip_rate[1],
                                   dest_purpose_rate=="Social/Recreation/Meal"~sum(.data$trip_rate[c(1,2)]),
                                   dest_purpose_rate=="Work"~sum(.data$trip_rate[c(1,2,3)]))
  )

ggplot(res, 
       aes(x=telecommute_status, y=trip_rate, fill=fct_rev(dest_purpose_rate))) +
  geom_col(width = 0.7) +
  geom_text(aes(y= trip_rate_position, label=scales::number(trip_rate, accuracy = 0.1),
                color=dest_purpose_rate),
            hjust = 1.1) +
  scale_fill_manual(values = c("#F05A28","#00A7A0","#8CC63E"))+
  scale_color_manual(values = c("#4C4C4C","#4C4C4C","#4C4C4C"), guide = "none")+
  coord_flip()+
  psrc_style() +
  theme(axis.title = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(color = "gray80", linewidth = 0.5),)+
  labs(caption = hts_source, title = "Workweek trips by telework")

res %>%
  mutate(trip_rate = scales::number(trip_rate, accuracy=0.01)) %>%
  pivot_wider(id_cols="telecommute_status", names_from="dest_purpose_rate", values_from="trip_rate")
```



### All trip mode share by telework status

- workers who work fully at home have significantly higher walk mode share
- hybrid workers have the highest transit mode share

```{r}
#| out-width: "50%"
#| fig-width: 5
res <- psrc_hts_stat(df_hts_data, "trip", 
                     c("telecommute_status", "mode_class_5"),
                     # exclude NAs in calculation: only considering workers
                     incl_na=FALSE) %>%
  filter(survey_year==2025) %>%
  group_by(telecommute_status) %>%
  mutate(
    prop_position = case_when(mode_class_5=="Drive"~0,
                              mode_class_5=="Transit"~.data$prop[1],
                              mode_class_5=="Walk"~sum(.data$prop[c(1,2)]),
                              mode_class_5=="Bike/Micromobility"~sum(.data$prop[c(1,2,3)]),
                              mode_class_5=="Other"~sum(.data$prop[c(1,2,3,4)]))
    ) #%>%
  # mutate(prop = scales::percent(prop, accuracy=0.01),
  #        mode_class_5 = factor(mode_class_5, levels=c("Drive", "Walk", "Transit", "Bike/Micromobility", "Other"))) %>%
  # arrange(mode_class_5) %>%
  # pivot_wider(id_cols = "telecommute_status3", names_from = "mode_class_5", values_from = "prop")


ggplot(res, 
       aes(x=telecommute_status, y=prop, fill=fct_rev(mode_class_5))) +
  geom_col(width = 0.7) +
  geom_text(aes(y= prop_position, label=scales::percent(prop, accuracy = 1),
                color=mode_class_5),
            hjust = -0.05) +
  scale_fill_manual(values = c("#F05A28","#00A7A0","#91268F","#8CC63E","#4C4C4C"))+
  scale_color_manual(values = c("white","#4C4C4C","white","#4C4C4C","#4C4C4C"), guide = "none")+
  scale_y_continuous(labels = scales::percent) +
  coord_flip()+
  psrc_style() +
  theme(axis.title = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(color = "gray80", linewidth = 0.5),)+
  labs(caption = hts_source, title = "Telework status by home location")
```

### VMT per workweek by telework status

- VMT per workweek: Fully At Home < Hybrid < Fully In Person

```{r}
psrc_hts_vmtrate(df_hts_data, "telecommute_status", incl_na=FALSE) %>%
  mutate(workweek_vmt = mean * 5,
         workday_vmt = mean) %>%
  filter(survey_year == 2025) %>%
  select(survey_year,telecommute_status,workday_vmt,workweek_vmt)
```

