---
title: "telework_for_presentation"
author: "suzanne"
date: "2024-04-08"
output: html_document
---


```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
library(forcats)
source('../survey-23-preprocess.R')
```


```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad,home_jurisdiction,survey_year, hhsize, hh_race_category, numchildren,hh_weight
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category, employment,
person_id, workplace, sexuality, person_weight, gender, age, industry, survey_year
                from HHSurvey.v_persons_labels")
                
#day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , #survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,
person_id, mode_1, dest_purpose, survey_year, trip_weight 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)

setDT(trip)
```
#race_category needs to be added to the codebook



## Read in Codebook
```{r pressure, echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```

## Set IDs as characters
TO DO: functionalize convert all ids to characters, or store as characters upfront
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]


trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))%>%filter(survey_year=='2023')
person <- person%>%mutate(survey_year=as.character(survey_year))%>%filter(survey_year=='2023')
trip <- trip%>%mutate(survey_year=as.character(survey_year))%>%filter(survey_year=='2023')

hh<-order_factors(hh, 'hhincome_broad', value_labels)


```


```{r}
variable_list<-add_variable(variable_list, 'sexuality_broad','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='sexuality_broad')
value_labels<-add_values_code(group_name='sexuality_broad')
person<-grp_to_tbl(tbl=person, ungrouped_name='sexuality', grouped_name='sexuality_broad')
```


```{r}
variable_list<-add_variable(variable_list, 'gender_grp','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='gender_grp')
value_labels<-add_values_code(group_name='gender_grp')
```

```{r}
person<-grp_to_tbl(tbl=person, ungrouped_name='gender', grouped_name='gender_grp')
```



```{r}
variable_list<-add_variable(variable_list, 'age_grp_2','person')
group_labels<-get_grouped_labels(group_id='group_3', group_name='age_grp_2')
value_labels<-add_values_code(group_name='age_grp_2')
```



check that survey is picking up demographics compare to pums first
```{r}
person<-grp_to_tbl(tbl=person, ungrouped_name='age', grouped_name='age_grp_2')
```



NOTE THIS HAS NO Non-response filtered- should be everyone
```{r}
hts_data = list(hh = hh,
                person = person,
                trip = trip)
id_cols=c('hh_id', 'person_id', 'trip_id')
wt_cols<- c('hh_weight', 'person_weight', 'trip_weight')
```

```{r}

summary = summarize_weighted(hts_data= hts_data,
                               summarize_var = 'age',
                               summarize_by=NULL,
                               id_cols=id_cols,
                               wt_cols=wt_cols,
                               wtname='person_weight'
                               )
write.csv(summary$summary$wtd, 'age.csv')
```
```{r}

```



```{r}

summary = summarize_weighted(hts_data= hts_data,
                               summarize_var = 'gender',
                               summarize_by=NULL,
                               id_cols=id_cols,
                               wt_cols=wt_cols,
                               wtname='person_weight'
                               )
write.csv(summary$summary$wtd,'gender.csv')
```

```{r}
summary = summarize_weighted(hts_data= hts_data,
                               summarize_var = 'sexuality',
                               summarize_by=NULL,
                               id_cols=id_cols,
                               wt_cols=wt_cols,
                               wtname='person_weight'
                               )
summary$summary$wtd
```


NOTE FILTERING OUT NONRESPONSE
```{r}
person<-person%>%filter(sexuality!='Prefer not to answer')
```



```{r}
hts_data = list(hh = hh,
                person = person,
                trip = trip)
id_cols=c('hh_id', 'person_id', 'trip_id')
wt_cols<- c('hh_weight', 'person_weight', 'trip_weight')
```

```{r}
hts_data_filter_race = list(hh = hh,
                person = person,
                trip = trip)
id_cols=c('hh_id', 'person_id', 'trip_id')
wt_cols<- c('hh_weight', 'person_weight', 'trip_weight')
```
Overall
```{r}

summary_overall = summarize_weighted(hts_data= hts_data,
                               summarize_var = 'sexuality',
                               summarize_by=NULL,
                               id_cols=id_cols,
                               wt_cols=wt_cols,
                               wtname='person_weight'
                               )
```

```{r}
summary_info<-summary_overall$summary$wtd%>%mutate(prop_moe=prop_se*1.645)%>%mutate(sexuality= fct_reorder(sexuality,-prop))%>%mutate(est=round(est,-4))
static_bar_chart(summary_info, x='prop', y='sexuality', fill='sexuality')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
```
```{r}
summary_info<-summary_overall$summary$wtd%>%mutate(est_moe=est_se*1.645)%>%mutate(sexuality= fct_reorder(sexuality,-prop))%>%mutate(est_total=round(3380000*prop, -4))%>%filter(sexuality!='Heterosexual (straight)')
static_bar_chart(summary_info, x='est_total', y='sexuality', fill='sexuality')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
```

Gender


```{r}
person_filter_gender<-person%>%filter(gender!='Prefer not to answer')
```

```{r}
hts_data_filter_gender = list(hh = hh,
                person = person_filter_gender,
                trip = trip)
id_cols=c('hh_id', 'person_id', 'trip_id')
wt_cols<- c('hh_weight', 'person_weight', 'trip_weight')
```

```{r}
summary = summarize_weighted(hts_data= hts_data_filter_gender,
                               summarize_var = 'gender_grp',
                               summarize_by='sexuality_broad',
                               id_cols=id_cols,
                               wt_cols=wt_cols,
                               wtname='person_weight'
                               )
```

```{r}
summary_info<-summary$summary$wtd%>%mutate(prop_moe=prop_se*1.645)
static_column_chart(summary_info, y='prop', x='sexuality_broad', fill='gender_grp', moe='prop_moe')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
```
Age





```{r}

summary = summarize_weighted(hts_data= hts_data,
                               summarize_var = 'age_grp_2',
                               summarize_by='sexuality',
                               id_cols=id_cols,
                               wt_cols=wt_cols,
                               wtname='person_weight'
                               )
summary$summary$wtd

```


Income

```{r}
hh_filter_inc<-hh%>%filter(hhincome_broad!='Prefer not to answer')
hts_data_filter_inc = list(hh = hh_filter_inc,
                person = person,
                trip = trip)
id_cols=c('hh_id', 'person_id', 'trip_id')
wt_cols<- c('hh_weight', 'person_weight', 'trip_weight')
```

```{r}

summary = summarize_weighted(hts_data= hts_data_filter_inc,
                               summarize_var = 'hhincome_broad',
                               summarize_by = c('sexuality_broad'),
                               id_cols=id_cols,
                               wt_cols=wt_cols,
                               wtname='person_weight'
                               )
```



```{r}

summary_w_moe<-summary$summary$wtd%>%mutate(prop_moe=prop_se*1.645)

static_column_chart(summary_w_moe, y='prop', x='sexuality_broad', fill='hhincome_broad', moe='prop_moe' )+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
```

Children in Household

Gender
```{r}

summary = summarize_weighted(hts_data= hts_data,
                               summarize_var = 'numchildren',
                               summarize_by='sexuality_broad',
                               id_cols=id_cols,
                               wt_cols=wt_cols,
                               wtname='person_weight'
                               )
summary$summary$wtd
```

Race
```{r}
person_filter_race<-person%>%filter(race_category!='Child')%>%filter(race_category!='Missing')%>%drop_na(race_category)%>%filter(sexuality_broad!='Missing')%>%drop_na(sexuality_broad)
```






```{r}
hts_data_filter_race = list(hh = hh,
                person = person_filter_race,
                trip = trip)
id_cols=c('hh_id', 'person_id', 'trip_id')
wt_cols<- c('hh_weight', 'person_weight', 'trip_weight')
```


```{r}
summary = summarize_weighted(hts_data= hts_data_filter_race,
                               summarize_var = 'race_category',
                               summarize_by = c('sexuality_broad'),
                               id_cols=id_cols,
                               wt_cols=wt_cols,
                               wtname='person_weight'
                               )
```





```{r}
summary$summary$wtd
summary_w_moe<-summary$summary$wtd%>%mutate(prop_moe=prop_se*1.645)%>%filter(sexuality_broad!='Heterosexual (straight)')

static_column_chart(summary_w_moe, y='prop', x='sexuality_broad', fill='race_category', moe='prop_moe' )+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
```
```{r}

summary_w_moe<-summary$summary$wtd%>%mutate(prop_moe=prop_se*1.645)%>%filter(sexuality_broad!='Heterosexual (straight)')

```

Q/C against PUMS

```{r}
library(psrccensus)
library(dplyr)
library(magrittr)

pums5_22 <- get_psrc_pums(5,2022,"p",vars=c("AGEP","HUPAC","BINCOME","BIN_AGE"), 
  dir="J:/Projects/Census/AmericanCommunitySurvey/Data/PUMS/pums_rds")
pums5_22 %<>% mutate(
  child_present=case_when(grepl("^No", as.character(HUPAC)) ~"No",
                          grepl("^With", as.character(HUPAC)) ~"Yes",
                          is.na(HUPAC) ~ NA_character_))
rs <- list()
rs$"Adults by inc cat" <-
    psrc_pums_count(filter(pums5_22, AGEP>17), group_vars="BINCOME")
rs$"Adults by child in hh" <-
    psrc_pums_count(filter(pums5_22, AGEP>17), group_vars="child_present")
rs$"Adults by age cat" <-
    psrc_pums_count(filter(pums5_22, AGEP>17), group_vars="BIN_AGE")
```



```


