---
title: "Regional Growth Center Analysis using 2017, 2019, and 2021 travel surveys"
author: "Joanne Lin"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
---


```{r rmarkdown setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE) # formatting
```

```{r library setup, include=FALSE}
# psrc packages
library(psrc.travelsurvey)
library(psrcplot)
library(psrctrends)

library(tidyverse)
# library(stringr)
# library(tidycensus)
# library(htmlwidgets)

# install fonts from psrcplot
install_psrc_fonts()
```


# Variables we might want to summarize
```{r Variable, echo=TRUE}

gen_vars<-c('sample_county', 'vehicle_count',  "hhincome_broad",  
              'hhincome_detailed', "age", "age_category", 'final_home_is_rgc', 'race_category', 'race_eth_broad', 
              'dest_purpose_cat', 'origin_purpose_cat', 'trip_path_distance', 'person_id')

# remember to filter out drivers for hov trips
# not a problem = the weights are 0 anyways
trip_vars = c("hhid",
              "survey_year",
              # "personid",
              # "pernum",
              "hhincome_broad",
              "age", 
              "age_category",
              'final_home_is_rgc',
              "dayofweek",
              # "tripid",
              "depart_time_hhmm",
              "travelers_total",
              # "o_purpose",
              # "o_purp_cat",
              # "d_purpose",
              # "d_purp_cat",
              "mode_1",
              # "mode_type",
              "driver",
              "pool_start",
              "change_vehicles",
              "toll",
              "toll_pay",
              "taxi_pay",
              "mode_acc",
              "mode_egr",
              "speed_mph")


hh_vars=c("hhid",
          "survey_year",
          "final_home_is_rgc",
          "hhsize",
          "vehicle_count",
          "numadults",
          "numchildren",
          "numworkers",
          "lifecycle",
          "num_trips",  
          "res_factors_30min",
          "res_factors_afford",
          "res_factors_closefam",
          "res_factors_hhchange",
          "res_factors_hwy",
          "res_factors_school",
          "res_factors_space",
          "res_factors_transit",
          "res_factors_walk",
          "res_factors_cultural",  
          "res_dur",
          "res_type",
          "res_months",
          "broadband",
          "offpark",
          "offpark_cost",
          "streetpark")
```

Get the data from Elmer.
```{r trip + household data, include=FALSE}
trip_data_17_19<- get_hhts("2017_2019", "t", vars=trip_vars)
trip_data_21<- get_hhts("2021", "t", vars=trip_vars)

hh_data_17_19<- get_hhts("2017_2019", "h", vars=hh_vars)
hh_data_21<- get_hhts("2021", "h", vars=hh_vars)

```


Group variables into meaningful categories with statistically large enough groups.First for trips, group the data.
```{r variable names}
unique(hh_data_21$res_dur)
```
```{r}
names(hh_data_21)
```

```{r Group variables}
hh_group_data <- function(.data){
  .data <- .data %>%
    mutate(
      vehicle_count = substring(vehicle_count,1,1),
      vehicle_count = factor(case_when(vehicle_count==0 ~ "No vehicle", 
                                       vehicle_count==1 ~ "1",
                                       vehicle_count==2 ~ "2",
                                       vehicle_count %in% c(3,4,5,6,7,8)~ "3 or more")),
      hhsize = substring(hhsize,1,1),
      hhsize = case_when(hhsize %in% c(5,6,7,8,9)~ "5 or more",
                         TRUE ~ hhsize),
      res_type = factor(case_when(res_type == "Single-family house (detached house)"~ "Single-family house",
                                  res_type == "Townhouse (attached house)"~ "Townhouse",
                                  res_type %in% c("Building with 4 or more apartments/condos",
                                                  "Building with 3 or fewer apartments/condos")~ "Apartment/Condo",
                                  res_type %in% c("Other (including boat, RV, van, etc.)","Mobile home/trailer",
                                                  "Dorm or institutional housing")~ "Others")),
      res_dur = case_when(res_dur %in% c("Between 2 and 3 years", "Between 3 and 5 years", 
                                         "Between 5 and 10 years","Between 10 and 20 years",
                                         "More than 20 years")~ "More than 2 years",
                          TRUE ~ res_dur)

      )
  .data$hhsize <- factor(.data$hhsize, levels=c("1","2","3","4","5 or more"))
  .data$res_dur <- factor(.data$res_dur, levels=c("Less than a year","Between 1 and 2 years","More than 2 years"))
  return(.data)
}


hh_group_21 <- hh_data_21 %>% hh_group_data()
hh_group_17_19 <- hh_data_17_19 %>% hh_group_data()
```

```{r}
plot  <- hhts_count(hh_group_21, group_vars=c("final_home_is_rgc","vehicle_count")) %>%
  filter(vehicle_count!="Total") %>%
  add_row(hhts_count(hh_group_17_19, group_vars=c("final_home_is_rgc","vehicle_count")) %>%
            filter(vehicle_count!="Total"))

plot$vehicle_count <- factor(plot$vehicle_count, levels=c("No vehicle","1", "2", "3 or more"))
plot$final_home_is_rgc <- factor(plot$final_home_is_rgc, levels=c("RGC","Not RGC"))


chart <- create_facet_bar_chart (t=plot, w.x="vehicle_count",w.y="share", 
                                 f="survey", g= "final_home_is_rgc",
                                 w.title="Number of vehicles",
                                 # source="Source: Household travel survey",
                                 w.moe="share_moe", w.color="psrc_light")
chart
```
```{r}
plot  <- hhts_count(hh_group_21 %>% filter(final_home_is_rgc=="RGC"), group_vars=c("vehicle_count","res_dur")) %>%
  filter(res_dur!="Total") %>%
  add_row(hhts_count(hh_group_17_19 %>% filter(final_home_is_rgc=="RGC"), group_vars=c("vehicle_count","res_dur")) %>%
            filter(res_dur!="Total"))

plot$vehicle_count <- factor(plot$vehicle_count, levels=c("No vehicle","1", "2", "3 or more"))
plot$final_home_is_rgc <- factor(plot$final_home_is_rgc, levels=c("RGC","Not RGC"))


chart <- create_facet_bar_chart (t=plot, w.x="vehicle_count",w.y="share", 
                                 f="survey", g= "res_dur",
                                 w.title="Number of vehicles",
                                 # source="Source: Household travel survey",
                                 w.moe="share_moe", w.color="psrc_light")
chart
```


```{r visualization functions}
# visuals ----
# hhts_count and prep for graphing
count_prep_fx <- function(data, grp1, grp2) {
  
  hhts_count(data,
             group_vars = c(grp1, grp2)) %>%
    mutate(period = data$label) %>%
    filter(!!sym(grp2)!="Total") %>% 
    filter(!!sym(grp1)!="NA")
}
```

```{r}
plot_chart <- function(.data, vars) {
  .data %>% 
    group_by(final_home_is_rgc,survey) %>% 
    mutate(num_hh = n()) %>% 
    group_by(final_home_is_rgc, survey, num_hh, {{vars}}) %>% 
    summarise(counts = n()) %>%
    ungroup() %>%
    mutate(per = counts*100/num_hh)
}

plot_chart_w <- function(.data, vars) {
  .data %>% 
    group_by(final_home_is_rgc,survey) %>% 
    mutate(num_hh = sum(hh_weight)) %>% 
    group_by(final_home_is_rgc, survey, num_hh, {{vars}}) %>% 
    summarise(counts = sum(hh_weight)) %>%
    ungroup() %>%
    mutate(per = counts*100/num_hh)
}
```

```{r bar chart for number of vehicles}
plot_vehicle_count_w <- hh_data %>% plot_chart_w(vehicle_count)
plot_vehicle_count <- hh_data %>% plot_chart(vehicle_count)
plot_vehicle_count$vehicle_count <- factor(plot_vehicle_count$vehicle_count, 
                             levels=c("No vehicle","1","2","3","4 or more"))
plot_vehicle_count_w$vehicle_count <- factor(plot_vehicle_count_w$vehicle_count, 
                             levels=c("No vehicle","1","2","3","4 or more"))

ggplot(plot_vehicle_count, 
       aes(x= vehicle_count, y= per)) +
  geom_col(aes(fill=survey),position=position_dodge()) +
  facet_wrap(~final_home_is_rgc) +
  xlab("Number of vehicles in each household") + 
  ylab("Percentage of households (%)")+
  theme_bw()


```
```{r}
# res_dur==
ggplot(plot_vehicle_count_w, 
       aes(x= vehicle_count, y= per)) +
  geom_col(aes(fill=survey),position=position_dodge()) +
  facet_wrap(~final_home_is_rgc) +
  xlab("Number of vehicles in each household") + 
  ylab("Percentage of households (%)")+
  theme_bw()
```


```{r hhsize}

plot_hhsize <- hh_data %>% plot_chart(hhsize)

ggplot(plot_hhsize, 
       aes(x= hhsize, y= per)) +
  geom_col(aes(fill=survey),position=position_dodge()) +
  facet_wrap(~final_home_is_rgc) +
  xlab("Household Sizes") + 
  ylab("Percentage of households (%)")+
  theme_bw()

```

```{r resident type}
plot_res_type <- test %>% plot_chart(res_type)
plot_res_type$res_type <- factor(plot_res_type$res_type, 
                             levels=c("Single-family house", "Townhouse", "Apartment/Condo", "Others"))

ggplot(plot_res_type, 
       aes(x= res_type, y= per)) +
  geom_col(position=position_dodge()) +
  facet_wrap(~final_home_is_rgc) +
  xlab("Types of residence") + 
  ylab("Percentage of households (%)")+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust = 0.8)) 

```




```{r Group variables}
trip_data_17_19 <- trip_data_17_19 %>% 
  mutate(NoVehicles = ifelse(vehicle_count == '0 (no vehicles)', 'No Vehicles', "Has Vehicles")) %>%
  mutate(
    hhincome_100 = case_when(
                     hhincome_broad == '$100,000-$199,000' ~ '$100,000 or more',
                     hhincome_broad == '$200,000 or more' ~ '$100,000 or more',
                     TRUE ~ hhincome_broad
    )
  ) %>%
  mutate(
    age_grp = case_when(
                age == '75-84 years' ~ '75 years or older',
                age == '85 or years older' ~ '75 years or older',
                TRUE ~ age
    )
  ) %>% 
  mutate(simple_purpose = ifelse(dest_purpose_cat == 'Home',
                                 origin_purpose_cat,
                                 dest_purpose_cat
  )) %>%
  mutate(
    simple_purpose = case_when(
                        simple_purpose == 'Work' ~ 'Work/School',
                        simple_purpose == 'School' ~ 'Work/School',
                        simple_purpose == 'Work-related' ~ 'Work/School',
                        simple_purpose == 'Shop' ~ 'Shop',
                        simple_purpose == 'Escort' ~ 'Errands',
                        simple_purpose == 'Errand/Other' ~ 'Errands',
                        simple_purpose == 'Change mode' ~ 'Errands',
                        simple_purpose == 'Social/Recreation' ~ 'Social/Recreation/Meal',
                        simple_purpose == 'Meal' ~ 'Social/Recreation/Meal',
                        simple_purpose == 'Meal' ~ 'Social/Recreation/Meal',
                        simple_purpose == 'Home' ~ 'Errands',
                        is.na(simple_purpose) ~ 'Errands',
                        TRUE ~ simple_purpose
    )
  )


trip_data_21 <-
  trip_data_21 %>% mutate(NoVehicles = ifelse(vehicle_count == '0 (no vehicles)', 'No Vehicles', "Has Vehicles")) %>%
  mutate(
    hhincome_100 = case_when(
      hhincome_broad == '$100,000-$199,000' ~ '$100,000 or more',
      hhincome_broad == '$200,000 or more' ~ '$100,000 or more',
      TRUE ~ hhincome_broad
    )
  ) %>%
  mutate(
    age_grp = case_when(
      age == '75-84 years' ~ '75 years or older',
      age == '85 or years older' ~ '75 years or older',
      TRUE ~ age
    )
  ) %>% mutate(simple_purpose = ifelse(
    dest_purpose_cat == 'Home',
    origin_purpose_cat,
    dest_purpose_cat
  )) %>%
  mutate(
    simple_purpose = case_when(
      simple_purpose == 'Work' ~ 'Work/School',
      simple_purpose == 'School' ~ 'Work/School',
      simple_purpose == 'Work-related' ~ 'Work/School',
      simple_purpose == 'Shop' ~ 'Shop',
      simple_purpose == 'Escort' ~ 'Errands',
      simple_purpose == 'Errand/Other' ~ 'Errands',
      simple_purpose == 'Change mode' ~ 'Errands',
      simple_purpose == 'Social/Recreation' ~ 'Social/Recreation/Meal',
      simple_purpose == 'Meal' ~ 'Social/Recreation/Meal',
      simple_purpose == 'Meal' ~ 'Social/Recreation/Meal',
      simple_purpose == 'Home' ~ 'Errands',
      is.na(simple_purpose) ~ 'Errands',
      TRUE ~ simple_purpose
    )
  ) 


trip_data_17_19$hhincome_100_f = factor(
  trip_data_17_19$hhincome_100,
  levels = c(
    "Under $25,000",
    "$25,000-$49,999",
    "$50,000-$74,999",
    "$75,000-$99,999",
    "$100,000 or more",
    "Prefer not to answer"
  )
)

trip_data_21$hhincome_100_f = factor(
  trip_data_21$hhincome_100,
  levels = c(
    "Under $25,000",
    "$25,000-$49,999",
    "$50,000-$74,999",
    "$75,000-$99,999",
    "$100,000 or more",
    "Prefer not to answer"
  )
)


trip_data_17_19$NoVehicles = factor(trip_data_17_19$NoVehicles,
                                    levels = c('No Vehicles', 'Has Vehicles'))

trip_data_21$NoVehicles = factor(trip_data_21$NoVehicles,
                                 levels = c('No Vehicles', 'Has Vehicles'))


simple_groupings <-
  c(
    'race_eth_broad' = 'Race/Ethnicity',
    'hhincome_100_f' = 'Household Income',
    'NoVehicles' = 'Household Vehicles',
    'age_grp' = 'Age Group',
    'sample_county' = 'Home County',
    'age_category' = 'Age Category',
    'final_home_is_rgc' = 'Home in Regional Growth Center'
  )
```

Now group the data into the same categories for people.
```{r}
person_data_17_19 <-
  person_data_17_19 %>% mutate(NoVehicles = ifelse(vehicle_count == '0 (no vehicles)', 'No Vehicles', "Has Vehicles")) %>%
  mutate(
    hhincome_100 = case_when(
      hhincome_broad == '$100,000-$199,000' ~ '$100,000 or more',
      hhincome_broad == '$200,000 or more' ~ '$100,000 or more',
      TRUE ~ hhincome_broad
    )
  ) %>%
  mutate(
    age_grp = case_when(
      age == '75-84 years' ~ '75 years or older',
      age == '85 or years older' ~ '75 years or older',
      TRUE ~ age
    )
  ) 

person_data_21 <-
  person_data_21 %>% mutate(NoVehicles = ifelse(vehicle_count == '0 (no vehicles)', 'No Vehicles', "Has Vehicles")) %>%
  mutate(
    hhincome_100 = case_when(
      hhincome_broad == '$100,000-$199,000' ~ '$100,000 or more',
      hhincome_broad == '$200,000 or more' ~ '$100,000 or more',
      TRUE ~ hhincome_broad
    )
  ) %>%
  mutate(
    age_grp = case_when(
      age == '75-84 years' ~ '75 years or older',
      age == '85 or years older' ~ '75 years or older',
      TRUE ~ age
    )
  ) 


person_data_17_19$hhincome_100_f = factor(
  person_data_17_19$hhincome_100,
  levels = c(
    "Under $25,000",
    "$25,000-$49,999",
    "$50,000-$74,999",
    "$75,000-$99,999",
    "$100,000 or more",
    "Prefer not to answer"
  )
)

person_data_21$hhincome_100_f = factor(
  person_data_21$hhincome_100,
  levels = c(
    "Under $25,000",
    "$25,000-$49,999",
    "$50,000-$74,999",
    "$75,000-$99,999",
    "$100,000 or more",
    "Prefer not to answer"
  )
)


person_data_17_19$NoVehicles = factor(person_data_17_19$NoVehicles,
                                      levels = c('No Vehicles', 'Has Vehicles'))

person_data_21$NoVehicles = factor(person_data_21$NoVehicles,
                                   levels = c('No Vehicles', 'Has Vehicles'))



simple_groupings <-
  c(
    'race_eth_broad' = 'Race/Ethnicity',
    'hhincome_100_f' = 'Household Income',
    'NoVehicles' = 'Household Vehicles',
    'age_grp' = 'Age Group',
    'sample_county' = 'Home County',
    'age_category' = 'Age Category',
    'final_home_is_rgc' = 'Home in Regional Growth Center'
  )
```


##How many trips do people make in RGCs and for what purposes?

For this we have to aggregate both the trips and the people, and join them to calculate trips per person.
We may want to functionalize some of this since we are doing things twice, just felt easier not to. (i get indecisive about functionalizing things i do only twice)

```{r}
trip_summs_2017_2019 <- hhts_count(
  trip_data_17_19,
  group_vars = c('final_home_is_rgc', 'simple_purpose'),
  spec_wgt = 'trip_weight_2017_2019_v2021_adult'
) %>% drop_na(c('simple_purpose', 'final_home_is_rgc')) %>% filter(final_home_is_rgc !=
                                                                     'Total')
trip_summs_2021 <- hhts_count(
  trip_data_21,
  group_vars = c('final_home_is_rgc', 'simple_purpose'),
  spec_wgt = 'trip_weight_2021_ABS_Panel_adult'
) %>% drop_na(c('final_home_is_rgc', 'simple_purpose')) %>% filter(final_home_is_rgc !=
                                                                     'Total')


person_summs_2017_2019 <- hhts_count(
  person_data_17_19,
  group_vars = c('final_home_is_rgc'),
  spec_wgt = 'hh_weight_2017_2019_v2021_adult'
) %>%
  filter(final_home_is_rgc != 'Total') %>% drop_na(final_home_is_rgc)

person_summs_2021 <- hhts_count(person_data_21,
                                group_vars = c('final_home_is_rgc'),
                                spec_wgt = 'person_weight_2021_ABS_Panel_adult') %>%
  filter(final_home_is_rgc != 'Total') %>% drop_na(final_home_is_rgc)


# Join persons and trips by group and calculate trips/person.

trips_persons_2017_2019 <-
  merge(
    trip_summs_2017_2019,
    person_summs_2017_2019,
    by = 'final_home_is_rgc',
    suffixes = c('_trips', '_people')
  )
trips_persons_2021 <-
  merge(
    trip_summs_2021,
    person_summs_2021,
    by = 'final_home_is_rgc',
    suffixes = c('_trips', '_people')
  )

trips_persons_2017_2019 <-
  trips_persons_2017_2019 %>% mutate(
    trips_per_person = count_trips / count_people,
    moe_trips_person = moe_ratio(count_trips, count_people, count_moe_trips, count_moe_people)
  )


trips_persons_2021 <-
  trips_persons_2021 %>% mutate(
    trips_per_person = count_trips / count_people,
    moe_trips_person = moe_ratio(count_trips, count_people, count_moe_trips, count_moe_people)
  )

trips_persons_long <-
  rbind(trips_persons_2017_2019, trips_persons_2021) %>%
  mutate(survey_trips = str_replace_all(survey_trips, "_", "/"))

trips_persons_long_all_trips <-
  trips_persons_long %>% filter(simple_purpose == 'Total')

trips_persons_2021_purpose <-
  trips_persons_2021 %>% filter(simple_purpose != 'Total')

trips_persons_long_shop_trips <-
  trips_persons_long %>% filter(simple_purpose == 'Shop')



```
Make charts about trips by purpose by RGC
```{r}
p0 <-
  create_bar_chart(
    t = trips_persons_2021_purpose ,
    x = 'simple_purpose',
    y = 'trips_per_person',
    f = 'final_home_is_rgc',
    moe = 'moe_trips_person',
    est = 'number',
    color = 'psrc_pairs',
    interactive = 'yes',
    dec=1
  )  


p1 <-
  create_facet_bar_chart(
    t = trips_persons_long,
    w.x = 'final_home_is_rgc',
    w.y = 'trips_per_person',
    f = 'survey_trips',
    g = 'simple_purpose',
    w.moe = 'moe_trips_person',
    est.type = 'number',
    w.color = 'psrc_light',
    w.interactive = 'yes'
  )

p2 <-
  create_column_chart(
    t = trips_persons_long_all_trips ,
    x = 'final_home_is_rgc',
    y = 'trips_per_person',
    f = 'survey_trips',
    moe = 'moe_trips_person',
    est = 'number',
    color = 'psrc_pairs',
    interactive = 'yes'
  ) 


p3 <-
  create_column_chart(
    t = trips_persons_long_shop_trips ,
    x = 'final_home_is_rgc',
    y = 'trips_per_person',
    f = 'survey_trips',
    moe = 'moe_trips_person',
    est = 'number',
    color = 'psrc_pairs',
    interactive = 'yes'
  ) 



htmlwidgets::saveWidget(p0, "trips_by_purpose_rgc_2021.html")
htmlwidgets::saveWidget(p1, "trips_by_purpose_rgc_2017_2019_2021.html")
htmlwidgets::saveWidget(p2, "trips_overall_rgc_2017_2019_2021.html")
htmlwidgets::saveWidget(p3, "shop_trips_rgc_2017_2019_2021.html")

```











Trip Mode Shares by Demographic
```{r }
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  # 'final_home_is_rgc' <- eval('final_home_is_rgc')
  'final_home_is_rgc' <- 'final_home_is_rgc'
  print('final_home_is_rgc')
  trip_summs_2017_2019 <- hhts_count(trip_data_17_19,
                                        group_vars=c('final_home_is_rgc','mode_simple'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter('final_home_is_rgc'!='Total') %>%filter(mode_simple!='Total')%>%drop_na('final_home_is_rgc')%>%mutate('final_home_is_rgc'=str_wrap('final_home_is_rgc', width=10))



  trip_summs_2021 <- hhts_count(trip_data_21,
                                   group_vars=c('final_home_is_rgc','mode_simple'),
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter('final_home_is_rgc'!='Total') %>%filter(mode_simple!='Total')%>%drop_na('final_home_is_rgc')%>%mutate('final_home_is_rgc'=str_wrap('final_home_is_rgc', width=10))

  if('final_home_is_rgc'=='race_eth_broad'){

     trip_summs_2017_2019 <- trip_summs_2017_2019 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')

        trip_summs_2021 <- trip_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  #trip_summs_2017_2019$race_eth_broad<-factor(trip_summs_2017_2019$race_eth_broad, levels=
                                                  #c('White', 'Other', 'Hispanic', 'Black', 'Asian'))

   #trip_summs_2021$race_eth_broad<-factor(trip_summs_2021$race_eth_broad, levels=
                                                  #c('White', 'Other', 'Hispanic', 'Black', 'Asian'))



  }

 #  if('final_home_is_rgc'=='hhincome_100_f'){

  #   trip_summs_2017_2019$hhincome_100_f=factor( trip_summs_2017_2019$hhincome_100_f, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))

 #trip_summs_2021$hhincome_100_f=factor(trip_summs_2021$hhincome_100_f, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))


#   }

  trip_summs_long<- rbind(trip_summs_2017_2019, trip_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))


   # factorize variable column. Adjust levels in function definition
  #trip_summs_long <- factor_variable_column('final_home_is_rgc', trip_summs_long)

  #trip_summs_2021_only<-  trip_summs_2021%>%filter(non_motorized_mode=='Walk/Bike')

 # p0<-create_bar_chart(t=trip_summs_2021_only , x='final_home_is_rgc', y='share',  f='final_home_is_rgc',moe='share_moe', est.type='percent', #color='psrc_light')+ t#heme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()


   p0<-create_column_chart(t=trip_summs_2021 , x='mode_simple', y='share', f='final_home_is_rgc', moe='share_moe', est.type='percent', color = 'psrc_light', interactive=TRUE)  +theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  
  psrc_pres <- add_bullet_plus_chart_slide (p=psrc_pres, 
                                         p.title="Walk, Bike, and trip Shares", 
                                         p.caption=paste0("2017/2019 vs 2021"), 

                                         p.chart=p0)

  p1<-create_facet_bar_chart(t=trip_summs_long , f='mode_simple',  x='final_home_is_rgc', y='share', g='survey', moe='share_moe', est.type='percent', color = 'psrc_light', interactive=TRUE) +theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)
psrc_pres <- add_bullet_plus_chart_slide (p=psrc_pres, 
                                         p.title="Walk, Bike, and trip Shares", 
                                         p.caption=paste0("2017/2019 vs 2021"), 

                                         p.chart=p1)
}
 # p<-create_facet_bar_chart(t=trip_summs_long , x='final_home_is_rgc', y='share', f='non_motorized_mode', g='survey', moe='share_moe', est.type='percent', color = 'psrc_light')+ theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()



```

# Walk Trip Mode Shares by Demographic
```{r }
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  # 'final_home_is_rgc' <- eval('final_home_is_rgc')
  'final_home_is_rgc' <- 'final_home_is_rgc'
  print('final_home_is_rgc')
  trip_summs_2017_2019 <- hhts_count(trip_data_17_19,
                                        group_vars=c('final_home_is_rgc','mode_simple'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter('final_home_is_rgc'!='Total') %>%
    filter(mode_simple=='Walk')%>%drop_na('final_home_is_rgc')%>%mutate('final_home_is_rgc'=str_wrap('final_home_is_rgc', width=10))



  trip_summs_2021 <- hhts_count(trip_data_21,
                                   group_vars=c('final_home_is_rgc','mode_simple'),
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter('final_home_is_rgc'!='Total') %>%
    filter(mode_simple=='Walk')%>%drop_na('final_home_is_rgc')%>%mutate('final_home_is_rgc'=str_wrap('final_home_is_rgc', width=10))

  if('final_home_is_rgc'=='race_eth_broad'){

     trip_summs_2017_2019 <- trip_summs_2017_2019 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')

        trip_summs_2021 <- trip_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  #trip_summs_2017_2019$race_eth_broad<-factor(trip_summs_2017_2019$race_eth_broad, levels=
                                                  #c('White', 'Other', 'Hispanic', 'Black', 'Asian'))

   #trip_summs_2021$race_eth_broad<-factor(trip_summs_2021$race_eth_broad, levels=
                                                  #c('White', 'Other', 'Hispanic', 'Black', 'Asian'))



  }

 #  if('final_home_is_rgc'=='hhincome_100_f'){

  #   trip_summs_2017_2019$hhincome_100_f=factor( trip_summs_2017_2019$hhincome_100_f, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))

 #trip_summs_2021$hhincome_100_f=factor(trip_summs_2021$hhincome_100_f, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))


#   }

  trip_summs_long<- rbind(trip_summs_2017_2019, trip_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))


   # factorize variable column. Adjust levels in function definition
  #trip_summs_long <- factor_variable_column('final_home_is_rgc', trip_summs_long)

  #trip_summs_2021_only<-  trip_summs_2021%>%filter(non_motorized_mode=='Walk/Bike')

 # p0<-create_bar_chart(t=trip_summs_2021_only , x='final_home_is_rgc', y='share',  f='final_home_is_rgc',moe='share_moe', est.type='percent', #color='psrc_light')+ t#heme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()


   p0<-create_column_chart(t=trip_summs_2021 , x='final_home_is_rgc', y='share', f='mode_simple', moe='share_moe', est.type='percent', color = 'psrc_light', interactive=TRUE) + 
  xlab(as.character(g['final_home_is_rgc'])) + ylab('Walk mode share')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  
  psrc_pres <- add_bullet_plus_chart_slide (p=psrc_pres, 
                                         p.title="Walk, Bike, and trip Shares", 
                                         p.caption=paste0("2017/2019 vs 2021"), 

                                         p.chart=p0)

  p1<-create_column_chart(t=trip_summs_long , x='final_home_is_rgc', y='share', f='survey', moe='share_moe', est.type='percent', color = 'psrc_light', interactive=TRUE)+ 
  xlab(as.character(g['final_home_is_rgc'])) + ylab("Walk mode share")+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)
psrc_pres <- add_bullet_plus_chart_slide (p=psrc_pres, 
                                         p.title="Walk, Bike, and trip Shares", 
                                         p.caption=paste0("2017/2019 vs 2021"), 

                                         p.chart=p1)
}
 # p<-create_facet_bar_chart(t=trip_summs_long , x='final_home_is_rgc', y='share', f='non_motorized_mode', g='survey', moe='share_moe', est.type='percent', color = 'psrc_light')+ theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()



#print(psrc_pres, target = "C:/Users/schildress/OneDrive - Puget Sound Regional Council/Documents/hhts_sntc_sept22_new2.pptx")
```
```{r }
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  # 'final_home_is_rgc' <- eval('final_home_is_rgc')
  'final_home_is_rgc' <- 'final_home_is_rgc'
  print('final_home_is_rgc')
  trip_summs_2017_2019 <- hhts_count(trip_data_17_19,
                                        group_vars=c('final_home_is_rgc','mode_simple'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter('final_home_is_rgc'!='Total') %>%
    filter(mode_simple=='trip')%>%drop_na('final_home_is_rgc')

  trip_summs_2021 <- hhts_count(trip_data_21,
                                   group_vars=c('final_home_is_rgc','mode_simple'),
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter('final_home_is_rgc'!='Total') %>%
    filter(mode_simple=='trip')%>%drop_na('final_home_is_rgc')
  if('final_home_is_rgc'=='race_eth_broad'){

     trip_summs_2017_2019 <- trip_summs_2017_2019 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')

        trip_summs_2021 <- trip_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  trip_summs_2017_2019$race_eth_broad<-factor(trip_summs_2017_2019$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))

   trip_summs_2021$race_eth_broad<-factor(trip_summs_2021$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))



  }



  trip_summs_long<- rbind(trip_summs_2017_2019, trip_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))


   # factorize variable column. Adjust levels in function definition
  #trip_summs_long <- factor_variable_column('final_home_is_rgc', trip_summs_long)

  #trip_summs_2021_only<-  trip_summs_2021%>%filter(non_motorized_mode=='Walk/Bike')

 # p0<-create_bar_chart(t=trip_summs_2021_only , x='final_home_is_rgc', y='share',  f='final_home_is_rgc',moe='share_moe', est.type='percent', #color='psrc_light')+ t#heme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()


   p0<-create_column_chart(t=trip_summs_2021 , x='final_home_is_rgc', y='share', f='mode_simple', moe='share_moe', est.type='percent', color = 'psrc_light', interactive=TRUE) + 
  xlab(as.character(g['final_home_is_rgc'])) + ylab('trip mode share')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  
  psrc_pres <- add_bullet_plus_chart_slide (p=psrc_pres, 
                                         p.title="Walk, Bike, and trip Shares", 
                                         p.caption=paste0("2017/2019 vs 2021"), 

                                         p.chart=p0)

  p1<-create_column_chart(t=trip_summs_long , x='final_home_is_rgc', y='share', f='survey', moe='share_moe', est.type='percent', color = 'psrc_light', interactive=TRUE)+ 
  xlab(as.character(g['final_home_is_rgc'])) + ylab("trip mode share")+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)
psrc_pres <- add_bullet_plus_chart_slide (p=psrc_pres, 
                                         p.title="Walk, Bike, and trip Shares", 
                                         p.caption=paste0("2017/2019 vs 2021"), 

                                         p.chart=p1)
}
```



```{r}
  all_trips_summs_2017_2019 <- hhts_count(trip_data_17_19,
                                        group_vars=c('simple_purpose'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter(simple_purpose!='Total')




  all_trips_summs_2021 <- hhts_count(trip_data_21,
                                   group_vars=c('simple_purpose'),
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter(simple_purpose!='Total')



  all_trips_summs_long<- rbind(all_trips_summs_2017_2019, all_trips_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))



p<-create_column_chart(t=all_trips_summs_long , x='simple_purpose', y='share', f='survey', moe='share_moe', est.type='percent', color='psrc_dark')+
  ylab('Share of All Trips') + xlab('Purpose')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+ theme(axis.title.x = element_text(size=50,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p)
  
  
psrc_pres <- add_bullet_plus_chart_slide (p=psrc_pres, 
                                         p.title="Access Mode to trip", 
                                         p.caption=paste0("2017/2019 vs 2021"), 

                                         p.chart=p)


  


```

```{r }
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  # 'final_home_is_rgc' <- eval('final_home_is_rgc')
  'final_home_is_rgc' <- 'final_home_is_rgc'
  print('final_home_is_rgc')
  trip_summs_2017_2019 <- hhts_count(trip_data_17_19,
                                        group_vars=c('final_home_is_rgc','simple_purpose'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter('final_home_is_rgc'!='Total') %>%drop_na(c('final_home_is_rgc', 'simple_purpose'))%>%filter(simple_purpose!='Total')

  trip_summs_2021 <- hhts_count(trip_data_21,
                                   group_vars=c('final_home_is_rgc','simple_purpose'),
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter('final_home_is_rgc'!='Total') %>%drop_na('final_home_is_rgc') %>%drop_na(c('final_home_is_rgc', 'simple_purpose'))%>%filter(simple_purpose!='Total')
  
  if('final_home_is_rgc'=='race_eth_broad'){

     trip_summs_2017_2019 <- trip_summs_2017_2019 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')

        trip_summs_2021 <- trip_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  trip_summs_2017_2019$race_eth_broad<-factor(trip_summs_2017_2019$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))

   trip_summs_2021$race_eth_broad<-factor(trip_summs_2021$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))



  }



  trip_summs_long<- rbind(trip_summs_2017_2019, trip_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))


   # factorize variable column. Adjust levels in function definition
  #trip_summs_long <- factor_variable_column('final_home_is_rgc', trip_summs_long)

  #trip_summs_2021_only<-  trip_summs_2021%>%filter(non_motorized_mode=='Walk/Bike')

 # p0<-create_bar_chart(t=trip_summs_2021_only , x='final_home_is_rgc', y='share',  f='final_home_is_rgc',moe='share_moe', est.type='percent', #color='psrc_dark')+ t#heme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()


   p0<-create_column_chart(t=trip_summs_2021 , x='simple_purpose', y='share', f='final_home_is_rgc', moe='share_moe', est.type='percent', color = 'psrc_dark', interactive=TRUE) + 
  xlab(as.character(g['final_home_is_rgc'])) + ylab('Trip Purpose')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  
  psrc_pres <- add_bullet_plus_chart_slide (p=psrc_pres, 
                                         p.title="Walk, Bike, and trip Shares", 
                                         p.caption=paste0("2017/2019 vs 2021"), 

                                         p.chart=p0)

  p1<-create_facet_bar_chart(t=trip_summs_long , x='final_home_is_rgc', y='share', f='survey',g='simple_purpose', moe='share_moe', est.type='percent', color = 'psrc_dark', interactive=TRUE)+ 
  xlab(as.character(g['final_home_is_rgc'])) + ylab("trip mode share")+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)
psrc_pres <- add_bullet_plus_chart_slide (p=psrc_pres, 
                                         p.title="Walk, Bike, and trip Shares", 
                                         p.caption=paste0("2017/2019 vs 2021"), 

                                         p.chart=p1)
}
```


```{r }
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  # 'final_home_is_rgc' <- eval('final_home_is_rgc')
  'final_home_is_rgc' <- 'final_home_is_rgc'
  print('final_home_is_rgc')
  trip_summs_2017_2019 <- hhts_median(trip_data_17_19,
                                         stat_var = 'trip_path_distance',
                                        group_vars=c('final_home_is_rgc'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
filter('final_home_is_rgc'!='Other')%>%drop_na('final_home_is_rgc')

  

  trip_summs_2021<- hhts_median(trip_data_21,
                                         stat_var = 'trip_path_distance',
                                        group_vars=c('final_home_is_rgc'),
                                        spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
filter('final_home_is_rgc'!='Other')%>%drop_na('final_home_is_rgc')
  
  if('final_home_is_rgc'=='race_eth_broad'){

     trip_summs_2017_2019 <- trip_summs_2017_2019 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')

        trip_summs_2021 <- trip_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  trip_summs_2017_2019$race_eth_broad<-factor(trip_summs_2017_2019$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))

   trip_summs_2021$race_eth_broad<-factor(trip_summs_2021$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))



  }



  trip_summs_long<- rbind(trip_summs_2017_2019, trip_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))


   # factorize variable column. Adjust levels in function definition
  #trip_summs_long <- factor_variable_column('final_home_is_rgc', trip_summs_long)

  #trip_summs_2021_only<-  trip_summs_2021%>%filter(non_motorized_mode=='Walk/Bike')

 # p0<-create_bar_chart(t=trip_summs_2021_only , x='final_home_is_rgc', y='share',  f='final_home_is_rgc',moe='share_moe', est.type='percent', #color='psrc_dark')+ t#heme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()


   p0<-create_column_chart(t=trip_summs_2021 , x='final_home_is_rgc', y='trip_path_distance_median', f='survey', moe='trip_path_distance_median_moe', est.type='number', color = 'psrc_pairs', interactive=TRUE) + 
  xlab(as.character(g['final_home_is_rgc'])) + ylab('Median Distance')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  
  psrc_pres <- add_bullet_plus_chart_slide (p=psrc_pres, 
                                         p.title="Walk, Bike, and trip Shares", 
                                         p.caption=paste0("2017/2019 vs 2021"), 

                                         p.chart=p0)

  p1<-create_column_chart(t=trip_summs_long , x='final_home_is_rgc', y='trip_path_distance_median', f='survey', moe='trip_path_distance_median_moe', est.type='number', color = 'psrc_pairs', interactive=TRUE)+ 
  xlab(as.character(g['final_home_is_rgc'])) + ylab("Median Distance")+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)
psrc_pres <- add_bullet_plus_chart_slide (p=psrc_pres, 
                                         p.title="Walk, Bike, and trip Shares", 
                                         p.caption=paste0("2017/2019 vs 2021"), 

                                         p.chart=p1)
}
```

```{r }
print(psrc_pres, target = "C:/Users/schildress/OneDrive - Puget Sound Regional Council/Documents/hhts_sntc_sept22_new6.pptx")
```




Trips per Person
```{r }


```{r}
trip_summs_long<- rbind(trip_summs_2017_2019, trip_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))


   # factorize variable column. Adjust levels in function definition
  #trip_summs_long <- factor_variable_column('final_home_is_rgc', trip_summs_long)

  #trip_summs_2021_only<-  trip_summs_2021%>%filter(non_motorized_mode=='Walk/Bike')

 # p0<-create_bar_chart(t=trip_summs_2021_only , x='final_home_is_rgc', y='share',  f='final_home_is_rgc',moe='share_moe', est.type='percent', #color='psrc_dark')+ t#heme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()


   p0<-create_column_chart(t=trip_summs_2021 , x='simple_purpose', y='share', f='final_home_is_rgc', moe='share_moe', est.type='percent', color = 'psrc_dark', interactive=TRUE) + 
  xlab(as.character(g['final_home_is_rgc'])) + ylab('Trip Purpose')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  
  psrc_pres <- add_bullet_plus_chart_slide (p=psrc_pres, 
                                         p.title="Walk, Bike, and trip Shares", 
                                         p.caption=paste0("2017/2019 vs 2021"), 

                                         p.chart=p0)

  p1<-create_facet_bar_chart(t=trip_summs_long , x='final_home_is_rgc', y='share', f='survey',g='simple_purpose', moe='share_moe', est.type='percent', color = 'psrc_dark', interactive=TRUE)+ 
  xlab(as.character(g['final_home_is_rgc'])) + ylab("trip mode share")+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)
psrc_pres <- add_bullet_plus_chart_slide (p=psrc_pres, 
                                         p.title="Walk, Bike, and trip Shares", 
                                         p.caption=paste0("2017/2019 vs 2021"), 

                                         p.chart=p1)
```

```{r}
library(psrcplot)
library(dplyr)
# This installs PSRC fonts for use in R. It has to be run at the start of each R sessions
install_psrc_fonts()
# Column Charts -----------------------------------------------------
mode_shares <- read.csv(system.file('extdata', 'example_data.csv', package='psrcplot')) %>% 
  filter(Category=="Mode to Work by Race") %>%
  filter(Geography=="Region" & Race=="Total") %>%
  mutate(Year = as.character(Year)) %>%
  mutate(axis_labels = stringr::str_wrap(Mode, width = 10))
# Static Chart with Error Bars 
static_ms_chart <- create_column_chart(t=mode_shares, x="Mode",y="share", f="Year",
                                       title="Mode Share to Work",
                                       source="Source: ACS 5yr Data Table B3002",
                                       moe="share_moe", color="psrc_light")
interactive_ms_chart <- plotly::ggplotly(static_ms_chart)

```
