---
title: "college_survey summary"
author: "weiran"
date: "June 20, 2016"
output: html_document
---

```{r,message=FALSE,echo=FALSE,warning=FALSE}

library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(lubridate)
library(weights)
library(reshape2)


setwd("C:\\Users\\weiran\\Documents\\HHSurvey\\college")
# setwd("~/Google Drive/Wei-PSRC-Intern")

```

##Data Preparing and Processing
this section read data and change the type of data

```{r,warning=FALSE}

# read data of college survey
df.person_c<- read.csv("1_PSRC_College_PersonDataset_2015-1-23_V1-3.csv")
df.trip_c<- read.csv("2_PSRC_College_TripDataset_2015-1-23_v2-1_LINKED.csv")
# read data of regional survey
df.person<- read.csv("2014-pr3-M-hhsurvey-persons.csv")
df.trip <- read.csv("2014-pr3-M-hhsurvey-trips-LINKED.csv")
df.person15 <- read.csv("~/HHSurvey/surveydat/2015-pr1-M-hhsurvey-person.csv")
#------------data processing------------

df_col <- df.trip_c %>%
  mutate(traveldate = dmy(as.character(traveldate))) %>%
  mutate(trip_commute = ifelse((o_purpose==1 & d_purpose==2)|(o_purpose==2 & d_purpose ==1)
                               ,1,0)) %>% #create a variable indicating whether it is a commute trip
  
  mutate_each(funs(as.factor),mode,college,trip_commute,o_purpose,d_purpose)


#trip dataset join with person dataset's variables
df_col.live <- df.person_c %>%
  select(personID=personid,live,h_district_name,h_districtnum,home_cnty,
  expwt,campus)

df_col<- df_col%>%
  left_join(df_col.live,by='personID') %>%    #join person dataset variable
  mutate_each(funs(as.factor),home_cnty)%>%
  filter(!is.na(expwt))

#sign names to factors
levels(df_col$home_cnty) <- c('King','Kitsap','Pierce','Snohomish','Other WA cnty','Outside WA')
df.person_c$college <- as.factor(df.person_c$college)
levels(df_col$college) <- c("Bellevue.C","Everett.C", "Green River.C", "Seattle.C", "U.W")
levels(df.person_c$college) <- c("Bellevue.C","Everett.C", "Green River.C", "Seattle.C", "U.W")
levels(df_col$h_district_name)[levels(df_col$h_district_name)==""] <- NA
```

## Sample size overview by college
This describe the sample size of each college and campus.

```{r,warning=FALSE}
#sample size by college and campus
#how many people by college in the survey
table(df.person_c$college)
# how many people by campus in the survey
# campus code:
# 1	Seattle campus
# 2	Bothell campus
# 3	Tacoma campus
# 4	North Seattle College
# 5	Seattle Central College
# 6	South Seattle College (Main Campus)
# 7	Main campus
# 8	Enumclaw campus
# 9	Kent campus
# 10	Downtown Auburn Center
# 11	Main campus
# 12	North campus
# 13	Online only - I never go to campus
# 14	South Seattle College (Georgetown Campus)
# 15	Seattle Vocational Institute

table(df.person_c$campus,df.person_c$college)
#how many trips by college in the survey
table(df_col$college)


```

## 1. Mode choice Share 

*mode* has been grouped into *newmode* with less catagories 

```{r, warning=FALSE}
#mode share of all trips and commute trips (unit:percentage)
# Here is the original mode code:
# 1	Drove alone
# 3	Drove/rode with people not in household
# 4	Motorcycle/moped/scooter
# 5	Vanpool
# 6	Bicycle
# 7	Walk, jog, or wheelchair
# 8	Bus (public transit)
# 9	Train (rail and monorail)
# 10	Ferry or water taxi
# 11	Streetcar
# 13	Taxi or other hired car service (e.g. Lyft, Uber)
# 14	Paratransit
# 16	Airplane or helicopter
# 15	Private bus or shuttle
# 17	Other (e.g. skateboard, kayak, motor home, etc.)

round(wpct(df_col$mode,df_col$expwt),digits = 3)
 

#weighted result
ggplot(df_col, aes(x=factor(mode)))+ 
  geom_bar(aes(y=..prop..,group = college,weight= expwt))+
  facet_grid(college~.)+
  ylab("Percentage")+
  scale_y_continuous(breaks=seq(0,0.6,0.1),labels=percent)

#grouped into *newmode* with less catagories 
library(plyr)
df_col<-df_col%>%
  mutate(newmode=ifelse(mode%in% c(1,2),'SOV',
                        ifelse(mode%in% c(3,5),'HOV',
                               ifelse(mode%in% c(8,9,10,11),'Transit',
                                      ifelse(mode==6,'Bike',
                                             ifelse(mode==7,'Walk','Other'))))))%>%
  mutate(d_purpose= factor(d_purpose))%>%
  mutate(d_purpose= revalue(d_purpose,
  c("1"="home","2"= "school","3" ="work", "4"= "work2", "5" ="grocery","6"= "shop","7"="medical","8"="personal","9"="drop","10"="exercise","11"="eat","12"="social","13"="recreation","14"="activity", "15"="transfer","16"="other","17"="school2")))

#compare with regional mode share
#trips by purpose

# purpose code
# 1	Go home/to dorm
# 2	Go to school (e.g. go to class, go to tutoring)
# 3	Go to work/job
# 4	Go to other work-related place (e.g. meeting, second job, delivery)
# 5	Go grocery shopping
# 6	Go to other shopping (e.g. mall, pet store)
# 7	Go to medical appointment (e.g. doctor, dentist)
# 8	Conduct personal business (e.g. bank, post office)
# 9	Drop off/pick up someone (e.g. drop off friend at bus stop)
# 10	Go exercise (e.g. gym, walk, jog, bike ride)
# 11	Go to restaurant to eat/get take-out
# 12	Attend social event (e.g. visit with friends, family, co-workers)
# 13	Attend recreational event (e.g. movies, sporting event)
# 14	Go to religious/community/volunteer activity
# 15	Transfer to another mode of transportation (e.g. change from ferry to bus)
# 16	Other
# 17	Go to school-related activity

df.trip<-df.trip%>%
  filter(mode !=-99)%>%
  mutate(newmode=ifelse(mode%in% c(1,2),'SOV',
                        ifelse(mode%in% c(3,5),'HOV',
                               ifelse(mode%in% c(8,9,10,11),'Transit',
                                      ifelse(mode==6,'Bike',
                                             ifelse(mode==7,'Walk','Other'))))))%>%
  mutate(d_purpose= factor(d_purpose))%>%
  mutate(d_purpose= revalue(d_purpose,c("1"="home","2"= "work","3" ="work2","4" ="grocery","5"= "shop","6"="school","7"="medical","8"="personal","9"="drop","10"="exercise","11"="eat","12"="social","13"="recreation","14"="activity","15"="transfer","16"="other")))
detach(package:plyr)

mode_c<-cbind(wpct(df_col$newmode,df_col$expwt),
                    wpct(df.trip$newmode,df.trip$expwt))
colnames(mode_c)<-c("college survey","regional survey")
round(mode_c,digits = 3)

ggplot(melt(mode_c),aes(Var1,value,fill= Var2))+geom_bar(stat="identity",position="dodge")+
    ylab("Percentage")+xlab("Mode")



ggplot(df_col,aes(x=college,weight= expwt))+
  geom_bar(aes(fill=newmode),position = "fill")+
  # geom_text(position="stack",aes(ymax=1),vjust=5)+
  ylab("Percentage")+
  ggtitle("Mode Choice Share by College")




#commute trip modeshare
df.person_c<-df.person_c%>%
  mutate(newmode=ifelse(mode_campus%in% c(1,3),'SOV',
                        ifelse(mode_campus%in% c(2,4,11),'HOV',
                               ifelse(mode_campus%in%
                                        c(8,9,10,7),'Transit',
                                      ifelse(mode_campus==5,'Bike',
                                    ifelse(mode_campus==6,'Walk','Other'))))))

round(wpct(df.person_c$newmode,df_col$expwt),digits = 3)

ggplot(df.person_c,aes(x=college,weight= expwt))+
  geom_bar(aes(fill=newmode),position = "fill")+
  ylab("Percentage")+
  ggtitle("Typical mode to/from campus by College(commuting)")



```

## 2.Trips per day, total  and by purpose


```{r,warning=FALSE}
# total and average daily trips
av_trip_c<-weighted.mean(df.person_c$numtrips,df.person_c$expwt)
av_trip<-weighted.mean(df.person$numtrips,df.person$expwt_final)

ggplot(df.person_c,aes(college,numtrips)) + geom_boxplot(aes(weight= expwt)) + 
  geom_hline(yintercept=c(av_trip_c,av_trip),  
             color=c("red","blue"), linetype="dashed", size=1)+
  ggtitle('average daily trips made by each student') 

av_trip_c 
av_trip
  


ggplot(df_col, aes(x=d_purpose))+ 
  geom_bar(aes(y=..prop..,group = college,weight= expwt))+
  facet_grid(college~.)+
  ylab('percentage') +
  ggtitle("trips by destination purpose")


t_c<-data.frame(college =wpct(df_col$d_purpose,df_col$expwt),d_purpose= levels(df_col$d_purpose))
t_r<-data.frame(regional=wpct(df.trip$d_purpose,df.trip$expwt_final),d_purpose= levels(df.trip$d_purpose))

ggplot(t_c)+ geom_bar(aes(d_purpose,college),stat="identity")+
  ylab('percentage') +
  ggtitle("trips by destination purpose(college)")

ggplot(t_r)+ geom_bar(aes(d_purpose,regional),stat="identity")+
  ylab('percentage') +
  ggtitle("trips by destination purpose(regional)")

```

##3. trip length/distance summary


```{r,warning=FALSE}
df.trip<- df.trip%>%
  filter(trip_dur_reported!=-99,gdist!=-99)

# Trip length/distance frequency and overall
trip_len_region<-weighted.mean(df.trip$trip_dur_reported, df.trip$expwt_final)
trip_len_college<-weighted.mean(df_col$trip_dur_reported, df_col$expwt)
round(data.frame(trip_len_region,trip_len_college),digits = 3)


ggplot(df_col,aes(trip_dur_reported)) +geom_histogram(colour="black", fill="white") +
  xlim(0,150) + 
  xlab('trip length distribution(showing up to 150 minutes)') + ylab('frequency')

# trip length by purpose

ggplot(df_col,aes(as.factor(d_purpose),trip_dur_reported)) + 
  geom_boxplot(aes(weight= expwt),fill= "red",alpha=0.2) +
  ylim(0,70) + ylab('trip length (showing up to 70 minutes)') + xlab('trip purpose')


# commute trip length by college
ggplot(df_col[df_col$trip_commute==1,],aes(college,trip_dur_reported)) + geom_boxplot(aes(weight= expwt)) +
  geom_hline(aes(yintercept=weighted.mean(trip_dur_reported, expwt),weight= expwt),  
             color="red", linetype="dashed", size=1)+
  ylim(0,75) + ggtitle("Length of Commuting Trips by College")+
  ylab('trip length (showing up to 70 minutes)') + xlab('college')


ggplot(df_col[df_col$trip_commute==1,],aes(college,gdist)) + geom_boxplot(aes(weight= expwt)) +
  geom_hline(aes(yintercept=weighted.mean(gdist, expwt)),  
             color="red", linetype="dashed", size=1)+
  ylim(0,30) + ggtitle("Distance of Commuting Trips by College")+
  ylab('trip distance (showing up to 30 miles)') + xlab('college')

#average trip distance vs regional
trip_dis<-weighted.mean(df.trip$gdist, df.trip$expwt_final,na.rm =T)
trip_dis_college<- weighted.mean(df_col$gdist, df_col$expwt,na.rm =T)
round(data.frame(trip_dis,trip_dis_college),digits = 1)



ggplot(df_col,aes(as.factor(d_purpose),gdist)) + 
  geom_boxplot(aes(weight= expwt),fill= "red",alpha=0.2) +
  geom_hline(yintercept=c(trip_dis_college,trip_dis),  
             color=c("red","blue"), linetype="dashed", size=0.8)+
  ylim(0,30) + ylab('trip dist(showing up to 30miles)') + xlab('trip purpose')



```


## 4. VMT per day per person
```{r,warning=FALSE}
# VMT per day per person by college
v_trip<- df.trip%>%
  filter(driver==1,mode %in% c(1,2,3,4,5))

v_trip_c<- df_col%>%
  filter(driver==1,mode %in% c(1,2,3,4,5))

vmt<-sum(v_trip$gdist*v_trip$expwt_final)/sum(df.person$expwt_final)
vmt_college<-sum(v_trip_c$gdist*v_trip_c$expwt)/sum(df.person_c$expwt)
round(data.frame(vmt,vmt_college),digits = 3)


df_col4.1 <- v_trip_c %>%
  group_by(college) %>%
  summarise(vmt = sum(gdist*expwt))
df_col4.1$vmt<-df_col4.1$vmt/summarise(group_by(df.person_c,college),n= sum(expwt))$n


```

##5.walk/bike/transit frequency


```{r,warning=FALSE}
df.person<- df.person%>%
  filter(walk_freq!=-99,bike_freq!=-99,transit_freq!=-99)

walk_freq<-wpct(df.person_c$walk_freq,df.person_c$expwt)
transit_freq<-wpct(df.person_c$transit_freq,df.person_c$expwt)
bike_freq<-wpct(df.person_c$bike_freq,df.person_c$expwt)
freq_c<-cbind(walk_freq,transit_freq,bike_freq)

walk_freq<-wpct(df.person$walk_freq,df.person_c$expwt_final)
transit_freq<-wpct(df.person$transit_freq,df.person_c$expwt_final)
bike_freq<-wpct(df.person$bike_freq,df.person_c$expwt_final)
freq<-cbind(walk_freq,transit_freq,bike_freq)
round(freq,digits = 3)#regional survey
round(freq_c,digit=3) #college survey

# round(cbind(walk_freq_c,walk_freq,transit_freq_c,transit_freq,bike_freq_c,bike_freq),digits = 3)
survey<-c(rep("college",21),rep("regional",21))
df.freq <-cbind(rbind(melt(freq_c), melt(freq)),survey)

ggplot(df.freq,aes(x= factor(Var1),y= value,fill=survey))+
  geom_bar(stat = "identity",position = "dodge")+facet_grid(Var2~.)+
  xlab("Freqency of usage")+ylab("percentage")

# walk freq by college.
df_col10 <- df.person_c %>%
  mutate(transit_freq=as.factor(transit_freq),
         bike_freq=as.factor(bike_freq),walk_freq=as.factor(walk_freq)) 

ggplot(df_col10,aes(x=college,weight= expwt)) + 
  geom_bar(aes(fill = walk_freq),position = "fill") + 
  scale_fill_brewer(palette = "Greens",direction = -1)+  theme_bw()+
  xlab('Walk frequency (1-7:from high to low)') + ylab('counts') + 
  ggtitle('Walk freq. by college')

# transit freq  by college.
ggplot(df_col10,aes(x=college,weight= expwt)) + 
  geom_bar(aes(fill = transit_freq),position = "fill") + 
  scale_fill_brewer(palette = "Greens",direction = -1)+  theme_bw()+
  xlab('transit frequency (1-7:from high to low)') + ylab('counts') + 
  ggtitle('transit freq. by college')

# bike freq  by college.
ggplot(df_col10,aes(x=college,weight= expwt)) + 
  geom_bar(aes(fill = bike_freq),position = "fill") + 
  scale_fill_brewer(palette = "Greens",direction = -1)+  theme_bw()+
  xlab('bike frequency (1-7:from high to low)') + ylab('counts') + 
  ggtitle('bike freq. by college')

```

## 6. Commute times by home location 


```{r, fig.height=7, fig.width=8, warning=FALSE}
# (h_district_name)
df_col6 <- df_col %>%
  filter(trip_commute==1) %>%
  select(trip_dur_reported,h_district_name,college)

ggplot(df_col6,aes(h_district_name,
               trip_dur_reported)) +
  # geom_point()+
  # stat_summary(fun.data = "mean_cl_boot", colour = "red", size = .8)+
  geom_dotplot(binaxis="y",binwidth = .5,stackdir = "center") + 
  stat_summary(fun.data = "mean_cl_boot", colour = "red", size = .6)+
  xlab('home district') +ylim(0,150) +
  ylab('trip time (in min)')+ 
  facet_grid(college~.,scales='free_y')+
  theme(axis.text.x = element_text(angle=45, hjust=1)) 

```

## 7. carshare data vs. 2014 regional results

```{r,warning=FALSE}
df.person_c<- df.person_c%>%
  filter(!is.na(carshare_car2go))

car2go<-wpct(df.person_c$carshare_car2go,df.person_c$expwt)
zipcar<-wpct(df.person_c$carshare_zipcar,df.person_c$expwt)
relayrides<-wpct(df.person_c$carshare_relayrides,df.person_c$expwt)
other<-wpct(df.person_c$carshare_other,df.person_c$expwt)
notmember<-wpct(df.person_c$carshare_none,df.person_c$expwt)

carshare_c<-round(cbind(car2go,zipcar,relayrides,other,notmember),digits = 3)
carshare_c#college survey

car2go<-wpct(df.person15$carshare_car2go,df.person15$expwt)
zipcar<-wpct(df.person15$carshare_zipcar,df.person15$expwt)
relayrides<-wpct(df.person15$carshare_relayrides,df.person15$expwt)
other<-wpct(df.person15$carshare_other,df.person15$expwt)
notmember<-wpct(df.person15$carshare_none,df.person15$expwt)

carshare15<-round(cbind(car2go,zipcar,relayrides,other,notmember),digits = 3)
carshare15#regional survey in 2015

# carshare membership by college
ggplot(df.person_c[!is.na(df.person_c$carshare_none),],aes(college,fill=factor(carshare_none))) +  
  geom_bar(position='fill',aes(weight = expwt)) +  xlab('home district') + ylab('carshare usage') +
  scale_fill_discrete(name='carshare membership',labels=c('Yes','No'))

```

## 8. Transit pass ownership and usage


```{r,warning=FALSE}
# transit percentage

round(wpct(df.person_c$transitpay_orca,df.person_c$expwt),digits = 3)
round(wpct(df.person_c$transitpay_cash,df.person_c$expwt),digits = 3)
round(wpct(df.person_c$transitpay_upass,df.person_c$expwt),digits = 3)

```

## 9. Parking pass ownership and lot usage by district or school


```{r,warning=FALSE}
df_col9 <- df.person_c %>%
  filter(!is.na(parking_pass)) %>%
  mutate(parking_pass=as.factor(ifelse(parking_pass==1,'Yes','No'))) %>%
  mutate(live=as.factor(ifelse(live==1,'on campus','off campus')))

ggplot(df_col9,aes(college,fill=parking_pass)) + 
geom_bar(position='fill',aes(weight = expwt)) 

```

##10  Time of day travel distribution for different schools vs 

```{r,warning=FALSE}
# regiondal distributions (hist of departure time by time of day)
df_col5 <- df_col %>%
  mutate(time_start = hm(as.character(time_start_hhmm))) %>%
  mutate(hour=hour(time_start),minute=minute(time_start)) %>%
  mutate(time=hour+minute/60) %>%
  mutate(cnty_start=as.factor(cnty_start),
         cnty_end=as.factor(cnty_end))
class(df_col5$time) <- 'POSIXct'

df.trip5 <- df.trip %>%
  mutate(time_start = hm(as.character(time_start_hhmm))) %>%
  mutate(hour=hour(time_start),minute=minute(time_start)) %>%
  mutate(time=hour+minute/60) 
class(df.trip5$time) <- 'POSIXct'

#regional vs college departure time
ggplot() + 
  # geom_histogram(data=df.trip5,aes(y=..density..,x=time),fill= "blue",alpha=0.3) +
  geom_density(data=df_col5,aes(time,y=..scaled..,weight= expwt),fill= "red",alpha=0.3) +
  geom_density(data=df.trip5,aes(time,y=..scaled..,weight= expwt_final),fill= "blue",alpha=0.3)+
  scale_x_datetime(labels=date_format("%S:00")) +
  xlab('departure time') + ylab('percentage') 

# by school
ggplot(df_col5) + geom_histogram(aes(time,y=..density..,weight= expwt,fill=college)) +
  scale_x_datetime(labels=date_format("%S:00")) +
  xlab('departure time') + ylab('Percentage') + 
  facet_grid(college~.)

# by county start?
levels(df_col5$cnty_start) <- c('King','Kitsap',
                            'Pierce','Snohomish',
                            'Other WA cnty',
                            'Outside WA')
ggplot(df_col5,aes(time)) + geom_histogram(aes(fill=cnty_start,weight= expwt)) +
  scale_x_datetime(labels=date_format("%S:00")) +
  xlab('departure time') + ylab('counts') + 
  facet_grid(cnty_start~.,scales='free_y')

```


## 11. driver's license vs region over all and by age groups (what is driver's license ownership of 


```{r,warning=FALSE}
# 18-24 old at UW main campus vs 18-24 at a community college, vs 18-24 year in the region over all)
df_col11 <- df.person_c %>%
  filter(!is.na(license)) %>%
  mutate(license=as.factor(ifelse(license==1,'Yes','No'))) %>%
  mutate(age_group = ifelse(age==1,'16-17',
                            ifelse(age %in% c(2,3,4), '18-24','>25')))

#ggplot(df_col11,aes(college,fill=license)) + geom_bar(position = 'dodge') + facet_grid(age_group~.,scales = 'free_y')
ggplot(df_col11,aes(age_group,fill=license)) + geom_bar(aes(weight = expwt),position = 'dodge') 
  
```

## 12. taxi_type vs. region (regular cab vs. uber/lyft)


```{r,warning=FALSE}
df_col12 <- df_col %>%
  filter(!is.na(taxi_type)) %>%
  mutate(taxi_type = as.factor(ifelse(taxi_type==1,'regular','uber/lyft'))) 

df.trip12 <- df.trip %>%
  filter(!is.na(taxi_type),taxi_type!=-99) %>%
  mutate(taxi_type = as.factor(ifelse(taxi_type==1,'regular','uber/lyft'))) 

taxi<-rbind(wpct(df.trip12$taxi_type,df.trip12$expwt_final),
            wpct(df_col12$taxi_type,df_col12$expwt))
rownames(taxi)<-c("reginal","college")
round(taxi,digits = 3)
```



```{r,warning=FALSE}

```



```{r,warning=FALSE}

```
