---
title: "part2telecommute"
author: "suzanne"
date: "2024-07-12"
output: html_document
---
First run telecommute_preprocess
These are the basic summaries below, and show work at home by demographic characteristics.

```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
library(forcats)
source('survey-23-preprocess.R')

```

```{r}

employment<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'workplace',
                               summarize_by = c('employment','survey_year'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```
```{r}
employment_workplace<-employment$summary$wtd
write.csv(employment_workplace, 'employment_workplace.csv')
```

```{r}
employment_workplace_23<-employment_workplace%>%filter(survey_year=='2023')%>%filter(employment!='Unpaid volunteer or intern')

write.csv(employment_workplace_23, 'employment_workplace_23.csv')
```

```{r}
library(scales)
static<-static_bar_chart(employment_workplace_23, x= 'prop' ,fill='workplace', y='employment', pos='stack')+ 

static
                                                                                                                
```


```{r}

telecommute_or_not<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_freq_cond',
                               summarize_by = 'survey_year',
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```

```{r}

telecommute_summary<-telecommute_or_not$summary$wtd%>%filter(telecommute_freq_cond=='1+ days per week')
static<-static_line_chart(telecommute_summary, y='prop', x='survey_year', fill='telecommute_freq_cond',lwidth=1.5) + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+ scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,0.6))

static

write.csv(telecommute_summary,'telecommuteornot_summary_trend.csv')
```

Who telecommuted Hybrid vs Not?

```{r}

telecommute_status <- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_status',
                               summarize_by = 'survey_year',
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```


```{r}

telecommute_summary<-telecommute_status$summary$wtd%>%mutate(telecommute_status=forcats::fct_reorder(telecommute_status, -prop))
static<-static_column_chart(telecommute_summary, y='prop', x='survey_year', fill='telecommute_status') + scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,0.9))+ theme(legend.text=element_text(size=10))

inter<-interactive_column_chart(telecommute_summary, y='prop', x='telecommute_status', fill='survey_year', xlabel='Work Location Status', ylabel='Worker Share')
static
inter
write.csv(telecommute_summary,'telecommute_summary_trend.csv')
```
```{r}

telecommute_status_broad <- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_status_broad',
                               summarize_by = 'survey_year',
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```


```{r}

telecommute_summary<-telecommute_status_broad$summary$wtd%>%mutate(telecommute_status_broad=forcats::fct_reorder(telecommute_status_broad, prop))
static<-static_line_chart(telecommute_summary, y='prop', x='survey_year', fill='telecommute_status_broad', lwidth=1.5) + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,.2))

inter<-interactive_column_chart(telecommute_summary, y='prop', x='telecommute_status_broad', fill='survey_year', xlabel='Work Location Status', ylabel='Worker Share')
static
inter
write.csv(telecommute_summary,'telecommute_summary_trend_broad.csv')
```

Location
Where did people work who telecommuted? What modes did they use prior in 2019 by location?

Where did people live who telecommuted? What modes did they use prior in 2019 by location?

Household Income

```{r}

telecommute_or_not_inc <- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_freq_cond',
                               summarize_by = c('survey_year', 'hhincome_broad'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```
```{r}

telecommute_summary_inc<-telecommute_or_not_inc$summary$wtd%>%filter(telecommute_freq_cond=='1+ days per week')%>%filter(survey_year=='2023')%>%filter(hhincome_broad!='Prefer not to answer')%>%mutate(moe=prop_se*1.645)%>%mutate(hhincome_broad=fct_relevel(hhincome_broad,
                                                  c('Under $75,000',
                                                      '$75,000-$199,999',
                                                      '$200,000 or more')))
                                                                                                                                                                                                           


static<-static_column_chart(telecommute_summary_inc, y='prop', x='hhincome_broad', fill='survey_year') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,.75))

static

write.csv(telecommute_summary_inc,'telecommute_summary_inc.csv')

```

Race and Ethnicity

Gender
```{r}

telecommute_or_not_gender <- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_freq_cond',
                               summarize_by = c('survey_year', 'gender'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```
```{r}

telecommute_summary_gend<-telecommute_or_not_gender$summary$wtd%>%filter(telecommute_freq_cond=='1+ days per week')%>%filter(survey_year=='2023')%>%mutate(gender=recode(gender, 	
'Boy/Man (cisgender or transgender)' = "Man/Boy", 'Girl/Woman (cisgender or transgender)'  = "Woman/Girl"))%>%filter(!gender %in% c('Non-binary/Something else fits better', 'Prefer not to answer'))
                                                                                                   
static<-static_column_chart(telecommute_summary_gend, y='prop', x='gender', fill='survey_year') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,.75))

static

write.csv(telecommute_summary_gend,'telecommute_summary_gender.csv')

```
Gender
```{r}

telecommute_or_not_gender <- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_status',
                               summarize_by = c('survey_year', 'gender'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```
```{r}

telecommute_summary_gend<-telecommute_or_not_gender$summary$wtd%>%filter(survey_year=='2023')%>%mutate(gender=recode(gender, 	
'Boy/Man (cisgender or transgender)' = "Man/Boy", 'Girl/Woman (cisgender or transgender)'  = "Woman/Girl"))%>%filter(!gender %in% c('Non-binary/Something else fits better', 'Prefer not to answer'))
                                                                                                   
static<-static_column_chart(telecommute_summary_gend, y='prop', x='gender', fill='telecommute_status', pos='stack') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,.75))

static

write.csv(telecommute_summary_gend,'telecommute_summary_gender.csv')

```

```{r}

telecommute_or_not_race<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_freq_cond',
                               summarize_by = c('survey_year', 'race_category'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```
```{r}

telecommute_summary_race<-telecommute_or_not_race$summary$wtd%>%filter(telecommute_freq_cond=='1+ days per week')%>%filter(survey_year=='2023')%>%filter(!race_category %in% c('Missing/No response', 'Child'))%>%mutate(moe=prop_se*1.645)
                                                                                                   
static<-static_column_chart(telecommute_summary_race, y='prop', x='race_category', fill='survey_year', moe='moe') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,.75))

static

write.csv(telecommute_summary_race,'telecommute_summary_race.csv')

```
```{r}

industryrace<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'industry_cond',
                               summarize_by = c('survey_year', 'race_category'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```

```{r}

industry_race<-industryrace$summary$wtd%>%filter(survey_year=='2023')%>%filter(!race_category%in% c('Child', 'Missing'))%>%mutate(moe=prop_se*1.645)
                                                                                                   
static<-static_column_chart(industry_race, y='prop', x='race_category', fill='industry_cond', moe='moe') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,.75))

static

write.csv(telecommute_summary_race,'telecommute_summary_race.csv')

```
Gender
```{r}

telecommute_or_not_age<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_freq_cond',
                               summarize_by = c('survey_year', 'age_grp_2'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```
```{r}

telecommute_summary_age<-telecommute_or_not_age$summary$wtd%>%filter(telecommute_freq_cond=='1+ days per week')%>%filter(survey_year=='2023')%>%mutate(moe=prop_se*1.645)%>%filter(!is.na(age_grp_2))%>%
  filter(age_grp_2!='Under 18')
                                                                                                   
static<-static_column_chart(telecommute_summary_age, y='prop', x='age_grp_2', fill='survey_year') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,.75))

static

write.csv(telecommute_summary_age,'telecommute_summary_age.csv')
```


Job Industry

```{r}

telecommute_or_not_ind <- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_freq_cond',
                               summarize_by = c('survey_year', 'industry_cond'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```
```{r}

telecommute_summary_ind<-telecommute_or_not_ind$summary$wtd%>%filter(telecommute_freq_cond=='1+ days per week')%>%filter(survey_year=='2023')%>%mutate(moe=prop_se*1.645)%>%mutate(industry_cond=forcats::fct_reorder(industry_cond, prop))%>%filter(industry_cond!="Missing: Skip Logic")
static<-static_bar_chart(telecommute_summary_ind, x='prop', y='industry_cond', fill='survey_year') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,1))

static

write.csv(telecommute_summary_ind,'telecommute_summary_industry.csv')

```
```{r}

telecommute_or_not_edu<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_freq_cond',
                               summarize_by = c('survey_year', 'edu_simple'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```

```{r}

telecommute_summary_edu<-telecommute_or_not_edu$summary$wtd%>%filter(telecommute_freq_cond=='1+ days per week')%>%filter(survey_year=='2023')%>%mutate(moe=prop_se*1.645)%>%filter(edu_simple!="Missing: Skip Logic")
static<-static_bar_chart(telecommute_summary_edu, x='prop', y='edu_simple', fill='survey_year') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,1))

static

write.csv(telecommute_summary_ind,'telecommute_summary_education.csv')

```
Sexuality

```{r}

telecommute_or_not_sex <- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_freq_cond',
                               summarize_by = c('survey_year', 'sexuality_broad'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```
```{r}

telecommute_summary_sex<-telecommute_or_not_sex$summary$wtd%>%filter(telecommute_freq_cond=='1+ days per week')%>%filter(survey_year=='2023')%>%mutate(moe=prop_se*1.645)%>%filter(sexuality_broad!="Missing: Skip Logic")
static<-static_bar_chart(telecommute_summary_sex, x='prop', y='sexuality_broad', fill='survey_year') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,1))

static

write.csv(telecommute_summary_ind,'telecommute_summary_sexuality.csv')

```


Vehicle Count


```{r}

telecommute_or_not_veh <- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_freq_cond',
                               summarize_by = c('survey_year', 'has_vehicle'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```
```{r}

telecommute_summary_veh<-telecommute_or_not_veh$summary$wtd%>%filter(telecommute_freq_cond=='1+ days per week')%>%filter(survey_year=='2023')%>%mutate(moe=prop_se*1.645)

static<-static_bar_chart(telecommute_summary_veh, x='prop', y='has_vehicle', fill='survey_year') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,1))

static

write.csv(telecommute_summary_veh,'telecommute_summary_vehicles.csv')

```

```{r}

industry_income<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'industry_cond',
                               summarize_by = c('survey_year', 'hhincome_broad'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```

```{r}

industry_income<-industry_income$summary$wtd%>%filter(survey_year=='2023')%>%mutate(moe=prop_se*1.645)

static<-static_column_chart(industry_income, y='prop', x='hhincome_broad', fill='industry_cond', moe='moe') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,1))

static

write.csv(industry_income, 'industry_income.csv')

```
```{r}

industry_race<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'industry_cond',
                               summarize_by = c('survey_year', 'race_category'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```

```{r}

industry_race<-industry_race$summary$wtd%>%filter(survey_year=='2023')%>%mutate(moe=prop_se*1.645)

static<-static_column_chart(industry_race, y='prop', x='race_category', fill='industry_cond', moe='moe') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,1))

static

write.csv(industry_race, 'industry_race.csv')

```
```{r}

telecommute_status_inc<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_status',
                               summarize_by = c('survey_year', 'hhincome_broad'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```

```{r}

telecommute_summary_inc<-telecommute_status_inc$summary$wtd%>%filter(survey_year=='2023')%>%filter(hhincome_broad!='Prefer not to answer')%>%mutate(moe=prop_se*1.645)%>%mutate(hhincome_broad=fct_relevel(hhincome_broad,
                                                  c('Under $75,000',
                                                      '$75,000-$199,999',
                                                      '$200,000 or more')))                                                                                                                              


static<-static_column_chart(telecommute_summary_inc, y='prop', x='hhincome_broad', fill='telecommute_status') +
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,1))

static

write.csv(telecommute_summary_inc,'telecommute_summary_inc.csv')

```

```{r}

telecommute_status_sector<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_status',
                               summarize_by = c('survey_year', 'industry_cond'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```

```{r}

telecommute_summary_sector<-telecommute_status_sector$summary$wtd%>%filter(survey_year=='2023')%>%mutate(moe=prop_se*1.645)%>%filter(industry_cond!="Missing: Skip Logic")%>%
  mutate(telecommute_status=fct_reorder(telecommute_status,est))%>%
  mutate(industry_cond=fct_reorder(industry_cond,est))%>%filter(telecommute_status!='Fully In Person')
                                                                                                                                                                                                           


static<-static_bar_chart(telecommute_summary_sector, x='est', y='industry_cond', fill='telecommute_status', pos='stack') + theme(axis.text.x=element_text(size=12), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
   
static

write.csv(telecommute_summary_sector,'telecommute_summary_industry_hybrid.csv')

```
```{r}

telecommute_status_sector<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_status',
                               summarize_by = c('survey_year', 'industry'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```

```{r}

telecommute_summary_sector<-telecommute_status_sector$summary$wtd%>%filter(survey_year=='2023')%>%mutate(moe=prop_se*1.645)%>%filter(industry!="Missing: Skip Logic")%>%
  mutate(telecommute_status=fct_reorder(telecommute_status,prop))
                                                                                                                                                                                                           


static<-static_column_chart(telecommute_summary_sector, y='prop', x='industry', fill='telecommute_status') + theme(axis.text.x=element_text(size=12), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,1))

static

write.csv(telecommute_summary_sector,'telecommute_summary_industrydetail_hybrid.csv')

```

```{r}

telecommute_status_sector<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_status',
                               summarize_by = c('survey_year', 'industry_usim'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```

```{r}

telecommute_summary_sector<-telecommute_status_sector$summary$wtd%>%filter(survey_year=='2023')%>%mutate(moe=prop_se*1.645)%>%filter(industry_usim!="Missing: Skip Logic")%>%
  mutate(telecommute_status=fct_reorder(telecommute_status,prop))
                                                                                                                                                                                                           


static<-static_column_chart(telecommute_summary_sector, y='prop', x='industry_usim', fill='telecommute_status') + theme(axis.text.x=element_text(size=12), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,1))

static

write.csv(telecommute_summary_sector,'telecommute_summary_industry_usim.csv')

```


##disability status
```{r}

telecommute_or_not_disab<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_freq_cond',
                               summarize_by = c('survey_year', 'disability_person'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```
```{r}

telecommute_summary_disab<-telecommute_or_not_disab$summary$wtd%>%filter(telecommute_freq_cond=='1+ days per week')%>%filter(survey_year=='2023')%>%mutate(moe=prop_se*1.645)
                                                                                                   
static<-static_column_chart(telecommute_summary_disab, y='prop', x='disability_person', fill='survey_year', moe='moe') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,.75))

static

write.csv(telecommute_summary_disab,'telecommute_summary_disab.csv')

```


```{r}

telecommute_status<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'telecommute_status',
                               summarize_by = c('survey_year', 'disability_person'),
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight'
                               )

```
```{r}

telecommute_summary_disab<-telecommute_status$summary$wtd%>%filter(survey_year=='2023')%>%mutate(moe=prop_se*1.645)
                                                                                                   
static<-static_column_chart(telecommute_summary_disab, y='prop', x='disability_person', fill='telecommute_status', moe='moe') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+
   scale_y_continuous(labels = scales::percent_format(scale = 100), limits=c(0,.75))

static

write.csv(telecommute_summary_disab,'telecommute_summary_disab.csv')

```
