---
title: "telecommute-summary"
author: "suzanne"
date: "2023-02-17"
output: pdf_document
---

I'm comparing estimates of telecommuting on an average weekday from the 2019/2021 household travel surveys to Replica data.

Differences- Replica data is Fall 2019 and Fall 2021; ours is from Spring 2019 and Spring 2021.
Their data is labeled with work from home (on an average weekday). We just have a field for telework_time. A person can telework at home, and go into work. In 2021, I'd guess that happened frequently.  For example, I usually work from home on days that I go into work.

```{r}
library(psrcplot)
library(psrc.travelsurvey)

```


# Read in Replica data for Working at home by income and race and ethnicity, 2019 and 2021
```{r setup, include=FALSE}
replica_data_fldr<- 'C:/GitHub/travel-studies/2021/analysis_in_progress/telecommuting/replica-estimates'
telecommute_2019_inc <- read.csv(file.path(replica_data_fldr, 'northwest-2019_Q4-thursday-household-income.csv'))
telecommute_2021_inc <- read.csv(file.path(replica_data_fldr, 'northwest-2021_Q4-thursday-household-income.csv'))
telecommute_2019_race_eth <- read.csv(file.path(replica_data_fldr, 'northwest-2019_Q4-thursday-race-and-ethnicity.csv'))
telecommute_2021_race_eth <- read.csv(file.path(replica_data_fldr, 'northwest-2021_Q4-thursday-race-and-ethnicity.csv'))


```

## R Markdown


# Read in Household Travel Survey data on the same items
```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
