---
title: "telecommute-summary"
author: "suzanne"
date: "2023-02-17"
output: pdf_document
---

I'm comparing estimates of telecommuting on an average weekday from the 2019/2021 household travel surveys to Replica data.

Differences- Replica data is Fall 2019 and Fall 2021; ours is from Spring 2019 and Spring 2021.
Their data is labeled with work from home (on an average weekday). We just have a field for telework_time. A person can telework at home, and go into work. In 2021, I'd guess that happened frequently.  For example, I usually work from home on days that I go into work.

```{r}
library(psrcplot)
library(psrc.travelsurvey)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
install_psrc_fonts()

```
```{r}
replica_data_fldr<- 'C:/GitHub/travel-studies/2021/analysis_in_progress/telecommuting/replica-estimates/'

telecommute_2019_inc_repl <- read.csv(paste0(replica_data_fldr, 'northwest-2019_Q4-thursday-household-income.csv'))
telecommute_2019_race_eth_repl <- read.csv(paste0(replica_data_fldr, 'northwest-2019_Q4-thursday-race-and-ethnicity.csv'))



telecommute_2021_inc_repl <- read.csv(paste0(replica_data_fldr, 'northwest-2021_Q4-thursday-household-income.csv'))
telecommute_2021_race_eth_repl <- read.csv(paste0(replica_data_fldr, 'northwest-2021_Q4-thursday-race-and-ethnicity.csv'))

telecommute_2019_inc_repl_p<-telecommute_2019_inc_repl%>%group_by(psrc_income)%>%
  summarize(count = sum(count))%>%
  mutate(year=2019)
telecommute_2021_inc_repl_p<-telecommute_2021_inc_repl%>%group_by(psrc_income)%>%
  summarize(count = sum(count))%>%
mutate(year=2021)

telecommute_2019_race_eth_repl_p<-telecommute_2019_race_eth_repl%>%group_by(race_eth_broad)%>%summarize(count = sum(count))%>%
  mutate(year=2019)
telecommute_2021_race_eth_repl_p<-telecommute_2021_race_eth_repl%>%group_by(race_eth_broad)%>%summarize(count = sum(count))%>%
mutate(year=2021)

telecommute_inc_repl<-rbind(telecommute_2019_inc_repl_p, telecommute_2021_inc_repl_p)%>%
  mutate(source='Replica')
telecommute_race_repl<-rbind(telecommute_2019_race_eth_repl_p, telecommute_2021_race_eth_repl_p)%>%
  mutate(source='Replica')
```



Read in Household Travel Survey data on the same items,selected teleworking more than one hour on a day realized that Repilica was using a stricter definition of always teleworking
# 2019
```{r }
days_2019<- get_hhts(survey ='2019', level='d', vars= c('hhincome_detailed', 'telework_time', 'race_eth_broad'))
telework_days_2019<-days_2019%>%drop_na(telework_time)%>%filter(!telework_time %in% c('None', '1 hour'))

# map income to same scale:
telework_days_2019<- telework_days_2019 %>%
  mutate(psrc_income = case_when(hhincome_detailed %in% c('Under $10,000', 	'$10,000-$24,999')~'Under $25,000',
                                hhincome_detailed %in% c('$25,000-$34,999', 	'$35,000-$49,999')~'$25,000-$49,999',
                                hhincome_detailed %in% c('$200,000-$249,999', 	'$250,000 or more')~'$200,000 or more',
                                .default = hhincome_detailed))

                                
telecommute_2019_inc_hhts<- hhts_count(df=telework_days_2019, group_vars='psrc_income', spec_wgt='day_weight_2019')%>%
  rename(year=survey)%>%mutate(source='Travel Survey Workers with any Telework Time')%>%select(count, psrc_income, year, source)


telecommute_2019_race_eth_hhts<- hhts_count(df=telework_days_2019, group_vars='race_eth_broad', spec_wgt='day_weight_2019')%>%
  rename(year=survey)%>%mutate(source='Travel Survey Workers with any Telework Time')%>%select(count, race_eth_broad, year, source)




```


# 2021
```{r}

days_2021<- get_hhts(survey ='2021', level='d', vars= c('hhincome_detailed', 'telework_time', 'race_eth_broad'))
telework_days_2021<-days_2021%>%filter(telework_time %in% c('1-6 hours', 'More than 6 hours'))



telework_days_2021<- telework_days_2021 %>%
  mutate(psrc_income = case_when(hhincome_detailed %in% c('Under $10,000', 	'$10,000-$24,999')~'Under $25,000',
                                hhincome_detailed %in% c('$25,000-$34,999', 	'$35,000-$49,999')~'$25,000-$49,999',
                                hhincome_detailed %in% c('$200,000-$249,999', 	'$250,000 or more')~'$200,000 or more',
                                .default = hhincome_detailed))

telecommute_2021_inc_hhts<- hhts_count(df=telework_days_2021, group_vars='psrc_income')%>%
  rename(year=survey)%>%mutate(source='Travel Survey Workers with any Telework Time')%>%select(count, psrc_income, year, source)

telecommute_2021_race_eth_hhts<- hhts_count(df=telework_days_2021, group_vars='race_eth_broad')%>%
  rename(year=survey)%>%mutate(source='Travel Survey Workers with any Telework Time')%>%select(count,race_eth_broad, year, source)




```

```{r}
telecommute_inc_hhts<-rbind(telecommute_2019_inc_hhts, telecommute_2021_inc_hhts)
telecommute_race_eth_hhts<-rbind(telecommute_2019_race_eth_hhts, telecommute_2021_race_eth_hhts)

```


Maybe the Replica data more aligns with the "workplace variable"

Usually the same location (outside home)
Workplace regularly varies (different offices or jobsites)
At home (telecommute or self-employed with home office)
Drives for a living (e.g., bus driver, salesperson)
Telework some days and travel to a work location some days


# Switch to the person table
# 2019
```{r }
days_2019<- get_hhts(survey ='2019', level='p', vars= c('hhincome_detailed', 'workplace', 'race_eth_broad'))
telework_days_2019<-days_2019%>%drop_na(workplace)%>%filter(workplace == 'At home (telecommute or self-employed with home office)')

# map income to same scale:
telework_days_2019<- telework_days_2019 %>%
  mutate(psrc_income = case_when(hhincome_detailed %in% c('Under $10,000', 	'$10,000-$24,999')~'Under $25,000',
                                hhincome_detailed %in% c('$25,000-$34,999', 	'$35,000-$49,999')~'$25,000-$49,999',
                                hhincome_detailed %in% c('$200,000-$249,999', 	'$250,000 or more')~'$200,000 or more',
                                .default = hhincome_detailed))

                                
telecommute_always_2019_inc_hhts<- hhts_count(df=telework_days_2019, group_vars='psrc_income', spec_wgt='hh_weight_2019')%>%
  rename(year=survey)%>%mutate(source='Travel Survey, Always Work at Home')%>%select(count, psrc_income, year, source)


telecommute_always_2019_race_eth_hhts<- hhts_count(df=telework_days_2019, group_vars='race_eth_broad', spec_wgt='hh_weight_2019')%>%
  rename(year=survey)%>%mutate(source='Travel Survey, Always Work at Home')%>%select(count, race_eth_broad, year, source)




```


# switch to the person table
```{r}

days_2021<- get_hhts(survey ='2021', level='p', vars= c('hhincome_detailed', 'workplace', 'race_eth_broad'))
telework_days_2021<-days_2021%>%filter(workplace == 'At home (telecommute or self-employed with home office)')



telework_days_2021<- telework_days_2021 %>%
  mutate(psrc_income = case_when(hhincome_detailed %in% c('Under $10,000', 	'$10,000-$24,999')~'Under $25,000',
                                hhincome_detailed %in% c('$25,000-$34,999', 	'$35,000-$49,999')~'$25,000-$49,999',
                                hhincome_detailed %in% c('$200,000-$249,999', 	'$250,000 or more')~'$200,000 or more',
                                .default = hhincome_detailed))

telecommute_always_2021_inc_hhts<- hhts_count(df=telework_days_2021, group_vars='psrc_income')%>%
  rename(year=survey)%>%mutate(source='Travel Survey, Always Work at Home')%>%select(count, psrc_income, year, source)

telecommute_always_2021_race_eth_hhts<- hhts_count(df=telework_days_2021, group_vars='race_eth_broad')%>%
  rename(year=survey)%>%mutate(source='Travel Survey, Always Work at Home')%>%select(count,race_eth_broad, year, source)




```

```{r}
telecommute_inc_hhts_always<-rbind(telecommute_always_2019_inc_hhts, telecommute_always_2021_inc_hhts)
telecommute_race_eth_hhts_always<-rbind(telecommute_always_2019_race_eth_hhts, telecommute_always_2021_race_eth_hhts)

```

put it together and what do we get?
```{r}
telecommute_inc<-rbind(telecommute_inc_repl,  telecommute_inc_hhts_always)%>%filter(psrc_income!='Total')%>%mutate(psrc_income= factor(psrc_income, levels=c('Under $25,000', '$25,000-$49,999', '$50,000-$74,999', '$75,000-$99,999', '$100,000-$149,999', '$150,000-$199,999', '$200,000 or more')))
telecommute_race_eth<-rbind(telecommute_race_repl, telecommute_race_eth_hhts_always)%>%filter(race_eth_broad!='Total')%>%mutate(race_eth_broad= word(race_eth_broad, 1))

```

```{r}

static_facet_column_chart(telecommute_inc, x ='psrc_income', y= 'count',  facet='year', fill='source', title='Always Telecommuters by Income', source='Replica, Fall 2019/2021 PSRC travel survey Spring 2019/2021', est='number')


```

```{r}

static_facet_column_chart(telecommute_race_eth, x ='race_eth_broad', y= 'count',  facet='year', fill='source', title='Always Telecommuters by Race and Ethnicity', source='Replica, Fall 2019/2021 PSRC travel survey Spring 2019/2021', est='number')


```

```{r}
telecommute_inc<-rbind(telecommute_inc_repl,  telecommute_inc_hhts)%>%filter(psrc_income!='Total')%>%mutate(psrc_income= factor(psrc_income, levels=c('Under $25,000', '$25,000-$49,999', '$50,000-$74,999', '$75,000-$99,999', '$100,000-$149,999', '$150,000-$199,999', '$200,000 or more')))
telecommute_race_eth<-rbind(telecommute_race_repl, telecommute_race_eth_hhts)%>%filter(race_eth_broad!='Total')%>%mutate(race_eth_broad= word(race_eth_broad, 1))

```

```{r}

static_facet_column_chart(telecommute_inc, x ='psrc_income', y= 'count',  facet='year', fill='source', title='Telecommuters by Income: Teleworking more than an hour on a Day from HHTS, Always from Replica ', source='Replica, Fall 2019/2021 PSRC travel survey Spring 2019/2021', est='number')


```

```{r}

static_facet_column_chart(telecommute_race_eth, x ='race_eth_broad', y= 'count',  facet='year', fill='source', title='Telecommuters by Race and Ethnicity: Teleworking more than an hour on a Day from HHTS, Always from Replica ', source='Replica, Fall 2019/2021 PSRC travel survey Spring 2019/2021', est='number')


```

