---
title: "test-new-package"
author: "suzanne"
date: "2024-09-27"
output: html_document
---
reading in libraries I usually need
```{r , include=FALSE}
library(psrc.travelsurvey)
library(travelSurveyTools)
library(data.table)
library(dplyr)
library(psrcplot)
library(stringr)
library(sf)
library(psrcelmer)
library(forcats)
```


I'm going to analyze sexuality and travel behavior data as a test case-won't be able to do a multi-year comparision though. So after that I'll do a summary of transit access modes.

```{r}
demo_summary_vars<-c('age', 'disability_person', 'education', 'employment','gender', 'industry',
               'hhincome_broad', 'home_county','home_lat', 'home_lng','hh_race_category','home_rgcname',
               'numadults', 'numchildren', 'numworkers','res_type', 'rent_own','sexuality',
               'student', 'vehicle_count')
travel_summary_vars<-c('bike_freq', 'benefits_3', 'dest_purpose', 'commute_freq', 'commute_mode',
                       'mode_characterization', 'mode_acc', 'transit_freq', 'tnc_freq','telecommute_freq','walk_freq','workplace')

all_vars<-c(demo_summary_vars, travel_summary_vars)
               

```

```{r}
hts_data <- get_psrc_hts(survey_vars = all_vars)

```

```{r}
hh<-hts_data$hh
hh_geo<-hh%>%
  st_as_sf(coords=c('home_lng','home_lat'), crs=4326)
```

#redo analysis with regional geographies
```{r}
regional.geogs<-st_read_elmergeo('REGIONAL_GEOGRAPHIES_evw')
```



```{r}
sf_use_s2(FALSE)
hh_reg_geog<-st_join(regional.geogs, hh_geo, join=st_intersects)%>%filter(!is.na(hh_id))

hh<-left_join(hh, hh_reg_geog, by='hh_id', suffix=c('','_geog'))%>%
  mutate(class_desc=if_else(is.na(class_desc), "Outside Urban Growth Area", class_desc))%>%
  mutate(home_loc= case_when(class_desc=='CitiesTowns' ~ 'Cities and Towns',
                             class_desc=='HCT' ~ 'High Capacity Transit',
                             class_desc=='UU' ~ 'Urban Unincorporated',
                             TRUE ~ class_desc
    
  ))%>%mutate(home_loc=factor(home_loc))
  
```

```{r}
hts_data$hh<-hh
```

```{r}
hts_data<-hts_data%>%hts_bin_sexuality()%>%hts_bin_age()
```

I realized I need a different categorization, recode the prefer not to answer to NA, so that we don't summarize them
```{r}
hts_data$person<-hts_data$person%>%
  mutate(sexuality_bin4=factor(case_when(sexuality=='Prefer not to answer' ~ NA,
                                  str_detect(sexuality_bin3, 'gay') ~'Gay/Don\'t Know/Other',
                                 .default=sexuality_bin3)))

```

```{r}
sexuality_age<-psrc_hts_stat(hts_data, "person", c('age_bin5','sexuality_bin4'), incl_na=FALSE)%>%
  filter(survey_year=='2023')%>% # only started asking in 2023%>%
filter(age_bin5!='Under 18 Years')

    
   
```

```{r}
static<-static_column_chart(sexuality_age, y='prop', x='age_bin5', fill='sexuality_bin4')
static
```

```{r}
sexuality_mode<-psrc_hts_stat(hts_data, "trip", c('sexuality_bin4','mode_characterization'), incl_na=FALSE)%>%
  filter(survey_year=='2023')%>%# only started asking in 2023%>%
mutate(mode_characterization=fct_reorder(mode_characterization, -prop))

    
   
```

```{r}
static<-static_column_chart(sexuality_mode, y='prop', x='mode_characterization', fill='sexuality_bin4', moe='prop_moe')
static
```


```{r}
sexuality_county<-psrc_hts_stat(hts_data, "person", c('home_county','sexuality_bin4'), incl_na=FALSE)%>%
  filter(survey_year=='2023')# only started asking in 2023%>%


    
   
```

```{r}
static<-static_column_chart(sexuality_county, y='prop', fill='home_county', x='sexuality_bin4')
static
```
```{r}
sexuality_income<-psrc_hts_stat(hts_data, "person", c('hhincome_broad','sexuality_bin4'), incl_na=FALSE)%>%
  filter(survey_year=='2023')# only started asking in 2023%>%


    
   
```

```{r}
static<-static_column_chart(sexuality_income, y='prop', fill='hhincome_broad', x='sexuality_bin4')
static
```
```{r}
sexuality_child<-psrc_hts_stat(hts_data, "person", c('sexuality_bin4','numchildren'), incl_na=FALSE)%>%
  filter(survey_year=='2023')# only started asking in 2023%>%


    
   
```

```{r}
static<-static_column_chart(sexuality_child, y='prop', fill='numchildren', x='sexuality_bin4')
static
```

```{r}
sexuality_education<-psrc_hts_stat(hts_data, "person", c('sexuality_bin4','education'), incl_na=FALSE)%>%
  filter(survey_year=='2023')# only started asking in 2023%>%


    
   
```

```{r}
static<-static_column_chart(sexuality_education, y='prop', fill='education', x='sexuality_bin4')
static
```
```{r}
sexuality_home_loc<-psrc_hts_stat(hts_data, "person", c('sexuality_bin4','home_loc'), incl_na=FALSE)%>%
  filter(survey_year=='2023')# only started asking in 2023


    
   
```

```{r}
static<-static_column_chart(sexuality_home_loc, y='prop', x='home_loc', fill='sexuality_bin4', moe='prop_moe')
static
```
