---
title: "summary-write-out"
author: "suzanne"
date: "2024-09-05"
output: html_document
---


topic_list<-c('mode_characterization', 'dest_purpose_simpler')


mode_characterization
dest_purpose_simpler
X
race_category
hhincome_broad
gender_grp
home_county
X
year


first run general preprocess

```{r}
trip_topic_list<-c('mode_characterization'='Trip Mode', 'dest_purpose_simpler'='Trip Purpose')
demo_geo_list<-c('hhincome_broad'='Household Income', 'gender_grp'='Person Gender', 'home_county'='Home County','race_category'= 'Person Race Category')
years_list<-hh%>%distinct(survey_year)%>%select(survey_year)
```



```{r setup, include=FALSE}

hh_data<-
  list(hh=hh, person=person, day=day, trip=trip)
ids = c('hh_id', 'person_id', 'day_id', 'trip_id')
wts = c('hh_weight', 'person_weight', 'day_weight', 'trip_weight')
```

```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```

```{r}
i<-0
for(topic_var in names(trip_topic_list)){
  
  for(demo_geo_var in names(demo_geo_list)){

      
      summarize_df<- summarize_weighted(hts_data= hh_data,
                               summarize_var = topic_var,
                               summarize_by = c('survey_year', demo_geo_var),
                               id_cols=ids,
                               wt_cols=wts,
                               wtname= 'trip_weight'
                               )

  
     crosstab_df<- summarize_df$summary$wtd
     
     crosstab_df$variable_name<-topic_var
     
     crosstab_df<-crosstab_df%>%rename(variable_value:={{topic_var}})
     crosstab_df<-crosstab_df%>%rename(demo_geo_value:={{demo_geo_var}})
     crosstab_df$demography_geography_name<-demo_geo_var
     crosstab_df$variable_nice_name<-trip_topic_list[[topic_var]]
     crosstab_df$demo_geography_nice_name<-demo_geo_list[[demo_geo_var]]
  
     crosstab_df<-crosstab_df%>%select(variable_name, variable_nice_name,variable_value, demography_geography_name,  demo_geography_nice_name,survey_year, count, prop, prop_se, est, est_se)
      print(crosstab_df)
      #crosstab_df<-crosstab_df%>%select(var_name, var_label, count,est, est_se, prop, prop_se)%>%
       # mutate(survey_year=the_year)
      
      if(i==0){
        crosstab_df_out<-crosstab_df
      }else{
        crosstab_df_out<-rbind(crosstab_df_out, crosstab_df)
      }
    
    i<-i+1
    
  }
}
write.csv(crosstab_df_out, 'data_for_app.csv')
```



