---
title: "Centers monitoring: 2023 mode share"
date: today
format: 
  html:
    mainfont: Poppins
    theme: [cosmo, custom.scss]
    toc: true
    df-print: kable
    warning: false
    echo: false
---

```{r}
source("region_dest_data.R")
source("region_dest_data_2023_new.R")

pal_rgc <- c(psrc_colors$purples_inc[2],psrc_colors$greens_inc[2])
pal_rgc <- setNames(pal_rgc, c("RGC","Not RGC"))

pal_rgc_mu <- c(psrc_colors$oranges_inc[2],psrc_colors$blues_inc[2],psrc_colors$greens_inc[2])
pal_rgc_mu <- setNames(pal_rgc_mu, c("Metro","Urban","Not RGC"))

```


## mode share for trips to centers

- Showing mode share for trips destined to centers to understand how do people get to centers 

### all trips


```{r}
#| layout: [[1,3], [1,3]]
#| out-height: 300px
#| out-width: 800px

prepped_dt <- hts_prep_variable(summarize_var = 'mode_simple',
                                summarize_by = c('survey_year', 'd_rgcname'),
                                data = hts_data,
                                id_cols='trip_id',
                                wt_cols='trip_weight',
                                weighted=TRUE)


summary_dt <- hts_summary(prepped_dt = prepped_dt$cat,
                            summarize_var = 'mode_simple',
                            summarize_by = c('survey_year', 'd_rgcname'),
                            summarize_vartype = 'categorical',
                            id_cols = 'trip_id',
                            wtname ='trip_weight',
                            weighted=TRUE,
                            se=TRUE)

df_plot <- summary_dt$summary$wtd %>%
  filter(survey_year=="2023") %>%
  rename(share = prop,
         sample = count,
         count = est)
df_plot$d_rgcname <- factor(df_plot$d_rgcname, levels=c("RGC","Not RGC"))
df_plot$mode_simple <- factor(df_plot$mode_simple, levels=c("Drive","Transit","Walk","Bike","Other"))



prepped_dt <- hts_prep_variable(summarize_var = 'mode_simple',
                                summarize_by = c('survey_year', 'category'),
                                data = hts_data,
                                id_cols='trip_id',
                                wt_cols='trip_weight',
                                weighted=TRUE)


summary_dt <- hts_summary(prepped_dt = prepped_dt$cat,
                            summarize_var = 'mode_simple',
                            summarize_by = c('survey_year', 'category'),
                            summarize_vartype = 'categorical',
                            id_cols = 'trip_id',
                            wtname ='trip_weight',
                            weighted=TRUE,
                            se=TRUE)

df_plot2 <- summary_dt$summary$wtd %>%
  filter(survey_year=="2023") %>%
  rename(share = prop,
         sample = count,
         count = est)
df_plot2$category <- factor(df_plot2$category, levels=c("Metro", "Urban", "Not RGC"))
df_plot2$mode_simple <- factor(df_plot2$mode_simple, levels=c("Drive","Transit","Walk","Bike","Other"))


df_drive <- df_plot %>% filter(mode_simple=="Drive")
df_drive2 <- df_plot2 %>% filter(mode_simple=="Drive")

df_plot3 <- df_plot %>%
  filter(mode_simple %in% c("Transit","Walk","Bike"))
df_plot3$mode_simple <- factor(df_plot3$mode_simple, levels=c("Transit","Walk","Bike"))
df_plot4 <- df_plot2 %>%
  filter(mode_simple %in% c("Transit","Walk","Bike"))
df_plot4$mode_simple <- factor(df_plot4$mode_simple, levels=c("Transit","Walk","Bike"))

plot_ly(data = df_drive,
        x = ~d_rgcname, y = ~share, type = "bar", text = ~share, textposition = 'outside',texttemplate =  "%{y:.0%}",
        # error_y = ~list(array = share_moe,color = '#000000'),
        color = ~d_rgcname, colors = pal_rgc, width=400) %>%
  layout(xaxis = list(title = 'mode'), yaxis = list(title = 'percent', tickformat=".0%", range = list(0, 1)),
         title = "drive mode share for trips to centers",
         font = list(family="Poppins"))

plot_ly(data = df_plot3,
        x = ~mode_simple, y = ~share, type = "bar", text = ~share, textposition = 'outside',texttemplate =  "%{y:.0%}",
        # error_y = ~list(array = share_moe,color = '#000000'),
        color = ~d_rgcname, colors = pal_rgc, width=600) %>%
  layout(xaxis = list(title = 'mode'), yaxis = list(title = 'percent', tickformat=".0%", range = list(0, 0.32)),
         title = "trip mode share to centers",
         font = list(family="Poppins"))
```


```{r}
#| layout: [[1,3], [1,3]]
#| out-height: 300px
#| out-width: 800px

plot_ly(data = df_drive2,
        x = ~category, y = ~share, type = "bar", text = ~share, textposition = 'outside',texttemplate =  "%{y:.0%}",
        # error_y = ~list(array = share_moe,color = '#000000'),
        color = ~category, colors = pal_rgc_mu, width=400) %>%
  layout(xaxis = list(title = 'mode'), yaxis = list(title = 'percent', tickformat=".0%", range = list(0, 1)),
         title = "drive mode share for trips to centers",
         font = list(family="Poppins"))
plot_ly(data = df_plot4,
        x = ~mode_simple, y = ~share, type = "bar", text = ~share, textposition = 'outside',texttemplate =  "%{y:.0%}",
        # error_y = ~list(array = share_moe,color = '#000000'),
        color = ~category, colors = pal_rgc_mu, width=600) %>%
  layout(xaxis = list(title = 'mode'), yaxis = list(title = 'percent', tickformat=".0%", range = list(0, 0.32)),
         title = "trip mode share to centers",
         font = list(family="Poppins"))
```

### Work/Non-work trips

:::{.panel-tabset}
#### Drive mode share

- non-work trips to centers have higher drive mode share than work trips

```{r}
#| out-width: 500px
#| out-height: 300px

prepped_dt <- hts_prep_variable(summarize_var = 'mode_simple',
                                summarize_by = c('survey_year', "trip_type",'d_rgcname'),
                                data = hts_data,
                                id_cols='trip_id',
                                wt_cols='trip_weight',
                                weighted=TRUE)


summary_dt <- hts_summary(prepped_dt = prepped_dt$cat,
                            summarize_var = 'mode_simple',
                            summarize_by = c('survey_year', "trip_type",'d_rgcname'),
                            summarize_vartype = 'categorical',
                            id_cols = 'trip_id',
                            wtname ='trip_weight',
                            weighted=TRUE,
                            se=TRUE)

df_plot <- summary_dt$summary$wtd %>%
  filter(survey_year=="2023") %>%
  rename(share = prop,
         sample = count,
         count = est)
df_plot$d_rgcname <- factor(df_plot$d_rgcname, levels=c("RGC","Not RGC"))
df_plot$mode_simple <- factor(df_plot$mode_simple, levels=c("Drive","Transit","Walk","Bike","Other"))

df_drive <- df_plot %>% filter(mode_simple=="Drive")
df_drive$trip_type <- factor(df_drive$trip_type, levels=c("Work","Non-work"))

plot_ly(data = df_drive,
        x = ~trip_type, y = ~share, type = "bar", text = ~share, textposition = 'outside',texttemplate =  "%{y:.0%}",
        # error_y = ~list(array = share_moe,color = '#000000'),
        color = ~d_rgcname, colors = pal_rgc) %>%
  layout(xaxis = list(title = 'mode'), yaxis = list(title = 'percent', tickformat=".0%", range = list(0, 1)),
         title = "drive mode share for trips to centers",
         font = list(family="Poppins"))
```

```{r}
#| out-width: 500px
#| out-height: 300px

prepped_dt <- hts_prep_variable(summarize_var = 'mode_simple',
                                summarize_by = c('survey_year', "trip_type",'category'),
                                data = hts_data,
                                id_cols='trip_id',
                                wt_cols='trip_weight',
                                weighted=TRUE)


summary_dt <- hts_summary(prepped_dt = prepped_dt$cat,
                            summarize_var = 'mode_simple',
                            summarize_by = c('survey_year', "trip_type",'category'),
                            summarize_vartype = 'categorical',
                            id_cols = 'trip_id',
                            wtname ='trip_weight',
                            weighted=TRUE,
                            se=TRUE)

df_plot_mu <- summary_dt$summary$wtd %>%
  filter(survey_year=="2023") %>%
  rename(share = prop,
         sample = count,
         count = est)
df_plot_mu$category <- factor(df_plot_mu$category, levels=c("Metro", "Urban", "Not RGC"))

df_drive_mu <- df_plot_mu %>% filter(mode_simple=="Drive")
df_drive_mu$trip_type <- factor(df_drive_mu$trip_type, levels=c("Work","Non-work"))

plot_ly(data = df_drive_mu,
        x = ~trip_type, y = ~share, type = "bar", text = ~share, textposition = 'outside',texttemplate =  "%{y:.0%}",
        # error_y = ~list(array = share_moe,color = '#000000'),
        color = ~category, colors = pal_rgc_mu) %>%
  layout(xaxis = list(title = 'mode'), yaxis = list(title = 'percent', tickformat=".0%", range = list(0, 1)),
         title = "drive mode share for trips to centers",
         font = list(family="Poppins"))
```

#### Transit/ Walk/ Bike

- work trips to both centers and non-centers have higher transit mode share than non-work trips
- non-work trips to non-centers have higher walk mode share than work trips

```{r}
#| layout-ncol: 2
#| out-height: 300px
#| out-width: 800px

df_plot2 <- df_plot %>%
  filter(mode_simple %in% c("Transit","Walk","Bike"))
df_plot2$mode_simple <- factor(df_plot2$mode_simple, levels=c("Transit","Walk","Bike"))

plot_ly(data = df_plot2 %>% filter(trip_type == "Work"),
        x = ~mode_simple, y = ~share, type = "bar", text = ~share, textposition = 'outside',texttemplate =  "%{y:.0%}",
        # error_y = ~list(array = share_moe,color = '#000000'),
        color = ~d_rgcname, colors = pal_rgc, width=470) %>%
  layout(xaxis = list(title = 'mode'), yaxis = list(title = 'percent', tickformat=".0%", range = list(0, 0.21)),
         title = "work trip mode share to centers",
         font = list(family="Poppins"))

plot_ly(data = df_plot2 %>% filter(trip_type == "Non-work"),
        x = ~mode_simple, y = ~share, type = "bar", text = ~share, textposition = 'outside',texttemplate =  "%{y:.0%}",
        # error_y = ~list(array = share_moe,color = '#000000'),
        color = ~d_rgcname, colors = pal_rgc, width=470) %>%
  layout(xaxis = list(title = 'mode'), yaxis = list(title = 'percent', tickformat=".0%", range = list(0, 0.21)),
         title = "non-work trip mode share to centers",
         font = list(family="Poppins"))
```


- work trips to metro centers and non-centers have higher transit mode share than non-work trips
- walk mode share for non-work trips to metro centers are very high

```{r}
#| layout-ncol: 2
#| out-height: 300px
#| out-width: 400px

df_plot_mu2 <- df_plot_mu %>%
  filter(mode_simple %in% c("Transit","Walk","Bike"))
df_plot_mu2$mode_simple <- factor(df_plot_mu2$mode_simple, levels=c("Transit","Walk","Bike"))

plot_ly(data = df_plot_mu2 %>% filter(trip_type == "Work"),
        x = ~mode_simple, y = ~share, type = "bar", text = ~share, textposition = 'outside',texttemplate =  "%{y:.0%}",
        # error_y = ~list(array = share_moe,color = '#000000'),
        color = ~category, colors = pal_rgc_mu, width=470) %>%
  layout(xaxis = list(title = 'mode'), yaxis = list(title = 'percent', tickformat=".0%", range = list(0, 0.41)),
         title = "work trip mode share to centers",
         font = list(family="Poppins"))

plot_ly(data = df_plot_mu2 %>% filter(trip_type == "Non-work"),
        x = ~mode_simple, y = ~share, type = "bar", text = ~share, textposition = 'outside',texttemplate =  "%{y:.0%}",
        # error_y = ~list(array = share_moe,color = '#000000'),
        color = ~category, colors = pal_rgc_mu, width=470) %>%
  layout(xaxis = list(title = 'mode'), yaxis = list(title = 'percent', tickformat=".0%", range = list(0, 0.41)),
         title = "non-work trip mode share to centers",
         font = list(family="Poppins"))
```

:::