---
title: "Centers monitoring: 2017/2019 and 2023 mode share"
date: today
format: 
  html:
    mainfont: Poppins
    theme: [cosmo, custom.scss]
    toc: true
    df-print: kable
    warning: false
    echo: false
---

```{r}
source("region_dest_data_2023_new.R")

pal_rgc <- c(psrc_colors$purples_inc[2],psrc_colors$greens_inc[2])
pal_rgc <- setNames(pal_rgc, c("RGC","Not RGC"))

pal_rgc_mu <- c(psrc_colors$oranges_inc[2],psrc_colors$blues_inc[2],psrc_colors$greens_inc[2])
pal_rgc_mu <- setNames(pal_rgc_mu, c("Metro","Urban","Not RGC"))

hts_data = list(# hh = hh,
                # person = person,
                # day = day,
                trip = df_trip_data_23_uga)

```


## all trips

```{r}
# rgc
df_plot <- trip_data_uga_17_19 %>% 
  hhts_count(group_vars=c("d_rgcname"),spec_wgt = "trip_weight_2017_2019",incl_na=FALSE) %>%
  # filter(d_rgcname!= "Total") %>%
  mutate(survey_year = "2017/2019") %>%
  select(survey_year,d_rgcname,share,count)

df_plot_23 <- get_hts_summary(dt_list=hts_data, 
                              summary_var=c('survey_year', 'd_rgcname'), 
                              variables_dt = variable_list, 
                              id_var='trip_id', wt_var='trip_weight') %>%
  filter(survey_year=="2023") %>%
  rename(share = prop,
         sample = count,
         count = est) %>%
  select(survey_year,d_rgcname,share,count) 

df_plot_23 <- df_plot_23 %>%
  add_row(data.frame(survey_year="2023",
                     d_rgcname="Total",
                     count=sum(df_plot_23$count),
                     share=sum(df_plot_23$share)))

# metro/urban
df_plot_mu <- trip_data_uga_17_19 %>% 
  hhts_count(group_vars=c("category"),spec_wgt = "trip_weight_2017_2019",incl_na=FALSE) %>%
  filter(category!= "Total", category!="Not RGC") %>%
  mutate(survey_year = "2017/2019") %>%
  select(survey_year,category,share,count)

df_plot_mu_23 <- get_hts_summary(dt_list=hts_data, 
                              summary_var=c('survey_year', 'category'), 
                              id_var='trip_id', wt_var='trip_weight') %>%
  filter(survey_year=="2023", category!="Not RGC") %>%
  rename(share = prop,
         sample = count,
         count = est) %>%
  select(survey_year,category,share,count) 

df_plot_all <- df_plot %>% 
  add_row(df_plot_23)%>% 
  add_row(df_plot_mu %>% rename(d_rgcname=category))%>% 
  add_row(df_plot_mu_23 %>% rename(d_rgcname=category))
df_plot_all$d_rgcname <- factor(df_plot_all$d_rgcname,
                                levels=c("Total","RGC","Metro","Urban","Not RGC"))
```


```{r}
# formatting
generate_trip_count_table <- function(.data){
  .data %>%
    mutate(`trip count change` = `trip count 2023`-`trip count 2017/2019`,
           `trip count % change` = `trip count 2023`/`trip count 2017/2019`-1) %>%
    mutate(`trip count change` = scales::number(`trip count change`, accuracy=1000, 
                                                big.mark = ",",
                                                style_positive="plus"),
           `trip count 2023` = scales::number(`trip count 2023`, accuracy=1000, 
                                              big.mark = ","),
           `trip count 2017/2019` = scales::number(`trip count 2017/2019`, accuracy=1000, 
                                                   big.mark = ","),
           `trip count % change` = scales::percent(`trip count % change`, accuracy=1,
                                                   style_positive="plus"))
}

generate_mode_share_table <- function(.data){
  .data %>%
    mutate(`mode share % change` = `mode share 2023`/`mode share 2017/2019`-1) %>%
    mutate(`mode share 2023` = scales::percent(`mode share 2023`,accuracy=0.1),
           `mode share 2017/2019` = scales::percent(`mode share 2017/2019`,accuracy=0.1),
           `mode share % change` = scales::percent(`mode share % change`,accuracy=1, style_positive="plus"))
}
```


```{r}
df_table <- df_plot_all %>%
  arrange(d_rgcname) %>%
  pivot_wider(id_cols = c(d_rgcname), names_from = survey_year, values_from = count, names_prefix = "trip count ") %>%
  generate_trip_count_table()

kbl(df_table)%>%
  kable_styling(bootstrap_options = c("hover"), html_font = "Poppins") %>%
  row_spec(c(2), background = psrc_colors$purples_inc[1]) %>%
  row_spec(c(3), background = psrc_colors$oranges_inc[1]) %>%
  row_spec(c(4), background = psrc_colors$blues_inc[1]) %>%
  row_spec(c(5), background = psrc_colors$greens_inc[1])
```

## mode share for trips to centers

- Showing mode share for trips destined to centers to understand how do people get to centers 

:::{.panel-tabset}

### RGC
```{r}

df_plot <- trip_data_uga_17_19 %>% 
  hhts_count(group_vars=c("d_rgcname","mode_simple"),spec_wgt = "trip_weight_2017_2019",incl_na=FALSE) %>%
  filter(mode_simple!= "Total") %>%
  mutate(survey_year = "2017/2019") %>%
  select(survey_year,d_rgcname,mode_simple,share,count)

df_plot_23 <- get_hts_summary(dt_list=hts_data, 
                              summary_var=c('survey_year', 'd_rgcname','mode_simple'), 
                              id_var='trip_id', wt_var='trip_weight') %>%
  filter(survey_year=="2023") %>%
  rename(share = prop,
         sample = count,
         count = est) %>%
  select(survey_year,d_rgcname,mode_simple,share,count)

df_plot_all <- df_plot %>% add_row(df_plot_23)
df_plot_all$d_rgcname <- factor(df_plot_all$d_rgcname, levels=c("RGC","Not RGC"))
df_plot_all$mode_simple <- factor(df_plot_all$mode_simple, levels=c("Drive","Transit","Walk","Bike","Other"))

df_table <- df_plot_all %>%
  arrange(d_rgcname,mode_simple) %>%
  pivot_wider(id_cols = c(d_rgcname, mode_simple), names_from = survey_year, values_from = count, names_prefix = "trip count ") %>%
  generate_trip_count_table()

df_table2 <- df_plot_all %>%
  arrange(d_rgcname,mode_simple) %>%
  pivot_wider(id_cols = c(d_rgcname, mode_simple), names_from = survey_year, values_from = share, names_prefix = "mode share ") %>%
  generate_mode_share_table()


kbl(df_table2 %>% 
      add_column(df_table[,-c(1,2)]))%>%
  kable_styling(bootstrap_options = c("hover"), html_font = "Poppins") %>%
  row_spec(c(1:5), background = psrc_colors$purples_inc[1]) %>%
  row_spec(c(6:10), background = psrc_colors$greens_inc[1])

```

### Metro/Urban

- drive mode share increased for all centers/non-centers
  - however, drive trips to metro centers dropped
- transit mode share dropped by half for metro centers
```{r}

df_plot <- trip_data_uga_17_19 %>% 
  hhts_count(group_vars=c("category","mode_simple"),spec_wgt = "trip_weight_2017_2019",incl_na=FALSE) %>%
  filter(mode_simple!= "Total") %>%
  mutate(survey_year = "2017/2019") %>%
  select(survey_year,category,mode_simple,share,count)

df_plot_23 <- get_hts_summary(dt_list=hts_data, 
                              summary_var=c('survey_year', 'category','mode_simple'), 
                              id_var='trip_id', wt_var='trip_weight') %>%
  filter(survey_year=="2023") %>%
  rename(share = prop,
         sample = count,
         count = est) %>%
  select(survey_year,category,mode_simple,share,count)

df_plot_all <- df_plot %>% add_row(df_plot_23)
df_plot_all$category <- factor(df_plot_all$category, levels=c("Metro", "Urban", "Not RGC"))
df_plot_all$mode_simple <- factor(df_plot_all$mode_simple, levels=c("Drive","Transit","Walk","Bike","Other"))

df_table <- df_plot_all %>%
  arrange(category,mode_simple) %>%
  pivot_wider(id_cols = c(category, mode_simple), names_from = survey_year, values_from = count, names_prefix = "trip count ") %>%
  generate_trip_count_table()


df_table2 <- df_plot_all %>%
  arrange(category,mode_simple) %>%
  pivot_wider(id_cols = c(category, mode_simple), names_from = survey_year, values_from = share, names_prefix = "mode share ") %>%
  generate_mode_share_table()


kbl(df_table2 %>% 
      add_column(df_table[,-c(1,2)]))%>%
  kable_styling(bootstrap_options = c("hover"), html_font = "Poppins") %>%
  row_spec(c(1:5), background = psrc_colors$oranges_inc[1]) %>%
  row_spec(c(6:10), background = psrc_colors$blues_inc[1]) %>%
  row_spec(c(11:15), background = psrc_colors$greens_inc[1])

```

:::

## mode share to centers by work/non-work trips

```{r}
# df_plot <- trip_data_uga_17_19 %>% 
#   hhts_count(group_vars=c("trip_type","d_rgcname"),spec_wgt = "trip_weight_2017_2019",incl_na=FALSE) %>%
#   mutate(survey_year = "2017/2019") %>%
#   select(survey_year,trip_type,d_rgcname,share,count)
# 
# df_plot_23 <- get_hts_summary(dt_list=hts_data, 
#                               summary_var=c('survey_year', "trip_type",'d_rgcname'), 
#                               id_var='trip_id', wt_var='trip_weight') %>%
#   filter(survey_year=="2023") %>%
#   rename(share = prop,
#          sample = count,
#          count = est) %>%
#   select(survey_year,trip_type,d_rgcname,share,count)
# 
# df_plot_23_total <- df_plot_23 %>% group_by(survey_year,trip_type) %>% summarise(share = sum(share),count = sum(count)) %>%
#   ungroup() %>%
#   mutate(d_rgcname = "Total", .after="trip_type")
# 
# df_plot_mu <- trip_data_uga_17_19 %>% 
#   hhts_count(group_vars=c("trip_type","category"),spec_wgt = "trip_weight_2017_2019",incl_na=FALSE) %>%
#   mutate(survey_year = "2017/2019") %>%
#   select(survey_year,trip_type,category,share,count) %>%
#   rename(d_rgcname=category) %>%
#   filter(d_rgcname %in% c("Metro", "Urban"))
# 
# df_plot_mu_23 <- get_hts_summary(dt_list=hts_data, 
#                               summary_var=c('survey_year', "trip_type",'category'), 
#                               id_var='trip_id', wt_var='trip_weight') %>%
#   filter(survey_year=="2023") %>%
#   rename(share = prop,
#          sample = count,
#          count = est,
#          d_rgcname=category) %>%
#   select(survey_year,trip_type,d_rgcname,share,count) %>%
#   filter(d_rgcname %in% c("Metro", "Urban"))
# 
# df_plot_all <- df_plot %>% add_row(df_plot_23) %>% add_row(df_plot_23_total) %>% add_row(df_plot_mu) %>% add_row(df_plot_mu_23)
# df_plot_all$d_rgcname <- factor(df_plot_all$d_rgcname, levels=c("Total","RGC","Metro", "Urban","Not RGC"))
# 
# df_table <- df_plot_all %>%
#   arrange(d_rgcname,trip_type) %>%
#   pivot_wider(id_cols = c(trip_type,d_rgcname), names_from = survey_year, values_from = count, names_prefix = "trip count ") %>%
#   filter(trip_type!="Total") %>%
#   generate_trip_count_table()
# 
# kbl(df_table)%>%
#   kable_styling(bootstrap_options = c("hover"), html_font = "Poppins") %>%
#   row_spec(c(3,4), background = psrc_colors$purples_inc[1]) %>%
#   row_spec(c(5,6), background = psrc_colors$oranges_inc[1]) %>%
#   row_spec(c(7,8), background = psrc_colors$blues_inc[1]) %>%
#   row_spec(c(9,10), background = psrc_colors$greens_inc[1])
```

:::{.panel-tabset}
### RGC

```{r}
df_plot <- trip_data_uga_17_19 %>% 
  hhts_count(group_vars=c("trip_type","d_rgcname","mode_simple"),spec_wgt = "trip_weight_2017_2019",incl_na=FALSE) %>%
  filter(mode_simple!= "Total") %>%
  mutate(survey_year = "2017/2019") %>%
  select(survey_year,trip_type,d_rgcname,mode_simple,share,count)

df_plot_23 <- get_hts_summary(dt_list=hts_data, 
                              summary_var=c('survey_year', "trip_type",'d_rgcname','mode_simple'), 
                              id_var='trip_id', wt_var='trip_weight') %>%
  filter(survey_year=="2023") %>%
  rename(share = prop,
         sample = count,
         count = est) %>%
  select(survey_year,trip_type,d_rgcname,mode_simple,share,count)

df_plot_all <- df_plot %>% add_row(df_plot_23)
df_plot_all$d_rgcname <- factor(df_plot_all$d_rgcname, levels=c("RGC","Not RGC"))
df_plot_all$mode_simple <- factor(df_plot_all$mode_simple, levels=c("Drive","Transit","Walk","Bike","Other"))

```

- mode share for work trips

```{r}

df_table <- df_plot_all %>% filter(trip_type == "Work") %>%
  arrange(d_rgcname,mode_simple) %>%
  pivot_wider(id_cols = c(trip_type,d_rgcname, mode_simple), names_from = survey_year, values_from = count, names_prefix = "trip count ") %>%
  generate_trip_count_table()

df_table2 <- df_plot_all %>% filter(trip_type == "Work") %>%
  arrange(d_rgcname,mode_simple) %>%
  pivot_wider(id_cols = c(trip_type,d_rgcname, mode_simple), names_from = survey_year, values_from = share, names_prefix = "mode share ") %>%
  generate_mode_share_table()

kbl(df_table2 %>% 
      add_column(df_table[,-c(1,2,3)]))%>%
  kable_styling(bootstrap_options = c("hover"), html_font = "Poppins") %>%
  row_spec(c(1:5), background = psrc_colors$purples_inc[1]) %>%
  row_spec(c(6:10), background = psrc_colors$greens_inc[1])
```


- mode share for non-work trips

```{r}

df_table <- df_plot_all %>% filter(trip_type == "Non-work") %>%
  arrange(d_rgcname,mode_simple) %>%
  pivot_wider(id_cols = c(trip_type,d_rgcname, mode_simple), names_from = survey_year, values_from = count, names_prefix = "trip count ") %>%
  generate_trip_count_table()

df_table2 <- df_plot_all %>% filter(trip_type == "Non-work") %>%
  arrange(d_rgcname,mode_simple) %>%
  pivot_wider(id_cols = c(trip_type,d_rgcname, mode_simple), names_from = survey_year, values_from = share, names_prefix = "mode share ") %>%
  generate_mode_share_table()

kbl(df_table2 %>% 
      add_column(df_table[,-c(1,2,3)]))%>%
  kable_styling(bootstrap_options = c("hover"), html_font = "Poppins") %>%
  row_spec(c(1:5), background = psrc_colors$purples_inc[1]) %>%
  row_spec(c(6:10), background = psrc_colors$greens_inc[1])
```

### Metro/Urban

```{r}
df_plot <- trip_data_uga_17_19 %>% 
  hhts_count(group_vars=c("trip_type","category","mode_simple"),spec_wgt = "trip_weight_2017_2019",incl_na=FALSE) %>%
  filter(mode_simple!= "Total") %>%
  mutate(survey_year = "2017/2019") %>%
  select(survey_year,trip_type,category,mode_simple,share,count)

df_plot_23 <- get_hts_summary(dt_list=hts_data, 
                              summary_var=c('survey_year', "trip_type",'category','mode_simple'), 
                              id_var='trip_id', wt_var='trip_weight') %>%
  filter(survey_year=="2023") %>%
  rename(share = prop,
         sample = count,
         count = est) %>%
  select(survey_year,trip_type,category,mode_simple,share,count)

df_plot_all <- df_plot %>% add_row(df_plot_23)
df_plot_all$category <- factor(df_plot_all$category, levels=c("Metro","Urban","Not RGC"))
df_plot_all$mode_simple <- factor(df_plot_all$mode_simple, levels=c("Drive","Transit","Walk","Bike","Other"))

```

- mode share for work trips

```{r}

df_table <- df_plot_all %>% filter(trip_type == "Work") %>%
  arrange(category,mode_simple) %>%
  pivot_wider(id_cols = c(trip_type,category, mode_simple), names_from = survey_year, values_from = count, names_prefix = "trip count ") %>%
  generate_trip_count_table()

df_table2 <- df_plot_all %>% filter(trip_type == "Work") %>%
  arrange(category,mode_simple) %>%
  pivot_wider(id_cols = c(trip_type,category, mode_simple), names_from = survey_year, values_from = share, names_prefix = "mode share ") %>%
  generate_mode_share_table()

kbl(df_table2 %>% 
      add_column(df_table[,-c(1,2,3)]))%>%
  kable_styling(bootstrap_options = c("hover"), html_font = "Poppins") %>%
  row_spec(c(1:5), background = psrc_colors$oranges_inc[1]) %>%
  row_spec(c(6:10), background = psrc_colors$blues_inc[1])%>%
  row_spec(c(11:15), background = psrc_colors$greens_inc[1])
```


- mode share for non-work trips

```{r}

df_table <- df_plot_all %>% filter(trip_type == "Non-work") %>%
  arrange(category,mode_simple) %>%
  pivot_wider(id_cols = c(trip_type,category, mode_simple), names_from = survey_year, values_from = count, names_prefix = "trip count ") %>%
  generate_trip_count_table()

df_table2 <- df_plot_all %>% filter(trip_type == "Non-work") %>%
  arrange(category,mode_simple) %>%
  pivot_wider(id_cols = c(trip_type,category, mode_simple), names_from = survey_year, values_from = share, names_prefix = "mode share ") %>%
  generate_mode_share_table()

kbl(df_table2 %>% 
      add_column(df_table[,-c(1,2,3)]))%>%
  kable_styling(bootstrap_options = c("hover"), html_font = "Poppins") %>%
  row_spec(c(1:5), background = psrc_colors$oranges_inc[1]) %>%
  row_spec(c(6:10), background = psrc_colors$blues_inc[1])%>%
  row_spec(c(11:15), background = psrc_colors$greens_inc[1])

```

:::