---
title: "Can Drive"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
    fig-width: 4.5
    fig-asp: .618
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  source(here::here("Analysis", "care-giving", f))
})
)

```

## Can Drive by Age

-   There were fewer 18-24 year olds who could drive in 2017/19 than those who couldn't. By 2023, there were significantly less who were driving.

-   There were more 25-34 year olds who could drive compared to those who couldn't in 2021, but it's hard to determine in other years.

-   Between 35-64 years in 2017/19 and 2025, there are more who could drive than couldn't. In 2023, there are more people who can drive than those who cannot in the 35-54 age range.

-   Those who are between 16-24 years of age have a higher share who cannot drive. The earliest a person can obtain a learner's permit is 15 in addition to people who delay learning how to drive.

```{r}
per_cda <- psrc_hts_stat(df_hts,
                    analysis_unit = "person",
                    group_vars = c("can_drive2", "age"),
                    incl_na = FALSE)

```

```{r}
#| out-width: "100%"
#| fig-align: "center"

per_cda |> 
  ggplot(aes(x = age, y = prop, fill = as.character(can_drive2))) +
  geom_col(position = "dodge") + 
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)
  ) +
  facet_wrap(~survey_year, ncol = 4, nrow = 1, strip.position = "top") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.7)
  ) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Persons by Age who Can or Cannot Drive")
```

```{r}
per_cda
```

## Care Trips by Can Drive and Age

-   Where care trips were taken, the vast majority of those between the ages of 25-84 years can drive.

-   Where there are more care trips by those who cannot drive can be found by those who are either too young to obtain a learners permit or those 85 years or older, depending on the survey year.

```{r}
t_cda <- create_candrive_tbl(geog = "Region")
```

```{r}
#| out-width: "100%"
#| fig-align: "center"

t_cda$long |> 
  ggplot(aes(x = age, y = prop, fill = as.character(can_drive2))) +
  geom_col(position = "dodge") + 
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)
  ) +
  facet_wrap(~survey_year, ncol = 4, nrow = 1, strip.position = "top") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.7)
  ) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Care Trips by Age")
```

```{r}
t_cda$long |> 
  select(-can_drive, -dest_loc) |> 
  relocate(can_drive2, .before = "age")

```

## Care Trips by Can Drive and Gender

```{r}
# Care trips by Gender

t_g <- create_candrive_tbl_gender(geog = "Region")

```

-   Of those who can drive, there are more care trips made by females than males.

```{r}
#| layout-ncol: 2
#| out-width: "50%"

t_g$long |> 
  ggplot(aes(x = gender2, y = prop, fill = can_drive2)) +
  geom_col(position = "dodge") + 
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)
  ) +
  facet_wrap(~survey_year, ncol = 4, nrow = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.7)) +
  labs(x = NULL,
       y = NULL,
       title = "Share of Care Trips by Can Drive and Gender")

t_g$long |> 
  ggplot(aes(x = as.character(survey_year), y = prop, fill = gender2)) +
  geom_col(position = "dodge") + 
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)
  ) +
  facet_wrap(~can_drive2, ncol = 2, nrow = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$blues_inc) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.7)) +
  labs(x = NULL,
       y = NULL,
       title = "Share of Care Trips by Can Drive and Gender")

```

```{r}
t_g$long
```
