---
title: "Trip Purpose"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  source(f)
  # source(here::here("Analysis", "care-giving", f))
})
)

```

## Trip Purpose

### PSRC Destination Purpose Assignment

Below are trips with their destination in the Region (King, Kitsap, Pierce, and Snohomish Counties) for all survey years. Care-giving is an aggregate of the following destination purposes (escort, personal business/errand/appointment, meal, and shopping). The care-giving analysis will also exclude HOME trips.

```{r}
#| label: tbl-trip-purp-all
#| 
# This section prints a table with weighted/unweighted x Home/Exclude Home x trip purpose

trip_purp_a <- create_trip_purpose_tbl(df_hts)

trip_purp_a_fmt <- trip_purp_a |> 
  pivot_longer(cols = contains("HOME")) |> 
  separate_wider_delim(cols = "name", names = c("home", "year"), delim = ".") |> 
  pivot_wider(id_cols = c("dest_purpose_cat", "home"), names_from = "year", values_from = "value")

knitr::kable(trip_purp_a_fmt)

```

### Care Purpose

Care is defined as the following purposes: Escort, Shopping, Meal, Personal Business/Errand/Appointment. All `HOME` trips are excluded. The remaining categories from the `dest_purpose_cat` either stand alone or have been aggregated to the following:

**Care**: `r care_purpose_cat`

**Work**: `r work_cat`

**Other**: `r other_cat`

**School**: `r school_cat`

Care related trips make up the majority of trips compared to other purpose categories across survey years.

```{r}
care_purp <- create_care_purpose_tbl(df_hts)
```

```{r}
#| label: fig-care-purp-cat-01
#| fig-height: 3
#| fig-width: 5.1
#| out-width: "350px"

plot_facet_wrap(care_purp, 
                survey_year, 
                care_purpose_cat, 
                care_purpose_cat, 
                "Care Purpose Categories by Year", 
                color_pal = psrc_colors$pognbgy_5) +
  theme(axis.text.x = element_blank())
```

```{r}
#| label: fig-care-purp-cat-02
#| fig-height: 3
#| fig-width: 5.1
#| out-width: "350px"
plot_facet_wrap(care_purp, 
                care_purpose_cat, 
                as.character(survey_year), 
                as.character(survey_year), 
                "Care Purpose Categories by Year", 
                color_pal = psrc_colors$blues_inc) +
  theme(axis.text.x = element_blank())
```

```{r}
#| label: tbl-care-purp-cat
#| 
knitr::kable(care_purp)
```
