---
title: "Mode and Income"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
    fig-width: 4.5
    fig-asp: .618
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  source(f)
  # source(here::here("Analysis", "care-giving", f))
})
)

```

## Analysis A

### Under \$50k and \$50k or More

Driving is the predominant mode of transportation for care trips. Those who earn under \$50,000 are using transit more than those who earn over \$50,000 across survey years.

```{r}
msi50bin_df <- mode_share_income_set_c(geog = "Region")$long |> 
  mutate(income_50 = factor(income_50, levels = c("Under $50,000", "$50,000 or more")))
```

::: panel-tabset
## Side-by-side

```{r}
#| out-width: "90%"
#| fig-align: "center"

msi50bin_df |> 
  ggplot() +
  geom_col(aes(x = as.character(survey_year), y = prop, fill = mode_class_5)) +
  facet_wrap(~income_50, strip.position = "top") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  psrc_style() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        # strip.placement = "outside",
        # panel.spacing = unit(0, "cm"),
        legend.position = "bottom",
        strip.text = element_text(hjust = 0.5)
        # strip.text = element_text(face = "bold")
        ) +
  labs(x = NULL,
       y = NULL)

```

## Under \$50,000

```{r}
#| layout-ncol: 2
#| out-width: "50%"
#| 
msi50bin_df_u <- msi50bin_df |>
  filter(income_50 == "Under $50,000")

plot_facet_wrap(table = msi50bin_df_u, 
                facet = as.character(survey_year), 
                var1 = mode_class_5, 
                var2  = mode_class_5, 
                title = "Income by Primary Trip Mode", 
                color_pal = psrc_colors$pognbgy_5)+
  theme(axis.text.x = element_blank())

plot_facet_wrap(table = msi50bin_df_u, 
                var1 = as.character(survey_year), 
                var2 = as.character(survey_year),
                facet = mode_class_5, 
                title = "Income by Primary Trip Mode", 
                color_pal = psrc_colors$blues_inc)+
  theme(axis.text.x = element_blank())
```

## \$50,000 or More

```{r}
#| layout-ncol: 2
#| out-width: "50%"
#| 
msi50bin_df_m <- msi50bin_df |>
  filter(income_50 == "$50,000 or more")

plot_facet_wrap(table = msi50bin_df_m, 
                facet = as.character(survey_year), 
                var1 = mode_class_5, 
                var2  = mode_class_5, 
                title = "Income by Primary Trip Mode", 
                color_pal = psrc_colors$pognbgy_5) +
  theme(axis.text.x = element_blank())

plot_facet_wrap(table = msi50bin_df_m, 
                var1 = as.character(survey_year), 
                var2 = as.character(survey_year),
                facet = mode_class_5, 
                title = "Income by Primary Trip Mode", 
                color_pal = psrc_colors$blues_inc) +
  theme(axis.text.x = element_blank())
```
:::

```{r}

mode_share_income_set_c(geog = "Region")$wide|>
  gt() |>
  tab_spanner_delim(delim = ".") |>
  cols_hide(c(dest_loc, care_purpose_cat)) |>
  cols_align(align = "left",
             columns = c(3)
)
```

### Under \$50k, \$50k-\$100k, \$100k or more

```{r}
msi_df <- mode_share_income_set_b(geog = "Region")$long  |> 
  mutate(income_comp = factor(income_comp, levels = c("Under $50,000", "$50,000-$99,999", "$100,000 or more")))

```

```{r}
#| out-width: "100%"
#| fig-align: "center"

msi_df |> 
  ggplot() +
  geom_col(aes(x = as.character(survey_year), y = prop, fill = mode_class_5)) +
  facet_wrap(~income_comp, strip.position = "bottom") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        strip.placement = "outside",
        panel.spacing = unit(0, "cm"),
        legend.position = "bottom",
        strip.text = element_text(face = "bold")) +
  labs(x = NULL,
       y = NULL)

```

```{r}

mode_share_income_set_b(geog = "Region")$wide|>
  gt() |>
  tab_spanner_delim(delim = ".") |>
  cols_hide(c(dest_loc, care_purpose_cat)) |>
  cols_align(align = "left",
             columns = c(3)
)
```

## Analysis B

### Under \$50k and \$50k or More

```{r}
mode_inc_b <- psrc_hts_stat(df_hts,
                            analysis_unit = "trip",
                            group_vars = c("dest_region", "income_50", "mode_class_5", "care_purpose_cat"),
                            incl_na = FALSE) |>
  mutate(prop_per = label_percent(accuracy = 0.1)(prop),
         moe_per = label_percent(accuracy = 0.1)(prop_moe))
```

::: panel-tabset
### Drive

```{r}
mode_inc_b_drive <- mode_inc_b |>
  filter(mode_class_5 == "Drive")
```

Of trips made by those who make under \$50,000 and driving was the primary trip mode, at least 50% were care-giving compared to other trip purposes. When compared to certain survey years, more care-giving drive trips were made than those who earn more than \$50,000.

```{r}
#| fig-align: "center"
#| out-width: "90%"

plot_facet_wrap(table = mode_inc_b_drive,
                facet = income_50,
                var1 = care_purpose_cat,
                var2  = as.character(survey_year),
                title = "Income by Primary Trip Mode (Drive)",
                color_pal = psrc_colors$pognbgy_5) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.7, size = 7))
```

### Transit

```{r}

mode_inc_b_trans <- mode_inc_b |>
  filter(mode_class_5 == "Transit")
```

```{r}
#| fig-align: "center"
#| out-width: "90%"

plot_facet_wrap(table = mode_inc_b_trans,
                facet = income_50,
                var1 = care_purpose_cat,
                var2  = as.character(survey_year),
                title = "Income by Primary Trip Mode (Transit)",
                color_pal = psrc_colors$pognbgy_5) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.7, size = 7))
```

### Walk

```{r}

mode_inc_b_walk <- mode_inc_b |>
  filter(mode_class_5 == "Walk")
```

Of trips made by those who earn $50,000 or more and where walking is the primary trip mode, it is not clear whether there are more care-giving trips or social/recreational trips aside from 2021 where there is a high share of walking trips for social/recreational purposes.

```{r}
#| fig-align: "center"
#| out-width: "90%"

plot_facet_wrap(table = mode_inc_b_walk,
                facet = income_50,
                var1 = care_purpose_cat,
                var2  = as.character(survey_year),
                title = "Income by Primary Trip Mode (Walk)",
                color_pal = psrc_colors$pognbgy_5) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.7, size = 7))
```

### Bike/Micromobility

```{r}

mode_inc_b_bike <- mode_inc_b |>
  filter(mode_class_5 == "Bike/Micromobility")
```

```{r}
#| fig-align: "center"
#| out-width: "90%"

plot_facet_wrap(table = mode_inc_b_bike,
                facet = income_50,
                var1 = care_purpose_cat,
                var2  = as.character(survey_year),
                title = "Income by Primary Trip Mode (Bike)",
                color_pal = psrc_colors$pognbgy_5) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.7, size = 7))
```

### Other

```{r}

mode_inc_b_other <- mode_inc_b |>
  filter(mode_class_5 == "Other")
```

```{r}
#| fig-align: "center"
#| out-width: "90%"

plot_facet_wrap(table = mode_inc_b_other,
                facet = income_50,
                var1 = care_purpose_cat,
                var2  = as.character(survey_year),
                title = "Income by Primary Trip Mode (Other)",
                color_pal = psrc_colors$pognbgy_5) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.7, size = 7))
```
:::
