---
title: "Can Drive"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
params:
  errorbar: FALSE    
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}
geog <- "Region"
geo_col <- case_when(geog == "Region" ~ "dest_region",
                     geog %in% c("King County", "Kitsap County", "Pierce County", "Snohomish County") ~ "dest_county")
```

```{r}

files <- c("data.R", "functions.R")

invisible(
  lapply(files, function(f) {
    source(f)
    # source(here::here("Analysis", "care-giving", f))
  })
)
source("../util.R")

```

## Can Drive by Age

- There were fewer 18-24 year olds who could drive in 2017/19 than those who couldn't. By 2023, there were significantly less who were driving.

- There were more 25-34 year olds who could drive compared to those who couldn't in 2021, but it's hard to determine in other years.

- Between 35-64 years in 2017/19 and 2025, there are more who could drive than couldn't. In 2023, there are more people who can drive than those who cannot in the 35-54 age range.

- Those who are between 16-24 years of age have a higher share who cannot drive. The earliest a person can obtain a learner's permit is 15 in addition to people who delay learning how to drive.

```{r}
per_cda <- psrc_hts_stat(df_hts,
                         analysis_unit = "person",
                         group_vars = c("can_drive2", "age"),
                         incl_na = FALSE)
```

```{r}
#| label: fig-person-can-drive-age
#| fig-cap: "Persons by Age who Can or Cannot Drive"
#| fig-width: 8
#| out-width: "80%"

per_cda |> 
  ggplot(aes(x = age, y = prop, fill = as.character(can_drive2))) +
  geom_col(position = "dodge", width=0.7) + 
  geom_text(aes(label=scales::percent(prop, accuracy = 1)),
            vjust = -0.3, 
            position = position_dodge(width = 0.9),
            check_overlap = TRUE
            # size = 8/.pt
  ) +
  errorbar_switch(geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe), width = 0.1), 
                  on=params$errorbar) +
  # geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
  #                orientation = "x",
  #                position = position_dodge(width = 0.9)
  # ) +
  facet_wrap(~survey_year, ncol = 5, nrow = 1, strip.position = "top") +
  # scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.7),
        axis.text.y = element_blank(),
        # Shrink the legend text and title
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 9),
        
        # Shrink the legend key boxes (both width and height)
        legend.key.size = unit(0.4, "cm") 
  ) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = NULL#, #"Persons by Age who Can or Cannot Drive"
  )
```

```{r}
#| label: tbl-person-can-drive-age
get_table_format(per_cda)
```

## Care Trips, Can Drive, Age

### Can Drive and Age by Trip Purpose

- Where care trips were taken, the vast majority of those between the ages of 25-84 years can drive.

- Where there are more care trips by those who cannot drive can be found by those who are either too young to obtain a learners permit or those 85 years or older, depending on the survey year.

```{r}
age_groups <- c("12-15 years", "16-17 years", "18-24 years", "25-34 years", "35-44 years", "45-54 years",
                "55-64 years", "65-74 years", "75-84 years", "85 years or older")

t_cda <- psrc_hts_stat(df_hts,
                       analysis_unit = "trip",
                       group_vars = c(geo_col, "care_purpose_cat", "age", "can_drive2"),
                       incl_na = FALSE) |>
  filter(.data[[geo_col]] == geog,
         care_purpose_cat == "Care") |>
  mutate(age = factor(age, levels = age_groups)) |>
  arrange(age)

```

<!-- ::: panel-tabset -->
#### Yes (can drive)

```{r}
#| label: fig-yes-can-drive-care-trips
#| fig-cap: "Age, Yes, can drive and Care Trips"
#| fig-width: 8
#| out-width: "70%"

plot_facet_wrap(table = t_cda |> filter(can_drive2 %in% "Yes"),
                facet = survey_year,
                var1 = age,
                var2 = age,
                title = NULL, #"Age, Yes, can drive and Care Trips",
                color_pal = psrc_colors$pognbgy_10)

```

#### No (can drive)

```{r}
#| label: fig-no-can-drive-care-trips
#| fig-cap: "Age, Yes, can drive and Care Trips"
#| fig-width: 8
#| out-width: "70%"

plot_facet_wrap(table = t_cda |> filter(can_drive2 %in% "No"),
                facet = survey_year,
                var1 = age,
                var2 = age,
                title = NULL, #"Age, No, cannot drive and Care Trips",
                color_pal = psrc_colors$pognbgy_10)
```
<!-- ::: -->

```{r}
#| label: tbl-can-drive-care-trips
get_table_format(t_cda)

```

<!-- ### Analysis B -->

```{r, eval=FALSE}

# Of those that can drive between ages X, X% are care-giving trips.

# This might be a difficult analysis because of all the different age groups and lower sample sizes

cda_t <- psrc_hts_stat(df_hts,
                       analysis_unit = "trip",
                       group_vars = c(geo_col, "can_drive2", "age", "care_purpose_cat"),
                       incl_na = FALSE) |>
  filter(.data[[geo_col]] == geog) |>
  mutate(age = factor(age, levels = age_groups)) |>
  arrange(age)

cda_t
```

## Care Trips, Can Drive, & Gender

All charts in this section focus on care trip purposes. The analysis is presented in two different ways. First, the charts are broken down by respondants’ ability to drive (Yes, No, NA). In the next section they are broken down by gender (Men, Women, Prefer not to answer, Other).

### Gender by Can Drive and Trip Purpose

The following charts and tables show the distribution of gender who can or cannot drive by trip purpose--specifically for care-giving.

- Of all care-trips in the Region and those who can drive, there is a higher share of Women than Men for most survey years

- Charts where respondents did not select an anwwer for Can Drive, those individuals are also ages 15 years and under.

```{r}
# Can Drive and Gender by Trip Purpose

t_cd_g <- psrc_hts_stat(df_hts,
                        analysis_unit = "trip",
                        group_vars = c(geo_col, "care_purpose_cat", "can_drive2", "gender2"),
                        incl_na = TRUE) 
```

```{r}
t_cd_g_f <- t_cd_g |>
  filter(.data[[geo_col]] == geog, 
         care_purpose_cat %in% "Care")
```

```{r, eval=FALSE}
# sumcheck
t_cd_g_f |> 
  group_by(survey_year, can_drive2) |> 
  summarise(sum_prop = sum(prop))
```

<!-- ::: panel-tabset -->

#### Yes (can drive)

```{r}
#| label: fig-yes-can-drive-trip-purp
#| fig-cap: "Care Trips, Yes-can drive, and Gender"
#| fig-width: 8
#| out-width: "70%"

plot_facet_wrap(table = t_cd_g_f |> filter(can_drive2 %in% "Yes") |> mutate(fifty_line = "50%"),
                facet = survey_year,
                var1 = gender2,
                var2 = gender2,
                title = NULL, #"Care Trips, Yes-can drive, and Gender",
                color_pal = psrc_colors$pognbgy_10) +
  theme(axis.text.x = element_blank())
```

#### No (can drive)

```{r}
#| label: fig-no-can-drive-trip-purp
#| fig-cap: "Care Trips, No-cannot drive, and Gender"
#| fig-width: 8
#| out-width: "70%"

plot_facet_wrap(table = t_cd_g_f |> filter(can_drive2 %in% "No"),
                facet = survey_year,
                var1 = gender2,
                var2 = gender2,
                title = NULL, #"Care Trips, No-cannot drive, and Gender",
                color_pal = psrc_colors$pognbgy_10) +
  theme(axis.text.x = element_blank())
```

#### NA (can drive)

NA represents persons who are ages 15 years and under.

```{r}
#| label: fig-na-can-drive-trip-purp
#| fig-cap: "Care Trips, NA Can Drive, and Gender"
#| fig-width: 8
#| out-width: "70%"

plot_facet_wrap(table = t_cd_g_f |> filter(is.na(can_drive2)),
                facet = survey_year,
                var1 = gender2,
                var2 = gender2,
                title = NULL, #"Care Trips, NA Can Drive, and Gender",
                color_pal = psrc_colors$pognbgy_10) +
  theme(axis.text.x = element_blank())
```

<!-- ::: -->

```{r}
#| label: tbl-gender-can-drive-trip-purp
#| 
get_table_format(t_cd_g_f)
```

### Trip Purpose by Can Drive and Gender

The following charts and tables show the distribution of trip purpose by those who can or cannot drive and gender.

```{r}
cd_g_t <- psrc_hts_stat(df_hts,
                        analysis_unit = "trip",
                        group_vars = c(geo_col, "gender2", "can_drive2", "care_purpose_cat"),
                        incl_na = FALSE) 
```

- Of all Men or Women who can drive, the predominant trip purpose is care-giving trips. For 4 out of the 5 survey years, over 50% of trips made by Women who can drive are care-related trips.

- For Men and Women who cannot drive, trips are primarily care-giving trips for most survey years.

#### Men

```{r}
#| label: fig-trip-purp-can-drive-men
#| fig-cap: "Men, Can Drive, and Trip Purpose"
#| fig-width: 8
#| out-width: "70%"

plot_facet_wrap(table = cd_g_t |> filter(gender2 %in% "Men"),
                facet = survey_year,
                var1 = can_drive2,
                var2 = care_purpose_cat,
                title = NULL, #"Men, Can Drive, and Trip Purpose",
                color_pal = psrc_colors$pognbgy_10)
```

```{r}
#| label: tbl-trip-purp-can-drive-men
#| 
get_table_format(cd_g_t |> filter(gender2 %in% "Men"))
```

#### Women

```{r}
#| label: fig-trip-purp-can-drive-women
#| fig-cap: "Women, Can Drive, and Trip Purpose"
#| fig-width: 8
#| out-width: "70%"

plot_facet_wrap(table = cd_g_t |> filter(gender2 %in% "Women"),
                facet = survey_year,
                var1 = can_drive2,
                var2 = care_purpose_cat,
                title = NULL, #"Women, Can Drive, and Trip Purpose",
                color_pal = psrc_colors$pognbgy_10)
```

```{r}
#| label: tbl-trip-purp-can-drive-women
#| 
get_table_format(cd_g_t |> filter(gender2 %in% "Women"))
```
