---
title: "Trip Rates"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
params:
  errorbar: FALSE    
---

```{r packages & sources, include=FALSE}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
library(kableExtra)
install_psrc_fonts()

files <- c("data.R", "functions.R", 
           "care-giving-ols.R", "triprate-functions.R")
lapply(files, source)
```

```{r custom variables & functions, include=FALSE}
df_hts <- df_hts |> 
  hts_bin_age() |>
  hts_bin_income() |>
  hts_bin_gender()

df_hts$trip <- df_hts$trip |> mutate(care_trip = if_else(care_purpose_cat == "Care", "Yes", NA))

df_hts$hh <- df_hts$hh |> mutate(low_income = case_when(
    is.na(hhincome_bin3) ~ NA_character_,
  hhincome_bin3 == "Less than $50,000"  ~ "Yes",
  TRUE ~ "No"
))

df_hts$person <- df_hts$person |>
  mutate(age_bin4 = factor(
    case_when(age_bin5 %in% c("25-44 Years", "45-64 Years") ~ "25-64 Years",
              TRUE ~ age_bin5),
    levels = c("Under 18 Years", "18-24 Years",
               "25-64 Years", "65 years or older")
    ),
    race_simpler = factor(case_when(
      race_simple == "Child" ~ NA_character_, 
      grepl("^(Asian|Hawaiian|Islander)", race_simple)  ~ "AANHPI",
      race_simple == "Black or African American" ~"Black",
      !is.na(race_simple) ~ race_simple),
levels = c("AANHPI", "Black","Hispanic", "White", "Some Other Race", NA))
  )

df_hts_care <- data.table::copy(df_hts)
df_hts_care$trip <- df_hts_care$trip |>
  filter(care_purpose_cat == "Care")

df_hts_ols_input <- data.table::copy(df_hts) 
df_hts_ols_input$person <- df_hts_ols_input$person |>
left_join(
  df_hts$hh |> select(hh_id, low_income),
  by = "hh_id"
)

trend_years <- c(2017, 2019, 2021, 2023, 2025)

# aggregate (all-group) caregiving trip rate by year, used as the reference line/series
overall_caregiving_trend <- psrc_hts_triprate(df_hts_care, group_vars = NULL)
```

# Caregiving Trip-Rate Regression Model

To disentangle the contribution of different demographic factors to caregiving trip rates, we fit an ordinary least squares (OLS) regression model, with age, gender, race, and income as covariates, as well as the survey year. Since each variable is specified as a categorical factor, the coefficients are on the same scale and can be directly compared.

The reference case is specified as survey year 2025, with reference demographic characteristics being white, male, age 25-64, income \$50K or more.

```{r caregiving-ols}
caregiving_ols <- fit_caregiving_ols(df_hts_ols_input)

coefficient_table <- caregiving_ols$coefficients |>
  mutate(
    significant = !is.na(p.value) & p.value < 0.05,
    across(
      c(estimate, std.error, statistic, conf.low, conf.high),
      ~ if_else(is.na(.x), NA_character_, sprintf("%.2f", .x))
    ),
    p.value = if_else(is.na(p.value), NA_character_, sprintf("%.2e", p.value)),
    across(
      c(estimate, std.error, statistic, p.value, conf.low, conf.high),
      ~ kableExtra::cell_spec(
        .x,
        color = if_else(significant, "black", "#808080")
      )
    )
  ) |>
  select(-significant)

kableExtra::kbl(coefficient_table, format = "html", escape = FALSE) |>
  kableExtra::kable_styling() |>
  kableExtra::row_spec(
    0,
    bold = TRUE,
    extra_css = "border-bottom: 2px solid #000;"
  )
```

Unsurprisingly, 2021 shows lower trip rates when compared to years unaffected by the Covid-19 pandemic. Variation across the remaining surveys is relatively minor.

Among demographic characteristics, age is particularly salient. Those under 25 take fewer daily caregiving trips per day, while those over 65 take more than the working-age reference group.

Even when controlling for other factors, women on average take more daily caregiving trips than men. Respondents with incomes under \$50K take more daily caregiving trips than those with higher incomes.

There are also residual differences by racial/ethnic groups, with Hispanic & Asian/HNPI respondents making fewer caregiving trips than the rest of the population.

# Caregiving Trip-Rate Trends

These demographic differences can also be observed in trend data. 

First, we'll look at the overall caregiving trip rate trend over time. 

```{r aggregate-trend}
plot_aggregate_trend(overall_caregiving_trend, trend_years, "Caregiving Trip Rate, 2017-2025")

```

There is a gradual downward trend in the caregiving triprate over time, punctuated by a larger decline during the pandemic and a partial recovery afterward.

Keeping that trend in mind as reference, we'll now plot the difference in each demographic from the aggregate trend line--i.e. zero difference would be interpreted as matching the overall trip rate shown above.

# Age

```{r age}
age_alltrips <- psrc_hts_triprate(df_hts, group_vars="age_bin5")

age_caregiving <- psrc_hts_triprate(df_hts_care, group_vars="age_bin5")

# age_tests <- pairwise_triprate_tests(df_hts_care, "age_bin5", age_adjust = FALSE)
# summarize_consistent_differences(age_tests) |>
#   select(group_1, group_2, same_direction_every_year,
#          average_triprate_difference, consistent_difference) |>
#   mutate(average_triprate_difference = round(average_triprate_difference, 2))

plot_triprate_comparison(age_alltrips, age_caregiving, "age_bin5", "2025 Trip Rates by Age", errorbar = TRUE)
```
```{r age-trend}
plot_difference_trend(age_caregiving, overall_caregiving_trend, "age_bin5", trend_years,
                      "Trip Rate by Age, Difference from Overall Rate")
```

Age displays the largest and most consistent differences among our demographic categories. 

Those under 25 make fewer caregiving trips relative to working-age adults. Those in standard retirement years have the highest caregiving trip rate, though the trend slopes downward.

# Gender

```{r gender}
gender_alltrips <- psrc_hts_triprate(df_hts, group_vars = "gender2")

gender_caregiving <- psrc_hts_triprate(df_hts_care, group_vars = "gender2")

# gender_tests <- pairwise_triprate_tests(df_hts_care, "gender2")
# summarize_consistent_differences(gender_tests) |>
#   select(group_1, group_2, same_direction_every_year,
#          average_triprate_difference, consistent_difference) |>
#   mutate(average_triprate_difference = round(average_triprate_difference, 2))

plot_triprate_comparison(gender_alltrips, gender_caregiving, "gender2", "2025 Trip Rates by Gender", errorbar = TRUE)
```
```{r gender-trend}
plot_difference_trend(gender_caregiving, overall_caregiving_trend, "gender2", trend_years,
                      "Trip Rate by Gender, Difference from Overall Rate")
```
In terms of gender, women have consistently made more caregiving trips than have men. The discrepancy narrowed significantly in 2019 and afterward has gradually widened, but not yet returned to 2017 levels. 

Caregiving triprates for the remainder, which comprise between 2 to eight percent of the population (and is thus more subject to sampling variability) have risen over time, although unchanged in the last two surveys. An explanation for the outlier in 2019 is not immediately clear.

# Income

```{r income}
income_alltrips <- psrc_hts_triprate(df_hts, group_vars="low_income")

income_caregiving <- psrc_hts_triprate(df_hts_care, group_vars="low_income")

# income_tests <- pairwise_triprate_tests(df_hts_care, "low_income")
# summarize_consistent_differences(income_tests) |>
#   select(group_1, group_2, same_direction_every_year,
#          average_triprate_difference, consistent_difference) |>
#   mutate(average_triprate_difference = round(average_triprate_difference, 2))

plot_triprate_comparison(income_alltrips, income_caregiving, "low_income",
                         "2025 Trip Rates by Low Income Status (Under $50K)", errorbar = TRUE)
```
```{r income-trend}
plot_difference_trend(income_caregiving, overall_caregiving_trend, "low_income", trend_years,
                      "Trip Rate by Low Income Status, Difference from Overall Rate")
```
A smaller difference is visible by income category: low-income earners (here defined as under $50,000 annually, comprising between 13 and 25 percent of the population) make more caregiving trips on average than higher earners. The disparity has varied; this likely reflects persistent low response rates among those with fewer economic resources.

# Race

```{r race}
race_alltrips <- psrc_hts_triprate(df_hts, group_vars="race_simpler", incl_na = FALSE)

race_caregiving <- psrc_hts_triprate(df_hts_care, group_vars="race_simpler", incl_na = FALSE)

# race_tests <- pairwise_triprate_tests(df_hts_care, "race_simpler")
# summarize_consistent_differences(race_tests) |>
#   select(group_1, group_2, same_direction_every_year,
#          average_triprate_difference, consistent_difference) |>
#   mutate(average_triprate_difference = round(average_triprate_difference, 2))

plot_triprate_comparison(race_alltrips, race_caregiving, "race_simpler", "2025 Trip Rates by Race/ethnicity", errorbar = TRUE)
```
```{r race-trend}
plot_difference_trend(race_caregiving, overall_caregiving_trend, "race_simpler", trend_years,
                      "Trip Rate by Race/ethnicity, Difference from Overall Rate")
```

While caregiving triprates among racial/ethnic groups vary, in 2025 these differences were not large enough or consistent enough to be meaningful. A larger future survey sample may be necessary to draw substantive conclusions.
