---
title: "Household Structure"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
    fig-width: 4.5
    fig-asp: .618
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  # source(here::here("Analysis", "care-giving", f))
  source(f)
})
)
```

## Children and Elders Definition

Elder is defined as a person 65 years or over.

Children is based on var `numchildren`.

## Overall Household Composition

- Generally there are predominately more households without children or elders than households with both children and elders.

```{r}
hhs_comp <- psrc_hts_stat(df_hts,
                     analysis_unit = "hh",
                     group_vars = "hh_composition",
                     incl_na = FALSE)
```

```{r}
#| layout-ncol: 2
#| out-width: "50%"

hhs_comp |> 
  ggplot(aes(y = prop, x = hh_composition, fill = hh_composition)) +
  geom_col(position = "dodge") +
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)) +
  facet_wrap(~survey_year, ncol = 5, nrow = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_10) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        # axis.text.x = element_text(angle = 90, vjust = 0.7),
        legend.position = "bottom",
        # Shrink the legend text and title
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 9),
        
        # Shrink the legend key boxes (both width and height)
        legend.key.size = unit(0.4, "cm")
  ) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Household Composition")

hhs_comp |> 
  mutate(hh_composition_label = str_trim(str_replace_all(hh_composition, ",", "\n")),
         hh_composition_label = str_to_title(hh_composition_label)) |> 
  ggplot(aes(y = prop, x = as.character(survey_year), fill = as.character(survey_year))) +
  geom_col(position = "dodge") +
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)) +
  facet_wrap(~hh_composition_label, ncol = 5, nrow = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$blues_inc) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        # axis.text.x = element_text(angle = 90, vjust = 0.7),
        axis.text.x = element_blank(),
        legend.position = "bottom",
        # Shrink the legend text and title
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 9),
        
        # Shrink the legend key boxes (both width and height)
        legend.key.size = unit(0.4, "cm")) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Household Composition")
```

```{r}
hhs_comp
```

```{r, eval=FALSE}
test <- psrc_hts_stat(df_hts,
                      analysis_unit = "hh",
                      group_vars = c("care_trip_taken", "hh_composition"),
                      incl_na = FALSE)

test_sum_prop <- test |> 
  group_by(survey_year, care_trip_taken) |> 
  summarize(sum(prop))

test_sum_c_e <- test |> 
  group_by(survey_year) |> 
  summarize(sum(count), sum(est))
```

```{r, eval=FALSE}
test_sum_prop
```

```{r, eval=FALSE}
test_sum_c_e
```

```{r, eval=FALSE}
hhc <- psrc_hts_stat(df_hts,
                     analysis_unit = "hh",
                     group_vars = c("hh_composition"),
                     incl_na = FALSE) 
hhc_sum <- hhc |> 
   mutate(survey_year = as.character(survey_year)) |>
  group_by(survey_year) |> 
  summarize(sum_prop = sum(prop), sum_est = sum(est), sum_count = sum(count))

```

## Care Trips Taken

### Analysis A

- Where care trips were taken, more were made by households with no children or elders followed by households that have children but no elders.

- There are few households that are multi-generational with both children and elders of households that took care trips.

```{r}
#| layout-ncol: 2
#| out-width: "50%"

hhstr_a <- psrc_hts_stat(df_hts,
                    analysis_unit = "hh",
                    group_vars = c("care_trip_taken", "hh_composition"),
                    incl_na = FALSE) |>
  mutate(prop_per = percent(prop),
         prop_moe_per = percent(prop_moe))

 hhstr_a |>
  filter(care_trip_taken == 1) |> 
  ggplot(aes(y = prop, x = hh_composition, fill = hh_composition)) +
  geom_col(position = "dodge") +
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)) +
  facet_wrap(~survey_year, ncol = 5, nrow = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_10) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        # axis.text.x = element_text(angle = 90, vjust = 0.7),
        legend.position = "bottom",
         # Shrink the legend text and title
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 9),
        
        # Shrink the legend key boxes (both width and height)
        legend.key.size = unit(0.4, "cm")) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Care Trips and Household Composition")

hhstr_a |>
  filter(care_trip_taken == 1) |> 
  mutate(hh_composition_label = str_trim(str_replace_all(hh_composition, ",", "\n")),
         hh_composition_label = str_to_title(hh_composition_label)) |> 
  ggplot(aes(y = prop, x = as.character(survey_year), fill = as.character(survey_year))) +
  geom_col(position = "dodge") +
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)) +
  facet_wrap(~hh_composition_label, ncol = 5, nrow = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$blues_inc) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        # axis.text.x = element_text(angle = 90, vjust = 0.7),
        legend.position = "bottom",
         # Shrink the legend text and title
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 9),
        
        # Shrink the legend key boxes (both width and height)
        legend.key.size = unit(0.4, "cm")) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Care Trips and Household Composition")
```


```{r}
hhstr_a
```

### Analysis B

Of households where there are either elders or children present, or none at all, there are more care-giving trips made. Especially of households where there are children and no elders, there are fewer non-care-giving trips compared to other household types.

There are few households with both elders and children. Their results have very large margins of error and cannot be determined.
```{r}
hhstr_b <- psrc_hts_stat(df_hts,
                    analysis_unit = "hh",
                    group_vars = c("hh_composition","care_trip_taken"),
                    incl_na = FALSE) |>
  mutate(prop_per = percent(prop),
         prop_moe_per = percent(prop_moe))
```

```{r}
#| fig-align: center
#| out-width: "80%"
 hhstr_b |>
  ggplot(aes(y = prop, x = as.character(survey_year), fill = care_trip_taken)) +
  geom_col(position = "dodge") +
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)) +
  facet_wrap(~hh_composition, ncol = 4, nrow = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_10, labels = c("1" = "Care Trip", "0" = "Non-care Trip")) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.7),
        legend.position = "bottom",
        # Shrink the legend text and title
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 9),
        
        # Shrink the legend key boxes (both width and height)
        legend.key.size = unit(0.4, "cm")) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Household Composition and Care Trips")
```

```{r}
hhstr_b
```

