---
title: "Race and Ethnicity"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
    fig-width: 4.5
    fig-asp: .618
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  source(f)
  # source(here::here("Analysis", "care-giving", f))
})
)

```

```{r}
geog <- "Region"
 geo_col <- case_when(geog == "Region" ~ "dest_region",
                       geog %in% c("King County", "Kitsap County", "Pierce County", "Snohomish County") ~ "dest_county")
```

<!-- ## All People by Race and Ethnicity -->

```{r, eval=FALSE}

re_rs <- psrc_hts_stat(df_hts,
                     analysis_unit = "person",
                     group_vars = "race_category",
                     incl_na = FALSE) |> 
  mutate(prop_per = percent(prop),
         moe_per = percent(prop_moe))


```

```{r, eval=FALSE}
#| out-width: "80%"

title <-  "Race and Ethnicity"
color_pal <-  psrc_colors$pognbgy_10

 re_rs |> 
   mutate(survey_year = as.character(survey_year)) |> 
    ggplot(aes(x = race_category, y = prop, fill = race_category)) + #gender2, x also = fill
    geom_col(position = "dodge") + 
    geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                   orientation = "x",
                   position = position_dodge(width = 0.9)
    ) +
    facet_wrap(vars(survey_year), ncol = 5, nrow = 1, strip.position = "top") + #survey_year
    labs(x = NULL,
         y = NULL,
         fill = NULL,
         title = title) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(values = color_pal) +
    psrc_style() +
    theme(panel.grid.major.y = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_blank(),
          # legend.position = "bottom",
          # Shrink the legend text and title
          legend.text = element_text(size = 8),
          legend.title = element_text(size = 9),
          
          # Shrink the legend key boxes (both width and height)
          legend.key.size = unit(0.4, "cm") 
          # axis.text.x = element_text(angle = 90, vjust = 0.7)
    ) 
```

```{r, eval=FALSE}
re_rs
```

<!-- ## All trips by Race and Ethnicity -->

<!-- Of all trips in 2025, 47% are made by White (non-Hispanic), 15% by AANHPI. -->

```{r, eval=FALSE}
tre_rs <- psrc_hts_stat(df_hts,
                        analysis_unit = "trip",
                        group_vars = "race_category",
                        incl_na = FALSE) |> 
  mutate(prop_per = percent(prop),
         moe_per = percent(prop_moe)) |>
  filter(survey_year == 2025)

```

```{r, eval=FALSE}
tre_rs
```

```{r, eval=FALSE}
## sumcheck

at_rs |> 
  group_by(survey_year) |> 
  summarize(person_sum_est = sum(est)) # 2025: 3.87 million

tre_rs |> 
  summarize(trip_sum_est = sum(est)) # 2025 w/o NA 15,879,518, w/NA 17,100,397	
```

## Care Trips by Race and Ethnicity

### Analysis B

<!-- Of all <race>, <x>% are care-giving trips -->

- There isn't a specific group that does more care-giving trips than others across survey years.

```{r}
# race simple refers to labels being simplified (not categories)
# incl_na is FALSE to exclude NAs in care_purpose_cat which are HOME trips

recg_rs <- psrc_hts_stat(df_hts,
                     analysis_unit = "trip",
                     group_vars = c(geo_col, "race_simple", "care_purpose_cat"),
                     incl_na = FALSE) |> 
  mutate(prop_per = percent(prop),
         moe_per = percent(prop_moe))
```

```{r, eval = FALSE}
recg_rs |> 
  group_by(survey_year, .data[[geo_col]], race_simple) |> 
  summarise(sum_prop = sum(prop))
```

::: {.panel-tabset}

#### By Survey Year
```{r}
#| out-width: "100%"

col_var <- "race_simple"

recg_rs_f <- recg_rs |> 
  filter(!.data[[col_var]] %in% c("Child") & !is.na(.data[[col_var]]),
         care_purpose_cat == "Care",
         .data[[geo_col]] %in% "Region")

plot_facet_wrap(table = recg_rs_f, 
                facet = survey_year, 
                var1 = .data[[col_var]], 
                var2 = .data[[col_var]], 
                title = "Race and Ethnicity & Care Trips", 
                color_pal = psrc_colors$pognbgy_10) +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "bottom",
         legend.text = element_text(size = 6),
          # legend.title = element_text(size = 9),
          
          # Shrink the legend key boxes (both width and height)
          legend.key.size = unit(0.2, "cm") 
  )
```

#### By Race Simple

```{r}
#| out-width: "100%"
plot_facet_wrap(table = recg_rs_f, 
                facet = .data[[col_var]], 
                var1 = as.character(survey_year), 
                var2 = as.character(survey_year), 
                title = "Race and Ethnicity & Care Trips", 
                color_pal = psrc_colors$blues_inc) +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "bottom",
         legend.text = element_text(size = 6),
          # legend.title = element_text(size = 9),
          
          # Shrink the legend key boxes (both width and height)
          legend.key.size = unit(0.2, "cm"),
        strip.text = element_text(size = 9)
  )
```
:::

```{r}
recg_rs_f 

```

```{r, eval=FALSE}
recg_rs |>
  group_by(survey_year, race_category) |>
  summarize(sum_prop = sum(prop))
```

#### Aggregate groups (People of Color)

- There isn't a group that takes more or less care-trips than the other.

```{r}
# race_binary refers to POC, White, Child

rebcg_rs <- psrc_hts_stat(df_hts,
                     analysis_unit = "trip",
                     group_vars = c(geo_col, "race_binary", "care_purpose_cat"),
                     incl_na = FALSE) |> 
  mutate(prop_per = percent(prop),
         moe_per = percent(prop_moe))
```

```{r, eval=FALSE}
# sumcheck

rebcg_rs_test <- psrc_hts_stat(df_hts,
                     analysis_unit = "trip",
                     group_vars = c(geo_col, "race_binary", "care_purpose_cat"),
                     incl_na = TRUE)

rebcg_rs_test |> 
  group_by(dest_region, survey_year, race_binary) |> 
  summarise(sum_prop = sum(prop))
```

::: {.panel-tabset}

#### By Survey Year
```{r}
#| out-width: "100%"

col_var <- "race_binary"

rebcg_rs_f <- rebcg_rs |> 
  filter(!.data[[col_var]] %in% c("Child") & !is.na(.data[[col_var]]),
         care_purpose_cat == "Care",
         .data[[geo_col]] %in% "Region")

plot_facet_wrap(table = rebcg_rs_f, 
                facet = survey_year, 
                var1 = .data[[col_var]], 
                var2 = .data[[col_var]], 
                title = "Race and Ethnicity & Care Trips", 
                color_pal = psrc_colors$pognbgy_10) +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "bottom",
         legend.text = element_text(size = 8),
          legend.title = element_text(size = 9),
          
          # Shrink the legend key boxes (both width and height)
          legend.key.size = unit(0.4, "cm") 
  )

```

#### By Race Binary

```{r}
#| out-width: "100%"

plot_facet_wrap(table = rebcg_rs_f, 
                facet = .data[[col_var]], 
                var1 = as.character(survey_year), 
                var2 = as.character(survey_year), 
                title = "Race and Ethnicity & Care Trips", 
                color_pal = psrc_colors$blues_inc) +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "bottom",
         legend.text = element_text(size = 8),
          legend.title = element_text(size = 9),
          
          # Shrink the legend key boxes (both width and height)
          legend.key.size = unit(0.4, "cm") 
  )
```
:::

```{r}
rebcg_rs_f
```

### Analysis A

- Of all care-giving trips in the Region across survey years, the largest share are made by those who are White Adults.

```{r}
col_var <- "race_category"

cgre_rs <- psrc_hts_stat(df_hts,
                     analysis_unit = "trip",
                     group_vars = c(geo_col, "care_purpose_cat", col_var),
                     incl_na = TRUE) |> 
  mutate(prop_per = percent(prop),
         moe_per = percent(prop_moe))
```

```{r, eval=FALSE}
# Do all the elements of the regional care-purpose facet equal to 100%?

cgre_rs |> 
  group_by(survey_year, .data[[geo_col]], care_purpose_cat) |> 
  summarise(sum_prop = sum(prop))
```

```{r}
#| out-width: "80%"

cgre_rs_f <- cgre_rs |> 
  filter(.data[[geo_col]] %in% "Region",
         care_purpose_cat == "Care")

plot_facet_wrap(table = cgre_rs_f, 
                facet = survey_year, 
                var1 = .data[[col_var]], 
                var2 = .data[[col_var]], 
                title = "Care Trips by Race and Ethnicity", 
                color_pal = psrc_colors$pognbgy_10) +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "bottom",
         legend.text = element_text(size = 8),
          legend.title = element_text(size = 9),
          
          # Shrink the legend key boxes (both width and height)
          legend.key.size = unit(0.4, "cm") 
  )

```

```{r}
cgre_rs_f
```

#### Aggregate groups (People of Color)

- Of all care-trips in the region across survey years, those who are White adults have a greater share compared to Adult People of Color; however, care-trip making among Adult People of Color has increased over time since 2019.

```{r}
col_var <- "race_binary"

cgreb_rs <- psrc_hts_stat(df_hts,
                     analysis_unit = "trip",
                     group_vars = c(geo_col, "care_purpose_cat", col_var),
                     incl_na = TRUE) |> 
  mutate(prop_per = percent(prop),
         moe_per = percent(prop_moe))
```

```{r}
#| out-width: "80%"

cgreb_rs_f <- cgreb_rs |> 
  filter(.data[[geo_col]] %in% "Region",
         care_purpose_cat == "Care")

plot_facet_wrap(table = cgreb_rs_f, 
                facet = survey_year, 
                var1 = .data[[col_var]], 
                var2 = .data[[col_var]], 
                title = "Care Trips by Race and Ethnicity", 
                color_pal = psrc_colors$pognbgy_10) +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "right",
         legend.text = element_text(size = 8),
          legend.title = element_text(size = 9),
          
          # Shrink the legend key boxes (both width and height)
          legend.key.size = unit(0.4, "cm") 
  )
```

```{r}
cgreb_rs_f
```
