---
title: "Trip Purpose"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  source(f)
  # source(here::here("Analysis", "care-giving", f))
})
)

```

## Trip Purpose

### PSRC Destination Purpose Assignment

Below are trips with their destination in the Region (King, Kitsap, Pierce, and Snohomish Counties) for all survey years. Care-giving is an aggregate of the following destination purposes (escort, personal business/errand/appointment, meal, and shopping). The care-giving analysis will also exclude HOME trips.

```{r}

# This section prints a table with weighted/unweighted x Home/Exclude Home x trip purpose

trip_purp_a <- create_trip_purpose_tbl(df_hts)

trip_purp_a |>
  gt() |>
  cols_align(
    align = "left",
    columns = dest_purpose_cat
  ) |>
  tab_spanner_delim(delim = ".") |>
  tab_style(
    style = cell_fill(color = "lightyellow"),
    locations = cells_body(
      columns = dest_purpose_cat,
      rows = dest_purpose_cat %in% care_purpose_cat  # Matches values in the vector
    )
  ) |>
  data_color(
    columns = contains("Exclude"),
    rows = dest_purpose_cat %in% care_purpose_cat,
    colors = "lightyellow"
  )


```

### Care Purpose

Care is defined as the following purposes: Escort, Shopping, Meal, Personal Business/Errand/Appointment. All `HOME` trips are excluded. The remaining categories from the `dest_purpose_cat` either stand alone or have been aggregated to the following:

**Care**: `r care_purpose_cat`

**Work**: `r work_cat`

**Other**: `r other_cat`

**School**: `r school_cat`

Care related trips make up the majority of trips compared to other purpose categories across survey years.

```{r}
care_purp <- create_care_purpose_tbl(df_hts)
```

```{r}
#| layout-ncol: 2
#| out-width: "50%"

plot_facet_wrap(care_purp, 
                survey_year, 
                care_purpose_cat, 
                care_purpose_cat, 
                "Care Purpose Categories by Year", 
                color_pal = psrc_colors$pognbgy_5) +
  theme(axis.text.x = element_blank())

plot_facet_wrap(care_purp, 
                care_purpose_cat, 
                as.character(survey_year), 
                as.character(survey_year), 
                "Care Purpose Categories by Year", 
                color_pal = psrc_colors$blues_inc) +
  theme(axis.text.x = element_blank())
```

```{r}
care_purp
```
