---
title: "Gender"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
params:
  errorbar: FALSE    
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  # print(here::here("Analysis", "care-giving", f))
  # source(here::here("Analysis", "care-giving", f))
  source(f)
})
)
source("../util.R")

```

```{r}
geog <- "Region"
 geo_col <- case_when(geog == "Region" ~ "dest_region",
                       geog %in% c("King County", "Kitsap County", "Pierce County", "Snohomish County") ~ "dest_county")
```

## All trips

```{r}
at_rs <- psrc_hts_stat(df_hts,
                     analysis_unit = "trip",
                     group_vars = c(geo_col, "gender2"),
                     incl_na = FALSE) |> 
  mutate(prop_per = percent(prop),
         moe_per = percent(prop_moe))

# Which group is doing more?
# How many trips overall are made by Females in the region?

# 17.1 million trips (for everyone), check persons/trips tables sum.
```

```{r}
#| layout-ncol: 2
#| out-width: "50%"

plot_facet_wrap(table = at_rs,
                facet = survey_year,
                var1 = gender2,
                var2 = gender2,
                title = "All Trips by Gender") +
  theme(axis.text.x = element_blank())

plot_facet_wrap(table = at_rs,
                facet = gender2,
                var1 = as.character(survey_year),
                var2 = as.character(survey_year),
                title = "All Trips by Gender",
                color_pal = psrc_colors$blues_inc) +
  theme(axis.text.x = element_blank())
```

```{r}
at_rs
```

```{r, eval=FALSE}
#### Check sums
rs2 |> 
  summarize(trips_sum_est = sum(est)) # M + F = 13,766,970	

sum_trips <- hts_data$trip |> 
  group_by(survey_year) |> 
  summarize(trip_weight_sum = sum(trip_weight)) # 2025: 17,100,397

sum_trips

rs3 <- psrc_hts_stat(df_hts,
                     analysis_unit = "trip",
                     group_vars = c(geo_col, "gender2"),
                     incl_na = TRUE) 

rs3 |> 
  group_by(survey_year) |> 
  summarize(trip_sum_est = sum(est)) # 2025: M + F + NA - 17,100,397

```

## Care Trips and Gender

The following charts and tables show the share of trips by gender for care-giving purposes.

### Gender by Trip Purpose

<!-- Comparing genders within each purpose -->

- Of all care trips, Women made over 50% of those trips across most survey years.

- Of all care trips in 2025, 53% are made by Women and 40% by Men.

```{r}
ctg_rs <- psrc_hts_stat(df_hts,
                    analysis_unit = "trip",
                    group_vars = c(geo_col, "care_purpose_cat", "gender2"),
                    incl_na = TRUE) |> 
  mutate(prop_per = percent(prop),
         moe_per = percent(prop_moe))

# Which group is doing more?
# Of all care trips made in , X are made by Females. Compared to all trips, X are made by Females.
```

```{r}
#| layout-ncol: 2
#| out-width: "50%"

ctg_rs_f <- ctg_rs |> 
  filter(.data[[geo_col]] %in% "Region",
         care_purpose_cat == "Care") |> 
  mutate(survey_year = as.character(survey_year))

plot_facet_wrap(table = ctg_rs_f,
                facet = survey_year,
                var1 = gender2,
                var2 = gender2,
                title = "Care Trips by Gender") +
 geom_hline(aes(yintercept = .5, linetype = "50%"), color = "red") +
  scale_linetype_manual(name = NULL, values = c("50%" = "dashed"))+
  theme(axis.text.x = element_blank())

plot_facet_wrap(table = ctg_rs_f,
                facet = gender2,
                var1 = survey_year,
                var2 = survey_year,
                title = "Care Trips by Gender",
                color_pal = psrc_colors$blues_inc) +
   geom_hline(aes(yintercept = .5, linetype = "50%"), color = "red") +
  scale_linetype_manual(name = NULL, values = c("50%" = "dashed"))+
  theme(axis.text.x = element_blank())

```

```{r}
ctg_rs_f 
```

### Trip Purpose by Gender

The following charts and table show the share of care-giving trips by gender.

- Of all trips made by Women in 2025, 53% are care-giving trips. Of all trips made by Men in 2025, 45% are care-giving trips.

```{r}
create_trip_purp_gender_chart <- function(table) {
  plot_facet_wrap(table = table,
                facet = survey_year,
                var1 = .data[[col_var]],
                var2 = care_purpose_cat,
                title = NULL, 
                color_pal = psrc_colors$pognbgy_10) +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "bottom"
  )
}

```


```{r}
# exclude where care_purpose_cat is NA because those are home trips, and b/c each gender is its own universe,
# NAs can also be filtered out

gct_rs <- psrc_hts_stat(df_hts,
                    analysis_unit = "trip",
                    group_vars = c(geo_col, "gender2", "care_purpose_cat"),
                    incl_na = FALSE) |> 
  mutate(prop_per = percent(prop),
         moe_per = percent(prop_moe))
```

<!-- ::: panel-tabset -->

```{r}
col_var <- "gender2"

gct_rs_f <- gct_rs |> 
  filter(.data[[geo_col]] %in% "Region",
         !is.na(.data[[col_var]])) |> 
  mutate(survey_year = as.character(survey_year))
```

#### Men

```{r}
#| label: fig-men-trip-purp
#| fig-cap: "Men and Trip Purpose"
#| fig-width: 8
#| out-width: "70%"

create_trip_purp_gender_chart(table = gct_rs_f |> filter(gender2 == "Men"))
  

```

```{r}
#| label: tbl-men-trip-purp
#| fig-cap: "Men and Trip Purpose Table"
knitr::kable(gct_rs_f |> 
  filter(gender2 == "Men"))
```

#### Women

```{r}
#| label: fig-women-trip-purp
#| fig-cap: "Women and Trip Purpose"
#| fig-width: 8
#| out-width: "70%"

create_trip_purp_gender_chart(table = gct_rs_f |> filter(gender2 == "Women"))

```

```{r}
#| label: tbl-women-trip-purp
#| fig-cap: "Women and Trip Purpose Table"
knitr::kable(gct_rs_f |> 
  filter(gender2 == "Women"))
```

#### Prefer not to say or Other

```{r}
#| label: fig-other-trip-purp
#| fig-cap: "Gender Prefer Not to Say and Trip Purpose"
#| fig-width: 8
#| out-width: "70%"

create_trip_purp_gender_chart(table = gct_rs_f |> filter(!gender2 %in% c("Men", "Women")))

```

```{r}
#| label: tbl-other-trip-purp
#| fig-cap: "Gender Prefer Not to Say and Trip Purpose Table"
knitr::kable(gct_rs_f |> 
  filter(!gender2 %in% c("Men", "Women")))
```

<!-- ::: -->
