---
title: "Care Giving Analysis"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}
source(here::here("Analysis", "care-giving", "scratch.R"))
```

## Trip Purpose

### PSRC destination purpose assignment

- Trips with destination in the Region: King, Kitsap, Pierce, and Snohomish Counties

```{r}

# This section prints a table with weighted/unweighted x Home/Exclude Home x trip purpose

rs1 <- psrc_hts_stat(df_hts,
                     analysis_unit = "trip",
                     group_vars = c("dest_region","dest_purpose_cat"),
                     incl_na = FALSE) |>
  mutate(type = "Include HOME")

rs1_no_home <- psrc_hts_stat(df_hts,
                             analysis_unit = "trip",
                             group_vars=c("dest_region","dest_purpose_cat_no_home"),
                             incl_na = FALSE)  |>
  mutate(type = "Exclude HOME") |>
  rename(dest_purpose_cat = dest_purpose_cat_no_home)

# Region

rs <- rs1 |>  
  add_row(rs1_no_home)

rs_reg <- rs |>
  mutate(unweighted = percent(count/sum(count)),
         weighted = percent(prop),
         .by = "type")

per_group <- c("survey_year", "dest_purpose_cat")

per_unweighted <- rs_reg |> 
  pivot_wider(id_cols = all_of(per_group), 
              names_from = "type", 
              values_from = "unweighted",
              names_prefix = "unweighted- ")

per_weighted <- rs_reg |> 
  pivot_wider(id_cols = all_of(per_group), 
              names_from = "type", 
              values_from = "weighted",
              names_prefix = "weighted- ")

per_unweighted %>%
  left_join(per_weighted, by = per_group) |> 
  gt() |> 
  tab_spanner(
    label = "Weighted",
    columns = starts_with("weighted")
  ) |> 
  tab_spanner(
    label = "Unweighted",
    columns = starts_with("Unweighted")
  ) |> 
  cols_label(
    ends_with("Include HOME") ~ "Include Home",
    ends_with("Exclude HOME") ~ "Exclude Home",
    dest_purpose_cat ~ "Destination Purpose Category"
  ) |> 
   cols_align(
    align = "left",
    columns = dest_purpose_cat
  ) |> 
  cols_hide(survey_year) |> 
  tab_caption(caption = paste("Survey Year:", survey_year))


```

### Care purpose

- Care is defined as the following purposes: Escort, Shopping, Meal, Personal Business/Errand/Appointment
  - all HOME trips are excluded
  
Care: `r care_purpose_cat`

Work: `r work_cat`

Other: `r other_cat`

School: `r school_cat`
  
```{r}
rs_care_purp <- psrc_hts_stat(df_hts, 
                       analysis_unit = "trip", 
                       group_vars = c("dest_region", "care_purpose_cat"),
                       incl_na = FALSE) %>%
  mutate(type = "percent") |>  
  mutate(unweighted = percent(count/sum(count)),
         weighted = percent(prop),
         moe = percent(prop_moe),
         .by = "type")

rs_care_purp |> 
  select(all_of(c("care_purpose_cat","unweighted","weighted","moe"))) |> 
  gt() |> 
  cols_label_with(fn = ~str_to_title(.)) |> 
  cols_label(
    care_purpose_cat ~ "Care Purpose Categories",
    moe ~ "MOE"
  ) |> 
   cols_align(
    align = "left",
    columns = 1
  )
```
## Care trips by mode and income
