---
title: "Can Drive"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
    fig-width: 4.5
    fig-asp: .618
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}
geog <- "Region"
 geo_col <- case_when(geog == "Region" ~ "dest_region",
                       geog %in% c("King County", "Kitsap County", "Pierce County", "Snohomish County") ~ "dest_county")
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  source(f)
  # source(here::here("Analysis", "care-giving", f))
})
)

```

## Can Drive by Age

- There were fewer 18-24 year olds who could drive in 2017/19 than those who couldn't. By 2023, there were significantly less who were driving.

- There were more 25-34 year olds who could drive compared to those who couldn't in 2021, but it's hard to determine in other years.

- Between 35-64 years in 2017/19 and 2025, there are more who could drive than couldn't. In 2023, there are more people who can drive than those who cannot in the 35-54 age range.

- Those who are between 16-24 years of age have a higher share who cannot drive. The earliest a person can obtain a learner's permit is 15 in addition to people who delay learning how to drive.

```{r}
per_cda <- psrc_hts_stat(df_hts,
                    analysis_unit = "person",
                    group_vars = c("can_drive2", "age"),
                    incl_na = FALSE)

# per_cda

```

```{r}
#| out-width: "100%"
#| fig-align: "center"

per_cda |> 
  ggplot(aes(x = age, y = prop, fill = as.character(can_drive2))) +
  geom_col(position = "dodge") + 
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)
  ) +
  facet_wrap(~survey_year, ncol = 5, nrow = 1, strip.position = "top") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.7),
        # Shrink the legend text and title
          legend.text = element_text(size = 8),
          legend.title = element_text(size = 9),
          
          # Shrink the legend key boxes (both width and height)
          legend.key.size = unit(0.4, "cm") 
  ) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Persons by Age who Can or Cannot Drive")
```

```{r}
per_cda
```

## Care Trips by Can Drive and Age

### Analysis A

- Where care trips were taken, the vast majority of those between the ages of 25-84 years can drive.

- Where there are more care trips by those who cannot drive can be found by those who are either too young to obtain a learners permit or those 85 years or older, depending on the survey year.

```{r}
age_groups <- c("12-15 years", "16-17 years", "18-24 years", "25-34 years", "35-44 years", "45-54 years",
                 "55-64 years", "65-74 years", "75-84 years", "85 years or older")

t_cda <- psrc_hts_stat(df_hts,
                      analysis_unit = "trip",
                      group_vars = c(geo_col, "care_purpose_cat", "age", "can_drive2"),
                      incl_na = FALSE) |>
    filter(.data[[geo_col]] == geog,
           care_purpose_cat == "Care") |>
    mutate(age = factor(age, levels = age_groups)) |>
    arrange(age)


```

::: panel-tabset
#### Yes (can drive)

```{r}
#| out-width: "100%"

plot_facet_wrap(table = t_cda |> filter(can_drive2 %in% "Yes"),
                facet = survey_year,
                var1 = age,
                var2 = age,
                title = "Age, Yes, can drive and Care Trips",
                color_pal = psrc_colors$pognbgy_10) +
  geom_hline(yintercept = .5, color = "red", linetype = "dashed")

```

#### No (can drive)

```{r}
#| out-width: "100%"

plot_facet_wrap(table = t_cda |> filter(can_drive2 %in% "No"),
                facet = survey_year,
                var1 = age,
                var2 = age,
                title = "Age, No, cannot drive and Care Trips",
                color_pal = psrc_colors$pognbgy_10) +
  geom_hline(yintercept = .5, color = "red", linetype = "dashed")
```
:::

```{r}
t_cda

```

<!-- ### Analysis B -->

```{r, eval=FALSE}

# Of those that can drive between ages X, X% are care-giving trips.

# This might be a difficult analysis because of all the different age groups and lower sample sizes

cda_t <- psrc_hts_stat(df_hts,
                      analysis_unit = "trip",
                      group_vars = c(geo_col, "can_drive2", "age", "care_purpose_cat"),
                      incl_na = FALSE) |>
    filter(.data[[geo_col]] == geog) |>
    mutate(age = factor(age, levels = age_groups)) |>
    arrange(age)

cda_t
```

## Care Trips by Can Drive and Gender

### Analysis A

- Of all care-trips in the Region and those who can drive, there is a higher share of Women than Men for most survey years

```{r}
# Care trips by Gender

t_cd_g <- psrc_hts_stat(df_hts,
                      analysis_unit = "trip",
                      group_vars = c(geo_col, "care_purpose_cat", "can_drive2", "gender2"),
                      incl_na = TRUE) 
```

```{r}
t_cd_g_f <- t_cd_g |>
  filter(.data[[geo_col]] == geog, 
         care_purpose_cat %in% "Care")
```

```{r, eval=FALSE}
# sumcheck
t_cd_g_f |> 
  group_by(survey_year, can_drive2) |> 
  summarise(sum_prop = sum(prop))
```

::: panel-tabset
#### Yes (can drive)

```{r}
#| out-width: "100%"

plot_facet_wrap(table = t_cd_g_f |> filter(can_drive2 %in% "Yes"),
                facet = survey_year,
                var1 = gender2,
                var2 = gender2,
                title = "Care Trips, Yes-can drive, and Gender",
                color_pal = psrc_colors$pognbgy_10) +
  geom_hline(yintercept = .5, color = "red", linetype = "dashed")
```

#### No (can drive)

```{r}
#| out-width: "100%"

plot_facet_wrap(table = t_cd_g_f |> filter(can_drive2 %in% "No"),
                facet = survey_year,
                var1 = gender2,
                var2 = gender2,
                title = "Care Trips, No-cannot drive, and Gender",
                color_pal = psrc_colors$pognbgy_10) +
  geom_hline(yintercept = .5, color = "red", linetype = "dashed")
```
:::

```{r}
t_cd_g_f
```

### Analysis B

```{r}
cd_g_t <- psrc_hts_stat(df_hts,
                      analysis_unit = "trip",
                      group_vars = c(geo_col, "can_drive2", "gender2", "care_purpose_cat"),
                      incl_na = FALSE) 

cd_g_t
```
Of all Men or Women who can drive, the predominant trip purpose is care-giving trips. For 4 out of the 5 survey years, over 50% of trips made by Women who can drive are care-related trips. 

For Men and Women who cannot drive, trips are primarily care-giving trips for most survey years.

::: panel-tabset
#### Yes (can drive)

```{r}
#| out-width: "100%"

plot_facet_wrap(table = cd_g_t |> filter(can_drive2 %in% "Yes"),
                facet = survey_year,
                var1 = gender2,
                var2 = care_purpose_cat,
                title = "Can Drive, Gender, Care Purpose Category",
                color_pal = psrc_colors$pognbgy_10) +
  geom_hline(yintercept = .5, color = "red", linetype = "dashed")
```

#### No (can drive)

```{r}
#| out-width: "100%"

plot_facet_wrap(table = cd_g_t |> filter(can_drive2 %in% "No"),
                facet = survey_year,
                var1 = gender2,
                var2 = care_purpose_cat,
                title = "Can Drive, Gender, Care Purpose Category",
                color_pal = psrc_colors$pognbgy_10) +
  geom_hline(yintercept = .5, color = "red", linetype = "dashed")
```
:::
