---
title: "Gender"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
    fig-width: 4.5
    fig-asp: .618
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  source(here::here("Analysis", "care-giving", f))
})
)

```

## Gender

Across survey years, a higher percentage of females are taking care trips compared to males.

```{r}
gender_df <- create_trips_gender_tbl(geog = "Region")

gender_df$wide |> 
  gt()|>
  tab_spanner_delim(delim = ".") |>
  cols_hide(c(dest_loc, care_purpose_cat)) |>
  cols_align(align = "left",
             columns = c(3))
```

```{r}
#| out-width: "100%"
#| fig-align: "center"

gender_df$long |> 
  ggplot(aes(x = as.character(survey_year), y = prop, fill = gender2)) +
  geom_col() +
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                orientation = "x") +
  facet_wrap(~gender2, strip.position = "bottom") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        strip.placement = "outside",
        panel.spacing = unit(0, "cm"),
        legend.position = "bottom",
        strip.text = element_text(face = "bold")) +
  labs(x = NULL,
       y = NULL,
       fill = NULL)

```

## Gender by Trip Mode

Driving was the primary mode of transportation for care-related trips in 2023, with a higher share among females than males. By 2025, this pattern becomes more balanced, with approximately 89% of both males and females using driving for care trips. Transit and walking account for a larger proportion of non-driving trips and are more commonly used by males than females.

```{r}
gender_mode_df <- create_mode_gender_tbl(geog = "Region")
```

```{r}
gender_mode_df$wide |> 
  gt()|> 
  tab_spanner_delim(delim = ".") |>
  cols_hide(c(dest_loc, care_purpose_cat)) |>
  cols_align(align = "left",
             columns = c(3))
```

```{r}
#| out-width: "100%"
#| fig-align: "center"

gender_mode_df$long |> 
  ggplot() +
  geom_col(aes(x = as.character(survey_year), y = prop, fill = mode_class_5)) +
  facet_wrap(~gender2, strip.position = "bottom") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        strip.placement = "outside",
        panel.spacing = unit(0, "cm"),
        legend.position = "bottom",
        strip.text = element_text(face = "bold")) +
  labs(x = NULL,
       y = NULL)

```
