---
title: "Household Structure"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
    fig-width: 4.5
    fig-asp: .618
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  source(here::here("Analysis", "care-giving", f))
})
)
```

## Children and Elders

Elder is defined as a person 65 years or over.

Of those who have taken care purpose trips, does their household have children and/or elders?

`has_children` refer to households where there are 1 or more children in the household based on var `numchildren`.

`has_elders` refer to households where there is someone 65 years or older in the household.

```{r}
create_hhstr_tbl(geog = "Region")$wide |>
  gt(rowname_col = "household_members") |> 
  tab_spanner_delim(delim = ".") |>
  cols_hide(c(has_children, has_elders)) |>
  cols_align(align = "left",
             columns = c(2))
```

```{r}
create_hhstr_tbl(geog = "Region")$long |>
  ggplot(aes(x = as.character(survey_year), y = prop, fill = household_members)) +
  geom_col() +
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                orientation = "x") +
  facet_grid(rows = vars(has_children), cols = vars(has_elders)) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        strip.placement = "outside",
        panel.spacing = unit(.5, "cm"),
        legend.position = "right",
        strip.text = element_blank()) +
  labs(x = NULL,
       y = NULL,
       fill = NULL)
```

The table below shows households where care trips were taken the share of households that have both children and elders.

```{r}
create_hhstr_tbl_b(geog = "Region")$wide |> 
  gt(rowname_col = "household_members") |> 
  tab_spanner_delim(delim = ".") |>
  cols_hide(c(children_or_elders)) |>
  cols_align(align = "left",
             columns = c(2))

```
```{r}
create_hhstr_tbl_b(geog = "Region")$long |>
  ggplot(aes(fill = as.character(survey_year), y = prop, x = household_members)) +
  geom_col(position = "dodge") +
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                orientation = "x",
                position = position_dodge(width = 0.9)) +
  # facet_grid(rows = vars(), cols = vars()) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        strip.placement = "outside",
        panel.spacing = unit(.5, "cm"),
        legend.position = "right",
        strip.text = element_blank()) +
  labs(x = NULL,
       y = NULL,
       fill = NULL)
```

