---
title: "Can Drive"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
    fig-width: 4.5
    fig-asp: .618
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  source(here::here("Analysis", "care-giving", f))
})
)

```

## Can Drive by Age

The table below shows of those who have taken care trips, the share of those who can or can't drive. Most drivers are between the ages of 16-84 years.A consistent share across survey years is between 25-84 years. There is increase drivers 85 years and older in 2025 than in 2023. For those between 18-24 years, there are less drivers compared to the previous survey year.

```{r}
df <- create_candrive_tbl(geog = "Region")

# make Age col wider or as row name, factor years.

df$wide |> 
  gt(rowname_col = "age")|>
  cols_width(
    stub() ~ px(500)
  ) |> 
  tab_spanner_delim(delim = ".") |>
  cols_hide(c(dest_loc)) |>
  cols_align(align = "left",
             columns = c(3))
```

```{r}
#| out-width: "100%"
#| fig-align: "center"

df$long |> 
  ggplot(aes(x = age, y = prop, fill = as.character(survey_year))) +
  geom_col(position = "dodge") + 
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                orientation = "x",
                position = position_dodge(width = 0.9)
                ) +
  facet_wrap(~can_drive2, ncol = 1, strip.position = "right") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5[1:2]) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        strip.placement = "outside",
        panel.spacing = unit(1, "cm"),
        legend.position = "bottom",
        strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, vjust = 0.7)) + #, hjust=1
  labs(x = NULL,
       y = NULL,
       fill = NULL)
```

## Can Drive by Age and Gender

```{r}
# Age and Gender as row names

df_ag <- create_candrive_tbl_b(geog = "Region")

```

```{r}

# make Age col wider or as row name, factor years.

df_ag$wide |> 
  gt(rowname_col = "age")|>
  cols_width(
    stub() ~ px(500)
  ) |> 
  tab_spanner_delim(delim = ".") |>
  cols_hide(c(dest_loc)) |>
  cols_align(align = "left",
             columns = c(3))
```

```{r}
#| out-width: "100%"
#| fig-align: "center"

df_ag$long |> 
  ggplot(aes(x = age, y = prop, fill = as.character(survey_year))) +
  geom_col(position = "dodge") + 
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                orientation = "x",
                position = position_dodge(width = 0.9)
                ) +
  facet_grid(cols = vars(can_drive2), rows = vars(gender2)) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5[1:2]) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        strip.placement = "outside",
        panel.spacing = unit(1, "cm"),
        legend.position = "bottom",
        strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 90, vjust = 0.7)) +
  labs(x = NULL,
       y = NULL,
       fill = NULL)
```

### Cannot Drive

A larger share of Females between 35-54 years who took care trips cannot not drive in 2023 compared to Males. In 2025, the share is not as high as compared to the previous survey year.

```{r}
#| out-width: "100%"
#| fig-align: "center"

df_ag$long |> 
  filter(survey_year == 2023 & can_drive2 == "No") |> 
  ggplot(aes(x = age, y = prop, fill = gender2)) +
  geom_col() + 
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5[1:2]) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        legend.position = "bottom",
        axis.text.x = element_text(angle = 45, vjust = 0.7)) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "2023 Cannot Drive")
```

```{r}
#| out-width: "100%"
#| fig-align: "center"

df_ag$long |> 
  filter(survey_year == 2025 & can_drive2 == "No") |> 
  ggplot(aes(x = age, y = prop, fill = gender2)) +
  geom_col() + 
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5[1:2]) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        legend.position = "bottom",
        axis.text.x = element_text(angle = 45, vjust = 0.7)) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "2025 Cannot Drive")
```
