---
title: "Household Structure"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
    fig-width: 4.5
    fig-asp: .618
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  source(here::here("Analysis", "care-giving", f))
})
)
```

## Children and Elders

Elder is defined as a person 65 years or over. 

Of those who have taken care purpose trips, does their household have children and/or elders?

`has_children` refer to households where there are 1 or more children in the household based on var `numchildren`.

`has_elders` refer to households where there is someone 65 years or older in the household.

```{r}
create_hhstr_tbl(geog = "Region") |> 
  gt()
```

The table below shows households where care trips were taken the share of households that have both children and elders.

```{r}
create_hhstr_tbl_b(geog = "Region") |> 
  gt()

```

