---
title: "Gender"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
    fig-width: 4.5
    fig-asp: .618
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  source(here::here("Analysis", "care-giving", f))
})
)

```

```{r}
geog <- "Region"
 geo_col <- case_when(geog == "Region" ~ "dest_region",
                       geog %in% c("King County", "Kitsap County", "Pierce County", "Snohomish County") ~ "dest_county")
```

## All trips

Of all trips, Women have usually made more trips than Men for all survey years except 2021.

```{r}
at_rs <- psrc_hts_stat(df_hts,
                     analysis_unit = "trip",
                     group_vars = c(geo_col, "gender2"),
                     incl_na = FALSE) |> 
  mutate(prop_per = percent(prop),
         moe_per = percent(prop_moe)) |> 
  select(-c(dest_region))

# Which group is doing more?
# How many trips overall are made by Females in the region?

# 17.1 million trips (for everyone), check persons/trips tables sum.
```

```{r}
#| layout-ncol: 2
#| out-width: "50%"

at_rs |> 
  ggplot(aes(x = as.character(survey_year), y = prop, fill = gender2)) +
  geom_col(position = "dodge") + 
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)
  ) +
  # facet_wrap(~survey_year, ncol = 4, nrow = 1, strip.position = "top") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.7)
  ) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "All Trips by Gender")

at_rs |> 
  ggplot(aes(x = gender2, y = prop, fill = as.character(survey_year))) +
  geom_col(position = "dodge") + 
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)
  ) +
  # facet_wrap(~survey_year, ncol = 4, nrow = 1, strip.position = "top") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.7)
  ) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "All Trips by Gender")
```
```{r}
at_rs
```



```{r, eval=FALSE}
#### Check sums
rs2 |> 
  summarize(trips_sum_est = sum(est)) # M + F = 13,766,970	

sum_trips <- hts_data$trip |> 
  group_by(survey_year) |> 
  summarize(trip_weight_sum = sum(trip_weight)) # 2025: 17,100,397

sum_trips

rs3 <- psrc_hts_stat(df_hts,
                     analysis_unit = "trip",
                     group_vars = c(geo_col, "gender2"),
                     incl_na = TRUE) 

rs3 |> 
  group_by(survey_year) |> 
  summarize(trip_sum_est = sum(est)) # 2025: M + F + NA - 17,100,397

```

## Care Trips by Gender

Of all care trips across survey years, over 50% are made by Women.

Of all care trips in 2025, 57% are made by Women and 42% by Men.

```{r}
ctg_rs <- psrc_hts_stat(df_hts,
                    analysis_unit = "trip",
                    group_vars = c(geo_col, "care_purpose_cat", "gender2"),
                    incl_na = FALSE) |> 
  mutate(prop_per = percent(prop),
         moe_per = percent(prop_moe)) |> 
  select(-c(dest_region))
# rs

# Which group is doing more?
# Of all care trips made in , X are made by Females. Compared to all trips, X are made by Females.
```

```{r}
#| layout-ncol: 2
#| out-width: "50%"

ctg_rs_f <- ctg_rs |> 
  filter(care_purpose_cat == "Care") |> 
  mutate(survey_year = as.character(survey_year))

plot_facet_wrap(table = ctg_rs_f, 
                facet = survey_year, 
                var1 = gender2, 
                var2 = gender2, 
                title = "Care Trips by Gender")

plot_facet_wrap(table = ctg_rs_f, 
                facet = gender2, 
                var1 = survey_year, 
                var2 = survey_year, 
                title = "Care Trips by Gender",
                color_pal = psrc_colors$blues_inc)

```

```{r}
ctg_rs
```

## Gender and Care Trips

Of all trips made by Women across survey years, more than 50% are care-giving trips.

Of all trips made by Women in 2025, 53% are care-giving trips. Of all trips made by Men in 2025, 45% were care-giving trips.

```{r}
gct_rs <- psrc_hts_stat(df_hts,
                    analysis_unit = "trip",
                    group_vars = c(geo_col, "gender2", "care_purpose_cat"),
                    incl_na = FALSE) |> 
  mutate(prop_per = percent(prop),
         moe_per = percent(prop_moe)) |> 
  select(-c(dest_region))
```

```{r}
#| layout-ncol: 2
#| out-width: "50%"

gct_rs_f <- gct_rs |> 
  filter(care_purpose_cat == "Care") |> 
  mutate(survey_year = as.character(survey_year))

plot_facet_wrap(table = gct_rs_f, 
                facet = survey_year, 
                var1 = gender2, 
                var2 = gender2, 
                title = "Gender and Care Trips")

plot_facet_wrap(table = gct_rs_f, 
                facet = gender2, 
                var1 = survey_year, 
                var2 = survey_year, 
                title = "Gender and Care Trips",
                color_pal = psrc_colors$blues_inc) +
  geom_hline(yintercept = .50, linetype = "dashed", color = "red")

```
```{r}
gct_rs <- gct_rs |> 
  mutate(lower_per = percent(prop - prop_moe),
         upper_per = percent(prop + prop_moe))

gct_rs
```


