---
title: "Care Giving"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  source(here::here("Analysis", "care-giving", f))
})
)

```

## Trip Purpose

### PSRC Destination Purpose Assignment

Below are trips with their destination in the Region (King, Kitsap, Pierce, and Snohomish Counties) for survey years 2023 and 2025.

```{r}

# This section prints a table with weighted/unweighted x Home/Exclude Home x trip purpose

trip_purp_a <- create_trip_purpose_tbl(df_hts)

trip_purp_a |>
  gt() |>
  cols_align(
    align = "left",
    columns = dest_purpose_cat
  ) |>
  tab_spanner_delim(delim = ".") |> 
  tab_style(
    style = cell_fill(color = "lightyellow"),
    locations = cells_body(
      columns = dest_purpose_cat,
      rows = dest_purpose_cat %in% care_purpose_cat  # Matches values in the vector
    )
  ) |> 
  data_color(
    columns = contains("weighted.Exclude"),
    rows = dest_purpose_cat %in% care_purpose_cat,
    colors = "lightyellow"
  )


```

### Care Purpose

Care is defined as the following purposes: Escort, Shopping, Meal, Personal Business/Errand/Appointment. All `HOME` trips are excluded.

Care: `r care_purpose_cat`

Work: `r work_cat`

Other: `r other_cat`

School: `r school_cat`

```{r}

care_purp <- create_care_purpose_tbl(df_hts)

care_purp |> 
  gt() |> 
   cols_align(
    align = "left",
    columns = 1
  ) |>
  tab_spanner_delim(delim = ".") |> 
    tab_style(
    style = cell_fill(color = "lightyellow"),
    locations = cells_body(
      columns = care_purpose_cat,
      rows = care_purpose_cat == "Care"  # Matches values in the vector
    )
  ) |> 
  data_color(
    columns = matches("^weighted"),
    rows = care_purpose_cat == "Care",
    colors = "lightyellow"
  )

```
