---
title: "Care Giving Analysis"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}
source(here::here("Analysis", "care-giving", "data.R"))
source(here::here("Analysis", "care-giving", "functions.R"))

```

## Trip Purpose

### PSRC Destination Purpose Assignment

Below are trips with their destination in the Region (King, Kitsap, Pierce, and Snohomish Counties) for survey years 2023 and 2025

```{r}

# This section prints a table with weighted/unweighted x Home/Exclude Home x trip purpose

trip_purp_a <- create_trip_purpose_tbl(df_hts)

trip_purp_a |>
  gt() |>
  cols_align(
    align = "left",
    columns = dest_purpose_cat
  ) |>
  tab_spanner_delim(delim = ".")


```

### Care Purpose

Care is defined as the following purposes: Escort, Shopping, Meal, Personal Business/Errand/Appointment. All `HOME` trips are excluded.

Care: `r care_purpose_cat`

Work: `r work_cat`

Other: `r other_cat`

School: `r school_cat`

```{r}
rs_care_purp <- psrc_hts_stat(df_hts, 
                              analysis_unit = "trip", 
                              group_vars = c("dest_region", "care_purpose_cat"),
                              incl_na = FALSE) %>%
  mutate(type = "percent") |>  
  mutate(unweighted = percent(count/sum(count)),
         weighted = percent(prop),
         moe = percent(prop_moe),
         .by = "type")

rs_care_purp |> 
  select(all_of(c("care_purpose_cat","unweighted","weighted","moe"))) |> 
  gt() |> 
  cols_align(align = "left",
             columns = 1
  )
```

## Care Trips by Mode and Income

::: panel-tabset
```{r}
mode_share_income <- function(income, geog){
  geo_col <- case_when(geog == "Region" ~ "dest_region",
                       geog %in% c("King County", "Kitsap County", "Pierce County", "Snohomish County") ~ "dest_county"
                    )
  
  rs <- psrc_hts_stat(df_hts, 
                       analysis_unit = "trip", 
                       group_vars = c(geo_col, "care_purpose_cat", income, "mode_class_5"),
                       incl_na = FALSE) |> 
    rename(dest_loc = sym(geo_col)) |>
    filter(dest_loc == geog,
           care_purpose_cat == "Care") |> 
    mutate(type = "percent") |>
    mutate(unweighted = percent(count/sum(count)),
           weighted = percent(prop),
           moe = percent(prop_moe),
           .by = c("type", income))

  v <- c("dest_loc", "care_purpose_cat", income, "mode_class_5", "unweighted", "weighted", "moe")
  rs |> select(all_of(v))
}
```

### \$50,000

```{r}
mode_share_income(income = "income_50", geog = "Region") #|> 
  # gt() #|> 
  # cols_hide(c(dest_loc, care_purpose_cat)) |> 
  # cols_align(align = "left",
  #            columns = c(3,4)
  # )
```

### \$75,000

```{r}
mode_share_income("income_75", geog = "Region")
```

### \$100,000

```{r}
mode_share_income("income_100", geog = "Region")
```
:::
