---
title: "Mode and Income"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
params:
  errorbar: FALSE    
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  source(f)
  # source(here::here("Analysis", "care-giving", f))
})
)

source("../util.R")

```

## Trip Mode and Income for Care-giving Trips

The following charts and tables analyze the share of trip modes across two to three income brackets for *all care-giving trips*.

### Under \$50k and \$50k or More

Driving is the predominant mode of transportation for care trips. Those who earn under \$50,000 are using transit more than those who earn over \$50,000 across survey years.

```{r}
msi50bin_df <- mode_share_income_set_c(geog = "Region")$long |> 
  mutate(income_50 = factor(income_50, levels = c("Under $50,000", "$50,000 or more")))
```

::: panel-tabset
## Side-by-side

```{r}
#| fig-width: 6
#| fig-align: "center"

msi50bin_df |> 
  ggplot() +
  geom_col(aes(x = as.character(survey_year), y = prop, fill = mode_class_5)) +
  facet_wrap(~income_50, strip.position = "top") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  psrc_style() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        # strip.placement = "outside",
        # panel.spacing = unit(0, "cm"),
        legend.position = "bottom",
        strip.text = element_text(hjust = 0.5)
        # strip.text = element_text(face = "bold")
        ) +
  labs(x = NULL,
       y = NULL)

```

## Under \$50,000

```{r}
#| layout-ncol: 2
#| fig-width: 7
#| fig-align: "center"

msi50bin_df_u <- msi50bin_df |>
  filter(income_50 == "Under $50,000")

plot_facet_wrap(table = msi50bin_df_u, 
                facet = as.character(survey_year), 
                var1 = mode_class_5, 
                var2  = mode_class_5, 
                title = "Income by Primary Trip Mode", 
                color_pal = psrc_colors$pognbgy_5)+
  theme(axis.text.x = element_blank())

plot_facet_wrap(table = msi50bin_df_u, 
                var1 = as.character(survey_year), 
                var2 = as.character(survey_year),
                facet = mode_class_5, 
                title = "Income by Primary Trip Mode", 
                color_pal = psrc_colors$blues_inc)+
  theme(axis.text.x = element_blank())
```

## \$50,000 or More

```{r}
#| layout-ncol: 2
#| fig-width: 7
#| fig-align: "center"

msi50bin_df_m <- msi50bin_df |>
  filter(income_50 == "$50,000 or more")

plot_facet_wrap(table = msi50bin_df_m, 
                facet = as.character(survey_year), 
                var1 = mode_class_5, 
                var2  = mode_class_5, 
                title = "Income by Primary Trip Mode", 
                color_pal = psrc_colors$pognbgy_5) +
  theme(axis.text.x = element_blank())

plot_facet_wrap(table = msi50bin_df_m, 
                var1 = as.character(survey_year), 
                var2 = as.character(survey_year),
                facet = mode_class_5, 
                title = "Income by Primary Trip Mode", 
                color_pal = psrc_colors$blues_inc) +
  theme(axis.text.x = element_blank())
```
:::

```{r}

mode_share_income_set_c(geog = "Region")$wide|>
  gt() |>
  tab_spanner_delim(delim = ".") |>
  cols_hide(c(dest_loc, care_purpose_cat)) |>
  cols_align(align = "left",
             columns = c(3)
)
```

### Under \$50k, \$50k-\$100k, \$100k or more

```{r}
msi_df <- mode_share_income_set_b(geog = "Region")$long  |> 
  mutate(income_comp = factor(income_comp, levels = c("Under $50,000", "$50,000-$99,999", "$100,000 or more")))

```

```{r}
#| fig-width: 8
#| fig-align: center

msi_df |> 
  ggplot() +
  geom_col(aes(x = as.character(survey_year), y = prop, fill = mode_class_5)) +
  facet_wrap(~income_comp, strip.position = "bottom") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_5) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        strip.placement = "outside",
        panel.spacing = unit(0, "cm"),
        legend.position = "bottom",
        strip.text = element_text(face = "bold")) +
  labs(x = NULL,
       y = NULL)

```

```{r}

mode_share_income_set_b(geog = "Region")$wide|>
  gt() |>
  tab_spanner_delim(delim = ".") |>
  cols_hide(c(dest_loc, care_purpose_cat)) |>
  cols_align(align = "left",
             columns = c(3)
)
```

## Trip purpose and Income by Primary Trip Mode

The following charts and tables analyze the share of trip purposes by primary trip mode across income levels. The analysis is first presented using two income brackets (Under \$50k and \$50k or More), followed by a more detailed breakdown using three income brackets (Under \$50k, \$50k–\$100k, and \$100k or More).

### Under \$50k and \$50k or More

```{r}
mode_inc_b <- psrc_hts_stat(df_hts,
                            analysis_unit = "trip",
                            group_vars = c("dest_region", "income_50", "mode_class_5", "care_purpose_cat"),
                            incl_na = FALSE) |>
  mutate(prop_per = label_percent(accuracy = 0.1)(prop),
         moe_per = label_percent(accuracy = 0.1)(prop_moe))
```

<!-- ::: panel-tabset -->

#### Drive

```{r}
mode_inc_b_drive <- mode_inc_b |>
  filter(mode_class_5 == "Drive")
```

- Of trips made by those who make under \$50,000 and where driving was the primary trip mode, at least 50% were care-giving compared to other trip purposes. The shares for care-giving trips by driving are higher for those making under \$50,000 compared to those who make more.

- Of those who earn more than \$50,000 and where driving is the primary trip mode, there are more care-giving trips than other trip purposes.

- When comparing across income levels, there are more work trips by driving of those who earn more than \$50,000 than those who earn less than \$50,000.

```{r}
#| label: fig-mode-inc-b-drive-01
#| fig-cap: "Trip Purpose and Income by Primary Trip Mode (Drive)"
#| fig-width: 8
#| out-width: "75%"

plot_facet_wrap(table = mode_inc_b_drive,
                facet = income_50,
                var1 = care_purpose_cat,
                var2  = as.character(survey_year),
                title = NULL, #"Income by Primary Trip Mode (Drive)",
                color_pal = psrc_colors$pognbgy_5) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.7, size = 7))
```

```{r}
#| label: tbl-mode-inc-b-drive-01
#| fig-cap: "Trip Purpose and Income by Primary Trip Mode (Drive) Table"

knitr::kable(mode_inc_b_drive)
```

#### Transit

```{r}

mode_inc_b_trans <- mode_inc_b |>
  filter(mode_class_5 == "Transit")
```

```{r}
#| label: fig-mode-inc-b-transit-01
#| fig-cap: "Trip Purpose and Income by Primary Trip Mode (Transit)"
#| fig-width: 8
#| out-width: "75%"

plot_facet_wrap(table = mode_inc_b_trans,
                facet = income_50,
                var1 = care_purpose_cat,
                var2  = as.character(survey_year),
                title = NULL, #"Income by Primary Trip Mode (Transit)",
                color_pal = psrc_colors$pognbgy_5) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.7, size = 7))
```

```{r}
#| label: tbl-mode-inc-b-transit-01
#| fig-cap: "Trip Purpose and Income by Primary Trip Mode (Transit) Table"
#| 
mode_inc_b_trans

```

#### Walk

```{r}

mode_inc_b_walk <- mode_inc_b |>
  filter(mode_class_5 == "Walk")
```

Of trips made by those who earn \$50,000 or more and where walking is the primary trip mode, it is not clear whether there are more care-giving trips or social/recreational trips aside from 2021 where there is a high share of walking trips for social/recreational purposes.

```{r}
#| label: fig-mode-inc-b-walk-01
#| fig-cap: "Trip Purpose and Income by Primary Trip Mode (Walk)"
#| fig-width: 8
#| out-width: "75%"

plot_facet_wrap(table = mode_inc_b_walk,
                facet = income_50,
                var1 = care_purpose_cat,
                var2  = as.character(survey_year),
                title = NULL, #"Income by Primary Trip Mode (Walk)",
                color_pal = psrc_colors$pognbgy_5) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.7, size = 7))
```

```{r}
#| label: tbl-mode-inc-b-walk-01
#| fig-cap: "Trip Purpose and Income by Primary Trip Mode (Walk) Table"
#| 
mode_inc_b_walk
```

#### Bike/Micromobility

```{r}

mode_inc_b_bike <- mode_inc_b |>
  filter(mode_class_5 == "Bike/Micromobility")
```

```{r}
#| label: fig-mode-inc-b-bike-01
#| fig-cap: "Trip Purpose and Income by Primary Trip Mode (Bike)"
#| fig-width: 8
#| out-width: "75%"

plot_facet_wrap(table = mode_inc_b_bike,
                facet = income_50,
                var1 = care_purpose_cat,
                var2  = as.character(survey_year),
                title = NULL,#"Income by Primary Trip Mode (Bike)",
                color_pal = psrc_colors$pognbgy_5) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.7, size = 7))
```

```{r}
#| label: tbl-mode-inc-b-bike-01
#| fig-cap: "Trip Purpose and Income by Primary Trip Mode (Bike) Table"
mode_inc_b_bike
```

#### Other

```{r}

mode_inc_b_other <- mode_inc_b |>
  filter(mode_class_5 == "Other")
```

```{r}
#| label: fig-mode-inc-b-other-01
#| fig-cap: "Trip Purpose and Income by Primary Trip Mode (Other)"
#| fig-width: 8
#| out-width: "75%"

plot_facet_wrap(table = mode_inc_b_other,
                facet = income_50,
                var1 = care_purpose_cat,
                var2  = as.character(survey_year),
                title = NULL, #"Income by Primary Trip Mode (Other)",
                color_pal = psrc_colors$pognbgy_5) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.7, size = 7))
```

```{r}
#| label: tbl-mode-inc-b-other-01
#| fig-cap: "Trip Purpose and Income by Primary Trip Mode (Other) Table"
mode_inc_b_other
```

<!-- ::: -->

### Under \$50k, \$50k-\$100k, \$100k or more

The following charts and tables analyze the share of care trips by primary trip mode across three income brackets: Under \$50k, \$50k–\$100k, and \$100k or More.

```{r}
mode_inc_b_3 <- psrc_hts_stat(df_hts,
                            analysis_unit = "trip",
                            group_vars = c("dest_region", "income_comp", "mode_class_5", "care_purpose_cat"),
                            incl_na = FALSE) |>
  mutate(prop_per = label_percent(accuracy = 0.1)(prop),
         moe_per = label_percent(accuracy = 0.1)(prop_moe))|> 
  mutate(income_comp = factor(income_comp, levels = c("Under $50,000", "$50,000-$99,999", "$100,000 or more")))

```

```{r}
create_mode_inc_chart <- function(table, mode) {
  table |> 
    ggplot(aes(x = as.character(survey_year), y = prop, fill = as.character(survey_year))) +
    geom_col(width=0.7) +
    geom_text(aes(label=scales::percent(prop, accuracy = 1)),
              vjust = -0.3, 
              position = position_dodge(width = 0.9),
              check_overlap = TRUE
              # size = 8/.pt
              ) +
    errorbar_switch(geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe), width = 0.1), 
                    on=params$errorbar) +
    # geom_linerange(aes(x = as.character(survey_year), ymin = prop - prop_moe, ymax = prop + prop_moe),
    #                orientation = "x",
    #                position = position_dodge(width = 0.9)) +
    facet_wrap(~income_comp, strip.position = "top") +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(values = psrc_colors$oranges_dec) +
    psrc_style() +
    theme(panel.grid.minor = element_blank(),
          panel.grid.major = element_blank(),
          axis.text.y = element_blank(),
          panel.grid.major.y = element_blank(),
          strip.placement = "outside",
          panel.spacing = unit(0, "cm"),
          legend.position = "none",
          strip.text = element_text(face = "bold")) +
    labs(x = NULL,
         y = NULL,
         title = NULL #paste(mode, "by Income and Trip Purpose")
         )
}
```

#### Drive

```{r}
#| label: fig-mode-inc-b3-drive-02
#| fig-cap: "Drive Trip Mode by Income and Care Trip Purpose"
#| fig-width: 8
#| out-width: "70%"

mode_type <- "Drive"

mode_inc_b_3 |> 
  filter(care_purpose_cat == "Care" & mode_class_5 == mode_type) |> 
  create_mode_inc_chart(mode = mode_type)


```

```{r}
#| label: tbl-mode-inc-b3-drive-02
#| fig-cap: "Drive Trip Mode by Income and Care Trip Purpose Table"
#| 
mode_inc_b_3 |> 
  filter(care_purpose_cat == "Care" & mode_class_5 == mode_type)
```

#### Transit

```{r}
#| label: fig-mode-inc-b3-transit-02
#| fig-cap: "Transit Trip Mode by Income and Care Trip Purpose"
#| fig-width: 8
#| out-width: "70%"

mode_type <- "Transit"

mode_inc_b_3 |> 
  filter(care_purpose_cat == "Care" &mode_class_5 == mode_type) |> 
   create_mode_inc_chart(mode = mode_type)
```

```{r}
#| label: tbl-mode-inc-b3-transit-02
#| fig-cap: "Transit Trip Mode by Income and Care Trip Purpose Table"
mode_inc_b_3 |> 
  filter(care_purpose_cat == "Care" &mode_class_5 == mode_type)
```

#### Walk

```{r}
#| label: fig-mode-inc-b3-walk-02
#| fig-cap: "Walk Trip Mode by Income and CareTrip Purpose"
#| fig-width: 8
#| out-width: "70%"

mode_type <- "Walk"

mode_inc_b_3 |> 
  filter(care_purpose_cat == "Care" &mode_class_5 == mode_type) |> 
   create_mode_inc_chart(mode = mode_type)

```

```{r}
#| label: tbl-mode-inc-b3-walk-02
#| fig-cap: "Walk Trip Mode by Income and Care Trip Purpose Table"
mode_inc_b_3 |> 
  filter(care_purpose_cat == "Care" &mode_class_5 == mode_type)
```

#### Bike/Micromobility

```{r}
#| label: fig-mode-inc-b3-bike-02
#| fig-cap: "Bike/Micromobility Trip Mode by Income and Care Trip Purpose"
#| fig-width: 8
#| out-width: "70%"
#| fig-align: center

mode_type <- "Bike/Micromobility"

mode_inc_b_3 |> 
  filter(care_purpose_cat == "Care" &mode_class_5 == mode_type) |>
   create_mode_inc_chart(mode = mode_type)

```

```{r}
#| label: tbl-mode-inc-b3-bike-02
#| fig-cap: "Bike/Micromobility Trip Mode by Income and Care Trip Purpose Table"
mode_inc_b_3 |> 
  filter(care_purpose_cat == "Care" &mode_class_5 == mode_type) 
```

#### Other

```{r}
#| label: fig-mode-inc-b3-other-02
#| fig-cap: "Other Trip Mode by Income and Care Trip Purpose"
#| fig-width: 8
#| out-width: "70%"

mode_type <- "Other"

mode_inc_b_3 |> 
  filter(care_purpose_cat == "Care" &mode_class_5 == mode_type) |> 
     create_mode_inc_chart(mode = mode_type)

```

```{r}
#| label: tbl-mode-inc-b3-other-02
#| fig-cap: "Other Trip Mode by Income and Care Trip Purpose Table"
#| 
mode_inc_b_3 |> 
  filter(care_purpose_cat == "Care" &mode_class_5 == mode_type) 
```
