---
title: "Household Structure"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
params:
  errorbar: FALSE    
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
  lapply(files, function(f) {
    # source(here::here("Analysis", "care-giving", f))
    source(f)
  })
)
source("../util.R")
```

## Children and Elders Definition

Elder is defined as a person 65 years or over.

Children is based on var `numchildren`.

## Overall Household Composition

- Generally there are predominately more households without children or elders than households with both children and elders.

```{r}
hhs_comp <- psrc_hts_stat(df_hts,
                          analysis_unit = "hh",
                          group_vars = "hh_composition",
                          incl_na = FALSE)
```

```{r}

create_hh_comp_chart <- function(table, x_var, y_var, fill_var, facet, colors) {
  table |> 
  ggplot(aes(y = {{y_var}}, x = {{x_var}}, fill = {{fill_var}})) +
  geom_col(position = "dodge") +
     geom_text(aes(label=scales::percent(prop, accuracy = 1)),
              vjust = -0.3, 
              position = position_dodge(width = 0.9),
              check_overlap = TRUE
              # size = 8/.pt
              ) +
  errorbar_switch(geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe), width = 0.1), 
                  on=params$errorbar) +
  facet_wrap(vars({{facet}}), ncol = 5, nrow = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = colors) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.position = "bottom",
        legend.text = element_text(size = 8), # Shrink the legend text and title
        legend.title = element_text(size = 9),
        legend.key.size = unit(0.4, "cm") # Shrink the legend key boxes (both width and height)
  ) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = NULL)
}
```

```{r}
#| label: fig-hh-comp-01
#| fig-cap: "Household Composition"
#| fig-width: 8
#| out-width: "70%"

create_hh_comp_chart(table = hhs_comp, 
                     x_var=hh_composition, 
                     y_var=prop, 
                     fill_var=hh_composition,
                     facet=survey_year,
                     colors=psrc_colors$pognbgy_10)
```

```{r}
#| label: fig-hh-comp-02
#| fig-cap: "Household Composition"
#| fig-width: 8
#| out-width: "70%"

hhs_comp2 <- hhs_comp |> 
  mutate(hh_composition_label = str_trim(str_replace_all(hh_composition, ",", "\n")),
         hh_composition_label = str_to_title(hh_composition_label))
  
create_hh_comp_chart(table = hhs_comp2, 
                     x_var=as.character(survey_year), 
                     y_var=prop, 
                     fill_var=as.character(survey_year), 
                     facet = hh_composition_label,
                     colors=psrc_colors$blues_dec)
```

```{r}
#| label: tbl-hh-comp

get_table_format(hhs_comp)
```

```{r, eval=FALSE}
test <- psrc_hts_stat(df_hts,
                      analysis_unit = "hh",
                      group_vars = c("care_trip_taken", "hh_composition"),
                      incl_na = FALSE)

test_sum_prop <- test |> 
  group_by(survey_year, care_trip_taken) |> 
  summarize(sum(prop))

test_sum_c_e <- test |> 
  group_by(survey_year) |> 
  summarize(sum(count), sum(est))
```

```{r, eval=FALSE}
test_sum_prop
```

```{r, eval=FALSE}
test_sum_c_e
```

```{r, eval=FALSE}
hhc <- psrc_hts_stat(df_hts,
                     analysis_unit = "hh",
                     group_vars = c("hh_composition"),
                     incl_na = FALSE) 
hhc_sum <- hhc |> 
  mutate(survey_year = as.character(survey_year)) |>
  group_by(survey_year) |> 
  summarize(sum_prop = sum(prop), sum_est = sum(est), sum_count = sum(count))

```

## Care Trips Taken

### Household Composition by Trip Purpose

The following charts and tables show the distribution of households and their composition by trip purpose -- specifically care-giving.

- Where care trips were taken, more were made by households with no children or elders followed by households that have children but no elders.

- There are few households that are multi-generational with both children and elders of households that took care trips.

```{r}
hhstr_a <- psrc_hts_stat(df_hts,
                         analysis_unit = "hh",
                         group_vars = c("care_trip_taken", "hh_composition"),
                         incl_na = FALSE) #|>
  # mutate(prop_per = percent(prop),
  #        prop_moe_per = percent(prop_moe))

```

```{r}
#| label: fig-trip-purp-hh-comp-01
#| fig-cap: "Share of Households with Care Trips by Household Composition"
#| fig-width: 8
#| out-width: "70%"

trip_purp_hhstr_a1 <- hhstr_a |>
  filter(care_trip_taken == 1)

create_hh_comp_chart(table = trip_purp_hhstr_a1, 
                     x_var=hh_composition, 
                     y_var=prop, 
                     fill_var=hh_composition,
                     facet=survey_year,
                     colors=psrc_colors$pognbgy_10)
```

```{r}
#| label: fig-trip-purp-hh-comp-02
#| fig-cap: "Share of Households with Care Trips by Household Composition"
#| fig-width: 8
#| out-width: "70%"

trip_purp_hhstr_a2 <- hhstr_a |>
  filter(care_trip_taken == 1) |> 
  mutate(hh_composition_label = str_trim(str_replace_all(hh_composition, ",", "\n")),
         hh_composition_label = str_to_title(hh_composition_label))

create_hh_comp_chart(table = trip_purp_hhstr_a2, 
                     x_var=as.character(survey_year), 
                     y_var=prop, 
                     fill_var=as.character(survey_year),
                     facet=hh_composition_label,
                     colors=psrc_colors$blues_dec)
```

```{r}
#| label: tbl-trip-purp-hh-comp-01

get_table_format(hhstr_a)
```

### Trip Purpose by Household Composition

The following charts and tables show the distribution of trip purposes by the different types of household composition.

- Of households where there are either elders or children present, or none at all, there are more care-giving trips made. Especially of households where there are children and no elders, there are fewer non-care-giving trips compared to other household types.

- There are few households with both elders and children. Their results have very large margins of error and cannot be determined.

```{r}
hhstr_b <- psrc_hts_stat(df_hts,
                         analysis_unit = "hh",
                         group_vars = c("hh_composition","care_trip_taken"),
                         incl_na = FALSE) #|>
  # mutate(prop_per = percent(prop),
  #        prop_moe_per = percent(prop_moe))
```

```{r}
#| label: fig-hhcomp-trip-purp
#| fig-cap: "Share of Households by Household Composition and Trip Purpose"
#| fig-width: 8
#| out-width: "70%"

create_hh_comp_chart(table = hhstr_b, 
                     x_var=as.character(survey_year), 
                     y_var=prop, 
                     fill_var=care_trip_taken,
                     facet=hh_composition,
                     colors=psrc_colors$pognbgy_10) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.7)) +
  scale_fill_manual(values = psrc_colors$pognbgy_10, labels = c("1" = "Care Trip", "0" = "Non-care Trip"))

```

```{r}
#| label: tbl-hhcomp-trip-purp
#| 
get_table_format(hhstr_b)
```
