---
title: "Household Structure"
editor: visual
format: 
  html:
    mainfont: Poppins
    toc: true
    warning: false
    echo: false
    fig-width: 4.5
    fig-asp: .618
---

```{r setup}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
library(scales)
library(gt)
install_psrc_fonts()
```

```{r}

files <- c("data.R", "functions.R")

invisible(
lapply(files, function(f) {
  # source(here::here("Analysis", "care-giving", f))
  source(f)
})
)
```

## Children and Elders Definition

Elder is defined as a person 65 years or over.

Children is based on var `numchildren`.

## Overall Household Composition

- Generally there are predominately more households without children or elders that households with both children and elders.

- In 2017/19 and 2021, there has been relatively more households with only children and no elders than only elders. In 2023 and 2025, it is undetermined if there are more or less.

```{r}
hhs_comp <- psrc_hts_stat(df_hts,
                     analysis_unit = "hh",
                     group_vars = "hh_composition",
                     incl_na = FALSE)
```

```{r}
#| layout-ncol: 2
#| out-width: "50%"

hhs_comp |> 
ggplot(aes(y = prop, x = hh_composition, fill = hh_composition)) +
  geom_col(position = "dodge") +
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)) +
  facet_wrap(~survey_year, ncol = 5, nrow = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_10) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.7),
        legend.position = "none") +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Household Composition")

hhs_comp |> 
mutate(hh_composition_label = str_trim(str_replace_all(hh_composition, ",", "\n")),
         hh_composition_label = str_to_title(hh_composition_label)) |> 
  ggplot(aes(y = prop, x = as.character(survey_year), fill = as.character(survey_year))) +
  geom_col(position = "dodge") +
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)) +
  facet_wrap(~hh_composition_label, ncol = 5, nrow = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$blues_inc) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.7),
        legend.position = "none") +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Household Composition")
```

```{r}
hhs_comp
```

```{r, eval=FALSE}
test <- psrc_hts_stat(df_hts,
                      analysis_unit = "hh",
                      group_vars = c("care_trip_taken", "hh_composition"),
                      incl_na = FALSE)

test_sum_prop <- test |> 
  group_by(survey_year, care_trip_taken) |> 
  summarize(sum(prop))

test_sum_c_e <- test |> 
  group_by(survey_year) |> 
  summarize(sum(count), sum(est))
```

```{r, eval=FALSE}
test_sum_prop
```

```{r, eval=FALSE}
test_sum_c_e
```

```{r, eval=FALSE}
hhc <- psrc_hts_stat(df_hts,
                     analysis_unit = "hh",
                     group_vars = c("hh_composition"),
                     incl_na = FALSE) 
hhc_sum <- hhc |> 
   mutate(survey_year = as.character(survey_year)) |>
  group_by(survey_year) |> 
  summarize(sum_prop = sum(prop), sum_est = sum(est), sum_count = sum(count))

```

### Care Trips Taken

#### Household Composition

- Where care trips were taken, there is a higher share of households with no children or elders followed by households that have children but no elders.

- There are few households that are multi-generational with both children and elders of households that took care trips.

```{r}
#| layout-ncol: 2
#| out-width: "50%"

create_hhstr_tbl(geog = "Region")$long |>
  filter(care_trip_taken == 1) |> 
  ggplot(aes(y = prop, x = hh_composition, fill = hh_composition)) +
  geom_col(position = "dodge") +
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)) +
  facet_wrap(~survey_year, ncol = 5, nrow = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$pognbgy_10) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.7),
        legend.position = "none") +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Households that Made Care Trips")

create_hhstr_tbl(geog = "Region")$long |>
  filter(care_trip_taken == 1) |> 
  mutate(hh_composition_label = str_trim(str_replace_all(hh_composition, ",", "\n")),
         hh_composition_label = str_to_title(hh_composition_label)) |> 
  ggplot(aes(y = prop, x = as.character(survey_year), fill = as.character(survey_year))) +
  geom_col(position = "dodge") +
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)) +
  facet_wrap(~hh_composition_label, ncol = 5, nrow = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$blues_inc) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.7),
        legend.position = "none") +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Households that Made Care Trips")
```

```{r, eval=FALSE}
#| out-width: "100%"
#| fig-align: "center"

create_hhstr_tbl(geog = "Region")$long |>
  filter(care_trip_taken == 1) |> 
  ggplot(aes(y = prop, x = as.character(survey_year), fill = as.character(survey_year))) +
  geom_col(position = "dodge") +
  geom_linerange(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                 orientation = "x",
                 position = position_dodge(width = 0.9)) +
  facet_wrap(~hh_composition, ncol = 5, nrow = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = psrc_colors$blues_inc) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.7),
        legend.position = "none") +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Households that Made Care Trips")
```

```{r}

create_hhstr_tbl(geog = "Region")$long

# create_hhstr_tbl_b(geog = "Region")$wide |> 
#   filter(care_trip_taken == 1) |> 
#   gt() |> 
#   tab_spanner_delim(delim = ".") |>
#   cols_align(align = "left",
#              columns = c(1,2))

```
