---
title: "2023 HTS data: KCM/ MIT data check"
date: today
format: 
  html:
    mainfont: Poppins
    theme: [cosmo, ../custom.scss]
    toc: true
    df-print: kable
    fig-asp: .7
    fig-width: 4
    warning: false
    echo: false
---

- I think the slides are using weighted shares

```{r}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcplot)
install_psrc_fonts()

percent_format <- function(.col){
  scales::percent(.col, accuracy=0.01)
}


# read hts data
vars <- c("dest_county","dest_purpose","dest_purpose_cat","dest_purpose_cat_5","mode_class","mode_class_5",
          "age","license","gender",
          "hhincome_broad","home_county")
hts_data <- get_psrc_hts(survey_vars = vars,survey_years = 2023)
```
  
```{r}
care_purpose_cat <- c(
  "Escort", "Shopping", "Meal", "Personal Business/Errand/Appointment"
)
work_cat <- c("Work","Work-related")
other_cat <- c("Other","Change mode","Overnight")
school_cat <- c("School","School-related")
```


```{r}
trip_mutate <- hts_data[["trip"]] %>%
  mutate(# remove home purpose: the pie chart doesn't include HOME trips
         dest_purpose_cat_no_home = case_when(dest_purpose_cat == "Home"~ NA,
                                              TRUE~dest_purpose_cat),
         # use dest_purpose: too complicated skip for now
         # care_purpose = case_when(dest_purpose %in% care_purpose~"Care",
         #                          TRUE~dest_purpose_cat),
         # care_purpose_no_home = case_when(dest_purpose_cat == "Home"~ NA,
         #                                  TRUE~care_purpose),
         care_purpose_cat = factor(
           case_when(dest_purpose_cat %in% care_purpose_cat~"Care",
                     dest_purpose_cat %in% work_cat~"Work",
                     dest_purpose_cat %in% other_cat~"Other",
                     dest_purpose_cat %in% school_cat~"School",
                     dest_purpose_cat == "Home"~ NA,
                     TRUE~dest_purpose_cat_5),
           levels = c("Care","Work","School","Social/Recreation","Other")))

low_50 <- c("Under $25,000","$25,000-$49,999")
high_50 <- c("$50,000-$74,999","$75,000-$99,999","$100,000-$199,999","$200,000 or more")
low_75 <- c("Under $25,000","$25,000-$49,999","$50,000-$74,999")
high_75 <- c("$75,000-$99,999","$100,000-$199,999","$200,000 or more")
low_100 <- c("Under $25,000","$25,000-$49,999","$50,000-$74,999","$75,000-$99,999")
high_100 <- c("$100,000-$199,999","$200,000 or more")

hh_mutate <- hts_data[["hh"]] %>%
  mutate(
    income_50 = factor(
      case_when(hhincome_broad %in% low_50~"Under $50,000",
                hhincome_broad %in% high_50~"$50,000 or more",
                TRUE~NA),
      levels=c("Under $50,000","$50,000 or more")),
    income_75 = factor(
      case_when(hhincome_broad %in% low_75~"Under $75,000",
                hhincome_broad %in% high_75~"$75,000 or more",
                TRUE~NA),
      levels=c("Under $75,000","$75,000 or more")),
    income_100 = factor(
      case_when(hhincome_broad %in% low_100~"Under $100,000",
                hhincome_broad %in% high_100~"$100,000 or more",
                TRUE~NA),
      levels=c("Under $100,000","$100,000 or more"))
  )

person_mutate <- hts_data[["person"]] %>%
  mutate(
    gender2 = factor(
      case_when(gender == "Boy/Man (cisgender or transgender)"~"Male",
                gender == "Girl/Woman (cisgender or transgender)"~"Female",
                TRUE~NA),
      levels=c("Male","Female"))
  )

# final HTS data
df_hts <- hts_data
df_hts[["trip"]] <- trip_mutate
df_hts[["hh"]] <- hh_mutate
df_hts[["person"]] <- person_mutate
```

## trip purpose

### PSRC destination purpose assignment

- only include trips with destination in King County
  
```{r}
rs1 <- psrc_hts_stat(df_hts, 
                     analysis_unit="trip", 
                     group_vars=c("dest_county","dest_purpose_cat"),
                     incl_na = FALSE) %>%
  mutate(type="Include HOME")

rs1_no_home <- psrc_hts_stat(df_hts, 
                     analysis_unit="trip", 
                     group_vars=c("dest_county","dest_purpose_cat_no_home"),
                     incl_na = FALSE) %>%
  mutate(type="Exclude HOME") %>%
  rename(dest_purpose_cat = dest_purpose_cat_no_home)

rs <- rs1 %>% 
  add_row(rs1_no_home) %>% 
  filter(dest_county == "King County")

rs_kc <- rs %>% filter(dest_county == "King County") %>%
  mutate(unweighted = percent_format(count/sum(count)),
         weighted = percent_format(prop),
         .by="type")

per_group <- c("survey_year","dest_county","dest_purpose_cat")
per_unweighted <- rs_kc %>%
  pivot_wider(id_cols = per_group, 
              names_from = "type", 
              values_from = "unweighted",
              names_prefix = "unweighted- ")
per_weighted <- rs_kc %>%
  pivot_wider(id_cols = per_group, 
              names_from = "type", 
              values_from = "weighted",
              names_prefix = "weighted- ")

per_unweighted %>%
  left_join(per_weighted, by=per_group)
```

### care purpose

- care purpose assignment
  - all HOME trips are excluded
- work and social trip share are close, but care share is higher than expected
  - (suspecting school and overnight trips are in care)
  
```{r}
#| echo: TRUE

care_purpose_cat <- c(
  "Escort", "Shopping", "Meal", "Personal Business/Errand/Appointment"
)
work_cat <- c("Work","Work-related")
other_cat <- c("Other","Change mode","Overnight")
school_cat <- c("School","School-related")
```

```{r}
rs_kc <- psrc_hts_stat(df_hts, 
                     analysis_unit="trip", 
                     group_vars=c("dest_county","care_purpose_cat"),
                     incl_na = FALSE) %>%
  mutate(type="percent") %>% 
  filter(dest_county == "King County") %>%
  mutate(unweighted = percent_format(count/sum(count)),
         weighted = percent_format(prop),
         moe = percent_format(prop_moe),
         .by="type")

rs_kc %>% select(all_of(c("dest_county","care_purpose_cat","unweighted","weighted","moe")))
```

## care trips in king county by mode and income

- dividing lower/higher income with $50,000 is closest to pie charts

:::{.panel-tabset}

### $50,000

```{r}
mode_share_income <- function(income){
  rs_kc <- psrc_hts_stat(df_hts, 
                       analysis_unit="trip", 
                       group_vars=c("dest_county","care_purpose_cat",income, "mode_class_5"),
                       incl_na = FALSE) %>%
    filter(dest_county == "King County", care_purpose_cat == "Care") %>%
    mutate(type="percent") %>%
    mutate(unweighted = percent_format(count/sum(count)),
           weighted = percent_format(prop),
           moe = percent_format(prop_moe),
           .by=c("type",income))
  
  rs_kc %>% select(all_of(c("dest_county","care_purpose_cat",income,"mode_class_5","unweighted","weighted","moe")))
}

mode_share_income("income_50")
```

### $75,000

```{r}
mode_share_income("income_75")
```

### $100,000

```{r}
mode_share_income("income_100")
```

:::

## license by age

- only king county residents are included
- weighted numbers are closer

```{r}
rs1 <- psrc_hts_stat(df_hts, 
                     analysis_unit="person", 
                     group_vars=c("home_county","age","license"),
                     incl_na = FALSE) %>%
  filter(home_county == "King County") %>%
  mutate(unweighted = percent_format(count/sum(count)),
         weighted = percent_format(prop),
         moe = percent_format(prop_moe),
         .by="age") %>%
  filter(license == "No, does not drive")

rs1 %>% select(all_of(c("home_county","age","unweighted","weighted","moe")))
```


## license by age and gender

- only king county residents are included
- weighted numbers are closer

```{r}
rs1 <- psrc_hts_stat(df_hts, 
                     analysis_unit="person", 
                     group_vars=c("home_county","age","gender2","license"),
                     incl_na = FALSE) %>%
  filter(home_county == "King County") %>%
  mutate(unweighted = percent_format(count/sum(count)),
         weighted = percent_format(prop),
         moe = percent_format(prop_moe),
         .by=c("age","gender2")) %>%
  filter(license == "No, does not drive")

per_group <- c("home_county","age")
per_unweighted <- rs1 %>%
  pivot_wider(id_cols = per_group, 
              names_from = "gender2", 
              values_from = "unweighted",
              names_prefix = "unweighted- ")
per_weighted <- rs1 %>%
  pivot_wider(id_cols = per_group, 
              names_from = "gender2", 
              values_from = "weighted",
              names_prefix = "weighted- ")

per_unweighted %>%
  left_join(per_weighted, by=per_group)

```

