---
title: "model shopping trips"
output:
  html_document:
    df_print: paged
---



```{r}
library(psrc.travelsurvey)
library(dplyr)
library(stringr)
library(ggplot2)
library(psrcplot)
library(forcats)
library(odbc)
library(DBI)
library(tidyr)
library(tidycensus)
#install.packages("psrcslides")
#library(psrcslides)
library(officer)
library(DBI)
library(foreign)
library(ggplot2)
library(MASS)
library(Hmisc)
library(reshape2)
library(jtools)
#install.packages("lessR")
library(lessR)
install_psrc_fonts()
```
Read in the data from Elmer

```{r}
 elmer_connect<-DBI::dbConnect(odbc::odbc(),
                 driver = "ODBC Driver 17 for SQL Server",
                 server = "AWS-PROD-SQL\\Sockeye",
                 database = "Elmer",
                 trusted_connection = "yes",
                 port = 1433)
  


```

We are getting household variables and trip variables to understand how
1. household characteristics relate to getting deliveries
2. trip making characteristics relate to getting deliveries
3. additional characteristics for delivery determination
```{r}
trip_vars<-c('mode_1', 'mode_simple','dest_purpose_cat', 'origin_purpose_cat', 'trip_path_distance', 'household_id', 'age_category', 'daynum', 'sample_source', 'hhgroup')
household_vars<-c('vehicle_count',  "hhincome_broad",  'hhincome_detailed', 'final_home_is_rgc', 'hhsize', 'household_id', 'lifecycle', 'sample_source', 'hhgroup', 'final_home_tract10', 'hh_race_apoc', 'numworkers', 'numchildren', 'numadults', 'broadband', 'mobile_device', 'res_factors_transit', 'res_factors_walk', 'rent_own')
```

Get all three years of data and bind it together.
```{r}
trip_data_17<- get_hhts("2017", "t", vars=trip_vars)%>%filter(age_category!='Under 18 years' & hhgroup=='rSurvey')%>%dplyr::select(-trip_weight_2017)

trip_data_19<- get_hhts("2019", "t", vars=trip_vars)%>%filter(age_category!='Under 18 years' & hhgroup=='rSurvey')%>%dplyr::select(-trip_weight_2019)

trip_data_21<- get_hhts("2021", "t", vars=trip_vars)%>%filter(age_category!='Under 18 years' & sample_source=='ABS')%>%dplyr::select(-c(trip_adult_weight_2021, trip_respondent_weight_2021))

trip_data_17_19_21<-rbind(trip_data_17, trip_data_19, trip_data_21)

household_data_17<- get_hhts("2017", "h", vars=household_vars)%>%filter( hhgroup=='rSurvey')%>%dplyr::select(-hh_weight_2017)
household_data_19<- get_hhts("2019", "h", vars=household_vars)%>%filter( hhgroup=='rSurvey')%>%dplyr::select(-hh_weight_2019)
household_data_21<- get_hhts("2021", "h", vars=household_vars)%>%filter(sample_source=='ABS')%>%dplyr::select(-hh_weight_2021)
household_data_17_19_21<-rbind(household_data_17, household_data_19, household_data_21)

#is.data.frame(household_data_17_19_21$broadband)

#library(dplyr)
household_data_17_19_21 %>% mutate_if(is.factor, as.character) -> household_data_17_19_21
#unique(household_data_17_19_21$hhincome_broad)
#unique(household_data_17_19_21$hhincome_detailed)
```


a function to group trips by purpose to understand how the number of household trips by a purpose relates to the propensity to receive deliveries
```{r}
trip_grouping <-function(trip_data){

trip_data<-trip_data%>%
  mutate(simple_purpose=ifelse(dest_purpose_cat=='Home', origin_purpose_cat, dest_purpose_cat))%>%
  mutate(simple_purpose=case_when(simple_purpose=='Work'~ 'Work/School',
                                  simple_purpose=='School'~ 'Work/School',
                                  simple_purpose=='Work-related'~ 'Work/School',
                                  simple_purpose=='Shop'~ 'Shop',
                                  simple_purpose=='Escort'~ 'Errands',
                                  simple_purpose=='Errand/Other'~ 'Errands',
                                  simple_purpose=='Change mode'~ 'Errands',
                                  simple_purpose=='Social/Recreation' ~ 'Social/Recreation',
                                  simple_purpose=='Meal' ~ 'Meal',
                                  simple_purpose=='Home' ~ 'Errands',
                                  is.na(simple_purpose) ~ 'Errands',
                                  TRUE ~ simple_purpose))

trip_data<-trip_data%>%filter(simple_purpose=='Shop')

trip_data_hhid<-trip_data%>% group_by(household_id,daynum)%>%tally()

}

```


```{r}
shop_trips_per_hh_day<-trip_grouping(trip_data_17_19_21)

```

A function for grouping household characteristics
```{r}

household_grouping<-function(household_data){
household_data<-household_data%>%
  mutate(NoVehicles=ifelse(vehicle_count=='0 (no vehicles)', 'No Vehicles', "Has Vehicles"))%>%
  mutate(hhincome_100= case_when(
    hhincome_broad=="Under $25,000" ~ "Under $75,000",
    hhincome_broad=="$25,000-$49,999" ~ "Under $75,000",
    hhincome_broad=="$50,000-$74,999" ~ "Under $75,000",
    hhincome_broad=='$75,000-$99,999' ~ '$75,000 or more',
    hhincome_broad=='$100,000 or more' ~ '$75,000 or more',
    hhincome_broad=='$200,000 or more' ~ '$75,000 or more',
                                 TRUE~hhincome_broad))%>%
  mutate(hhincome_detailed2 = case_when(hhincome_detailed == "Under $10,000" |
                                         hhincome_detailed == "$10,000-$24,999" |
                                         hhincome_detailed == "$25,000-$34,999" |
                                         hhincome_detailed == "$35,000-$49,999" ~ "Under $50,000",
                                          TRUE~ hhincome_detailed))%>%
  mutate(hhsize= case_when(hhsize == "1 person" ~ '1 person', 
                                  hhsize == "2 people"  ~ '2 people', 
                                  hhsize == "3 people" ~ '3 people',
                                  hhsize == "4 people" | 
                                  hhsize == "5 people" | 
                                  hhsize == "6 people" | 
                                  hhsize == "7 people" |
                                  hhsize == "8 people" | 
                                  hhsize == "12 people" ~ "4+ people"))%>%
  mutate(lifecycle2= case_when(lifecycle == "Household size > 1, Householder age 65+" | 
                                  lifecycle == "Household size = 1, Householder age 65+" ~ '65 years or older', 
                                   lifecycle == "Household size > 1, Householder age 35 - 64" |
                                   lifecycle == "Household size = 1, Householder age 35 - 64"  ~ '35-64, no kids',
                                   lifecycle == "Household size > 1, Householder under age 35" | 
                                   lifecycle == "Household size = 1, Householder under age 35" ~ 'Under 35 years, no kids',
                                   lifecycle == "Household includes children age 5-17" | 
                                   lifecycle == "Household includes children under 5" ~ 'Household has kids')) %>%
  mutate(broadband2=case_when(broadband == 'Broadband service is available where I live, but I DO NOT have broadband service at my residence'| 
                               broadband == 'Broadband service is not available where I live' | 
                               broadband == 'Don’t know' | 
                               broadband == "NA" |
                               broadband == 'No' ~ 'No',
                               broadband == 'Yes, my current residence has broadband service' | 
                               broadband == 'Yes' ~ 'Yes'))%>%
           mutate(rent_own= case_when(rent_own == 'Own/paying mortgage' ~ 'Own',
                                      rent_own == 'Rent' ~ 'Rent'))

household_data$hhincome_100_f=factor(household_data$hhincome_100,levels=c("Under $75,000","$75,000 or more"))

#household_data$hhincome_detailed = factor(household_data$hhincome_detailed, 
 #                                         levels = c("$Under $50,000",
  #                                        "$50,000-$74,999", 
   #                                       "$75,000-$99,999", 
    #                                      "$100,000-$149,999",
     #                                     "$150,000-$199,999",
      #                                    "$200,000-$249,999",
       #                                   "$250,000 or more"))

household_data$hhincome_broad = factor(household_data$hhincome_broad, 
                                          levels = c("$Under $25,000",
                                                     "$25,000-$49,999",
                                                     "$50,000-$74,999",
                                                     "$75,000-$99,999",
                                                     "$100,000 or more",
                                                     "$200,000 or more"))

household_data$lifecycle2 =factor(household_data$lifecycle2, levels= c('Household has kids', 'Under 35 years, no kids', '35-64, no kids', '65 years or older' ))

household_data
                                  
}
```

```{r}
household_data_17_19_21<-household_grouping(household_data_17_19_21)

```


Joining households to the number of shopping trips they made

```{r}
household_data_17_19_21_shop<-full_join(household_data_17_19_21, shop_trips_per_hh_day, by='household_id')%>%mutate_all(~replace(., is.na(.), 0)) 

household_data_17_19_21_shop%>%group_by(survey)%>%summarise(mean=mean(n,na.rm=TRUE))
```
```{r}
household_data_17_19_21_shop<- household_data_17_19_21_shop%>%mutate(shop_trips=ifelse(n>0, 1, 0))
```

```{r}
ggplot(household_data_17_19_21_shop, aes(x=shop_trips, fill=survey))+geom_histogram(stat='count')


```

This set of code is for summarizing whether a household received a delivery or not.
```{r}


survey_a <- list(survey = '2017_2019', label = '2017/2019')
survey_b <- list(survey = '2021', label = '2021')
survey_c <- list(survey = '2017', label = '2017')
survey_d <- list(survey = '2019', label = '2019')

# look at variable names and which years they were collected/documented

# hhts_varsearch("delivery")
# hhts_varsearch("traveldate")
# hhts_varsearch("race")
# hhts_varsearch('income')
# hhts_varsearch('age')
# hhts_varsearch('county')
# hhts_varsearch('home')
# hhts_varsearch('race')



# create variables that would like to group by for analysis of deliveries by household

delivery_type <- c("household_id", "delivery_food_freq", "delivery_grocery_freq", "delivery_pkgs_freq","delivery_work_freq", "deliver_package", 'deliver_work', 'deliver_grocery', 'deliver_food')
days <- c("dayofweek", "typical_day", "daynum", "pernum")


```



```{r, getting survey data with pre-identified variables for households}

# -- How frequently is someone having a delivery made?? 
# -- 2017, 2019, 2017/2019, 2021

# pull datasets from two separate dataframes per year

dsurvey_17 <- get_hhts(survey = survey_c$survey, 
                        level = "d", 
                        vars = c(delivery_type, days)) 



dsurvey_19 <- get_hhts(survey = survey_d$survey, 
                       level = "d", 
                       vars = c(delivery_type, days)) 



dsurvey_1719 <- get_hhts(survey = survey_a$survey, 
                       level = "d", 
                       vars = c(delivery_type, days)) 



dsurvey_21 <- get_hhts(survey = survey_b$survey, 
                       level = "d", 
                       vars = c(delivery_type, days)) 



```



```{r, visualization functions}

# function combining the delivery and household functions into one and ordering income by levels

smp_delivery_combo <- function(data, year) {
  ## rewriting labels of responses to be more concise
  temp_table <- data %>%
    mutate(delivery_food_all= case_when((pernum==1 & is.na(delivery_food_freq) & is.na(deliver_food)) ~ 'No HH Response',
                                        # pernum == 1 removes households where multiple members answered the question
                                        (pernum>1) ~ 'Not Person One, not the responder',
                                        delivery_food_freq == "0 (none)"  ~ 'No Delivery',
                                        deliver_food=='No' ~ 'No Delivery',
                                        
                                        TRUE ~ 'Delivery Received'))%>%
    mutate(delivery_pkgs_all= case_when((pernum==1 & is.na(delivery_pkgs_freq) & is.na(deliver_package)) ~ 'No HH Response',
                                        (pernum>1) ~ 'Not Person One, not the responder',
                                        deliver_package=='No' ~ 'No Delivery',
                                        delivery_pkgs_freq == "0 (none)"  ~ 'No Delivery',
                                        TRUE ~ 'Delivery Received'))%>%
    mutate(delivery_grocery_all=case_when((pernum==1 & is.na(delivery_grocery_freq) & is.na(deliver_grocery)) ~ 'No HH Response',
                                        (pernum>1) ~ 'Not Person One, not the responder',
                                        delivery_grocery_freq == "0 (none)"  ~ 'No Delivery',
                                        deliver_grocery=='No' ~ 'No Delivery',
                                        TRUE ~ 'Delivery Received'))%>%
    mutate(delivery_work_all= case_when((pernum==1 & is.na(delivery_work_freq) & is.na(deliver_work)) ~ 'No HH Response',
                                        (pernum>1) ~ 'Not Person One, not the responder',
                                        deliver_work =='No' ~ 'No Delivery',
                                        delivery_work_freq == "0 (none)"  ~ 'No Delivery',
                                         TRUE ~ 'Delivery Received'))
  temp_table<-temp_table%>%filter(pernum==1)%>% dplyr::select(survey,household_id,daynum, delivery_pkgs_all, delivery_grocery_all, delivery_food_all )
  temp_table
}
```

```{r, generate datasets}


delivery_17 <- smp_delivery_combo(dsurvey_17, '2017')


delivery_19 <- smp_delivery_combo(dsurvey_19, '2019')


delivery_21 <- smp_delivery_combo(dsurvey_21, '2021')
```

```{r}


delivery_17_19_21<- rbind(delivery_17, delivery_19, delivery_21)
```

```{r}
household_data_17_19_21_shop<- household_data_17_19_21_shop %>%mutate(daynum=ifelse(daynum==0,1,daynum))
```

Finally joining the households with demographics and number of shopping trips, to the delivery info.
```{r}
household_data_17_19_21_shop_del<-merge(household_data_17_19_21_shop, delivery_17_19_21, by=c('household_id', 'daynum'))%>%
  filter(delivery_food_all%in% c('Delivery Received', 'No Delivery'))%>%
  mutate('delivery_food'=ifelse(delivery_food_all=='Delivery Received', 1, 0))

```

** Reading in data about Census Tracts to understand how characteristics of the home geography relate to the propensity to receive deliveries.
```{r}
displ_index_data<- '/Coding/CURRENT_REPOS_GITHUB/travel-studies/2021/modeling/displacement_risk_estimation.csv'
tract_data<-read.csv(displ_index_data)

```

```{r}
household_data_17_19_21_shop_del$final_home_tract10<-as.character(household_data_17_19_21_shop_del$final_home_tract10)
```

Join tract level data to household data at the home end.
```{r}
hh_shop_tract<-merge(household_data_17_19_21_shop_del, tract_data,by.x='final_home_tract10', by.y='GEOID', all.x=TRUE) 

```

** Make a big old model and see what stands out
```{r, eval = FALSE}
# packages - what we'd like to look at but need to break down further with bivariate analysis
hh_shop_tract1 <- hh_shop_tract %>%
  filter(hhincome_detailed2 != c("Prefer not to answer"))%>%
  filter(hh_race_apoc != "Missing")

deliv_model<-glm(formula= delivery_food~hhincome_detailed2+lifecycle2+final_home_is_rgc+hh_race_apoc+poor_english+severe_cost_burdened+poverty_200+prox_high_inc+transit_qt_mile+dist_super+ln_jobs_auto_30+rent_own+NoVehicles+broadband2,family=binomial(), data=hh_shop_tract1)
summary(deliv_model)
plot_summs(deliv_model, scale=TRUE)
```

```{r, separate by years 2017, 2019, 2021}

hh_shop_tract_17 <- hh_shop_tract %>%
  filter(survey.x == '2017')

hh_shop_tract_19 <- hh_shop_tract %>%
  filter(survey.x == '2019')

hh_shop_tract_21 <- hh_shop_tract %>%
  filter(survey.x == '2021')

```

* A smaller model with only statistically significant variables.
```{r, income}
# packages + income levels
# default 100-149,000

hh_shop_tract2 <- hh_shop_tract %>%
  filter(hhincome_detailed2 != c("Prefer not to answer"))

# changing the default in the regression model
# use ~ to show what you are determining the relationship to (for instance, relating food deliveries to household income)
# new dataset <- glm(formula = delivery_food ~ hhincome_detailed2)
# DEFAULT VARIABLE/CATEGORY IS CHOSEN BY FIRST ALPHABETICAL OR NUMERICAL

# look at frequency of variable responses
table(hhincome_detailed2)

# relevel the variables
hhincome_detailed2 <- relevel(hhincome_detailed2, ref ="100,000-$149,999")

# run model with new relevel
  
deliv_model_income<-glm(formula = delivery_food ~ hhincome_detailed2,
                        family=binomial(), 
                        data=hh_shop_tract2)
summary(deliv_model_income)

plot_summs(deliv_model_income, scale=TRUE)
```

```{r, kids or not}
# default under 35 with kids

deliv_model_lifecycle<-glm(formula= delivery_food~lifecycle2, family=binomial(), data=hh_shop_tract)
summary(deliv_model_lifecycle)
plot_summs(deliv_model_lifecycle, scale=TRUE)
```

```{r, lifecycle}
# default under 35 with kids

deliv_model_lifecycle<-glm(formula= delivery_food~lifecycle2, family=binomial(), data=hh_shop_tract)
summary(deliv_model_lifecycle)
plot_summs(deliv_model_lifecycle, scale=TRUE)
```

```{r, numworkers}
# unsure why not showing range of numeric 0-5???

deliv_model_numworkers<-glm(formula= delivery_food~numworkers, family=binomial(), data=hh_shop_tract)
summary(deliv_model_numworkers)
plot_summs(deliv_model_numworkers, scale=TRUE)
```

```{r, shop_trips}
# default 0 or 1, and unsure of default

deliv_model_shop<-glm(formula= delivery_food~shop_trips, family=binomial(), data=hh_shop_tract)
summary(deliv_model_shop)
plot_summs(deliv_model_shop, scale=TRUE)
```

```{r, final home is rgc}
#default is NOT in rgc

deliv_model_rgc<-glm(formula= delivery_food~final_home_is_rgc, family=binomial(), data=hh_shop_tract)
summary(deliv_model_rgc)
plot_summs(deliv_model_rgc, scale=TRUE)
```

```{r, race}
# default is Asian-POC

hh_shop_tract3 <- hh_shop_tract %>%
  filter(hh_race_apoc != c("Missing"))

deliv_model_race<-glm(formula= delivery_food~hh_race_apoc, family=binomial(), data=hh_shop_tract3)
summary(deliv_model_race)
plot_summs(deliv_model_race, scale=TRUE)
```

```{r, broadband}
deliv_model_sig<-glm(formula= delivery_food~+broadband2,family=binomial(), data=hh_shop_tract)
summary(deliv_model_sig)
plot_summs(deliv_model_sig, scale=TRUE)

```

```{r, mobile}
deliv_model_sig<-glm(formula= delivery_food~mobile_device,family=binomial(), data=hh_shop_tract)
summary(deliv_model_sig)
plot_summs(deliv_model_sig, scale=TRUE)

```

```{r, rent own}
deliv_model_sig<-glm(formula= delivery_food~rent_own,family=binomial(), data=hh_shop_tract)
summary(deliv_model_sig)
plot_summs(deliv_model_sig, scale=TRUE)

```

```{r, vehicles}
deliv_model_sig<-glm(formula= delivery_food~NoVehicles,family=binomial(), data=hh_shop_tract)
summary(deliv_model_sig)
plot_summs(deliv_model_sig, scale=TRUE)

```
https://www.cdc.gov/training/SIC_CaseStudy/Interpreting_Odds_ptversion.pdf

https://stats.stackexchange.com/questions/412668/how-to-interpret-a-negative-coefficient-in-logistic-regression
https://mmuratarat.github.io/2019-09-05/odds-ratio-logistic-regression#:~:text=The%20odds%20of%20an%20event,0.9%3D0.111%20(recurring).
https://www.polyu.edu.hk/cbs/sjpolit/logisticregression.html

```{r}
odds<- exp(deliv_model_sig$coefficients)
odds
```
Interpretation:
* All else being equal, the probability of receiving a package on a given day is less than 50%
* The strongest association between receiving a package and not is for household income; as income increases the odds of receiving a package increase.
* Older households are less likely to get deliveries.
* Smaller households are less likely to get deliveries.
** having more workers in the household is associated with more deliveries.
* People were more likely to get deliveries in 2019 and even more likely in 2021 than 2017.
* People living within a quarter mile of transit are more likely to get deliveries.
* Living in a Census tract with a high share of people with limited English profiency is associated with getting fewer deliveries
* A household having shopping trips in their day is associated with getting more deliveries.