---
title: Delivery Data Dump
date: today
---

```{r}
#| include: false

# get data and load standard data processing
source("delivery_processing.R")  

hts_source <- "Puget Sound Regional Council (PSRC) 2017-2025 Household Travel Survey"

```

## delivery trends

- About a third of regional households got packages on an average weekday since 2019
- Food delivery became more popular since 2021. About 1 out of 25 households get food delivery on an average weekday
- In 2025, only 3% of households are getting service work on an average weekday, fewer than year 2021 and earlier
- No significant trend for grocery deliveries and deliveries outside of home

```{r}
stat_delivery <- function(delivery_var, title){
  
  res <- psrc_hts_stat(df_hts_data, "day", c(delivery_var), incl_na = TRUE) %>% 
    filter(.[[delivery_var]]=="Yes") 
  
  # result table with formatting
  df_plot <- res %>%
    # function sourced from Analysis/util.R
    get_table_format()
  
  # print your charts
  print(
    ggplot(res, 
           aes(x=factor(survey_year), y=prop)) +
      geom_col(fill=psrc_colors$pgnobgy_5[1]) +
      geom_text(aes(label=scales::percent(prop, accuracy = 1)),
                vjust = -0.3) +
      geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe), width = 0.4) +
      scale_y_continuous(labels = scales::percent) +
      psrc_style() +
      theme(panel.grid.major.y = element_blank(),
            axis.title = element_blank())+
      labs(caption = hts_source, title = title)
  )
  
  res %>%
    # function sourced from Analysis/util.R
    get_table_format()
  
  # print result table by calling df_plot
  # return(df_plot)
  
}

fig_delivery <- function(delivery_var, title){
  
  res <- psrc_hts_stat(df_hts_data, "day", c(delivery_var), incl_na = TRUE) %>% 
    filter(.[[delivery_var]]=="Yes") 
  
  print(
    ggplot(res, 
           aes(x=factor(survey_year), y=prop)) +
      geom_col(fill=psrc_colors$pgnobgy_5[1]) +
      geom_text(aes(label=scales::percent(prop, accuracy = 1)),
                vjust = -0.3) +
      geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe), width = 0.4) +
      scale_y_continuous(labels = scales::percent) +
      psrc_style() +
      theme(panel.grid.major.y = element_blank(),
            axis.title = element_blank())+
      labs(caption = hts_source, title = title)
  )
}

tbl_delivery <- function(delivery_var, title){
  
  psrc_hts_stat(df_hts_data, "day", c(delivery_var), incl_na = TRUE) %>% 
    filter(.[[delivery_var]]=="Yes")  %>%
    
    # function sourced from Analysis/util.R
    get_table_format()
  
}

```

### package delivery

```{r}
#| label: fig-deliver-package
#| fig-height: 3
#| fig-width: 5.1
#| out-width: "350px"

fig_delivery("deliver_package", "package delivery")

```

```{r}
#| label: tbl-deliver-package

knitr::kable(tbl_delivery("deliver_package", "package delivery"))

```

### food delivery

```{r}
#| label: fig-deliver-food
#| fig-height: 3
#| fig-width: 5.1
#| out-width: "450px"

fig_delivery("deliver_food", "food delivery")

```

```{r}
#| label: tbl-deliver-food

knitr::kable(tbl_delivery("deliver_food", "food delivery"))

```

### service work

```{r}
#| label: fig-deliver-work
#| fig-height: 3
#| fig-width: 5.1
#| out-width: "550px"

fig_delivery("deliver_work", "service work")
```

```{r}
#| label: tbl-deliver-work

knitr::kable(tbl_delivery("deliver_work", "service work"))

```

### grocery delivery

```{r}
#| label: fig-deliver-grocery
#| fig-height: 3
#| fig-width: 5.1
#| out-width: "650px"

fig_delivery("deliver_grocery", "grocery delivery")

```

```{r}
#| label: tbl-deliver-grocery

knitr::kable(tbl_delivery("deliver_grocery", "grocery delivery"))

```

### delivery outside home

```{r}
#| label: fig-deliver-outside
#| fig-height: 3
#| fig-width: 5.1
#| out-width: "40%"

fig_delivery("deliver_outside_any", "delivery outside home (inc. at office)")

```

```{r}
#| label: tbl-deliver-outside

knitr::kable(tbl_delivery("deliver_outside_any", "delivery outside home (inc. at office)"))

```

### delivery outside home and at office separated

```{r}
#| label: fig-deliver-elsewhere
#| fig-height: 3
#| fig-width: 5.1

fig_delivery("deliver_elsewhere", "delivery outside home")
```

```{r}
#| label: tbl-deliver-elsewhere

knitr::kable(tbl_delivery("deliver_elsewhere", "delivery outside home"))

```

```{r}
#| label: fig-deliver-office
#| fig-height: 3
#| fig-width: 5.1

fig_delivery("deliver_office", "delivery at office")
```

```{r}
#| label: tbl-deliver-office

knitr::kable(tbl_delivery("deliver_office", "delivery at office"))

```



## delivery by demographic characteristics

```{r}

plot_delivery_type <- function(delivery_var, group_var, title){

  res <- psrc_hts_stat(df_hts_data, "day", c(group_var,delivery_var), incl_na = TRUE) %>%
    filter(.[[delivery_var]]=="Yes", .[[group_var]]!="Prefer not to answer")

  # print your charts
  print(
    ggplot(res,
           aes(x=delivery_var, y=prop, fill=!!sym(group_var))) +
      geom_col(position = position_dodge(),width = 0.8) +
      geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                    position = position_dodge(0.8),
                    width = 0.4) +
      facet_wrap(~ survey_year, ncol=5)+
      scale_fill_manual(values = psrc_colors$pgnobgy_5)+
      scale_y_continuous(labels = scales::percent) +
      psrc_style() +
      theme(panel.grid.major.y = element_blank(),
            axis.title = element_blank(),
            axis.text.x = element_blank())+
      labs(caption = hts_source,
           title = title)
  )

  print(
    ggplot(res,
           aes(x=delivery_var, y=prop, fill=factor(survey_year))) +
      geom_col(position = position_dodge(),width = 0.8) +
      geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                    position = position_dodge(0.8),
                    width = 0.4) +
      facet_wrap(as.formula(paste("~", group_var)), ncol=5)+
      scale_fill_manual(values = psrc_colors$blues_inc)+
      scale_y_continuous(labels = scales::percent) +
      psrc_style() +
      theme(panel.grid.major.y = element_blank(),
            axis.title = element_blank(),
            axis.text.x = element_blank())+
      labs(caption = hts_source,
           title = title)
  )
  
}

table_delivery_type <- function(){
  
  res <- psrc_hts_stat(df_hts_data, "day", c(group_var,delivery_var), incl_na = TRUE) %>%
    filter(.[[delivery_var]]=="Yes", .[[group_var]]!="Prefer not to answer")

  # result table with formatting
  df_plot <- res %>%
    # function sourced from Analysis/util.R
    get_table_format()
  
}
```


<!-- ### household income -->

<!-- - households with higher income have been consistently receiving more packages than others -->
<!-- - no noticeable trends -->

<!-- :::{.panel-tabset} -->

<!-- #### package -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_package","income_3group", -->
<!--                               "package delivery by income") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### grocery -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_grocery","income_3group", -->
<!--                               "grocery delivery by income") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### food -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_food","income_3group", -->
<!--                               "food delivery by income") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- ::: -->


<!-- ### number of workers -->

<!-- :::{.panel-tabset} -->

<!-- #### package -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_package","numworkers_3cap", -->
<!--                               "package delivery by number of workers") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### grocery -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_grocery","numworkers_3cap", -->
<!--                               "grocery delivery by number of workers") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### food -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_food","numworkers_3cap", -->
<!--                               "food delivery by number of workers") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- ::: -->

<!-- ### vehicle availability -->

<!-- Households with no cars have more grocery deliveries -> grocery deliveries are accessible for households without cars -->

<!-- Households with enough cars have more food deliveries -> food deliveries serve households with higher income -->

<!-- :::{.panel-tabset} -->

<!-- #### package -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_package","have_cars", -->
<!--                               "package delivery by auto availability") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### grocery -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_grocery","have_cars", -->
<!--                               "grocery delivery by auto availability") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### food -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_food","have_cars", -->
<!--                               "food delivery by auto availability") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- ::: -->


<!-- ### household race categories -->

<!-- :::{.panel-tabset} -->

<!-- #### package -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_package","hh_race_category", -->
<!--                               "package delivery by household race") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### grocery -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_grocery","hh_race_category", -->
<!--                               "grocery delivery by household race") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### food -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_food","hh_race_category", -->
<!--                               "food delivery by household race") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- ::: -->

<!-- ### household composition -->

<!-- :::{.panel-tabset} -->

<!-- #### package -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_package","lifecycle_2group", -->
<!--                               "package delivery by household composition") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### grocery -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_grocery","lifecycle_2group", -->
<!--                               "grocery delivery by household composition") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### food -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_food","lifecycle_2group", -->
<!--                               "food delivery by household composition") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- ::: -->

<!-- ## delivery by home location and type -->


<!-- ### home county -->

<!-- :::{.panel-tabset} -->

<!-- #### package -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_package","home_county", -->
<!--                               "package delivery by home location - County") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### grocery -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_grocery","home_county", -->
<!--                               "grocery delivery by home location - County") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### food -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_food","home_county", -->
<!--                               "food delivery by home location - County") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- ::: -->

<!-- ### home in centers -->

<!-- :::{.panel-tabset} -->

<!-- #### package -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_package","home_rgc", -->
<!--                               "package delivery by home location - Center") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### grocery -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_grocery","home_rgc", -->
<!--                               "grocery delivery by home location - Center") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### food -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_food","home_rgc", -->
<!--                               "food delivery by home location - Center") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- ::: -->


<!-- ### residence type -->

<!-- :::{.panel-tabset} -->

<!-- #### package -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_package","res_type_group", -->
<!--                               "package delivery by residence type") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### grocery -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_grocery","res_type_group", -->
<!--                               "grocery delivery by residence type") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### food -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_food","res_type_group", -->
<!--                               "food delivery by residence type") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- #### service work -->

<!-- ```{r} -->
<!-- #| layout-ncol: 2 -->
<!-- #| out-width: "50%" -->

<!-- df_plot <- stat_delivery_type("deliver_work","res_type_group", -->
<!--                               "service work by residence type") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_plot -->
<!-- ``` -->

<!-- ::: -->

## shopping trip rates over the years

```{r}
#| label: fig-shop-trip-rate
#| fig-height: 3.8
#| fig-width: 5.1

df_plot <- psrc_hts_triprate(hts_data, group_vars="dest_purpose_cat") %>%
  filter(dest_purpose_cat == "Shopping")

ggplot(df_plot,
       aes(x=factor(survey_year), y=mean), fill=psrc_colors$pgnobgy_5[1]) +
  geom_col() +
  geom_text(aes(label=scales::number(mean, accuracy = 0.1)),
            vjust = -0.3) +
  scale_y_continuous(labels = scales::number) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank()) +
  labs(title = "shopping trip rates over years",
       caption = hts_source)
```


