---
title: "Delivery: Explore Landuse"
date: today
---

```{r}
#| include: false

# get data and load standard data processing
source("delivery_processing.R")  

hts_source <- "Puget Sound Regional Council (PSRC) 2017-2025 Household Travel Survey"
sc_source <- "Soundcast RTP 2023 baseyear outputs"

```

```{r}
landuse_vars <- c("home_hh_1", "home_hh_2", "home_emptot_1", "home_emptot_2", 
         "home_auto_jobs_access", "home_transit_jobs_access")

df_landuse <- df_hts_data$hh %>%
  select(hh_id,
         # categorical vars
         hhincome_num, have_cars, home_rgcname,
         # locations
         home_lat, home_lng, 
         # landuse vars
         any_of(landuse_vars)) %>%
  mutate(home_actunit_1 = home_hh_1 + home_emptot_1,
         home_actunit_2 = home_hh_2 + home_emptot_2)
day_landuse <- df_hts_data$day %>%
  left_join(df_landuse, by="hh_id") %>%
  filter(survey_year==2025, !is.na(deliver_package)) %>%
  mutate(deliver_package_bi = case_when(deliver_package=="Yes"~1,
                                        deliver_package=="No"~0),
         home_rgcname_bi = case_when(home_rgcname=="Not RGC" | is.na(home_rgcname)~0,
                                     TRUE~1),
         have_cars_bi = case_when(have_cars=="no cars"~0,
                                  have_cars=="own cars"~1,
                                     TRUE~NA))
```

::: {.callout-note}
These are 2025 unweighted results. The 19 households with missing delivery values are excluded

- value counts for household that received a package on an average weekday

```{r}
df <- data.frame(table(day_landuse$deliver_package, useNA = "always"))
colnames(df) <- c("deliver_package","value count")
df
```

:::




## correlation matrix for landuse variables

the landuse variables are highly correlated 

```{r}
#| out-width: "60%"
res <- cor(
  day_landuse %>% 
    select(any_of(landuse_vars),home_rgcname_bi,hhincome_num,have_cars_bi) %>%
    filter(!is.na(hhincome_num))
  )

library(corrplot)
corrplot(res, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```


## histograms

```{r}
#| layout-ncol: 3
#| out-width: "33%"
for (var in landuse_vars){
  
  print(
    ggplot(day_landuse %>% filter(!is.na(deliver_package)), 
           aes(x=!!sym(var), fill = deliver_package, color = deliver_package)) + 
      geom_histogram(alpha=0.5,bins = 20)+
      geom_density(alpha=.2) +
      psrc_style() +
      theme(axis.title = element_blank(),
            axis.text.x = element_blank())+
      labs(caption = paste(hts_source, "and", sc_source),
           title = paste("histogram for", var))
  )
  
}

```

## density histograms

```{r}
#| layout-ncol: 3
#| out-width: "33%"
for (var in landuse_vars){
  
  print(
    ggplot(day_landuse %>% filter(!is.na(deliver_package)), 
           aes(x=!!sym(var), fill = deliver_package, color = deliver_package)) + 
      geom_histogram(aes(y=after_stat(density)), alpha=0.5,bins = 20)+
      geom_density(alpha=.2) +
      psrc_style() +
      theme(axis.title = element_blank(),
            axis.text.x = element_blank())+
      labs(caption = paste(hts_source, "and", sc_source),
           title = paste("density histogram for", var, "by package delivery"))
  )
  
}

```


```{r}
#| out-width: "45%"
ggplot(day_landuse, aes(home_hh_1, fill = factor(home_rgcname_bi))) +
  geom_histogram(aes(y=after_stat(density)), position="identity", alpha=0.5,bins = 20)+
  geom_density(alpha=0.6) +
      psrc_style() +
      theme(axis.title = element_blank(),
            axis.text.x = element_blank())+
      labs(caption = paste(hts_source, "and", sc_source),
           title = paste("histogram for home_hh_1 by home in centers"))
```

## mapping

```{r}
library(psrcelmer)
library(leaflet)
library(sf)
# 2. Convert to an sf object (assuming WGS84, EPSG: 4326)
sf_points <- day_landuse %>%
  st_as_sf(coords = c("home_lat", "home_lng"), crs = 4326) %>%
  st_transform(sf_points, crs = 2285)

region_center_plot <- st_read_elmergeo('urban_centers_evw')

pal <- colorFactor(c("navy", "red"), domain = day_landuse$deliver_package)
pal_home_hh_1 <- colorNumeric("YlOrRd", domain = day_landuse$home_hh_1)
pal_access <- colorNumeric("YlOrRd", domain = day_landuse$home_transit_jobs_access)
pal_access_auto <- colorNumeric("YlOrRd", domain = day_landuse$home_auto_jobs_access)

leaflet(region_center_plot) %>%
  addPolygons(
    stroke = FALSE) %>%  
  addProviderTiles(providers$CartoDB.Positron, group = "Positron (minimal)") %>%
  addCircleMarkers(day_landuse$home_lng,day_landuse$home_lat,
                   group = "deliver_package",
                   color = pal(day_landuse$deliver_package),
                   stroke = FALSE,
                   fillOpacity = 0.5) %>%
  addLegend("bottomleft", pal = pal, values = ~day_landuse$deliver_package,
            group = "deliver_package",
            opacity = 1
  ) %>%
  addCircleMarkers(day_landuse$home_lng,day_landuse$home_lat,
                   group = "home_hh_1",
                   color = pal_home_hh_1(day_landuse$home_hh_1),
                   stroke = FALSE,
                   fillOpacity = 0.5) %>%
  addLegend("bottomleft", pal = pal_home_hh_1, values = ~day_landuse$home_hh_1,
            group = "home_hh_1",
            opacity = 1
  ) %>%
  addCircleMarkers(day_landuse$home_lng,day_landuse$home_lat,
                   group = "home_transit_jobs_access",
                   color = pal_access(day_landuse$home_transit_jobs_access),
                   stroke = FALSE, 
                   fillOpacity = 0.5) %>%
  addLegend("bottomleft", pal = pal_access, values = ~day_landuse$home_transit_jobs_access,
            group = "home_transit_jobs_access",
            opacity = 1
  ) %>%
  addCircleMarkers(day_landuse$home_lng,day_landuse$home_lat,
                   group = "home_auto_jobs_access",
                   color = pal_access_auto(day_landuse$home_auto_jobs_access),
                   stroke = FALSE, 
                   fillOpacity = 0.5) %>%
  addLegend("bottomleft", pal = pal_access_auto, values = ~day_landuse$home_auto_jobs_access,
            group = "home_auto_jobs_access",
            opacity = 1
  ) %>%
  # Layers control
  addLayersControl(
    baseGroups = c(
      "Positron (minimal)"
    ),
    overlayGroups = c("deliver_package","home_hh_1", "home_auto_jobs_access","home_transit_jobs_access"),
    options = layersControlOptions(collapsed = FALSE)
  )
```



## binary logit models

### current model

```{r}
model <- glm(deliver_package_bi ~ home_auto_jobs_access + home_rgcname_bi + hhincome_num, 
             data = day_landuse, 
             family = "binomial")
summary(model)
```

### other testing models

:::{.panel-tabset}

#### full model

```{r}
model <- glm(deliver_package_bi ~ home_actunit_1 + home_auto_jobs_access + home_rgcname_bi + hhincome_num + have_cars_bi, 
             data = day_landuse, 
             family = "binomial")
summary(model)
```

#### hhincome_num

```{r}
model <- glm(deliver_package_bi ~ hhincome_num, 
             data = day_landuse, 
             family = "binomial")
summary(model)
```

#### have_cars_bi

```{r}
model <- glm(deliver_package_bi ~ have_cars_bi, 
             data = day_landuse, 
             family = "binomial")
summary(model)
```

#### home_emptot_1

```{r}
model <- glm(deliver_package_bi ~ home_emptot_1, 
             data = day_landuse, 
             family = "binomial")
summary(model)
```

#### home_hh_1

```{r}
model <- glm(deliver_package_bi ~ home_hh_1, 
             data = day_landuse, 
             family = "binomial")
summary(model)
```

#### home_transit_jobs_access

```{r}
model <- glm(deliver_package_bi ~ home_transit_jobs_access, 
             data = day_landuse, 
             family = "binomial")
summary(model)
```

#### home_auto_jobs_access

```{r}
model <- glm(deliver_package_bi ~ home_auto_jobs_access, 
             data = day_landuse, 
             family = "binomial")
summary(model)
```

#### home_rgcname_bi

```{r}
model <- glm(deliver_package_bi ~ home_rgcname_bi, 
             data = day_landuse, 
             family = "binomial")
summary(model)
```

:::
