---
title: Delivery Data Dump
date: today
---

```{r}
#| include: false

# get data and load standard data processing
source("delivery_processing.R")  

hts_source <- "Puget Sound Regional Council (PSRC) 2017-2025 Household Travel Survey"

```

## delivery trends

- a third of regional households got a package on an average weekday in 2023
- about 1 out of 25 households get food delivery on an average weekday.
- food delivery became more popular since 2021

:::{.panel-tabset}

```{r}
stat_delivery <- function(delivery_var, title){
  
  res <- psrc_hts_stat(df_hts_data, "day", c(delivery_var), incl_na = TRUE) %>% 
    filter(.[[delivery_var]]=="Yes") 
  
  # result table with formatting
  df_plot <- res %>%
    # function sourced from Analysis/util.R
    get_table_format()
  
  # print your charts
  print(
    ggplot(res, 
           aes(x=factor(survey_year), y=prop)) +
      geom_col(fill=psrc_colors$pgnobgy_5[1]) +
      geom_text(aes(label=scales::percent(prop, accuracy = 0.1)),
                vjust = -0.3) +
      geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe), width = 0.4) +
      scale_y_continuous(labels = scales::percent) +
      psrc_style() +
      theme(panel.grid.major.y = element_blank(),
            axis.title = element_blank())+
      labs(caption = hts_source, title = title)
  )
  
  # print result table by calling df_plot
  return(df_plot)
  
}
```

### package

```{r}
#| out-width: "37%"
#| fig-width: 3
df_plot <- stat_delivery("deliver_package", "package delivery")
```

```{r}
# print table
df_plot
```

### service work

```{r}
#| out-width: "37%"
#| fig-width: 3
df_plot <- stat_delivery("deliver_work", "service work")
```

```{r}
df_plot
```

### food delivery

```{r}
#| out-width: "37%"
#| fig-width: 3
df_plot <- stat_delivery("deliver_food", "food delivery")
```

```{r}
df_plot
```

### grocery delivery

```{r}
#| out-width: "37%"
#| fig-width: 3
df_plot <- stat_delivery("deliver_grocery", "grocery delivery")
```

```{r}
df_plot
```

### delivery outside home + at office

```{r}
#| out-width: "37%"
#| fig-width: 3
df_plot <- stat_delivery("deliver_outside_any", "delivery outside home (inc. at office)")
```

```{r}
df_plot
```

- delivery outside home and at office separated

```{r}
#| layout-ncol: 2
#| out-width: "37%"
#| fig-width: 3

df_plot <- stat_delivery("deliver_elsewhere", "delivery outside home")
df_plot2 <- stat_delivery("deliver_office", "delivery at office")
```

```{r}
#| layout-ncol: 2
df_plot
df_plot2
```


:::


## delivery by demographic characteristics

```{r}

stat_delivery_type <- function(delivery_var, group_var, title){
  
  res <- psrc_hts_stat(df_hts_data, "day", c(group_var,delivery_var), incl_na = TRUE) %>% 
    filter(.[[delivery_var]]=="Yes", .[[group_var]]!="Prefer not to answer")
  
  # result table with formatting
  df_plot <- res %>%
    # function sourced from Analysis/util.R
    get_table_format()
  
  # print your charts
  print(
    ggplot(res, 
           aes(x=delivery_var, y=prop, fill=!!sym(group_var))) +
      geom_col(position = position_dodge(),width = 0.8) +
      geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                    position = position_dodge(0.8),
                    width = 0.4) +
      facet_wrap(~ survey_year, ncol=5)+
      scale_fill_manual(values = psrc_colors$pgnobgy_5)+
      scale_y_continuous(labels = scales::percent) +
      psrc_style() +
      theme(panel.grid.major.y = element_blank(),
            axis.title = element_blank(),
            axis.text.x = element_blank())+
      labs(caption = hts_source,
           title = title)
  )
  
  print(
    ggplot(res, 
           aes(x=delivery_var, y=prop, fill=factor(survey_year))) +
      geom_col(position = position_dodge(),width = 0.8) +
      geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                    position = position_dodge(0.8),
                    width = 0.4) +
      facet_wrap(as.formula(paste("~", group_var)), ncol=5)+
      scale_fill_manual(values = psrc_colors$blues_inc)+
      scale_y_continuous(labels = scales::percent) +
      psrc_style() +
      theme(panel.grid.major.y = element_blank(),
            axis.title = element_blank(),
            axis.text.x = element_blank())+
      labs(caption = hts_source,
           title = title)
  )
  
  # print result table by calling df_plot
  return(df_plot)
}
```


### household income

- households with higher income have been consistently receiving more packages than others
- no noticeable trends

:::{.panel-tabset}

#### package

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_package","income_3group", 
                              "package delivery by income")
```

```{r}
df_plot
```

#### grocery

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_grocery","income_3group", 
                              "grocery delivery by income")
```

```{r}
df_plot
```

#### food

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_food","income_3group", 
                              "food delivery by income")
```

```{r}
df_plot
```

:::


### number of workers

:::{.panel-tabset}

#### package

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_package","numworkers_3cap", 
                              "package delivery by number of workers")
```

```{r}
df_plot
```

#### grocery

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_grocery","numworkers_3cap", 
                              "grocery delivery by number of workers")
```

```{r}
df_plot
```

#### food

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_food","numworkers_3cap", 
                              "food delivery by number of workers")
```

```{r}
df_plot
```

:::

### vehicle availability

Households with no cars have more grocery deliveries -> grocery deliveries are accessible for households without cars

Households with enough cars have more food deliveries -> food deliveries serve households with higher income

:::{.panel-tabset}

#### package

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_package","have_cars", 
                              "package delivery by auto availability")
```

```{r}
df_plot
```

#### grocery

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_grocery","have_cars", 
                              "grocery delivery by auto availability")
```

```{r}
df_plot
```

#### food

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_food","have_cars", 
                              "food delivery by auto availability")
```

```{r}
df_plot
```

:::


### household race categories

:::{.panel-tabset}

#### package

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_package","hh_race_category", 
                              "package delivery by household race")
```

```{r}
df_plot
```

#### grocery

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_grocery","hh_race_category", 
                              "grocery delivery by household race")
```

```{r}
df_plot
```

#### food

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_food","hh_race_category", 
                              "food delivery by household race")
```

```{r}
df_plot
```

:::

### household composition

:::{.panel-tabset}

#### package

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_package","lifecycle_2group", 
                              "package delivery by household composition")
```

```{r}
df_plot
```

#### grocery

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_grocery","lifecycle_2group", 
                              "grocery delivery by household composition")
```

```{r}
df_plot
```

#### food

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_food","lifecycle_2group", 
                              "food delivery by household composition")
```

```{r}
df_plot
```

:::

## delivery by home location and type


### home county

:::{.panel-tabset}

#### package

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_package","home_county", 
                              "package delivery by home location - County")
```

```{r}
df_plot
```

#### grocery

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_grocery","home_county", 
                              "grocery delivery by home location - County")
```

```{r}
df_plot
```

#### food

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_food","home_county", 
                              "food delivery by home location - County")
```

```{r}
df_plot
```

:::

### home in centers

:::{.panel-tabset}

#### package

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_package","home_rgc", 
                              "package delivery by home location - Center")
```

```{r}
df_plot
```

#### grocery

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_grocery","home_rgc", 
                              "grocery delivery by home location - Center")
```

```{r}
df_plot
```

#### food

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_food","home_rgc", 
                              "food delivery by home location - Center")
```

```{r}
df_plot
```

:::


### residence type

:::{.panel-tabset}

#### package

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_package","res_type_group", 
                              "package delivery by residence type")
```

```{r}
df_plot
```

#### grocery

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_grocery","res_type_group", 
                              "grocery delivery by residence type")
```

```{r}
df_plot
```

#### food

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_food","res_type_group", 
                              "food delivery by residence type")
```

```{r}
df_plot
```

#### service work

```{r}
#| layout-ncol: 2
#| out-width: "50%"

df_plot <- stat_delivery_type("deliver_work","res_type_group", 
                              "service work by residence type")
```

```{r}
df_plot
```

:::

## shopping trip rates over the years

```{r}
#| echo: true
#| out-width: "50%"

# number of shopping trips
shopping_trip <- hts_data$trip %>%
  filter(dest_purpose_cat=="Shopping") %>%
  summarise(trip_est = sum(trip_weight), .by = c(dest_purpose_cat, survey_year))

person_count <- hts_data$person %>%
  summarise(person_est = sum(person_weight), .by = survey_year)

test <- shopping_trip %>%
  left_join(person_count, by="survey_year") %>%
  mutate(trip_rate = trip_est/person_est,
         survey_year = factor(survey_year))

ggplot(test, 
       aes(x=factor(survey_year), y=trip_rate), fill=psrc_colors$pgnobgy_5[1]) +
  geom_col() +
  geom_text(aes(label=scales::number(trip_rate, accuracy = 0.01)),
            vjust = -0.3) +
  scale_y_continuous(labels = scales::number) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank()) +
  labs(title = "shopping trip rates over years",
       caption = hts_source)
```


