---
title: Delivery Analysis
date: today

params:
  errorbar: false
---

The delivery analysis explores how many households in the region received packages on an average weekday 

- @sec-delivery-trends looks at the percentage of households getting different types of deliveries or service work on an average weekday
- @sec-delivery-demo shows delivery patterns for households with different demographic characteristics
- @sec-trip-rates shows shopping trip rates over the years

```{r}
#| include: false

# get data and load standard data processing
source("delivery_processing.R")  

hts_source <- "Puget Sound Regional Council (PSRC) 2017-2025 Household Travel Survey"

# assign fill colors to delivery types
psrc_fill_dec <- list(
  deliver_package = psrc_colors$purples_dec[1:5],
  deliver_food = psrc_colors$oranges_dec[1:5],
  deliver_grocery = psrc_colors$blues_dec[1:5],
  deliver_work = psrc_colors$greens_dec[1:5],
  deliver_outside_any = psrc_colors$grays_dec[4:5],
  deliver_elsewhere = psrc_colors$grays_dec[4:5],
  deliver_office = psrc_colors$grays_dec[4:5]
  )

psrc_fill_segment <- list(
  income_2group = c( "#91268F", "#8CC63E"),
  have_cars = c( "#F05A28", "#00A7A0"),
  lifecycle_2group = c( "#91268F", "#00A7A0"),
  numworkers_3cap = psrc_colors$pgnobgy_5,
  hh_race_category = psrc_colors$gnbopgy_5,
  home_county = psrc_colors$obgnpgy_10,
  home_rgc = c( "#00A7A0", "#8CC63E"),
  res_type_group = c( "#91268F", "#F05A28")
  )
```

## delivery trends {#sec-delivery-trends}

```{r}

fig_delivery <- function(delivery_var){
  
  res <- psrc_hts_stat(df_hts_data, "day", c(delivery_var), incl_na = TRUE) %>% 
    filter(.[[delivery_var]]=="Yes") 
  
  print(
    ggplot(res, 
           aes(x=factor(survey_year), y=prop, group = 1)) +
      geom_col(aes(fill = factor(survey_year)), width=0.7) +
      geom_text(aes(label=scales::percent(prop, accuracy = 1, group = 1)),
                vjust = -0.3) +
      geom_hline(yintercept = 0, color = "#cbcbcb", linewidth = 0.3) +
      errorbar_switch(geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe), width = 0.1), 
                      on=params$errorbar) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
      scale_fill_manual(values = psrc_fill_dec[[delivery_var]]) +
      psrc_style() +
      theme(legend.position = "none",
            panel.grid.major.y = element_blank(),
            axis.text.y = element_blank(),
            axis.title = element_blank())
  )
}

tbl_delivery <- function(delivery_var){
  
  psrc_hts_stat(df_hts_data, "day", c(delivery_var), incl_na = TRUE) %>% 
    filter(.[[delivery_var]]=="Yes")  %>%
    
    # function sourced from Analysis/util.R
    get_table_format()
  
}

```

### package delivery

- Package delivery peaked in 2021 at 39\%. 
- The rate in 2025 reached a similar level at about a third of regional households got packages on an average weekday.

```{r}
#| label: fig-deliver-package
#| fig-cap: package delivery
#| fig-height: 2.5
#| fig-width: 3


fig_delivery("deliver_package")

```

```{r}
#| label: tbl-deliver-package

tbl_delivery("deliver_package")
```


### food delivery

- Food delivery dramatically increased in 2021 and stayed the same since at 4\%.

```{r}
#| label: fig-deliver-food
#| fig-cap: food delivery
#| fig-height: 2.5
#| fig-width: 3


fig_delivery("deliver_food")

```

```{r}
#| label: tbl-deliver-food

tbl_delivery("deliver_food")

```

### service work

- Service work was higher before the end of COVID at 6\% and dropped to about half in 2023 and 2025.


```{r}
#| label: fig-deliver-work
#| fig-cap: service work
#| fig-height: 2.5
#| fig-width: 3


fig_delivery("deliver_work")
```

```{r}
#| label: tbl-deliver-work

tbl_delivery("deliver_work")

```

### grocery delivery

- Grocery delivery more than doubled in 2021 and reached 5\%, but dropped to near pre-COVID level after.

```{r}
#| label: fig-deliver-grocery
#| fig-cap: grocery delivery
#| fig-height: 2.5
#| fig-width: 3


fig_delivery("deliver_grocery")

```

```{r}
#| label: tbl-deliver-grocery

tbl_delivery("deliver_grocery")

```

### delivery outside home

- 2\% of households have deliveries outside of home (incl. at office) in 2025. The rate is similar in 2023.

#### outside home + at office

```{r}
#| label: fig-deliver-outside
#| fig-cap: "delivery outside home (inc. at office)"
#| fig-height: 2.5
#| fig-width: 3


fig_delivery("deliver_outside_any")

```

```{r}
#| label: tbl-deliver-outside

tbl_delivery("deliver_outside_any")

```


#### outside home & at office separated

- deliveries outside of home was at 2\%
- deliveries received at office was low

```{r}
#| label: fig-deliver-elsewhere-office
#| fig-cap: "delivery outside home"
#| fig-height: 2.5
#| fig-width: 3
#| layout-ncol: 2


fig_delivery("deliver_elsewhere")
fig_delivery("deliver_office")
```

```{r}
#| label: tbl-deliver-elsewhere-office
#| layout-ncol: 2

tbl_delivery("deliver_elsewhere")
tbl_delivery("deliver_office")
```

## delivery by demographic characteristics {#sec-delivery-demo}

Each subsection includes two figures:

- percentage of households receiving delivery across household types in each year
- delivery trends over years for each household type

```{r}

plot_delivery_type <- function(delivery_var, group_var, type="both"){

  res <- psrc_hts_stat(df_hts_data, "day", c(group_var,delivery_var), incl_na = TRUE) %>%
    filter(.[[delivery_var]]=="Yes", .[[group_var]]!="Prefer not to answer")

  # print your charts
  if( type %in% c("year","both") ){

    df_plot <- ggplot(res,
                      aes(x=delivery_var, y=prop, fill=!!sym(group_var))) +
      geom_col(position = position_dodge(), width=0.9) +
      geom_text(aes(label=scales::percent(prop, accuracy = 1)),
                    position = position_dodge(0.9),
                vjust = -0.3, 
                size = 10, size.unit="pt") +
      errorbar_switch(geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                                    position = position_dodge(0.9),
                                    width = 0.1), 
                      on=params$errorbar) +
      geom_hline(yintercept = 0, color = "#cbcbcb", linewidth = 0.3) +
      facet_wrap(~ survey_year, ncol=5)+
      scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, 0.1))) +
      scale_fill_manual(values = psrc_fill_segment[[group_var]])+
      psrc_style() +
      theme(panel.grid.major.y = element_blank(),
            legend.text = element_text(size = 11),
            axis.title = element_blank(),
            axis.text.x = element_blank(),
            axis.text.y = element_blank())

    print(df_plot)

  }

  if( type %in% c("cat","both") ){

    df_plot <- ggplot(res,
                      aes(x=factor(survey_year), y=prop, fill=factor(survey_year))) +
      geom_col(width = 0.8) +
      geom_text(aes(label=scales::percent(prop, accuracy = 1)),
                vjust = -0.3) +
      errorbar_switch(geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                                    width = 0.1), 
                      on=params$errorbar) +
      geom_hline(yintercept = 0, color = "#cbcbcb", linewidth = 0.3) +
      facet_wrap(as.formula(paste("~", group_var)), ncol=5)+
      scale_fill_manual(values = psrc_fill_dec[[delivery_var]])+
      scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, 0.1))) +
      psrc_style() +
      theme(legend.position = "none",
            strip.text = element_text(size = 10.5),
            panel.grid.major.y = element_blank(),
            panel.spacing.x = unit(0.5, "lines"),
            axis.title = element_blank(),
            axis.text.x = element_text(size = 10.5),
            axis.text.y = element_blank())

    print(df_plot)

  }

}

table_delivery_type <- function(delivery_var, group_var){

  psrc_hts_stat(df_hts_data, "day", c(group_var,delivery_var), incl_na = TRUE) %>%
    filter(.[[delivery_var]]=="Yes", .[[group_var]]!="Prefer not to answer") %>%
    # function sourced from Analysis/util.R
    get_table_format()


}
```

### household income

#### package delivery by household income

- higher income households had been consistently receiving more packages at around 40\% in 2019 and later years

```{r}
#| label: fig-package-income
#| fig-cap: "package delivery by household income"
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_package","income_2group",
                   type="year")
```

```{r}
#| label: fig-package-income-year
#| fig-cap: "household income: package delivery by year"
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_package","income_2group",
                   type="cat")
```


```{r}
#| label: tbl-package-income
table_delivery_type("deliver_package","income_2group")
```

#### grocery delivery by household income

- No significant difference observed in grocery delivery across income levels
- Sample sizes for were low, ranging 33 to 79 counts

```{r}
#| layout-ncol: 2
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_grocery","income_2group")
```

```{r}
table_delivery_type("deliver_grocery","income_2group")
```


#### food delivery by household income

- No significant difference observed in food delivery across income levels

```{r}
#| layout-ncol: 2
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_food","income_2group")
```

```{r}
table_delivery_type("deliver_food","income_2group")
```



### vehicle availability

#### package delivery by vehicle availability

- More households that owned cars had packages delivered compare to no-car households (however, auto ownership may be highly correlated with household income)

```{r}
#| label: fig-package-vehicle
#| fig-cap: "package delivery by household income"
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_package","have_cars",
                   type="year")
```

```{r}
#| fig-cap: "household income: package delivery by year"
#| fig-height: 3.9
#| fig-width: 5.2
#| fif-align: center


plot_delivery_type("deliver_package","have_cars",
                   type="cat")
```


```{r}
#| label: tbl-package-vehicle
table_delivery_type("deliver_package","have_cars")
```

#### grocery delivery by vehicle availability

- Grocery deliveries peaked at 5\% during COVID for household that owned vehicles and the rates dropped in 2023 and 2025
- Grocery deliveries for households without cars stayed high from 2021 to 2025. This could be because the service was an accessible way to get groceries for households without cars
- Sample sizes for were low

```{r}
#| label: fig-grocery-vehicle
#| fig-cap: "grocery delivery by auto availability"
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_grocery","have_cars",
                   type="year")
```

```{r}
#| label: fig-grocery-vehicle-year
#| fig-cap: "auto availability: grocery delivery by year"
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_grocery","have_cars",
                   type="cat")
```

```{r}
#| label: tbl-grocery-vehicle

table_delivery_type("deliver_grocery","have_cars")
```


#### food delivery by vehicle availability

- Sample sizes for were low

```{r}
#| layout-ncol: 2
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_food","have_cars")
```

```{r}
table_delivery_type("deliver_food","have_cars")
```

### household composition

#### package delivery by household composition

```{r}
#| label: fig-package-hh
#| fig-cap: "package delivery by household composition"
#| fig-height: 3.9
#| fig-width: 5.6


plot_delivery_type("deliver_package","lifecycle_2group",
                   type="year")
```

```{r}
#| fig-cap: "household composition: package delivery by year"
#| fig-height: 3.9
#| fig-width: 5.6


plot_delivery_type("deliver_package","lifecycle_2group",
                   type="cat")
```

```{r}
#| label: tbl-package-hh

table_delivery_type("deliver_package","lifecycle_2group")
```

#### grocery delivery by household composition

```{r}
#| layout-ncol: 2
#| fig-height: 3.9
#| fig-width: 5.6


plot_delivery_type("deliver_grocery","lifecycle_2group")
```

```{r}

table_delivery_type("deliver_grocery","lifecycle_2group")
```

#### food delivery by household composition

```{r}
#| layout-ncol: 2
#| fig-height: 3.9
#| fig-width: 5.6


plot_delivery_type("deliver_food","lifecycle_2group")
```

```{r}

table_delivery_type("deliver_food","lifecycle_2group")
```

### number of workers

#### package delivery by number of workers

```{r}
#| fig-height: 3
#| fig-width: 8
#| fig-align: center

plot_delivery_type("deliver_package","numworkers_3cap",type="year")
```

```{r}
#| fig-height: 3
#| fig-width: 8.5
#| fig-align: center

plot_delivery_type("deliver_package","numworkers_3cap",type="cat")
```

```{r}
table_delivery_type("deliver_food","lifecycle_2group")
```

#### grocery delivery by number of workers

```{r}
#| fig-height: 3.9
#| fig-width: 7.5
#| fig-align: center

plot_delivery_type("deliver_grocery","numworkers_3cap",type="year")
```

```{r}
#| fig-height: 3.9
#| fig-width: 8.5
#| fig-align: center

plot_delivery_type("deliver_grocery","numworkers_3cap",type="cat")
```


```{r}
table_delivery_type("deliver_grocery","numworkers_3cap")
```

#### food delivery by number of workers

```{r}
#| fig-height: 3.9
#| fig-width: 7.5
#| fig-align: center

plot_delivery_type("deliver_food","numworkers_3cap",type="year")
```

```{r}
#| fig-height: 3.9
#| fig-width: 8.5
#| fig-align: center

plot_delivery_type("deliver_food","numworkers_3cap",type="cat")
```

```{r}
table_delivery_type("deliver_food","numworkers_3cap")
```

### household race categories

#### package delivery by household race categories

```{r}
#| fig-height: 3
#| fig-width: 12
#| fig-align: center

plot_delivery_type("deliver_package","hh_race_category",type="year")
```

```{r}
#| fig-height: 3
#| fig-width: 13
#| fig-align: center

plot_delivery_type("deliver_package","hh_race_category",type="cat")
```


```{r}
table_delivery_type("deliver_package","hh_race_category")
```

#### grocery delivery by household race categories

```{r}
#| fig-height: 3
#| fig-width: 12
#| fig-align: center

plot_delivery_type("deliver_grocery","hh_race_category",type="year")
```

```{r}
#| fig-height: 3
#| fig-width: 13
#| fig-align: center

plot_delivery_type("deliver_grocery","hh_race_category",type="cat")
```


```{r}
table_delivery_type("deliver_grocery","hh_race_category")
```

#### food delivery by household race categories

```{r}
#| fig-height: 3
#| fig-width: 12
#| fig-align: center

plot_delivery_type("deliver_food","hh_race_category",type="year")
```

```{r}
#| fig-height: 3
#| fig-width: 13
#| fig-align: center

plot_delivery_type("deliver_food","hh_race_category",type="cat")
```

```{r}
table_delivery_type("deliver_food","hh_race_category")
```

## delivery by home location and type


### home county

#### package delivery by home county

```{r}
#| fig-height: 3
#| fig-width: 8
#| fig-align: center

plot_delivery_type("deliver_package","home_county",type="year")
```

```{r}
#| fig-height: 3
#| fig-width: 8.5
#| fig-align: center

plot_delivery_type("deliver_package","home_county",type="cat")
```

```{r}
table_delivery_type("deliver_package","home_county")
```

#### grocery delivery by home county

```{r}
#| fig-height: 3
#| fig-width: 7.5
#| fig-align: center

plot_delivery_type("deliver_grocery","home_county",type="year")
```

```{r}
#| fig-height: 3
#| fig-width: 8.5
#| fig-align: center

plot_delivery_type("deliver_grocery","home_county",type="cat")
```

```{r}
table_delivery_type("deliver_grocery","home_county")
```

#### food delivery by home county

```{r}
#| fig-height: 3
#| fig-width: 7.5
#| fig-align: center

plot_delivery_type("deliver_food","home_county",type="year")
```

```{r}
#| fig-height: 3
#| fig-width: 8.5
#| fig-align: center

plot_delivery_type("deliver_food","home_county",type="cat")
```


```{r}
table_delivery_type("deliver_food","home_county")
```

### home in centers

#### package delivery by home centers

```{r}
#| layout-ncol: 2
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_package","home_rgc")
```

```{r}
table_delivery_type("deliver_package","home_rgc")
```

#### grocery delivery by home centers

```{r}
#| layout-ncol: 2
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_grocery","home_rgc")
```

```{r}
table_delivery_type("deliver_grocery","home_rgc")
```

#### food delivery by home centers

```{r}
#| layout-ncol: 2
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_food","home_rgc")
```

```{r}
table_delivery_type("deliver_food","home_rgc")
```

### residence type

#### package delivery by residence type

```{r}
#| layout-ncol: 2
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_package","res_type_group")
```

```{r}
table_delivery_type("deliver_food","res_type_group")
```

#### grocery delivery by residence type

```{r}
#| layout-ncol: 2
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_grocery","res_type_group")
```

```{r}
table_delivery_type("deliver_grocery","res_type_group")
```

#### food delivery by residence type

```{r}
#| layout-ncol: 2
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_food","res_type_group")
```

```{r}
table_delivery_type("deliver_food","res_type_group")
```

#### service work delivery by residence type

```{r}
#| layout-ncol: 2
#| fig-height: 3.9
#| fig-width: 5.2


plot_delivery_type("deliver_work","res_type_group")
```

```{r}
table_delivery_type("deliver_work","res_type_group")
```

## shopping trip rates over the years {#sec-trip-rates}

- Compared to be pre-pandemic years, the average weekday shopping rate in 2025 was almost 1/5 lower.

```{r}
#| label: fig-shop-trip-rate
#| fig-cap: "shopping trip rates over years"
#| fig-height: 2.5
#| fig-width: 3


# update travelSurveyTools for correct calculation
# remotes::install_github("RSGInc/travelSurveyTools", ref = "fix_trip_vmt_rate")

df_plot <- psrc_hts_triprate(hts_data, group_vars="dest_purpose_cat") %>%
  filter(dest_purpose_cat == "Shopping")

ggplot(df_plot,
       aes(x=factor(survey_year), y=mean, fill=factor(survey_year))) +
  geom_col(width=0.7) +
  geom_text(aes(label=scales::number(mean, accuracy = 0.1)),
            vjust = -0.3) +
  geom_hline(yintercept = 0, color = "#cbcbcb", linewidth = 0.3) +
  errorbar_switch(geom_errorbar(aes(ymin = mean - mean_moe, ymax = mean + mean_moe), width = 0.1), 
                  on=params$errorbar) +
  scale_y_continuous(labels = scales::number, expand = expansion(mult = c(0, 0.1))) +
  scale_fill_manual(values = psrc_colors$grays_dec) +
  psrc_style() +
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.y = element_blank())
```


