---
title: "movers_why_2021"
author: "suzanne"
date: "3/31/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Recent Movers in 2021

This document looks at why people in moved who answered the survey in 2021. We will group the data into how long ago they moved to try to understand pre and post covid.

To lookup variable names the codebook is here: J:\Projects\Surveys\HHTravel\Survey2021\\Data\\Combined_Codebook_2021_With_Weights.xlsx

We will use the psrc.travesurvey package, documentation is here: <https://psrc.github.io/psrc.travelsurvey/articles/calculate_hhts_summaries.html>

Code is here: <https://github.com/psrc/psrc.travelsurvey>

First, let's understand the universe of respondents on the movers question.


```{r get_started}
library(psrc.travelsurvey)
library(dplyr)
library(stringr)
library(data.table)
library(tidyr)

res_dur_2019<-get_hhts(dyear=2019, level="h", vars="res_dur")
res_dur_2019_counts<- hhts_count(df=res_dur_2019, group_vars='res_dur')


res_dur_2021<-get_hhts(dyear=2021, level="h", vars="res_dur")
res_dur_2021_counts<- hhts_count(df=res_dur_2021, group_vars='res_dur')

res_dur_2019_2021_counts<-merge(res_dur_2019_counts, res_dur_2021_counts, by='res_dur', suffixes =c('2019', '2021'))



write.table(res_dur_2019_2021_counts, "clipboard", sep="\t", row.names=FALSE)

res_dur_2019_2021_counts

```
Differences are generally within MOE across 2019, 2021 surveys. Looks like maybe more in the "less than a year" category but can't tell.

We're really most interested in the movers in the past year vs 1 and 2 years, or 2 and 3 years ago.

Why moved to current residence

```{r moved_recently}
hhs_2021<-get_hhts(dyear=2021, level="h", vars=c("res_dur",
"res_factors_30min",
"res_factors_afford",
"res_factors_closefam",
"res_factors_hwy",
"res_factors_school",
"res_factors_space",
"res_factors_transit",
"res_factors_walk",
"res_factors_cultural"

))



vars_to_summarize<-c(
"res_factors_30min",
"res_factors_afford",
"res_factors_closefam",
"res_factors_hwy",
"res_factors_school",
"res_factors_space",
"res_factors_transit",
"res_factors_walk",
"res_factors_cultural")



mover_sum<- list()
mover_sum <- lapply(vars_to_summarize, FUN=function(x) hhts_count(df=hhs_2021,group_vars=c(x, "res_dur"))) %>%
    lapply(FUN=function(y){mutate(y, "var_name"=colnames(y)[1])}) %>% rbindlist(use.names=FALSE) %>% 
    rename(var_value=colnames(.)[1]) %>% relocate(var_name)%>%filter(res_dur!='Total')

mover_shares<-mover_sum%>% select(var_name, var_value, res_dur, count)%>%pivot_wider(names_from=c(res_dur), values_from=count,names_prefix='res_dur') %>%group_by(var_name)%>%mutate(across(starts_with("res_dur"),~ .x/sum(.x)))%>%filter(var_value=='Very important')%>%select(var_value,var_name,'res_durLess than a year')

write.table(mover_shares, "clipboard", sep="\t", row.names=FALSE)

```

```{r}
hhs_2019<-get_hhts(dyear=2019, level="h", vars=c("res_dur",
"res_factors_30min",
"res_factors_afford",
"res_factors_closefam",
"res_factors_hwy",
"res_factors_school",
"res_factors_space",
"res_factors_transit",
"res_factors_walk"
))



vars_to_summarize<-c(
"res_factors_30min",
"res_factors_afford",
"res_factors_closefam",
"res_factors_hwy",
"res_factors_school",
"res_factors_space",
"res_factors_transit",
"res_factors_walk")



mover_sum<- list()
mover_sum <- lapply(vars_to_summarize, FUN=function(x) hhts_count(df=hhs_2019,group_vars=c(x, "res_dur"))) %>%
    lapply(FUN=function(y){mutate(y, "var_name"=colnames(y)[1])}) %>% rbindlist(use.names=FALSE) %>% 
    rename(var_value=colnames(.)[1]) %>% relocate(var_name)%>%filter(res_dur!='Total')

mover_shares<-mover_sum%>% select(var_name, var_value, res_dur, count)%>%pivot_wider(names_from=c(res_dur), values_from=count,names_prefix='res_dur') %>%group_by(var_name)%>%mutate(across(starts_with("res_dur"),~ .x/sum(.x)))%>%filter(var_value=='Very important')%>%select(var_value,var_name,'res_durLess than a year')

write.table(mover_shares, "clipboard", sep="\t", row.names=FALSE)

```

Why did you move from previous residence



```{r prev-2021}
hhs_2019<-get_hhts(dyear=2019, level="h", vars=c("res_dur",
"prev_res_factors_housing_cost",
"prev_res_factors_income_change",
"prev_res_factors_community_change",
"prev_res_factors_hh_size",
"prev_res_factors_more_space",
"prev_res_factors_less_space",
"prev_res_factors_employment",
"prev_res_factors_covid_flexibility",
"prev_res_factors_school",
"prev_res_factors_crime",
"prev_res_factors_covid_health",
"prev_res_factors_quality",
"prev_res_factors_forced",
"prev_res_factors_no_answer",
"prev_res_factors_other",
"prev_res_factors_specify"

))



vars_to_summarize<-c(
"prev_res_factors_housing_cost",
"prev_res_factors_income_change",
"prev_res_factors_community_change",
"prev_res_factors_hh_size",
"prev_res_factors_more_space",
"prev_res_factors_less_space",
"prev_res_factors_employment",
"prev_res_factors_covid_flexibility",
"prev_res_factors_school",
"prev_res_factors_crime",
"prev_res_factors_covid_health",
"prev_res_factors_quality",
"prev_res_factors_forced",
"prev_res_factors_no_answer",
"prev_res_factors_other",
"prev_res_factors_specify")



mover_sum<- list()
mover_sum <- lapply(vars_to_summarize, FUN=function(x) hhts_count(df=hhs_2019,group_vars=c(x, "res_dur"))) %>%
    lapply(FUN=function(y){mutate(y, "var_name"=colnames(y)[1])}) %>% rbindlist(use.names=FALSE) %>% 
    rename(var_value=colnames(.)[1]) %>% relocate(var_name)%>%filter(res_dur!='Total')%>%drop_na()

mover_shares<-mover_sum%>% select(var_name, var_value, res_dur, count)%>%pivot_wider(names_from=c(res_dur), values_from=count,names_prefix='res_dur') %>%group_by(var_name)%>%mutate(across(starts_with("res_dur"),~ .x/sum(.x)))%>%select(var_value,var_name,'res_durLess than a year')%>%filter(var_value=='Selected')

write.table(mover_shares, "clipboard", sep="\t", row.names=FALSE)



```

```{r prev-2021}
hhs_2021<-get_hhts(dyear=2021, level="h", vars=c("res_dur",
"prev_res_factors_housing_cost",
"prev_res_factors_income_change",
"prev_res_factors_community_change",
"prev_res_factors_hh_size",
"prev_res_factors_more_space",
"prev_res_factors_less_space",
"prev_res_factors_employment",
"prev_res_factors_covid_flexibility",
"prev_res_factors_school",
"prev_res_factors_crime",
"prev_res_factors_covid_health",
"prev_res_factors_quality",
"prev_res_factors_forced",
"prev_res_factors_no_answer",
"prev_res_factors_other",
"prev_res_factors_specify"

))



vars_to_summarize<-c(
"prev_res_factors_housing_cost",
"prev_res_factors_income_change",
"prev_res_factors_community_change",
"prev_res_factors_hh_size",
"prev_res_factors_more_space",
"prev_res_factors_less_space",
"prev_res_factors_employment",
"prev_res_factors_covid_flexibility",
"prev_res_factors_school",
"prev_res_factors_crime",
"prev_res_factors_covid_health",
"prev_res_factors_quality",
"prev_res_factors_forced",
"prev_res_factors_no_answer",
"prev_res_factors_other",
"prev_res_factors_specify")



mover_sum<- list()
mover_sum <- lapply(vars_to_summarize, FUN=function(x) hhts_count(df=hhs_2021,group_vars=c(x, "res_dur"))) %>%
    lapply(FUN=function(y){mutate(y, "var_name"=colnames(y)[1])}) %>% rbindlist(use.names=FALSE) %>% 
    rename(var_value=colnames(.)[1]) %>% relocate(var_name)%>%filter(res_dur!='Total')%>%drop_na()

mover_shares<-mover_sum%>% select(var_name, var_value, res_dur, count)%>%pivot_wider(names_from=c(res_dur), values_from=count,names_prefix='res_dur') %>%group_by(var_name)%>%mutate(across(starts_with("res_dur"),~ .x/sum(.x)))%>%select(var_value,var_name,'res_durLess than a year')%>%filter(var_value=='Selected')

write.table(mover_shares, "clipboard", sep="\t", row.names=FALSE)

```



```{r get_started}

displaced_hh_2019<-get_hhts(dyear=c(2017,2019), level="h", vars="res_factors_displaced")
displaced_hh_2019_counts<- hhts_count(df=displaced_hh_2019, group_vars='res_factors_displaced')


displaced_hh_2021<-get_hhts(dyear=2021, level="h", vars="res_factors_displaced")
displaced_hh_2021_counts<- hhts_count(df=displaced_hh_2021, group_vars='res_factors_displaced')



displaced_2017_2019<-merge(displaced_hh_2019_counts, displaced_hh_2021_counts, by='res_factors_displaced', suffixes =c('2019', '2021'))

write.table(displaced_hh_2021_counts, "clipboard", sep="\t", row.names=FALSE)
displaced_2017_2019
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
library(psrc.travelsurvey)
library(dplyr)
carshare_hh_2019<-get_hhts(dyear=2019, level="h", vars="car_share")
carshare_hh_2019_counts<- hhts_count(df=carshare_hh_2019, group_vars='car_share')
```


```{r pressure, echo=FALSE}
hhs_2021 <- get_hhts(dyear=2021, level="h", vars=c("res_dur", unlist(vars_to_summarize)) %>%
    mutate(covid_mover=factor(case_when(res_dur=="Less than a year" ~1, !is.na(res_dur) ~0)))
df <- list()
df <- lapply(my_vars, FUN=function(x) hhts_count(df=hhs_2021,group_vars=c(x, "covid_mover"))) %>%
    lapply(FUN=function(y){mutate(y, "var_name"=colnames(y)[1])}) %>% rbindlist(use.names=FALSE) %>% 
    rename(var_value=colnames(.)[1]) %>% relocate(var_name)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
