---
title: "more-transit"
author: "suzanne"
date: "2024-09-16"
output: html_document
---

A few more questions came up about transit for TOC and TOD.

Transit Access Mode
- 2023
and trend
Transit Trips by purpose
-2023
and trend


### Read in packages

```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
library(forcats)
library(sf)
```

### Helper functions
```{r}

# get new labels associated with a group  
get_grouped_labels <- function(.value_labels = value_labels, group_id, group_name) {
  group_title <- paste0(group_id, '_title')
  group_value <- paste0(group_id, '_value')
  
  cols <- c('label', group_value)
  group_labels <- .value_labels[get(eval(group_title)) == group_name, ..cols]
  group_labels<-group_labels%>%distinct(label, .keep_all=TRUE)
  return(group_labels)
}


# function for grouping variable and adding it to the values list
add_values_code <- function(.group_labels = group_labels, .value_labels = value_labels, group_name) {

  names_labels <- names(.group_labels)
  group_id_value <- names_labels[grepl('group', names_labels)]

  v <- unique(.group_labels, by = group_id_value)
  v <- v[, value := .I][, .(variable = group_name, value, label = get(eval(group_id_value)))]

  new_cols <- map( c('title', 'value'), ~paste('group', 1:3, .x, sep = '_')) |> unlist()
  v[, (new_cols) := NA]

  new_val_labels <- rbindlist(list(.value_labels, v), fill=TRUE)
  new_val_labels[, val_order := .I]
  
  return(new_val_labels) 
}

grp_to_tbl <- function(tbl, .group_labels = group_labels, ungrouped_name, grouped_name) {
  tbl <- left_join(tbl, .group_labels, by = setNames('label', ungrouped_name))
  #find what the group id is based on the group_labels
  # find the group text based on the group_labels
  names_labels <- names(.group_labels)
  group_id_value <- names_labels[grepl('group', names_labels)]
  
  setnames(tbl, group_id_value, grouped_name)
  return(tbl)
}

order_factors<-function(tbl, variable_name, value_labels){
  var_val_labels<-value_labels%>%filter(variable==variable_name)
  tbl<-tbl%>%left_join(var_val_labels, by=join_by(!!sym(variable_name)==label))%>%
    arrange(val_order)%>%
    mutate({{variable_name}}:=factor(!!sym(variable_name), levels=unique(!!sym(variable_name))))
  return(tbl)
}

summarize_weighted <- function(hts_data, summarize_var, summarize_by, id_cols, wt_cols,wtname,summarize_vartype='categorical'){
  
  
  prepped_dt <- hts_prep_variable(summarize_var = summarize_var,
                                  summarize_by = summarize_by,
                                  data = hts_data,
                                  id_cols=id_cols,
                                  wt_cols=wt_cols,
                                  weighted=TRUE,
                                  remove_missing=TRUE)

  
  if(summarize_vartype=='categorical'){
      summary<-hts_summary(prepped_dt = prepped_dt$cat,
                       summarize_var = summarize_var,
                       summarize_by = summarize_by,
                       summarize_vartype = summarize_vartype,
                       id_cols= id_cols,
                       wtname=wtname,
                       weighted=TRUE,
                       se=TRUE)
  }else{
    summary<-hts_summary(prepped_dt = prepped_dt$num,
                         summarize_var = summarize_var,
                         summarize_by = summarize_by,
                         summarize_vartype = summarize_vartype,
                         id_cols= id_cols,
                         wtname=wtname,
                         weighted=TRUE,
                         se=TRUE)
    
  }
  
  
  return(summary)
}

add_variable<-function(variable_list,variable_name, table_name, data_type='integer/categorical'){

  new_var_tbl<-
  data.table(
    variable = variable_name,
    is_checkbox = 0,
    hh = 0,
    person = 0,
    day = 0,
    trip = 0,
    vehicle = 0,
    location = 0,
    description = variable_name,
    logic = variable_name,
    data_type =data_type,
    shared_name = variable_name
  )
  
  new_var_tbl<-new_var_tbl%>%mutate({{table_name}}:=1)
  print(new_var_tbl)
  variable_list<-rbind(variable_list, new_var_tbl)
  
  
}


# Add associated values to value table
new_value_tbl <- function(variable_name_list, variable_value_list,order_start = 0){
  
  len <- length(variable_name_list)-1
  
  add_value_tbl <- data.frame(variable = variable_name_list,
                              value = c(0:len),
                              label = variable_value_list,
                              val_order = c(order_start:(order_start+len)),
                              group_1_title = NA,
                              group_1_value = NA,
                              group_2_title = NA,
                              group_2_value = NA,
                              group_3_title = NA,
                              group_3_value = NA)
}

# Add variables from existing grouping
get_var_grouping <- function(value_tbl, group_number, grouping_name){
  
  group_title <- paste0("group_",group_number,"_title")
  group_value <- paste0("group_",group_number,"_value")
  #TODO: fix value order
  value_order_start <- max(value_tbl[,c("val_order")]) +1
  
  grouping_tbl <- value_tbl %>% 
    filter(!!sym(group_title) == grouping_name)
  
  variable_name <- unname(unlist(grouping_tbl[1,c("variable")]))
  group_name <- unname(unlist(grouping_tbl[1,c(group_title)]))
  
  grouping_value <- grouping_tbl %>%
    select(all_of(c("label", group_value))) %>%
    rename(!!variable_name := label,
           !!grouping_name  := !!sym(group_value))
  distinct_value <- grouping_tbl %>%
    select(all_of(c(group_title, group_value))) %>% 
    distinct()
  
  add_value_tbl <- new_value_tbl(distinct_value[[group_title]],distinct_value[[group_value]], value_order_start)
  
  final <- list(add_value_tbl,grouping_value)
  
  return(final)
}

# Add custom variable 
create_custom_variable <- function(value_tbl, variable_name,label_vector){
  
  value_order_start <- max(value_tbl[,c("val_order")]) +1
  add_value_tbl <- new_value_tbl(rep(variable_name, times=length(label_vector)),label_vector, value_order_start)
  
  return(add_value_tbl)
}

add_variable_to_data <- function(hts_data, value_map) {
  
  ungroup_name <- names(value_map)[1]
  group_name <- names(value_map)[2]
  
  tbl <- hts_data %>% left_join(value_map, by = ungroup_name)
  
  return(tbl)
}




```



```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhsize, hhincome_broad,home_jurisdiction, home_county, hh_weight,vehicle_count, home_rgcname, home_lat, home_lng, survey_year, hh_race_category
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category,
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, sexuality, industry,office_available, commute_freq, education, work_lat, work_lng , work_rgcname, work_jurisdiction, work_county
                from HHSurvey.v_persons_labels")
                
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,mode_acc,
person_id, mode_1, dest_purpose, survey_year, trip_weight, distance_miles, dest_rgcname, mode_characterization, travelers_total 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)
setDT(day)
setDT(trip)
```

#  Pre-processing
#race_category needs to be added to the codebook

# to do: add Seattle, Bellevue geographies

## Read in Codebook
```{r pressure, echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```

## Set IDs as characters
TO DO: functionalize convert all ids to characters, or store as characters upfront
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
day[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
day[, person_id := as.character(person_id)]
trip[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
trip[, day_id := as.character(day_id)]

trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
day <- day%>%mutate(survey_year=as.character(survey_year))
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```







```{r}
variable_list<-add_variable(variable_list, 'telecommute_freq_cond','person')
group_labels<-get_grouped_labels(group_id='group_2', group_name='telecommute_freq_cond')
value_labels<-add_values_code(group_name='telecommute_freq_cond')
person<-grp_to_tbl(tbl=person, ungrouped_name='telecommute_freq', grouped_name='telecommute_freq_cond')
```

```{r}
variable_list<-add_variable(variable_list, 'telecommute_freq_broad','person')
group_labels<-get_grouped_labels(group_id='group_3', group_name='telecommute_freq_broad')
value_labels<-add_values_code(group_name='telecommute_freq_broad')
person<-grp_to_tbl(tbl=person, ungrouped_name='telecommute_freq', grouped_name='telecommute_freq_broad')
```

```{r}
variable_list<-add_variable(variable_list, 'commute_freq_simple','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='commute_freq_simple')
value_labels<-add_values_code(group_name='commute_freq_simple')
person<-grp_to_tbl(tbl=person, ungrouped_name='commute_freq', grouped_name='commute_freq_simple')
```
```{r}
variable_list<-add_variable(variable_list, 'industry_cond','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='industry_cond')
value_labels<-add_values_code(group_name='industry_cond')
person<-grp_to_tbl(tbl=person, ungrouped_name='industry', grouped_name='industry_cond')
```
```{r}
variable_list<-add_variable(variable_list, 'industry_usim','person')
group_labels<-get_grouped_labels(group_id='group_2', group_name='industry_usim')
value_labels<-add_values_code(group_name='industry_usim')
person<-grp_to_tbl(tbl=person, ungrouped_name='industry', grouped_name='industry_usim')
```


```{r}
variable_list<-add_variable(variable_list, 'age_grp_2','person')
group_labels<-get_grouped_labels(group_id='group_3', group_name='age_grp_2')
value_labels<-add_values_code(group_name='age_grp_2')
person<-grp_to_tbl(tbl=person, ungrouped_name='age', grouped_name='age_grp_2')
```

```{r}
variable_list<-add_variable(variable_list, 'home_county','hh')

value<- c(0,1,2,3)
label=c('Kitsap', 'Pierce', 'Snohomish', 'King')
workgeo_val_labels<-data.frame(value, label)
workgeo_val_labels<-workgeo_val_labels%>%mutate(variable='home_county')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, workgeo_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  


```


```{r}
variable_list<-add_variable(variable_list, 'rgc_cat_home','hh')

value<- c(0,1,2,3)
label=c('NotRGC', 'MetroRGC', 'UrbanRGC', 'Missing')
homegeo_val_labels<-data.frame(value, label)
homegeo_val_labels<-homegeo_val_labels%>%mutate(variable='rgccat_home')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, homegeo_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  


```

```{r}
uga.lyr<-st_read_elmergeo('URBAN_GROWTH_AREA_evw')
```


```{r}
center.lyr <- st_read_elmergeo('URBAN_CENTERS_evw')
```



```{r}
hh_geo<-hh%>%
  st_as_sf(coords=c('home_lng','home_lat'), crs=4326)
```


```{r}
hh_in_uga<-st_filter(hh_geo, uga.lyr)
```

```{r}
hh_in_uga<-hh_in_uga%>%mutate(in_uga=1)
```

```{r}
hh<-left_join(hh, hh_in_uga)
hh<-hh%>%  mutate(in_uga=if_else(is.na(in_uga), 0, 1))
```




```{r}
hh<-left_join(hh, center.lyr, by=join_by(home_rgcname==name), suffix= c("", '.origin'))
```








```{r }
hh<-hh%>%mutate(rgc_cat_home= 
                      case_when(is.na(category) ~ 'NotRGC',
                                category=='Metro' ~ 'MetroRGC',
                                category=='Urban' ~ 'UrbanRGC',
                                TRUE ~'Missing')
)
variable_list<-add_variable(variable_list, 'rgc_cat_home','hh')
```
```{r}
variable_list<-add_variable(variable_list, 'homegeog','hh')

value<- c(0,1,2,3,4,5)
label=c('Kitsap', 'Pierce', 'Seattle Outside Downtown', 'Bellevue', 'Seattle Downtown', 'Snohomish-King Suburban' )
homegeo_val_labels<-data.frame(value, label)
homegeo_val_labels<-homegeo_val_labels%>%mutate(variable='homegeog')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, homegeo_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  


```

```{r}


hh<-hh%>%mutate(homegeog=case_when(
  home_rgcname=='Seattle Downtown'~ 'Seattle Downtown',
  (home_rgcname!='Seattle Downtown' & home_jurisdiction=='Seattle') ~ 'Seattle Outside Downtown',
   home_jurisdiction=='Bellevue' ~ 'Bellevue',
  ((home_county=='King' &!home_jurisdiction %in% c('Seattle', 'Bellevue'))|home_county=='Snohomish')~'Snohomish-King Suburban',
  .default=home_county
))
```


```{r}
variable_list<-add_variable(variable_list, 'gender_grp','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='gender_grp')
value_labels<-add_values_code(group_name='gender_group')
person<-grp_to_tbl(tbl=person, ungrouped_name='gender', grouped_name='gender_grp')
```

```{r}
variable_list<-add_variable(variable_list, 'telecommute_freq_simple','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='telecommute_freq_simple')
value_labels<-add_values_code(group_name='telecommute_freq_simple')
person<-grp_to_tbl(tbl=person, ungrouped_name='telecommute_freq', grouped_name='telecommute_freq_simple')%>%
  mutate(telecommute_freq_simple=gsub("[\r\n]","", telecommute_freq_simple))# there's some weird characters in there i can't figure out how to remove
```
```{r}
variable_list<-add_variable(variable_list, 'edu_simple','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='edu_simple')
value_labels<-add_values_code(group_name='edu_simple')
person<-grp_to_tbl(tbl=person, ungrouped_name='education', grouped_name='edu_simple')
```

```{r}
variable_list<-add_variable(variable_list, 'sexuality_broad','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='sexuality_broad')
value_labels<-add_values_code(group_name='sexuality_broad')
person<-grp_to_tbl(tbl=person, ungrouped_name='sexuality', grouped_name='sexuality_broad')
```
```{r}
variable_list<-add_variable(variable_list, 'vehicle_count_simple','hh')
group_labels<-get_grouped_labels(group_id='group_1', group_name='vehicle_count_simple')
value_labels<-add_values_code(group_name='vehicle_count_simple')
hh<-grp_to_tbl(tbl=hh, ungrouped_name='vehicle_count', grouped_name='vehicle_count_simple')
```
```{r}
variable_list<-add_variable(variable_list, 'has_vehicle','hh')
group_labels<-get_grouped_labels(group_id='group_2', group_name='has_vehicle')
value_labels<-add_values_code(group_name='has_vehicle')
hh<-grp_to_tbl(tbl=hh, ungrouped_name='vehicle_count', grouped_name='has_vehicle')
```


```{r}
variable_list<-add_variable(variable_list, 'telecommute_status', 'person')
variable_list<-add_variable(variable_list, 'telecommute_broad', 'person')
variable_list<-add_variable(variable_list, 'telecommute_status_broad', 'person')
variable_list<-add_variable(variable_list, 'commute_freq_all', 'person')

```

# add necessary variables to the persons table
# update telecommute_freq_cond to also include at home people






```{r}
value<- c(0,3)
label=c('Under $75,000', '$75,000-$199,999')
inc_val_labels<-data.frame(value, label)
inc_val_labels<-inc_val_labels%>%mutate(variable='hhincome_broad')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, inc_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels 
hh<- order_factors(hh, 'hhincome_broad', value_labels)
```

```{r}
person<- order_factors(person, 'age', value_labels)
person<-person%>%mutate(telecommute_freq_cond=case_when(
                  workplace=='At home (telecommute or self-employed with home office)'
                  ~ '1+ days per week',
                  .default=telecommute_freq_cond))

```



```{r}

value<- c(0,1,2,3)
label=c('Not Worker', 'Fully at Home', 'Hybrid', 'Fully in Person')
tele_val_labels<-data.frame(value, label)
tele_val_labels<-tele_val_labels%>%mutate(variable='telecommute_status')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, tele_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  


```


```{r}

value<- c(0,1,2,3,4)
label=c('Not Worker', 'Fully at Home', 'Mostly In Person', 'Mostly at Home', 'Fully in Person')
tele_val_labels<-data.frame(value, label)
tele_val_labels<-tele_val_labels%>%mutate(variable='telecommute_status_broad')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, tele_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  


```

```{r}

value<- c(0,1,2,3, 4, 5)
label=c('Never or less than weekly', '1 days', '2 days', '3 days', '4 days', '5+ days')
val_labels<-data.frame(value, label)
val_labels<-tele_val_labels%>%mutate(variable='commute_freq_all')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, tele_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  

```


More detailed variable definitions
```{r}
person<-person%>%mutate(telecommute_status=
        case_when(workplace %in%c('Missing: Skip logic', "Missing: Skip Logic") 
                  ~ 'Not Worker',
                  (is.na(workplace)) ~ 'Not Worker',
                  (workplace=='At home (telecommute or self-employed with home office)' )
                  ~ 'Fully At Home',
                  (telecommute_freq_cond == '1+ days per week')
                  ~'Hybrid',
                  .default = 'Fully In Person'
        )
)

```




```{r}
person<-person%>%mutate(telecommute_status_broad=
        case_when(
                  (telecommute_freq_broad == '1-2 days')
                  ~'Mostly In Person',
                  (telecommute_freq_broad == '3+ days')
                  ~'Mostly At Home',
                  .default = telecommute_status
        )
)

```

How much are people commuting?
Combine At home with other workers

```{r}
person<-person%>%mutate(commute_freq_all=
                           if_else(workplace== "At home (telecommute or self-employed with home office)", 
                                   'Never or less than weekly', commute_freq))



person<- person%>%mutate(commute_freq_all=case_match(commute_freq_all,
                               "1-3 days a month" ~ "Never or less than weekly",
                               'Never' ~ "Never or less than weekly",
                                "A few times per month"   ~  "Never or less than weekly",  
                                "Less than monthly"   ~  "Never or less than weekly",  
                               "None, did not commute to workplace last week" ~"Never or less than weekly",
                                '5 days a week' ~ '5+ days',
                                '6-7 days a week' ~ '5+ days',
                               .default= commute_freq_all
                               ))


person<- person%>%mutate(commute_freq_all=str_replace(commute_freq_all, ' a week', ''))
                               
person<- person%>%mutate(commute_freq_all=case_when(
  commute_freq_all %in% c('1 day', '2 days') ~ '1-2 days',
  commute_freq_all %in% c('3 days', '4 days') ~ '3-4 days',
  .default = commute_freq_all
  
))  

#person<- person%>%mutate(commute_freq_all=factor(commute_freq_all, levels=c("Never or less than #weekly", '1-2 days', '3-4 days', '5+ days')))


```


```{r}
person<-person%>%mutate(telecommute_status=
        case_when(workplace %in%c('Missing: Skip logic', "Missing: Skip Logic") 
                  ~ 'Not Worker',
                  (is.na(workplace)) ~ 'Not Worker',
                  (workplace=='At home (telecommute or self-employed with home office)' )
                  ~ 'Fully At Home',
                  (telecommute_freq_cond == '1+ days per week')
                  ~'Hybrid',
                  .default = 'Fully In Person'
        )
)

```




```{r}
person<-person%>%mutate(telecommute_status_broad=
        case_when(
                  (telecommute_freq_broad == '1-2 days')
                  ~'1-2 days',
                  (telecommute_freq_broad == '3+ days')
                  ~'3-4 days',
                  .default = telecommute_status
        )
)

```


```{r}
variable_list<-add_variable(variable_list, 'dest_purpose_simpler','trip')
group_labels<-get_grouped_labels(group_id='group_1', group_name='dest_purpose_simpler')
value_labels<-add_values_code(group_name='dest_purpose_simpler')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='dest_purpose', grouped_name='dest_purpose_simpler')
```

```{r}
variable_list<-add_variable(variable_list, 'dest_purpose_simplest','trip')
group_labels<-get_grouped_labels(group_id='group_2', group_name='dest_purpose_simplest')
value_labels<-add_values_code(group_name='dest_purpose_simplest')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='dest_purpose', grouped_name='dest_purpose_simplest')
```


```{r}
variable_list<-add_variable(variable_list, 'workgeog','person')

value<- c(0,1,2,3,4,5)
label=c('Kitsap', 'Pierce', 'Seattle Outside Downtown', 'Bellevue', 'Seattle Downtown', 'Snohomish-King Suburban' )
workgeo_val_labels<-data.frame(value, label)
workgeo_val_labels<-workgeo_val_labels%>%mutate(variable='workgeog')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, workgeo_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  


```
```{r}
variable_list<-add_variable(variable_list, 'work_county','person')

value<- c(0,1,2,3)
label=c('Kitsap', 'Pierce', 'Snohomish', 'King')
workgeo_val_labels<-data.frame(value, label)
workgeo_val_labels<-workgeo_val_labels%>%mutate(variable='work_county')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, workgeo_val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  


```



```{r}
workers<-person%>%filter(!workplace %in% c('Missing: Skip logic', "Missing: Skip Logic"))%>%drop_na(workplace)
worker_list<-list(hh=hh, person=workers)

```

```{r}

workers<-workers%>%mutate(workgeog=case_when(
  work_rgcname=='Seattle Downtown' ~ 'Seattle Downtown',
  (work_rgcname!='Seattle Downtown' &work_jurisdiction=='Seattle') ~ 'Seattle Outside Downtown',
  work_jurisdiction=='Bellevue' ~ 'Bellevue',
  ((work_county=='King' &!work_jurisdiction %in% c('Seattle', 'Bellevue'))|work_county=='Snohomish')~'Snohomish-King Suburban',
  .default=work_county
))
```


# Transit Analysis Start
```{r setup, include=FALSE}

hh_data<-
  list(hh=hh, person=person, day=day, trip=trip)
ids = c('hh_id', 'person_id', 'day_id', 'trip_id')
wts = c('hh_weight', 'person_weight', 'day_weight', 'trip_weight')
```



```{r}
variable_list<-add_variable(variable_list, 'mode_acc_1','trip')
group_labels<-get_grouped_labels(group_id='group_1', group_name='mode_acc_1')
value_labels<-add_values_code(group_name='mode_acc_1')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='mode_acc', grouped_name='mode_acc_1')
```

```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```

```{r}
transit_trips<-trip%>%filter(mode_characterization=='Transit')#%>%filter(!mode_acc_1 %in% c("Missing", #"Transfer"))%>%drop_na(mode_acc_1)
## very many missing data points...



```

```{r}
access_mode<-transit_trips%>%group_by(survey_year, mode_acc, mode_characterization)%>%summarize(n=n())
write.csv(access_mode, 'access_mode.csv')
```
```{r}
hh_transit_data<-
  list(hh=hh, person=person, day=day, trip=transit_trips)
ids = c('hh_id', 'person_id', 'day_id', 'trip_id')
wts = c('hh_weight', 'person_weight', 'day_weight', 'trip_weight')
```




```{r}
output <- summarize_weighted(hts_data= hh_transit_data,
                               summarize_var = 'mode_acc_1',
                               summarize_by = c('survey_year'),
                               id_cols=ids,
                               wt_cols=wts,
                               wtname= 'trip_weight'
                               )
```


```{r}

output_summary<-output$summary$wtd%>%mutate(moe=prop_se*1.645)%>%filter(!is.na(mode_acc_1))
static<-static_column_chart(output_summary, y='prop', x='survey_year', fill='mode_acc_1', color='pognbgy_10', pos='stack') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))


static
write.csv(output_summary, 'mode_access.csv')
```

```{r}
output <- summarize_weighted(hts_data= hh_transit_data,
                               summarize_var = 'mode_acc_1',
                               summarize_by = c('survey_year', 'dest_purpose_simpler'),
                               id_cols=ids,
                               wt_cols=wts,
                               wtname= 'trip_weight'
                               )
```


```{r}

output_summary<-output$summary$wtd%>%mutate(moe=prop_se*1.645)%>%filter(survey_year=='2023')
static<-static_column_chart(output_summary, y='prop', x='dest_purpose_simpler', fill='mode_acc_1', color='pognbgy_10', pos='stack') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))


static
write.csv(output_summary, 'mode_purpose_access.csv')
```
```{r}
output <- summarize_weighted(hts_data= hh_transit_data,
                               summarize_var = 'mode_acc_1',
                               summarize_by = c('survey_year', 'homegeog'),
                               id_cols=ids,
                               wt_cols=wts,
                               wtname= 'trip_weight'
                               )
```


```{r}

output_summary<-output$summary$wtd%>%mutate(moe=prop_se*1.645)%>%filter(survey_year=='2023')
static<-static_column_chart(output_summary, y='prop', x='homegeog', fill='mode_acc_1', color='pognbgy_10', pos='stack') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))


static

write.csv(output_summary, 'mode_access_geog.csv')
```

```{r}
transit_trips<-trip%>%filter(mode_characterization=='Transit')
```

```{r}
hh_transit_data<-
  list(hh=hh, person=person, day=day, trip=transit_trips)
ids = c('hh_id', 'person_id', 'day_id', 'trip_id')
wts = c('hh_weight', 'person_weight', 'day_weight', 'trip_weight')
```


```{r}
output <- summarize_weighted(hts_data= hh_transit_data,
                               summarize_var = 'dest_purpose_simplest',
                               summarize_by = c('survey_year'),
                               id_cols=ids,
                               wt_cols=wts,
                               wtname= 'trip_weight'
                               )
```
# Percent of transit trips by broad purpose
```{r}

output_summary<-output$summary$wtd%>%mutate(moe=prop_se*1.645)%>%filter(dest_purpose_simplest!='Missing')
static<-static_column_chart(output_summary, y='prop', x='survey_year', fill='dest_purpose_simplest', color='pognbgy_10', pos='stack') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))


static

write.csv(output_summary, 'mode_access_broad_purpose.csv')
```
# Percent of transit trips by broad purpose
```{r}

output_summary<-output$summary$wtd%>%mutate(moe=prop_se*1.645)%>%filter(dest_purpose_simplest!='Missing')
static<-static_column_chart(output_summary, y='est', x='survey_year', fill='dest_purpose_simplest', color='pognbgy_10', pos='stack') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))


static
write.csv(output_summary, 'mode_broad_purpose_count_access.csv')
```

