---
title: "Transit Use"
author: "Kristin Mitchell"
format: html
editor: visual
---

# Travel Mode: Transit

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
library(ggplot2)

source("C:/temp/2023/summary/survey-23-preprocess.R")
```

## Read in codebook

```{r codebook}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```

## Read in data from Elmer

```{r elmer_data}
hh <- get_query(sql = "SELECT household_id as hh_id, num_trips AS h_num_trips, hhsize, vehicle_count, hhincome_broad,
                       home_rgcname, home_county, survey_year, hh_weight
                       FROM HHSurvey.v_households_labels;")

person <- get_query(sql = "SELECT person_id, household_id as hh_id, num_trips AS p_num_trips, age, gender, race_category, disability_person, commute_subsidy_1, industry,
                           employment, workplace, transit_freq, commute_freq, commute_mode, survey_year, person_weight
                           FROM HHSurvey.v_persons_labels;")

trip <- get_query(sql = "SELECT trip_id, household_id as hh_id, person_id, origin_x_coord, origin_y_coord, origin_county, origin_rgcname,
                         dest_x_coord, dest_y_coord, dest_county, dest_rgcname, distance_miles, duration_minutes,
                         origin_purpose_cat, origin_purpose, dest_purpose_cat, dest_purpose,
                         mode_type, mode_1, mode_characterization, mode_simple, travelers_total, travelers_hh,
                         survey_year, trip_weight 
                         FROM HHSurvey.v_trips_labels
                         WHERE travel_dow NOT IN ('Friday', 'Saturday', 'Sunday');")

setDT(hh)
setDT(person)
setDT(trip)
```

## Set IDs as character type

```{r set_ids}
hh[, hh_id := as.character(hh_id)]

person[, hh_id := as.character(hh_id)]
person[, person_id := as.character(person_id)]

trip[, hh_id := as.character(hh_id)]
trip[, person_id := as.character(person_id)]
trip[, trip_id := as.character(trip_id)]

hh[, survey_year := as.character(survey_year)]
person[, survey_year := as.character(survey_year)]
trip[, survey_year := as.character(survey_year)]
```

## Get unique modes from trips

```{r trip_modes}
trip_modes <- unique(trip[, c("mode_1", "mode_type", "mode_characterization", "mode_simple")])

# add mode_simple to variables list
variable_list <- add_variable(variable_list, "mode_simple", "trip")
group_labels <- get_grouped_labels(group_id = "group_1", group_name = "mode_simple")
value_labels <- add_values_code(group_name = "mode_simple")
```

Based on the modes in the trip data, will need to filter on `mode_characterization` to remove "Airplane"

## Remove Airplane

```{r}
trip <- filter(trip, mode_characterization != "Airplane")
```

## Adding a new variable to the codebook

make a function with bunch of default #to do make this easier

```{r}

variable_list<-rbind(
    variable_list,
    data.table(
      variable = "mode_characterization",
      is_checkbox = c(0,0),
      hh = c(0,0),
      person = c(0,0),
      day = c(0,0),
      trip = c(1,1),
      vehicle = c(0,0),
      location = c(0,0),
      description = "mode aggregation",
      logic = "mode aggregation",
      data_type = "integer/categorical",
      shared_name = "mode_characterization"
    )
  )

```

# Add associated values

```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='mode_characterization')
value_labels<-add_values_code(group_name='mode_characterization')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='mode_char1', grouped_name='mode_characterization')
```

```{r}

value<- c(0,1,2,3,4,5)
label=c('SOV', 'HOV', 'Walk', 'Transit', 'Bike/Micromobility')
val_labels<-data.frame(value, label)
val_labels<-val_labels%>%mutate(variable='mode_drive')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  

```

# add mode_drive to the trip table

```{r}
trip<- trip%>%mutate(travelers_total_fix= ifelse(travelers_total!='1 traveler', 'More than 1', '1 traveler'))
trip<-trip%>%mutate(mode_simple= replace_na(mode_simple, 'Drive'))%>%
mutate(mode_w_sov=case_when(
  mode_simple=="Drive"& travelers_total=='1 traveler' ~ 'SOV',
  is.na(travelers_total) ~ 'SOV',
  mode_simple=="Drive"& travelers_total!='1 traveler'~  'HOV',
  .default= mode_simple
))
```

## Get transit_freq

```{r}
# add transit_freq to variables list

variable_list <- add_variable(variable_list, "transit_freq", "person") 
group_labels <- get_grouped_labels(group_id = "group_1", group_name = "transit_freq") 
value_labels <- add_values_code(group_name = "transit_freq") 
person<-grp_to_tbl(tbl=person, ungrouped_name="transit_freq", grouped_name="transit_freq")

# add transit_freq_grp
variable_list <- add_variable(variable_list, "transit_freq_grp", "person") 
group_labels <- get_grouped_labels(group_id = "group_1", group_name = "transit_freq_grp") 
value_labels <- add_values_code(group_name = "transit_freq_grp") 
person<-grp_to_tbl(tbl=person, ungrouped_name="transit_freq", grouped_name="transit_freq_grp")

```

## Get dest_purpose_cat

```{r}

# add dest_purpose_simple to variables list

variable_list <- add_variable(variable_list, "dest_purpose_cat", "trip")  
group_labels <- get_grouped_labels(group_id = "group_1", group_name = "dest_purpose_cat") 
value_labels <- add_values_code(group_name = "dest_purpose_cat")  
trip<-grp_to_tbl(tbl=trip, ungrouped_name="dest_purpose_cat", grouped_name="dest_purpose_cat") 

```

## Get Industry

```{r}
# add mode_simple to variables list 
variable_list <- add_variable(variable_list, "industry_cond", "person")
group_labels <- get_grouped_labels(group_id = "group_1", group_name = "industry_cond")
value_labels <- add_values_code(group_name = "industry_cond")
person<-grp_to_tbl(tbl=person, ungrouped_name="industry", grouped_name="industry_cond")

```

## Get Commute Pass

```{r}

# add commute_subsidy_1 to variables list  
variable_list <- add_variable(variable_list, "commute_subsidy_1", "person")
group_labels <- get_grouped_labels(group_id = "group_1", group_name = "commute_subsidy_1") 
value_labels <- add_values_code(group_name = "commute_subsidy_1") 
person<-grp_to_tbl(tbl=person, ungrouped_name="commute_subsidy_1", grouped_name="commute_subsidy_1")
```

## Summaries

```{r age_cond_variable}

#remove duplicates
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)

variable_list <- add_variable(variable_list, 
                              variable_name = "age_condensed",
                              table_name = "person", 
                              data_type = "character")  

person <- person %>%    
  mutate(age_condensed = case_when(age %in% c("Under 5 years old", "5-11 years", "12-15 years", "16-17 years") ~ "Under 18 years old", age %in% c("18-24 years", "25-34 years") ~ "18-34 years", age %in% c("35-44 years", "45-54 years", "55-64 years") ~ "35-64 years", age %in% c("65-74 years", "75-84 years", "85 or years older") ~ "65 years or older")) %>% mutate(age_condensed = factor(age_condensed, levels = c("Under 18 years old", "18-34 years", "35-64 years", "65 years or older")))
```

```{r mode_condensed_variable}

variable_list <- add_variable(variable_list, 
                              variable_name = "mode_condensed",
                              table_name = "trip", 
                              data_type = "character")  

trip <- trip %>%   
   mutate(mode_condensed = case_when(mode_characterization %in% c("Drive HOV2", "Drive HOV3+", "Drive SOV") ~ "Drive", mode_characterization %in% c("Bike/Micromobility") ~ "Bike/Micromobility", mode_characterization %in% c("Transit") ~ "Transit", mode_characterization %in% c("Walk") ~ "Walk")) %>% mutate(mode_condensed = factor(mode_condensed, levels = c("Bike/Micromobility", "Drive", "Transit", "Walk")))
 
```

```{r trip_purpose_condensed_variable}

variable_list <- add_variable(variable_list, 
                              variable_name = "purp_condensed",
                              table_name = "trip", 
                              data_type = "character")  

trip <- trip %>%   
   mutate(purp_condensed = case_when(dest_purpose_cat %in% c("School", "School-related") ~ "School", dest_purpose_cat %in% c("Work", "Work-related") ~ "Work", dest_purpose_cat %in% c("Shop", "Errand/Other") ~ "Shopping/Errands", dest_purpose_cat %in% c("Meal", "Social/Recreation") ~ "Social/Rec/Meal")) %>% mutate(purp_condensed = factor(purp_condensed, levels = c("School", "Work", "Social/Rec/Meal", "Shop/Errands")))
```

```{r income_variable}

variable_list <- add_variable(variable_list,
                              variable_name = "hhincome_broad_combined", 
                              table_name = "hh", 
                              data_type = "character")  

hh <- hh %>%    
      mutate(hhincome_broad_combined = ifelse(hhincome_broad %in% c("$100,000-$199,000", "$200,000 or more"), "$100,000 or more", hhincome_broad)) %>% mutate(hhincome_broad_combined = factor(hhincome_broad_combined, levels = c("Under $25,000", "$25,000-$49,999", "$50,000-$74,999", "$75,000-$99,999", "$100,000 or more", "Prefer not to answer")))

variable_list <- add_variable(variable_list,
                              variable_name = "hhincome_broad_combined2", 
                              table_name = "hh", 
                              data_type = "character")  

```

```{r vehicle_owner_variable}

variable_list <- add_variable(variable_list, 
                              variable_name = "vehicle_ownership", 
                              table_name = "hh", 
                              data_type = "character")  

hh <- hh %>%    
  mutate(vehicle_ownership = ifelse(vehicle_count == "0 (no vehicles)", "Not a vehicle owner", "Vehicle owner"))

## data exploration for zero vehicle households

# hh_test <- hh %>%
#   filter(vehicle_count == "0 (no vehicles)")
# 
# aggregate(hh_test$hh_weight, by=list(survey_year=hh_test$survey_year), FUN=sum)
# 
# # count zero vehicle households
# hh_count <- hh %>%
#    count(vehicle_count == "0 (no vehicles)")

# hh_total <- hh_test %>%
#   mutate(Num = readr::parse_number(hhsize))
# 
# aggregate(hh_total$Num, by=list(survey_year=hh_total$survey_year), FUN=sum)
# 
# person_w <- person %>%
#   select("person_id", "hh_id", "person_weight")
# 
# person_W_noveh <- left_join(hh_total, person_w, by = "hh_id")
# 
# aggregate(person_W_noveh$person_weight, by=list(survey_year=person_W_noveh$survey_year), FUN=sum)

```

```{r summary}

hts_data <- list(hh = hh, 
                 person = person,
                 trip = trip) 
ids <- c("hh_id", "person_id", "trip_id") 
wts <- c("hh_weight", "person_weight", "trip_weight") 

#remove duplicates
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)

#mode_simple
summary_trips_simple <- summarize_weighted(hts_data = hts_data, 
                                    summarize_var = "mode_simple", 
                                    summarize_by = "survey_year",
                                    id_cols = ids, 
                                    wt_cols = wts, 
                                    wtname = "trip_weight")  

mode_summary_simple <- summary_trips_simple$summary$wtd %>% 
  mutate(moe = prop_se * 1.645)

mode_summary_simple <- mode_summary_simple %>%
  filter(mode_simple != "NA")

#mode_characterization
summary_trips <- summarize_weighted(hts_data = hts_data,
                                    summarize_var = "mode_condensed",
                                    summarize_by = "survey_year",
                                    id_cols = ids,
                                    wt_cols = wts,
                                    wtname = "trip_weight"
                                    )

mode_summary <- summary_trips$summary$wtd %>% 
  mutate(moe = prop_se * 1.645) %>%
  filter(mode_condensed != "NA") %>%
  filter(survey_year == "2023")

#transit_frequency
summary_transit <- summarize_weighted(hts_data = hts_data,
                                    summarize_var = "transit_freq",
                                    summarize_by = "survey_year",
                                    id_cols = ids,
                                    wt_cols = wts,
                                    wtname = "trip_weight"
                                    )

transit_summary <- summary_transit$summary$wtd %>% 
  mutate(moe = prop_se * 1.645) 

#transit and walk
tandw <- c("Transit", "Walk")

transitandwalk <- mode_summary %>% filter(mode_characterization %in% tandw)
 
# #rgcs
# summary_trips_centers <- summarize_weighted(hts_data = hts_data,
#                                     summarize_var = "mode_characterization",
#                                     summarize_by=c("survey_year","home_rgcname"),
#                                     id_cols = ids,
#                                     wt_cols = wts,
#                                     wtname = "trip_weight"
#                                     )
# 
# mode_summary_centers <- summary_trips_centers$summary$wtd %>% 
#   mutate(moe = prop_se * 1.645)
# 
# t <- get_table(schema = "HHSurvey", tbl = "households_2021")
# tbltrips <- get_table(schema = "HHSurvey", tbl = "trips_2021")
# 
# teverett <- t %>%
#   filter(final_home_rgcnum == "Kirkland Totem Lake")
# 
# tripseverett <- right_join(tbltrips, teverett, by = "household_id")
# 
# t23 <- get_table(schema = "HHSurvey", tbl = "household_fact_2023")
# 
# transit_centers <- mode_summary_centers %>%
#   filter(mode_characterization == "Transit")
# 
# walk_centers <- mode_summary_centers %>%
#   filter(mode_characterization == "Walk")
```

## Demographics

```{r}

#income with $100K+
mode_by_income <- summarize_weighted(hts_data = hts_data,
                                       summarize_var = "mode_characterization",
                                       summarize_by = c("survey_year", "hhincome_broad_combined"),
                                       id_cols = ids,
                                       wt_cols = wts,
                                       wtname = "trip_weight"
                                       )

transit_by_income <- mode_by_income$summary$wtd %>%
  filter(mode_characterization == "Transit") %>%
  filter(survey_year == "2023") %>%
  mutate(moe = prop_se * 1.645)

#income with $100K-199K and $200K+ 
mode_by_income2 <- summarize_weighted(hts_data = hts_data,
                                       summarize_var = "mode_characterization",
                                       summarize_by = c("survey_year", "hhincome_broad"),
                                       id_cols = ids,
                                       wt_cols = wts,
                                       wtname = "trip_weight"
                                       )

transit_by_income2 <- mode_by_income2$summary$wtd %>%
  filter(mode_characterization == "Transit") %>%
  filter(survey_year == "2023") %>%
  mutate(moe = prop_se * 1.645)

transit_by_income2 <- transit_by_income2 %>%    
      mutate(hhincome_broad = factor(hhincome_broad, levels = c("Under $25,000", "$25,000-$49,999", "$50,000-$74,999", "$75,000-$99,999", "$100,000-$199,000", "$200,000 or more", "Prefer not to answer")))

# aggregate to fewer age groups
# mode_by_age_cond <- summarize_weighted(hts_data = hts_data,
#                                        summarize_var = "mode_characterization",
#                                        summarize_by = c("survey_year", "age_condensed"),
#                                        id_cols = ids,
#                                        wt_cols = wts,
#                                        wtname = "trip_weight"
#                                        )
# 
# transit_by_age_cond <- mode_by_age_cond$summary$wtd %>% 
#   filter(mode_characterization == "Transit", age_condensed != "NA") %>% 
#   mutate(moe = prop_se * 1.645)

#race
mode_by_race <- summarize_weighted(hts_data = hts_data,
                                   summarize_var = "mode_characterization",
                                   summarize_by = c("survey_year", "race_category"),
                                   id_cols = ids,
                                   wt_cols = wts,
                                   wtname = "trip_weight"
                                   )

transit_by_race <- mode_by_race$summary$wtd %>% 
  filter(mode_characterization == "Transit",  race_category != "Child", race_category != "Missing") %>% 
  mutate(race_category = recode(race_category, 'White Only' = 'White')) %>%
  mutate(moe = prop_se * 1.645)

transit_by_race_23 <- transit_by_race %>%
  filter(survey_year == "2023") %>%
  filter(race_category != "Missing/No response")

share_by_race <- transit_by_race %>%
  group_by(survey_year) %>%
  mutate(total = sum(est)) %>%
  mutate(share = (est/total))

#disability
mode_by_dis <- summarize_weighted(hts_data = hts_data,
                                   summarize_var = "mode_characterization",
                                   summarize_by = c("survey_year", "disability_person"),
                                   id_cols = ids,
                                   wt_cols = wts,
                                   wtname = "trip_weight"
                                   )

transit_by_dis <- mode_by_dis$summary$wtd %>% 
  filter(mode_characterization == "Transit", survey_year == "2023") %>% 
  mutate(moe = prop_se * 1.645)

#destination purpose
mode_by_purp <- summarize_weighted(hts_data = hts_data,
                                   summarize_var = "mode_characterization",
                                   summarize_by = c("survey_year", "dest_purpose_cat"),
                                   id_cols = ids,
                                   wt_cols = wts,
                                   wtname = "trip_weight"
                                   )

transit_by_purp <- mode_by_purp$summary$wtd %>% 
  filter(mode_characterization == "Transit", survey_year == "2023") %>%
  mutate(moe = prop_se * 1.645)

#job industry
summary_trips_ind <- summarize_weighted(hts_data = hts_data,
                                    summarize_var = "mode_characterization",
                                    summarize_by = c("survey_year", "industry_cond"),
                                    id_cols = ids,
                                    wt_cols = wts,
                                    wtname = "trip_weight"
                                    )

mode_summary_ind <- summary_trips_ind$summary$wtd %>% 
  mutate(moe = prop_se * 1.645)

transit_by_ind <- mode_summary_ind %>% 
    filter(mode_characterization == "Transit", industry_cond != "NA", industry_cond != "Missing: Skip Logic")

#vehicle ownership in centers
summary_trips_veh <- summarize_weighted(hts_data = hts_data,
                                    summarize_var = "mode_condensed",
                                    summarize_by = c("survey_year", "vehicle_ownership"),
                                    id_cols = ids,
                                    wt_cols = wts,
                                    wtname = "trip_weight"
                                    )

mode_summary_veh <- summary_trips_veh$summary$wtd %>% 
  mutate(moe = prop_se * 1.645)

transit_by_veh <- mode_summary_veh %>% 
    filter(mode_condensed == "Transit")

No_veh <- mode_summary_veh %>% 
    filter(vehicle_ownership == "Not a vehicle owner", mode_condensed != "NA")

#vehicle ownership in centers
summary_trips_veh_cen <- summarize_weighted(hts_data = hts_data,
                                    summarize_var = "mode_characterization",
                                    summarize_by = c("survey_year", "vehicle_ownership", "home_rgcname"),
                                    id_cols = ids,
                                    wt_cols = wts,
                                    wtname = "trip_weight"
                                    )

mode_summary_veh_cen <- summary_trips_veh_cen$summary$wtd %>% 
  mutate(moe = prop_se * 1.645)

#write.csv(mode_summary_veh, 'noveh2.csv')

transit_by_veh_Cen <- mode_summary_veh_cen %>% 
    filter(mode_characterization == "Transit")

No_veh_cen <- mode_summary_veh %>% 
    filter(vehicle_ownership == "Not a vehicle owner", mode_characterization != "NA")

#, industry_cond != "NA", industry_cond != "Missing: Skip Logic")

```

## Initial Mode Charts

```{r urban_rail_chart_est, echo=FALSE}

#chart for simple mode
# mode_chart <- static_column_chart(mode_summary_simple,
#                       x = "survey_year", y = "prop", fill = "mode_simple",
#                     ylabel = "% of Trips", xlabel = "Survey Year", title = "Trips - Prop",
#                     moe = "prop_se") + theme(
#                       axis.text.x = element_text(size = 14),
#                       axis.text.y = element_text(size = 14),
#                       axis.title.y = element_text(size = 20),
#                       axis.title.x = element_text(size = 20),
#                       plot.title = element_text(size = 24)
#                       )

#chart for mode characterization
mode_char_chart <- static_column_chart(mode_summary,
                    x = "mode_condensed", y = "prop", fill = "survey_year",
                    ylabel = "Share of Trips", xlabel = "Mode", 
                    #title = "Trips - Prop",  
                    moe = "moe") + theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )
mode_char_chart



#chart for transit frequency
transit_chart <- static_bar_chart(transit_summary,
                    x = "prop", y = "transit_freq", fill = "survey_year",
                    ylabel = "% of Trips", xlabel = "Survey Year", title = "Trips - Prop",  moe = "moe") + theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )
transit_chart
```

## Transit Charts

```{r}

#transit by race
transitbyrace_chart <- static_bar_chart(transit_by_race_23,
                    x = "prop", y = "race_category", fill = "survey_year",
                    ylabel = "Race/Ethnicity", xlabel = "Share of Trips", #title = "Transit Mode Share by Race for 2023", 
                    moe = "moe"
                    ) + theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )


transitsharebyrace_chart <- static_line_chart(share_by_race,
                    x = "survey_year", y = "share", fill = "race_category",
                    ylabel = "Share", xlabel = "Survey Year", title = "Transit Trips by Race - Share"
                    #moe = "moe"
                    ) + theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )

#transit by disability
transitbydis_chart <- static_column_chart(transit_by_dis,
                    x = "disability_person", y = "prop", fill = "disability_person",
                    ylabel = "Share of Trips", 
                    #xlabel = "Survey Year", #title = "Transit Trips by Disability - Estimate",
                    moe = "moe") + theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )

#transit by destination purpose
transitbypurp_chart <- static_column_chart(transit_by_purp,
                    x = "dest_purpose_cat", y = "prop", fill = "survey_year",
                    ylabel = "Share of Trips", 
                    #xlabel = "Survey Year", #title = "Transit Trips by Disability - Estimate",
                    moe = "moe") + theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )
#transit by industry
transitbyind_chart <- static_line_chart(transit_by_ind,
                    x = "survey_year", y = "prop", fill = "industry_cond", 
                    color = "gnbopgy_10",
                    ylabel = "# of Trips", xlabel = "Survey Year", title = "Transit Trips by Job Industry - Prop",
                    #moe = "prop_se"
                    ) + 
                      theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )

transitbyincome_chart <- static_bar_chart(transit_by_income2,
                    x = "prop", y = "hhincome_broad", fill = "survey_year",
                    ylabel = "Household Income", xlabel = "Transit Mode Share", moe = "moe") + theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )

transitbyage_chart <- static_column_chart(transit_by_age_cond,
                    x = "survey_year", y = "est", fill = "age_condensed",
                    ylabel = "# of Trips", xlabel = "Survey Year", title = "Transit Trips - Estimate",
                    moe = "est_se") + theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )

transitbyveh_chart <- static_column_chart(transit_by_veh,
                    x = "vehicle_ownership", y = "prop", fill = "survey_year",
                    ylabel = "Share of Trips", xlabel = "Vehicle Ownership", 
                    , moe = "moe"
                    ) + theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )

noveh_chart_mode <- static_bar_chart(No_veh,
                    x = "prop", y = "mode_condensed", fill = "survey_year",
                    ylabel = "Mode", xlabel = "Share of Trips", 
                    #title = "Trips by Mode for Zero Vehicle HH - Prop, 
                    moe = "moe"
                    ) + theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )

centers_chart_mode <- static_column_chart(transit_centers,
                    x = "home_rgcname", y = "prop", fill = "survey_year",
                    ylabel = "% of Trips", xlabel = "Survey Year", title = "Mode by Center - Prop"
                    , moe = "moe"
                    ) + theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )

#chart for transit and walking
transit_walk_chart <- static_column_chart(transitandwalk,
                    x = "mode_characterization", y = "prop", fill = "survey_year",
                    ylabel = "% of Trips", xlabel = "Survey Year", title = "Transit and Walking - Prop"
                    , moe = "moe"
                    ) + theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )

centers_chart_walk <- static_column_chart(walk_centers,
                    x = "home_rgcname", y = "prop", fill = "survey_year",
                    ylabel = "% of Trips", xlabel = "Survey Year", title = "Walking by Center - Prop"
                    , moe = "moe"
                    ) + theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )

```

## Summary for Simple Trips

```{r summary}

hts_data <- list(hh = hh, 
                 person = person, 
                 trip = trip) 
ids <- c("hh_id", "person_id", "trip_id") 
wts <- c("hh_weight", "person_weight", "trip_weight")

summary_trips_simp <- summarize_weighted(hts_data = hts_data, 
                                    summarize_var = "mode_simple", 
                                    summarize_by = "survey_year", 
                                    id_cols = ids, wt_cols = wts, 
                                    wtname = "trip_weight" )

mode_summary_simp <- summary_trips_simp$summary$wtd %>% mutate(moe = prop_se * 1.645)

#filter for Urban Rail

#mode_summary_rail <- mode_summary |> filter(grepl('Urban Rail', mode_1))
```

```{r chart_share, echo=FALSE}

static_column_chart(na.omit(mode_summary_simp),
                    x = "mode_simple", y = "prop", fill = "survey_year",
                    ylabel = "% of Trips", xlabel = "Survey Year", title = "Trips - Share",
                    moe = "moe") + theme(
                      axis.text.x = element_text(size = 14),
                      axis.text.y = element_text(size = 14),
                      axis.title.y = element_text(size = 20),
                      axis.title.x = element_text(size = 20),
                      plot.title = element_text(size = 24)
                      )
```
