---
title: "VMT Data Description Notes"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    df_print: paged
    fig_width: 7
    fig_height: 3
    fig_caption: true
date: "2023-06-22"
---
<!-- find source code: data_description_script.R -->
<!-- # checklist -->
<!-- - [ ] person/household-level vmt -->
<!-- - [ ] correlation plot -->
<!-- - [ ] lm with one variable: find significant variables -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
               warning=FALSE, 
               message=FALSE,
               fig.align = "center")
```


```{r, include=FALSE}
library(echarts4r)
library(psrcplot)
library(psrc.travelsurvey)
library(psrccensus)
library(psrcelmer)
library(tidyverse)
library(wdm)
library(survey)
library(htmlwidgets)
options(scipen=999)

install_psrc_fonts()

Sys.setenv(CENSUS_API_KEY = '3fc20d0d6664692c0becc323b82c752408d843d9')
Sys.getenv("CENSUS_API_KEY")

# Common options
e_common(
  font_family = "Poppins",
  theme = "infographic"
)

source("hhts_var_function.R")
```

```{r find hhts variables}
# hhts_codebook <- get_table(schema = 'HHSurvey', tbl_name = 'variables_codebook')
# trip_path_distance: Bing-estimated trip distance (miles)

hh_vars=c("sample_county", "final_home_rgcnum", "final_home_is_rgc",
          "hhsize", "hhincome_broad", #"hhincome_detailed", 
          "numadults", "numchildren", "numworkers", "lifecycle",
          # "res_dur",  "res_months",
          "res_type",
          #"broadband", 
          "vehicle_count", "offpark", "offpark_cost", "streetpark")

trip_vars = c("trip_id","mode_simple",
              'dest_purpose_cat', 'origin_purpose_cat',
              'd_tract10',"d_rgcname",
              "reported_duration", 
              'trip_path_distance', # Bing-estimated trip distance (miles)
              "depart_time_hhmm", "arrival_time_hhmm", 
              "dayofweek", "travelers_total",
              "park_pay" # if mode = auto (non-taxi)
              )

person_vars=c("gender", 
              "age", "age_category", "race_eth_broad",
              "education", "workplace", "industry", "employment",
              "worker", # for calculating number of workers in the household
              
              "license",
              "commute_freq", # How often commuted to workplace last week
              "commute_mode", # Method of commuting to work location/office last week
              "telecommute_freq",
              
              "mode_freq_1", # Times ridden transit in past 30 days
              "mode_freq_2", # Times ridden a bike in past 30 days
              "mode_freq_3", # Times gone for a walk in past 30 days
              "mode_freq_4", # Times used carshare in past 30 days
              "mode_freq_5", # Times used rideshare in past 30 days
              "benefits_3", # Employer commuter benefits: Free/partially subsidized passes/fares
              "work_park_type","workpass","workpass_cost" # if parks at work
              )
              

# var_set <- hhts_codebook %>% filter(grepl("trip ID", description, ignore.case = TRUE))

hh_data_17_19<- get_hhts("2017_2019", "h", vars=hh_vars) %>% hh_group_data() 
per_data_17_19<- get_hhts("2017_2019", "p", vars=person_vars) %>% per_group_data(hh_data_17_19)
trip_data_17_19<- get_hhts("2017_2019", "t", vars=trip_vars) %>% trip_group_data()
day_data_17_19<- get_hhts("2017_2019", "d", vars=c("telework_time")) %>%day_group_data()
```


```{r}
# vmt data for analysis and modeling
working_vars <- c("person_id","household_id","sample_segment","survey",
                  "sample_county","final_home_rgcnum","urban_metro",
                  "gender","age_category","race_eth_broad","have_child",
                  "employment","license","hhincome_broad","vehicle_count",
                  "transit_pass","work_parking","commute_mode_simple",
                  "freq_transit","telecommute_freq",
                  "hh_weight_2017_2019","hh_weight_2017_2019_adult")


# new calculation with weighted trips
vmt_trip <- trip_data_17_19 %>% 
  filter(mode_simple=='Drive') %>%
  mutate(trip_path_distance=replace_na(trip_path_distance,0),
         travelers_total=case_when(travelers_total>10~1, 
                                   travelers_total<1~1,
                                   is.na(travelers_total)~1,
                                   TRUE~travelers_total),
         trip_adult_weight_2017_2019=replace_na(trip_adult_weight_2017_2019,0),
         vmt= trip_path_distance/travelers_total, 
         vmt_trip_weighted=((trip_path_distance*trip_adult_weight_2017_2019)/travelers_total))




# final person data for data description 
person_final <- per_data_17_19 %>% 
  select(all_of(working_vars)) %>%
  mutate(hh_weight_2017_2019_adult=replace_na(hh_weight_2017_2019_adult,0),
         vehicle_count_num = as.numeric(substr(vehicle_count,1,1)),
         vehicle_bi = ifelse(vehicle_count_num>0, "1+ vehicle(s)", "0 vehicle"),
         freq_transit_simple = factor(recode(freq_transit,
                                             "1 day/week" = "1-4 days/week",
                                             "2-4 days/week" = "1-4 days/week",
                                             "5 days/week" = "5+ days/week",
                                             "6-7 days/week" = "5+ days/week"),
                                      c("5+ days/week",
                                        "1-4 days/week",
                                        "1-3 times in the past 30 days",
                                        "I do this, but not in the past 30 days",
                                        "I never do this")),
         freq_telecommute_simple = recode(telecommute_freq,
                                      "Never" = "Never",
                                      "1 day a week" = "1-2 days",
                                      "2 days a week" = "1-2 days",
                                      "3 days a week" = "3-4 days",
                                      "4 days a week" = "3-4 days",
                                      "5 days a week" = "5+ days",
                                      "6-7 days a week" = "5+ days",
                                      "Not applicable" = as.character(NA))) %>% 
  group_by(urban_metro,final_home_rgcnum) %>%
  mutate(big_rgc_grp = as.character(case_when(n()<30 & urban_metro =="Metro"~"Other Metro centers",
                                              n()<30 & urban_metro =="Urban"~"Other Urban centers",
                                              TRUE~final_home_rgcnum))) %>%
  ungroup()


vmt_final <- vmt_trip %>%
  left_join(person_final, by = c("survey","sample_segment","household_id","person_id")) %>%
  left_join(day_data_17_19, by = c("survey","sample_segment","household_id","person_id","daynum"))

day_final <- day_data_17_19 %>%
  left_join(person_final, by = c("survey","sample_segment","household_id","person_id"))
  
```


## Regional VMT stat

- total VMT in the region: `r sum(vmt_final$vmt_trip_weighted)`
- total adults in the region: `r sum(vmt_final$hh_weight_2017_2019_adult)`
- total employed adults in the region: `r sum(vmt_final[vmt_final$employment %in% c("Employed full time (35+ hours/week, paid)","Employed part time (fewer than 35 hours/week, paid)"),"hh_weight_2017_2019_adult"])`

- average VMT per person in the region: `r sum(vmt_final$vmt_trip_weighted)/sum(vmt_final$hh_weight_2017_2019_adult)`



## vmt by geographics {.tabset}


```{r}
vmt_stat_function <- function(df_vmt, df_person, group_var, group_var_str){
  return(
    df_vmt %>%
      group_by({{group_var}}) %>%
      summarise(sample_size = n(),
                vmt_sum = sum(vmt_trip_weighted)) %>%
      ungroup() %>%
      left_join(df_person %>%
                  group_by({{group_var}}) %>%
                  summarise(pop_sum = sum(hh_weight_2017_2019_adult)) %>%
                  ungroup(), by = group_var_str) %>%
      mutate(vmt_mean = vmt_sum/pop_sum)
  )
}

```

### vmt by counties

```{r}
df_stat <- vmt_stat_function(vmt_final,person_final, sample_county,"sample_county") 
df_stat

df_stat %>% 
  mutate(vmt_mean = round(vmt_mean,2)) %>%
  e_charts(sample_county) |> 
  e_bar(vmt_mean, name = "mean of vmt in county",
        label=list(show = TRUE)) |> 
  e_color(psrc_colors$pognbgy_5) |>
  e_axis_labels(x = "county", y = "vmt") |>
  e_grid(height = "55%", top = "30%", containLabel= TRUE) |>
  e_legend(top='10%') |>
  e_title("vmt distribution by counties")
```

### vmt by centers

```{r}
df_stat <- vmt_stat_function(vmt_final,person_final, big_rgc_grp,"big_rgc_grp") 
df_stat

df_stat%>%
  left_join(urban_metro[,c("name", "category")], by=c("big_rgc_grp"="name")) %>% 
  mutate(category = case_when(big_rgc_grp == "Other Metro centers"~"Metro",
                              big_rgc_grp == "Other Urban centers"~"Urban",
                              big_rgc_grp == "Not RGC"~"Not RGC",
                              TRUE~category
                              ),
         vmt_mean = round(vmt_mean,2)) %>%
  group_by(category) |> 
  e_charts(big_rgc_grp) |>  
  e_bar(vmt_mean, 
        bind = category,
        label=list(show = TRUE),
        barGap='-100%',
        barMaxWidth='60%') |>
  e_color(psrc_colors$pognbgy_5) |>
  e_axis_labels(x = "center", y = "mean of vmt") |>
  e_x_axis(axisLabel=list(rotate= 45),
           axisTick=list(alignWithLabel=TRUE)) |>
  e_grid(height = "55%", top = "30%", containLabel= TRUE)  |> 
  e_legend(top='12%') |>
  e_title("vmt distribution by centers",
          "centers with sample size smaller than 30 are grouped in other metro and other urban centers") |> 
  e_tooltip(formatter = htmlwidgets::JS("
      function(params){
        return('<strong>' + params.name + 
                '</strong><br />RGC name: ' + params.value[0] + '<br />vmt mean: ' + params.value[1])
      }
    "))

```

## vmt by demographic characteristics {.tabset}

### vmt by income group

```{r}
df_stat <- vmt_stat_function(vmt_final,person_final, hhincome_broad,"hhincome_broad") 
df_stat

df_stat%>%
  mutate(vmt_mean = round(vmt_mean,2)) %>%
  e_charts(hhincome_broad) |>
  e_bar(vmt_mean,
        label=list(show = TRUE),
        name = "mean of vmt") |>
  e_color(psrc_colors$pognbgy_5) |>
  e_x_axis(axisLabel = list(interval = 0L)) |>
  e_flip_coords() |>
  e_grid(height = "75%", top = "20%", containLabel= TRUE)  |>
  e_legend(top='10%') |>
  e_axis_labels(x = "income level") |>
  e_title("vmt distribution by income level")
```

## vmt by travel behavior {.tabset}

### vmt by transit frequency
- frequent transit users have significantly lower vmt
```{r vmt transit frequency}
df_stat <- vmt_stat_function(vmt_final,person_final, freq_transit_simple,"freq_transit_simple")
df_stat

df_stat %>%
  filter(!is.na(freq_transit_simple)) %>%
  e_charts(freq_transit_simple) |>
  e_bar(vmt_mean,
        label=list(show = TRUE),
        name = "mean of vmt") |>
  e_color(psrc_colors$pognbgy_5) |>
  e_x_axis(axisLabel = list(interval = 0L)) |>
  e_flip_coords() |>
  e_grid(height = "75%", top = "20%", containLabel= TRUE)  |>
  e_legend(top='10%') |>
  e_axis_labels(x = "transit frequency", y = "mean of vmt") |>
  e_title("vmt distribution by transit frequency")
```


### vmt by telecommute frequency
- people who telework 3-4 days/week have the highest vmt, while people who telework 5 days/week have the lowest
```{r}
# only include full/part-time employment
df_stat <- vmt_stat_function(vmt_final%>% 
  filter(employment %in% c("Employed full time (35+ hours/week, paid)",
                           "Employed part time (fewer than 35 hours/week, paid)")),
  person_final%>% 
  filter(employment %in% c("Employed full time (35+ hours/week, paid)",
                           "Employed part time (fewer than 35 hours/week, paid)")), 
  telecommute_freq,"telecommute_freq")
df_stat

df_stat %>%
  filter(!is.na(telecommute_freq)) %>%
  mutate(vmt_mean = round(vmt_mean,2)) %>%
  e_charts(telecommute_freq) |>
  e_bar(vmt_mean,
        label=list(show = TRUE),
        name = "mean of vmt") |>
  e_color(psrc_colors$pognbgy_5) |>
  e_x_axis(axisLabel = list(interval = 0L)) |>
  e_flip_coords() |>
  e_grid(height = "75%", top = "20%", containLabel= TRUE)  |>
  e_legend(top='10%') |>
  e_axis_labels(x = "telecommute \nfrequency") |>
  e_title("vmt distribution by telecommute frequency",
          "only include travelers with full/part-time employment")
```

- **only including high income group** (household income higher than $75,000)
```{r}
# only include full/part-time employment and high income households
df_stat <- vmt_stat_function(vmt_final%>% 
  filter(hhincome_broad %in% c("$75,000-$99,999","$100,000 or more"),
         employment %in% c("Employed full time (35+ hours/week, paid)",
                           "Employed part time (fewer than 35 hours/week, paid)")),
  person_final%>% 
  filter(hhincome_broad %in% c("$75,000-$99,999","$100,000 or more"),
         employment %in% c("Employed full time (35+ hours/week, paid)",
                           "Employed part time (fewer than 35 hours/week, paid)")), 
  telecommute_freq,"telecommute_freq")
df_stat

df_stat %>%
  filter(!is.na(telecommute_freq)) %>%
  mutate(vmt_mean = round(vmt_mean,2)) %>%
  e_charts(telecommute_freq) |>
  e_bar(vmt_mean,
        label=list(show = TRUE),
        name = "mean of vmt") |>
  e_color(psrc_colors$pognbgy_5) |>
  e_x_axis(axisLabel = list(interval = 0L)) |>
  e_flip_coords() |>
  e_grid(height = "75%", top = "20%", containLabel= TRUE)  |>
  e_legend(top='10%') |>
  e_axis_labels(x = "telecommute \nfrequency") |>
  e_title("vmt distribution by telecommute frequency",
          "only include travelers with full/part-time employment household income > $75,000")

```

### vmt by teleworking hours
```{r}
# only include full/part-time employment
df_stat <- vmt_final%>% 
  filter(employment %in% c("Employed full time (35+ hours/week, paid)",
                           "Employed part time (fewer than 35 hours/week, paid)")) %>%
      group_by(telework_2hour) %>%
      summarise(sample_size = n(),
                vmt_sum = sum(vmt_trip_weighted)) %>%
      ungroup() %>%
  left_join(day_final%>% 
              filter(employment %in% c("Employed full time (35+ hours/week, paid)",
                                       "Employed part time (fewer than 35 hours/week, paid)")) %>%
              group_by(telework_2hour) %>%
              summarise(pop_sum = sum(day_weight_2017_2019)) %>%
              ungroup(), by = "telework_2hour") %>%
  mutate(vmt_mean = vmt_sum/pop_sum)

df_stat

df_stat %>%
  filter(!is.na(telework_2hour)) %>%
  mutate(vmt_mean = round(vmt_mean,2)) %>%
  e_charts(telework_2hour) |>
  e_bar(vmt_mean,
        label=list(show = TRUE),
        name = "mean of vmt") |>
  e_color(psrc_colors$pognbgy_5) |>
  e_x_axis(axisLabel = list(interval = 0L)) |>
  e_flip_coords() |>
  e_grid(height = "75%", top = "20%", containLabel= TRUE)  |>
  e_legend(top='10%') |>
  e_axis_labels(y = "teleworking hours") |>
  e_title("vmt distribution by teleworking hours",
          "only include travelers with full/part-time employment")
```



<!-- # Basic stat for household-level vmt {.tabset} -->

<!-- ## vmt in counties and centers -->
<!-- ```{r hh vmt geography [county and centers]} -->
<!-- # vmt in county -->
<!-- vmt_hh %>%  -->
<!--   hhts_mean(stat_var="vmt_day", group_vars="sample_county") %>% -->
<!--   mutate(lower = vmt_day_mean-vmt_day_mean_moe, -->
<!--          upper = vmt_day_mean+vmt_day_mean_moe) %>%  -->
<!--   e_charts(sample_county) |>  -->
<!--   e_bar(vmt_day_mean, name = "mean of vmt in county") |>  -->
<!--   e_error_bar(lower, upper) |>  -->
<!--   e_color(psrc_colors$pognbgy_5) |> -->
<!--   e_axis_labels(x = "county", y = "vmt") |> -->
<!--   e_grid(height = "55%", top = "30%", containLabel= TRUE) |> -->
<!--   e_legend(top='10%') |> -->
<!--   e_title("vmt distribution by counties") -->

<!-- # group centers with small sample size -->
<!-- rgc_sample_size <- vmt_hh %>%  -->
<!--   hhts_count(group_vars=c("urban_metro","final_home_rgcnum")) %>% -->
<!--   mutate(big_rgc_grp = as.character(case_when(sample_size<30 & urban_metro =="Metro"~"Other Metro centers", -->
<!--                                               sample_size<30 & urban_metro =="Urban"~"Other Urban centers", -->
<!--                                               TRUE~final_home_rgcnum))) %>% -->
<!--   filter(final_home_rgcnum!="Total") -->

<!-- # vmt in RGCs -->
<!-- vmt_hh %>% -->
<!--   left_join(rgc_sample_size %>% select(final_home_rgcnum,big_rgc_grp), by = "final_home_rgcnum") %>%  -->
<!--   hhts_mean(stat_var="vmt_day", group_vars=c("urban_metro","big_rgc_grp")) %>% -->
<!--   filter(big_rgc_grp!="Total")  %>%  -->
<!--   arrange(vmt_day_mean) %>% -->
<!--   select(urban_metro,big_rgc_grp,vmt_day_mean,sample_size#,lower,upper -->
<!--          ) %>% -->
<!--   mutate(vmt_day_mean = round(vmt_day_mean,1)) |>  -->
<!--   group_by(urban_metro) |>  -->
<!--   e_charts(big_rgc_grp) |>   -->
<!--   e_bar(vmt_day_mean,  -->
<!--         bind = urban_metro, -->
<!--         label=list(show = TRUE), -->
<!--         barGap='-100%', -->
<!--         barMaxWidth='60%') |> -->
<!--   e_color(psrc_colors$pognbgy_5) |> -->
<!--   e_axis_labels(x = "center", y = "mean of vmt") |> -->
<!--   e_x_axis(axisLabel=list(rotate= 45), -->
<!--            axisTick=list(alignWithLabel=TRUE)) |> -->
<!--   e_grid(height = "55%", top = "30%", containLabel= TRUE)  |>  -->
<!--   e_legend(top='12%') |> -->
<!--   e_title("vmt distribution by centers", -->
<!--           "centers with sample size smaller than 30 are grouped in other metro and other urban centers") |>  -->
<!--   e_tooltip(formatter = htmlwidgets::JS(" -->
<!--       function(params){ -->
<!--         return('<strong>' + params.name +  -->
<!--                 '</strong><br />RGC name: ' + params.value[0] + '<br />vmt mean: ' + params.value[1]) -->
<!--       } -->
<!--     ")) -->
<!-- ``` -->

<!-- ## income group -->
<!-- ```{r} -->
<!-- # only high income  -->
<!-- vmt_hh %>%  -->
<!--   hhts_count(group_vars=c("hhincome_broad"), incl_na = FALSE) %>% -->
<!--   filter(hhincome_broad!="Total") %>% -->
<!--   select(2,5,7) -->


<!-- vmt_hh %>%  -->
<!--   hhts_mean(stat_var="vmt_day", group_vars="hhincome_broad", incl_na = FALSE) %>% -->
<!--   mutate(vmt_day_mean = round(vmt_day_mean,2)) %>% -->
<!--   filter(hhincome_broad!="Total") %>% -->
<!--   e_charts(hhincome_broad) |> -->
<!--   e_bar(vmt_day_mean, -->
<!--         label=list(show = TRUE), -->
<!--         name = "mean of vmt") |> -->
<!--   e_color(psrc_colors$pognbgy_5) |> -->
<!--   e_x_axis(axisLabel = list(interval = 0L)) |> -->
<!--   e_flip_coords() |> -->
<!--   e_grid(height = "75%", top = "20%", containLabel= TRUE)  |> -->
<!--   e_legend(top='10%') |> -->
<!--   e_axis_labels(x = "income level") |> -->
<!--   e_title("vmt distribution by income level") -->
<!-- ``` -->

<!-- ## vehicle count -->
<!-- ```{r} -->
<!-- # only high income  -->
<!-- vmt_hh %>%  -->
<!--   hhts_count(group_vars=c("vehicle_count_simple4"), incl_na = FALSE) %>% -->
<!--   filter(vehicle_count_simple4!="Total") %>% -->
<!--   select(2,5,7) -->


<!-- vmt_hh %>%  -->
<!--   hhts_mean(stat_var="vmt_day", group_vars="vehicle_count_simple4", incl_na = FALSE) %>% -->
<!--   mutate(vmt_day_mean = round(vmt_day_mean,2)) %>% -->
<!--   filter(vehicle_count_simple4!="Total") %>% -->
<!--   e_charts(vehicle_count_simple4) |> -->
<!--   e_bar(vmt_day_mean, -->
<!--         label=list(show = TRUE), -->
<!--         name = "mean of vmt") |> -->
<!--   e_color(psrc_colors$pognbgy_5) |> -->
<!--   e_x_axis(axisLabel = list(interval = 0L)) |> -->
<!--   # e_flip_coords() |> -->
<!--   e_grid(height = "75%", top = "20%", containLabel= TRUE)  |> -->
<!--   e_legend(top='10%') |> -->
<!--   e_axis_labels(x = "vehicle count") |> -->
<!--   e_title("vmt distribution by vehicle count") -->
<!-- ``` -->

## vmt by transit & parking subsidies {.tabset}
  - how does vmt correlate with transit pass and free/subsidized parking at work?
      * people that are offered a transit pass have lower vmt than those not offered one
      * people that are offered free/subsidized parking at work have higher vmt than those not offered
  - transit pass vs transit frequency: people that are offered a free transit pass tend to use transit more frequently
  - transit frequency vs drive trips: 50% of people who takes transit 5+ days/week also drive 

### vmt by commute benefits
```{r vmt benefits [transit and parking subsidy]}
# transit pass
df_stat <- vmt_stat_function(vmt_final%>% 
  filter(employment %in% c("Employed full time (35+ hours/week, paid)",
                           "Employed part time (fewer than 35 hours/week, paid)")),
  person_final%>% 
  filter(employment %in% c("Employed full time (35+ hours/week, paid)",
                           "Employed part time (fewer than 35 hours/week, paid)")), 
  transit_pass,"transit_pass")%>%
  rename(benefits = transit_pass) %>%
  filter(!is.na(benefits), benefits!="I don't know") %>%
  mutate(type = "transit pass")

df_stat2 <- vmt_stat_function(vmt_final%>% 
  filter(employment %in% c("Employed full time (35+ hours/week, paid)",
                           "Employed part time (fewer than 35 hours/week, paid)")),
  person_final%>% 
  filter(employment %in% c("Employed full time (35+ hours/week, paid)",
                           "Employed part time (fewer than 35 hours/week, paid)")), 
  work_parking,"work_parking") %>%
  rename(benefits = work_parking) %>%
  filter(!is.na(benefits)) %>%
  mutate(type = "free/reimbursed parking at work",
         benefits = case_when(benefits=="free/reimbursed parking"~"Offered",
                              benefits=="paid parking"~"Not offered",
                              TRUE~benefits)) %>%
  add_row(df_stat) %>%
  mutate(benefits = factor(benefits, levels=c("Offered","Not offered")))
df_stat2

df_stat2 %>%
  mutate(vmt_mean = round(vmt_mean,2)) %>%
  group_by(benefits) %>%
  e_charts(type) |>  
  e_bar(vmt_mean, 
        label=list(show = TRUE)) |>
  e_color(psrc_colors$pognbgy_5) |>
  e_x_axis(axisLabel = list(interval = 0L)) |>
  e_grid(height = "55%", top = "30%", containLabel= TRUE)  |> 
  e_legend(top='12%') |>
  e_axis_labels(x = "commute \nbenefits", y = "mean of vmt") |>
  e_title("vmt effects",
          "only include travelers with full/part-time employment")
```

### transit pass vs transit frequency
```{r transit pass vs transit frequency}
df<- vmt_final %>%
  group_by(transit_pass,freq_transit_simple) %>%
  summarise(sample_size = n(),
            vmt_sum = sum(vmt_trip_weighted)) %>%
  ungroup() %>%
  left_join(person_final %>%
              group_by(transit_pass,freq_transit_simple) %>%
              summarise(pop_sum = sum(hh_weight_2017_2019_adult)) %>%
              ungroup(), by = c("transit_pass","freq_transit_simple")) %>%
  mutate(vmt_mean = vmt_sum/pop_sum) %>%
  group_by(transit_pass) %>%
  mutate(total_pop = sum(pop_sum)) %>%
  ungroup() %>%
  mutate(share = round(pop_sum/total_pop,3)*100) %>%
  filter(!is.na(transit_pass), !is.na(freq_transit_simple), transit_pass!="I don't know")


df %>%
  group_by(freq_transit_simple) %>%
  e_charts(transit_pass) |>  
  e_bar(share, 
        label=list(show = TRUE)) |>
  e_color(c(psrc_colors$blues_dec[1:4],psrc_colors$purples_inc[3])) |>
  e_grid(height = "55%", top = "30%", containLabel= TRUE)  |> 
  e_legend(top='10%') |>
  e_axis_labels(x = "transit pass", y = "share (%)") |>
  e_title("transit pass vs transit frequency")
```

<!-- ## transit frequency vs drive trips -->
<!-- - 50% of transit users also drive -->
<!-- ```{r transit frequency vs drive trips} -->
<!-- df <- vmt_per %>% -->
<!--   mutate(n_trip_drive = n_trip_drive/num_day, -->
<!--          dum_drive = case_when(n_trip_drive>0~"drive", -->
<!--                                TRUE~"no drive"), -->
<!--          dum_transit_5plus = case_when(freq_transit %in% c("6-7 days/week","5 days/week")~"5+ days/week", -->
<!--                                TRUE~"other")) %>% -->
<!--   hhts_count(group_vars=c("dum_transit_5plus","dum_drive"), -->
<!--              spec_wgt = "hh_weight_2017_2019_adult", -->
<!--              incl_na = FALSE)  -->
<!-- df %>%  -->
<!--   filter(dum_drive=="Total") %>% -->
<!--   select(1,2,4,5,8) -->

<!-- df %>% -->
<!--   mutate(share = round(share,2)*100) %>% -->
<!--   filter(dum_drive!="Total") %>% -->
<!--   group_by(dum_drive) %>% -->
<!--   e_charts(dum_transit_5plus) |>   -->
<!--   e_bar(share,  -->
<!--         label=list(show = TRUE)) |> -->
<!--   e_color(psrc_colors$pognbgy_5) |> -->
<!--   e_grid(height = "55%", top = "30%", containLabel= TRUE)  |>  -->
<!--   e_legend(top='10%') |> -->
<!--   e_axis_labels(x = "transit \nfrequency", y = "share (%)") |> -->
<!--   e_title("transit frequency vs drive trips") -->
<!-- ``` -->


<!-- # correlation matrix -->

<!-- ## correlation matrix for continuous variables -->
<!-- ```{r} -->
<!-- # cor(cor_test %>% select(vmt_day, vehicle_count_num), method="kendall", use = "pairwise.complete.obs") -->
<!-- ``` -->


<!-- ## correlation matrix for ordinal variables -->
<!-- ```{r weighted correlation matrix} -->
<!-- # cor_test for testing variable correlation -->

<!-- kendall_cor_vars <-  c("vmt_day","hhincome_ord","telecommute_freq_ord","freq_transit_ord",  -->
<!--                        "vehicle_count_num") -->
<!-- point_biserial_cor_vars <- c("vmt_day","transit_pass_binary","work_parking_binary","have_child") -->
<!-- #  -->
<!-- # print("unweighted correlation matrix with ordinal") -->
<!-- # cor(cor_test %>% select(kendall_cor_vars),  -->
<!-- #     method="kendall", use = "pairwise.complete.obs") -->

<!-- # print("weighted correlation matrix") -->

<!-- cor_matrix <-wdm(cor_test %>% select(all_of(kendall_cor_vars)), method = "kendall",  -->
<!--                          weights = cor_test$hh_weight_2017_2019) -->

<!-- corrplot::corrplot(cor_matrix, method = 'number', type = 'upper') -->
<!-- ``` -->

<!-- - correlation significance -->
<!-- ```{r correlation significance} -->

<!-- cor_sig_test <- function(var){ -->
<!--   cor.test(cor_test$vmt_day, pull(cor_test, {{var}}), method="kendall")$p.value -->
<!-- } -->

<!-- cor_results <- as.data.frame(cor_matrix) %>%  -->
<!--   select(1) %>% -->
<!--   mutate(vmt_day_p.value=map_dbl(kendall_cor_vars, cor_sig_test)) # experiment with purrr: mission complete -->
<!-- cor_results -->
<!-- ``` -->

<!-- ## correlation matrix for binary variables -->
<!-- ```{r} -->
<!-- cor_matrix_bi <-wdm(cor_test %>% select(all_of(point_biserial_cor_vars)),  -->
<!--                          weights = cor_test$hh_weight_2017_2019) -->


<!-- corrplot::corrplot(cor_matrix_bi, method = 'number', type = 'upper') -->
<!-- ``` -->

<!-- - correlation significance -->
<!-- ```{r} -->

<!-- cor_sig_test <- function(var){ -->
<!--   cor.test(cor_test$vmt_day, pull(cor_test, {{var}}))$p.value -->
<!-- } -->

<!-- cor_results_bi <- as.data.frame(cor_matrix_bi) %>%  -->
<!--   select(1) %>% -->
<!--   mutate(vmt_day_p.value=map_dbl(point_biserial_cor_vars, cor_sig_test)) # experiment with purrr: mission complete -->
<!-- cor_results_bi -->
<!-- ``` -->


<!-- ```{r} -->
<!-- test <- cor_test %>% -->
<!--   mutate(SDMVSTRA=1) %>% -->
<!--   filter(hh_weight_2017_2019>0, -->
<!--          vmt_day>0) -->
<!-- design <- svydesign(strata=~SDMVSTRA, id=~person_id, weights=~hh_weight_2017_2019, -->
<!--                         nest=FALSE, data=test) -->
<!-- # Unweighted boxplot -->
<!-- boxplot(test$vmt_day~test$work_parking, -->
<!--        ylab = "vmt", -->
<!--        main = "Unweighted") -->
<!-- # Weighted boxplot -->
<!-- svyboxplot(~vmt_day~work_parking, design, all.outliers = T, -->
<!--      ylab = "vmt", -->
<!--      main = "Work Parking Weighted") -->
<!-- # Weighted boxplot -->
<!-- svyboxplot(~vmt_day~transit_pass, design, all.outliers = T, -->
<!--      ylab = "vmt", -->
<!--      main = "Transit Pass Weighted") -->

<!-- ``` -->

