---
title: "delivery-analysis"
author: "suzanne"
date: "2024-09-11"
output: html_document
---

Summarizing the delivery data is extremely complex because:
1. The question wording has changed over time
2. The question is not asked of all household members
3. The question is a checkbox now
4. Rmove asks the question on multiple days


I'm going to work through the steps very manually to understand what is going on.


```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
library(forcats)
library(sf)
source('survey-23-preprocess.R')
```

Documented question wording

[delivery]
if rMove or (rMove for Web and person 1)
We want to know how delivery and household services contribute to travel. 
On <traveldate> , which of the following occurred? 
Select all that apply.
	Received packages at home (e.g., USPS, FedEx, UPS)
Received personal packages at work if employment = full/part/self/volunteer and job_type = "only one work location" or "teleworks some days and travels to a work location some days"
Received packages at another location (e.g., Amazon locker, package pick-up point)
Take-out/prepared food delivered to home
Groceries delivered to home
Someone came to do work at home (e.g., babysitter, housecleaning, lawn)
Other item delivered to home (e.g., appliance)
None of the above

Documented Variable Values
deliver_elsewhere	0	Not selected
deliver_elsewhere	1	Selected
deliver_elsewhere	995	Missing Response
deliver_food	0	No
deliver_food	1	Yes
deliver_food	995	Missing Response
deliver_grocery	0	No
deliver_grocery	1	Yes
deliver_grocery	995	Missing Response
deliver_none	0	Not selected
deliver_none	1	Selected
deliver_none	995	Missing Response
deliver_office	0	Not selected
deliver_office	1	Selected
deliver_office	995	Missing Response
deliver_other	0	Not selected
deliver_other	1	Selected
deliver_other	995	Missing Response
deliver_package	0	No
deliver_package	1	Yes
deliver_package	995	Missing Response
deliver_work	0	No
deliver_work	1	Yes
deliver_work	995	Missing Response



```{r}
hh<- get_query(sql= "select  num_days_complete_weekday, household_id as hh_id,hh_weight, survey_year, hhgroup
                from HHSurvey.v_households_labels")
```


```{r}
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , daynum, hh_day_iscomplete, deliver_grocery,travel_dow, deliver_elsewhere, deliver_none, deliver_office, deliver_other, deliver_food, deliver_package, survey_year from HHSurvey.v_days_labels")
```

On an average weekday, what is the proportion of households that received a package?
for now

1. Only include 2023 data
2. Filter to Weekdays (travel_dow)- can be done via day weight
3. if anyone got a package (aggregate across person days in hh), count they received
4. Account for multiple weekdays in RMove num_days_complete_weekday
```{r}
hh_day<-hh%>%right_join(day, by='hh_id')
hh_day_23<-hh_day%>%filter(survey_year.x=='2023')


```
```{r}
hh_day_23_wkday<-hh_day_23%>%filter(travel_dow %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday'))
```

this is number of people in the region
```{r}
sum(hh_day_23_wkday$day_weight)
```

I'm going to use day weights from now on

```{r}
hhdays_pkg<-hh_day_23_wkday%>%group_by(hh_id,daynum,deliver_package)%>%summarize(person_days=sum(day_weight))

```

```{r}
hhdays_pkg_1<-hhdays_pkg%>%pivot_wider(names_from=deliver_package, values_from=person_days)%>% replace(is.na(.), 0)
```

```{r}
hh_pkg_agg<-hhdays_pkg_1%>%mutate(corrected_no=if_else(Yes>0, 0, No),
                                  corrected_missing=if_else((Yes>0 | No>0), 0, `Missing Response`))%>%
  mutate('Total'=sum(Yes+corrected_no+corrected_missing))
```
```{r}
summary<-hh_pkg_agg%>%ungroup()%>%summarize(HasPkg=sum(Yes), NoPkg=sum(corrected_no), Total=sum(Total), MissingData=sum(corrected_missing))%>%mutate(share=HasPkg/Total)
```

```{r}
summary
```
What if I do the analysis with household_weights, num of complete weekday and household day is complete?

```{r}
hh_day_23_wkday<-hh_day_23%>%filter(travel_dow %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday'))%>%filter(hh_day_iscomplete=='Yes')%>%mutate(
  num_days_complete_weekday=as.numeric(num_days_complete_weekday), hh_weight=as.numeric(hh_weight))%>%mutate(day_weight_new=hh_weight/num_days_complete_weekday)


hhdays_pkg<-hh_day_23_wkday%>%
  group_by(hh_id,daynum,deliver_package)%>%
  summarize(person_days=sum(day_weight_new), day_weight_old=sum(day_weight))

```

```{r}
hhdays_pkg_1<-hhdays_pkg%>%pivot_wider(names_from=deliver_package, values_from=person_days)%>% replace(is.na(.), 0)
```

```{r}
hh_pkg_agg<-hhdays_pkg_1%>%mutate(corrected_no=if_else(Yes>0, 0, No),
                                  corrected_missing=if_else((Yes>0 | No>0), 0, `Missing Response`))%>%
  mutate('Total'=sum(Yes+corrected_no+corrected_missing))
```
```{r}
summary<-hh_pkg_agg%>%ungroup()%>%summarize(HasPkg=sum(Yes), NoPkg=sum(corrected_no), Total=sum(Total), MissingData=sum(corrected_missing))%>%mutate(share=HasPkg/Total)
```

```{r}
summary
```
