---
title: "modeling vehicle miles traveled"
output:
  html_document:
    df_print: paged
---




```{r }
library(psrc.travelsurvey)
library(dplyr)
library(stringr)
library(ggplot2)
library(psrcplot)
library(forcats)
library(odbc)
library(DBI)
library(tidyr)
library(tidycensus)
library(psrcslides)
library(officer)
library(DBI)
library(stargazer)
library(MASS)
library(stats)
install_psrc_fonts()
```

### Initial list of variables to read in
```{r}
mode_vars<-c('mode_1', 'mode_simple')
other_vars<-c('sample_county',  'dest_purpose_cat', 'origin_purpose_cat', 'trip_path_distance', 'household_id',  'travelers_total', 'person_id', 'trip_adjustment_17_19', 'trip_adjustment_21', 'trip_weight_2017_2019','trip_adult_weight_2021', 'hhgroup', 'nwkdays', 'age_category')

all_vars<-c(mode_vars, other_vars)
household_vars=c('hhid', 'final_home_tract_10', 'hhgroup' )
```

##  Handle 2017/2019 and 2021 separately
## to do: combine them where possible
# 2017 and 2019 data handling

```{r}
trip_data_17_19<- get_hhts(survey="2017_2019", level="t", vars=all_vars)


```

## Accounting for several observed weekdays, need to count observed weekdays i 
```{r}
trip_data_17_19<-trip_data_17_19[,!duplicated(colnames(trip_data_17_19))]%>%mutate(numwkdays=as.numeric(nwkdays))%>%filter(age_category!='Under 18 years')%>%filter(daynum==1)
```
# calculate vmt for 2017/2019
```{r}

driving_data<-trip_data_17_19 %>% dplyr::filter(mode_simple=='Drive') %>%
  mutate(travelers_total=ifelse(is.na(travelers_total), 1, travelers_total))%>%
  mutate(travelers_total=ifelse(travelers_total>10, 1, travelers_total))%>%
  
  mutate(travelers_total=ifelse(travelers_total<1, 1, travelers_total))%>%
  mutate(vmt= trip_path_distance*trip_adjustment_17_19/travelers_total)%>%
  mutate(weighted_vmt= trip_path_distance*trip_weight_2017_2019*numwkdays/travelers_total)
```
# vmt per day, for people who have vmt, restrict to the first observed day for ease of analysis
```{r}
vmt_per<- driving_data%>%group_by(person_id, daynum)%>%
          summarize(vmt_day=sum(vmt),weighted_vmt=sum(weighted_vmt))%>%
          filter(row_number()==1)

```
# get person demographics, restrict to adults for comparability
```{r}

persons_list<- get_hhts(survey="2017_2019", level="p", vars=c('person_id', 'numadults', 'employment', 'final_cnty', 'vehicle_count',  "hhincome_broad",  'hhincome_detailed', 'final_home_is_rgc','hhsize', 'lifecycle', 'age_category', 'survey_year', 'rent_own', 'res_dur', 'numworkers', 'numchildren', 'gender_grp', 'jobs_count', 'student', 'education', 'industry', 'license', 'commute_freq', 'commute_dur', 'telecommute_freq', 'work_county', 'benefits_1', 'benefits_2', 'benefits_3', 'benefits_4', 'rent_own', 'commute_auto_distance', 'res_dur','age', 'car_share', 'telecommute_freq_simple', 'workplace', 'commute_freq', 'hh_weight_2017_2019'))
persons_list<-persons_list[,!duplicated(colnames(persons_list))]%>%filter('age_category'!='Under 18 years')

```

### get the home location information
```{r}
hh_tract<- get_hhts(survey='2017_2019', level = 'h', vars= c( 'final_home_tract10', 'hhgroup'))

persons_tract<-left_join(persons_list,hh_tract,by='household_id')%>%mutate(tract_char= as.character(final_home_tract10))

```

### join the home location information to information about transportation accessibility
```{r}
t_access<- read.csv('transportation_accessibility.csv')%>%mutate(tract_char=as.character(geoid_nm))

persons_access<- left_join(persons_tract,t_access, by='tract_char')
```

## join the persons with their accessibility data and demographic data to observed vmt, if not vmt in the day,
fill in zeros
```{r}
persons_vmt<-left_join(persons_access,vmt_per, by='person_id')%>%
  mutate(vmt_day=replace_na(vmt_day,0))%>%
  mutate(hh_weight_2017_2019_adult=replace_na(hh_weight_2017_2019_adult,0))%>%
  mutate(weighted_vmt=replace_na(weighted_vmt,0))%>%
  mutate(ln_vmt_day=log(1+vmt_day))%>%
  filter(vmt_day<400)

```

```{r}
summary(persons_vmt$vmt_day)
summary(persons_vmt$ln_vmt_day)
ggplot(persons_vmt, aes(ln_vmt_day))+
geom_histogram()+
xlim(0, 10)

ggplot(persons_vmt, aes(vmt_day))+
geom_histogram()
```


```{r}
persons_vmt_new<-persons_vmt%>%mutate(inc100k=case_when(hhincome_broad=='$75,000-$99,999' ~ 'Under 100K',
                                                 hhincome_broad=='$100,000 or more' ~ 'Over 100K',
                                                 hhincome_broad=='Prefer not to answer' ~ 'No Answer',
                                                 hhincome_broad=='$25,000-$49,999' ~ 'Under 100K',
                                                 hhincome_broad=='$50,000-$74,999' ~ 'Under 100K'))%>%
  mutate(vehicle_count_simple= case_when(vehicle_count=substr(vehicle_count,1,1) %in% c("3","4", "5", "6", "7", "8") ~ '3 or more vehicles',
                                         .default = vehicle_count))%>%mutate(telecommute_freq_new = case_when(telecommute_freq %in% c("1 day a week", "2 days a week") ~ "1-4 days per week", 
                                          telecommute_freq %in% c("3 days a week", "4 days a week") ~ "1-4 days per week", 
                                          telecommute_freq %in% c("5 days a week", "6-7 days a week") ~ "5+ days",
                                           telecommute_freq %in% c("Never", "Not applicable", "Less than monthly", "A few times per month") ~ "Never or Less than Every Week",
                                          !is.na(telecommute_freq) ~ telecommute_freq))%>%
  mutate(Kitsap_home= ifelse(final_cnty=='Kitsap', 'Kitsap', 'Not Kitsap'))%>%
           mutate(transit_ben = ifelse(benefits_3 %in% c('Offered, and I use', 'Offered and I don\'t use'), 'Offered', 'Not Offered or do not know'))%>%
           mutate(compressed_ben = ifelse(benefits_2 %in%  c('Offered, and I use', 'Offered and I don\'t use'), 'Offered', 'Not Offered or do not know'))%>%
  mutate(grad_school=ifelse(education=='Graduate/post-graduate degree', 'Graduate degree', 'No graduate degree'))%>%
           mutate(flextime_ben = ifelse(benefits_1 %in%  c('Offered, and I use', 'Offered and I don\'t use'), 'Offered', 'Not Offered or do not know'))%>%
           mutate(oth_ben = ifelse(benefits_4 %in%  c('Offered, and I use', 'Offered and I don\'t use'), 'Offered', 'Not Offered or do not know'))%>%
  mutate(survey_year_factor=as.factor(survey_year))

           

```





## read in 2021 data

```{r}
trip_data__21<- get_hhts(survey="2021", level="t", vars=all_vars)
```

```{r}
trip_data__21<-trip_data__21[,!duplicated(colnames(trip_data__21))]
```
# to do - join back to the person data to get people who don't have any driving
```{r}

driving_data_21<-trip_data__21 %>% dplyr::filter(mode_simple=='Drive')  %>%
  mutate(travelers_total=ifelse(is.na(travelers_total), 1, travelers_total))%>%
  mutate(travelers_total=ifelse(travelers_total>10, 1, travelers_total))%>%
  mutate(travelers_total=ifelse(travelers_total<1, 1, travelers_total))%>%
              mutate(vmt= trip_path_distance*trip_adjustment_21/travelers_total,
                     weighted_vmt= trip_path_distance*trip_adult_weight_2021/travelers_total)
    
```

```{r}
vmt_per_21<- driving_data_21%>%group_by(person_id, daynum)%>%
          summarize(vmt_day=sum(vmt),weighted_vmt=sum(weighted_vmt))%>%
          filter(row_number()==1)
```

```{r}

persons_list_21<- get_hhts(survey="2021", level="p", vars=c('person_id', 'numadults', 'employment', 'final_cnty', 'vehicle_count',  "hhincome_broad",  'hhincome_detailed', 'final_home_is_rgc','hhsize', 'lifecycle', 'age_category', 'survey_year', 'rent_own', 'res_dur', 'numworkers', 'numchildren', 'gender_grp', 'jobs_count', 'student', 'education', 'industry', 'license', 'commute_freq', 'commute_dur', 'telecommute_freq', 'work_county', 'benefits_1', 'benefits_2', 'benefits_3', 'benefits_4', 'rent_own', 'commute_auto_distance', 'res_dur','age', 'car_share', 'telecommute_freq_simple', 'workplace', 'person_adult_weight_2021'))


```

```{r}
persons_list_21<-persons_list_21[,!duplicated(colnames(persons_list_21))]
```

```{r}
hh_tract_21<- get_hhts(survey='2021', level = 'h', vars= c( 'final_home_tract10'))

persons_tract_21<-left_join(persons_list_21,hh_tract_21,by='household_id')%>%mutate(tract_char= as.character(final_home_tract10))

```

```{r}
t_access_21<- read.csv('transportation_accessibility.csv')%>%mutate(tract_char=as.character(geoid_nm))

persons_access_21<- left_join(persons_tract_21,t_access_21, by='tract_char')
```

```{r}
persons_vmt_21<-left_join(persons_access_21,vmt_per_21, by='person_id')%>%
  mutate(vmt_day=replace_na(vmt_day,0))%>%
  mutate(weighted_vmt=replace_na(weighted_vmt,0))%>%
  mutate(ln_vmt_day=log(1+vmt_day))%>%
  filter(vmt_day<400)

```

```{r}
summary(persons_vmt_21$vmt_day)
summary(persons_vmt_21$ln_vmt_day)
ggplot(persons_vmt_21, aes(ln_vmt_day))+
geom_histogram()+
xlim(0, 10)
```




```{r}
persons_vmt_new_21<-persons_vmt_21%>%mutate(inc100k=case_when(hhincome_broad=='$75,000-$99,999' ~ 'Under 100K',
                                                 hhincome_broad=='$100,000-$199,000' ~ 'Over 100K',
                                                 hhincome_broad=='$200,000 or more' ~ 'Over 100K',
                                                 hhincome_broad=='Prefer not to answer' ~ 'No Answer',
                                                 hhincome_broad=='$25,000-$49,999' ~ 'Under 100K',
                                                 hhincome_broad=='$50,000-$74,999' ~ 'Under 100K'))%>%
  mutate(vehicle_count_simple= case_when(vehicle_count=substr(vehicle_count,1,1) %in% c("3","4", "5", "6", "7", "8") ~ '3 or more vehicles',
                                         .default = vehicle_count))%>%
  mutate(telecommute_freq_new = case_when(telecommute_freq %in% c("1-2 days") ~ "1-4 days per week", 
                                          telecommute_freq %in% c("3-4 days") ~ "1-4 days per week", 
                                          telecommute_freq %in% c("5+ days") ~ "5+ days",
                                          telecommute_freq %in% c("Never / None")  ~ "Never or Less than Every Week",
                                          !is.na(telecommute_freq) ~ telecommute_freq))%>%
  mutate(Kitsap_home= ifelse(final_cnty=='Kitsap', 'Kitsap', 'Not Kitsap'))%>%
           mutate(transit_ben = ifelse(benefits_3 %in% c('Offered, and I use', 'Offered and I don\'t use'), 'Offered', 'Not Offered or do not know'))%>%
           mutate(compressed_ben = ifelse(benefits_2 %in%  c('Offered, and I use', 'Offered and I don\'t use'), 'Offered', 'Not Offered or do not know'))%>%
  mutate(grad_school=ifelse(education=='Graduate/post-graduate degree', 'Graduate degree', 'No graduate degree'))%>%
           mutate(flextime_ben = ifelse(benefits_1 %in%  c('Offered, and I use', 'Offered and I don\'t use'), 'Offered', 'Not Offered or do not know'))%>%
           mutate(oth_ben = ifelse(benefits_4 %in%  c('Offered, and I use', 'Offered and I don\'t use'), 'Offered', 'Not Offered or do not know'))%>%
  mutate(survey_year_factor=as.factor(survey_year))


           

```

```{r}
common_cols <- intersect(colnames(persons_vmt_new), colnames(persons_vmt_new_21))
persons_vmt_17_19_21<-rbind(persons_vmt_new[, common_cols], persons_vmt_new_21[, common_cols])
```


```{r}
table(persons_vmt_17_19_21$survey_year)

persons_vmt_17_19_21%>%group_by(survey_year)%>%summarize(mean_vmt=mean(vmt_day))

```




# https://data.library.virginia.edu/interpreting-log-transformations-in-a-linear-model/

```{r}

```


Summarize weighted VMT by weighted characteristic
```{r}
vmt_summary <- function(person_day_df_1, person_weight_name_1,survey_year_1, 
                           person_day_df_2, person_weight_name_2, survey_year_2,
                           group) {
  
  vmt_mean_group_1<-person_day_df_1%>%
    group_by(!!ensym(group))%>%
    summarize(sum_per_weights=sum(!!ensym(person_weight_name_1)), sum_weighted_vmt=sum(weighted_vmt))%>%
    mutate(vmt_mean=sum_weighted_vmt/sum_per_weights)%>%
    mutate(survey_year=survey_year_1)
  
    vmt_mean_group_2<-person_day_df_2%>%
    group_by(!!ensym(group))%>%
    summarize(sum_per_weights=sum(!!ensym(person_weight_name_2)), sum_weighted_vmt=sum(weighted_vmt))%>%
    mutate(vmt_mean=sum_weighted_vmt/sum_per_weights)%>%
    mutate(survey_year=survey_year_2)
    
    vmt_mean_group<-rbind(vmt_mean_group_1, vmt_mean_group_2)%>%
      drop_na()

  return(vmt_mean_group)
}

```

```{r}
person_summary <- function(person_day_df_1, person_weight_name_1,survey_year_1, 
                           person_day_df_2, person_weight_name_2, survey_year_2,
                           group) {
  
  shares_group_1<-person_day_df_1%>%
    group_by(!!ensym(group))%>%
    summarize(sum_per_weights=sum(!!ensym(person_weight_name_1)))%>%
    drop_na()%>%
    mutate(share= sum_per_weights/sum(sum_per_weights))%>%
    mutate(survey_year=survey_year_1)
  
    shares_group_2<-person_day_df_2%>%
    group_by(!!ensym(group))%>%
    summarize(sum_per_weights=sum(!!ensym(person_weight_name_2)))%>%
    drop_na()%>%
    mutate(share= sum_per_weights/sum(sum_per_weights))%>%
    mutate(survey_year=survey_year_2)
    
    shares_group<-rbind(shares_group_1, shares_group_2)
  return(shares_group)
}

```


```{r}
persons_list<-persons_list%>%
  mutate(telecommute_freq_new = case_when(telecommute_freq %in% c("1 day a week", "2 days a week") ~ "1-4 days per week", 
                                          telecommute_freq %in% c("3 days a week", "4 days a week") ~ "1-4 days per week", 
                                          telecommute_freq %in% c("5 days a week", "6-7 days a week") ~ "5+ days",
                                           telecommute_freq %in% c("Never", "Not applicable", "Less than monthly", "A few times per month") ~ "Never or Less than Every Week",
                                          !is.na(telecommute_freq) ~ telecommute_freq,
                                          is.na(telecommute_freq) ~ "Not Worker"))%>%filter(telecommute_freq_new!='Not Worker')
  
persons_list_21<-persons_list_21%>%
  mutate(telecommute_freq_new = case_when(telecommute_freq %in% c("1-2 days") ~ "1-4 days per week", 
                                          telecommute_freq %in% c("3-4 days") ~ "1-4 days per week", 
                                          telecommute_freq %in% c("5+ days") ~ "5+ days",
                                          telecommute_freq %in% c("Never / None")  ~ "Never or Less than Every Week",
                                          !is.na(telecommute_freq) ~ telecommute_freq,
                                          is.na(telecommute_freq) ~ 'Not Worker'))%>%filter(telecommute_freq_new!='Not Worker')
  

telework_share<-person_summary(persons_list,'hh_weight_2017_2019', '2017/2019',
                            persons_list_21,'person_adult_weight_2021', '2021',
                                  'telecommute_freq_new' )%>%
mutate(telecommute_freq_new=factor(telecommute_freq_new, levels=c("Never or Less than Every Week", "1-4 days per week" ,"5+ days")))%>%
         filter(telecommute_freq_new %in% c("Never or Less than Every Week", "1-4 days per week", "5+ days"))


```

```{r}
static_column_chart(t=telework_share, x='telecommute_freq_new', y='share', fill='survey_year', 
                    xlabel='Telecommute Frequency', ylabel='Share of Workers')+ theme(axis.text.x = element_text(size = 14),axis.text.y = element_text(size = 14)) + theme(axis.title = element_text(size=18))
```

```{r}
employment_vmt<-vmt_summary(persons_vmt_new,'hh_weight_2017_2019_adult', '2017/2019',
                            persons_vmt_new_21,'person_adult_weight_2021', '2021',
                                  'employment')%>%filter(employment!='Employed but not currently working (e.g., on leave, furloughed 100%)'
)
```

```{r}
static_column_chart(t=employment_vmt, x='employment', y='vmt_mean', fill='survey_year', 
                    xlabel='Employment Status', ylabel='Vehicle Miles Traveled (mean)')+ theme(axis.text.x = element_text(size = 12),axis.text.y = element_text(size = 12)) + theme(axis.title = element_text(size=12))
```



```{r}
hhinc_vmt<-vmt_summary(persons_vmt_new,'hh_weight_2017_2019_adult', '2017/2019',
                            persons_vmt_new_21,'person_adult_weight_2021', '2021',
                                  'inc100k')
```

```{r}
static_column_chart(t=hhinc_vmt, x='inc100k', y='vmt_mean', fill='survey_year', 
                    xlabel='Household Income', ylabel='vehicle miles traveled (mean)')

```
```{r}
vehicle_count_vmt<-vmt_summary(persons_vmt_new,'hh_weight_2017_2019_adult', '2017/2019',
                            persons_vmt_new_21,'person_adult_weight_2021', '2021',
                                  'vehicle_count_simple')


```

```{r}
telework_vmt<-vmt_summary(persons_vmt_new,'hh_weight_2017_2019_adult', '2017/2019',
                            persons_vmt_new_21,'person_adult_weight_2021', '2021',
                                  'telecommute_freq_new')

```

```{r}
static_column_chart(t=vehicle_count_vmt, x='vehicle_count_simple', y='vmt_mean', fill='survey_year', 
                    xlabel='Household Vehicles', ylabel='vehicle miles traveled (mean)')

```
```{r}
static_column_chart(t=telework_vmt, x='telecommute_freq_new', y='vmt_mean', fill='survey_year', 
                    xlabel='Telecommute Frequency', ylabel='Vehicle Miles Traveled (mean)')+ theme(axis.text.x = element_text(size = 14),axis.text.y = element_text(size = 14)) + theme(axis.title = element_text(size=18))
```
```{r}
transitben_vmt<-vmt_summary(persons_vmt_new,'hh_weight_2017_2019_adult', '2017/2019',
                            persons_vmt_new_21,'person_adult_weight_2021', '2021',
                                  'benefits_3')

```

```{r}
static_column_chart(t=transitben_vmt, x='benefits_3', y='vmt_mean', fill='survey_year', 
                    xlabel='Transit Pass Work Benefit', ylabel='vehicle miles traveled (mean)')


```





```{r}
vmt_model_best_fit_21<-lm(ln_vmt_day ~ employment  + vehicle_count_simple + 
    gender_grp + numworkers + numchildren + numadults + grad_school + 
    student + license + telecommute_freq_new + Kitsap_home +
    transit_ben + flextime_ben + compressed_ben + rent_own + 
     log(1 + jobs_transit/10000) + log(1 + 
    pharmacy) + log(1 + park) +workplace + 
    log(1 + pharmacy)+ survey_year_factor  , data = persons_vmt_17_19_21)

output<-stargazer(vmt_model_best_fit_21, type='text', out='vmt_model_best_21.txt')

```


```{r}
library(broom)
tidy_vmt<-tidy(vmt_model_best_fit_21)
write.csv(tidy_vmt, 'vmt_model.csv')

```

```{r}
library(jtools)
plot_coeff<-plot_summs(vmt_model_best_fit_21)
```




```{r}
library(jtools)
plot_coeff<-plot_summs(vmt_model_best_fit_21)
plot_coeff
```
