---
title: "vmt"
author: "suzanne"
date: "2024-05-28"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
source('../survey-23-preprocess.R')
```



## Read in data items from Elmer

The travelSurveyTools package expects everything to be a data.table, so run setDT all over the place.


```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad, survey_year, hh_weight,home_rgcname
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category, 
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, industry,office_available, commute_freq, employment
                from HHSurvey.v_persons_labels")
                
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,driver,
person_id, mode_1, dest_purpose, travelers_total, survey_year, distance_miles,trip_weight 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)
setDT(day)
setDT(trip)
```
#race_category needs to be added to the codebook

# to do: add Seattle, Bellevue geographies

## Read in Codebook
```{r , echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```


## Set IDs as characters
I guess for joining?, functionalize convert all ids to charcets
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
day[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
day[, person_id := as.character(person_id)]
trip[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
trip[, day_id := as.character(day_id)]

trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
day <- day%>%mutate(survey_year=as.character(survey_year))
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```

## Make a new variable

If your variable is going to be a transformation of existing data in the codebook- for example grouping- you need to add the new variable and its new values to the internal to code, working codebook.


```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='dest_purpose_simple')
value_labels<-add_values_code(group_name='dest_purpose_simple')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='dest_purpose', grouped_name='dest_purpose_simple')
```


```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='mode_simple')
value_labels<-add_values_code(group_name='mode_simple')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='mode_1', grouped_name='mode_simple')
```

```{r}
variable_list<-add_variable(variable_list, 'vmt','day',data_type='numeric')
```


```{r}
variable_list<-add_variable(variable_list, 'travelers_total_num','trip',data_type='numeric')
```



```{r}

driving_trip<-trip %>% dplyr::filter(mode_simple=='Drive') %>%
  filter(!is.na(travelers_total) | !is.na(distance_miles)| 
          !(travelers_total=='Missing Response'))%>% # remove missing data
  mutate(travelers_total_num=as.numeric(substring(travelers_total,1,1)))%>%
  mutate(travelers_total_num=if_else(is.na(travelers_total_num), 1,travelers_total_num))%>%
  mutate(travelers_total_num=
            ifelse(travelers_total_num>=5, 7, travelers_total_num)) # have to make some assumption for 5+
           


```

# this will account for the adjustments on trips as well.

```{r}
driving_trip<-left_join(driving_trip, hh, by='hh_id')
driving_trip<-driving_trip%>%mutate(weighted_vmt=distance_miles*trip_weight/travelers_total_num)%>%
  mutate(weighted_vmt=ifelse(is.na(weighted_vmt), 0, weighted_vmt))%>%
  mutate(non_adj_vmt=(distance_miles/travelers_total_num))%>%
  mutate(unweighted_vmt=(distance_miles*trip_weight)/(travelers_total_num*hh_weight))%>%
  mutate(unweighted_vmt=ifelse(is.na(unweighted_vmt), 0, unweighted_vmt))%>%
  filter(distance_miles<200)

```

```{r}
driving_day<-driving_trip%>%group_by(day_id)%>%summarize(total_vmt=sum(weighted_vmt), vmt_est_day=sum(unweighted_vmt))
```




```{r}
all_days<-left_join(day,driving_day, by='day_id' )%>%mutate(total_vmt=if_else(is.na(total_vmt), 0, total_vmt))%>%mutate(vmt=if_else(is.na(vmt_est_day), 0, vmt_est_day))%>%filter(day_weight!=0)
```


```{r}
center.lyr <- st_read_elmergeo('URBAN_CENTERS')
```

```{r}
variable_list<-add_variable(variable_list, 'rgc_cat_home','hh')
```

```{r}
variable_list<-add_variable(variable_list, 'home_rgc','hh')
```

```{r}

value<- c(0,1,2)
label=c( 'Not RGC','Urban RGC', 'Metro RGC')
val_labels<-data.frame(value, label)
val_labels<-val_labels%>%mutate(variable='rgc_cat_home')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  
```

```{r}

value<- c(0,1)
label=c( 'Not RGC','RGC')
val_labels<-data.frame(value, label)
val_labels<-val_labels%>%mutate(variable='home_rgc')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  
```

```{r}
hh<-left_join(hh, center.lyr, by=join_by(home_rgcname==name), suffix= c("", '.origin'))
```

```{r }
hh<-hh%>%mutate(rgc_cat_home= 
                      case_when(is.na(category) ~ 'NotRGC',
                                category=='Metro' ~ 'MetroRGC',
                                category=='Urban' ~ 'UrbanRGC',
                                TRUE ~'Missing')
)%>%mutate(home_rgc=
  if_else(rgc_cat_home %in% c('MetroRGC','UrbanRGC'), 'RGC', 'Not RGC'))
```


                    

```{r}
hh_person_day<-list(hh=hh, day=all_days)

```

# by applying the weights in this manner, so we can look at individual days
# the resulting estimates are lower than if you apply the trip weights, and day weights together,
# I think because of sum product impacts
```{r}
all_days<-all_days%>%mutate(vmt_day_wt=vmt_est_day*day_weight)%>%mutate(vmt_day_wt=if_else(is.na(vmt_day_wt), 0, vmt_day_wt))
vmt_year<-all_days%>%group_by(survey_year)%>%summarize(VMT_Year=sum(vmt_day_wt))
people_year<-person%>%group_by(survey_year)%>%summarize(all_people=sum(person_weight))

vmt_people_year<-merge(vmt_year, people_year, by='survey_year')%>%mutate(vmt_per_person=VMT_Year/all_people)

vmt_people_year
```

```{r}
days_hh<-left_join(all_days, hh, by='hh_id')
vmt_year_rgc<-days_hh%>%group_by(survey_year.x, rgc_cat_home)%>%summarize(VMT_Year=sum(vmt_day_wt))
person_hh<-merge(person,hh, by=('hh_id'))

people_year_rgc<-person_hh%>%group_by(survey_year.x, rgc_cat_home)%>%summarize(all_people=sum(person_weight))

vmt_people_year<-merge(vmt_year_rgc, people_year_rgc, by=c('survey_year.x', 'rgc_cat_home'))%>%mutate(vmt_per_person=VMT_Year/all_people)

```




