---
title: "vmt"
author: "suzanne"
date: "2024-05-28"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
source('../survey-23-preprocess.R')
```



## Read in data items from Elmer

The travelSurveyTools package expects everything to be a data.table, so run setDT all over the place.


```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad,survey_year, hh_weight
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category,
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, industry,office_available, commute_freq, employment
                from HHSurvey.v_persons_labels")
                
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,driver,
person_id, mode_1, dest_purpose, travelers_total, survey_year, distance_miles,trip_weight 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)
setDT(day)
setDT(trip)
```
#race_category needs to be added to the codebook

# to do: add Seattle, Bellevue geographies

## Read in Codebook
```{r , echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```


## Set IDs as characters
I guess for joining?, functionalize convert all ids to charcets
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
day[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
day[, person_id := as.character(person_id)]
trip[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
trip[, day_id := as.character(day_id)]

trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
day <- day%>%mutate(survey_year=as.character(survey_year))
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```

## Make a new variable

If your variable is going to be a transformation of existing data in the codebook- for example grouping- you need to add the new variable and its new values to the internal to code, working codebook.


```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='dest_purpose_simple')
value_labels<-add_values_code(group_name='dest_purpose_simple')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='dest_purpose', grouped_name='dest_purpose_simple')
```


```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='mode_simple')
value_labels<-add_values_code(group_name='mode_simple')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='mode_1', grouped_name='mode_simple')
```

```{r}
variable_list<-add_variable(variable_list, 'vmt','trip',data_type='numeric')
```


```{r}
variable_list<-add_variable(variable_list, 'travelers_total_num','trip',data_type='numeric')
```

```{r}

```

```{r}

driving_trip<-trip %>% dplyr::filter(mode_simple=='Drive') %>%
  filter(!is.na(travelers_total) | !is.na(distance_miles)| 
          !(travelers_total=='Missing Response'))%>% # remove missing data
  mutate(travelers_total_num=as.numeric(substring(travelers_total,1,1)))%>%
  filter(!is.na(travelers_total_num))%>%
  mutate(travelers_total_num=
            ifelse(travelers_total_num>=5, 7, travelers_total_num)) # have to make some assumption for 5+
           


```

```{r}
travelers_summ<-driving_trip%>%group_by(survey_year, travelers_total, travelers_total_num)%>%summarize(weighted_trips=sum(trip_weight), count_trips=n())
travelers_summ
write.csv(travelers_summ, 'travelers_per_drive_trip.csv')
```
```{r}
travelers_sum_chart<-travelers_summ%>%mutate(travelers_total_char=as.character(travelers_total_num))
static_bar_chart(travelers_sum_chart%>%ungroup(), x='weighted_trips', y='travelers_total_char', fill='survey_year')

```

```{r}
driving_trip<-driving_trip%>%mutate(weighted_vmt=distance_miles*trip_weight/travelers_total_num)%>%
  mutate(weighted_vmt=ifelse(is.na(weighted_vmt), 0, weighted_vmt))%>%
  mutate(unweighted_vmt=distance_miles/travelers_total_num)%>%
  mutate(unweighted_vmt=ifelse(is.na(weighted_vmt), 0, unweighted_vmt))%>%
  filter(distance_miles<200)

```


```{r}

driving_year<-driving_trip%>%group_by(survey_year)%>%
  summarize(total_vmt=sum(weighted_vmt), total_trips=sum(trip_weight))%>%
  filter(survey_year %in% c('2019', '2023'))
  
```

```{r}
people<-person%>%group_by(survey_year)%>%filter(survey_year %in% c('2019', '2023'))%>%
  summarise(people=sum(person_weight))
people
  
```




```{r}
driving_per_person<-merge(driving_year, people, by = 'survey_year')%>%mutate(vmt_per_person=total_vmt/people, trips_per_person=total_trips/people)
driving_per_person
write.csv(driving_per_person, 'driving_per_person.csv')
```




```{r}
vmt_year<-driving_trip%>%
group_by(survey_year,dest_purpose_simple)%>%
summarize(sum_vmt=sum(weighted_vmt), sum_trips=sum(trip_weight), sum_pmt=sum(distance_miles*trip_weight),  mean_travelers=mean(travelers_total_num))%>%filter(survey_year%in% c('2019','2023'))%>%
mutate(sum_vmt=round(sum_vmt, -3))
vmt_year
write.csv(vmt_year, 'vmt_purpose.csv')
```
```{r}

static_bar_chart(t=vmt_year,y='dest_purpose_simple', x='sum_vmt', fill='survey_year')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
```

```{r}
vmt_per_person<-merge(vmt_year, people, by = 'survey_year')%>%mutate(vmt_per_person=sum_vmt/people, pmt_per_person=sum_pmt/people)
```

```{r}
vmt_per_person_agg<-vmt_per_person%>%group_by(survey_year)%>%
  summarize(total_vmt=sum(sum_vmt),vmt_per_capita=sum(vmt_per_person), total_pmt=sum(sum_pmt), pmt_per_capita=sum(pmt_per_person))
                                                                       
vmt_per_person_agg                                                                     
```




```{r}
hts_data = list(hh=hh,
                person=person,
                day=day,
                trip = driving_trip)
ids = c('hh_id', 'person_id','day_id', 'trip_id')
wts = c('hh_weight', 'person_weight', 'day_weight', 'trip_weight')
```
```{r}
travellers_summ<-summarize_weighted(hts_data= hts_data,
                               summarize_var = 'travelers_total_num',
                               summarize_by = 'survey_year',
                               id_cols=ids,
                               wt_cols=wts,
                               wtname='trip_weight',
                               summarize_vartype='numeric'
                               )

```
```{r}
travellers_summ$summary$wtd
```


```{r}
length_summ<-summarize_weighted(hts_data= hts_data,
                               summarize_var = 'distance_miles',
                               summarize_by = 'survey_year',
                               id_cols=ids,
                               wt_cols=wts,
                               wtname='trip_weight',
                               summarize_vartype='numeric'
                               )

```


```{r}
length_summ$summary$wtd
```


```{r}
length_purpose_summ<-summarize_weighted(hts_data= hts_data,
                               summarize_var = 'distance_miles',
                               summarize_by = c('survey_year', 'dest_purpose_simple'),
                               id_cols=ids,
                               wt_cols=wts,
                               wtname='trip_weight',
                               summarize_vartype='numeric'
                               )

```
```{r}
write.csv( length_purpose_summ[["summary"]][["wtd"]]%>%filter(survey_year %in% c('2019', '2023')), 'purpose_length.csv')
```

```{r}
person<-merge(person, hh, by='hh_id')

```


```{r}

persons_vmt<-left_join(person, driving_trip, by = 'person_id')%>%
  mutate(weighted_vmt=replace_na(weighted_vmt, 0))


person_vmt_summary<-persons_vmt%>%group_by(survey_year.x, hhincome_broad)%>%
  summarize(total_vmt=sum(weighted_vmt))%>%
  mutate(total_vmt=replace_na(total_vmt,0))%>%
  filter(survey_year.x %in% c('2019', '2023'))%>%rename(survey_year=survey_year.x)


person_inc<-person%>%group_by(hhincome_broad,survey_year.x)%>%summarize(people=sum(person_weight))%>%rename(survey_year=survey_year.x)

person_vmt_summary_inc<-merge(person_vmt_summary, person_inc, by=c('hhincome_broad', 'survey_year'))%>%mutate(vmt_per_person=total_vmt/people)

person_vmt_inc<-order_factors(person_vmt_summary_inc, 'hhincome_broad', value_labels)%>%filter(hhincome_broad!='Prefer not to answer')
```

```{r}
static_column_chart(person_vmt_inc, x='hhincome_broad', y='vmt_per_person', fill='survey_year')
```


```{r}

persons_vmt<-left_join(person, driving_trip, by = 'person_id')%>%
  mutate(weighted_vmt=replace_na(weighted_vmt, 0))


person_vmt_summary<-persons_vmt%>%group_by(survey_year.x, race_category)%>%
  summarize(total_vmt=sum(weighted_vmt))%>%
  mutate(total_vmt=replace_na(total_vmt,0))%>%
  filter(survey_year.x %in% c('2019', '2023'))%>%rename(survey_year=survey_year.x)


person_inc<-person%>%group_by(race_category,survey_year.x)%>%summarize(people=sum(person_weight))%>%rename(survey_year=survey_year.x)

person_vmt_summary_inc<-merge(person_vmt_summary, person_inc, by=c('race_category', 'survey_year'))%>%mutate(vmt_per_person=total_vmt/people)

person_vmt_inc<-order_factors(person_vmt_summary_inc, 'race_category', value_labels)
```

```{r}
static_column_chart(person_vmt_inc, x='race_category', y='vmt_per_person', fill='survey_year')
```

```{r}
people<-person%>%group_by(survey_year)%>%filter(survey_year %in% c('2019', '2023'))%>%
  summarise(people=sum(person_weight))
people
  
```
