---
title: "vmt"
author: "suzanne"
date: "2024-05-28"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
source('../survey-23-preprocess.R')
```



## Read in data items from Elmer

The travelSurveyTools package expects everything to be a data.table, so run setDT all over the place.


```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad, survey_year, hh_weight,home_rgcname
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category, 
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, industry,office_available, commute_freq, employment
                from HHSurvey.v_persons_labels")
                
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,driver,
person_id, mode_1, dest_purpose, travelers_total, survey_year, distance_miles,trip_weight 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)
setDT(day)
setDT(trip)
```
#race_category needs to be added to the codebook

# to do: add Seattle, Bellevue geographies

## Read in Codebook
```{r , echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```


## Set IDs as characters
I guess for joining?, functionalize convert all ids to charcets
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
day[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
day[, person_id := as.character(person_id)]
trip[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
trip[, day_id := as.character(day_id)]

trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
day <- day%>%mutate(survey_year=as.character(survey_year))
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```

## Make a new variable

If your variable is going to be a transformation of existing data in the codebook- for example grouping- you need to add the new variable and its new values to the internal to code, working codebook.


```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='dest_purpose_simple')
value_labels<-add_values_code(group_name='dest_purpose_simple')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='dest_purpose', grouped_name='dest_purpose_simple')
```


```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='mode_simple')
value_labels<-add_values_code(group_name='mode_simple')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='mode_1', grouped_name='mode_simple')
```

```{r}
variable_list<-add_variable(variable_list, 'vmt','day',data_type='numeric')
```


```{r}
variable_list<-add_variable(variable_list, 'travelers_total_num','trip',data_type='numeric')
```



```{r}

driving_trip<-trip %>% dplyr::filter(mode_simple=='Drive') %>%
  filter(!is.na(travelers_total) | !is.na(distance_miles)| 
          !(travelers_total=='Missing Response'))%>% # remove missing data
  mutate(travelers_total_num=as.numeric(substring(travelers_total,1,1)))%>%
  mutate(travelers_total_num=if_else(is.na(travelers_total_num), 1,travelers_total_num))%>%
  mutate(travelers_total_num=
            ifelse(travelers_total_num>=5, 7, travelers_total_num)) # have to make some assumption for 5+
           


```

# this will account for the adjustments on trips as well.

```{r}
driving_trip<-left_join(driving_trip, hh, by='hh_id')
driving_trip<-driving_trip%>%mutate(weighted_vmt=distance_miles*trip_weight/travelers_total_num)%>%
  mutate(weighted_vmt=ifelse(is.na(weighted_vmt), 0, weighted_vmt))%>%
  mutate(non_adj_vmt=(distance_miles/travelers_total_num))%>%
  mutate(unweighted_vmt=(distance_miles*trip_weight)/(travelers_total_num*hh_weight))%>%
  mutate(unweighted_vmt=ifelse(is.na(unweighted_vmt), 0, unweighted_vmt))%>%
  filter(distance_miles<200)

```

```{r}
driving_day<-driving_trip%>%group_by(day_id)%>%summarize(total_vmt=sum(weighted_vmt), vmt_est_day=sum(unweighted_vmt))
```


```{r}
all_days<-left_join(day,driving_day, by='day_id' )%>%mutate(total_vmt=if_else(is.na(total_vmt), 0, total_vmt))%>%mutate(vmt=if_else(is.na(vmt_est_day), 0, vmt_est_day))%>%filter(day_weight!=0)
```


```{r}
hh_person<-merge(hh, person, by ='hh_id')
```


```{r}
all_days_person<-merge(hh_person, all_days, by ='person_id')%>%filter(vmt<200)
```

```{r}
summary(all_days_person$vmt)
```



```{r}
all_days_person<-all_days_person%>%mutate(ln_vmt=log(vmt+1))
```


```{r}
model<-lm(ln_vmt ~survey_year+hhincome_broad+race_category+workplace+telecommute_freq+gender+commute_freq,data=all_days_person)
summary(model)

```

```{r}
hhincome_broad+race_category+workplace+ age+industry, 
```



