---
title: "vmt"
author: "suzanne"
date: "2024-05-28"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
library(sf)
source('../survey-23-preprocess.R')
```



## Read in data items from Elmer

The travelSurveyTools package expects everything to be a data.table, so run setDT all over the place.


```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad, home_lat, home_lng,survey_year, hh_weight,home_rgcname
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category, 
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, industry,office_available, commute_freq, employment
                from HHSurvey.v_persons_labels")
                
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,driver,
person_id, mode_1, dest_purpose, travelers_total, survey_year, distance_miles,trip_weight 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)
setDT(day)
setDT(trip)
```
#race_category needs to be added to the codebook

# to do: add Seattle, Bellevue geographies

## Read in Codebook
```{r , echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```


## Set IDs as characters
I guess for joining?, functionalize convert all ids to charcets
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
day[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
day[, person_id := as.character(person_id)]
trip[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
trip[, day_id := as.character(day_id)]

trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
day <- day%>%mutate(survey_year=as.character(survey_year))
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```

## Make a new variable

If your variable is going to be a transformation of existing data in the codebook- for example grouping- you need to add the new variable and its new values to the internal to code, working codebook.


```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='dest_purpose_simple')
value_labels<-add_values_code(group_name='dest_purpose_simple')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='dest_purpose', grouped_name='dest_purpose_simple')
```


```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='mode_simple')
value_labels<-add_values_code(group_name='mode_simple')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='mode_1', grouped_name='mode_simple')
```

```{r}
variable_list<-add_variable(variable_list, 'vmt','trip',data_type='numeric')
```


```{r}
variable_list<-add_variable(variable_list, 'travelers_total_num','trip',data_type='numeric')
```



```{r}

driving_trip<-trip %>% dplyr::filter(mode_simple=='Drive') %>%
  filter(!is.na(travelers_total) | !is.na(distance_miles)| 
          !(travelers_total=='Missing Response'))%>% # remove missing data
  mutate(travelers_total_num=as.numeric(substring(travelers_total,1,1)))%>%
  mutate(travelers_total_num=if_else(is.na(travelers_total_num), 1,travelers_total_num))%>%
  mutate(travelers_total_num=
            ifelse(travelers_total_num>=5, 7, travelers_total_num)) # have to make some assumption for 5+
           


```



```{r}

driving_trip<-driving_trip%>%mutate(weighted_vmt=distance_miles*trip_weight/travelers_total_num)%>%
  mutate(weighted_vmt=ifelse(is.na(weighted_vmt), 0, weighted_vmt))%>%
  mutate(unweighted_vmt=distance_miles/travelers_total_num)%>%
  mutate(unweighted_vmt=ifelse(is.na(unweighted_vmt), 0, unweighted_vmt))%>%
  filter(distance_miles<200)

```

```{r}
driving_day<-driving_trip%>%group_by(day_id)%>%summarize(total_vmt=sum(weighted_vmt), vmt_est_day=sum(unweighted_vmt))
```




```{r}

driving_year<-driving_trip%>%group_by(survey_year)%>%
  summarize(total_vmt=sum(weighted_vmt), total_trips=sum(trip_weight))
```

```{r}
people<-person%>%group_by(survey_year)%>%
  summarise(people=sum(person_weight))
people
  
```




```{r}
driving_per_person<-merge(driving_year, people, by = 'survey_year')%>%mutate(vmt_per_person=total_vmt/people, trips_per_person=total_trips/people)
driving_per_person
write.csv(driving_per_person, 'driving_per_person.csv')
```

# read in all the geographies

```{r}
uga.lyr<-st_read_elmergeo('URBAN_GROWTH_AREA_evw')
```


```{r}
center.lyr <- st_read_elmergeo('URBAN_CENTERS_evw')
```



```{r}
hh_geo<-hh%>%
  st_as_sf(coords=c('home_lng','home_lat'), crs=4326)
```


```{r}
hh_in_uga<-st_filter(hh_geo, uga.lyr)
```

```{r}
hh_in_uga<-hh_in_uga%>%mutate(in_uga=1)
```

```{r}
hh<-left_join(hh, hh_in_uga)
hh<-hh%>%  mutate(in_uga=if_else(is.na(in_uga), 0, 1))
```




```{r}
hh<-left_join(hh, center.lyr, by=join_by(home_rgcname==name), suffix= c("", '.origin'))
```







```{r}
hh_in_rural<-st_filter(hh_geo, rural)
```










```{r }
hh<-hh%>%mutate(rgc_cat_home= 
                      case_when(is.na(category) ~ 'NotRGC',
                                category=='Metro' ~ 'MetroRGC',
                                category=='Urban' ~ 'UrbanRGC',
                                TRUE ~'Missing')
)
```
                    


```{r}
hh%>%group_by(rgc_cat_home, survey_year)%>%summarize( count=n())
```






```{r}
hh<-hh%>%mutate(rgc_uga=paste(in_uga, rgc_cat_home, sep='_'))
hh_day<-right_join(hh, day, by='hh_id')%>%mutate(survey_year=survey_year.x)
```


```{r}
day_rgc_cat<-hh_day%>%group_by(survey_year,rgc_cat_home)%>%summarize(person_days=sum(day_weight))%>%mutate(home_rgc=
 if_else(rgc_cat_home %in% c('MetroRGC','UrbanRGC'), 'RGC', 'Not RGC'))
```
```{r}
day_rgc_uga<-hh_day%>%group_by(survey_year,rgc_uga)%>%summarize(person_days=sum(day_weight))
```

```{r}
day_vmt<-left_join(hh, driving_trip, by = 'hh_id')%>%
  mutate(weighted_vmt=replace_na(weighted_vmt, 0))%>%rename(survey_year=survey_year.x)
day_vmt<-day_vmt%>%mutate(home_rgc=
 if_else(rgc_cat_home %in% c('MetroRGC','UrbanRGC'), 'RGC', 'Not RGC'))

```

```{r}
day_vmt_rgc<-day_vmt%>%group_by(survey_year, rgc_cat_home)%>%
  summarize(total_vmt=sum(weighted_vmt))%>%
  mutate(total_vmt=replace_na(total_vmt,0))



day_vmt_summary_rgc<-merge(day_vmt_rgc, day_rgc_cat, by=c( 'survey_year', 'rgc_cat_home'))%>%mutate(vmt_per_person=total_vmt/person_days)

day_vmt_summary_rgc


```

```{r}

day_vmt_summary_2<-day_vmt_summary_rgc%>%group_by(survey_year, home_rgc)%>%
  summarize(vmt_rgc=sum(total_vmt), person_days_rgc=sum(person_days))%>%
  mutate(vmt_home_rgc=vmt_rgc/person_days_rgc)

day_vmt_summary_2




```



```{r}
day_vmt_rgc_uga<-day_vmt%>%group_by(survey_year, rgc_uga)%>%
  summarize(total_vmt=sum(weighted_vmt))%>%
  mutate(total_vmt=replace_na(total_vmt,0))



day_vmt_summary_rgc_uga<-merge(day_vmt_rgc_uga, day_rgc_uga, by=c( 'survey_year', 'rgc_uga'))%>%mutate(vmt_per_person=total_vmt/person_days)

day_vmt_summary_rgc_uga


```

do a couple more geographies






#redo analysis with regional geographies
```{r}
regional.geogs<-st_read_elmergeo('REGIONAL_GEOGRAPHIES_evw')
```



```{r}
sf_use_s2(FALSE)
hh_reg_geog<-st_join(regional.geogs, hh_geo, join=st_intersects)%>%filter(!is.na(hh_id))
```

```{r}
hh<-left_join(hh, hh_reg_geog, by='hh_id')
```
```{r}
hh_reg_geog_year<-hh%>%group_by(class_desc, survey_year)%>%summarize( count=n())
```

```{r}

hh_day<-right_join(hh, day, by='hh_id')%>%mutate(survey_year=survey_year.x)
```

```{r}
day_reg_geo<-hh_day%>%group_by(survey_year,class_desc)%>%summarize(person_days=sum(day_weight))
```

```{r}
day_vmt_reg_geo<-day_vmt%>%group_by(survey_year, class_desc)%>%
  summarize(total_vmt=sum(weighted_vmt))%>%
  mutate(total_vmt=replace_na(total_vmt,0))



day_vmt_summary_reg_geo<-merge(day_vmt_reg_geo, day_reg_geo, by=c( 'survey_year', 'class_desc'))%>%mutate(vmt_per_person=total_vmt/person_days)

day_vmt_summary_reg_geo


```
