---
title: "telework stats"
date: today
---

```{r}
library(psrc.travelsurvey)
library(tidyverse)
library(psrcelmer)
library(data.table)
library(magrittr)

telework_vars <- c("telecommute_freq","workplace","industry","commute_freq")

# Retrieve the data
hts_data <- get_psrc_hts(survey_years = 2023, survey_vars = telework_vars)
```

```{r}
goods_producing <- c("Natural resources (e.g., forestry, fishery, energy)",
  "Manufacturing (e.g., aerospace & defense, electrical, machinery)",
  "Construction")
transportation <- c("Transportation and utilities")

full_hts_data <- hts_data
full_hts_data$person <- full_hts_data$person %>%
  mutate(telecommute_freq_mine = case_when(telecommute_freq %in% c("5 days a week","6-7 days a week")~"5+ days a week",
                                           telecommute_freq %in% c("A few times per month","Less than monthly",
                                                                   "Never")~"Never or Less than 1 day a week",  
                                           TRUE~telecommute_freq),
         commute_freq_mine = case_when(commute_freq %in% c("5 days a week","6-7 days a week")~"5+ days a week",
                                           commute_freq %in% c("A few times per month","Less than monthly",
                                                               "1-3 days a month","Never")~"Never or Less than 1 day a week",  
                                           TRUE~commute_freq),
         industry_sector = case_when(industry %in% goods_producing~"Goods Producing industry",
                                     industry %in% transportation~"Trade, Transportation, and Utilities industry sectors",
                                     industry=="Missing: Skip Logic"~"Missing: Skip Logic",
                                     is.na(industry)~NA,
                                     TRUE~"All Other Services industry sectors"))


hybrid_hts_data <- full_hts_data
hybrid_hts_data$person <- hybrid_hts_data$person %>% 
  mutate(workplace= case_when(workplace!="Telework some days and travel to a work location some days"~"not hybrid",
                              TRUE~"Telework some days"),
         telecommute_freq= case_when(workplace!="Telework some days"~"not hybrid",
                              TRUE~telecommute_freq),
         commute_freq= case_when(workplace!="Telework some days"~"not hybrid",
                              TRUE~commute_freq))
```

## basic summary for telework-related variables

- `workplace`
```{r}
psrc_hts_stat(full_hts_data, "person", c("workplace"),incl_na = FALSE) %>%
  mutate_at(c("prop","prop_moe"), ~scales::percent(., accuracy=0.01))
```

- `commute_freq`

_if employment = full/part/self/volunteer and job_type IS NOT "work only from home" or “drive/bike/travel for work”_

```{r}
psrc_hts_stat(full_hts_data, "person", c("commute_freq_mine"),incl_na = FALSE) %>%
  mutate_at(c("prop","prop_moe"), ~scales::percent(., accuracy=0.01))
```
## Hybrid worker only

- keep only workers with workplace as `"Telework some days and travel to a work location some days"`
- how many days a week are hybrid workers commuting?

```{r}
psrc_hts_stat(hybrid_hts_data, "person", c("commute_freq_mine"),incl_na = FALSE) %>%
  mutate_at(c("prop","prop_moe"), ~scales::percent(., accuracy=0.01)) %>%
  filter(commute_freq_mine!="not hybrid")
```
- hybrid worker commute frequency by industry

```{r}
psrc_hts_stat(hybrid_hts_data, "person", c("industry_sector","commute_freq_mine"),incl_na = FALSE) %>%
  mutate_at(c("prop","prop_moe"), ~scales::percent(., accuracy=0.01)) %>%
  filter(commute_freq_mine!="not hybrid") %>%
  pivot_wider(id_cols = c("commute_freq_mine"), names_from = "industry_sector", values_from = "prop")
```

- industry list in HTS data

```{r}
full_hts_data$person %>% 
  select(industry_sector,industry) %>% 
  unique() %>%
  arrange(industry_sector,industry)
```

