---
title: "other variables"
date: today
---

## residential variables

```{r}
library(tidyverse)
library(psrcelmer)

# read data
hh <- get_query(sql= "select * from HHSurvey.v_households_labels where survey_year=2023")
day <- get_query(sql= "select * from HHSurvey.v_days_labels where survey_year=2023")
hh_delivered <- read_csv("J:/Projects/Surveys/HHTravel/Survey2023/Data/old_stuff/data_deliverable_81823/2023/delivered_230907/hh.csv") %>%
  mutate(hhid = as.integer(hhid))

# households with missing residential information
hh_weight_ok <- hh %>% filter(hh_weight>0)
hh_res_miss <- hh_weight_ok %>% filter(rent_own=="Missing Response")
hh_delivered_res_miss <- hh_delivered %>% filter(rent_own==995)

# test <- day %>% filter(household_id==23058160)%>%
#   select(day_id:num_complete_trip_surveys,trips_yesno,day_pct_trips_flagged,survey_year,day_weight)
# day_res_miss <- day %>% filter(household_id %in% hh_res_miss$household_id,
#                                pernum==1#, daynum==4
#                                ) %>%
#   select(day_id:num_complete_trip_surveys,trips_yesno,day_pct_trips_flagged,survey_year,day_weight) %>%
#   group_by(household_id) %>%
#   mutate(valid_day4 = sum(daynum==4 & summary_complete=="Yes"))
```


- issue: `{r} table(hh_res_miss$rent_own)[["Missing Response"]]` households with missing responses in residential variables (`rent_own`, `res_dur`, `res_months`, `res_type`) 
- question logic: if rMove or (rMove for Web and person 1)
  - delivery questions have the same logic but do not have the same missing response issue
- Missing response happened directly in data delivered by RSG: `{r} table(hh_delivered$rent_own)[["995"]]` households
- all households with missing residential variables skipped pre_home variables
  - `{r} table(hh_res_miss$rent_own)[["Missing Response"]]` out of `{r} table(hh_res_miss$prev_rent_own)[["Missing: Skip Logic"]]` households with `prev_rent_own=="Missing: Skip Logic"` are missing residential variables (hh_weight=0 households removed)


```{r}
# write csv for RSG
hh_delivered_miss <- hh_delivered %>% filter(rent_own==995)
hh_miss <- hh %>% filter(rent_own=="Missing Response")
# write.csv(hh_miss, "missing_residential.csv")
hh_miss %>% group_by(hhgroup) %>% summarise(hh_count = n())
```


```{r}
test <- hh_miss %>% filter(hhgroup %in% c("Signup survey completed via rMove, Diary completed via rMove"))
test %>% group_by(hhgroup,numdayscomplete) %>% summarise(hh_count = n())
```









