---
title: "RGC assignment"
date: today
---

- RGC variables:`dest_rgcname`,	`origin_rgcname`,	`home_rgcname`,	`prev_home_rgcname`,  `school_rgcname`,	`work_rgcname`
- Question: how is the RGC assignment performed?
  - temporary method: add a 300ft buffer to RGC ploygons
- Fixing
  - need to reassign Greater Downtown Kirkland for all RGC related variables
  - person_data variables look okay: `school_rgcname`,	`work_rgcname`
  - trip data and household data need fixing:`dest_rgcname`,	`origin_rgcname`,	`home_rgcname`,	`prev_home_rgcname`

```{r}
# 2024/10/18 RGC check
library(psrc.travelsurvey)
library(tidyverse)
library(psrcelmer)
library(sf)
library(leaflet)
sf_use_s2(TRUE)

# list of table names and view names for each data table
table_names <- c('hh','person','day','trip','vehicle')
view_names <- c('v_households_labels','v_persons_labels','v_days_labels','v_trips_labels','v_vehicles_labels')
names(view_names) <- table_names

df_view_name <- data.frame(table = table_names,
                           Elmer.view.name = view_names,
                           row.names = NULL)

# RGC layer
region_center_plot <- st_read_elmergeo('urban_centers_evw')
region_center <- st_read_elmergeo('urban_centers_evw') %>%
  # projection
  st_transform(2926) %>%
  select(name)

region_center_buffer <- st_buffer(region_center, dist = 300) %>%
  rename(name_buffer = name)
region_center_plot_buffer <- region_center_buffer %>%
   st_transform(4326) 


# read HTS data
hh_data <- get_query(sql= paste0("select * from HHSurvey.", view_names['hh']))
person_data <- get_query(sql= paste0("select * from HHSurvey.", view_names['person']))
trip_data <- get_query(sql= paste0("select * from HHSurvey.", view_names['trip']))
```

```{r}
match_rgc_assign <- function(data,rgc_var,weight,point_lng,point_lat,vars_list){
  result <- data %>% 
    filter(.[[weight]]>0, !is.na(.[[point_lng]])) %>%
    select(all_of(vars_list)) %>%
    # create sf layer
    st_as_sf(coords = c(point_lng, point_lat), crs = 4326, remove=FALSE) %>%
    st_transform(2926) %>% 
    # spatial join RGC layer
    st_join(region_center, join = st_intersects) %>%
    st_join(region_center_buffer, join = st_intersects) %>%
    mutate(rgc_check = case_when(.[[rgc_var]]==name~"correct",
                                 is.na(.[[rgc_var]])+is.na(name)==2~"correct",
                                 .[[rgc_var]]=="Not RGC" & is.na(name)~"correct",
                                 .[[rgc_var]]==name_buffer~"correct",
                                 TRUE~"wrong")) %>%
    arrange(desc(rgc_check))
  st_geometry(result) <- NULL
  
  # print(head(result))
  
  return(result)
}

# plot RGC assignment
plot_location <- function(rgc_lyr,rgc_name,lat,lng){
  leaflet(rgc_lyr%>%filter(name==rgc_name)) %>%
    addPolygons(stroke = FALSE) %>%
    addTiles() %>%
    addMarkers(lng,lat)
}

# plot_location("Bothell Canyon Park",47.80259,-122.2181)

```

## trip data

:::{.panel-tabset}
### `dest_rgcname`

```{r}
vars_list <- c("trip_id","survey_year",
               "dest_rgcname","dest_lat","dest_lng")

trip_dest_rgc <- match_rgc_assign(trip_data,"dest_rgcname","trip_weight","dest_lng","dest_lat",vars_list)
```

- a total of `{r} nrow(trip_dest_rgc %>% filter(rgc_check=="wrong"))` mistakes in RGC home assignment
- unassigned Greater Downtown Kirkland records: `{r} nrow(trip_dest_rgc %>% filter(rgc_check=="wrong", name=="Greater Downtown Kirkland"))`
  - unassigned Kirkland distributed across survey years

```{r}
table(trip_dest_rgc %>% filter(rgc_check=="wrong", name=="Greater Downtown Kirkland") %>% select(survey_year))
```

- mistakes in RGC home assignment distributed across survey years (exclude Kirkland)
```{r}
table(trip_dest_rgc %>% filter(rgc_check=="wrong", name!="Greater Downtown Kirkland" | is.na(name)) %>% select(survey_year))
```

```{r}
head(trip_dest_rgc)
```


```{r}
plot_location(region_center_plot,"Lakewood",	47.15033,	-122.5029)
```

### `origin_rgcname`

```{r}
vars_list <- c("trip_id","survey_year",
               "origin_rgcname","origin_lat","origin_lng")
# origin rgc: correct
trip_origin_rgc <- match_rgc_assign(trip_data,"origin_rgcname","trip_weight","origin_lng","origin_lat",vars_list)
```


- a total of `{r} nrow(trip_origin_rgc %>% filter(rgc_check=="wrong"))` mistakes in RGC home assignment
- unassigned Greater Downtown Kirkland records: `{r} nrow(trip_origin_rgc %>% filter(rgc_check=="wrong", name=="Greater Downtown Kirkland"))`  
  - unassigned Kirkland distributed across survey years

```{r}
table(trip_origin_rgc %>% filter(rgc_check=="wrong", name=="Greater Downtown Kirkland") %>% select(survey_year))
```

- mistakes in RGC home assignment distributed across survey years (exclude Kirkland)
```{r}
table(trip_origin_rgc %>% filter(rgc_check=="wrong", name!="Greater Downtown Kirkland" | is.na(name)) %>% select(survey_year))
```

```{r}
head(trip_origin_rgc)
```

```{r}
plot_location(region_center_plot,"Greater Downtown Kirkland",47.67586,	-122.2066)
```
:::

## home data

:::{.panel-tabset}
### `home_rgcname`
```{r}
# home rgc: wrong (192 households)
vars_list <- c("household_id","survey_year","home_rgcname","home_lat","home_lng")

home_rgc <- match_rgc_assign(hh_data,"home_rgcname","hh_weight","home_lng","home_lat",vars_list)
```


- a total of `{r} nrow(home_rgc %>% filter(rgc_check=="wrong"))` mistakes in RGC home assignment
- unassigned Greater Downtown Kirkland records: `{r} nrow(home_rgc %>% filter(rgc_check=="wrong", name=="Greater Downtown Kirkland"))`
  - unassigned Kirkland distributed across survey years

```{r}
table(home_rgc %>% filter(rgc_check=="wrong", name=="Greater Downtown Kirkland") %>% select(survey_year))
```

- mistakes in RGC home assignment distributed across survey years (exclude Kirkland)
```{r}
table(home_rgc %>% filter(rgc_check=="wrong", name!="Greater Downtown Kirkland" | is.na(name)) %>% select(survey_year))
```

```{r}
head(home_rgc)
```

```{r}
plot_location(region_center_plot,"Seattle South Lake Union",	47.61426,	-122.3435	)
```




### `prev_home_rgcname`
```{r}
# home rgc: wrong (192 households)
vars_list <- c("household_id","survey_year","prev_home_rgcname","prev_home_lat","prev_home_lng")

prev_home_rgc <- match_rgc_assign(hh_data,"prev_home_rgcname","hh_weight","prev_home_lng","prev_home_lat",vars_list)
```


- a total of `{r} nrow(prev_home_rgc %>% filter(rgc_check=="wrong"))` mistakes in RGC home assignment
- unassigned Greater Downtown Kirkland records: `{r} nrow(prev_home_rgc %>% filter(rgc_check=="wrong", name=="Greater Downtown Kirkland"))`
    - unassigned Kirkland distributed across survey years

```{r}
table(prev_home_rgc %>% filter(rgc_check=="wrong", name=="Greater Downtown Kirkland") %>% select(survey_year))
```

- mistakes in RGC home assignment distributed across survey years (exclude Kirkland)
```{r}
table(prev_home_rgc %>% filter(rgc_check=="wrong", name!="Greater Downtown Kirkland" | is.na(name)) %>% select(survey_year))
```

```{r}
head(prev_home_rgc)
```

```{r}
plot_location(region_center_plot,"Greater Downtown Kirkland",47.67372,	-122.2026)
```

:::

## person data
:::{.panel-tabset}

### `school_rgcname`
```{r}
vars_list <- c("person_id","survey_year","school_rgcname","school_loc_lat","school_loc_lng")

school_rgc <- match_rgc_assign(person_data,"school_rgcname","person_weight","school_loc_lng","school_loc_lat",vars_list)
```


- a total of `{r} nrow(school_rgc %>% filter(rgc_check=="wrong"))` mistakes in RGC home assignment
- unassigned Greater Downtown Kirkland records: `{r} nrow(school_rgc %>% filter(rgc_check=="wrong", name=="Greater Downtown Kirkland"))`
    - unassigned Kirkland distributed across survey years

```{r}
table(school_rgc %>% filter(rgc_check=="wrong", name=="Greater Downtown Kirkland") %>% select(survey_year))
```

- mistakes in RGC home assignment distributed across survey years (exclude Kirkland)
```{r}
table(school_rgc %>% filter(rgc_check=="wrong", name!="Greater Downtown Kirkland" | is.na(name)) %>% select(survey_year))
```

```{r}
head(school_rgc)
```

```{r}
plot_location(region_center_plot,"Lakewood",47.16091,	-122.5132)
```


### `work_rgcname`

```{r}
vars_list <- c("person_id","survey_year","work_rgcname","work_lat","work_lng")

work_rgc <- match_rgc_assign(person_data,"work_rgcname","person_weight","work_lng","work_lat",vars_list)
```


- a total of `{r} nrow(work_rgc %>% filter(rgc_check=="wrong"))` mistakes in RGC home assignment
- unassigned Greater Downtown Kirkland records: `{r} nrow(work_rgc %>% filter(rgc_check=="wrong", name=="Greater Downtown Kirkland"))`
    - unassigned Kirkland distributed across survey years

```{r}
table(work_rgc %>% filter(rgc_check=="wrong", name=="Greater Downtown Kirkland") %>% select(survey_year))
```

- mistakes in RGC home assignment distributed across survey years (exclude Kirkland)
```{r}
table(work_rgc %>% filter(rgc_check=="wrong", name!="Greater Downtown Kirkland" | is.na(name)) %>% select(survey_year))
```

```{r}
head(work_rgc)
```

```{r}
plot_location(region_center_plot,"Greater Downtown Kirkland",47.67792,	-122.1981)
```
:::