---
title: "student and worker status"
date: today
---

- issues:
  - school location lat/lng misplaced in work location lat/lng
  - incorrect student status for under 18
    - 2017,2019,2023: all school grade students are not assigned student status
    - 2021: Under 5 years old and some 5-11 years old were wrongly assigned as student
  - SOV mode for under 16

```{r}
library(tidyverse)
library(psrcelmer)

# list of table names and view names for each data table
table_names <- c('hh','person','day','trip','vehicle')
view_names <- c('v_households_labels','v_persons_labels','v_days_labels','v_trips_labels','v_vehicles_labels')
names(view_names) <- table_names

df_view_name <- data.frame(table = table_names,
                           Elmer.view.name = view_names,
                           row.names = NULL)

# import all views
hh_data <- get_query(sql= paste0("select * from HHSurvey.", view_names['hh']))
person_data <- get_query(sql= paste0("select * from HHSurvey.", view_names['person']))
# day_data <- get_query(sql= paste0("select * from HHSurvey.", view_names['day']))
trip_data <- get_query(sql= paste0("select * from HHSurvey.", view_names['trip']))
# vehicle_data <- get_query(sql= paste0("select * from HHSurvey.", view_names['vehicle']))

# remove weight==0
hh_data_valid <- hh_data %>% filter(hh_weight>0)
person_data_valid <- person_data %>% filter(person_weight>0)
```


```{r}
person_vars <- c("person_id","household_id","survey_year","age")
school_loc <- c("student","schooltype","school_loc_lat","school_loc_lng","school_in_region","school_county",
                "school_jurisdiction","school_rgcname","school_tract20")
work_loc <- c("employment","work_lat","work_lng","work_in_region","work_county",
              "work_jurisdiction","work_rgcname","work_tract20")

non_school <- c("Daycare","Preschool","Cared for at home")
school_grade_type <- c("Elementary school (public, private, charter)",
                       "K-12 public school",
                       "K-12 private school",
                       "K-12 home school (full-time or part-time)",
                       "Middle school (public, private, charter)",
                       "High school (public, private, charter)")
school_college <- c("College, graduate, or professional school" ,
                    "Vocational/technical school",
                    "Graduate or professional school",                                 
                    "4-year college",
                    "2-year college")
school_other <- c("Other",
                  "Home school")

person_school_work <- person_data_valid %>%
  select(any_of(c(person_vars,"student","schooltype","employment","school_loc_lat","school_loc_lng","work_lat","work_lng"))) %>%
  mutate(student_simple = case_when(grepl("Full-time student",student) | grepl("Part-time student",student)~"Student",
                                    is.na(student)~ "NA",
                                    TRUE~student),
         schooltype_simple = case_when(schooltype %in% non_school~"Preschool/Daycare/Cared for at home",
                                       schooltype %in% school_grade_type~"Grade school",
                                       schooltype %in% school_college~"University/College/Graduate school/Vocational school",
                                       TRUE~schooltype)) %>%
  mutate(age_simple = case_when(age %in% c("Under 5 years old")~"Under 5 years old",
                                age %in% c("5-11 years")~"5-11",
                                age %in% c("12-15 years")~"12-15",
                                age %in% c("16-17 years")~"16-17",
                              TRUE~"18+"),
         employment_simple = case_when(grepl("^(Self-)?[eE]mployed", as.character(employment))~"Employed",
                                       # is.na(employment)~ "NA",
                                       TRUE~"No")
         ) %>% 
  mutate(work_lat_lng = case_when(!is.na(work_lat)~"have_work_loc",
                                  TRUE~"no_work_loc"),
         school_loc_lat_lng = case_when(!is.na(school_loc_lat)~"have_school_loc",
                                        TRUE~"no_school_loc"))
```


```{r}
person_under5 <- person_school_work %>% filter(age_simple=="Under 5 years old")
person_5_11 <- person_school_work %>% filter(age_simple=="5-11")
person_12_15 <- person_school_work %>% filter(age_simple=="12-15")

person_5_15 <- person_school_work %>% filter(age_simple %in% c("5-11","12-15"))

person_16_17 <- person_school_work %>% filter(age_simple=="16-17")
person_18plus <- person_school_work %>% filter(age_simple=="18+")
```



    
## issue 1: (only in 2019) school location misplaced as work location

- 5-15 years old: all work locations should be school locations

```{r}
person_5_15 %>% filter(is.na(school_loc_lat) & !is.na(work_lat)) %>%
  group_by(survey_year,age) %>%
  summarise(person_count = n()) %>%
  arrange(desc(age))
```

- 16-17 years old: all work locations should be school locations
  - no `employment_simple == "Employed"` records

```{r}
person_16_17 %>% 
  filter(is.na(school_loc_lat) & !is.na(work_lat), employment_simple!="Employed") %>%
  group_by(survey_year,age,employment_simple, work_lat_lng) %>%
  summarise(person_count = n()) %>%
  pivot_wider(id_cols = c("age","survey_year","employment_simple"), names_from = "work_lat_lng", values_from = "person_count")
```

## issue 2: student status recode for children (divide by age groups)

:::{.panel-tabset}
### Under 5 years old: not student

- double check: do we want babies to be coded as students?
- current data
  - 2021: Under 5 years old are assigned as students
  - only 2023: some Under 5 years old have school locations
    
```{r}
person_under5 %>% 
  filter(survey_year==2023) %>%
  mutate(school_loc_lat_lng = case_when(!is.na(school_loc_lat)~"have_school_loc",
                                        TRUE~"no_school_loc")) %>%
  group_by(age_simple,survey_year,school_loc_lat_lng) %>%
  summarise(person_count = n()) %>%
  arrange(desc(school_loc_lat_lng))
```

### 5-11 years old: assign student status in grade schools

- additional step because 5 and 6 years old are not students
- not student if their school type is `"Daycare"`,`"Preschool"`,`"Cared for at home"` or missing
- student if school type is `"Elementary school (public, private, charter)"`,
                            `"K-12 public school"`,
                            `"K-12 private school"`,
                            `"K-12 home school (full-time or part-time)"`,
                            `"Middle school (public, private, charter)"`,
                            `"High school (public, private, charter)"`
- current data
    
```{r}
person_5_11 %>% 
  group_by(age,survey_year,schooltype_simple) %>%
  summarise(person_count = n()) %>%
  pivot_wider(id_cols = c("age","survey_year"), names_from = "schooltype_simple", values_from = "person_count")
```

### 12-17 years old: all children should be students
    - current data
```{r}
person_12_15 %>%
  add_row(person_16_17) %>%
  group_by(survey_year, student) %>%
  reframe(person_count = n()) %>%
  pivot_wider(id_cols = "survey_year", names_from = "student", values_from = "person_count")
```

### 18 and above: keep as it is

:::

### temporary recode results

```{r}
# fix 2019 school location
fix_school_location <- function(.data){
  .data %>%
  mutate(school_loc_lat_recode = case_when(employment_simple!="Employed" & is.na(school_loc_lat) & !is.na(work_lat)~work_lat,
                                    TRUE~school_loc_lat),
         school_loc_lng_recode = case_when(employment_simple!="Employed" & is.na(school_loc_lng) & !is.na(work_lng)~work_lng,
                                    TRUE~school_loc_lng))
}

# recode student by age group
student_under5 <- person_under5 %>%
  mutate(school_loc_lat_recode = school_loc_lat,
         school_loc_lng_recode = school_loc_lng,
         student_recode = "No, not a student")
student_5_11 <- person_5_11 %>%
  fix_school_location() %>%
  # mutate(student_recode = case_when(!is.na(school_loc_lat_recode)~"Student",
  #                                   TRUE~"No, not a student"))
  mutate(student_recode = case_when(schooltype_simple=="Grade school"~"Student",
                                    TRUE~"No, not a student"))
student_12_15 <- person_12_15 %>%
  fix_school_location() %>%
  mutate(student_recode = "Student")
student_16_17 <- person_16_17 %>%
  fix_school_location() %>%
  mutate(student_recode = "Student")
student_18plus <- person_18plus %>%
  mutate(school_loc_lat_recode = school_loc_lat,
         school_loc_lng_recode = school_loc_lng,
         student_recode = case_when(student_simple=="Student"~"Student",
                                    TRUE~"No, not a student"))
# combine all age groups
student_recoded <- student_under5 %>%
  add_row(student_5_11) %>%
  add_row(student_12_15) %>%
  add_row(student_16_17) %>%
  add_row(student_18plus)

# for counting errors
# count_error <- student_recoded %>%
#   mutate(error = case_when(student_simple=="Student" & student_recode=="Student"~0,
#                            student_simple=="Student" & student_recode!="Student"~1,
#                            student_recode=="Student"~1,
#                            TRUE~0)) %>%
#   filter(error==1)
```


- student count: original and recode
  - 2021 count dropped: Under 5 years old and some 5-11 years old were wrongly assigned as student
  - distribution across years:
  
```{r}
student_recoded %>%
  filter(student_simple=="Student") %>%
  group_by(survey_year) %>%
  summarise(student_count=n()) %>%
  left_join(student_recoded %>%
    filter(student_recode=="Student") %>%
    group_by(survey_year) %>%
    summarise(recode_count=n()),
    by="survey_year") %>%
  mutate(error_count = recode_count-student_count)


```

  
```{r, fig.width=8, fig.height=3}
df_plot <- student_recoded %>% 
  group_by(survey_year,student_simple) %>%
  reframe(count=n()) %>%
  group_by(survey_year) %>%
  mutate(percent=count/sum(count)) %>%
  ungroup() %>%
  rename(student = student_simple) %>%
  mutate(type = "original")
df_plot2 <- student_recoded %>% 
  group_by(survey_year,student_recode) %>%
  reframe(count=n()) %>%
  group_by(survey_year) %>%
  mutate(percent=count/sum(count)) %>%
  ungroup() %>%
  rename(student = student_recode) %>%
  mutate(type = "recode")

df_plot3 <- rbind(df_plot,df_plot2) %>%
  filter(student!="No, not a student") %>%
  mutate(student = factor(student, levels=c("Student","NA","Missing: Skip Logic","Missing Response")))
ggplot(df_plot3, aes(x=factor(survey_year),fill=type)) +
  geom_col(aes(y=count), position="dodge")+ 
  facet_wrap(~student, ncol=4) +
  labs(title="number of students") +
  theme_bw()
```

- student share: original and recode
```{r, fig.width=5, fig.height=3}
ggplot(df_plot3 %>% filter(student=="Student"), aes(x=factor(survey_year),fill=type)) +
  geom_col(aes(y=percent), position="dodge")+ 
  scale_y_continuous(labels = scales::percent_format())+
  labs(title="share of students") +
  theme_bw()
```

## issue 3: SOV for children (only children under 16 and `trip_weight>0`)

- number of children trips with Drive SOV mode

```{r, fig.height=4, fig.width=8}
children_student_recoded <- person_data %>%
  filter(age %in% c("Under 5 years old","5-11 years","12-15 years")) %>%
  left_join(student_recoded %>% select(person_id,student_recode), by="person_id")

trip_children <- trip_data %>%
  filter(trip_weight>0,
         person_id %in% children_student_recoded$person_id) %>%
  left_join(student_recoded %>% select(person_id,age,student_recode), by="person_id")

trip_children_SOV <- trip_children %>% filter(mode_class=="Drive SOV") %>%
  select(all_of(c("trip_id","household_id","person_id","survey_year",
"distance_miles","duration_minutes","speed_mph","speed_flag",
"origin_purpose_cat","dest_purpose_cat",
"mode_class","mode_class_5","mode_other_specify",
"travelers_total","travelers_hh","travelers_nonhh","driver","trip_weight",
"age","student_recode")))

trip_children_SOV %>%
  group_by(survey_year) %>%
  summarise(person_count = n())
  
```

- possible error source 1: `travelers_total`

```{r}
trip_children_SOV %>%
  group_by(survey_year,travelers_total) %>%
  summarise(person_count = n())
```

