---
title: "HTS delivery analysis: new weights"
format: 
  html:
    fig-width: 4
    fig-height: 3
    df-print: kable
  
toc: true
echo: FALSE
warning: FALSE
---


```{r}
library(tidyverse)
library(psrc.travelsurvey)
library(psrcelmer)
library(travelSurveyTools)
library(data.table)
library(psrcplot)
library(sf)
install_psrc_fonts()

# show moe or numbers
show_moe <- NULL
# show_moe <- "prop_moe"

source("survey-23-preprocess_JLin.R")
cb_path <- "J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx"
variable_list <- readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
delivery_vars <- c("deliver_food", "deliver_grocery", "deliver_package", "deliver_work", "deliver_other",
                   "deliver_none", "deliver_elsewhere", "deliver_office")

# get data
hh_day_delivery <- readRDS("hh_day_delivery.rds")%>%
  mutate(day_id = paste0(as.character(hh_id),as.character(daynum))) %>%
  mutate_at(vars(hh_id,day_id,survey_year), ~as.character(.)) %>%
  mutate_at(vars(deliver_elsewhere:deliver_home_any), ~ifelse(is.na(.),"No",.))
setDT(hh_day_delivery)

hh <- get_query(sql= "select  household_id as hh_id, hhsize, hhincome_broad,home_jurisdiction, home_county, hh_weight,vehicle_count, home_rgcname, home_lat, home_lng, survey_year, numadults, numchildren, numworkers, num_students, rent_own, hh_race_category
                from HHSurvey.v_households_labels") %>% 
  mutate_at(vars(hh_id,survey_year), ~as.character(.)) %>%
  mutate(
    hhincome_broad_pre2023 = factor(hhincome_broad,
                            levels=c("Under $25,000","$25,000-$49,999", "$50,000-$74,999", 
                                     "$75,000-$99,999","$100,000-$199,999", 
                                     "$100,000 or more",
                                     "$200,000 or more", "Prefer not to answer")),
    hhincome_broad = factor(hhincome_broad,
                            levels=c("Under $25,000","$25,000-$49,999", "$50,000-$74,999", 
                                     "$75,000-$99,999","$100,000-$199,999", 
                                     "$200,000 or more", "Prefer not to answer")),
    hhincome = factor(case_when(
      hhincome_broad_pre2023 %in% c("Under $25,000","$25,000-$49,999")~'Under $50,000',
      hhincome_broad_pre2023 %in% c("$50,000-$74,999", "$75,000-$99,999")~"$50,000-$99,999",
      hhincome_broad_pre2023 %in% c("$100,000 or more","$100,000-$199,999", "$200,000 or more")~"$100,000 or more",
      TRUE~hhincome_broad_pre2023),
      levels = c('Under $50,000',"$50,000-$99,999","$100,000 or more","Prefer not to answer"))
    ) %>%
  mutate(
    wkr_grp = case_when(
      numworkers %in% c("1 worker",  "0 workers")~numworkers,
      numworkers %in% c("2 workers", "3 workers", "4 workers", "5 workers", "6 workers")~"2+ workers",))

setDT(hh)

hts_hh_day <- list("day" = hh_day_delivery %>% left_join(hh, by=c("hh_id","survey_year")))
hts_hh_day_2023 <- list("day" = hh_day_delivery %>% 
                          filter(survey_year==2023) %>%
                          left_join(hh, by=c("hh_id","survey_year")))
```


```{r}
my_variable_list <- new_add_variable(
  variables_dt = variable_list,
  variable_names = c("deliver_home_any","hhincome",
                     #"hhincome2","hhincome3","hhincome4","hhincome5",
                     "hh_day_weight", "homegeog", 'rgc_cat_home',
                     'class_desc',"wkr_grp","hhincome_broad_pre2023"),
  table_name = "day") %>%
  mutate(is_checkbox = ifelse(variable %in% delivery_vars, 0, is_checkbox),
         shared_name = ifelse(variable %in% delivery_vars, variable, shared_name),
         day = ifelse(variable %in% c("hhincome_broad", "home_county","numchildren", "numworkers", "numadults", "hhsize_simple"),1,day),
         data_type = ifelse(variable == "hh_day_weight", "numeric", data_type))
```

## delivery trends

```{r}
delivery_summ <- function(delivery_varname){
  # get table
  test <- get_hts_summary(hts_hh_day,
                          summary_var = c("survey_year",delivery_varname),
                          variables_dt = my_variable_list,
                          id_var="day_id",
                          wt_var="hh_day_weight",
                          wt_name="hh_day_weight")

  df_delivery <- test %>% filter(.[[delivery_varname]] == "Yes")
  df_delivery<- df_delivery%>%mutate(Year=as.character(survey_year))

  df_delivery
}

chart_style <- list(
  theme(axis.text.x=element_text(size=16), 
        axis.text.y=element_text(size=14),
        legend.text = element_text(size=16), 
        axis.title.y=element_text(size=12), 
        axis.title.x=element_text(size=14))
  )

plot_column <- function(summary_data, fill, moe=NULL){
  static_column_chart(summary_data, 
                      x='Year', fill=fill, y='prop', 
                      dec=0, color='pgnobgy_10', moe=moe) +
  chart_style
}

```

- package
```{r}

df_deliver_pkg<-delivery_summ('deliver_package')

plot_column(df_deliver_pkg, fill='deliver_package', moe=show_moe)

# write.csv(df_deliver_pkg, 'deliver_pkg_trend.csv')
```

- other deliveries

:::{.panel-tabset}

### service work

```{r}

df_deliver_work<-delivery_summ('deliver_work')

plot_column(df_deliver_work, fill='deliver_work', moe=show_moe)

# write.csv(df_deliver_work, 'deliver_work_trend.csv')
```

### food delivery

```{r}

df_deliver_food<-delivery_summ('deliver_food')

plot_column(df_deliver_food, fill='deliver_food', moe=show_moe)
# write.csv(df_deliver_food, 'deliver_food_trend.csv')
```

### grocery delivery

```{r}

df_deliver_grocery<-delivery_summ('deliver_grocery')

plot_column(df_deliver_grocery, fill='deliver_grocery', moe=show_moe)

# write.csv(df_deliver_grocery, 'deliver_grocery_trend.csv')
```

### delivery outside home

```{r}

df_deliver_else<-delivery_summ('deliver_elsewhere')
df_deliver_else
```
```{r}

df_deliver_office<-delivery_summ('deliver_office')
df_deliver_office
```
:::


## 2023 delivery by demographic characteristics

```{r}
delivery_summ_demo <- function(delivery_varname,demoname){
  # get table
  test <- get_hts_summary(hts_hh_day_2023, 
                          summary_var = c(demoname,delivery_varname),
                          variables_dt = my_variable_list, 
                          id_var="day_id",
                          wt_var="hh_day_weight",
                          wt_name="hh_day_weight")
  
  df_delivery <- test %>% filter(.[[delivery_varname]] == "Yes")

  
  df_delivery
}

plot_bar <- function(summary_data, y_fill, moe=NULL){
  static_bar_chart(summary_data, y=y_fill, fill=y_fill, x='prop', 
                   dec=0, color='pgnobgy_10', moe=moe) +
  chart_style
}
```

### household income

- package

```{r}
#| layout-ncol: 2

summary<-delivery_summ_demo('deliver_package', 'hhincome_broad' )%>%
  filter(hhincome_broad!="Prefer not to answer")

plot_bar(summary, y_fill='hhincome_broad', moe=show_moe)
# write.csv(summary, 'deliver_package_income_trend.csv')


summary<-delivery_summ_demo('deliver_package', 'hhincome' )%>%
  filter(hhincome!="Prefer not to answer")

plot_bar(summary, y_fill='hhincome', moe=show_moe)
# write.csv(summary, 'deliver_package_income_ls.csv')
```

- other delivery

:::{.panel-tabset}

#### grocery

```{r}
#| layout-ncol: 2

summary<-delivery_summ_demo('deliver_grocery', 'hhincome_broad' )%>%
  filter(hhincome_broad!="Prefer not to answer")

plot_bar(summary, y_fill='hhincome_broad', moe=show_moe)
# write.csv(summary, 'deliver_grocery_income_trend.csv')

summary<-delivery_summ_demo('deliver_grocery', 'hhincome' )%>%
  filter(hhincome!="Prefer not to answer")

plot_bar(summary, y_fill='hhincome', moe=show_moe)
```

#### food

```{r}
#| layout-ncol: 2

summary<-delivery_summ_demo('deliver_food', 'hhincome_broad' )%>%
  filter(hhincome_broad!="Prefer not to answer")

plot_bar(summary, y_fill='hhincome_broad', moe=show_moe)


summary<-delivery_summ_demo('deliver_food', 'hhincome' ) %>%
  filter(hhincome!="Prefer not to answer")

plot_bar(summary, y_fill='hhincome', moe=show_moe)
```

:::

### home county

- package

```{r}

summary<-delivery_summ_demo('deliver_package', 'home_county' )%>%
  filter(!is.na(home_county))

plot_bar(summary, y_fill='home_county', moe=show_moe)

```

- other delivery

:::{.panel-tabset}

#### grocery

```{r}
summary<-delivery_summ_demo('deliver_grocery', 'home_county' )

plot_bar(summary, y_fill='home_county', moe=show_moe)
```

#### food

```{r}

summary<-delivery_summ_demo('deliver_food', 'home_county' )

plot_bar(summary, y_fill='home_county', moe=show_moe)
# write.csv(summary, 'deliver_food_income_trend.csv')
```

:::

### number of workers

- package

```{r}

summary<-delivery_summ_demo('deliver_package', 'wkr_grp' )

plot_bar(summary, y_fill='wkr_grp', moe=show_moe)
```

- other delivery

:::{.panel-tabset}

#### grocery

```{r}

summary<-delivery_summ_demo('deliver_grocery', 'wkr_grp' )

plot_bar(summary, y_fill='wkr_grp', moe=show_moe)
# write.csv(summary, 'deliver_grocery_wkr_grp.csv')
```

#### food

```{r}

summary<-delivery_summ_demo('deliver_food', 'wkr_grp' )

plot_bar(summary, y_fill='wkr_grp', moe=show_moe)
# write.csv(summary, 'deliver_food_wkr_grp.csv')
```

:::

## explore income trend

```{r}
delivery_summ_demo_year <- function(delivery_varname,demoname){
  # get table
  test <- get_hts_summary(hts_hh_day, 
                          summary_var = c("survey_year", demoname,delivery_varname),
                          variables_dt = my_variable_list, 
                          id_var="day_id",
                          wt_var="hh_day_weight",
                          wt_name="hh_day_weight")
  
  df_delivery <- test %>% filter(.[[delivery_varname]] == "Yes")

  
  df_delivery
}

plot_income_trend <- function(summary_data, fill, moe=NULL){
  static_column_chart(summary_data, x='survey_year', fill=fill, y='prop', 
                      dec=0, color='pgnobgy_10', moe=moe) +
    chart_style
}

```

- package

```{r}
summary<-delivery_summ_demo_year('deliver_package', 'hhincome' ) %>%
  filter(hhincome!="Prefer not to answer")
plot_income_trend(summary, fill='hhincome', moe=show_moe)

```

- other delivery

:::{.panel-tabset}

### grocery

```{r}
summary<-delivery_summ_demo_year('deliver_grocery', 'hhincome' ) %>%
  filter(hhincome!="Prefer not to answer")
plot_income_trend(summary, fill='hhincome', moe=show_moe)

```

### food

```{r}
summary<-delivery_summ_demo_year('deliver_food', 'hhincome' ) %>%
  filter(hhincome!="Prefer not to answer")
plot_income_trend(summary, fill='hhincome', moe=show_moe)

```

:::

## person trip rate

```{r}
# number of shopping trips
shopping_trip <- hts_data_hhday$trip %>%
  filter(dest_purpose_cat=="Shopping") %>%
  group_by(dest_purpose_cat,survey_year) %>%
  summarise(trip_est = sum(trip_weight))

person_count <- hts_data_hhday$person %>%
  group_by(survey_year) %>%
  summarise(person_est = sum(person_weight))

test <- shopping_trip %>%
  left_join(person_count, by="survey_year") %>%
  mutate(trip_rate = trip_est/person_est,
         survey_year = factor(survey_year))
```


```{r}
static_column_chart(test, 
                    x='survey_year', fill="dest_purpose_cat", y='trip_rate', 
                    dec=0, color='pgnobgy_10') +
  chart_style
```



