---
title: "Understanding the impact of data cleaning on model fit and accuracy"
author: "Parastoo Jabbari"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document: default
  word_document: default
---


```{r global-options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE) # formatting
```



This report compares models that are built using 2021 household travel survey data delivered by RSG on August 2 and the same data after cleaning by PSRC team.

```{r libraries, include=FALSE}
#Load functions script
#source("travel_survey_analysis_functions.R")
library(ggplot2)
library(spatstat)
library(pander)
library(reshape2)
library(table1)
library(knitr)
#library(kableExtra)
library(odbc)
library(DBI)
library(data.table)
library(tidyverse)
library(mlogit)
library(dfidx)
library(stargazer)
library(rmarkdown)
library(htmlTable)
#library(flextable)
```



```{r, message=FALSE}

#Connect to database
db.connect <- function() {
  elmer_connection <- dbConnect(odbc(),
                                driver = "SQL Server",
                                server = "AWS-PROD-SQL\\Sockeye",
                                database = "Elmer",
                                trusted_connection = "yes"
  )
}
con = db.connect()

# a function to read tables and queries from Elmer
read.dt <- function(astring, type =c('table_name', 'sqlquery')) {
  elmer_connection <- db.connect()
  if (type == 'table_name') {
    dtelm <- dbReadTable(elmer_connection, SQL(astring))
  } else {
    dtelm <- dbGetQuery(elmer_connection, SQL(astring))
  }
  dbDisconnect(elmer_connection)
  dtelm
}



sql.query = paste("SELECT [recid]
      ,[hhid]
      ,[personid]
      ,[pernum]
      ,[tripid]
      ,[tripnum]
      ,[traveldate]
      ,[daynum]
      ,[dayofweek]
      ,[data_source]
      ,[hhgroup]
      ,[depart_time_mam]
      ,[depart_time_timestamp]
      ,[arrival_time_mam]
      ,[arrival_time_timestamp]
      ,[origin_lat]
      ,[origin_lng]
      ,[dest_lat]
      ,[dest_lng]
      ,[trip_path_distance]
      ,[google_duration]
      ,[reported_duration]
      ,[travelers_total]
      ,[origin_purpose]
      ,[origin_purpose_cat]
      ,[dest_purpose]
      ,[dest_purpose_cat]
      ,[mode_1]
      ,[transit_systems]
      ,[transit_lines]
      ,[psrc_inserted]
      ,[driver]
      ,[mode_type]
      ,[origin_name]
      ,[dest_name] 
  FROM [hhts_cleaning].[HHSurvey].[Trip]")
trips = read.dt(sql.query, 'sqlquery')

sql.query2 = paste("SELECT [age], [personid], [gender], [employment] ,[race_afam]
      ,[student]
      ,[worker]
      ,[education]
      ,[license]
      ,[race_aiak]
      ,[race_asian]
      ,[race_hapi]
      ,[race_hisp]
      ,[race_white]
      ,[race_other]
      ,[race_noanswer]
      ,[industry]
       FROM [hhts_cleaning].[HHSurvey].[Person]")
persons = read.dt(sql.query2, 'sqlquery')

sql.query3 = paste("SELECT [hhincome_detailed]
      ,[hhid]
      ,[vehicle_count]
      ,[rent_own]
      ,[numchildren]
      ,[numworkers]
      ,[hhsize]
      ,[final_home_tract]
      ,[reported_lat]
      ,[reported_lng] FROM [hhts_cleaning].[HHSurvey].[Household]")
hhs = read.dt(sql.query3, 'sqlquery')



#Merge age, race, and income data to trips dataset
df<-merge(trips,persons, by = "personid", all.x = T, all.y = F)
df_cleaned<-merge(df,hhs, by = "hhid", all.x = T, all.y = F)

remove(trips,persons,hhs)
```





```{r, message=FALSE}

# a function to read tables and queries from Elmer
read.dt <- function(astring, type =c('table_name', 'sqlquery')) {
  elmer_connection <- db.connect()
  if (type == 'table_name') {
    dtelm <- dbReadTable(elmer_connection, SQL(astring))
  } else {
    dtelm <- dbGetQuery(elmer_connection, SQL(astring))
  }
  dbDisconnect(elmer_connection)
  dtelm
}



sql.query = paste("SELECT [hhid]
      ,[personid]
      ,[pernum]
      ,[tripid]
      ,[tripnum]
      ,[traveldate]
      ,[daynum]
      ,[dayofweek]
      ,[data_source]
      ,[hhgroup]
      ,[depart_time_mam]
      ,[depart_time_timestamp]
      ,[arrival_time_mam]
      ,[arrival_time_timestamp]
      ,[origin_lat]
      ,[origin_lng]
      ,[dest_lat]
      ,[dest_lng]
      ,[trip_path_distance]
      ,[google_duration]
      ,[reported_duration]
      ,[travelers_total]
      ,[origin_purpose]
      ,[origin_purpose_cat]
      ,[dest_purpose]
      ,[dest_purpose_cat]
      ,[mode_1]
      ,[driver]
      ,[mode_type]
      ,[origin_name]
      ,[dest_name] 
  FROM [hhts_cleaning].[dbo].[hts_trip]")
trips = read.dt(sql.query, 'sqlquery')

sql.query2 = paste("SELECT [age], [personid], [gender], [employment] ,[race_afam]
      ,[student]
      ,[worker]
      ,[education]
      ,[license]
      ,[race_aiak]
      ,[race_asian]
      ,[race_hapi]
      ,[race_hisp]
      ,[race_white]
      ,[race_other]
      ,[race_noanswer]
      ,[industry]
       FROM [hhts_cleaning].[dbo].[hts_person]")
persons = read.dt(sql.query2, 'sqlquery')

sql.query3 = paste("SELECT [hhincome_detailed]
      ,[hhid]
      ,[vehicle_count]
      ,[rent_own]
      ,[numchildren]
      ,[numworkers]
      ,[hhsize]
      ,[final_home_tract]
      ,[reported_lat]
      ,[reported_lng] FROM [hhts_cleaning].[dbo].[hts_hh]")
hhs = read.dt(sql.query3, 'sqlquery')



#Merge age, race, and income data to trips dataset
df<-merge(trips,persons, by = "personid", all.x = T, all.y = F)
df_raw<-merge(df,hhs, by = "hhid", all.x = T, all.y = F)

# Remove datasets that are no longer needed
remove(trips,persons,hhs)
```

```{r, message=FALSE}

## Changing labels
df_cleaned$mode_1<-ifelse(df_cleaned$mode_1 == 1,"Walk",(ifelse(df_cleaned$mode_1 == 2, "Bike",ifelse(df_cleaned$mode_1 %in%  c(3:16,33,34), "Car",ifelse(df_cleaned$mode_1 %in% c(21,22,26,32,36,37), "Other","Transit")))))


df_raw$mode_1<-ifelse(df_raw$mode_1 == 1,"Walk",(ifelse(df_raw$mode_1 == 2, "Bike",ifelse(df_raw$mode_1 %in%  c(3:16,33,34), "Car",ifelse(df_raw$mode_1  %in% c(21,22,26,32,36,37), "Other","Transit")))))

df_cleaned$age <- dplyr::recode(df_cleaned$age, '1' = "Under 5 years old", '2' = "5-11 years", '3' = "12-15 years", '4' = "16-17 years", '5' = "18-24 years", '6' = "25-34 years", '7' = "35-44 years", '8' = "45-54 years", '9' = "55-64 years", '10' = "65-74 years", '11' = "75-84 years", '12' = "85 years or older")

df_raw$age <- dplyr::recode_factor(df_raw$age, '1' = "Under 5 years old", '2' = "5-11 years", '3' = "12-15 years", '4' = "16-17 years", '5' = "18-24 years", '6' = "25-34 years", '7' = "35-44 years", '8' = "45-54 years", '9' = "55-64 years", '10' = "65-74 years", '11' = "75-84 years", '12' = "85 years or older")

#df_cleaned$age <- ordered(df_cleaned$age, levels = c("Under 5 years old", "5-11 years", "12-15 years", "16-17 years", "18-24 years", "25-34 years", "35-44 years", "45-54 years", "55-64 years",  "65-74 years", "75-84 years", "85 years or older"))

#df_raw$age <- ordered(df_raw$age, levels = c("Under 5 years old", "5-11 years", "12-15 years", "16-17 years", "18-24 years", "25-34 years", "35-44 years", "45-54 years", "55-64 years",  "65-74 years", "75-84 years", "85 years or older"))

df_cleaned$age_group <- dplyr::recode(df_cleaned$age, "16-17 years" = "Under 25 years old", "18-24 years" = "Under 25 years old", "25-34 years" = "25-44 years", "35-44 years" = "25-44 years", "45-54 years" = "45-64 years", "55-64 years" = "45-64 years", "65-74 years" = "65 years or older", "75-84 years" = "65 years or older", "85 or years older" = "65 years or older")

df_raw$age_group <- dplyr::recode_factor(df_raw$age, "16-17 years" = "Under 25 years old", "18-24 years" = "Under 25 years old", "25-34 years" = "25-44 years", "35-44 years" = "25-44 years", "45-54 years" = "45-64 years", "55-64 years" = "45-64 years", "65-74 years" = "65 years or older", "75-84 years" = "65 years or older", "85 or years older" = "65 years or older")

df_cleaned$age_group <- ordered(df_cleaned$age_group, levels = c("Under 25 years old", "25-44 years", "45-64 years", "65 years or older"))

df_raw$age_group <- ordered(df_raw$age_group, levels = c("Under 25 years old", "25-44 years", "45-64 years", "65 years or older"))

df_cleaned$gender <- dplyr::recode(df_cleaned$gender, "1" = "Male", "2" = "Female", "3" = "Another", "4" = "Not listed here", "5" = "Non-Binary")

df_raw$gender <- dplyr::recode(df_raw$gender, "1" = "Male", "2" = "Female", "3" = "Another", "4" = "Not listed here", "5" = "Non-Binary")


df_cleaned$hhincome_detailed <- dplyr::recode(df_cleaned$hhincome_detailed,	"1" = "Under $75,000",
	"2" =	"Under $75,000", "3" = "Under $75,000", "4" = "Under $75,000", "5" = "Under $75,000", "6" = "Over $75,000",
	"7" = "Over $75,000", "8" = "Over $75,000", "9" = "Over $75,000", "10" = "Over $75,000","11" = "Prefer not to answer")

df_raw$hhincome_detailed <- dplyr::recode(df_raw$hhincome_detailed,	"1" = "Under $75,000",
	"2" =	"Under $75,000", "3" = "Under $75,000", "4" = "Under $75,000", "5" = "Under $75,000", "6" = "Over $75,000",
	"7" = "Over $75,000", "8" = "Over $75,000", "9" = "Over $75,000", "10" = "Over $75,000","11" = "Prefer not to answer")

df_cleaned$hhincome_detailed <- ordered(df_cleaned$hhincome_detailed, levels = c("Under $75,000", "Over $75,000", "Prefer not to answer"))

df_raw$hhincome_detailed <- ordered(df_raw$hhincome_detailed, levels = c("Under $75,000", "Over $75,000", "Prefer not to answer"))

```

## 1. Commute Mode Choice Model

I extracted commute trips starting from home and ending at work/work related destination from cleaned and raw version of 2021 HHTS. For this analysis I only used one trip per person.

```{r, message=FALSE}

# Extracting commute trips for cleaned data
commute<-df_cleaned[(df_cleaned$dest_purpose == 10 | df_cleaned$dest_purpose == 11 | df_cleaned$dest_purpose == 14) & df_cleaned$origin_purpose == 1 & is.na(df_cleaned$dest_purpose_cat) == F & is.na(df_cleaned$origin_purpose_cat) == F,]

personids<-unique(commute$personid)

commute_sh <- commute[1,]
commute_sh[,] <- NA

## giving priority to "Work" trips (not "work-related")
for (i in personids){
  links <- commute[commute$personid == i,]
  
    if(any(links$dest_purpose_cat == 2)){
      n = min(links$tripnum[which(links$dest_purpose_cat == 2)])}
    else {
      n = min(links$tripnum)
    }
  commute_sh <- rbind(commute_sh,commute[which(commute$personid == i & commute$tripnum == n),])
}

commute_sh<-commute_sh[-1,]

commute_cleaned <- commute_sh[,!(names(commute_sh) %in% c("origin_name","dest_name","transit_systems","transit_lines","psrc_inserted","traveldate","depart_time_timestamp","arrival_time_timestamp"))]
#Save data to local folder
#write.csv(commute_cleaned,"home to work trips cleaned.csv",row.names = FALSE)
```


```{r, message=FALSE}
# Extracting non commute trips for cleaned data
ncommute<-df_cleaned[(df_cleaned$dest_purpose != 10 & df_cleaned$dest_purpose != 11 & df_cleaned$dest_purpose != 14) & is.na(df_cleaned$dest_purpose_cat) == F & is.na(df_cleaned$origin_purpose_cat) == F,]

ncommute_cleaned <- ncommute[,!(names(ncommute) %in% c("origin_name","dest_name","transit_systems","transit_lines","psrc_inserted","traveldate","depart_time_timestamp","arrival_time_timestamp"))]

#Save data to local folder
#write.csv(ncommute_cleaned,"Non-commute trips cleaned.csv",row.names = FALSE)
```

```{r, message=FALSE}
# Extracting commute trips for raw data

commute<-df_raw[(df_raw$dest_purpose == 10 | df_raw$dest_purpose == 11 | df_raw$dest_purpose == 14) & df_raw$origin_purpose == 1 & is.na(df_raw$dest_purpose_cat) == F & is.na(df_raw$origin_purpose_cat) == F,]


personids<-unique(commute$personid)

commute_sh <- commute[1,]
commute_sh[,] <- NA

for (i in personids){
  links <- commute[commute$personid == i,]
  
    if(any(links$dest_purpose_cat == 2)){
      n = min(links$tripnum[which(links$dest_purpose_cat == 2)])}
    else {
      n = min(links$tripnum)
    }
  commute_sh <- rbind(commute_sh,commute[which(commute$personid == i & commute$tripnum == n),])
}

commute_sh<-commute_sh[-1,]

commute_raw <- commute_sh[,!(names(commute_sh) %in% c("origin_name","dest_name","transit_systems","transit_lines","psrc_inserted","traveldate","depart_time_timestamp","arrival_time_timestamp"))]

#Save data to local folder
#write.csv(commute_raw,"home to work trips raw.csv",row.names = FALSE)
```


```{r, message=FALSE}

# Extracting non commute trips for raw data

ncommute<-df_raw[(df_raw$dest_purpose != 10 & df_raw$dest_purpose != 11 & df_raw$dest_purpose != 14) & is.na(df_raw$dest_purpose_cat) == F & is.na(df_raw$origin_purpose_cat) == F,]

ncommute_raw <- ncommute[,!(names(ncommute) %in% c("origin_name","dest_name","transit_systems","transit_lines","psrc_inserted","traveldate","depart_time_timestamp","arrival_time_timestamp"))]


#Save data to local folder
#write.csv(ncommute_raw,"Non-commute trips raw.csv",row.names = FALSE)
```


```{r, message=FALSE}

# Create dataframe for mlogit choice modeling 
choice_data_cleaned<-dfidx(commute_cleaned, shape = "wide", choice = c("mode_1"),idnames = c(NA, "alt"), idx = list(c("recid")))

choice_data_raw<-dfidx(commute_raw, shape = "wide", choice = c("mode_1"),idnames = c(NA, "alt"), idx = list(c("tripid")))

choice_data_cleaned$alt <- factor(choice_data_cleaned$alt, levels = c("Car", "Bike", "Walk", "Transit","Other"))

choice_data_raw$alt <- factor(choice_data_raw$alt, levels = c("Car", "Bike", "Walk", "Transit","Other"))



```

```{r, message=FALSE}
# Update destination purpose categories

ncommute_cleaned$dest_purpose_cats <- ifelse(ncommute_cleaned$dest_purpose == 1,	"Went home", ifelse(ncommute_cleaned$dest_purpose == 6, "Went to school/daycare", ifelse(ncommute_cleaned$dest_purpose == 9,	"Dropped off/picked up someone", ifelse(ncommute_cleaned$dest_purpose == 10, "Went to primary workplace", ifelse(ncommute_cleaned$dest_purpose == 11, "Went to work-related place", ifelse(ncommute_cleaned$dest_purpose == 14,	"Went to other work-related activity", ifelse(ncommute_cleaned$dest_purpose == 30,  "Went grocery shopping", ifelse(ncommute_cleaned$dest_purpose == 32, "Went to other shopping", ifelse(ncommute_cleaned$dest_purpose == 33,	"Conducted personal business", ifelse(ncommute_cleaned$dest_purpose == 34, "Went to medical appointment", ifelse(ncommute_cleaned$dest_purpose == 50, "Went to restaurant to eat/get take-out", ifelse(ncommute_cleaned$dest_purpose == 51,	"Went to exercise", ifelse(ncommute_cleaned$dest_purpose == 52, "Attended social event", ifelse(ncommute_cleaned$dest_purpose == 53, "Attended recreational event", ifelse(ncommute_cleaned$dest_purpose == 54, "Went to religious/community/volunteer activity", ifelse(ncommute_cleaned$dest_purpose == 55,	"Vacation/Traveling (rMove only)", ifelse(ncommute_cleaned$dest_purpose == 56, "Went to a family activity", ifelse(ncommute_cleaned$dest_purpose == 60, "Transferred to another mode of transportation", ifelse(ncommute_cleaned$dest_purpose == 61, "Other appointment/errands (rMove only)", ifelse(ncommute_cleaned$dest_purpose == 62, "Other social/leisure (rMove only)", ifelse(ncommute_cleaned$dest_purpose == 97,	"Other purpose", "NA")))))))))))))))))))))

ncommute_raw$dest_purpose_cats <- ifelse(ncommute_raw$dest_purpose == 1,	"Went home", ifelse(ncommute_raw$dest_purpose == 6, "Went to school/daycare", ifelse(ncommute_raw$dest_purpose == 9,	"Dropped off/picked up someone", ifelse(ncommute_raw$dest_purpose == 10, "Went to primary workplace", ifelse(ncommute_raw$dest_purpose == 11, "Went to work-related place", ifelse(ncommute_raw$dest_purpose == 14,	"Went to other work-related activity", ifelse(ncommute_raw$dest_purpose == 30,  "Went grocery shopping", ifelse(ncommute_raw$dest_purpose == 32, "Went to other shopping", ifelse(ncommute_raw$dest_purpose == 33,	"Conducted personal business", ifelse(ncommute_raw$dest_purpose == 34, "Went to medical appointment", ifelse(ncommute_raw$dest_purpose == 50, "Went to restaurant to eat/get take-out", ifelse(ncommute_raw$dest_purpose == 51,	"Went to exercise", ifelse(ncommute_raw$dest_purpose == 52, "Attended social event", ifelse(ncommute_raw$dest_purpose == 53, "Attended recreational event", ifelse(ncommute_raw$dest_purpose == 54, "Went to religious/community/volunteer activity", ifelse(ncommute_raw$dest_purpose == 55,	"Vacation/Traveling (rMove only)", ifelse(ncommute_raw$dest_purpose == 56, "Went to a family activity", ifelse(ncommute_raw$dest_purpose == 60, "Transferred to another mode of transportation", ifelse(ncommute_raw$dest_purpose == 61, "Other appointment/errands (rMove only)", ifelse(ncommute_raw$dest_purpose == 62, "Other social/leisure (rMove only)", ifelse(ncommute_raw$dest_purpose == 97,	"Other purpose", "NA")))))))))))))))))))))


ncommute_cleaned$dest_purpose_cat1 <- ifelse(ncommute_cleaned$dest_purpose_cat == 1 , "Home", ifelse(ncommute_cleaned$dest_purpose_cat == 2, "Work", ifelse(ncommute_cleaned$dest_purpose_cat == 3, "Work-related",  ifelse(ncommute_cleaned$dest_purpose_cat == 4, "School", ifelse(ncommute_cleaned$dest_purpose_cat == 5, "Escort",  ifelse(ncommute_cleaned$dest_purpose_cat == 6, "Shop", ifelse(ncommute_cleaned$dest_purpose_cat == 7, "Meal",  ifelse(ncommute_cleaned$dest_purpose_cat == 8, "Social/Recreation", ifelse(ncommute_cleaned$dest_purpose_cat == 9, "Errand/Other",  ifelse(ncommute_cleaned$dest_purpose_cat == 10, "Change mode","NA"))))))))))   

ncommute_raw$dest_purpose_cat1 <- ifelse(ncommute_raw$dest_purpose_cat == 1 , "Home", ifelse(ncommute_raw$dest_purpose_cat == 2, "Work", ifelse(ncommute_raw$dest_purpose_cat == 3, "Work-related", ifelse(ncommute_raw$dest_purpose_cat == 4, "School", ifelse(ncommute_raw$dest_purpose_cat == 5, "Escort",  ifelse(ncommute_raw$dest_purpose_cat == 6, "Shop", ifelse(ncommute_raw$dest_purpose_cat == 7, "Meal", ifelse(ncommute_raw$dest_purpose_cat == 8, "Social/Recreation", ifelse(ncommute_raw$dest_purpose_cat == 9, "Errand/Other",  ifelse(ncommute_raw$dest_purpose_cat == 10, "Change mode","NA"))))))))))       


ncommute_cleaned$dest_purpose_cat1 <- factor(ncommute_cleaned$dest_purpose_cat1, 
                                                     levels = c("Home", "School", "Escort", "Shop", "Meal",
                                                                "Social/Recreation", "Errand/Other","Change mode"))                                                                 

ncommute_raw$dest_purpose_cat1 <- factor(ncommute_raw$dest_purpose_cat1, 
                                                     levels = c("Home", "School", "Escort", "Shop", "Meal",
                                                                "Social/Recreation", "Errand/Other","Change mode"))     


choice_data_cleaned_ncommute<-dfidx(ncommute_cleaned, shape = "wide", choice = c("mode_1"),idnames = c(NA, "alt"), idx = list(c("recid")))

choice_data_raw_ncommute<-dfidx(ncommute_raw, shape = "wide", choice = c("mode_1"),idnames = c(NA, "alt"), idx = list(c("tripid")))



choice_data_cleaned_ncommute$alt <- factor(choice_data_cleaned_ncommute$alt, levels = c("Car", "Bike", "Walk", "Transit","Other"))

choice_data_raw_ncommute$alt <- factor(choice_data_raw_ncommute$alt, levels = c("Car", "Bike", "Walk", "Transit","Other"))


```



```{r}
mod_cleaned <- commute_cleaned %>% count(mode_1, sort = TRUE)

mod_raw <- commute_raw %>% count(mode_1, sort = TRUE)

mod_sum <- merge(mod_cleaned,mod_raw, by = c("mode_1"),all = T)

mod_sum$n.x<- ifelse(is.na(mod_sum$n.x),0,mod_sum$n.x)

names(mod_sum) <- c("Mode", paste0("Cleand data (N = ",sum(mod_sum$n.x),")"),paste0("Raw data (N = ",sum(mod_sum$n.y),")"))

#mod_sum<-rbind(mod_sum,c("Total", sum(mod_sum$`Cleand data`),sum(mod_sum$`Raw data`)))

htmlTable(mod_sum, caption = 'Number of Observations per Mode', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')

```


### 1.1 Vehicle count

The number of vehicles in the household is a significant variable for all the modes relative car. It is more likely for people to choose car as a commute mode if there are more cars available to them. This variable is significant in both models. _R^2^_ is slightly higher in the model with cleaned data.
```{r mylatextable_vehiclecount, results = "asis"}


commute_cleaned$vehicle_count_cat <- as.character(ifelse(commute_cleaned$vehicle_count >= 3, "3 or more", commute_cleaned$vehicle_count))
commute_raw$vehicle_count_cat <- as.character(ifelse(commute_raw$vehicle_count >= 3, "3 or more", commute_raw$vehicle_count))
   
  
mod_veh_cleaned <- commute_cleaned %>% count(mode_1, vehicle_count_cat, sort = TRUE)

mod_veh_raw <- commute_raw %>% count(mode_1, vehicle_count_cat, sort = TRUE)

mod_veh_sum <- merge(mod_veh_cleaned,mod_veh_raw, by = c("mode_1", "vehicle_count_cat"),all = T)

mod_veh_sum$n.x<- ifelse(is.na(mod_veh_sum$n.x),0,mod_veh_sum$n.x)

names(mod_veh_sum) <- c("Mode","Vehicle Count", "Cleand data","Raw data")

htmlTable(mod_veh_sum, caption = 'Number of Observations per Mode and Vehicle Count', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')

m_cleaned <- mlogit(mode_1 ~  0| vehicle_count  | 0, choice_data_cleaned)

m_raw <- mlogit(mode_1 ~  0|vehicle_count | 0, choice_data_raw)

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other"), digits = 2)

```
### 1.2 Trip distance
Trip distance is almost statistically significant for all modes except for other. As distance increases, travelers are more likely to choose car over walk, bike, and transit.  The two models are pretty close and _R^2^_ is about the same in both of the models.

```{r mylatextable_dist, results = "asis"}
dist_cleaned <- commute_cleaned %>% group_by(mode_1) %>%  summarise(
  N = n(), Mean = round(mean(trip_path_distance, na.rm = T),2), Median = round(median(trip_path_distance, na.rm = T),2), qs_0.25 = round(quantile(trip_path_distance, c(0.25), na.rm = T),2)[1], qs_0.75 = round(quantile(trip_path_distance, c(0.75), na.rm = T),2)[1],
    Max = max(trip_path_distance, na.rm = T),
    Min = min(trip_path_distance, na.rm = T)
  ) %>% mutate(
  Dataset = "Cleaned")

dist_raw <- commute_raw  %>% group_by(mode_1) %>% summarise(
  N = n(), Mean = round(mean(trip_path_distance, na.rm = T),2), Median = round(median(trip_path_distance, na.rm = T),2), qs_0.25 = round(quantile(trip_path_distance, c(0.25), na.rm = T),2)[1], qs_0.75 = round(quantile(trip_path_distance, c(0.75), na.rm = T),2)[1],
    Max = max(trip_path_distance, na.rm = T),
    Min = min(trip_path_distance, na.rm = T)
  ) %>% mutate(
  Dataset = "Raw")

dist_sum <- cbind(rbind(dist_cleaned,dist_raw))
dist_sum <- dist_sum[order(dist_sum [,"mode_1"] ),c(9,1:8)]


htmlTable(dist_sum, cgroup = 'Statistic', n.cgroup = 9, caption = 'Summary of Commute Distances', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')

#stargazer(dist_sum, header=FALSE, type = 'html', summary=FALSE, rownames = FALSE, digits = 2,  column.sep.width = "3pt", font.size = "small")

m_cleaned <- mlogit(mode_1 ~  0| trip_path_distance  | 0, choice_data_cleaned)

m_raw <- mlogit(mode_1 ~  0|trip_path_distance  | 0, choice_data_raw)

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other"), digits = 2)


```
### 1.3 Age

The reference age group is 25-34 years, and compared to this group, individuals in age categories between 35 and 64 are less likely to choose walking over car. There are several age categories in this model that have few observations, and as a result, the standard error is too high for them.

```{r mylatextable_ageonly, results = "asis"}
age1<-commute_cleaned %>% count(age) %>% group_by(age)
age2<-commute_raw %>% count(age) %>% group_by(age)
age_n <- merge(age1, age2, by="age", all=TRUE)

age_n$age <- ordered(age_n$age, levels = c("Under 5 years old", "5-11 years", "12-15 years", "16-17 years", "18-24 years", "25-34 years", "35-44 years", "45-54 years", "55-64 years",  "65-74 years", "75-84 years", "85 years or older"))
age_n<-age_n[order(age_n$age),]

age_n$n.x <- ifelse(is.na(age_n$n.x), 0, age_n$n.x )
names(age_n)<- c("Age", "n (cleaned data)", "n (raw data)")

htmlTable(age_n, caption = 'Number of Observations per Age group', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')


age1<-commute_cleaned %>% count(mode_1, age)
age2<-commute_raw %>% count(mode_1, age)
age_n <- merge(age1, age2, by=c("mode_1","age"), all=TRUE)

age_n$age <- ordered(age_n$age, levels = c("Under 5 years old", "5-11 years", "12-15 years", "16-17 years", "18-24 years", "25-34 years", "35-44 years", "45-54 years", "55-64 years",  "65-74 years", "75-84 years", "85 years or older"))
age_n<-age_n[order(age_n$age),]

age_n$n.x <- ifelse(is.na(age_n$n.x), 0, age_n$n.x )

names(age_n)<- c("Mode","Age", "n (cleaned data)", "n (raw data)")


htmlTable(age_n, caption = 'Number of Observations per Mode and Age', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')

choice_data_cleaned$age <- factor(choice_data_cleaned$age, levels = c("25-34 years", "16-17 years", "18-24 years",  "35-44 years", "45-54 years", "55-64 years",  "65-74 years", "75-84 years"))

choice_data_raw$age <- factor(choice_data_raw$age, levels = c("25-34 years", "12-15 years", "16-17 years", "18-24 years",  "35-44 years", "45-54 years", "55-64 years",  "65-74 years", "75-84 years"))

m_cleaned <- mlogit(mode_1 ~  0| age | 0, choice_data_cleaned)


m_raw <- mlogit(mode_1 ~  0| age | 0, choice_data_raw)

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other"), digits = 2)

```

### 1.4 Race

Individuals who identify as white are less likely to use transit relative to other modes compared to non-white individuals. Individuals who identified their race as Other (people who didn’t choose any of the available race options) are more likely to choose bike over car.

```{r mylatextable_raceonly, results = "asis"}


race_cleaned<-rbind(c("African American or Black",length(commute_cleaned$race_afam[which(commute_cleaned$race_afam == 1)])),
            c("American Indian or Alaska Native",length(commute_cleaned$race_aiak[which(commute_cleaned$race_aiak == 1)])),
            c("Asian",length(commute_cleaned$race_asian[which(commute_cleaned$race_asian == 1)])),
            c("Native Hawaiian or Pacific Islander",length(commute_cleaned$race_hapi[which(commute_cleaned$race_hapi == 1)])),
            c("Hispanic",length(commute_cleaned$race_hisp[which(commute_cleaned$race_hisp == 1)])),
            c("White",length(commute_cleaned$race_white[which(commute_cleaned$race_white == 1)])),
            c("Other",length(commute_cleaned$race_other[which(commute_cleaned$race_other == 1)])),
            c("No answer",length(commute_cleaned$race_noanswer[which(commute_cleaned$race_noanswer == 1)])))

colnames(race_cleaned) <- c("Race","n")


race_raw<-rbind(c("African American or Black",length(commute_raw$race_afam[which(commute_raw$race_afam == 1)])),
            c("American Indian or Alaska Native",length(commute_raw$race_aiak[which(commute_raw$race_aiak == 1)])),
            c("Asian",length(commute_raw$race_asian[which(commute_raw$race_asian == 1)])),
            c("Native Hawaiian or Pacific Islander",length(commute_raw$race_hapi[which(commute_raw$race_hapi == 1)])),
            c("Hispanic",length(commute_raw$race_hisp[which(commute_raw$race_hisp == 1)])),
            c("White",length(commute_raw$race_white[which(commute_raw$race_white == 1)])),
            c("Other",length(commute_raw$race_other[which(commute_raw$race_other == 1)])),
            c("No answer",length(commute_raw$race_noanswer[which(commute_raw$race_noanswer == 1)])))

colnames(race_raw) <- c("Race", "n")

race_n <- merge(race_cleaned, race_raw, by="Race", all=TRUE)

names(race_n)<- c("Race", "n (cleaned data)", "n (raw data)")

htmlTable(race_n, caption = 'Number of Observations per Race group', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')


race_cleaned <- commute_cleaned %>% count(mode_1, race_afam , race_aiak , race_asian , race_hapi , race_hisp , race_white , race_other,race_noanswer  )

race_cleaned$race<-paste0(ifelse(race_cleaned$race_afam == 1,"African American or Black,",''), ifelse(race_cleaned$race_aiak == 1,"American Indian or Alaska Native,",''),
                         ifelse(race_cleaned$race_asian == 1,"Asian,",''), ifelse(race_cleaned$race_hapi == 1,"Native Hawaiian or Pacific Islander,",''), 
                         ifelse(race_cleaned$race_hisp == 1,"Hispanic,",''), ifelse(race_cleaned$race_white == 1,"White,",''),ifelse(race_cleaned$race_other == 1,"Other,",''),
                         ifelse(race_cleaned$race_noanswer == 1 | (race_cleaned$race_afam !=1 & race_cleaned$race_aiak !=1 & race_cleaned$race_asian !=1 & race_cleaned$race_hapi !=1
                                                                   & race_cleaned$race_hisp !=1 & race_cleaned$race_white != 1 & race_cleaned$race_other != 1),"No answer,",''))

race_cleaned <- race_cleaned[,c(1,11,10)]

race_cleaned$n[race_cleaned$mode_1 == race_cleaned$mode_1[which(duplicated(race_cleaned[,c(1,2)]))] & race_cleaned$race == race_cleaned$race[which(duplicated(race_cleaned[,c(1,2)]))]]<-sum(race_cleaned$n[race_cleaned$mode_1 == race_cleaned$mode_1[which(duplicated(race_cleaned[,c(1,2)]))] & race_cleaned$race == race_cleaned$race[which(duplicated(race_cleaned[,c(1,2)]))]]  )

race_cleaned <-unique(race_cleaned)

race_raw <- commute_raw %>% count(mode_1, race_afam , race_aiak , race_asian , race_hapi , race_hisp , race_white , race_other,race_noanswer  )

race_raw$race<-paste0(ifelse(race_raw$race_afam == 1,"African American or Black,",''), ifelse(race_raw$race_aiak == 1,"American Indian or Alaska Native,",''),
                         ifelse(race_raw$race_asian == 1,"Asian,",''), ifelse(race_raw$race_hapi == 1,"Native Hawaiian or Pacific Islander,",''), 
                         ifelse(race_raw$race_hisp == 1,"Hispanic,",''), ifelse(race_raw$race_white == 1,"White,",''),ifelse(race_raw$race_other == 1,"Other,",''),
                         ifelse(race_raw$race_noanswer == 1 | (race_raw$race_afam !=1 & race_raw$race_aiak !=1 & race_raw$race_asian !=1 & race_raw$race_hapi !=1
                                                                   & race_raw$race_hisp !=1 & race_raw$race_white != 1 & race_raw$race_other != 1),"No answer,",''))

race_raw <- race_raw[,c(1,11,10)]

race_raw$n[race_raw$mode_1 == "Car" & race_raw$race == race_raw$race[which(duplicated(race_raw[,c(1,2)]))]]<-sum(race_raw$n[race_raw$mode_1 == "Car" & race_raw$race == race_raw$race[which(duplicated(race_raw[,c(1,2)]))]]  )


race_raw$n[race_raw$mode_1 == "Transit" & race_raw$race == race_raw$race[which(duplicated(race_raw[,c(1,2)]))]]<-sum(race_raw$n[race_raw$mode_1 == "Transit" & race_raw$race == race_raw$race[which(duplicated(race_raw[,c(1,2)]))]]  )

race_raw <-unique(race_raw)

race_n <- merge(race_cleaned, race_raw, by=c("mode_1","race"))

race_n$n.x <- ifelse(is.na(race_n$n.x), 0, race_n$n.x )
race_n$n.y <- ifelse(is.na(race_n$n.y), 0, race_n$n.y )

names(race_n)<- c("Mode","Race", "n (cleaned data)", "n (raw data)")

htmlTable(race_n, caption = 'Number of Observations per Mode and Race group', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')


m_cleaned <- mlogit(mode_1 ~  0|  race_afam + race_aiak + race_asian + race_hapi + race_hisp + race_white + race_other | 0, choice_data_cleaned)

m_raw <- mlogit(mode_1 ~  0| race_afam + race_aiak + race_asian + race_hapi + race_hisp + race_white + race_other | 0, choice_data_raw)

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other","African American or Black:Bike", "African American or Black:Walk", "African American or Black:Transit", "African American or Black:Other", "American Indian or Alaska Native:Bike", "American Indian or Alaska Native:Walk", "American Indian or Alaska Native:Transit", "American Indian or Alaska Native:Other", "Asian:Bike", "Asian:Walk", "Asian:Transit", "Asian:Other", "Native Hawaiian or Pacific Islander:Bike", "Native Hawaiian or Pacific Islander:Walk", "Native Hawaiian or Pacific Islander:Transit", "Native Hawaiian or Pacific Islander:Other", "Hispanic:Bike",  "Hispanic:Walk",  "Hispanic:Transit",  "Hispanic:Other", "White:Bike", "White:Walk", "White:Transit", "White:Other", "Other:Bike", "Other:Walk",  "Other:Transit",  "Other:Other"), digits = 2)

```


### 1.5 Gender

We found that individuals who identify as male are more likely to use bike than other genders in the cleaned data model (significant at the 10 percent level).

```{r mylatextable_genderonly, results = "asis"}
gen1<-commute_cleaned %>% count(gender) %>% group_by(gender)
gen2<-commute_raw %>% count(gender) %>% group_by(gender)
gen_n <- merge(gen1, gen2, by="gender", all=TRUE)

names(gen_n)<- c("Gender", "n (cleaned data)", "n (raw data)")
gen_n<-gen_n[order(gen_n$Gender),]

htmlTable(gen_n, caption = 'Number of Observations per Gender group', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')

gen1<-commute_cleaned %>% count(mode_1, gender)
gen2<-commute_raw %>% count(mode_1, gender)
gen_n <- merge(gen1, gen2, by=c("mode_1","gender"), all=TRUE)

gen_n$n.x <- ifelse(is.na(gen_n$n.x), 0, gen_n$n.x )

names(gen_n)<- c("Mode","Gender", "n (cleaned data)", "n (raw data)")

htmlTable(gen_n, caption = 'Number of Observations per Gender group', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')

choice_data_cleaned$male<-ifelse( choice_data_cleaned$gender == "Male", "Male","Non-Male")
choice_data_raw$male<-ifelse( choice_data_raw$gender == "Male", "Male","Non-Male")

choice_data_cleaned$male <- ordered(choice_data_cleaned$male, levels = c("Non-Male", "Male"))
choice_data_raw$male <- ordered(choice_data_raw$male, levels = c("Non-Male", "Male"))


m_cleaned <- mlogit(mode_1 ~  0| male | 0, choice_data_cleaned)

m_raw <- mlogit(mode_1 ~  0| male | 0, choice_data_raw)

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other", "Male:Bike", "Male:Walk", "Male:Transit", "Male:Other"), digits = 2)

```

### 1.6 Income

Individuals whose household income is below \$75,000 are more likely to use transit for work trips than other modes. 

```{r mylatextable_incomeonly, results = "asis"}

inc1<-commute_cleaned %>% count(hhincome_detailed) %>% group_by(hhincome_detailed)
inc2<-commute_raw %>% count(hhincome_detailed) %>% group_by(hhincome_detailed)
inc_n <- merge(inc1, inc2, by="hhincome_detailed", all=TRUE)

names(inc_n)<- c("Income", "n (cleaned data)", "n (raw data)")
inc_n<-inc_n[order(inc_n$Income),]


#stargazer(inc_n,  type = 'html', summary=FALSE, rownames = FALSE)

htmlTable(inc_n, caption = 'Income', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 13px 0px;')

m_cleaned <- mlogit(mode_1 ~  0| hhincome_detailed | 0, choice_data_cleaned)


m_raw <- mlogit(mode_1 ~  0| hhincome_detailed | 0, choice_data_raw)

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other","Prefer not to answer:Bike","Prefer not to answer:Walk","Prefer not to answer:Transit","Prefer not to answer:Other","Under $75,000:Bike","Under $75,000:Walk","Under $75,000:Transit","Under $75,000:Other"), digits = 2)

```

### 1.7 Significant variables only

Including some of the significant variables from the previous models:

```{r, results = "asis"}

m_cleaned <- mlogit(mode_1 ~  0|male + trip_path_distance + vehicle_count + race_white | 0, choice_data_cleaned)

m_raw <- mlogit(mode_1 ~  0| male + trip_path_distance + vehicle_count + race_white | 0, choice_data_raw)

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other", "Male:Bike", "Male:Walk", "Male:Transit", "Male:Other", "Trip distance:Bike", "Trip distance:Walk", "Trip distance:Transit", "Trip distance:Other", "Vehicle count:Bike", "Vehicle count:Walk", "Vehicle count:Transit", "Vehicle count:Other", "White:Bike", "White:Walk", "White:Transit", "White:Other"))

freq_cleaned<-as.data.frame(m_cleaned$freq/dim(commute_cleaned)[1])
freq_cleaned$Freq<-round(freq_cleaned$Freq,4)
predicted_freq_cleaned <- as.data.frame(round(apply(fitted(m_cleaned, outcome=FALSE), 2, mean),4))
predicted_freq_cleaned <- cbind(rownames(predicted_freq_cleaned),predicted_freq_cleaned)
colnames(predicted_freq_cleaned)<-colnames(freq_cleaned)
df_cleaned<-merge(freq_cleaned,predicted_freq_cleaned, by = "choice")
colnames(df_cleaned)<-c("Choice","Choice Frequency","Fitted Probability")
df_cleaned$change<-round((df_cleaned$`Fitted Probability`-df_cleaned$`Choice Frequency`)*100/df_cleaned$`Choice Frequency`,3)
colnames(df_cleaned)[4]<-"Change (%)"
htmlTable(df_cleaned, caption = 'Cleaned data model frequencies vs. probabilities', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding-left: 3.5em; padding-right: 3.5em;')


freq_raw<-as.data.frame(m_raw$freq/dim(commute_raw)[1])
freq_raw$Freq<-round(freq_raw$Freq,4)
predicted_freq_raw <- as.data.frame(round(apply(fitted(m_raw, outcome=FALSE), 2, mean),4))
predicted_freq_raw <- cbind(rownames(predicted_freq_raw),predicted_freq_raw)
colnames(predicted_freq_raw)<-colnames(freq_raw)
df_raw<-merge(freq_raw,predicted_freq_raw, by = "choice")
colnames(df_raw)<-c("Choice","Choice Frequency","Fitted Probability")
df_raw$change<-round((df_raw$`Fitted Probability`-df_raw$`Choice Frequency`)*100/df_raw$`Choice Frequency`,3)
colnames(df_raw)[4]<-"Change (%)"
htmlTable(df_raw, caption = 'Raw data model frequencies vs. probabilities', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding-left: 3.5em; padding-right: 3.5em;')


odd_ratios <- cbind(round(exp(coefficients(m_cleaned)),2),round(exp(coefficients(m_raw)),2))


colnames(odd_ratios) <- c("Cleand data","Raw data")
rownames(odd_ratios)[5:16] <- c("Male:Bike","Male:Walk","Male:Transit","Male:Other", "Trip distance:Bike", "Trip distance:Walk", "Trip distance:Transit", "Trip distance:Other", "Vehicle count:Bike", "Vehicle count:Walk", "Vehicle count:Transit", "Vehicle count:Other")
rownames(odd_ratios)[17:20] <- c("White:Bike","White:Walk","White:Transit","White:Other")

htmlTable(odd_ratios, caption = 'Odd Ratios', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')
```

## 2. Non-commute Mode Choice Model


```{r}
mod_cleaned <- ncommute_cleaned %>% count(mode_1, sort = TRUE)

mod_raw <- ncommute_raw %>% count(mode_1, sort = TRUE)

mod_sum <- merge(mod_cleaned,mod_raw, by = c("mode_1"),all = T)

mod_sum$n.x<- ifelse(is.na(mod_sum$n.x),0,mod_sum$n.x)

names(mod_sum) <- c("Mode", paste0("Cleand data (N = ",sum(mod_sum$n.x),")"),paste0("Raw data (N = ",sum(mod_sum$n.y),")"))

#mode_sum<-rbind(mod_sum,c("Total", sum(mod_sum$`Cleand data`),sum(mod_sum$`Raw data`)))

htmlTable(mod_sum, caption = 'Number of Observations per Mode', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')
          
```

### 2.1 Vehicle count

Number of vehicles in the household is a significant variable for choosing bike, walk, transit, and other mode relative to car. Higher number of cars in the household increases the likelihood of choosing choose car over all other modes as a non-commute mode. Both models are consistent.

```{r mylatextable_vehiclecount_cn, results = "asis"}

ncommute_cleaned$vehicle_count_cat <- as.character(ifelse(ncommute_cleaned$vehicle_count >= 3, "3 or more", ncommute_cleaned$vehicle_count))
ncommute_raw$vehicle_count_cat <- as.character(ifelse(ncommute_raw$vehicle_count >= 3, "3 or more", ncommute_raw$vehicle_count))
   
  
mod_veh_cleaned <- ncommute_cleaned %>% count(mode_1, vehicle_count_cat, sort = TRUE)

mod_veh_raw <- ncommute_raw %>% count(mode_1, vehicle_count_cat, sort = TRUE)

mod_veh_sum <- merge(mod_veh_cleaned,mod_veh_raw, by = c("mode_1", "vehicle_count_cat"),all = T)

mod_veh_sum$n.x<- ifelse(is.na(mod_veh_sum$n.x),0,mod_veh_sum$n.x)

names(mod_veh_sum) <- c("Mode","Vehicle Count", "Cleand data","Raw data")

htmlTable(mod_veh_sum, caption = 'Trip Mode vs. Household Vehicle Count', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')

m_cleaned <- mlogit(mode_1 ~  0| vehicle_count  | 0, choice_data_cleaned_ncommute)

m_raw <- mlogit(mode_1 ~  0|vehicle_count | 0, choice_data_raw_ncommute)

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other"), digits = 2)

```


### 2.2 Trip distance
In the cleaned model, trip distance is statistically significant for all modes relative to car except other. Thus, as distance increases, travelers are more likely to choose car over biking and walking. However, the coefficient for transit is small but positive, which means that as travel distance increases, people prefer transit slightly more than car. For the uncleaned model, travel distance coefficient for bike is not statistically significant but it is significant for other. _R^2^_ is higher in the model with cleaned data.

```{r mylatextable_dist_cn, results = "asis"}
dist_cleaned <- ncommute_cleaned %>% group_by(mode_1) %>%  summarise(
  N = n(), Mean = round(mean(trip_path_distance, na.rm = T),2), Median = round(median(trip_path_distance, na.rm = T),2), qs_0.25 = round(quantile(trip_path_distance, 0.25, na.rm = T),2)[1], qs_0.75 = round(quantile(trip_path_distance, c(0.75), na.rm = T),2)[1],
    Max = max(trip_path_distance, na.rm = T),
    Min = min(trip_path_distance, na.rm = T)
  ) %>% mutate(
  Dataset = "Cleaned")

dist_raw <- ncommute_raw  %>% group_by(mode_1) %>%  summarise(
  N = n(), Mean = round(mean(trip_path_distance, na.rm = T),2), Median = round(median(trip_path_distance, na.rm = T),2), qs_0.25 = round(quantile(trip_path_distance, 0.25, na.rm = T),2)[1], qs_0.75 = round(quantile(trip_path_distance, c(0.75), na.rm = T),2)[1],
    Max = max(trip_path_distance, na.rm = T),
    Min = min(trip_path_distance, na.rm = T)
  ) %>% mutate(
  Dataset = "Raw")


dist_sum <- cbind(rbind(dist_cleaned,dist_raw))
dist_sum <- dist_sum[order(dist_sum [,"mode_1"] ),c(9,1:8)]


htmlTable(dist_sum, cgroup = 'Statistic', n.cgroup = 9, caption = 'Summary of Non-commute Distances', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')

#stargazer(dist_sum, header=FALSE, type = 'html', summary=FALSE, rownames = FALSE, digits = 2,  column.sep.width = "3pt", font.size = "small")

m_cleaned <- mlogit(mode_1 ~  0| trip_path_distance  | 0, choice_data_cleaned_ncommute)

m_raw <- mlogit(mode_1 ~  0|trip_path_distance  | 0, choice_data_raw_ncommute)

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other"), digits = 2)


```
### 2.3 Age

The reference age group is 25-34 years. Walking is significant for all the age groups and they all tend to choose car over walk compared to the reference age group of 25-35.
```{r mylatextable_ageonly_cn, results = "asis"}
age1<-ncommute_cleaned %>% count(age) %>% group_by(age)
age2<-ncommute_raw %>% count(age) %>% group_by(age)
age_n <- merge(age1, age2, by="age", all=TRUE)

age_n$n.x <- ifelse(is.na(age_n$n.x), 0, age_n$n.x )

names(age_n)<- c("Age", "n (cleaned data)", "n (raw data)")

htmlTable(age_n, caption = 'Number of Observations per Age group', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')

age1<-ncommute_cleaned %>% count(mode_1, age)
age2<-ncommute_raw %>% count(mode_1, age)
age_n <- merge(age1, age2, by=c("mode_1","age"), all=TRUE)

age_n$n.x <- ifelse(is.na(age_n$n.x), 0, age_n$n.x )

names(age_n)<- c("Mode","Age", "n (cleaned data)", "n (raw data)")

htmlTable(age_n, caption = 'Number of Observations per Mode and Age group', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')

choice_data_cleaned_ncommute$age <- factor(choice_data_cleaned_ncommute$age, levels = c("25-34 years", "Under 5 years old", "5-11 years", "12-15 years", "16-17 years", "18-24 years", "35-44 years", "45-54 years", "55-64 years",  "65-74 years", "75-84 years", "85 years or older"))

choice_data_raw_ncommute$age <- factor(choice_data_raw_ncommute$age, levels = c("25-34 years", "Under 5 years old", "5-11 years", "12-15 years", "16-17 years", "18-24 years", "35-44 years", "45-54 years", "55-64 years",  "65-74 years", "75-84 years", "85 years or older"))

m_cleaned <- mlogit(mode_1 ~  0| age | 0, choice_data_cleaned_ncommute)


m_raw <- mlogit(mode_1 ~  0| age | 0, choice_data_raw_ncommute)

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other"), digits = 2)

```

### 2.4 Race

*	Identifying as African American or Black increases the likelihood of choosing transit over.
* Individuals who identify themselves as Asian are more likely to choose walking or other modes relative to cars.
* Individuals who identify as Native Hawaiian or Pacific Islander are less likely to choose walking over car.
* Individuals who identify as Hispanic are more likely to choose bike over car and less likely to choose transit over car.
* Individuals who identify as white are more likely to choose bike and walk over car and less likely to choose transit over car.
* Individuals who selected other as their race are more likely to choose bike, and use other modes relative to car.

```{r mylatextable_raceonly_cn, results = "asis"}


race_cleaned<-rbind(c("African American or Black",length(ncommute_cleaned$race_afam[which(ncommute_cleaned$race_afam == 1)])),
            c("American Indian or Alaska Native",length(ncommute_cleaned$race_aiak[which(ncommute_cleaned$race_aiak == 1)])),
            c("Asian",length(ncommute_cleaned$race_asian[which(ncommute_cleaned$race_asian == 1)])),
            c("Native Hawaiian or Pacific Islander",length(ncommute_cleaned$race_hapi[which(ncommute_cleaned$race_hapi == 1)])),
            c("Hispanic",length(ncommute_cleaned$race_hisp[which(ncommute_cleaned$race_hisp == 1)])),
            c("White",length(ncommute_cleaned$race_white[which(ncommute_cleaned$race_white == 1)])),
            c("Other",length(ncommute_cleaned$race_other[which(ncommute_cleaned$race_other == 1)])),
            c("No answer",length(ncommute_cleaned$race_noanswer[which(ncommute_cleaned$race_noanswer == 1)])))

colnames(race_cleaned) <- c("Race","n")


race_raw<-rbind(c("African American or Black",length(ncommute_raw$race_afam[which(ncommute_raw$race_afam == 1)])),
            c("American Indian or Alaska Native",length(ncommute_raw$race_aiak[which(ncommute_raw$race_aiak == 1)])),
            c("Asian",length(ncommute_raw$race_asian[which(ncommute_raw$race_asian == 1)])),
            c("Native Hawaiian or Pacific Islander",length(ncommute_raw$race_hapi[which(ncommute_raw$race_hapi == 1)])),
            c("Hispanic",length(ncommute_raw$race_hisp[which(ncommute_raw$race_hisp == 1)])),
            c("White",length(ncommute_raw$race_white[which(ncommute_raw$race_white == 1)])),
            c("Other",length(ncommute_raw$race_other[which(ncommute_raw$race_other == 1)])),
            c("No answer",length(ncommute_raw$race_noanswer[which(ncommute_raw$race_noanswer == 1)])))

colnames(race_raw) <- c("Race", "n")

race_n <- merge(race_cleaned, race_raw, by="Race", all=TRUE)

names(race_n)<- c("Race", "n (cleaned data)", "n (raw data)")

htmlTable(race_n, caption = 'Number of Observations per Race group', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')


race_cleaned <- ncommute_cleaned %>% count(mode_1, race_afam , race_aiak , race_asian , race_hapi , race_hisp , race_white , race_other,race_noanswer  )

race_cleaned$race<-paste0(ifelse(race_cleaned$race_afam == 1,"African American or Black,",''), ifelse(race_cleaned$race_aiak == 1,"American Indian or Alaska Native,",''),
                         ifelse(race_cleaned$race_asian == 1,"Asian,",''), ifelse(race_cleaned$race_hapi == 1,"Native Hawaiian or Pacific Islander,",''), 
                         ifelse(race_cleaned$race_hisp == 1,"Hispanic,",''), ifelse(race_cleaned$race_white == 1,"White,",''),ifelse(race_cleaned$race_other == 1,"Other,",''),
                         ifelse(race_cleaned$race_noanswer == 1 | (race_cleaned$race_afam !=1 & race_cleaned$race_aiak !=1 & race_cleaned$race_asian !=1 & race_cleaned$race_hapi !=1
                                                                   & race_cleaned$race_hisp !=1 & race_cleaned$race_white != 1 & race_cleaned$race_other != 1),"No answer,",''))

race_cleaned <- race_cleaned[,c(1,11,10)]

race_cleaned$n[race_cleaned$mode_1 ==  "Car" & race_cleaned$race == race_cleaned$race[which(duplicated(race_cleaned[,c(1,2)]))]]<-sum(race_cleaned$n[race_cleaned$mode_1 ==  "Car" & race_cleaned$race == race_cleaned$race[which(duplicated(race_cleaned[,c(1,2)]))]]  )

race_cleaned$n[race_cleaned$mode_1 ==  "Walk" & race_cleaned$race == race_cleaned$race[which(duplicated(race_cleaned[,c(1,2)]))]]<-sum(race_cleaned$n[race_cleaned$mode_1 ==  "Walk" & race_cleaned$race == race_cleaned$race[which(duplicated(race_cleaned[,c(1,2)]))]]  )

race_cleaned$n[race_cleaned$mode_1 ==  "Transit" & race_cleaned$race == race_cleaned$race[which(duplicated(race_cleaned[,c(1,2)]))]]<-sum(race_cleaned$n[race_cleaned$mode_1 ==  "Transit" & race_cleaned$race == race_cleaned$race[which(duplicated(race_cleaned[,c(1,2)]))]]  )

race_cleaned$n[race_cleaned$mode_1 ==  "Other" & race_cleaned$race == race_cleaned$race[which(duplicated(race_cleaned[,c(1,2)]))]]<-sum(race_cleaned$n[race_cleaned$mode_1 ==  "Other" & race_cleaned$race == race_cleaned$race[which(duplicated(race_cleaned[,c(1,2)]))]]  )

race_cleaned <-unique(race_cleaned)

race_raw <- ncommute_raw %>% count(mode_1, race_afam , race_aiak , race_asian , race_hapi , race_hisp , race_white , race_other,race_noanswer  )

race_raw$race<-paste0(ifelse(race_raw$race_afam == 1,"African American or Black,",''), ifelse(race_raw$race_aiak == 1,"American Indian or Alaska Native,",''),
                         ifelse(race_raw$race_asian == 1,"Asian,",''), ifelse(race_raw$race_hapi == 1,"Native Hawaiian or Pacific Islander,",''), 
                         ifelse(race_raw$race_hisp == 1,"Hispanic,",''), ifelse(race_raw$race_white == 1,"White,",''),ifelse(race_raw$race_other == 1,"Other,",''),
                         ifelse(race_raw$race_noanswer == 1 | (race_raw$race_afam !=1 & race_raw$race_aiak !=1 & race_raw$race_asian !=1 & race_raw$race_hapi !=1
                                                                   & race_raw$race_hisp !=1 & race_raw$race_white != 1 & race_raw$race_other != 1),"No answer,",''))

race_raw <- race_raw[,c(1,11,10)]

race_raw$n[race_raw$mode_1 == "Car" & race_raw$race == race_raw$race[which(duplicated(race_raw[,c(1,2)]))]]<-sum(race_raw$n[race_raw$mode_1 == "Car" & race_raw$race == race_raw$race[which(duplicated(race_raw[,c(1,2)]))]]  )


race_raw$n[race_raw$mode_1 == "Transit" & race_raw$race == race_raw$race[which(duplicated(race_raw[,c(1,2)]))]]<-sum(race_raw$n[race_raw$mode_1 == "Transit" & race_raw$race == race_raw$race[which(duplicated(race_raw[,c(1,2)]))]]  )

race_raw$n[race_raw$mode_1 == "Walk" & race_raw$race == race_raw$race[which(duplicated(race_raw[,c(1,2)]))]]<-sum(race_raw$n[race_raw$mode_1 == "Transit" & race_raw$race == race_raw$race[which(duplicated(race_raw[,c(1,2)]))]]  )

race_raw <-unique(race_raw)

race_n <- merge(race_cleaned, race_raw, by=c("mode_1","race"))

race_n$n.x <- ifelse(is.na(race_n$n.x), 0, race_n$n.x )
race_n$n.y <- ifelse(is.na(race_n$n.y), 0, race_n$n.y )

names(race_n)<- c("Mode","Race", "n (cleaned data)", "n (raw data)")

htmlTable(race_n, caption = 'Number of Observations per Mode and Race group', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')

m_cleaned <- mlogit(mode_1 ~  0|  race_afam + race_aiak + race_asian + race_hapi + race_hisp + race_white + race_other | 0, choice_data_cleaned_ncommute)


m_raw <- mlogit(mode_1 ~  0| race_afam + race_aiak + race_asian + race_hapi + race_hisp + race_white + race_other | 0, choice_data_raw_ncommute)

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other","African American or Black:Bike", "African American or Black:Walk", "African American or Black:Transit", "African American or Black:Other", "American Indian or Alaska Native:Bike", "American Indian or Alaska Native:Walk", "American Indian or Alaska Native:Transit", "American Indian or Alaska Native:Other", "Asian:Bike", "Asian:Walk", "Asian:Transit", "Asian:Other", "Native Hawaiian or Pacific Islander:Bike", "Native Hawaiian or Pacific Islander:Walk", "Native Hawaiian or Pacific Islander:Transit", "Native Hawaiian or Pacific Islander:Other", "Hispanic:Bike",  "Hispanic:Walk",  "Hispanic:Transit",  "Hispanic:Other", "White:Bike", "White:Walk", "White:Transit", "White:Other", "Other:Bike", "Other:Walk",  "Other:Transit",  "Other:Other"), digits = 2)

```


### 2.5 Gender

Individuals who identify as male are more likely to use bike, walk, and transit than other genders.

```{r mylatextable_genderonly_cn, results = "asis"}
gen1<-ncommute_cleaned %>% count(gender) %>% group_by(gender)
gen2<-ncommute_raw %>% count(gender) %>% group_by(gender)
gen_n <- merge(gen1, gen2, by="gender", all=TRUE)

names(gen_n)<- c("Gender", "n (cleaned data)", "n (raw data)")
gen_n<-gen_n[order(gen_n$Gender),]

htmlTable(gen_n, caption = 'Number of Observations per Gender group', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')

gen1<-ncommute_cleaned %>% count(mode_1, gender)
gen2<-ncommute_raw %>% count(mode_1, gender)
gen_n <- merge(gen1, gen2, by=c("mode_1","gender"), all=TRUE)

gen_n$n.x <- ifelse(is.na(gen_n$n.x), 0, gen_n$n.x )

names(gen_n)<- c("Mode","Gender", "n (cleaned data)", "n (raw data)")

htmlTable(gen_n, caption = 'Number of Observations per Mode and Gender group', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')

choice_data_cleaned_ncommute$male<-ifelse( choice_data_cleaned_ncommute$gender == "Male", "Male","Non-Male")
choice_data_raw_ncommute$male<-ifelse( choice_data_raw_ncommute$gender == "Male", "Male","Non-Male")

choice_data_cleaned_ncommute$male <- ordered(choice_data_cleaned_ncommute$male, levels = c("Non-Male", "Male"))
choice_data_raw_ncommute$male <- ordered(choice_data_raw_ncommute$male, levels = c("Non-Male", "Male"))


m_cleaned <- mlogit(mode_1 ~  0| male | 0, choice_data_cleaned_ncommute)

m_raw <- mlogit(mode_1 ~  0| male | 0, choice_data_raw_ncommute)

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other", "Male:Bike", "Male:Walk", "Male:Transit", "Male:Other"), digits = 2)

```

### 2.6 Income

Individuals whose income is under \$75,000 are more likely to choose transit and other modes relative to car compared to individuals whose income is over \$75,000.


```{r mylatextable_incomeonly_cn, results = "asis"}

inc1<-ncommute_cleaned %>% count(hhincome_detailed) %>% group_by(hhincome_detailed)
inc2<-ncommute_raw %>% count(hhincome_detailed) %>% group_by(hhincome_detailed)
inc_n <- merge(inc1, inc2, by="hhincome_detailed", all=TRUE)

names(inc_n)<- c("Income", "n (cleaned data)", "n (raw data)")
inc_n<-inc_n[order(inc_n$Income),]


#stargazer(inc_n,  type = 'html', summary=FALSE, rownames = FALSE)

htmlTable(inc_n, caption = 'Income', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 13px 0px;')


inc1<-ncommute_cleaned %>% count(mode_1, hhincome_detailed) 
inc2<-ncommute_raw %>% count(mode_1, hhincome_detailed)
inc_n <- merge(inc1, inc2, by=c("mode_1","hhincome_detailed"), all=TRUE)

inc_n$n.x <- ifelse(is.na(inc_n$n.x), 0, inc_n$n.x )

names(inc_n)<- c("Income", "n (cleaned data)", "n (raw data)")
inc_n<-inc_n[order(inc_n$Income),]

htmlTable(inc_n, caption = 'Number of Obervations per Mode and Income group', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 13px 0px;')



m_cleaned <- mlogit(mode_1 ~  0| hhincome_detailed | 0, choice_data_cleaned_ncommute)


m_raw <- mlogit(mode_1 ~  0| hhincome_detailed | 0, choice_data_raw_ncommute)

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other","Prefer not to answer:Bike","Prefer not to answer:Walk","Prefer not to answer:Transit","Prefer not to answer:Other","Under $75,000:Bike","Under $75,000:Walk","Under $75,000:Transit","Under $75,000:Other"), digits = 2)

```


### 2.7 Trip Purpose

Reference purpose category is "Home". 
*	Individuals are less likely to walk to school compared to driving relative to trips ending at home.
*	For escort trips people are more likely to use car rather than any other modes compared to home trips.
*	People are less likely to bike or walk for shopping.
*	People are less likely to walk for getting a meal compared to going home relative to using a car.
*	People are more likely to bike and walk for social/recreation purposes than going home compared to car.
*	People are less likely to walk for errand purposes.


```{r mylatextable_purposeonly_cn, results = "asis"}
mod_dest_cleaned <- ncommute_cleaned %>% count(mode_1, dest_purpose_cat1, sort = TRUE)

mod_dest_raw <- ncommute_raw %>% count(mode_1, dest_purpose_cat1, sort = TRUE)

mod_dest_sum <- merge(mod_dest_cleaned,mod_dest_raw, by = c("mode_1", "dest_purpose_cat1"),all = T)

mod_dest_sum$n.x<- ifelse(is.na(mod_dest_sum$n.x),0,mod_dest_sum$n.x)
mod_dest_sum$n.y<- ifelse(is.na(mod_dest_sum$n.y),0,mod_dest_sum$n.y)

names(mod_dest_sum) <- c("Mode","Purpose", "Cleand data","Raw data")

htmlTable(mod_dest_sum, caption = 'Trip Mode vs. Trip Purpose', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')

m_cleaned <- mlogit(mode_1 ~  0| dest_purpose_cat1 | 0, choice_data_cleaned_ncommute)


m_raw <- mlogit(mode_1 ~  0| dest_purpose_cat1 | 0, choice_data_raw_ncommute)

# If model specification changes, covariate labels need to be updated

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other", "School:Bike","School:Walk","School:Transit","School:Other","Escort:Bike","Escort:Walk","Escort:Transit","Escort:Other","Shop:Bike","Shop:Walk","Shop:Transit","Shop:Other","Meal:Bike","Meal:Walk","Meal:Transit","Meal:Other","Social/Recreation:Bike","Social/Recreation:Walk","Social/Recreation:Transit","Social/Recreation:Other","Errand/Other:Bike","Errand/Other:Walk","Errand/Other:Transit","Errand/Other:Other","Change mode:Bike","Change mode:Walk","Change mode:Transit","Change mode:Other"), digits = 2)

```


### 2.8 Significant variables only


```{r, results = "asis"}

m_cleaned <- mlogit(mode_1 ~  0|male + trip_path_distance + vehicle_count + race_white + race_hisp | 0, choice_data_cleaned_ncommute)

m_raw <- mlogit(mode_1 ~  0| male + trip_path_distance + vehicle_count + race_white + race_hisp | 0, choice_data_raw_ncommute)

stargazer(m_cleaned, m_raw, title="Results", align=TRUE, type = 'html', column.labels = c("Cleaned data", "Raw data"), style = "qje", covariate.labels = c("ASC:Bike", "ASC:Walk", "ASC:Transit", "ASC:Other", "Male:Bike", "Male:Walk", "Male:Transit", "Male:Other", "Trip distance:Bike", "Trip distance:Walk", "Trip distance:Transit", "Trip distance:Other", "Vehicle count:Bike", "Vehicle count:Walk", "Vehicle count:Transit", "Vehicle count:Other", "White:Bike", "White:Walk", "White:Transit", "White:Other", "Hispanic:Bike", "Hispanic:Other", "Hispanic:Transit", "Hispanic:Walk"))


freq_cleaned<-as.data.frame(m_cleaned$freq/dim(ncommute_cleaned)[1])
freq_cleaned$Freq<-round(freq_cleaned$Freq,4)
predicted_freq_cleaned <- as.data.frame(round(apply(fitted(m_cleaned, outcome=FALSE), 2, mean),4))
predicted_freq_cleaned <- cbind(rownames(predicted_freq_cleaned),predicted_freq_cleaned)
colnames(predicted_freq_cleaned)<-colnames(freq_cleaned)
df_cleaned<-merge(freq_cleaned,predicted_freq_cleaned, by = "choice")
colnames(df_cleaned)<-c("Choice","Choice Frequency","Fitted Probability")
df_cleaned$change<-round((df_cleaned$`Fitted Probability`-df_cleaned$`Choice Frequency`)*100/df_cleaned$`Choice Frequency`,3)
colnames(df_cleaned)[4]<-"Change (%)"
htmlTable(df_cleaned, caption = 'Cleaned data model frequencies vs. probabilities', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding-left: 3.5em; padding-right: 3.5em;')


freq_raw<-as.data.frame(m_raw$freq/dim(ncommute_raw)[1])
freq_raw$Freq<-round(freq_raw$Freq,4)
predicted_freq_raw <- as.data.frame(round(apply(fitted(m_raw, outcome=FALSE), 2, mean),4))
predicted_freq_raw <- cbind(rownames(predicted_freq_raw),predicted_freq_raw)
colnames(predicted_freq_raw)<-colnames(freq_raw)
df_raw<-merge(freq_raw,predicted_freq_raw, by = "choice")
colnames(df_raw)<-c("Choice","Choice Frequency","Fitted Probability")
df_raw$change<-round((df_raw$`Fitted Probability`-df_raw$`Choice Frequency`)*100/df_raw$`Choice Frequency`,3)
colnames(df_raw)[4]<-"Change (%)"
htmlTable(df_raw, caption = 'Raw data model frequencies vs. probabilities', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding-left: 3.5em; padding-right: 3.5em;')


odd_ratios <- cbind(round(exp(coefficients(m_cleaned)),2),round(exp(coefficients(m_raw)),2))


colnames(odd_ratios) <- c("Cleand data","Raw data")
rownames(odd_ratios)[5:16] <- c("Male:Bike","Male:Walk","Male:Transit","Male:Other", "Trip distance:Bike", "Trip distance:Walk", "Trip distance:Transit", "Trip distance:Other", "Vehicle count:Bike", "Vehicle count:Walk", "Vehicle count:Transit", "Vehicle count:Other")
rownames(odd_ratios)[17:24] <- c("White:Bike","White:Walk","White:Transit","White:Other","Hispanic:Bike","Hispanic:Walk","Hispanic:Transit","Hispanic:Other")

htmlTable(odd_ratios, caption = 'Odd Ratios', rownames = FALSE,
          ## padding to cells: top side bottom
          css.cell = 'padding: 0px 10px 0px;')
```
