```{r}

```

```{r}

```

---
title: "Race and Ethnicity Analysis"
format: 
  html:
    toc: true
editor: visual
execute:
  warning: false
---

# General setup 

PSRC and RSG have built a new package called travelSurveyTools to work with travel survey data. It is recommended to read through this documentation before you start using the package. <https://rsginc.github.io/travelSurveyTools/>

## Read in Libraries
```{r setup}
# YOU MIGHT need to download some special packages on github: travelSurveyTools, psrcelmer, and psrcplot. IF you need to download them, here's how:
# library(devtools)
# 
# devtools::install_github('rsgInc/travelSurveyTools') 
# devtools::install_github('psrc/psrcelmer')
# devtools::install_github('psrc/psrcplot')

library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
library(table1)
# source('../survey-23-preprocess.R')
source('C:/Users/mrichards/Documents/GitHub/travel-studies/2023/summary/survey-23-preprocess.R')
```

## Read in Codebook
```{r}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```

## Read in Data from Elmer
```{r }
# hh<- get_query(sql= "select  household_id as hh_id,
#                res_dur, prev_home_wa, vehicle_count, numadults, hhincome_detailed,
#                hhincome_broad,survey_year, prev_home_tract20,
#                prev_rent_own, rent_own, prev_res_type, res_type, hh_weight, hhsize
#                 from HHSurvey.v_households_labels")

person<- get_query(sql= "select household_id as hh_id, race_category,
                   race_afam, race_aiak, race_asian, race_hapi, race_noanswer, race_other, race_other_specify, race_white,
                   ethnicity_1, ethnicity_2, ethnicity_3, ethnicity_4, ethnicity_997, ethnicity_999, ethnicity_other,
                   person_id, survey_year, person_weight, gender, age
                   from HHSurvey.v_persons_labels")
                
# setDT(hh)
setDT(person)
```

## Set IDs as characters
```{r}

# hh[, hh_id:=as.character(hh_id)]
# person[, hh_id:=as.character(hh_id)]
person[, person_id := as.character(person_id)]

# hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
```

```         
```

The survey instrument is here: J:/Projects/Surveys/HHTravel/Survey2023/Documents/Survey_Instrument/Puget_Sound_HTS_Questionnaire_2023_Final.docx

Open the codebook and review the variable and it's values. J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx

# Data setup
```{r}
# somehow a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```


```{r}
# nrow(person) #25364
# nrow(person$ethnicity_999) #prefer not to answer: NULL
# nrow(person$race_noanswer) #NULL

table(person$age)
adult <- person %>%
  filter(age%in% c('18-24 years', '25-34 years', '35-44 years', '45-54 years', '55-64 years', '65-74 years', '75-84 years', '85 years or older'))
# table(adult$age)
# nrow(adult) #21492
# length(unique(adult$person_id)) #21492 - just to check

# unique(adult$race_other_specify) #checking 'other' responses
# length(unique(adult$race_other_specify)) #99 people who selected 'other' 

adult_23 <- adult %>% 
  filter(survey_year=="2023") #5931 - to know how many respondents total
```

# Race and Ethnicity
## Race
```{r}
# pivot table so that each row is a different race column
adult_pivot_race <- adult_23 %>% 
  pivot_longer(cols = c(race_afam, race_aiak, race_asian, race_hapi, race_noanswer, race_other, race_white),
               names_to = "race_group",
               values_to = "select")
# unique(adult_pivot_race$race_other_specify)
# table(adult_pivot_race$survey_year)

adult_pivot_race_select <- adult_pivot_race %>% 
  filter(select %in% 'Selected')

person_id_multiple_23_race <- adult_pivot_race_select %>% 
  group_by(person_id) %>% summarize(n=n()) %>% #count how many racial groups were selected per person
  right_join(adult_pivot_race_select, by = ("person_id")) #join to main table so number connected to races selected

# nrow(person_id_multiple_23_race) #22056
table(person_id_multiple_23_race$n)
```

```{r, include = FALSE, eval=FALSE}
## before making function (for checking) ----
adult_1race <- person_id_multiple %>%
  filter(n %in% 1)
adult_2race <- person_id_multiple %>%
  filter(n %in% 2)
adult_3race <- person_id_multiple %>%
  filter(n %in% 3)
adult_4race <- person_id_multiple %>%
  filter(n %in% 4)
adult_5race <- person_id_multiple %>%
  filter(n %in% 5)
adult_6race <- person_id_multiple %>%
  filter(n %in% 6)

adult_2race_summary <- adult_2race %>%
  pivot_wider(names_from = race_group, values_from = n, values_fill = 0) %>%
  dplyr::select(race_asian:race_other) %>%
  group_by(across(race_asian:race_other)) %>% count %>% arrange(desc(n)) %>%
  mutate(total=sum(adult_2race_wide$n)) %>%
  mutate(prop=(round(n/total, digits = 3)))
```

```{r}
# create simplified table for combining with ethnicity data (eventually)
adult_race <- person_id_multiple_23_race %>%
  dplyr::select(person_id, race_group, n)

## after making function ----
race_list <- c('race_afam', 'race_aiak', 'race_asian', 'race_hapi', 'race_white', 'race_noanswer', 'race_other')

count_combination_race <- function(group, n_races){
  temp <- person_id_multiple_23_race %>% 
    filter(n %in% n_races)
  
  pivot <- temp %>% 
    pivot_wider(names_from = group, values_from = n, values_fill = 0) %>% 
    dplyr::select(any_of(race_list)) %>% 
    group_by(across(any_of(race_list))) %>% count %>% arrange(desc(n))
  
  output <- pivot %>% 
    mutate(total=sum(pivot$n)) %>% 
    mutate(prop=(round(n/total, digits = 3)))
}

adult_1race_summary <- count_combination_race('race_group', 1) #5655 people
adult_2race_summary <- count_combination_race('race_group', 2) #234 people 
adult_3race_summary <- count_combination_race('race_group', 3) #32 people
adult_4race_summary <- count_combination_race('race_group', 4) #2 people
adult_5race_summary <- count_combination_race('race_group', 5) #0 people
adult_6race_summary <- count_combination_race('race_group', 6) #0 people
```
```{r, echo=FALSE}
adult_1race_summary
```
1 RACE:

* 68.7%: white
* 15.5%: asian
* 11.1%: no answer
* 2.7%: african american/black
* 1.1%: 'Other'
* 0.5%: american indian/alaskan native
* 0.4%: hawaiian/pacific islander

```{r, echo=FALSE}
adult_2race_summary
```
2 RACES:

* 51.3%: asian AND white 
* 18.8%: american indian/alaskan native AND white
* 9%: african american/black AND white
* 6%: other AND white
* 5.1%: asian AND hawaiian/pacific islander
* 4.3%: hawaiian/pacific islander AND white
* 2.1%: african american/black AND white

```{r, echo=FALSE}
adult_3race_summary
```
3 RACES:

* 40.6%: african american/black AND american indian/alaskan native AND white
* 21.9: asian AND hawaiian/pacific islander AND white
* 9.4%: asian AND other AND white
* 9.4%: american indian/alaskan native AND asian AND white
* 9.4%: african american/black AND asian AND white


```{r, include=FALSE, eval=FALSE}
# testing numbers
race_afam <- person_id_multiple_23_race %>% 
  filter(n==1,
         race_group=="race_afam")

check <- person_id_multiple_23_race %>%
  filter(ethnicity_999=="Selected") #789
length(unique(check$person_id)) #776 

pivot <- check %>% 
    pivot_wider(names_from = race_group, values_from = n, values_fill = 0) %>%
  dplyr::select(any_of(race_list)) %>%
  group_by(across(any_of(race_list))) %>% 
    count %>% arrange(desc(n))

output <- pivot %>% 
    mutate(total=sum(pivot$n)) %>% 
    mutate(prop=(round(n/total, digits = 3)))
```


## Ethnicity
```{r}
# pivot table so that each row is a different ethnicity column
adult_pivot_ethnicity <- adult_23 %>% 
  pivot_longer(cols = c(ethnicity_1, ethnicity_2, ethnicity_3, ethnicity_4, ethnicity_997, ethnicity_999),
               names_to = "ethnicity_group",
               values_to = "select") #%>% 
  # mutate(eth_expand=case_when(ethnicity_group=="ethnicity_1"~"Not of Hispanic, Latino, or Spanish origin",
  #                             ethnicity_group=="ethnicity_2"~"Mexican, Mexican American, Chicano",
  #                             ethnicity_group=="ethnicity_3"~"Puerto Rican",
  #                             ethnicity_group=="ethnicity_4"~"Cuban",
  #                             ethnicity_group=="ethnicity_997"~"Another Hispanic, Latino, or Spanish origin",
  #                             ethnicity_group=="ethnicity_999"~"Prefer not to answer"))

# unique(adult$ethnicity_1)
table(adult$ethnicity_1, adult$ethnicity_2)
# unique(adult_pivot_ethnicity$ethnicity_other) #83
# unique(adult_pivot_ethnicity$survey_year) #83
# length(unique(adult_pivot_ethnicity$person_id)) #5931 - correct number of people
# table(adult_pivot_ethnicity$select)

adult_pivot_ethnicity_select <- adult_pivot_ethnicity %>%
  filter(select %in% 'Selected')

# length(unique(adult_pivot_ethnicity_select$person_id)) #5923 - 8 people missing


person_id_multiple_23_eth <- adult_pivot_ethnicity_select %>% 
  group_by(person_id) %>% summarize(n=n()) %>% #count how many ethnicity groups were selected per person
  right_join(adult_pivot_ethnicity_select, by = ("person_id")) #join to main table so number connected to ethnicity selected

# length(unique(person_id_multiple_23_eth$person_id)) #5923 - 8 people missing
table(person_id_multiple_23_eth$n)

# unique(person_id_multiple_23_eth$ethnicity_other)
# length(unique(person_id_multiple_23_eth$person_id)) #5923 - 8 people missing

# where are 8 missing people?
check_join <- adult_pivot_ethnicity %>% 
  anti_join(adult_pivot_ethnicity_select, by="person_id") 

# length(unique(check_join$person_id)) #8
missing_ids <- unique(check_join$person_id)

missing_people <- adult_23 %>% 
  filter(person_id %in% missing_ids) %>%
  print
```
The 2023 adult dataset contains 5,931 people, but when looking at the ethnicity data, there were 5,923 people. The table above shows the 8 people who are missing - they didn't respond to the race or ethnicity questions?

```{r, include=FALSE, eval=FALSE}
## before making function (for checking) ----
adult_2eth <- person_id_multiple_23 %>%
  filter(n %in% 2)
adult_3eth <- person_id_multiple_23 %>%
  filter(n %in% 3)

adult_3eth_summary <- adult_3eth %>%
  pivot_wider(names_from = ethnicity_group, values_from = n, values_fill = 0) %>%
  dplyr::select(any_of(c('ethnicity_1', 'ethnicity_2', 'ethnicity_3', 'ethnicity_4', 'ethnicity_997', 'ethnicity_999'))) %>%
  group_by(across(any_of(c('ethnicity_1', 'ethnicity_2', 'ethnicity_3', 'ethnicity_4', 'ethnicity_997', 'ethnicity_999')))) %>% count %>% arrange(desc(n))

adult_2eth_summary <- adult_2eth %>%
  pivot_wider(names_from = eth_expand, values_from = n, values_fill = 0) %>%
  dplyr::select(any_of(c('Not of Hispanic, Latino, or Spanish origin',
                           'Mexican, Mexican American, Chicano',
                           'Puerto Rican',
                           'Cuban',
                           'Another Hispanic, Latino, or Spanish origin',
                           'Prefer not to answer'))) %>%
  group_by(across(any_of(c('Not of Hispanic, Latino, or Spanish origin',
                           'Mexican, Mexican American, Chicano',
                           'Puerto Rican',
                           'Cuban',
                           'Another Hispanic, Latino, or Spanish origin',
                           'Prefer not to answer')))) %>% count %>% arrange(desc(n))
```

```{r}
# create simplified table for combining with race data (eventually)
adult_eth <- person_id_multiple_23_eth %>% 
  dplyr::select(person_id, ethnicity_group, n)

## after making function ----
ethnicity_list <- c('ethnicity_1', 'ethnicity_2', 'ethnicity_3', 'ethnicity_4', 'ethnicity_997', 'ethnicity_999')

count_combination_eth <- function(group, n_ethnicities){
  temp <- person_id_multiple_23_eth %>% 
    filter(n %in% n_ethnicities)
  
  pivot <- temp %>% 
    pivot_wider(names_from = group, values_from = n, values_fill = 0) %>%
  dplyr::select(any_of(ethnicity_list)) %>%
  group_by(across(any_of(ethnicity_list))) %>% 
    count %>% arrange(desc(n))
  
  output <- pivot %>% 
    mutate(total=sum(pivot$n)) %>% 
    mutate(prop=(round(n/total, digits = 3)))
}

adult_1eth_summary <- count_combination_eth('ethnicity_group', 1) #5902 people
adult_2eth_summary <- count_combination_eth('ethnicity_group', 2) #19 people 
adult_3eth_summary <- count_combination_eth('ethnicity_group', 3) #2 people
adult_4eth_summary <- count_combination_eth('ethnicity_group', 4) #0 people
adult_5eth_summary <- count_combination_eth('ethnicity_group', 5) #0 people
adult_6eth_summary <- count_combination_eth('ethnicity_group', 6) #0 people
```

```{r, echo=FALSE}
adult_1eth_summary
```
1 ETHNICITY:

* 81.3%: Not of Hispanic/Latino/Spanish origin (4801/5902)
* 13.1%: 'Prefer not to answer' (776/5902)
* 3.1%: Mexican, Mexican American, Chicano (182/5902)
* 1.9%: 'Another' (112/5902)
* 0.4%: Puerto Rican (22/5902)
* 0.2%: Cuban (9/5902)

```{r, echo=FALSE}
adult_2eth_summary
```
2 ETHNICITIES:

* 63.2%: Mexican, Mexican American, Chicano AND 'Another'
* 21.1%: Puerto Rican AND 'Another'
* 5.3%: Cuban AND 'Another'
* 5.3%: Puerto Rican AND Cuban
* 5.3%: Mexican, Mexican American, Chicano AND Cuban
* 0.2%: Cuban

```{r, echo=FALSE}
adult_3eth_summary
```
3 ETHNICITIES:

* 100%: Mexican, Mexican American, Chicano AND Puero Rican AND 'Another'

```{r, include=FALSE, eval=FALSE}
# test
eth_3 <- person_id_multiple_23_eth %>% 
  filter(n==1,
         ethnicity_group=="ethnicity_3")

table(person_id_multiple_23_eth$n) #5902 with 1 ethnicity
filter <- person_id_multiple_23_eth %>% 
  filter(n==1) #5902
filter2 <- filter %>% 
  filter(ethnicity_group=="ethnicity_1") #4801 people reporting 1 ethnicity are not hispanic/latino/Spanish origin
```


## Race + ethnicity combinations
### For recoding exercise
```{r}
# 1. generating table for re-coding exercise 
# nrow(adult_23) #5931

other_race <- adult_23 %>% 
  filter(race_other=="Selected") #133
other_race_specify <- adult_23 %>% 
  filter(!is.na(race_other_specify)) #133
other_eth_specify <- adult_23 %>% 
  filter(!is.na(ethnicity_other)) #132

other_race_eth <- adult_23 %>% 
  filter(!is.na(race_other_specify) |
           !is.na(ethnicity_other)) #226
# length(unique(other_race_eth$person_id)) #226 - just to make sure no duplicates

# for writing a data.frame or list of data.frames to an xlsx file
library(openxlsx)
write.xlsx(other_race_eth, 'T:/2024July/Mary/HHTS/2023/adult_2023_race_eth_recode.xlsx')
```

### Understanding general distribution
```{r}
# create simplified table that 
race_temp <- adult_race %>% 
  dplyr::group_by(person_id) %>% 
  summarise(n_race=mean(n)) #5923 - making sure each row is a unique respondent
eth_temp <- adult_eth %>% 
  dplyr::group_by(person_id) %>% 
  summarise(n_eth=mean(n)) #5923 - making sure each row is a unique respondent

race_eth_temp <- race_temp %>% 
  right_join(eth_temp, by = "person_id") #5923 - making sure each row is a unique respondent

# checking to make sure that there are the correct numbers of respondents with 1,2,3 races/ethnicities
# table(race_eth_temp$n_race)

# join to original wide table with all race/ethnicity
adult_race_eth_count <- adult %>% 
  dplyr::filter(survey_year=="2023") %>% 
  dplyr::select_if(grepl("person_id", names(.)) | 
                     grepl("race_", names(.)) | 
                     grepl("ethnicity_", names(.))) %>% #5931
  right_join(race_eth_temp, by = "person_id")

# length(unique(adult_race_eth_count$person_id)) #5931 - making sure each row is a unique respondent

# detail_names <- c("Not of Hispanic, Latino, or Spanish origin"="ethnicity_1",
#                   "Mexican, Mexican American, Chicano"="ethnicity_2",
#                   "Puerto Rican"="ethnicity_3",
#                   "Cuban"="ethnicity_4",
#                   "Another ethnicity"="ethnicity_997",
#                   "Prefer not to answer"="ethnicity_999")
# rename(adult_race_eth_count, all_of(detail_names))

# table(adult_race_eth_count$n_race, adult_race_eth_count$n_eth)
table1(~ factor(n_eth) | as.factor(n_race), data=adult_race_eth_count) #formatted table
```

The table shows general trends and combinations

* 5,637/5,923 respondents : 1 race and 1 ethnicity (95.2%)
* 17/5,923 respondents : 1 race and 2 ethnicities (0.29%)
* 1/5,923 respondents : 1 race and 3 ethnicities (0%)
* etc. 

### Exploring different combinations
Based on the table, the vast majority of respondents chose 1 race and 1 ethnicity
```{r, include=FALSE, eval=FALSE}
# before making function (for checking) ----
race_eth_1 <- adult_race_eth_count %>%
  dplyr::filter(n_race==1 & n_eth==1) %>%
  dplyr::select(any_of(ethnicity_list)) %>%
  group_by(across(any_of(ethnicity_list))) %>%
    count %>% arrange(desc(n))
```

```{r}
## after making function ----
race_eth_combination <- function(num_race, num_eth, category_list){
  output <- adult_race_eth_count %>% 
    dplyr::filter(n_race==num_race & n_eth==num_eth) %>% 
    dplyr::select(any_of(ethnicity_list)) %>%
    group_by(across(any_of(ethnicity_list))) %>% 
    count %>% arrange(desc(n))
  
  final <- output %>% 
    mutate(total=sum(output$n)) %>% 
    mutate(prop=(round(n/total, digits = 3)))
}

race_1_eth_1 <- race_eth_combination(1,1,ethnicity_list) #5637
race_2_eth_1 <- race_eth_combination(2,1,ethnicity_list) #233
race_3_eth_1 <- race_eth_combination(3,1,ethnicity_list) #30
race_4_eth_1 <- race_eth_combination(4,1,ethnicity_list) #2
race_5_eth_1 <- race_eth_combination(5,1,ethnicity_list) #0

race_1_eth_2 <- race_eth_combination(1,2,ethnicity_list) #17
race_1_eth_3 <- race_eth_combination(1,3,ethnicity_list) #1
race_1_eth_4 <- race_eth_combination(1,4,ethnicity_list) #0
```
1 RACE/ 1 ETHNICITY: 5,637 respondents

* 81.3%: not Hispanic/Latino/Spanish origin
* 13.4%: 'Prefer not to answer'
* 3%: Mexican, Mexican American, Chicano
* 1.6%: 'Another'
* 0.3%: Puerto Rican
* 0.1%: Cuban

2 RACES/1 ETHNICITY: 233 respondents

* 85.4%: not Hispanic/Latino/Spanish origin
* 5.6%: 'Another'
* 5.2%: Mexican, Mexican American, Chicano
* 2.6%: 'Prefer not to answer'
* 0.9%: Puerto Rican
* 0.4%: Cuban

...

1 RACE / 2 ETHNICITIES: 17 respondents

* 64.7%: Mexican, Mexican American, Chicano AND 'Another'
* 17.6%: Puerto Rican AND 'Another'
* 5.9%: Cuban AND 'Another'
* 5.9%: Puerto Rican AND Cuban
* 5.9%: Mexican, Mexican American, Chicano AND Cuban

```{r, include=FALSE, eval=FALSE}
# before making function (for checking) ----
eth_1 <- person_id_multiple_23_eth %>%
  filter(n==1,
         ethnicity_group=="ethnicity_1") #4801
pivot <- eth_1 %>%
  pivot_wider(names_from = 'ethnicity_group', values_from = n, values_fill = 0) %>%
  dplyr::select(any_of(race_list)) %>%
  group_by(across(any_of(race_list))) %>% count %>% arrange(desc(n))
output <- pivot %>%
  mutate(total=sum(pivot$n)) %>%
  mutate(prop=(round(n/total, digits = 3)))

length(unique(eth_1$person_id))
```

```{r}
## after making functions ----
count_combination_race_by_ethnicitiy <- function(n_ethnicities, ethnicity_field, group){
  data <- person_id_multiple_23_eth %>% 
    filter(n==n_ethnicities,
           ethnicity_group==ethnicity_field)
  
  pivot <- data %>% 
    pivot_wider(names_from = group, values_from = n, values_fill = 0) %>% 
    dplyr::select(any_of(race_list)) %>% 
    group_by(across(any_of(race_list))) %>% count %>% arrange(desc(n))
  
  output <- pivot %>% 
    mutate(total=sum(pivot$n)) %>% 
    mutate(prop=(round(n/total, digits = 3)))
}

count_combination_race_by_ethnicitiy2 <- function(ethnicity_field, group){
  data <- person_id_multiple_23_eth %>% 
    filter(ethnicity_group!=ethnicity_field)
  
  pivot <- data %>% 
    pivot_wider(names_from = group, values_from = n, values_fill = 0) %>% 
    dplyr::select(any_of(race_list)) %>% 
    group_by(across(any_of(race_list))) %>% count %>% arrange(desc(n))
  
  output <- pivot %>% 
    mutate(total=sum(pivot$n)) %>% 
    mutate(prop=(round(n/total, digits = 3)))
}
```

#### Not Hispanic/Latino/Spanish Origin
```{r}
eth1_race <- count_combination_race_by_ethnicitiy(1, 'ethnicity_1', 'ethnicity_group')

eth1_race
```
1 ETHNICITY (Not Hispanic/Latino/Spanish origin):

* 74.2%: white (3562/4801)
* 16.4%: asian (786/4801)
* 2.7% african american (129/4801)
* 2.4% white AND asian
* 1.2% didn't provide an answer

```{r, number check, include=FALSE, eval=FALSE}
## CHECKING NUMBERS
check <- adult_23 %>%
  filter(ethnicity_1=="Selected",
         ethnicity_2=="Not selected",
         ethnicity_3=="Not selected",
         ethnicity_4=="Not selected",
         ethnicity_997=="Not selected",
         ethnicity_999=="Not selected") #4750 
length(unique(check$person_id))

check <- adult_23 %>%
  filter(ethnicity_1=="Selected",
         is.na(ethnicity_2),
         is.na(ethnicity_3),
         is.na(ethnicity_4),
         is.na(ethnicity_997),
         is.na(ethnicity_999)) #51 
length(unique(check$person_id))

# why are these numbers not lining up? something is wrong with the person_id_multiple_23_eth dataset?
check_join <- eth_1 %>% 
  anti_join(check, by="person_id") 

length(unique(check_join$person_id)) #51
missing_ids <- unique(check_join$person_id)

missing_people <- adult_23 %>% 
  filter(person_id %in% missing_ids)
# the missing 51 people - a recoding issue - NA vs Not Selected after 51 respondents were updated 
# the values have been updated 7/25, so all set
```

#### Hispanic/Latino/Spanish Origin
```{r}
not_eth1_race <- count_combination_race_by_ethnicitiy2('ethnicity_1', 'ethnicity_group') #1122

#check to make sure function works
# adult_23 #5931 people in dataset
# person_id_multiple_23_eth #5946 ethnicities (some people have multiple ethnicities)
# length(unique(adult_23$person_id)) #5931 people in dataset
# length(unique(person_id_multiple_23_race$person_id)) #5923
# length(unique(person_id_multiple_23_eth$person_id)) #5923
# 
# check <- person_id_multiple_23_eth %>% 
#   filter(ethnicity_group!="ethnicity_1") #1145
# length(unique(check$person_id)) #1122

not_eth1_race
```
1 ETHNICITY (Hispanic/Latino/Spanish origin):

* 51.1%: no answer (573/1122)
* 28.6%: white (321/1122)
* 7.9% asian (89/1122)
* 5.1% other 
* 2.1% african american/black
* 1.0%: white AND other
* 0.8%: american indian/alaskan native AND white
* 0.6%: african american/black AND american indian/alaskan native AND white
* 0.5%: asian AND white
* 0.4%: american indian/alaskan native
