```{r}

```

```{r}

```

---
title: "Race and Ethnicity Analysis"
format: 
  html:
    toc: true
editor: visual
execute:
  warning: false
---

# General setup 

PSRC and RSG have built a new package called travelSurveyTools to work with travel survey data. It is recommended to read through this documentation before you start using the package. <https://rsginc.github.io/travelSurveyTools/>

## Read in Libraries
```{r setup}
# YOU MIGHT need to download some special packages on github: travelSurveyTools, psrcelmer, and psrcplot. IF you need to download them, here's how:
# library(devtools)
# 
# devtools::install_github('rsgInc/travelSurveyTools') 
# devtools::install_github('psrc/psrcelmer')
# devtools::install_github('psrc/psrcplot')
devtools::install_github('psrc/psrc.travelsurvey')

library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
library(table1)
library(psrc.travelsurvey)
# source('../survey-23-preprocess.R')
source('C:/Users/mrichards/Documents/GitHub/travel-studies/2023/summary/survey-23-preprocess.R')
```

## Read in Codebook
```{r}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```

## Read in Data from Elmer
```{r }
hh<- get_query(sql= "select  household_id as hh_id, hh_race_category, hh_weight, survey_year
                from HHSurvey.v_households_labels")

person<- get_query(sql= "select household_id as hh_id, race_category, survey_year,
                   race_afam, race_aiak, race_asian, race_hapi, race_noanswer, race_other, race_other_specify, race_white,
                   ethnicity_1, ethnicity_2, ethnicity_3, ethnicity_4, ethnicity_997, ethnicity_999, ethnicity_other,
                   person_id, person_weight, gender, age
                   from HHSurvey.v_persons_labels")
                
setDT(hh)
setDT(person)
```

## Set IDs as characters
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
person[, person_id := as.character(person_id)]

hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
```

```         
```

The survey instrument is here: J:/Projects/Surveys/HHTravel/Survey2023/Documents/Survey_Instrument/Puget_Sound_HTS_Questionnaire_2023_Final.docx

Open the codebook and review the variable and it's values. J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx

# Data setup
```{r}
# somehow a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```


```{r}
# nrow(person) #25364
# nrow(person$ethnicity_999) #prefer not to answer: NULL
# nrow(person$race_noanswer) #NULL

table(person$age)
adult <- person %>%
  filter(age%in% c('18-24 years', '25-34 years', '35-44 years', '45-54 years', '55-64 years', '65-74 years', '75-84 years', '85 years or older'))
# table(adult$age)
# nrow(adult) #21492
# length(unique(adult$person_id)) #21492 - just to check

# check <- adult %>% 
#   filter(hh_id=="23226964")
unique(person$race_category)
table(person$race_category, person$age)

# unique(adult$race_other_specify) #checking 'other' responses
# length(unique(adult$race_other_specify)) #99 people who selected 'other' 

adult_23 <- adult %>% 
  filter(survey_year=="2023") #5931 - to know how many respondents total

# some_other <- adult_23  %>% 
#   filter(race_category=="Some Other Race non-Hispanic")
check_race_eth <- adult_23 %>%
  filter(ethnicity_999=="Selected",
         race_noanswer=="Selected")
```

# Person Race and Ethnicity
## Race
```{r}
# pivot table so that each row is a different race column
adult_pivot_race <- adult_23 %>% 
  pivot_longer(cols = c(race_afam, race_aiak, race_asian, race_hapi, race_noanswer, race_other, race_white),
               names_to = "race_group",
               values_to = "select")
# unique(adult_pivot_race$race_other_specify)
# table(adult_pivot_race$survey_year)

adult_pivot_race_select <- adult_pivot_race %>% 
  filter(select %in% 'Selected')

person_id_multiple_23_race <- adult_pivot_race_select %>% 
  group_by(person_id) %>% dplyr::summarise(n=n()) %>% #count how many racial groups were selected per person
  right_join(adult_pivot_race_select, by = ("person_id")) #join to main table so number connected to races selected

# nrow(person_id_multiple_23_race) #22056
table(person_id_multiple_23_race$n)
table1(~race_category, data=adult_23)
```

```{r, include = FALSE, eval=FALSE}
## before making function (for checking) ----
adult_1race <- person_id_multiple_23_race %>%
  filter(n %in% 1)
adult_2race <- person_id_multiple_23_race %>%
  filter(n %in% 2)
adult_3race <- person_id_multiple_23_race %>%
  filter(n %in% 3)
adult_4race <- person_id_multiple_23_race %>%
  filter(n %in% 4)
adult_5race <- person_id_multiple_23_race %>%
  filter(n %in% 5)
adult_6race <- person_id_multiple_23_race %>%
  filter(n %in% 6)

adult_2race_summary <- adult_2race %>%
  pivot_wider(names_from = race_group, values_from = n, values_fill = 0) %>%
  dplyr::select(race_asian:race_other) %>%
  group_by(across(race_asian:race_other)) %>% count %>% arrange(desc(freq)) %>%
  mutate(total=sum(freq)) %>%
  mutate(prop=(round(freq/total, digits = 3)))
```

```{r}
# create simplified table for combining with ethnicity data (eventually)
adult_race <- person_id_multiple_23_race %>%
  dplyr::select(person_id, race_group, n)

## after making function ----
race_list <- c('race_afam', 'race_aiak', 'race_asian', 'race_hapi', 'race_white', 'race_noanswer', 'race_other')

count_combination_race <- function(group, n_races){
  temp <- person_id_multiple_23_race %>% 
    filter(n %in% n_races)
  
  pivot <- temp %>% 
    pivot_wider(names_from = group, values_from = n, values_fill = 0) %>% 
    dplyr::select(any_of(race_list)) %>% 
    group_by(across(any_of(race_list))) %>% count %>% arrange(desc(n))
  
  output <- pivot %>% 
    mutate(total=sum(pivot$n)) %>% 
    mutate(prop=(round(n/total, digits = 3)))
}

adult_1race_summary <- count_combination_race('race_group', 1) #5655 people
adult_2race_summary <- count_combination_race('race_group', 2) #234 people 
adult_3race_summary <- count_combination_race('race_group', 3) #32 people
adult_4race_summary <- count_combination_race('race_group', 4) #2 people
# adult_5race_summary <- count_combination_race('race_group', 5) #0 people
# adult_6race_summary <- count_combination_race('race_group', 6) #0 people
```

```{r, echo=FALSE}
adult_1race_summary
```
1 RACE:

* 68.7%: white
* 15.5%: asian
* 11.1%: no answer
* 2.7%: african american/black
* 1.1%: 'Other'
* 0.5%: american indian/alaskan native
* 0.4%: hawaiian/pacific islander

```{r, echo=FALSE}
adult_2race_summary
```
2 RACES:

* 51.3%: asian AND white 
* 18.8%: american indian/alaskan native AND white
* 9%: african american/black AND white
* 6%: other AND white
* 5.1%: asian AND hawaiian/pacific islander
* 4.3%: hawaiian/pacific islander AND white
* 2.1%: african american/black AND white

```{r, echo=FALSE}
adult_3race_summary
```
3 RACES:

* 40.6%: african american/black AND american indian/alaskan native AND white
* 21.9: asian AND hawaiian/pacific islander AND white
* 9.4%: asian AND other AND white
* 9.4%: american indian/alaskan native AND asian AND white
* 9.4%: african american/black AND asian AND white


```{r, include=FALSE, eval=FALSE}
# testing numbers
race_afam <- person_id_multiple_23_race %>% 
  filter(n==1,
         race_group=="race_afam")

check <- person_id_multiple_23_race %>%
  filter(ethnicity_999=="Selected") #789
length(unique(check$person_id)) #776 

pivot <- check %>% 
    pivot_wider(names_from = race_group, values_from = n, values_fill = 0) %>%
  dplyr::select(any_of(race_list)) %>%
  group_by(across(any_of(race_list))) %>% 
    count %>% arrange(desc(n))

output <- pivot %>% 
    mutate(total=sum(pivot$n)) %>% 
    mutate(prop=(round(n/total, digits = 3)))
```


## Ethnicity
```{r}
# pivot table so that each row is a different ethnicity column
adult_pivot_ethnicity <- adult_23 %>% 
  pivot_longer(cols = c(ethnicity_1, ethnicity_2, ethnicity_3, ethnicity_4, ethnicity_997, ethnicity_999),
               names_to = "ethnicity_group",
               values_to = "select") #%>% 
  # mutate(eth_expand=case_when(ethnicity_group=="ethnicity_1"~"Not of Hispanic, Latino, or Spanish origin",
  #                             ethnicity_group=="ethnicity_2"~"Mexican, Mexican American, Chicano",
  #                             ethnicity_group=="ethnicity_3"~"Puerto Rican",
  #                             ethnicity_group=="ethnicity_4"~"Cuban",
  #                             ethnicity_group=="ethnicity_997"~"Another Hispanic, Latino, or Spanish origin",
  #                             ethnicity_group=="ethnicity_999"~"Prefer not to answer"))

# unique(adult$ethnicity_1)
table(adult$ethnicity_1, adult$ethnicity_2)
# unique(adult_pivot_ethnicity$ethnicity_other) #83
# unique(adult_pivot_ethnicity$survey_year) #83
# length(unique(adult_pivot_ethnicity$person_id)) #5931 - correct number of people
# table(adult_pivot_ethnicity$select)

adult_pivot_ethnicity_select <- adult_pivot_ethnicity %>%
  filter(select %in% 'Selected')

# length(unique(adult_pivot_ethnicity_select$person_id)) #5923 - 8 people missing


person_id_multiple_23_eth <- adult_pivot_ethnicity_select %>% 
  group_by(person_id) %>% dplyr::summarise(n=n()) %>% #count how many ethnicity groups were selected per person
  right_join(adult_pivot_ethnicity_select, by = ("person_id")) #join to main table so number connected to ethnicity selected

# length(unique(person_id_multiple_23_eth$person_id)) #5923 - 8 people missing
table(person_id_multiple_23_eth$n)

# unique(person_id_multiple_23_eth$ethnicity_other)
# length(unique(person_id_multiple_23_eth$person_id)) #5923 - 8 people missing

# where are 8 missing people?
check_join <- adult_pivot_ethnicity %>% 
  anti_join(adult_pivot_ethnicity_select, by="person_id") 

# length(unique(check_join$person_id)) #8
missing_ids <- unique(check_join$person_id)

missing_people <- adult_23 %>% 
  filter(person_id %in% missing_ids) %>%
  print
```
The 2023 adult dataset contains 5,931 people, but when looking at the ethnicity data, there were 5,923 people. The table above shows the 8 people who are missing - they didn't respond to the race or ethnicity questions?

```{r, include=FALSE, eval=FALSE}
## before making function (for checking) ----
adult_2eth <- person_id_multiple_23_eth %>%
  filter(n %in% 2)
adult_3eth <- person_id_multiple_23_eth %>%
  filter(n %in% 3)

adult_3eth_summary <- adult_3eth %>%
  pivot_wider(names_from = ethnicity_group, values_from = n, values_fill = 0) %>%
  dplyr::select(any_of(c('ethnicity_1', 'ethnicity_2', 'ethnicity_3', 'ethnicity_4', 'ethnicity_997', 'ethnicity_999'))) %>%
  group_by(across(any_of(c('ethnicity_1', 'ethnicity_2', 'ethnicity_3', 'ethnicity_4', 'ethnicity_997', 'ethnicity_999')))) %>% count %>% arrange(desc(n))

adult_2eth_summary <- adult_2eth %>%
  pivot_wider(names_from = ethnicity_group, values_from = n, values_fill = 0) %>%
  dplyr::select(any_of(c('Not of Hispanic, Latino, or Spanish origin',
                           'Mexican, Mexican American, Chicano',
                           'Puerto Rican',
                           'Cuban',
                           'Another Hispanic, Latino, or Spanish origin',
                           'Prefer not to answer'))) %>%
  group_by(across(any_of(c('Not of Hispanic, Latino, or Spanish origin',
                           'Mexican, Mexican American, Chicano',
                           'Puerto Rican',
                           'Cuban',
                           'Another Hispanic, Latino, or Spanish origin',
                           'Prefer not to answer')))) %>% count %>% arrange(desc(n))
```

```{r}
# create simplified table for combining with race data (eventually)
adult_eth <- person_id_multiple_23_eth %>% 
  dplyr::select(person_id, ethnicity_group, n)

## after making function ----
ethnicity_list <- c('ethnicity_1', 'ethnicity_2', 'ethnicity_3', 'ethnicity_4', 'ethnicity_997', 'ethnicity_999')

count_combination_eth <- function(group, n_ethnicities){
  temp <- person_id_multiple_23_eth %>% 
    filter(n %in% n_ethnicities)
  
  pivot <- temp %>% 
    pivot_wider(names_from = group, values_from = n, values_fill = 0) %>%
  dplyr::select(any_of(ethnicity_list)) %>%
  group_by(across(any_of(ethnicity_list))) %>% 
    count %>% arrange(desc(n))
  
  output <- pivot %>% 
    mutate(total=sum(pivot$n)) %>% 
    mutate(prop=(round(n/total, digits = 3)))
}

adult_1eth_summary <- count_combination_eth('ethnicity_group', 1) #5902 people
adult_2eth_summary <- count_combination_eth('ethnicity_group', 2) #19 people 
adult_3eth_summary <- count_combination_eth('ethnicity_group', 3) #2 people
# adult_4eth_summary <- count_combination_eth('ethnicity_group', 4) #0 people
# adult_5eth_summary <- count_combination_eth('ethnicity_group', 5) #0 people
# adult_6eth_summary <- count_combination_eth('ethnicity_group', 6) #0 people
```

```{r, echo=FALSE}
adult_1eth_summary
```
1 ETHNICITY:

* 81.3%: Not of Hispanic/Latino/Spanish origin (4801/5902)
* 13.1%: 'Prefer not to answer' (776/5902)
* 3.1%: Mexican, Mexican American, Chicano (182/5902)
* 1.9%: 'Another' (112/5902)
* 0.4%: Puerto Rican (22/5902)
* 0.2%: Cuban (9/5902)

```{r, echo=FALSE}
adult_2eth_summary
```
2 ETHNICITIES:

* 63.2%: Mexican, Mexican American, Chicano AND 'Another'
* 21.1%: Puerto Rican AND 'Another'
* 5.3%: Cuban AND 'Another'
* 5.3%: Puerto Rican AND Cuban
* 5.3%: Mexican, Mexican American, Chicano AND Cuban
* 0.2%: Cuban

```{r, echo=FALSE}
adult_3eth_summary
```
3 ETHNICITIES:

* 100%: Mexican, Mexican American, Chicano AND Puero Rican AND 'Another'

```{r, include=FALSE, eval=FALSE}
# test
eth_3 <- person_id_multiple_23_eth %>% 
  filter(n==1,
         ethnicity_group=="ethnicity_3")

table(person_id_multiple_23_eth$n) #5902 with 1 ethnicity
filter <- person_id_multiple_23_eth %>% 
  filter(n==1) #5902
filter2 <- filter %>% 
  filter(ethnicity_group=="ethnicity_1") #4801 people reporting 1 ethnicity are not hispanic/latino/Spanish origin
```


## Race + ethnicity combinations
### For recoding exercise
```{r}
# 1. generating table for re-coding exercise 
# nrow(adult_23) #5931

other_race <- adult_23 %>% 
  filter(race_other=="Selected") #133
other_race_specify <- adult_23 %>% 
  filter(!is.na(race_other_specify)) #133
other_eth_specify <- adult_23 %>% 
  filter(!is.na(ethnicity_other)) #132

other_race_eth <- adult_23 %>% 
  filter(!is.na(race_other_specify) |
           !is.na(ethnicity_other)) #226
# length(unique(other_race_eth$person_id)) #226 - just to make sure no duplicates

# for writing a data.frame or list of data.frames to an xlsx file
library(openxlsx)
# write.xlsx(other_race_eth, 'T:/2024July/Mary/HHTS/2023/adult_2023_race_eth_recode.xlsx')
```

### Understanding general distribution
```{r}
# create simplified table that 
race_temp <- adult_race %>% 
  group_by(person_id) %>% 
  dplyr::summarise(n_race=mean(n)) #5923 - making sure each row is a unique respondent
eth_temp <- adult_eth %>% 
  group_by(person_id) %>% 
  dplyr::summarise(n_eth=mean(n)) #5923 - making sure each row is a unique respondent

race_eth_temp <- race_temp %>% 
  right_join(eth_temp, by = "person_id") #5923 - making sure each row is a unique respondent

# checking to make sure that there are the correct numbers of respondents with 1,2,3 races/ethnicities
# table(race_eth_temp$n_race)

# join to original wide table with all race/ethnicity
adult_race_eth_count <- adult %>% 
  dplyr::filter(survey_year=="2023") %>% 
  dplyr::select_if(grepl("person_id", names(.)) | 
                     grepl("race_", names(.)) | 
                     grepl("ethnicity_", names(.))) %>% #5931
  right_join(race_eth_temp, by = "person_id")

# length(unique(adult_race_eth_count$person_id)) #5931 - making sure each row is a unique respondent

# detail_names <- c("Not of Hispanic, Latino, or Spanish origin"="ethnicity_1",
#                   "Mexican, Mexican American, Chicano"="ethnicity_2",
#                   "Puerto Rican"="ethnicity_3",
#                   "Cuban"="ethnicity_4",
#                   "Another ethnicity"="ethnicity_997",
#                   "Prefer not to answer"="ethnicity_999")
# rename(adult_race_eth_count, all_of(detail_names))

# table(adult_race_eth_count$n_race, adult_race_eth_count$n_eth)
table1(~ factor(n_eth) | as.factor(n_race), data=adult_race_eth_count) #formatted table
```

The table shows general trends and combinations

* 5,637/5,923 respondents : 1 race and 1 ethnicity (95.2%)
* 17/5,923 respondents : 1 race and 2 ethnicities (0.29%)
* 1/5,923 respondents : 1 race and 3 ethnicities (0%)
* etc. 

### Exploring different combinations
Based on the table, the vast majority of respondents chose 1 race and 1 ethnicity
```{r, include=FALSE, eval=FALSE}
# before making function (for checking) ----
race_eth_1 <- adult_race_eth_count %>%
  dplyr::filter(n_race==1 & n_eth==1) %>%
  dplyr::select(any_of(ethnicity_list)) %>%
  group_by(across(any_of(ethnicity_list))) %>%
    count %>% arrange(desc(n))
```

```{r}
## after making function ----
race_eth_combination <- function(num_race, num_eth, category_list){
  output <- adult_race_eth_count %>% 
    dplyr::filter(n_race==num_race & n_eth==num_eth) %>% 
    dplyr::select(any_of(ethnicity_list)) %>%
    group_by(across(any_of(ethnicity_list))) %>% 
    count %>% arrange(desc(n))
  
  final <- output %>% 
    mutate(total=sum(output$n)) %>% 
    mutate(prop=(round(n/total, digits = 3)))
}

race_1_eth_1 <- race_eth_combination(1,1,ethnicity_list) #5637
race_2_eth_1 <- race_eth_combination(2,1,ethnicity_list) #233
race_3_eth_1 <- race_eth_combination(3,1,ethnicity_list) #30
race_4_eth_1 <- race_eth_combination(4,1,ethnicity_list) #2
# race_5_eth_1 <- race_eth_combination(5,1,ethnicity_list) #0

race_1_eth_2 <- race_eth_combination(1,2,ethnicity_list) #17
race_1_eth_3 <- race_eth_combination(1,3,ethnicity_list) #1
# race_1_eth_4 <- race_eth_combination(1,4,ethnicity_list) #0
```
1 RACE/ 1 ETHNICITY: 5,637 respondents

* 81.3%: not Hispanic/Latino/Spanish origin
* 13.4%: 'Prefer not to answer'
* 3%: Mexican, Mexican American, Chicano
* 1.6%: 'Another'
* 0.3%: Puerto Rican
* 0.1%: Cuban

2 RACES/1 ETHNICITY: 233 respondents

* 85.4%: not Hispanic/Latino/Spanish origin
* 5.6%: 'Another'
* 5.2%: Mexican, Mexican American, Chicano
* 2.6%: 'Prefer not to answer'
* 0.9%: Puerto Rican
* 0.4%: Cuban

...

1 RACE / 2 ETHNICITIES: 17 respondents

* 64.7%: Mexican, Mexican American, Chicano AND 'Another'
* 17.6%: Puerto Rican AND 'Another'
* 5.9%: Cuban AND 'Another'
* 5.9%: Puerto Rican AND Cuban
* 5.9%: Mexican, Mexican American, Chicano AND Cuban

```{r, include=FALSE, eval=FALSE}
# before making function (for checking) ----
eth_1 <- person_id_multiple_23_eth %>%
  filter(n==1,
         ethnicity_group=="ethnicity_1") #4801
pivot <- eth_1 %>%
  pivot_wider(names_from = 'ethnicity_group', values_from = n, values_fill = 0) %>%
  dplyr::select(any_of(race_list)) %>%
  group_by(across(any_of(race_list))) %>% count %>% arrange(desc(n))
output <- pivot %>%
  mutate(total=sum(pivot$n)) %>%
  mutate(prop=(round(n/total, digits = 3)))

length(unique(eth_1$person_id))
```

```{r}
## after making functions ----
count_combination_race_by_ethnicitiy <- function(n_ethnicities, ethnicity_field, group){
  data <- person_id_multiple_23_eth %>% 
    filter(n==n_ethnicities,
           ethnicity_group==ethnicity_field)
  
  pivot <- data %>% 
    pivot_wider(names_from = group, values_from = n, values_fill = 0) %>% 
    dplyr::select(any_of(race_list)) %>% 
    group_by(across(any_of(race_list))) %>% count %>% arrange(desc(n))
  
  output <- pivot %>% 
    mutate(total=sum(pivot$n)) %>% 
    mutate(prop=(round(n/total, digits = 3)))
}

count_combination_race_by_ethnicitiy2 <- function(ethnicity_field, group){
  data <- person_id_multiple_23_eth %>% 
    filter(ethnicity_group!=ethnicity_field)
  
  pivot <- data %>% 
    pivot_wider(names_from = group, values_from = n, values_fill = 0) %>% 
    dplyr::select(any_of(race_list)) %>% 
    group_by(across(any_of(race_list))) %>% count %>% arrange(desc(n))
  
  output <- pivot %>% 
    mutate(total=sum(pivot$n)) %>% 
    mutate(prop=(round(n/total, digits = 3)))
}
```

#### Not Hispanic/Latino/Spanish Origin
```{r}
eth1_race <- count_combination_race_by_ethnicitiy(1, 'ethnicity_1', 'ethnicity_group')

eth1_race
```
1 ETHNICITY (Not Hispanic/Latino/Spanish origin):

* 74.2%: white (3562/4801)
* 16.4%: asian (786/4801)
* 2.7% african american (129/4801)
* 2.4% white AND asian
* 1.2% didn't provide an answer

```{r, number check, include=FALSE, eval=FALSE}
## CHECKING NUMBERS
check <- adult_23 %>%
  filter(ethnicity_1=="Selected",
         ethnicity_2=="Not selected",
         ethnicity_3=="Not selected",
         ethnicity_4=="Not selected",
         ethnicity_997=="Not selected",
         ethnicity_999=="Not selected") #4750 
length(unique(check$person_id))

check <- adult_23 %>%
  filter(ethnicity_1=="Selected",
         is.na(ethnicity_2),
         is.na(ethnicity_3),
         is.na(ethnicity_4),
         is.na(ethnicity_997),
         is.na(ethnicity_999)) #51 
length(unique(check$person_id))

# why are these numbers not lining up? something is wrong with the person_id_multiple_23_eth dataset?
check_join <- eth_1 %>% 
  anti_join(check, by="person_id") 

length(unique(check_join$person_id)) #51
missing_ids <- unique(check_join$person_id)

missing_people <- adult_23 %>% 
  filter(person_id %in% missing_ids)
# the missing 51 people - a recoding issue - NA vs Not Selected after 51 respondents were updated 
# the values have been updated 7/25, so all set
```

#### Hispanic/Latino/Spanish Origin
```{r}
not_eth1_race <- count_combination_race_by_ethnicitiy2('ethnicity_1', 'ethnicity_group') #1122

#check to make sure function works
# adult_23 #5931 people in dataset
# person_id_multiple_23_eth #5946 ethnicities (some people have multiple ethnicities)
# length(unique(adult_23$person_id)) #5931 people in dataset
# length(unique(person_id_multiple_23_race$person_id)) #5923
# length(unique(person_id_multiple_23_eth$person_id)) #5923
# 
# check <- person_id_multiple_23_eth %>% 
#   filter(ethnicity_group!="ethnicity_1") #1145
# length(unique(check$person_id)) #1122

not_eth1_race
```
1 ETHNICITY (Hispanic/Latino/Spanish origin):

* 51.1%: no answer (573/1122)
* 28.6%: white (321/1122)
* 7.9% asian (89/1122)
* 5.1% other 
* 2.1% african american/black
* 1.0%: white AND other
* 0.8%: american indian/alaskan native AND white
* 0.6%: african american/black AND american indian/alaskan native AND white
* 0.5%: asian AND white
* 0.4%: american indian/alaskan native

### Spot checking race category designations
```{r, include=FALSE, eval=FALSE}
table(adult_23$race_category)
table1(~race_category, data = adult_23)

no_missing_race_cat_2023_b <- adult_23 %>% 
  filter(race_category!="Missing/No response")
table1(~race_category, data = no_missing_race_cat_2023_b)

# spot check - random sample
check_df <- adult_23 %>% 
  dplyr::select(-c(hh_id, race_category, survey_year, person_weight, 
                   gender, age,  
                   race_other_specify, ethnicity_other)) %>% 
  dplyr::sample_frac(size=0.005) 

check_df[check_df== 'Not Selected' | #to account for race categories
           check_df== 'Not selected'] <- 0 #to account for ethnicity categories

# hapi_check <- adult_23 %>% 
#   dplyr::select(-c(hh_id, race_category, survey_year, person_weight, 
#                    gender, age, 
#                    race_other_specify, ethnicity_other)) %>%
#   dplyr::filter(race_hapi=="Selected" & race_asian=="Selected")
#   dplyr::sample_frac(size=0.5)


# ISSUES FOUND AFTER FIRST ROUND OF SPOT CHECKS -------------------------------------------
# case when asian and ethnicity_1 were selected but was grouped as Hispanic, person_id=2301887102
# case when aiak, white, and ethnicity_1 were selected but was grouped as white, person_id=2341004601
# case when afam, white, and ethnicity_1 were selected but was grouped as white, person_id=2333290001
# case when asian, white, and ethnicity_1 were selected but grouped as white, person_id=2310134001, 2306473501, 2316162403
# case when asian, hapi, and white were selected but grouped as white, person_id=2302756402, 2304251301, 2304947401

check_df_ids <- adult_23 %>% 
  dplyr::select(-c(hh_id, race_category, survey_year, person_weight, 
                   gender, age, race_category,  
                   race_other_specify, ethnicity_other)) %>% 
  dplyr::filter(person_id=="2301887102")

# SECOND CHECK -------------------------------------------
# 30 cases when race_noanswer and ethnicity_2 were selected and was grouped as Hispanic, person_id=2334480902, 2314196601, 2305926901, 2301957402, 2311501401
missing_norace_eth <- adult_23 %>% 
  filter(ethnicity_2=="Selected",
         race_noanswer=="Selected") %>% #30 select ethnicity_2, but do provide race
  dplyr::select(race_afam:race_white, starts_with("ethnicity"), race_category, person_id)
# 2 cases when race_noanswer and ethnicity_3 were selected and was grouped as Hispanic, person_id=2334480902, 2335237701 
# 2 cases when race_noanswer and ethnicity_997 were selected and was grouped as Hispanic, person_id=2334480902, 2315385501   
missing_norace_eth <- adult_23 %>% 
  filter(ethnicity_1!="Selected",
         ethnicity_999!="Selected",
         race_noanswer=="Selected") %>% #30 select ethnicity_2, but do provide race
  dplyr::select(race_afam:race_white, starts_with("ethnicity"), race_category, person_id)

# THIRD CHECK -------------------------------------------
# case when race_white and ethnicity_999 were selected and was grouped as missing/no response, but should be white? person_id=2302993902, 2334376002, 2324371402, 2316054601, 2324640701
white_999 <- adult_23 %>% 
  filter(race_white=="Selected",
         ethnicity_999=="Selected") %>% #124
  dplyr::select(race_afam:race_white, starts_with("ethnicity"), race_category, person_id)

# 838 missing/no response ---
# looking into overlap between `race_noanswer` and `ethnicity_999`
# how many people identify as ethnicity_999, but provide a race? like afam (2304254201) or asian (2326142801)
missing_eth <- adult_23 %>% 
  filter(ethnicity_999=="Selected", #776
         race_noanswer!="Selected") %>% #235 don't provide ethnicity, but do provide race
  dplyr::select(race_afam:race_white, race_category, person_id)

# how many of that group was classified as "Missing/No response" but had a race chosen?
missing_eth_999 <- missing_eth %>% 
  filter(race_category=="Missing/No response") #119
table(missing_eth_999$race_white) #119

# looking into each race -----
# table1(~race_afam, data=missing_eth) #13 afam
missing_eth_afam <- adult_23 %>% 
  filter(ethnicity_999=="Selected", #776
         race_afam=="Selected") %>% #13
  dplyr::select(race_afam:race_white, race_category, person_id, ethnicity_999)
# person_id = 2322282101 selected ethnicity_999, but was categorized as 2+ races (selected afam, aiak, white) 

# table1(~race_asian, data=missing_eth) #88 asian
missing_eth_asian <- adult_23 %>% 
  filter(ethnicity_999=="Selected", #776
         race_asian=="Selected") %>% #88
  dplyr::select(race_afam:race_white, race_category, person_id, ethnicity_999)
# person_id = 2315970702 selected ethnicity_999, but was categorized as 2+ races (asian, hapi, other)

# table1(~race_aiak, data=missing_eth) #3 aiak
missing_eth_aiak <- adult_23 %>% 
  filter(ethnicity_999=="Selected", #776
         race_aiak=="Selected") %>% #3
  dplyr::select(race_afam:race_white, race_category, person_id, ethnicity_999)
# person_id = 2322282101 selected ethnicity_999, but was categorized as 2+ races (selected afam, aiak, white) 

# table1(~race_hapi, data=missing_eth) #5 hapi
missing_eth_hapi <- adult_23 %>% 
  filter(ethnicity_999=="Selected", #776
         race_hapi=="Selected") %>% #5
  dplyr::select(race_afam:race_white, race_category, person_id, ethnicity_999)
# person_id = 2315970702 selected ethnicity_999, but was categorized as 2+ races (asian, hapi, other)

# table1(~race_white, data=missing_eth) #124 white
missing_eth_white <- adult_23 %>% 
  filter(ethnicity_999=="Selected", #776
         race_white=="Selected") %>% #124
  dplyr::select(race_afam:race_white, race_category, person_id, ethnicity_999)
# person_id = 2322282101 selected ethnicity_999, but was categorized as 2+ races (selected afam, aiak, white) 
# person_id = 2304197102 selected white and ethnicity_999, but was categorized as other

# table1(~race_other, data=missing_eth) #8 other
missing_eth_other <- adult_23 %>% 
  filter(ethnicity_999=="Selected", #776
         race_other=="Selected") %>% #8
  dplyr::select(race_afam:race_white, race_category, person_id, ethnicity_999)
# person_id = 2315970702 selected ethnicity_999, but was categorized as 2+ races (asian, hapi, other)

table1(~race_category, data=missing_eth) # mistake - shouldn't have any race categories other than 'missing/no response' because ethnicity_999 was selected for this subgroup of 235 people

# how many people identify as race_noanswer but provide ethnicity
missing_race <- adult_23 %>% 
  filter(race_noanswer=="Selected", #630
         ethnicity_999!="Selected") #89 provide ethnicity, 541 choose no to provide 

# how many of these 89 people identify as hispanic/latino/spanish origin (not ethnicity_1)
missing_race_hispanic <- missing_race %>% 
  filter(ethnicity_1!= "Selected") #32, 57 chose NOT hispanic/latino/spanish origin

# how many people identify as race_noanswer and ethnicity_999
missing_both <- adult_23 %>% 
  filter(ethnicity_999=="Selected",
         race_noanswer=="Selected") #541

# table1(~race_category, data=missing_both) #to check - result as expected
```

# race_category_2023_c > race_category
## Person Race > Household Race 
### By size
The new person race variable: 'race_category_2023' and 'race_category_2023_b'
```{r}
# table(adult_23$race_category)
adult_23_simp <- adult_23 %>% 
  select(hh_id, person_id, race_category, age) 

hh_id_multiple_23_prace <- adult_23_simp %>% 
  group_by(hh_id) %>% dplyr::summarise(n_members=n()) #count how many people per hh - total hh is 3661, total people is 5931
```
total households: 3,661
total people: 5,931
```{r}
# share of households by number of people
table1(~as.factor(n_members), data=hh_id_multiple_23_prace) 
```

```{r}
# full data set (each row is different person) includes number of people in household
combined <- adult_23_simp %>% 
  right_join(hh_id_multiple_23_prace, by = ("hh_id"))

updated_race_list <- (unique(combined$race_category))
```

```{r, include=FALSE, eval=FALSE}
# before making function (for checking) ----
# select households with two people
hh_2people <- combined %>% 
  filter(n_members==2)

hh_2people_wider <- combined %>% 
  filter(n_members==2) %>% 
  pivot_wider(names_from = race_category, values_from = n_members, values_fill = 0) %>% 
  group_by(hh_id) %>% dplyr::summarise(across(any_of(updated_race_list), list(sum = sum))) %>%
  rename_with(~ sub("_sum$", "", .x), everything()) %>% #remove suffix 'sum' 
  group_by(across(any_of(updated_race_list))) %>% count %>% arrange(desc(n)) %>% 
  mutate(total=sum(pivot$n)) %>%
  mutate(prop=(round(n/total, digits = 3)))

hh_2people_wider[,1:ncol(hh_2people_wider)]/2
```

```{r}
## after making function ----
count_combination_prace <- function(n){
  temp <- combined %>% 
    filter(n_members==n) 
  
  pivot <- temp %>% 
    pivot_wider(names_from = race_category, values_from = n_members, values_fill = 0) %>% 
    group_by(hh_id) %>% dplyr::summarise(across(any_of(updated_race_list), list(sum = sum))) %>%
    rename_with(~ sub("_sum$", "", .x), everything()) %>% #remove suffix 'sum' 
    group_by(across(any_of(updated_race_list))) %>% count %>% arrange(desc(n)) 
  
  output <- pivot %>% 
    mutate(total=sum(pivot$n),
           prop=(round(n/total, digits = 3)))
  }

hh1 <- count_combination_prace(1)
hh2 <- count_combination_prace(2)
hh3 <- count_combination_prace(3)
hh4 <- count_combination_prace(4)
hh5 <- count_combination_prace(5)
hh6 <- count_combination_prace(6)
```

#### 1-Person Households
```{r}
hh1
```

* 60%: White alone non-Hispanic
* 13%: Asian alone non-Hispanic
* 12%: Missing/No response
* 7%: Hispanic

#### 2-Person Households
```{r}
hh2
```

#### 3-Person Households
```{r}
hh3
```

#### 4-Person Households
```{r}
hh4
```

#### 5-Person Households
```{r}
hh5
```

#### 6-Person Households
```{r}
hh6
```




### By race
```{r}
combined
updated_race_list

combined_race_number <- combined %>% 
  dplyr::mutate(race_num=case_when(race_category=="White non-Hispanic"~1,
                                   race_category=="Hispanic"~2,
                                   race_category=="Black or African American non-Hispanic"~3,
                                   race_category=="Asian non-Hispanic"~4,
                                   race_category=="Some Other Race non-Hispanic"~5,
                                   race_category=="Two or More Races non-Hispanic"~6,
                                   race_category=="Missing/No response"~7))

hh_race_num <- combined_race_number %>% 
  group_by(hh_id) %>% 
  dplyr::summarise(product=prod(race_num)) 

join <- hh_id_multiple_23_prace%>% 
  right_join(hh_race_num, by="hh_id")%>% 
  mutate(hh_grouping=case_when(product==1~"white hh",
                               product>1~"other")) #3661
table1(~hh_grouping|n_members, data=join)
```


#### All people
```{r}
all_people <- combined %>% 
  dplyr::mutate(person_cat=case_when(race_category=="Black or African American non-Hispanic"~"African American",
                                 race_category=="Asian non-Hispanic"~"AANHPI",
                                 race_category=="Hispanic"~"Hispanic",
                                 race_category=="Missing/No response"~"Missing",
                                 race_category=="White non-Hispanic"~"White",
                                 race_category=="Some Other Race non-Hispanic" |
                                   race_category=="Two or More Races non-Hispanic"~"Other"))

table1(~person_cat, data=all_people)

unique(all_people$person_cat)
all_people_nomissing <- all_people %>% 
  filter(person_cat!="Missing")

table1(~person_cat, data=all_people_nomissing)

```

#### All white households (person-level)
```{r}
join_justwhite <- join %>% 
  filter(hh_grouping=="white hh") %>% #2049 households
  inner_join(combined_race_number, by="hh_id") #3192 person

# length(unique(join_justwhite$hh_id)) #2049 to check
table1(~as.factor(n_members.x), data=join_justwhite)
```

#### Other households
##### One-person hh, non-white (person-level)
```{r}
one_person_hh_nonwhite <- join %>% 
  filter(n_members==1,
         hh_grouping == "other") %>% #652
  left_join(combined_race_number, by="hh_id") #652
  
table1(~race_category, data=one_person_hh_nonwhite)
```

##### More than one-person hh, not just white
```{r}
other_hh <- join %>% 
  filter(n_members>1,
         hh_grouping=="other") #960 households
# length(unique(other_hh$hh_id)) #960 to check

other_hh_race <- other_hh %>% 
  inner_join(combined_race_number, by="hh_id") #2202 person

# checks ---
length(unique(other_hh_race$hh_id)) #960 - confirming that this is the correct number of hh (out of all 3661 households, 2049 hh are just white, 652 are one person-other race hh)
# table(other_hh$hh_grouping) #just to check

# test_hh <- other_hh_race %>% 
#   filter(hh_id==23000858 |
#            hh_id==23001496 |
#            hh_id==23001700)

pivot <- other_hh_race %>% 
  pivot_wider(names_from = race_category, values_from = n_members.x, values_fill = 0) %>% 
  group_by(hh_id, n_members.y) %>% 
  dplyr::summarise(across(any_of(updated_race_list), list(sum = sum))) %>% 
  rename_with(~ sub("_sum$", "", .x), everything()) #remove suffix 'sum'

twoplus_person_notjustwhite <- pivot %>% 
  group_by(hh_id, n_members.y) %>% 
  dplyr::summarise(across(any_of(updated_race_list), ~.x/n_members.y)) %>% ungroup()
```
#### Household combinations
##### All combinations (with hh size)
```{r}
#to look at distribution
library(plyr) #for count/pivot function
library(dplyr)
hh_comp_2plus_count <- count(twoplus_person_notjustwhite[, -1])

hh_comp_2plus <- hh_comp_2plus_count %>% 
  arrange(desc(freq)) %>% 
  mutate(total=sum(hh_comp_2plus_count$freq)) %>% 
  mutate(prop=(round(freq/total, digits = 3)))

hh_comp_2plus

# check
# sum(hh_comp_2plus$n) #939
```

##### All combinations (no hh size)
```{r}
columns <- colnames(hh_comp_2plus[1:7])
grouping_hh <- function(x) {
  ifelse(x >0, 1, 0)}

hh_comp_2plus_groups <- hh_comp_2plus %>% mutate_at(c(columns), grouping_hh)

hh_comp_2plus_simp <- hh_comp_2plus_groups[2:9]

hh_comp_2plus_race_groups <- hh_comp_2plus_simp %>% 
  count(1:7, wt= 'freq') %>% #https://dplyr.tidyverse.org/reference/count.html
  arrange(desc(freq)) %>% 
  mutate(total=sum(hh_comp_2plus_simp$freq)) %>% 
  mutate(prop=(round(freq/total, digits = 3)))

hh_comp_2plus_race_groups
# check
# sum(hh_comp_2plus_race_groups$freq) #939
```


## Household Race/Ethnicity Categorization
The categories and logic are based on documentation from 2020: https://www.psrc.org/media/3205

### All white households, regardless of size
```{r}
all_hh_justwhite_categories <- join_justwhite %>% 
  distinct(hh_id) %>% #keep unique hh id - 2049 households
  dplyr::mutate(hh_cat="White") %>% #add field to match other dfs
  inner_join(join_justwhite, by="hh_id", multiple="first") #join to base data to keep all fields

# all_hh_justwhite_categories <- all_person_justwhite_categories %>% 
  
```


### Non-white, 1 person households
```{r}
# unique(one_person_hh_nonwhite$race_category)
# table(one_person_hh_nonwhite$race_category)

oneperson_hh_nonwhite_categories <- one_person_hh_nonwhite %>% 
  dplyr::mutate(hh_cat=case_when(race_category=="Black or African American non-Hispanic"~"African American",
                                 race_category=="Asian non-Hispanic"~"AANHPI",
                                 race_category=="Hispanic"~"Hispanic",
                                 race_category=="Missing/No response"~"Missing",
                                 race_category=="Some Other Race non-Hispanic" |
                                   race_category=="Two or More Races non-Hispanic"~"Other"))

# to check
# table(oneperson_hh_nonwhite_categories$hh_cat)
# nrow(oneperson_hh_nonwhite_categories) #652 - households
# length(unique(oneperson_hh_nonwhite_categories$hh_id)) #652 - households
```

### Non-white only, 2+ households
```{r}
twoplus_person_nonwhite <- twoplus_person_notjustwhite %>% 
  dplyr::mutate(hh_cat=case_when(`Missing/No response`>0~"Missing",
                                 Hispanic>0 |
                                   Hispanic>0 & 
                                   `White non-Hispanic`>0~"Hispanic",
                                 `Black or African American non-Hispanic` >0 |
                                   `Black or African American non-Hispanic` >0 &
                                   `White non-Hispanic`>0~"African American",
                                 `Asian non-Hispanic`>0 |
                                   `Asian non-Hispanic`>0 &
                                   `White non-Hispanic`>0 ~"AANHPI",
                                 `Some Other Race non-Hispanic`>0 | 
                                   `Two or More Races non-Hispanic`>0 ~ "Other")) %>% 
  dplyr::select(hh_id, hh_cat)

length(unique(twoplus_person_nonwhite$hh_id)) #960 - households
nrow(twoplus_person_nonwhite) #960 - persons


# getting data in same format as other data sets
colnames(twoplus_person_nonwhite)
colnames(other_hh_race)

twoplus_hh_nonwhite_categories <- twoplus_person_nonwhite %>% 
  inner_join(other_hh_race, by=c('hh_id'), multiple="first") #join to base data to keep all fields but at hh level - 960
```

### All households
```{r}
colnames(all_hh_justwhite_categories) #2049
colnames(oneperson_hh_nonwhite_categories) #652
colnames(twoplus_hh_nonwhite_categories) #960


all_hh_extra <- rbind(all_hh_justwhite_categories, #2049
                      oneperson_hh_nonwhite_categories, #652
                      twoplus_hh_nonwhite_categories) #960

all_hh <- all_hh_extra %>%
  dplyr::select(hh_id:hh_grouping)

#check
# table(all_hh$hh_grouping) #2049 white
# table(all_hh$hh_cat) #2049 white
```
#### With Missing
```{r}
table1(~ hh_cat, data=all_hh)
table1(~ hh_cat| n_members.x, data=all_hh)
# table1(~ hh_cat| race_category, data=all_hh)
```

#### Without Missing
```{r}
all_hh_nomissing <- all_hh %>% 
  dplyr::filter(hh_cat!="Missing") #4843


table1(~ hh_cat, data=all_hh_nomissing)
table1(~ hh_cat| n_members.x, data=all_hh_nomissing)
# table1(~ hh_cat| race_category_c, data=all_hh_nomissing)
```

### Spot checking hh race category designations
```{r}
# need to look at household view for 2023 data
hh_2023 <- hh %>% 
  dplyr::filter(survey_year=="2023")

table1(~hh_race_category, data = hh_2023)
table1(~ hh_cat, data=all_hh) #checking against my manually coded categorizations
# Both tables include 3,661 which is a good initial check
# Missing and Non-hispanic white include the same numbers: 434 and 2049
# African American, Asian, Hispanic, and Other values are slightly off



# check
hispanic_1hh_check <- oneperson_hh_nonwhite_categories %>% 
  dplyr::filter(hh_cat=="Hispanic") #117
hispanic_hh_check <- twoplus_person_notjustwhite %>% 
  dplyr::filter(Hispanic>0) #171


# check rows
# join hh data with person level data - one person hh and 2+ person hh
hh_1person_2023 <- one_person_hh_nonwhite %>% 
  left_join(hh_2023, by="hh_id") #652
hh_2person_2023 <- twoplus_person_notjustwhite %>% 
  left_join(hh_2023, by="hh_id") #960
# correct number of columns when added with just white

# check 1 person hh - make sure the race_category aligns with the hh raca
table(hh_1person_2023$race_category, hh_1person_2023$hh_race_category)

# check 2 person hh - make sure the race_categories align with the hh_race_category
check_df_2person <- hh_2person_2023 %>% 
  dplyr::select(c(-survey_year, -n_members.y)) %>%
  dplyr::sample_frac(size=0.04) 

check_df_2person
```

### Weighted hh_race_category
```{r}
table1(~hh_race_category, data=hh_2023)

# RSG function
summary_hh_race <-  hts_summary(
  prepped_dt = hh_2023,
  summarize_var = "hh_race_category",
  summarize_vartype = "categorical",
  wtname = "hh_weight",
  weighted = TRUE,
  se=TRUE)

hts_data=list(hh=hh)
hh_race<- summarize_weighted(hts_data= hts_data,
                               summarize_var = 'hh_race_category',
                               summarize_by = 'survey_year',
                               id_cols=c('hh_id'),
                               wt_cols=c('hh_weight'),
                               wtname= 'hh_weight'
                               )


# using PSRC package -- not working
hts_data <- get_psrc_hts(survey_vars = c("hh_race_category"))  # default includes all survey_years
# Calculate a categorical summary, i.e. count/share
rs2 <- psrc_hts_stat(hts_data, "household", c("hh_race_category"))

dplyr::select(rs2[survey_year==2023], 
                   -c(survey_year))
```