---
title: "trip modes by county'
author: "suzanne"
date: "2024-08-30"
output: html_document
---

Data request for the light rail opening.
Initial request
•	How many Snohomish County residents travel to Seattle (anywhere in the city) each day (average weekday)? Also, how many trips is that?
•	Of these trips, what are the shares for driving, transit, and other modes?

Data from LEHD on the first question:

 
Of the 602,400 people who worked in the City of Seattle in 2021:
•	12.5% (75,300) live in Snohomish County, the largest share outside of King County
•	Pierce County is next at 6.6% (39,800), or half the proportion of Snohomish County
•	Kitsap County is third at 2.6% (15,700)


This request is mainly focused on modes -
Summary to do home county to Seattle trip mode share
for all trips, for work and work-related (not requested but I'll get it anyway)

```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
library(psrccensus)
source('survey-23-preprocess.R')
```

YOU MIGHT need to download some special packages on github: travelSurveyTools, psrcelmer, and psrcplot. IF you need to download them, here's how:

library(devtools)

devtools::install_github('rsgInc/travelSurveyTools') devtools::install_github('psrc/psrcelmer')

devtools::install_github('psrc/psrcplot')

## Read in Codebook

```{r , echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```

variable_list and value_labels will be used by the travelSurveyTools package whenever you summarize.

## Read in Data from Elmer

The 2017, 2019, 2021, and 2023 travel survey data is in Elmer. You will need to select exactly which columns you want to summarize. Note that the variables and values you pull in must match the codebook. (if for some reason in error, there is inconsistency we have to fix it).

The package uses DT and data.table for most operation, so you will usually want inputs to be in data.table format.


```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad,home_jurisdiction, home_county, hh_weight,vehicle_count, home_rgcname, home_lat, home_lng, survey_year, hh_race_category
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category, walk_freq, bike_freq, transit_freq,
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, sexuality, industry,office_available, commute_freq, education, work_lat, work_lng , work_rgcname, work_jurisdiction, work_county, share_1, share_2, share_3, share_4, share_5
                from HHSurvey.v_persons_labels")
                
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,
person_id, mode_1, dest_purpose, survey_year, trip_weight, distance_miles, mode_characterization, travelers_total, dest_jurisdiction, dest_county 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)
setDT(day)
setDT(trip)
```

## Set IDs as characters

We should fix this by writing a function. But for now set all ids to characters for easy joining.

```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
day[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
day[, person_id := as.character(person_id)]
trip[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
trip[, day_id := as.character(day_id)]

trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
day <- day%>%mutate(survey_year=as.character(survey_year))
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```


```{r}
variable_list<-add_variable(variable_list, 'mode_char1','trip')
group_labels<-get_grouped_labels(group_id='group_1', group_name='mode_char1')
value_labels<-add_values_code(group_name='mode_char1')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='mode_characterization',grouped_name='mode_char1')
```

```{r}
hh_23<-hh%>%filter(survey_year=='2023')
person_23<-person%>%filter(survey_year=='2023')
day_23<-day%>%filter(survey_year=='2023')
trip_23<-trip%>%filter(survey_year=='2023')
trip_23<-trip_23%>%filter(mode_characterization!='Airplane')

ids = c('hh_id', 'person_id', 'day_id', 'trip_id')
wts = c('hh_weight', 'person_weight', 'day_weight', 'trip_weight')
```




```{r}
hh_data_23<-
  list(hh=hh_23, person=person_23, day=day_23, trip=trip_23)
```


```{r}
output <- summarize_weighted(hts_data= hh_data_23,
                               summarize_var = 'mode_char1',
                               summarize_by = c('home_county'),
                               id_cols=ids,
                               wt_cols=wts,
                               wtname= 'trip_weight'
                               )
```
```{r}
est_mode<-output$summary$wtd%>%group_by(home_county, mode_char1)%>%summarize(est=sum(est))%>%pivot_wider(names_from= home_county, values_from=est)
```


```{r}
write.csv(est_mode, 'est_mode.csv')
```






```{r}
mode_to_work<-get_acs_recs(geography = 'county',
             table.names = 'B08301',
             years = 2022,
             acs.type = 'acs5')
```

do not include worked from home

```{r}
mode_to_work_summary<-mode_to_work%>%filter(label %in% c(	
'Estimate!!Total:!!Car, truck, or van:!!Drove alone', 'Estimate!!Total:!!Car, truck, or van:!!Carpooled:','Estimate!!Total:!!Public transportation (excluding taxicab):
', 'Estimate!!Total:!!Taxicab','Estimate!!Total:!!Motorcycle','Estimate!!Total:!!Bicycle',
'Estimate!!Total:!!Walked','Estimate!!Total:!!Other means'))%>%
  mutate(commute_mode= str_replace(label,'Estimate!!Total:!!',''))

```

```{r}
mode_to_work_wide<-mode_to_work_summary%>%group_by(name, commute_mode)%>%summarize(estimate=sum(estimate))%>%mutate(commute_mode1=if_else(commute_mode %in% c('Motorcycle', 'Taxicab') , 'Car, truck, or van:!!Drove alone', commute_mode))%>%group_by(name,commute_mode1)%>%summarize(estimate=sum(estimate))%>%pivot_wider(names_from=name, values_from=estimate)
                                                                                                          
```

```{r}
write.csv(mode_to_work_wide, 'mode_to_work_wide.csv')
```

