---
title: '2021 PSRC HTS: Unweighted Frequency Tables by Sample Source'
# editor_options:
#   chunk_output_type: console
author: PSRC
date: "Documentation last compiled on `r Sys.Date()`"
output:
  html_document:
    collapsed: yes
    self_contained: yes
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
    toc_depth: '4'
params:
  codebook_path: J:/Projects/Surveys/HHTravel/Survey2021/Data/data_from_rsg_20210803/PSRC_2021_HTS_Combined_Codebook.xlsx
  tab_by: sample_source
  tables: [hts_hh, hts_vehicle, hts_person, hts_day, hts_trip]
  variables_to_exclude: ["hhid", "personid", "tripid", "loc_start_other", "loc_end_other", "arrival_time_hhmm", "depart_time_hhmm", "arrival_time_mam", "depart_time_mam", "final_home_puma10", "survey_year", "hhmember1", "hhmember2", "hhmember3", "hhmember4", "hhmember5", "hhmember6", "hhmember7", "hhmember8", "hhmember9", "hhmember10", "hhmember11", "hhmember12"]
  missing_values: !r c(994, 995, -9998, -9999)
  weight_column: 
  study_name: "PSRC HTS 2021"
---

```{r psrc functions, include = FALSE}
# connecting to db
db.connect <- function() {
  db_connection <- dbConnect(odbc(),
                                driver = "SQL Server",
                                server = "AWS-PROD-SQL\\Sockeye",
                                database = "HouseholdTravelSurvey2023",
                                trusted_connection = "yes"
  )
}

# a function to read tables and queries from Elmer
read.dt <- function(astring, type =c('table_name', 'sqlquery')) {
  elmer_connection <- db.connect()
  if (type == 'table_name') {
    dtelm <- dbReadTable(db_connection, SQL(astring))
  } else {
    dtelm <- dbGetQuery(db_connection, SQL(astring))
  }
  dbDisconnect(elmer_connection)
  dtelm
}

two_vars_by_hhgroup= function(dataset,col_name,var_string){
  
  col_name = enquo(col_name)
  
  
  trips_by_var = dataset %>% 
  group_by(hhgroup.x,( !!col_name), hh_is_complete_b) %>% 
  summarise (n_trips = n_distinct(tripid)) 
  
  browser_mode = trips_by_mode %>% 
    filter(hhgroup %in% c(3, 4, 5, 6,7,8)) %>% 
    rename (browser_n_trips = n_trips) %>% 
    mutate(browser_trips_share = round(browser_n_trips/sum(browser_n_trips)*100,1))
    
    
  smartphone_mode = trips_by_mode %>% 
    filter(hhgroup %in% c(9, 10,11))%>%
    group_by(hh_is_complete_b)%>%
    rename (smartphone_n_trips = n_trips) %>% 
    mutate(smartphone_trips_share = round(panel_n_trips/sum(panel_n_trips)*100,1))
  
  temp = ABS_mode %>% 
    full_join(panel_mode,  by = var_string) %>% 
    dplyr::select(-c(sample_source.x.x,sample_source.x.y)) %>% 
    rename( "value" := !!col_name)
   # mutate(value = !!col_name)
  
  var_label_table = as.data.frame(value_labels[value_labels$variable %in% c(var_string)] %>% mutate(value = as.double(value))) %>% 
    dplyr::select(value, value_label)
  
  #browser()
  
  temp = temp %>% left_join(var_label_table, by = "value") %>% 
    dplyr::select(-c(value)) %>%
    relocate(value_label) %>% 
    arrange(desc(ABS_trips_share))

  return(temp)
  
}


get_labels = function(dataset,col_name,var_string){
  col_name = enquo(col_name)
  
  var_label_table = as.data.frame(value_labels[value_labels$variable %in% c(var_string)] %>% mutate(value = as.double(value))) %>% 
    dplyr::select(value, value_label)
  
  dataset = dataset %>%
    rename( "value" := !!col_name)
  
  temp = dataset %>% left_join(var_label_table, by = "value") %>% 
    dplyr::select(-c(value)) %>%
    relocate(value_label) 

  return(temp)
  
}
```

```{r setup, include = FALSE, echo = FALSE, message = FALSE, warning = FALSE, results = 'asis'}

# Update codebook path above & data input paths below if running outside of RSG!

options(scipen = 99)
options(knitr.kable.NA = '')
library(data.table)
library(knitr)
library(kableExtra)
library(readxl)
library(lubridate)
#library(RPostgres)
library(stringr)
library(ggplot2)
library(odbc)
library(DBI)
library(tidyverse)
source('C:/Users/schildress/Documents/GitHub/travel-survey-rsg-psrc-2021/rsg_qaqc/_crosstabs_functions.R')

```

------------------------------------------------------------------------

First we will look at the UNCLEANED data by itself, then the CLEANED data, and then together compare the two.

# Uncleaned data

```{r get_inputs, include = FALSE, echo = FALSE, cache = FALSE}

# UPDATE folder path if running outside of RSG
# may also need to update file formats/read functions (e.g., read from a database)
elmer_connection <- dbConnect(odbc::odbc(),
                              driver = "SQL Server",
                              server = "AWS-PROD-SQL\\Sockeye",
                              database = "hhts_cleaning",
                              trusted_connection = "yes")
h <- dbGetQuery(elmer_connection,
                "SELECT * FROM dbo.hts_hh")
p <- dbGetQuery(elmer_connection,
                "SELECT * FROM dbo.hts_person")

d <- dbGetQuery(elmer_connection,
                "SELECT * FROM dbo.hts_day")

t <- dbGetQuery(elmer_connection,
                "SELECT * FROM dbo.hts_trip")
```

```{r varvals, cache = TRUE, echo = FALSE, message = FALSE}

variable_labels = 
  read_codebook(
    params$codebook_path,
    varvals = FALSE,
    sheet = 'Variable_List')

setnames(variable_labels, 'label', 'description', skip_absent = TRUE)

value_labels = 
    read_codebook(
    params$codebook_path,
    varvals = TRUE, 
    sheet = 'Value_Labels')

setnames(value_labels, 'label', 'value_label', skip_absent = TRUE)

variable_labels[, common_name := gsub(':.*', '', description)]

value_labels[, value := as.character(value)]

# Change checkbox logic to be the same for each variable

variable_labels[variable %like% 'delivery_', logic := '']
variable_labels[variable %like% 'no_travel_', logic := 'If made no trips on travel day']
variable_labels[variable %like% 'ethnicity_', logic := 'If related adult household member']
variable_labels[variable %like% 'race_', logic := 'If related adult household member']


```

```{r join trips and person, include = FALSE ,message=FALSE,warning=FALSE}
trips_person = left_join(t, p, by = "personid")

trips_person_under17 = trips_person %>% 
                        filter(age < 5)

trips_hh = left_join(t,h, by = "hhid")
```

------------------------------------------------------------------------

### average trips per person by panel vs ABS

Here, average is calculated for people who made at least one trip (people who are present in trip table )

```{r, echo=FALSE,message=FALSE,warning=FALSE}


var_label_table = as.data.frame(value_labels[value_labels$variable %in% c("sample_source")] %>% mutate(value = as.double(value)))


temp = trips_person %>% 
  group_by(sample_source.x) %>% 
  summarise (n_trips = n_distinct(tripid), n_people = n_distinct(personid)) %>% 
  mutate(average_trips_per_person = n_trips*1.0/n_people) %>% 
  left_join(var_label_table, by = c("sample_source.x" = "value")) %>% 
  dplyr::select(-c(variable, sample_source.x , label_value, val_order)) %>% 
  relocate(value_label)

avg_trips_uncleaned <- temp

  temp %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))



```

Here, average is calculated for all people (including people that didnt do any trips)

```{r, echo=FALSE,message=FALSE,warning=FALSE}
day_panel_pernum1 = d %>% filter(sample_source == 2, pernum == 1)

var_label_table = as.data.frame(value_labels[value_labels$variable %in% c("sample_source")] %>% mutate(value = as.double(value)))

person_panel = subset(p, (personid %in% day_panel_pernum1$personid)) 

person_panel_summary = person_panel %>% 
    group_by(sample_source) %>%  summarise(ppl_count = n())

ppl_ABS = p %>% 
          filter(sample_source == 1)%>% 
    group_by(sample_source) %>%  summarise(ppl_count = n())

all_ppl = bind_rows(ppl_ABS,person_panel_summary)


temp = t %>% 
  group_by(sample_source) %>% 
  summarise (n_trips = n_distinct(tripid)) %>% 
  full_join(all_ppl, by = "sample_source") %>% 
  mutate(average_trips_per_person = n_trips*1.0/ppl_count) %>% 
  left_join(var_label_table, by = c("sample_source" = "value")) %>% 
  dplyr::select(-c(variable, sample_source , label_value, val_order)) %>% 
  relocate(value_label)
 
avg_trips_all_uncleaned<-temp
temp %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))



```

------------------------------------------------------------------------

### trips by mode (mode_1) by panel vs ABS; and overall

```{r,echo=FALSE,message=FALSE,warning=FALSE}


trips_by_mode = trips_person %>% 
  group_by(sample_source.x, mode_1) %>% 
  summarise (n_trips = n_distinct(tripid)) 

ABS_mode = trips_by_mode %>% 
  filter(sample_source.x == 1) %>% 
  rename (ABS_n_trips = n_trips) %>% 
  mutate(ABS_trips_share = round(ABS_n_trips/sum(ABS_n_trips)*100,1))
  
  
panel_mode = trips_by_mode %>% 
  filter(sample_source.x == 2)%>% 
  rename (panel_n_trips = n_trips) %>% 
  mutate(panel_trips_share = round(panel_n_trips/sum(panel_n_trips)*100,1))

temp = ABS_mode %>% 
  full_join(panel_mode,  by = "mode_1") %>% 
  dplyr::select(-c(sample_source.x.x,sample_source.x.y))

var_label_table = as.data.frame(value_labels[value_labels$variable %in% c("mode_1")] %>% mutate(value = as.double(value))) %>% 
  dplyr::select(value, value_label)

temp = temp %>% left_join(var_label_table, by = c("mode_1" = "value")) %>% 
  dplyr::select(-c(mode_1)) %>% 
  relocate(value_label) %>% 
  arrange(desc(ABS_trips_share))
trips_by_mode_uncleaned<-temp

  temp %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

------------------------------------------------------------------------

### trips by origin purpose by panel vs ABS; overall

```{r,echo=FALSE,message=FALSE,warning=FALSE}
orig_purpose_all = two_vars_average_by_sample_source(trips_person, origin_purpose,"origin_purpose")

orig_purpose_all_uncleaned<- orig_purpose_all
orig_purpose_all %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

------------------------------------------------------------------------

### trips by destination purpose by panel vs ABS; overall

```{r,echo=FALSE,message=FALSE,warning=FALSE}
dest_purpose_all = two_vars_average_by_sample_source(trips_person, dest_purpose,"dest_purpose")
dest_purpose_all_uncleaned<-dest_purpose_all
dest_purpose_all %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

------------------------------------------------------------------------

### total number of trips made by people under 18

```{r,echo=FALSE,message=FALSE,warning=FALSE}
var_label_table = as.data.frame(value_labels[value_labels$variable %in% c("sample_source")] %>% mutate(value = as.double(value)))

temp = trips_person_under17 %>% 
  group_by(sample_source.x) %>% 
  summarise (n_trips = n_distinct(tripid), n_people = n_distinct(personid)) %>% 
  mutate(average_trips_per_person = n_trips*1.0/n_people) %>% 
  left_join(var_label_table, by = c("sample_source.x" = "value")) %>% 
  dplyr::select(-c(variable, sample_source.x , label_value, val_order)) %>% 
  relocate(value_label)

trips_und_17_uncleaned <-temp
 temp %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
trips_und_17_uncleaned

```

------------------------------------------------------------------------

### total number of trips made by people under 18 by mode;

```{r,echo=FALSE,message=FALSE,warning=FALSE}
mode_under17 = two_vars_average_by_sample_source(trips_person_under17, mode_1,"mode_1")

mode_under17_uncleaned<-mode_under17
mode_under17 %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
mode_under17_uncleaned
```

------------------------------------------------------------------------

### total number of trips made by people under 18 by origin purpose

```{r,echo=FALSE,message=FALSE,warning=FALSE}
orig_purpose_all = two_vars_average_by_sample_source(trips_person_under17, origin_purpose,"origin_purpose")

origin_purpose_all_und18_uncleaned<-orig_purpose_all
orig_purpose_all %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

origin_purpose_all_und18_uncleaned
```

------------------------------------------------------------------------

### total number of trips made by people under 18 by destination purpose

```{r,echo=FALSE,message=FALSE,warning=FALSE}
dest_purpose_all = two_vars_average_by_sample_source(trips_person_under17, dest_purpose,"dest_purpose")

dest_purpose_all_und18_uncleaned<-dest_purpose_all
dest_purpose_all %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
dest_purpose_all_und18_uncleaned
```

### people with no trips by sample source

Online Panel Participants are assigned retrospective travel dates (most recent Monday, Tuesday, Wednesday, Thursday) and provide diary information for ONLY person 1. To find share of 0 trips for the panel, we need to subset pernum = 1. There are 896 people from the online panel that were asked o provide travel diary.

```{r,echo=FALSE,message=FALSE,warning=FALSE}
day_panel_pernum1 = d %>% filter(sample_source == 2, pernum == 1)

person_panel = subset(p, (personid %in% day_panel_pernum1$personid)) 

person_panel_summary = person_panel %>% 
    group_by(sample_source) %>%  summarise(ppl_count = n())

person_panel_summary_uncleaned<-person_panel_summary

person_panel_summary_formatted<-get_labels(person_panel_summary, sample_source, "sample_source") %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
person_panel_summary_formatted

```

```{r,echo=FALSE,message=FALSE,warning=FALSE}

#for ABS panel

personid_with_trips = t %>% distinct(personid)

ppl_ABS = p %>% 
          filter(sample_source == 1)

ppl_with_no_trips_ABS = subset(ppl_ABS, !(personid %in% personid_with_trips$personid))

ppl_with_no_trips_panel = subset(person_panel, !(personid %in% personid_with_trips$personid))

no_trips_by_ss_ABS = ppl_with_no_trips_ABS %>% 
                group_by(sample_source) %>% 
                summarise(n_ppl_no_trips=n())

no_trips_by_ss_panel = ppl_with_no_trips_panel %>% 
                group_by(sample_source) %>% 
                summarise(n_ppl_no_trips=n())

no_trips_by_ss = bind_rows(no_trips_by_ss_ABS, no_trips_by_ss_panel)

people_with_trips_panel = subset(person_panel, (personid %in% personid_with_trips$personid)) %>% 
                          group_by(sample_source) %>% 
                          summarise(n_ppl_with_trips =n())

people_with_trips = t %>%
  filter(sample_source == 1) %>% 
  group_by(sample_source) %>% 
  summarise (n_ppl_with_trips = n_distinct(personid)) %>% 
  bind_rows(people_with_trips_panel)

all_people = p %>% 
            group_by(sample_source) %>% 
            summarise (all_ppl = n_distinct(personid))

full_trips = people_with_trips %>% 
              full_join( no_trips_by_ss, by = "sample_source") %>%
              full_join(all_people, by = "sample_source") %>% 
              mutate(share_ppl_no_trips = n_ppl_no_trips*1.0/(n_ppl_no_trips+n_ppl_with_trips)*100)

full_trips_uncleaned<-full_trips

full_trips_formatted<-get_labels(full_trips, sample_source, "sample_source") %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
full_trips_formatted

```

# Cleaned Data

```{r}


# UPDATE folder path if running outside of RSG
# may also need to update file formats/read functions (e.g., read from a database)

elmer_connection <- dbConnect(odbc::odbc(),
                              driver = "SQL Server",
                              server = "AWS-PROD-SQL\\Sockeye",
                              database = "hhts_cleaning",
                              trusted_connection = "yes",
                                port = 1433)
#h <- dbGetQuery(elmer_connection,
#                "SELECT * FROM HHSurvey.HouseholdV")
p <- dbGetQuery(elmer_connection,
                "SELECT * FROM HHSurvey.Person")

d <- dbGetQuery(elmer_connection,
                "SELECT * FROM HHSurvey.Day")

t <- dbGetQuery(elmer_connection,
                "SELECT [recid]
      ,[survey_year]
      ,[hhid]
      ,[sample_source]
      ,[hhgroup]
      ,[nwkdays]
      ,[personid]
      ,[pernum]
      ,[daynum]
      ,[dayofweek]
      ,[traveldate]
      ,[data_source]
      ,[tripid]
      ,[tripnum]
      ,[travelers_total]
      ,[origin_purpose]
      ,[origin_purpose_cat]
      ,[dest_purpose]
      ,[dest_purpose_cat]
      ,[mode_1]
      ,[mode_type]
      ,[driver]
       FROM HHSurvey.Trip")

```

```{r, cache = TRUE, echo = FALSE, message = FALSE}

variable_labels = 
  read_codebook(
    params$codebook_path,
    varvals = FALSE,
    sheet = 'Variable_List')

setnames(variable_labels, 'label', 'description', skip_absent = TRUE)

value_labels = 
    read_codebook(
    params$codebook_path,
    varvals = TRUE, 
    sheet = 'Value_Labels')

setnames(value_labels, 'label', 'value_label', skip_absent = TRUE)

variable_labels[, common_name := gsub(':.*', '', description)]

value_labels[, value := as.character(value)]

# Change checkbox logic to be the same for each variable

variable_labels[variable %like% 'delivery_', logic := '']
variable_labels[variable %like% 'no_travel_', logic := 'If made no trips on travel day']
variable_labels[variable %like% 'ethnicity_', logic := 'If related adult household member']
variable_labels[variable %like% 'race_', logic := 'If related adult household member']


```

```{r, include = FALSE ,message=FALSE,warning=FALSE}
p<-p%>%mutate(personid=as.numeric(personid))
trips_person = left_join(t, p, by = "personid")

trips_person_under17 = trips_person %>% 
                        filter(age < 5)

trips_hh = left_join(t,h, by = "hhid")
```

### average trips per person by panel vs ABS

Here, average is calculated for people who made at least one trip (people who are present in trip table )

```{r, echo=FALSE,message=FALSE,warning=FALSE}


var_label_table = as.data.frame(value_labels[value_labels$variable %in% c("sample_source")] %>% mutate(value = as.double(value)))


temp = trips_person %>% 
  group_by(sample_source.x) %>% 
  summarise (n_trips = n_distinct(tripid), n_people = n_distinct(personid)) %>% 
  mutate(average_trips_per_person = n_trips*1.0/n_people) %>% 
  left_join(var_label_table, by = c("sample_source.x" = "value")) %>% 
  dplyr::select(-c(variable, sample_source.x , label_value, val_order)) %>% 
  relocate(value_label)

avg_trips_cleaned <- temp

  temp %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))



```

Here, average is calculated for all people (including people that didnt do any trips)

```{r, echo=FALSE,message=FALSE,warning=FALSE}
day_panel_pernum1 = d %>% filter(sample_source == 2, pernum == 1)

var_label_table = as.data.frame(value_labels[value_labels$variable %in% c("sample_source")] %>% mutate(value = as.double(value)))

person_panel = subset(p, (personid %in% day_panel_pernum1$personid)) 

person_panel_summary = person_panel %>% 
    group_by(sample_source) %>%  summarise(ppl_count = n())

ppl_ABS = p %>% 
          filter(sample_source == 1)%>% 
    group_by(sample_source) %>%  summarise(ppl_count = n())

all_ppl = bind_rows(ppl_ABS,person_panel_summary)


temp = t %>% 
  group_by(sample_source) %>% 
  summarise (n_trips = n_distinct(tripid)) %>% 
  full_join(all_ppl, by = "sample_source") %>% 
  mutate(average_trips_per_person = n_trips*1.0/ppl_count) %>% 
  left_join(var_label_table, by = c("sample_source" = "value")) %>% 
  dplyr::select(-c(variable, sample_source , label_value, val_order)) %>% 
  relocate(value_label)
 
avg_trips_all_cleaned<-temp
temp %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))



```

------------------------------------------------------------------------

### trips by mode (mode_1) by panel vs ABS; and overall

```{r,echo=FALSE,message=FALSE,warning=FALSE}


trips_by_mode = trips_person %>% 
  group_by(sample_source.x, mode_1) %>% 
  summarise (n_trips = n_distinct(tripid)) 

ABS_mode = trips_by_mode %>% 
  filter(sample_source.x == 1) %>% 
  rename (ABS_n_trips = n_trips) %>% 
  mutate(ABS_trips_share = round(ABS_n_trips/sum(ABS_n_trips)*100,1))
  
  
panel_mode = trips_by_mode %>% 
  filter(sample_source.x == 2)%>% 
  rename (panel_n_trips = n_trips) %>% 
  mutate(panel_trips_share = round(panel_n_trips/sum(panel_n_trips)*100,1))

temp = ABS_mode %>% 
  full_join(panel_mode,  by = "mode_1") %>% 
  dplyr::select(-c(sample_source.x.x,sample_source.x.y))

var_label_table = as.data.frame(value_labels[value_labels$variable %in% c("mode_1")] %>% mutate(value = as.double(value))) %>% 
  dplyr::select(value, value_label)

temp = temp %>% left_join(var_label_table, by = c("mode_1" = "value")) %>% 
  dplyr::select(-c(mode_1)) %>% 
  relocate(value_label) %>% 
  arrange(desc(ABS_trips_share))
trips_by_mode_cleaned<-temp

  temp %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

------------------------------------------------------------------------

### trips by origin purpose by panel vs ABS; overall

```{r,echo=FALSE,message=FALSE,warning=FALSE}
orig_purpose_all = two_vars_average_by_sample_source(trips_person, origin_purpose,"origin_purpose")

orig_purpose_all_cleaned<- orig_purpose_all
orig_purpose_all %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

------------------------------------------------------------------------

### trips by destination purpose by panel vs ABS; overall

```{r,echo=FALSE,message=FALSE,warning=FALSE}
dest_purpose_all = two_vars_average_by_sample_source(trips_person, dest_purpose,"dest_purpose")
dest_purpose_all_cleaned<-dest_purpose_all
dest_purpose_all %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

------------------------------------------------------------------------

### total number of trips made by people under 18

```{r,echo=FALSE,message=FALSE,warning=FALSE}
var_label_table = as.data.frame(value_labels[value_labels$variable %in% c("sample_source")] %>% mutate(value = as.double(value)))

temp = trips_person_under17 %>% 
  group_by(sample_source.x) %>% 
  summarise (n_trips = n_distinct(tripid), n_people = n_distinct(personid)) %>% 
  mutate(average_trips_per_person = n_trips*1.0/n_people) %>% 
  left_join(var_label_table, by = c("sample_source.x" = "value")) %>% 
  dplyr::select(-c(variable, sample_source.x , label_value, val_order)) %>% 
  relocate(value_label)

trips_und_17_cleaned <-temp
 temp %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
trips_und_17_cleaned

```

------------------------------------------------------------------------

### total number of trips made by people under 18 by mode;

```{r,echo=FALSE,message=FALSE,warning=FALSE}
mode_under17 = two_vars_average_by_sample_source(trips_person_under17, mode_1,"mode_1")

mode_under17_cleaned<-mode_under17
mode_under17 %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
mode_under17_cleaned
```

------------------------------------------------------------------------

### total number of trips made by people under 18 by origin purpose

```{r,echo=FALSE,message=FALSE,warning=FALSE}
orig_purpose_all = two_vars_average_by_sample_source(trips_person_under17, origin_purpose,"origin_purpose")

origin_purpose_all_und18_cleaned<-orig_purpose_all
orig_purpose_all %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

origin_purpose_all_und18_cleaned
```

------------------------------------------------------------------------

### total number of trips made by people under 18 by destination purpose

```{r,echo=FALSE,message=FALSE,warning=FALSE}
dest_purpose_all = two_vars_average_by_sample_source(trips_person_under17, dest_purpose,"dest_purpose")

dest_purpose_all_und18_cleaned<-dest_purpose_all
dest_purpose_all %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
dest_purpose_all_und18_cleaned
```

### people with no trips by sample source

Online Panel Participants are assigned retrospective travel dates (most recent Monday, Tuesday, Wednesday, Thursday) and provide diary information for ONLY person 1. To find share of 0 trips for the panel, we need to subset pernum = 1. There are 896 people from the online panel that were asked o provide travel diary.

```{r,echo=FALSE,message=FALSE,warning=FALSE}
day_panel_pernum1 = d %>% filter(sample_source == 2, pernum == 1)

person_panel = subset(p, (personid %in% day_panel_pernum1$personid)) 

person_panel_summary = person_panel %>% 
    group_by(sample_source) %>%  summarise(ppl_count = n())

person_panel_summary_cleaned<-person_panel_summary

person_panel_summary_formatted<-get_labels(person_panel_summary, sample_source, "sample_source") %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
person_panel_summary_formatted

```

```{r,echo=FALSE,message=FALSE,warning=FALSE}

#for ABS panel

personid_with_trips = t %>% distinct(personid)

ppl_ABS = p %>% 
          filter(sample_source == 1)

ppl_with_no_trips_ABS = subset(ppl_ABS, !(personid %in% personid_with_trips$personid))

ppl_with_no_trips_panel = subset(person_panel, !(personid %in% personid_with_trips$personid))

no_trips_by_ss_ABS = ppl_with_no_trips_ABS %>% 
                group_by(sample_source) %>% 
                summarise(n_ppl_no_trips=n())

no_trips_by_ss_panel = ppl_with_no_trips_panel %>% 
                group_by(sample_source) %>% 
                summarise(n_ppl_no_trips=n())

no_trips_by_ss = bind_rows(no_trips_by_ss_ABS, no_trips_by_ss_panel)

people_with_trips_panel = subset(person_panel, (personid %in% personid_with_trips$personid)) %>% 
                          group_by(sample_source) %>% 
                          summarise(n_ppl_with_trips =n())

people_with_trips = t %>%
  filter(sample_source == 1) %>% 
  group_by(sample_source) %>% 
  summarise (n_ppl_with_trips = n_distinct(personid)) %>% 
  bind_rows(people_with_trips_panel)

all_people = p %>% 
            group_by(sample_source) %>% 
            summarise (all_ppl = n_distinct(personid))

full_trips = people_with_trips %>% 
              full_join( no_trips_by_ss, by = "sample_source") %>%
              full_join(all_people, by = "sample_source") %>% 
              mutate(share_ppl_no_trips = n_ppl_no_trips*1.0/(n_ppl_no_trips+n_ppl_with_trips)*100)

full_trips_cleaned<-full_trips

full_trips_formatted<-get_labels(full_trips, sample_source, "sample_source") %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
full_trips_formatted


```

```{r,echo=FALSE,message=FALSE,warning=FALSE}
compare_df<-function(df1, df2){
  df1_num<- df1 %>% select(where(is.numeric))
  df2_num<- df2 %>% select(where(is.numeric))
  df_diff<- df2_num -df1_num
  df_differs <- cbind(df1%>%select(1), df_diff)%>%rename_with(~paste0(.,' cleaned-uncleaned difference'))
  df_percent_diff<-df_diff/df1_num
  df_percent_diff<-  df_percent_diff%>%mutate_at(names(.),funs(scales::percent(.)))%>%rename_with(~paste0(.,' cleaned-uncleaned percent difference'))
  df_percent_differs<- cbind(df_differs, df_percent_diff)
  df_percent_differs %>%
  kbl() %>%
  #kable_material(c("striped", "hover"))
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
}
```

# Compare Cleaned and Uncleaned data

## average trips per person

```{r}
compare_df(avg_trips_all_uncleaned,avg_trips_all_cleaned)

```

## trips by mode

```{r}
compare_df(trips_by_mode_uncleaned,trips_by_mode_cleaned)

```

## trips by origin purpose

```{r}
orig_purpose_all_uncleaned
orig_purpose_all_cleaned
```

## trips by destination purpose

```{r}
dest_purpose_all_uncleaned
dest_purpose_all_cleaned
```

## trips made by people under 18

```{r}
compare_df(trips_und_17_cleaned,trips_und_17_uncleaned)
```

## people with no trips

```{r}
full_trips_uncleaned
full_trips_cleaned
```
