---
title: "PSRC Weight Checks"
author: "RSG, Inc"
date: "1 Feb 2022"
editor_options: 
  chunk_output_type: console
---


```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, comment = '', fig.width=8, fig.asp=0.8)
options(scipen = 99999)

library(data.table)
library(forcats)
library(stringr)
library(janitor)
library(flextable)
library(ggplot2)
library(scales)

# directories and paths -------------------------------------------------------

# Directory for PSRC survey data

user = Sys.info()['user']
project_dir = str_glue(
  'C:/Users/{user}/Resource Systems Group, Inc/',
  'Transportation MR - Documents/PSRC Survey Program/21016_PSRC_HTS/',
  'Internal/3.DataAnalysis'
)

psrc_dir = file.path(project_dir, '1.Data/ClientData')

# Directory for RSG 2021 delivered weights
wt_dir = file.path(project_dir, '3.Weighting/for_psrc')

```


This document compares summary statistics using several different weights.  Each summary is calculated using the following:

1.  Unweighted
2.  Combined 2017/2019 weights delivered in 2019 (variable names end with `_wt_combined`).
3.  Combined 2017/2019 weights delivered in 2021 (variable names end with `_wt_1719`).


### PSRC suggested checks:

1.  num_trips by hhsize (household)
2.  prev_res_factors_housing_cost(household)
3.  telecommute_freq (person)
4.  wbt_transitmore_1 by race_afam (person)
5.  delivery_pkgs_freq by hhincome_detailed (day)
6.  mode_1(trips)
7.  d_purpose (trips)
8.  trip_path_distance histogram (trips)


### Calculate combined weights 

(join to PSRC data not shown)

```{r load_weights, echo=TRUE}

hh_wt = fread(file.path(wt_dir, 'hh_weights.csv'))

person_wt = fread(file.path(wt_dir, 'person_weights.csv'))

day_wt = fread(file.path(wt_dir, 'day_weights.csv'))

trip_wt = fread(file.path(wt_dir, 'trip_weights.csv'))

# Calculate combined weights --------------------------------------------------

hh_wt[, hh_wt_1719 := hh_weight / 2]

person_wt[, person_wt_1719 := person_weight / 2]

day_wt[, day_wt_1719 := day_weight / 2]

trip_wt[, trip_wt_1719 := trip_weight / 2]

```


```{r check_sums, eval=FALSE, include=FALSE}

# Check sums ------------------------------------------------------------------

# combined weight should be intermediate between the two years.

hh_wt[, hh_weight_2017 := hh_weight * (survey_year == 2017)]
hh_wt[, hh_weight_2019 := hh_weight * (survey_year == 2019)]

hh_wt[, lapply(.SD, sum), .SDcols = c('hh_weight_2017', 'hh_weight_2019', 'hh_wt_1719')]

person_wt[, person_weight_2017 := person_weight * (survey_year == 2017)]
person_wt[, person_weight_2019 := person_weight * (survey_year == 2019)]

person_wt[, lapply(.SD, sum), .SDcols = c('person_weight_2017', 'person_weight_2019', 'person_wt_1719')]

day_wt[, day_weight_2017 := day_weight * (survey_year == 2017)]
day_wt[, day_weight_2019 := day_weight * (survey_year == 2019)]

day_wt[, lapply(.SD, sum), .SDcols = c('day_weight_2017', 'day_weight_2019', 'day_wt_1719')]

trip_wt[, trip_weight_2017 := trip_weight * (survey_year == 2017)]
trip_wt[, trip_weight_2019 := trip_weight * (survey_year == 2019)]

trip_wt[, lapply(.SD, sum), .SDcols = c('trip_weight_2017', 'trip_weight_2019', 'trip_wt_1719')]

```


```{r load_psrc_data, include=FALSE}

# Read in PSRC data to join with weights


# Load PSRC data --------------------------------------------------------------

hh = readRDS(file.path(psrc_dir, 'hh.rds'))
person = readRDS(file.path(psrc_dir, 'person.rds'))
day = readRDS(file.path(psrc_dir, 'day.rds'))
trip = readRDS(file.path(psrc_dir, 'trip.rds'))


# Join weights ---------------------------------------------------------------

hh_wt[, household_id := as.character(household_id)]

hh[
  hh_wt,
  hh_wt_1719 := hh_wt_1719,
  on = 'household_id']

person_wt[, person_id := as.character(person_id)]
person[
  person_wt,
  person_wt_1719 := person_wt_1719,
  on = 'person_id']

day_wt[, person_id := as.character(person_id)]
day[
  day_wt,
  day_wt_1719 := day_wt_1719,
  on = c('person_id', 'daynum')]

trip_wt[, trip_id := as.character(trip_id)]
trip[
  trip_wt,
  trip_wt_1719 := trip_wt_1719,
  on = 'trip_id']


# Check sums of weights have not changed due to join -------------------------

stopifnot(
  hh_wt[, sum(hh_wt_1719)] == hh[, sum(hh_wt_1719)],
  person_wt[, sum(person_wt_1719)] == person[, sum(person_wt_1719)],
  day_wt[, sum(day_wt_1719)] == day[, sum(day_wt_1719)],
  trip_wt[, sum(trip_wt_1719)] == trip[, sum(trip_wt_1719)]
)


```



### Number of trips by household size

```{r}

# Num trips by hhsize

hh[, num_trips := as.numeric(num_trips)]
tbl = hh[, .(
  Unweighted = mean(num_trips),
  hh_wt_combined = sum(num_trips * hh_wt_combined)/sum(hh_wt_combined),
  hh_wt_1719 = sum(num_trips * hh_wt_1719)/sum(hh_wt_1719)),
keyby = hhsize]

tbl %>%
  flextable() %>%
  colformat_double(digits=2) %>%
  autofit()

```


```{r}

dt = tbl %>%
  melt(id.vars = 'hhsize', variable.name = 'Weight', value.name = 'num_trips')
dt[, Weight := fct_inorder(Weight)]
  
ggplot(dt, aes(x = hhsize, y = num_trips, fill = Weight)) +
  geom_col(position='dodge') +
  coord_flip()


```


### Previous residence housing cost 

```{r}

tbl = hh[prev_res_factors_housing_cost %in% c('Selected', 'Not Selected'),
  .(Unweighted = .N, 
    hh_wt_combined = sum(hh_wt_combined),
    hh_wt_1719 = sum(hh_wt_1719)), 
  by = prev_res_factors_housing_cost]


tbl = tbl %>%
  adorn_percentages(denominator='col')

tbl %>%
  adorn_pct_formatting(digits=2) %>%
  flextable() %>%
  autofit()
```


### Telecommute frequency

```{r}

label_order = c(
  'Never', 'Less than monthly', 'A few times per month',
  '1 day a week', paste(c(2:5, '6-7'), 'days a week'))

person[, telecommute_freq_og := telecommute_freq]
person[, telecommute_freq := factor(telecommute_freq_og, levels=label_order)]
# person[, .N, keyby=.(telecommute_freq, telecommute_freq_og)]

tbl = person[
  !is.na(telecommute_freq),
  .(Unweighted = .N, 
    hh_wt_combined = sum(hh_wt_combined),
    person_wt_1719 = sum(person_wt_1719)), 
  keyby= telecommute_freq]

tbl = tbl %>%
  adorn_percentages(denominator='col')

tbl  %>%
  adorn_pct_formatting(digits=2) %>%
  flextable() %>%
  autofit()

```


```{r}

dt = tbl %>%
  melt(
    id.vars = 'telecommute_freq',
    variable.name = 'Weight',
    value.name = 'fraction')

dt[, Percent := fraction * 100]
dt[, Weight := fct_inorder(Weight)]
  
ggplot(dt, aes(x = telecommute_freq, y = Percent, fill = Weight)) +
  geom_col(position='dodge') +
  coord_flip()

```


### "Use more transit -- safer ways to get to stops" by Race: African American


```{r}

person[, wbt_transitmore_1_og := wbt_transitmore_1]

level_order = c(
  'Never', 'Rarely (1-3 times per month)', 'Occasionally (1-2 days per week)',
  'Somewhat often (3-5 days per week)', 'Regularly (6-7 days per week)')

person[, wbt_transitmore_1 := factor(wbt_transitmore_1_og, levels=level_order)]
# person[, .N, keyby = .(wbt_transitmore_1, wbt_transitmore_1_og)]

levels(person$wbt_transitmore_1) = str_replace(
  levels(person$wbt_transitmore_1),
  ' [(]', '\n(')
# person[, .N, keyby = wbt_transitmore_1]


person[, race_afam_og := race_afam]
person[, race_afam := fifelse(race_afam_og == 'Selected', 'African-American', 'Other race')]

tbl = person[
  !is.na(wbt_transitmore_1),
  .(Unweighted = as.numeric(.N),
    hh_wt_combined = sum(hh_wt_combined),
    person_wt_1719 = sum(person_wt_1719)),
  keyby = .(wbt_transitmore_1, race_afam)] %>%
  
  melt(
    id.vars = c('wbt_transitmore_1', 'race_afam'),
    variable.name = 'Weight',
    value.name = 'Count')

tbl[, Fraction := Count / sum(Count), by = .(Weight, race_afam)]
tbl[, Percent := Fraction * 100]
# tbl[, sum(Fraction), by = .(Weight, race_afam)]

tbl %>%
  dcast(race_afam + wbt_transitmore_1 ~ Weight, value.var = 'Percent') %>%
  flextable() %>%
  colformat_double(digits = 2, suffix='%') %>%
  autofit()

```


```{r, fig.asp=0.6}

ggplot(tbl, aes(x = wbt_transitmore_1, y=Percent, fill = Weight)) +
  geom_col(position='dodge') +
  facet_wrap(~ race_afam) +
  coord_flip()

```

### "Package deliveries on travel day" by Household Income


```{r}

dt = day[
  , .(household_id, delivery_pkgs_freq, hh_day_wt_combined, day_wt_1719)]

dt = dt[
  hh,
  hhincome_detailed := hhincome_detailed,
  on = 'household_id']

income_levels = c(
  'Prefer not to answer', 'Under $10,000', '$10,000-$24,999', '$25,000-$34,999',
  '$35,000-$49,999', '$50,000-$74,999', '$75,000-$99,999', '$100,000-$149,999',
  '$150,000-$199,999', '$200,000-$249,999', '$250,000 or more')

dt[, hhincome_detailed_og := hhincome_detailed]
dt[, hhincome_detailed := factor(hhincome_detailed_og, levels = income_levels)]
# dt[, .N, keyby = .(hhincome_detailed, hhincome_detailed_og)]

tbl = dt[
  !(is.na(delivery_pkgs_freq) | delivery_pkgs_freq %like% 'Missing'),
  .(Unweighted = as.numeric(.N),
    hh_wt_combined = sum(hh_day_wt_combined),
    person_wt_1719 = sum(day_wt_1719)),
  keyby = .(hhincome_detailed, delivery_pkgs_freq)] %>%

  melt(
    id.vars = c('delivery_pkgs_freq', 'hhincome_detailed'),
    variable.name = 'Weight',
    value.name = 'Count')

tbl[, Fraction := Count / sum(Count), by = .(Weight, hhincome_detailed)]
tbl[, Percent := Fraction * 100]
# tbl[, sum(Fraction), by = .(Weight, hhincome_detailed)]

tbl %>%
  dcast(hhincome_detailed + delivery_pkgs_freq ~ Weight, value.var = 'Percent') %>%
  flextable() %>%
  colformat_double(digits = 2, suffix='%') %>%
  autofit()

```



```{r}

ggplot(tbl, aes(x = delivery_pkgs_freq, y=Percent, fill = Weight)) +
  geom_col(position='dodge') +
  facet_wrap(~ hhincome_detailed, as.table=FALSE) +
  coord_flip()

```


### Travel mode 


```{r}

trip[, mode_1_og := mode_1]
trip[, mode_1 := str_wrap(mode_1_og, width = 25)]
trip[, mode_1 := fct_infreq(mode_1)]
trip[, mode_1 := fct_lump_prop(mode_1, prop=0.01)]

tbl = trip[,
  .(Unweighted = .N, 
    trip_wt_combined = sum(trip_wt_combined),
    trip_wt_1719 = sum(trip_wt_1719)), 
  keyby=mode_1]

tbl = tbl %>%
  adorn_percentages(denominator='col')

tbl %>%
  adorn_pct_formatting(digits=2) %>%
  flextable() %>%
  autofit()
```



```{r}

dt = tbl %>%
  melt(
    id.vars = 'mode_1',
    variable.name = 'Weight',
    value.name = 'fraction')

dt[, Percent := fraction * 100]
dt[, Weight := fct_inorder(Weight)]
  
dt[, mode_1 := fct_rev(mode_1)]
ggplot(dt, aes(x = mode_1, y = Percent, fill = Weight)) +
  geom_col(position='dodge') +
  coord_flip()

```

### Destination purpose


```{r}

trip[, dest_purpose_og := dest_purpose]
trip[, dest_purpose := fct_infreq(dest_purpose_og)]
trip[, dest_purpose := fct_lump_prop(dest_purpose, prop = 0.01)]

tbl = trip[!is.na(dest_purpose),
  .(Unweighted = .N, 
    trip_wt_combined = sum(trip_wt_combined),
    trip_wt_1719 = sum(trip_wt_1719)), 
  keyby = dest_purpose] %>%
  adorn_percentages(denominator = 'col')

tbl %>%
  adorn_pct_formatting(digits=2) %>%
  flextable() %>%
  autofit()

```



```{r}

trip[, dest_purpose := str_wrap(dest_purpose_og, width = 35)]
trip[, dest_purpose := fct_infreq(dest_purpose)]
trip[, dest_purpose := fct_lump_prop(dest_purpose, prop = 0.01)]

tbl = trip[!is.na(dest_purpose),
  .(Unweighted = .N, 
    trip_wt_combined = sum(trip_wt_combined),
    trip_wt_1719 = sum(trip_wt_1719)), 
  keyby = dest_purpose] %>%
  adorn_percentages(denominator = 'col')

dt = tbl %>%
  melt(
    id.vars = 'dest_purpose',
    variable.name = 'Weight',
    value.name = 'fraction')

dt[, Percent := fraction * 100]
dt[, Weight := fct_inorder(Weight)]
dt[, dest_purpose := fct_rev(dest_purpose)]

ggplot(dt, aes(x = dest_purpose, y = Percent, fill = Weight)) +
  geom_col(position='dodge') +
  coord_flip()
```

### Trip distance 

```{r, fig.width=6, fig.asp = 1.4}

dt = trip[, .(trip_id, trip_path_distance, trip_wt_combined, trip_wt_1719)]
dt[, Unweighted := 1]

dt = melt(dt, id.vars=c('trip_id', 'trip_path_distance'), variable.name = 'Weight', value.name = 'wt_value')

weight_levels = c('Unweighted', 'trip_wt_combined', 'trip_wt_1719')
dt[, Weight := factor(Weight, levels = weight_levels)]

ggplot(
  dt[!is.na(wt_value) & !is.na(trip_path_distance)],
  aes(
    x = log10(trip_path_distance + 0.01),
    fill = Weight,
    weight=wt_value)) + 
  geom_histogram(binwidth=0.1, color='white') +
  facet_wrap(~ Weight, ncol=1, scales='free_y') + 
  scale_y_continuous(labels = comma) +
  theme(legend.position='none')
  
```

