---
title: "Regional Growth Center Analysis using 2017, 2019, and 2021 travel surveys"
output:
  html_document:
    df_print: paged
---

Recent analysis has uncovered that people who live in RGCs have substantially different travel choices than those who do not.

```{r }
library(psrc.travelsurvey)
library(tidyr)
library(dplyr)
library(plotly)
library(stringr)
library(psrcplot)
library(tidycensus)
library(htmlwidgets)
install_psrc_fonts()
```

Clearly living in RGC does not directly cause this behavior. People who live in RGCs are already different from people who do not- they have smaller household sizes, are younger, and have higher incomes (at least I think).

But it shows that people in RGCs are able to take advantage of better transportation options and closer activities.

# Variables we might want to summarize
```{r}
mode_vars<-c('mode_1', 'mode_simple')
other_vars<-c('sample_county', 'vehicle_count',  "hhincome_broad",  'hhincome_detailed', "age", "age_category", 'final_home_is_rgc', 'race_category', 'race_eth_broad', 'dest_purpose_cat', 'origin_purpose_cat', 'trip_path_distance', 'person_id')
all_vars<-c(mode_vars, other_vars)
person_vars=c('vehicle_count',  "hhincome_broad",  'hhincome_detailed', "age", "age_category", 'final_home_is_rgc', 'race_category', 'race_eth_broad', 'person_id', 'workplace')
```

Get the data from Elmer.
 
```{r}
trip_data_17_19<- get_hhts("2017_2019", "t", vars=all_vars)
trip_data_21<- get_hhts("2021", "t", vars=all_vars)
```


```{r}
person_data_17_19<- get_hhts("2017_2019", "p", vars=person_vars)
person_data_21<- get_hhts("2021", "p", vars=person_vars)
```


Group variables into meaningful categories with statistically large enough groups.First for trips, group the data.
```{r}
trip_data_17_19 <-
  trip_data_17_19 %>% mutate(NoVehicles = ifelse(vehicle_count == '0 (no vehicles)', 'No Vehicles', "Has Vehicles")) %>%
  mutate(
    hhincome_100 = case_when(
      hhincome_broad == '$100,000-$199,000' ~ '$100,000 or more',
      hhincome_broad == '$200,000 or more' ~ '$100,000 or more',
      TRUE ~ hhincome_broad
    )
  ) %>%
  mutate(
    age_grp = case_when(
      age == '75-84 years' ~ '75 years or older',
      age == '85 or years older' ~ '75 years or older',
      TRUE ~ age
    )
  ) %>% mutate(simple_purpose = ifelse(
    dest_purpose_cat == 'Home',
    origin_purpose_cat,
    dest_purpose_cat
  )) %>%
  mutate(
    simple_purpose = case_when(
      simple_purpose == 'Work' ~ 'Work/School',
      simple_purpose == 'School' ~ 'Work/School',
      simple_purpose == 'Work-related' ~ 'Work/School',
      simple_purpose == 'Shop' ~ 'Shop',
      simple_purpose == 'Escort' ~ 'Errands',
      simple_purpose == 'Errand/Other' ~ 'Errands',
      simple_purpose == 'Change mode' ~ 'Errands',
      simple_purpose == 'Social/Recreation' ~ 'Social/Recreation/Meal',
      simple_purpose == 'Meal' ~ 'Social/Recreation/Meal',
      simple_purpose == 'Meal' ~ 'Social/Recreation/Meal',
      simple_purpose == 'Home' ~ 'Errands',
      is.na(simple_purpose) ~ 'Errands',
      TRUE ~ simple_purpose
    )
  )


trip_data_21 <-
  trip_data_21 %>% mutate(NoVehicles = ifelse(vehicle_count == '0 (no vehicles)', 'No Vehicles', "Has Vehicles")) %>%
  mutate(
    hhincome_100 = case_when(
      hhincome_broad == '$100,000-$199,000' ~ '$100,000 or more',
      hhincome_broad == '$200,000 or more' ~ '$100,000 or more',
      TRUE ~ hhincome_broad
    )
  ) %>%
  mutate(
    age_grp = case_when(
      age == '75-84 years' ~ '75 years or older',
      age == '85 or years older' ~ '75 years or older',
      TRUE ~ age
    )
  ) %>% mutate(simple_purpose = ifelse(
    dest_purpose_cat == 'Home',
    origin_purpose_cat,
    dest_purpose_cat
  )) %>%
  mutate(
    simple_purpose = case_when(
      simple_purpose == 'Work' ~ 'Work/School',
      simple_purpose == 'School' ~ 'Work/School',
      simple_purpose == 'Work-related' ~ 'Work/School',
      simple_purpose == 'Shop' ~ 'Shop',
      simple_purpose == 'Escort' ~ 'Errands',
      simple_purpose == 'Errand/Other' ~ 'Errands',
      simple_purpose == 'Change mode' ~ 'Errands',
      simple_purpose == 'Social/Recreation' ~ 'Social/Recreation/Meal',
      simple_purpose == 'Meal' ~ 'Social/Recreation/Meal',
      simple_purpose == 'Meal' ~ 'Social/Recreation/Meal',
      simple_purpose == 'Home' ~ 'Errands',
      is.na(simple_purpose) ~ 'Errands',
      TRUE ~ simple_purpose
    )
  ) 


trip_data_17_19$hhincome_100_f = factor(
  trip_data_17_19$hhincome_100,
  levels = c(
    "Under $25,000",
    "$25,000-$49,999",
    "$50,000-$74,999",
    "$75,000-$99,999",
    "$100,000 or more",
    "Prefer not to answer"
  )
)

trip_data_21$hhincome_100_f = factor(
  trip_data_21$hhincome_100,
  levels = c(
    "Under $25,000",
    "$25,000-$49,999",
    "$50,000-$74,999",
    "$75,000-$99,999",
    "$100,000 or more",
    "Prefer not to answer"
  )
)


trip_data_17_19$NoVehicles = factor(trip_data_17_19$NoVehicles,
                                    levels = c('No Vehicles', 'Has Vehicles'))

trip_data_21$NoVehicles = factor(trip_data_21$NoVehicles,
                                 levels = c('No Vehicles', 'Has Vehicles'))


simple_groupings <-
  c(
    'race_eth_broad' = 'Race/Ethnicity',
    'hhincome_100_f' = 'Household Income',
    'NoVehicles' = 'Household Vehicles',
    'age_grp' = 'Age Group',
    'sample_county' = 'Home County',
    'age_category' = 'Age Category',
    'final_home_is_rgc' = 'Home in Regional Growth Center'
  )
```

Now group the data into the same categories for people.
```{r}
person_data_17_19 <-
  person_data_17_19 %>% mutate(NoVehicles = ifelse(vehicle_count == '0 (no vehicles)', 'No Vehicles', "Has Vehicles")) %>%
  mutate(
    hhincome_100 = case_when(
      hhincome_broad == '$100,000-$199,000' ~ '$100,000 or more',
      hhincome_broad == '$200,000 or more' ~ '$100,000 or more',
      TRUE ~ hhincome_broad
    )
  ) %>%
  mutate(
    age_grp = case_when(
      age == '75-84 years' ~ '75 years or older',
      age == '85 or years older' ~ '75 years or older',
      TRUE ~ age
    )
  ) 

person_data_21 <-
  person_data_21 %>% mutate(NoVehicles = ifelse(vehicle_count == '0 (no vehicles)', 'No Vehicles', "Has Vehicles")) %>%
  mutate(
    hhincome_100 = case_when(
      hhincome_broad == '$100,000-$199,000' ~ '$100,000 or more',
      hhincome_broad == '$200,000 or more' ~ '$100,000 or more',
      TRUE ~ hhincome_broad
    )
  ) %>%
  mutate(
    age_grp = case_when(
      age == '75-84 years' ~ '75 years or older',
      age == '85 or years older' ~ '75 years or older',
      TRUE ~ age
    )
  ) 


person_data_17_19$hhincome_100_f = factor(
  person_data_17_19$hhincome_100,
  levels = c(
    "Under $25,000",
    "$25,000-$49,999",
    "$50,000-$74,999",
    "$75,000-$99,999",
    "$100,000 or more",
    "Prefer not to answer"
  )
)

person_data_21$hhincome_100_f = factor(
  person_data_21$hhincome_100,
  levels = c(
    "Under $25,000",
    "$25,000-$49,999",
    "$50,000-$74,999",
    "$75,000-$99,999",
    "$100,000 or more",
    "Prefer not to answer"
  )
)


person_data_17_19$NoVehicles = factor(person_data_17_19$NoVehicles,
                                      levels = c('No Vehicles', 'Has Vehicles'))

person_data_21$NoVehicles = factor(person_data_21$NoVehicles,
                                   levels = c('No Vehicles', 'Has Vehicles'))



simple_groupings <-
  c(
    'race_eth_broad' = 'Race/Ethnicity',
    'hhincome_100_f' = 'Household Income',
    'NoVehicles' = 'Household Vehicles',
    'age_grp' = 'Age Group',
    'sample_county' = 'Home County',
    'age_category' = 'Age Category',
    'final_home_is_rgc' = 'Home in Regional Growth Center'
  )
```


##How many trips do people make in RGCs and for what purposes?

For this we have to aggregate both the trips and the people, and join them to calculate trips per person.
We may want to functionalize some of this since we are doing things twice, just felt easier not to. (i get indecisive about functionalizing things i do only twice)

```{r}
trip_summs_2017_2019 <- hhts_count(
  trip_data_17_19,
  group_vars = c('final_home_is_rgc', 'simple_purpose'),
  spec_wgt = 'trip_weight_2017_2019_v2021_adult'
) %>% drop_na(c('simple_purpose', 'final_home_is_rgc')) %>% filter(final_home_is_rgc !=
                                                                     'Total')
trip_summs_2021 <- hhts_count(
  trip_data_21,
  group_vars = c('final_home_is_rgc', 'simple_purpose'),
  spec_wgt = 'trip_weight_2021_ABS_Panel_adult'
) %>% drop_na(c('final_home_is_rgc', 'simple_purpose')) %>% filter(final_home_is_rgc !=
                                                                     'Total')


person_summs_2017_2019 <- hhts_count(
  person_data_17_19,
  group_vars = c('final_home_is_rgc'),
  spec_wgt = 'hh_weight_2017_2019_v2021_adult'
) %>%
  filter(final_home_is_rgc != 'Total') %>% drop_na(final_home_is_rgc)

person_summs_2021 <- hhts_count(person_data_21,
                                group_vars = c('final_home_is_rgc'),
                                spec_wgt = 'person_weight_2021_ABS_Panel_adult') %>%
  filter(final_home_is_rgc != 'Total') %>% drop_na(final_home_is_rgc)


# Join persons and trips by group and calculate trips/person.

trips_persons_2017_2019 <-
  merge(
    trip_summs_2017_2019,
    person_summs_2017_2019,
    by = 'final_home_is_rgc',
    suffixes = c('_trips', '_people')
  )
trips_persons_2021 <-
  merge(
    trip_summs_2021,
    person_summs_2021,
    by = 'final_home_is_rgc',
    suffixes = c('_trips', '_people')
  )

trips_persons_2017_2019 <-
  trips_persons_2017_2019 %>% mutate(
    trips_per_person = count_trips / count_people,
    moe_trips_person = moe_ratio(count_trips, count_people, count_moe_trips, count_moe_people)
  )


trips_persons_2021 <-
  trips_persons_2021 %>% mutate(
    trips_per_person = count_trips / count_people,
    moe_trips_person = moe_ratio(count_trips, count_people, count_moe_trips, count_moe_people)
  )

trips_persons_long <-
  rbind(trips_persons_2017_2019, trips_persons_2021) %>%
  mutate(survey_trips = str_replace_all(survey_trips, "_", "/"))

trips_persons_long_all_trips <-
  trips_persons_long %>% filter(simple_purpose == 'Total')

trips_persons_2021_purpose <-
  trips_persons_2021 %>% filter(simple_purpose != 'Total')

trips_persons_long_shop_trips <-
  trips_persons_long %>% filter(simple_purpose == 'Shop')



```
Make charts about trips by purpose by RGC
```{r}
p0 <-
  create_bar_chart(
    t = trips_persons_2021_purpose ,
    x = 'simple_purpose',
    y = 'trips_per_person',
    f = 'final_home_is_rgc',
    moe = 'moe_trips_person',
    est = 'number',
    color = 'psrc_pairs',
    interactive = 'no'
  )  


p1 <-
  create_facet_bar_chart(
    t = trips_persons_long,
    w.x = 'final_home_is_rgc',
    w.y = 'trips_per_person',
    f = 'survey_trips',
    g = 'simple_purpose',
    w.moe = 'moe_trips_person',
    est= 'number',
    w.color = 'psrc_light',
    w.interactive = 'no'
  )

p2 <-
  create_column_chart(
    t = trips_persons_long_all_trips ,
    x = 'final_home_is_rgc',
    y = 'trips_per_person',
    f = 'survey_trips',
    moe = 'moe_trips_person',
    est = 'number',
    color = 'psrc_pairs',
    interactive = 'no'
  ) 


p3 <-
  create_column_chart(
    t = trips_persons_long_shop_trips ,
    x = 'final_home_is_rgc',
    y = 'trips_per_person',
    f = 'survey_trips',
    moe = 'moe_trips_person',
    est = 'number',
    color = 'psrc_pairs',
    interactive = 'no'
  ) 


p0
p1
p2
p3

```






Trip Mode Shares by RGCs or not
```{r }
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]

  trip_summs_2017_2019 <- hhts_count(trip_data_17_19,
                                        group_vars=c('final_home_is_rgc','mode_simple'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter('final_home_is_rgc'!='Total') %>%filter(mode_simple!='Total')%>%drop_na('final_home_is_rgc')%>%mutate('final_home_is_rgc'=str_wrap('final_home_is_rgc', width=10))



  trip_summs_2021 <- hhts_count(trip_data_21,
                                   group_vars=c('final_home_is_rgc','mode_simple'),
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter('final_home_is_rgc'!='Total') %>%filter(mode_simple!='Total')%>%drop_na('final_home_is_rgc')%>%mutate('final_home_is_rgc'=str_wrap('final_home_is_rgc', width=10))

  
  trip_summs_long<- rbind(trip_summs_2017_2019, trip_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))




   p0<-create_column_chart(t=trip_summs_2021 , x='mode_simple', y='share', f='final_home_is_rgc', moe='share_moe', est='percent', color = 'psrc_light', interactive=TRUE)  +theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  

  p1<-create_facet_bar_chart(t=trip_summs_long , f='mode_simple',  x='final_home_is_rgc', y='share', g='survey', moe='share_moe', est='percent', color = 'psrc_light', interactive=TRUE) +theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)

}

p0
p1



```

# Walk Trip Mode Shares by RGC or not; just zooming in on walk
```{r }
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  trip_summs_2017_2019 <- hhts_count(trip_data_17_19,
                                        group_vars=c('final_home_is_rgc','mode_simple'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter('final_home_is_rgc'!='Total') %>%
    filter(mode_simple=='Walk')%>%drop_na('final_home_is_rgc')%>%mutate('final_home_is_rgc'=str_wrap('final_home_is_rgc', width=10))



  trip_summs_2021 <- hhts_count(trip_data_21,
                                   group_vars=c('final_home_is_rgc','mode_simple'),
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter('final_home_is_rgc'!='Total') %>%
    filter(mode_simple=='Walk')%>%drop_na('final_home_is_rgc')%>%mutate('final_home_is_rgc'=str_wrap('final_home_is_rgc', width=10))

  trip_summs_long<- rbind(trip_summs_2017_2019, trip_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))


   p0<-create_column_chart(t=trip_summs_2021 , x='final_home_is_rgc', y='share', f='mode_simple', moe='share_moe', est='percent', color = 'psrc_light', interactive=TRUE) + 
  xlab(as.character(g['final_home_is_rgc'])) + ylab('Walk mode share')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  


  p1<-create_column_chart(t=trip_summs_long , x='final_home_is_rgc', y='share', f='survey', moe='share_moe', est='percent', color = 'psrc_light', interactive=TRUE)+ 
  xlab(as.character(g['final_home_is_rgc'])) + ylab("Walk mode share")+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)


```




Median Trip Distances, rgc or not, People in RGCs have shorter trips
```{r }

  trip_summs_2017_2019 <- hhts_median(trip_data_17_19,
                                         stat_var = 'trip_path_distance',
                                        group_vars=c('final_home_is_rgc'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
filter('final_home_is_rgc'!='Other')%>%drop_na('final_home_is_rgc')

  

  trip_summs_2021<- hhts_median(trip_data_21,
                                         stat_var = 'trip_path_distance',
                                        group_vars=c('final_home_is_rgc'),
                                        spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
filter('final_home_is_rgc'!='Other')%>%drop_na('final_home_is_rgc')
  
  

  trip_summs_long<- rbind(trip_summs_2017_2019, trip_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))



   p0<-create_column_chart(t=trip_summs_2021 , x='final_home_is_rgc', y='trip_path_distance_median', f='survey', moe='trip_path_distance_median_moe', est='number', color = 'psrc_pairs', interactive=TRUE) + 
  xlab(as.character(g['final_home_is_rgc'])) + ylab('Median Distance')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  


  p1<-create_column_chart(t=trip_summs_long , x='final_home_is_rgc', y='trip_path_distance_median', f='survey', moe='trip_path_distance_median_moe', est='number', color = 'psrc_pairs', interactive=TRUE)+ 
  xlab(as.character(g['final_home_is_rgc'])) + ylab("Median Distance")+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)

}
```

