---
title: "Seattle mode share analysis for Comp Plan EIS"
author: "suzanne"
date: '2022-06-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Stephen Fesler
The Urbanist
I'm hoping to obtain contemporary data for the Paine Field/Everett MIC. I'm hoping to understand more current employment levels in that area and the mode share in that area. I'm especially interested in understanding the share of transit use versus the share of SOV by employees in the area, but a full mode share breakdown would be useful, too. I was only able to find a PDF with 2010 estimates on this data, which is obviously considerably out of date. If the PSRC also has data on common commute travel patterns to/from the MIC, that would also be helpful. Our intended use of this is strictly educational.


Hi Stephen, here are the latest data we can provide for your request.

Employment - year 2020 total employment is 51,622 for the Paine Field MIC subarea

Mode Share - year 2018 travel demand model (SoundCast), mode for work trips ending in the Paine Field MIC subarea

        - SOV           75%
        - HOV           17%
        - Walk/Bike      4%
        - Transit       4%

We can get observed mode shares from the household travel survey, but there aren't enough records for the Paine Field MIC area to be statistically valid.  So just keep in mind that the mode shares I've provided are not observed but modeled estimates.
Kris
## read in shape file with geographies

```{r }
library(sf)
library(dplyr)
library(psrc.travelsurvey)
library(leaflet)
library(tidyr)
library(ggplot2)
library(raster)
library(ggspatial)
library(RColorBrewer)
library(lubridate)
library(scales)
```



Read in the Sector layer

```{r}

spn<- 2285
wgs_84<-4326

sectors_paine<-"C:/GitHub/travel-studies/2021/analysis_in_progress/paine_field/PaineMIC.shp"

sector_layer<-st_read(sectors_paine)
sector_layer<- st_transform(sector_layer, crs=wgs_84)
```

Read in the travel survey data and do some work to clean up modes and times of day.
```{r}
mode_geo_df<-get_hhts('2017_2019', "t", vars= c('trip_id','main_mode', 'travelers_total', 'origin_lat', 'origin_lng', 'dest_lat', 'dest_lng', 'mode_simple', 'dest_purpose_cat', 'origin_purpose_cat', 'trip_path_distance'))%>%
drop_na(c(origin_lat, dest_lat))%>%
mutate(mode_incl_hov = ifelse(mode_simple !='Drive', mode_simple, ifelse(travelers_total==1, 'SOV', 'HOV')))%>%
filter(dest_purpose_cat=='Work' | origin_purpose_cat=='Work')

```



Turn the trip ends into a geographic data layer.
```{r}
mode_geo_sf_origin<-mode_geo_df%>% st_as_sf(coords=c("origin_lng","origin_lat"), crs= "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
mode_geo_sf_dest<-mode_geo_df%>% st_as_sf(coords=c("dest_lng","dest_lat"), crs= "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")


```

Read in Seattle Sectors Layer



make a map of the sectors with the trip ends
```{r}
map.lat<-47.615
map.lon<--122.257
map.zoom<-10
pal <- leaflet::colorFactor(palette="Set2", domain = sector_layer$area)

travel_survey_map<- leaflet()%>%
     addProviderTiles(providers$CartoDB.Positron) %>%
  addLayersControl(baseGroups = c("Base Map"),
                      overlayGroups = c("Travel Survey Origins", "Paine MIC"),
                      options = layersControlOptions(collapsed = TRUE))%>%

  addPolygons(data=sector_layer,
               fillColor = pal(sector_layer$COUNTY),
               weight = 4,
               opacity = 0.1,
               color = "#91268F",
               dashArray = "4",
               fillOpacity = 1.0,
               group = "Seattle Sectors") %>%
     addCircles(data=mode_geo_sf_origin,
                  weight = .1, 
                 radius = .1,
                  fill = TRUE,
                  opacity = 0.01,
                  group = ("Travel Survey Origins")) %>%

    leaflet::addLegend(pal = pal,
                       values = sector_layer$COUNTY,
                       position = "bottomright",
                       title = paste('Paine MIC', '<br>', 'for Mode Share')) %>%

    leaflet::setView(lng=map.lon, lat=map.lat, zoom=map.zoom)

#travel_survey_map
```

Overlay the trip origins and destinations with the Seattle sectors so they can be put into them.
```{r}
sf::sf_use_s2(FALSE)
origin_sector_paine<- st_join(mode_geo_sf_origin, sector_layer, join=st_within)
dest_sector_paine<-st_join(mode_geo_sf_dest,sector_layer, join=st_within)


```

```{r}
origin_sector_paine<- as.data.frame(st_drop_geometry(origin_sector_paine))
dest_sector_paine<- as.data.frame(st_drop_geometry(dest_sector_paine))

```

```{r}
od_sector_paine<-merge(origin_sector_paine, dest_sector_paine, by='trip_id', suffixes=c('', '_dest'))%>% mutate(mode_nonm=ifelse((mode_incl_hov=='Walk' | mode_incl_hov=='Bike') , 'Non-Motorized', mode_incl_hov))%>%filter_at(vars('COUNTY', 'COUNTY_dest'), any_vars(!is.na(.)))
```

Summarize all trips by whether they are in the PM Peak, what mode they are, what Seattle sector the origin is in, and what Seattle sector the destination is in. Includes all times and places outside Seattle.
```{r}
```


```{r}
trips_by_mode<-hhts_count(od_sector_paine, group_vars=c('mode_nonm'), spec_wgt='trip_weight_2017_2019_v2021')
```

```{r}
work_dist<-hhts_median(od_sector_paine, stat_var='trip_path_distance', spec_wgt='trip_weight_2017_2019_v2021')
```



```{r}
write.csv(trips_by_mode, 'trips_by_mode_paine.csv')
```

repeat for 2021

``

Read in the travel survey data and do some work to clean up modes and times of day.
```{r}
mode_geo_df<-get_hhts('2021', "t", vars= c('trip_id','main_mode', 'travelers_total', 'origin_lat', 'origin_lng', 'dest_lat', 'dest_lng', 'mode_simple', 'dest_purpose_cat', 'origin_purpose_cat'))%>%
drop_na(c(origin_lat, dest_lat))%>%
mutate(mode_incl_hov = ifelse(mode_simple !='Drive', mode_simple, ifelse(travelers_total==1, 'SOV', 'HOV')))%>%
filter(dest_purpose_cat=='Work' | origin_purpose_cat=='Work')


```



Turn the trip ends into a geographic data layer.
```{r}
mode_geo_sf_origin<-mode_geo_df%>% st_as_sf(coords=c("origin_lng","origin_lat"), crs= "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
mode_geo_sf_dest<-mode_geo_df%>% st_as_sf(coords=c("dest_lng","dest_lat"), crs= "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")


```


Overlay the trip origins and destinations with the Seattle sectors so they can be put into them.
```{r}
sf::sf_use_s2(FALSE)
origin_sector_paine<- st_join(mode_geo_sf_origin, sector_layer, join=st_within)
dest_sector_paine<-st_join(mode_geo_sf_dest,sector_layer, join=st_within)


```

```{r}
origin_sector_paine<- as.data.frame(st_drop_geometry(origin_sector_paine))
dest_sector_paine<- as.data.frame(st_drop_geometry(dest_sector_paine))

```

```{r}
od_sector_paine<-merge(origin_sector_paine, dest_sector_paine, by='trip_id', suffixes=c('', '_dest'))%>% mutate(mode_nonm=ifelse((mode_incl_hov=='Walk' | mode_incl_hov=='Bike') , 'Non-Motorized', mode_incl_hov))%>%filter_at(vars('COUNTY', 'COUNTY_dest'), any_vars(!is.na(.)))
```

Summarize all trips by whether they are in the PM Peak, what mode they are, what Seattle sector the origin is in, and what Seattle sector the destination is in. Includes all times and places outside Seattle.
```{r}
trips_by_mode_2021<-hhts_count(od_sector_paine, group_vars=c('mode_nonm'), spec_wgt='trip_weight_2021_ABS_Panel_adult')
```

```{r}
write.csv(trips_by_mode_2021, 'trips_by_mode_paine_2021.csv')
```

