---
title: "trip-behaviors"
author: "suzanne"
date: "2024-04-18"
output:
  pdf_document: default
  html_document: default
---

Trips per person per day
by Purpose
by Mode

by Worker/Non-Worker
---
title: "trip info"
author: "suzanne"
date: "2024-03-20"
output:
  pdf_document: default
  html_document: default
---

## Load in Packages
IF YOU need to download some packages on github: travelSurveyTools, psrcelmer, and psrcplot.
For this use the syntax:
library(devtools)
devtools::install_github('rsgInc/travelSurveyTools')
devtools::install_github('psrc/psrcelmer')
devtools::install_github('psrc/psrcplot')

```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
source('../survey-23-preprocess.R')
```



## Read in data items from Elmer

The travelSurveyTools package expects everything to be a data.table, so run setDT all over the place.


```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad,survey_year, hh_weight
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category,
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, industry,office_available, commute_freq, employment
                from HHSurvey.v_persons_labels")
                
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,
person_id, mode_1, dest_purpose, travelers_total, survey_year, distance_miles,trip_weight 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)
setDT(day)
setDT(trip)
```
#race_category needs to be added to the codebook

# to do: add Seattle, Bellevue geographies

## Read in Codebook
```{r , echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```


## Set IDs as characters
I guess for joining?, functionalize convert all ids to charcets
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
day[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
day[, person_id := as.character(person_id)]
trip[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
trip[, day_id := as.character(day_id)]

trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
day <- day%>%mutate(survey_year=as.character(survey_year))
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```

## Make a new variable

If your variable is going to be a transformation of existing data in the codebook- for example grouping- you need to add the new variable and its new values to the internal to code, working codebook.


```{r}
variable_list<-add_variable(variable_list, 'worker','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='worker')
value_labels<-add_values_code(group_name='worker')
person<-grp_to_tbl(tbl=person, ungrouped_name='employment', grouped_name='worker')
person<-person%>%mutate(worker=if_else(is.na(worker), 'Not Worker', worker))
```


```{r}
variable_list<-add_variable(variable_list, 'dest_purpose_5','trip')
group_labels<-get_grouped_labels(group_id='group_3', group_name='dest_purpose_5')
value_labels<-add_values_code(group_name='dest_purpose_5')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='dest_purpose', grouped_name='dest_purpose_5')
```

```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```



```{r}
ids = c('hh_id', 'person_id', 'day_id', 'trip_id')
wts = c('hh_weight', 'person_weight', 'day_weight', 'trip_weight')



hts_data = list(hh = hh,
                person = person,
                day = day,
                trip = trip)
```

```{r}
dist_summ<-summarize_weighted(hts_data= hts_data,
                               summarize_var = 'distance_miles',
                               summarize_by = 'survey_year',
                               id_cols=ids,
                               wt_cols=wts,
                               wtname='trip_weight'
                               )$summary$wtd%>%filter(distance_miles!='Exactly 0')%>%
                                filter(survey_year!='2017')

```




```{r}
static_column_chart(t=dist_summ,y='prop', x='distance_miles', fill='survey_year' ,ylabel= 'Percent of Trips', xlabel='Trip Distance (miles)', dec=0)+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=14), axis.title.x=element_text(size=14))+ylim(0, 0.65)


```


```{r}
mean_dist_summ<-summarize_weighted(hts_data= hts_data,
                               summarize_var = 'distance_miles',
                               summarize_by = 'survey_year',
                               summarize_vartype='numeric',
                               id_cols=ids,
                               wt_cols=wts,
                               wtname='trip_weight'
                               )$summary$wtd%>%filter(survey_year!='2017')


```




```{r}
z_score<-1.645
mean_dist_summ<-mean_dist_summ%>%mutate(moe=z_score*mean_se)
static_column_chart(t=mean_dist_summ,y='mean', x='survey_year', fill='survey_year' ,ylabel= 'Mean Distance', xlabel='Survey Year', dec=0, moe='moe')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=14), axis.title.x=element_text(size=14))+ylim(0, 6)

```


```{r}


```




```{r}

static_column_chart(t=mean_dist_summ,y='median', x='survey_year', fill='survey_year' ,ylabel= 'Median Distance', xlabel='Survey Year', dec=1)+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=14), axis.title.x=element_text(size=14))+ylim(0, 3.5)

```

```{r}

triprate_prep<-hts_prep_triprate(summarize_by = c('survey_year'),
                    variables_dt = variable_list,
                    hts_data = hts_data,
                    ids = ids,
                    wts = wts,
                    weighted = TRUE,
                    remove_outliers=FALSE)


```
```{r}
output <- hts_summary(triprate_prep$num,
                       summarize_var = 'num_trips_wtd',
                       summarize_by = c('survey_year'),
                       summarize_vartype = 'numeric',
                       id_cols = ids,
                       wtname = 'day_weight')
```

```{r}
 wtd_triprate = output$summary$wtd[, 
                                    .(
                                      summarize_by = c(get('survey_year')),
                                      'Weighted Trip Rate' = round(mean, 2)
                                    )
  ]
```
```{r}
wtd_triprate

wtd_trip_rate<-output$summary$wtd%>%filter(survey_year!='2017')
```


```{r}
static_column_chart(t=wtd_trip_rate,y='mean', x='survey_year', fill='survey_year' ,ylabel= 'Trips per Day', xlabel='Year', dec=1)+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+ylim(0, 5)


```



```{r}
triprate_prep<-hts_prep_triprate(summarize_by = c('survey_year', 'worker'),
                    variables_dt = variable_list,
                    hts_data = hts_data,
                    ids = ids,
                    wts = wts,
                    weighted = TRUE,
                    remove_outliers=FALSE)


```

```{r}
output <- hts_summary(triprate_prep$num,
                       summarize_var = 'num_trips_wtd',
                       summarize_by = c('survey_year','worker'),
                       summarize_vartype = 'numeric',
                       id_cols = ids,
                       wtname = 'day_weight')
```



```{r}
wtd_trip_rate<-output$summary$wtd%>%filter(survey_year!='2017')
```


```{r}
static_line_chart(t=wtd_trip_rate,y='mean', x='survey_year', fill='worker' ,ylabel= 'Trips per Day', xlabel='Year')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))+ylim(0, 5)

```
```{r}

hts_data_23 = list(hh = hh%>%filter(survey_year=='2023'),
                person = person%>%filter(survey_year=='2023'),
                day = day%>%filter(survey_year=='2023'),
                trip = trip%>%filter(survey_year=='2023'))

```


```{r}
triprate_prep<-hts_prep_triprate(summarize_by = c('dest_purpose_5'),
                    variables_dt = variable_list,
                    hts_data = hts_data_23,
                    ids = ids,
                    wts = wts,
                    weighted = TRUE,
                    remove_outliers=FALSE)


```

```{r}
output <- hts_summary(triprate_prep$num,
                       summarize_var = 'num_trips_wtd',
                       summarize_by = c('dest_purpose_5'),
                       summarize_vartype = 'numeric',
                       id_cols = ids,
                       wtname = 'day_weight',
                       se=TRUE )
```





```{r}
wtd_trip_rate<-output$summary$wtd
wtd_trip_rate_23<-output$summary$wtd%>%mutate(times=5*mean)%>%filter(dest_purpose_5!='NA')%>%
  mutate(dest_purpose_5= fct_reorder(dest_purpose_5,-times))%>%mutate(survey_year='2023')
```


```{r}

static<-static_column_chart(wtd_trip_rate, x='dest_purpose_5', fill='dest_purpose_5', y='mean', dec=, color='pgnobgy_10')+
  theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
static
write.csv(wtd_trip_rate, 'trip_rate_purpose.csv')
```

```{r}

hts_data_17 = list(hh = hh%>%filter(survey_year=='2017'),
                person = person%>%filter(survey_year=='2017'),
                day = day%>%filter(survey_year=='2017'),
                trip = trip%>%filter(survey_year=='2017'))

```


```{r}
triprate_prep<-hts_prep_triprate(summarize_by = c('dest_purpose_5'),
                    variables_dt = variable_list,
                    hts_data = hts_data_17,
                    ids = ids,
                    wts = wts,
                    weighted = TRUE,
                    remove_outliers=FALSE)


```

```{r}
output <- hts_summary(triprate_prep$num,
                       summarize_var = 'num_trips_wtd',
                       summarize_by = c('dest_purpose_5'),
                       summarize_vartype = 'numeric',
                       id_cols = ids,
                       wtname = 'day_weight',
                       se=TRUE )
```



```{r}
wtd_trip_rate_17<-output$summary$wtd
wtd_trip_rate_17<-output$summary$wtd%>%mutate(times=5*mean)%>%filter(dest_purpose_5!='NA')%>%
                                                                    
  mutate(dest_purpose_5= fct_reorder(dest_purpose_5,-times))%>%mutate(survey_year='2017')
```



```{r}

hts_data_19 = list(hh = hh%>%filter(survey_year=='2019'),
                person = person%>%filter(survey_year=='2019'),
                day = day%>%filter(survey_year=='2019'),
                trip = trip%>%filter(survey_year=='2019'))

```


```{r}
triprate_prep<-hts_prep_triprate(summarize_by = c('dest_purpose_5'),
                    variables_dt = variable_list,
                    hts_data = hts_data_19,
                    ids = ids,
                    wts = wts,
                    weighted = TRUE,
                    remove_outliers=FALSE)


```

```{r}
output <- hts_summary(triprate_prep$num,
                       summarize_var = 'num_trips_wtd',
                       summarize_by = c('dest_purpose_5'),
                       summarize_vartype = 'numeric',
                       id_cols = ids,
                       wtname = 'day_weight',
                       se=TRUE )
```



```{r}
wtd_trip_rate_19<-output$summary$wtd
wtd_trip_rate_19<-output$summary$wtd%>%mutate(times=5*mean)%>%filter(dest_purpose_5!='NA')%>%
                                                                    
  mutate(dest_purpose_5= fct_reorder(dest_purpose_5,-times))%>%mutate(survey_year='2019')
```




```{r}

hts_data_21 = list(hh = hh%>%filter(survey_year=='2021'),
                person = person%>%filter(survey_year=='2021'),
                day = day%>%filter(survey_year=='2021'),
                trip = trip%>%filter(survey_year=='2021'))

```


```{r}
triprate_prep<-hts_prep_triprate(summarize_by = c('dest_purpose_5'),
                    variables_dt = variable_list,
                    hts_data = hts_data_21,
                    ids = ids,
                    wts = wts,
                    weighted = TRUE,
                    remove_outliers=TRUE)


```

```{r}
output <- hts_summary(triprate_prep$num,
                       summarize_var = 'num_trips_wtd',
                       summarize_by = c('dest_purpose_5'),
                       summarize_vartype = 'numeric',
                       id_cols = ids,
                       wtname = 'day_weight',
                       se=TRUE )
```



```{r}
wtd_trip_rate_21<-output$summary$wtd
wtd_trip_rate_21<-output$summary$wtd%>%mutate(times=5*mean)%>%filter(dest_purpose_5!='NA')%>%
  mutate(dest_purpose_5= fct_reorder(dest_purpose_5,-times))%>%mutate(survey_year='2021')
```

```{r}
z_score<-1.65
wtd_trip_rate_all<-rbind(wtd_trip_rate_17,wtd_trip_rate_19, wtd_trip_rate_21, wtd_trip_rate_23)%>%
  mutate(moe_time=5*z_score*mean_se)%>%filter(dest_purpose_5!='Missing')

```

```{r}
static<-static_column_chart(wtd_trip_rate_all, x='survey_year', y='mean', fill='dest_purpose_5', pos='stack', dec=2, color='pgnobgy_10') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

static
```
```{r}
z_score<-1.65
wtd_trip_rate_shopping<-rbind(wtd_trip_rate_17,wtd_trip_rate_19, wtd_trip_rate_21, wtd_trip_rate_23)%>%
  mutate(mean_moe=z_score*mean_se)%>%filter(dest_purpose_5!='Missing')%>%filter(dest_purpose_5=='Shopping')

```

```{r}
static<-static_column_chart(wtd_trip_rate_shopping, x='survey_year', y='mean', fill='survey_year', pos='stack', est="number", dec=2, color='pgnobgy_10')+
#+geom_errorbar(aes(ymin=mean-mean_moe, ymax=mean+mean_moe),
#                 width=0.2, position = position_dodge(0.9)) 
theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

static
```

work school
```{r}
wtd_trip_rate_mand<-wtd_trip_rate_all%>%filter(dest_purpose_simple %in% c('Primary Work', 'School', 'Work-related'))

```


break down by primary work and other purpose

```{r}

static_facet_column_chart(t=wtd_trip_rate_mand,y='times', x='survey_year', fill='survey_year' ,ylabel= 'Activity Type', xlabel='Times per Work Week went to Do Activity', moe='moe_time', facet='dest_purpose_simple',dec=1,scales='fixed')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
```
```{r}
wtd_trip_rate_soc<-wtd_trip_rate_all%>%filter(dest_purpose_simple %in% c('Social/Recreation', 'Eat Meal'))

```




```{r}

static_facet_column_chart(t=wtd_trip_rate_soc,y='times', x='survey_year', fill='survey_year' ,ylabel= 'Activity Type', xlabel='Times per Work Week went to Do Activity', moe='moe_time', facet='dest_purpose_simple',dec=1,scales='fixed')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
```
```{r}
wtd_trip_rate_errand<-wtd_trip_rate_all%>%filter(dest_purpose_simple %in% c('Errands/Appointments/Other', 'Pick up/Drop off', 'Shopping'))

```




```{r}

static_facet_column_chart(t=wtd_trip_rate_errand,y='times', x='survey_year', fill='survey_year' ,ylabel= 'Activity Type', xlabel='Times per Work Week went to Do Activity', moe='moe_time', facet='dest_purpose_simple',dec=1,scales='fixed')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
```


```{r}
mean_dist_summ<-summarize_weighted(hts_data= hts_data,
                               summarize_var = 'distance_miles',
                               summarize_by = c('survey_year', 'dest_purpose_simple'),
                               summarize_vartype='numeric',
                               id_cols=ids,
                               wt_cols=wts,
                               wtname='trip_weight'
                               )$summary$wtd%>%filter(survey_year!='2017')


```




```{r}
z_score<-1.645
mean_dist_summ<-mean_dist_summ%>%mutate(moe=z_score*mean_se)
chart<-static_facet_column_chart(t=mean_dist_summ,y='mean', x='survey_year', fill='dest_purpose_simple' ,ylabel= 'Mean Distance', xlabel='Trip Destination Purpose', dec=0, moe='moe', color='pgnobgy_10', facet='dest_purpose_simple')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=14), axis.title.x=element_text(size=14))

```
