---
title: "trip-behaviors"
author: "suzanne"
date: "2024-04-18"
output:
  pdf_document: default
  html_document: default
---

Trips per person per day
by Purpose
by Mode

by Worker/Non-Worker
---
title: "first-example-mode-summary"
author: "suzanne"
date: "2024-03-20"
output:
  pdf_document: default
  html_document: default
---

## Load in Packages
IF YOU need to download some packages on github: travelSurveyTools, psrcelmer, and psrcplot.
For this use the syntax:
library(devtools)
devtools::install_github('rsgInc/travelSurveyTools')
devtools::install_github('psrc/psrcelmer')
devtools::install_github('psrc/psrcplot')

```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
source('survey-23-preprocess.R')
```

## Debugging travelSurveyTools

travelSurveyTools still has many cryptic error messages usually related to data or codebook formatting and naming. You may need to debug to find where or why the problem exists.

You can use: devtools::load_all('C:/GitHub/travelSurveyTools/R') and then browser() and put breakpoints in the code to find the problem.


## Read in data items from Elmer

The travelSurveyTools package expects everything to be a data.table, so run setDT all over the place.


```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad,survey_year, hh_weight
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category,
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, industry,office_available, commute_freq, employment
                from HHSurvey.v_persons_labels")
                
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,
person_id, mode_1, dest_purpose, travelers_total, survey_year, trip_weight 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)
setDT(day)
setDT(trip)
```
#race_category needs to be added to the codebook

# to do: add Seattle, Bellevue geographies

## Read in Codebook
```{r , echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```


## Set IDs as characters
I guess for joining?, functionalize convert all ids to charcets
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
day[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
day[, person_id := as.character(person_id)]
trip[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
trip[, day_id := as.character(day_id)]

trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
day <- day%>%mutate(survey_year=as.character(survey_year))
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```

## Make a new variable

If your variable is going to be a transformation of existing data in the codebook- for example grouping- you need to add the new variable and its new values to the internal to code, working codebook.


#make this even easier?
```{r}
variable_list<-add_variable(variable_list, 'worker','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='worker')
value_labels<-add_values_code(group_name='worker')
person<-grp_to_tbl(tbl=person, ungrouped_name='employment', grouped_name='worker')
person<-person%>%mutate(worker=if_else(is.na(worker), 'Not Worker', worker))
```



```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```



```{r}
ids = c('hh_id', 'person_id', 'day_id', 'trip_id')
wts = c('hh_weight', 'person_weight', 'day_weight', 'trip_weight')



hts_data = list(hh = hh,
                person = person,
                day = day,
                trip = trip)
```

```{r}
triprate_prep<-hts_prep_triprate(summarize_by = c('survey_year'),
                    variables_dt = variable_list,
                    hts_data = hts_data,
                    ids = ids,
                    wts = wts,
                    weighted = TRUE,
                    remove_outliers=FALSE)


```
```{r}
output <- hts_summary(triprate_prep$num,
                       summarize_var = 'num_trips_wtd',
                       summarize_by = c('survey_year'),
                       summarize_vartype = 'numeric',
                       id_cols = ids,
                       wtname = 'day_weight')
```

```{r}
 wtd_triprate = output$summary$wtd[, 
                                    .(
                                      summarize_by = c(get('survey_year')),
                                      'Weighted Trip Rate' = round(mean, 2)
                                    )
  ]
```
```{r}
wtd_triprate
```

```{r}
triprate_prep<-hts_prep_triprate(summarize_by = c('survey_year', 'worker'),
                    variables_dt = variable_list,
                    hts_data = hts_data,
                    ids = ids,
                    wts = wts,
                    weighted = TRUE,
                    remove_outliers=FALSE)


```

```{r}
output <- hts_summary(triprate_prep$num,
                       summarize_var = 'num_trips_wtd',
                       summarize_by = c('survey_year','worker'),
                       summarize_vartype = 'numeric',
                       id_cols = ids,
                       wtname = 'day_weight')
```

```{r}
 wtd_triprate = output$summary$wtd[, 
                                    .(
                                      summarize_by = c(get('survey_year', 'worker')),
                                      'Weighted Trip Rate' = round(mean, 2)
                                    )
  ]
```

```{r}
wtd_triprate
```

```{r}
```{r}
static_line_chart(t=wtd_triprate,y='prop', x='survey year', fill='worker' ,ylabel= 'Trips per Day', xlabel='Year')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

```
```

