---
title: "trip-behaviors"
author: "suzanne"
date: "2024-04-18"
output:
  pdf_document: default
  html_document: default
---

Trips per person per day
by Purpose
by Mode

by Worker/Non-Worker
---
title: "trip info"
author: "suzanne"
date: "2024-03-20"
output:
  pdf_document: default
  html_document: default
---

## Load in Packages
IF YOU need to download some packages on github: travelSurveyTools, psrcelmer, and psrcplot.
For this use the syntax:
library(devtools)
devtools::install_github('rsgInc/travelSurveyTools')
devtools::install_github('psrc/psrcelmer')
devtools::install_github('psrc/psrcplot')

```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
source('../survey-23-preprocess.R')
```



## Read in data items from Elmer

The travelSurveyTools package expects everything to be a data.table, so run setDT all over the place.


```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad,survey_year, hh_weight
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category,
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, industry,office_available, commute_freq, employment
                from HHSurvey.v_persons_labels")
                
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,
person_id, mode_1, dest_purpose, travelers_total, survey_year, distance_miles,trip_weight 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)
setDT(day)
setDT(trip)
```
#race_category needs to be added to the codebook

# to do: add Seattle, Bellevue geographies

## Read in Codebook
```{r , echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```


## Set IDs as characters
I guess for joining?, functionalize convert all ids to charcets
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
day[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
day[, person_id := as.character(person_id)]
trip[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
trip[, day_id := as.character(day_id)]

trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
day <- day%>%mutate(survey_year=as.character(survey_year))
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```

## Make a new variable

If your variable is going to be a transformation of existing data in the codebook- for example grouping- you need to add the new variable and its new values to the internal to code, working codebook.



```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```



```{r}
ids = c('hh_id', 'person_id', 'day_id', 'trip_id')
wts = c('hh_weight', 'person_weight', 'day_weight', 'trip_weight')



hts_data = list(hh = hh,
                person = person,
                day = day,
                trip = trip)
```


```{r}

triprate_prep<-hts_prep_triprate(summarize_by = c('survey_year'),
                    variables_dt = variable_list,
                    hts_data = hts_data,
                    ids = ids,
                    wts = wts,
                    weighted = TRUE,
                    remove_outliers=FALSE)


```
```{r}
output <- hts_summary(triprate_prep$num,
                       summarize_var = 'num_trips_wtd',
                       summarize_by = c('survey_year'),
                       summarize_vartype = 'numeric',
                       id_cols = ids,
                       wtname = 'day_weight')
```

```{r}
 wtd_triprate = output$summary$wtd[, 
                                    .(
                                      summarize_by = c(get('survey_year')),
                                      'Weighted Trip Rate' = round(mean, 2)
                                    )
  ]
```

```{r}
 wtd_triprate 

```

```{r}
 unwtd_triprate = output$summary$unwtd[, 
                                    .(
                                      summarize_by = c(get('survey_year')),
                                      'Unweighted Trip Rate' = round(mean, 2)
                                    )
  ]
unwtd_triprate
```

wtd_trip_rate<-output$summary$wtd%>%filter(survey_year!='2017')
```


