---
title: "trip-behaviors"
author: "suzanne"
date: "2024-04-18"
output:
  pdf_document: default
  html_document: default
---

Trips per person per day
by Purpose
by Mode

by Worker/Non-Worker
---
title: "trip info"
author: "suzanne"
date: "2024-03-20"
output:
  pdf_document: default
  html_document: default
---

## Load in Packages
IF YOU need to download some packages on github: travelSurveyTools, psrcelmer, and psrcplot.
For this use the syntax:
library(devtools)
devtools::install_github('rsgInc/travelSurveyTools')
devtools::install_github('psrc/psrcelmer')
devtools::install_github('psrc/psrcplot')

```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
source('../survey-23-preprocess.R')
```



## Read in data items from Elmer

The travelSurveyTools package expects everything to be a data.table, so run setDT all over the place.


```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad,survey_year, hh_weight
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category,
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, industry,office_available, commute_freq, employment
                from HHSurvey.v_persons_labels")
                
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,
person_id, mode_1, dest_purpose, travelers_total, survey_year, distance_miles,origin_jurisdiction, dest_jurisdiction,trip_weight 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)
setDT(day)
setDT(trip)
```
#race_category needs to be added to the codebook

# to do: add Seattle, Bellevue geographies

## Read in Codebook
```{r , echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```


## Set IDs as characters
I guess for joining?, functionalize convert all ids to charcets
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
day[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
day[, person_id := as.character(person_id)]
trip[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
trip[, day_id := as.character(day_id)]

trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
day <- day%>%mutate(survey_year=as.character(survey_year))
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```

## Make a new variable

If your variable is going to be a transformation of existing data in the codebook- for example grouping- you need to add the new variable and its new values to the internal to code, working codebook.


```{r}
variable_list<-add_variable(variable_list, 'worker','person')
group_labels<-get_grouped_labels(group_id='group_1', group_name='worker')
value_labels<-add_values_code(group_name='worker')
person<-grp_to_tbl(tbl=person, ungrouped_name='employment', grouped_name='worker')
person<-person%>%mutate(worker=if_else(is.na(worker), 'Not Worker', worker))
```


```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='dest_purpose_simple')
value_labels<-add_values_code(group_name='dest_purpose_simple')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='dest_purpose', grouped_name='dest_purpose_simple')
```

```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```



```{r}
ids = c('hh_id', 'person_id', 'day_id', 'trip_id')
wts = c('hh_weight', 'person_weight', 'day_weight', 'trip_weight')

trip<-trip%>%mutate(Seattle_trip=
                      if_else((origin_jurisdiction=='Seattle' | dest_jurisdiction=='Seattle'), 'Seattle', 'Not Seattle'))

hts_data = list(hh = hh,
                person = person,
                day = day,
                trip = trip)
```

```{r}
variable_list<-add_variable(variable_list, 'Seattle_trip', trip)
```

```{r}

value<- c(0,1)
label=c('Seattle', 'Not Seattle')
val_labels<-data.frame(value, label)
val_labels<-val_labels%>%mutate(variable='Seattle_trip')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  

```

```{r}
dist_summ<-summarize_weighted(hts_data= hts_data,
                               summarize_var = 'distance_miles',
                               summarize_by = c('survey_year', 'Seattle_trip'),
                               id_cols=ids,
                               wt_cols=wts,
                               wtname='trip_weight'
                               )$summary$wtd%>%filter(distance_miles!='Exactly 0')%>%
                                filter(survey_year!='2017')

```




```{r}
static_column_chart(t=dist_summ,y='prop', x='distance_miles', fill='survey_year' ,ylabel= 'Percent of Trips', xlabel='Trip Distance (miles)', dec=0)+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=14), axis.title.x=element_text(size=14))+ylim(0, 0.65)


```


```{r}
mean_dist_summ<-summarize_weighted(hts_data= hts_data,
                               summarize_var = 'distance_miles',
                               summarize_by = c('survey_year', 'Seattle_trip'),
                               summarize_vartype='numeric',
                               id_cols=ids,
                               wt_cols=wts,
                               wtname='trip_weight'
                               )$summary$wtd%>%filter(survey_year!='2017')


```




```{r}
z_score<-1.645
mean_dist_summ<-mean_dist_summ%>%mutate(moe=z_score*mean_se)
static_column_chart(t=mean_dist_summ,y='mean', x='survey_year', fill='Seattle_trip' ,ylabel= 'Mean Distance', xlabel='Survey Year', dec=0, moe='moe')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=14), axis.title.x=element_text(size=14))+ylim(0, 6)

```







```{r}

static_column_chart(t=mean_dist_summ,y='median', x='survey_year', fill='Seattle_trip' ,ylabel= 'Median Distance', xlabel='Survey Year', dec=1)+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=14), axis.title.x=element_text(size=14))+ylim(0, 3.5)

```




```{r}
trip_summ<-summarize_weighted(hts_data= hts_data,
                               summarize_var = 'Seattle_trip',
                               summarize_by =c('survey_year'),
                               id_cols=ids,
                               wt_cols=wts,
                               wtname='trip_weight'
                               )$summary$wtd%>%drop_na(Seattle_trip)%>%drop_na(survey_year)

```


```{r}
trip_summ<-trip_summ%>%filter(survey_year!='2017' & survey_year !='2021')%>%mutate(rounded_est=round(est, -4))
static_column_chart(t=trip_summ,y='rounded_est', x='Seattle_trip', fill='survey_year', ylabel= 'Number of Trips', xlabel='Survey Year',  color='pognbgy_10', dec=-1)+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=14), axis.title.x=element_text(size=14))

```

# Add associated values
```{r}
variable_list<-add_variable(variable_list, 'mode_w_sov','trip')
```

```{r}
variable_list<-add_variable(variable_list, 'mode_simple','trip')
group_labels<-get_grouped_labels(group_id='group_1', group_name='mode_simple')
value_labels<-add_values_code(group_name='mode_simple')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='mode_1', grouped_name='mode_simple')
```

```{r}

value<- c(0,1,2,3,4,5)
label=c('SOV', 'HOV', 'Walk', 'Transit', 'Bike', 'Other')
val_labels<-data.frame(value, label)
val_labels<-val_labels%>%mutate(variable='mode_w_sov')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  

```

```{r}

value<- c(0,1,2,3,4,5)
label=c('SOV', 'Not SOV')
val_labels<-data.frame(value, label)
val_labels<-val_labels%>%mutate(variable='mode_sov_bin')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  

```

```{r}
variable_list<-add_variable(variable_list, 'mode_sov_bin', trip)
```

# add mode_w_sov to the trip table
```{r}
trip<- trip%>%mutate(travelers_total_fix= ifelse(travelers_total!='1 traveler', 'More than 1', '1 traveler'))
trip<-trip%>%mutate(mode_simple= replace_na(mode_simple, 'Drive'))%>%
mutate(mode_w_sov=case_when(
  mode_simple=="Drive"& travelers_total=='1 traveler' ~ 'SOV',
  is.na(travelers_total) ~ 'SOV',
  mode_simple=="Drive"& travelers_total!='1 traveler'~  'HOV',
  .default= mode_simple
))%>%mutate(mode_sov_bin=if_else(mode_w_sov=='SOV', 'SOV', 'Not SOV'))
```


The package expects the data to be in a list of data.tables.
```{r}

hts_data = list(#hh = hh,
                #person = person,
                #day = day,
                trip = trip)
```


```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```



##summarize data
```{r}

mode_summary = summarize_weighted(hts_data= hts_data,
                               summarize_var = 'mode_sov_bin',
                               summarize_by = c('survey_year', 'Seattle_trip'),
                               id_cols='trip_id',
                               wt_cols='trip_weight',
                               wtname='trip_weight'
                               )
```


```{r}
z_score=1.645

common_modes<-mode_summary$summary$wtd%>%mutate(moe=prop_se*z_score)%>%mutate(moe_est=est_se*z_score)%>%
  mutate(mode_sov_bin= fct_reorder(mode_sov_bin,-prop))%>%filter(survey_year!='2017')%>%drop_na(Seattle_trip)%>%mutate(mode_sov_bin=case_match(mode_sov_bin, 
                                                                                                                                               "SOV" ~ "Drive Alone",
                                                    "Not SOV" ~ "Carpooling, Walking, Biking ,Transit" ))%>%filter(survey_year!='2021')


static_facet_column_chart(common_modes, x='Seattle_trip', y='est', fill='survey_year',ylabel= 'Share', xlabel='Trip Mode', facet='mode_sov_bin',moe='moe_est', scales='fixed',color='pognbgy_10')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
```
##summarize data
```{r}

mode_summary = summarize_weighted(hts_data= hts_data,
                               summarize_var = 'mode_w_sov',
                               summarize_by = c('survey_year', 'Seattle_trip'),
                               id_cols='trip_id',
                               wt_cols='trip_weight',
                               wtname='trip_weight'
                               )
```


## Visualize
```{r}
z_score=1.645

common_modes<-mode_summary$summary$wtd%>%mutate(moe=prop_se*z_score)%>%mutate(moe_est=est_se*z_score)%>%
  mutate(mode_w_sov= fct_reorder(mode_w_sov,-prop))%>%filter(survey_year!='2017')%>%drop_na(Seattle_trip)

sov_modes<-common_modes%>%filter(mode_w_sov=='SOV')%>%filter(survey_year!='2017' & survey_year !='2021')

static_column_chart(sov_modes,x='Seattle_trip', y='est', fill='survey_year',ylabel= 'Number of Trips', xlabel='SOV trips', moe='moe_est')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))


walk_transit_modes<-common_modes%>%filter(mode_w_sov %in% c('Walk', 'Transit'))%>%filter(!survey_year %in% c('2017', '2021'))%>%filter(survey_year!='2017' & survey_year !='2021')

static_facet_column_chart(walk_transit_modes,x='Seattle_trip', y='est', facet='mode_w_sov', fill='survey_year',ylabel= 'Number of Trips', xlabel='Trip Mode', moe='moe_est', scales='fixed')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

static_facet_column_chart(common_modes, x='Seattle_trip', y='prop', fill='survey_year',ylabel= 'Share', xlabel='Trip Mode', facet='mode_w_sov', moe='moe')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))



static_facet_column_chart(common_modes, x='Seattle_trip', y='est', fill='survey_year',ylabel= 'Share', xlabel='Trip Mode', facet='mode_w_sov',moe='moe_est')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
```

```{r}
drive_trip<-trip%>%filter(mode_simple=='Drive')

hts_data = list(#hh = hh,
                #person = person,
                #day = day,
                trip = drive_trip)

mode_travelers_summary = summarize_weighted(hts_data= hts_data,
                               summarize_var = 'travelers_total',
                               summarize_by = 'survey_year',
                               id_cols='trip_id',
                               wt_cols='trip_weight',
                               wtname='trip_weight'
                               )
```

## Visualize
```{r}

mode_travelers<-mode_travelers_summary$summary$wtd

static_column_chart(mode_travelers, x='travelers_total', y='prop', fill='survey_year')

```
