---
title: "TDM and Travel Behavior"
output: html_notebook
---
The TDM team noted the following, so I thought we could look into relationships between travel behavior and TDM.

“More data is needed to evaluate the effectiveness and efficiency of TDM programs across the region and of individual strategies implemented by TDM programs. Effectiveness measures how well strategies and/or programs are achieving their stated goals using quantitative metrics. Efficiency determines how well funds are spent on TDM programming. (i.e., What is the return on investment for dollars spent on TDM programs? What are its impacts?) ii. It is difficult to directly relate reductions in greenhouse gas (GHG) emissions or vehicle miles traveled (VMT) or improved roadway level of service (LOS) directly to TDM efforts. Further data is needed to evaluate causality and measure TDM’s impacts in the region.”

Ideas:
I can build a commute mode choice model that includes accessibility values from skims, or not? I could just build a basic commute mode choice model.

I can build a model of vmt for workers, with the following components: 
* household income
* household size
* number of children
* characteristics of home geography
  -number of jobs within 30 min
  -parcel attributes(household density)
* auto travel time from home to work
  



```{r libraries}
# You will probably need to install a few libraries to get this to work.
source('travel_survey_analysis_functions.R')
library(data.table)
library(tidyverse)
library(DT)
library(openxlsx)
library(odbc)
library(DBI)
library(fastDummies)
library(aod)
library(BMA)
library(MASS)
library(jtools)
library(ggstance)
library(sjPlot)
library(effects)
library(DescTools)


```



```{r read_in_data}
#The codebook is checked in with these code files, named Combined_Codebook_022020.xlsx


################# Input Data File Reading
# This file contains information by Census Tract such as percent of households https://www.psrc.org/displacement-risk-mapping
# in poverty, coming from the displacement risk analysis work.
# it is checked in on github. You will need to point this variable to where it is
# on your computer.


# This commented out file contains very detailed information about parcels.  For example,
# it contains information for each parcel about the number of jobs within a half mile.
# It is nearly 1 GB. You may wish to move this file locally for speed reasons. This is 2018 parcel data
#parcel_data<- 'C:/Users/SChildress/Documents/HHSurvey/displace_estimate/buffered_parcels.txt'
# I've moved my locally, as you can see:
parcel_data <- 'C:/Users/SChildress/Documents/HHSurvey/displace_estimate/buffered_parcels.dat'
displ_index_data<- 'J:/Projects/Surveys/HHTravel/Survey2019/Analysis/displacement/estimation_files/displacement_risk_estimation.csv'




## Read person-displacement data from Elmer, other travel survey data as well
db.connect <- function() {
  elmer_connection <- dbConnect(odbc(),
                                driver = "SQL Server",
                                server = "AWS-PROD-SQL\\SOCKEYE",
                                database = "Elmer",
                                trusted_connection = "yes"
  )
}

read.dt <- function(astring, type =c('table_name', 'sqlquery')) {
  elmer_connection <- db.connect()
  if (type == 'table_name') {
    dtelm <- dbReadTable(elmer_connection, SQL(astring))
  } else {
    dtelm <- dbGetQuery(elmer_connection, SQL(astring))
  }
  dbDisconnect(elmer_connection)
  setDT(dtelm)
}


dbtable.person.query<- paste("SELECT *  FROM HHSurvey.v_persons_2017_2019_in_house")
person_dt<-read.dt(dbtable.person.query, 'tablename')
parcel_df<-read.table(parcel_data, header=TRUE, sep='')
displ_risk_df <- read.csv(displ_index_data)
```


```{r merge_data}
person_dt<-person_dt %>% filter(worker != 'No jobs')

missing_codes <- c('Missing: Technical Error', 'Missing: Non-response', 
                   'Missing: Skip logic', 'Children or missing')


# switch over to using df syntax for simplicity
person_df <- setDF(person_dt)


person_df$census_2010_tract <- as.character(person_df$final_home_tract)

#join people to the displacement risk data
person_df_dis <- merge(person_df,displ_risk_df, by.x='census_2010_tract', by.y='GEOID', all.x=TRUE)
```


```{r merge_data}

# a list of parcel -based variables I'd like to try in the model, there are more on the file
parcel_based_vars<-c('hh_2', 'stugrd_2', 'stuhgh_2', 'stuuni_2', 'empedu_2', 'empfoo_2', 'empgov_2', 'empind_2',
                      'empmed_2', 'empofc_2', 'empret_2', 'empsvc_2', 'emptot_2', 'ppricdy2', 'pprichr2',
                     'tstops_2', 'nparks_2', 'aparks_2', 'dist_lbus', 'dist_ebus', 'dist_crt', 'dist_fry',
                      'dist_lrt')

person_df_dis$parcel_id = as.character(person_df_dis$parcel_id)

parcel_df$parcelid <- as.character(parcel_df$parcelid)
#merging by the previous residence parcel from the person table and by the parcel id in the parcel table
person_df_dis_parcel<- merge(person_df_dis, parcel_df, by.x='parcel_id', by.y='parcelid', all.x = TRUE)

```
```{r}
# build a simple commute mode choice model


```


do some summaries commute mode choice by transportation benefits
```{r}

person_no_na <- person_df_dis_parcel %>% filter(!is.na(commute_mode))

create_table_one_var("commute_mode", person_no_na,"person" )

create_table_one_var("benefits_1", person_no_na,"person" )

create_table_one_var("benefits_2", person_no_na,"person" )
create_table_one_var("benefits_3", person_no_na,"person" )
create_table_one_var("benefits_4", person_no_na,"person" )



```
