---
title: 'Household Travel Survey: Commuting Data'
author: "Suzanne Childress & Mary Richards"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 6
    toc_float: true
  word_document: default
  pdf_document: default
---

This report presents preliminary findings for commuting trends based on the Household Travel Survey (2017 and 2019), a survey conducted every other year to determine travel patterns for individuals living within PSRC's four-county jurisdiction.

```{r global-options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE) # formatting
```

```{r libraries, include=FALSE}
# source('travel_survey_analysis_functions.R')
library(tidyverse)
library(openxlsx)
library(odbc)
library(DBI)
library(dplyr)
library(ggplot2)

library(table1)
library(knitr)
library(kableExtra)
library(summarytools)
```

```{r functions, include=FALSE}
# connecting to Elmer
db.connect <- function() {
  elmer_connection <- dbConnect(odbc(),
                                driver = "SQL Server",
                                server = "AWS-PROD-SQL\\Sockeye",
                                database = "Elmer",
                                trusted_connection = "yes"
  )
}

# a function to read tables and queries from Elmer
read.dt <- function(astring, type =c('table_name', 'sqlquery')) {
  elmer_connection <- db.connect()
  if (type == 'table_name') {
    dtelm <- dbReadTable(elmer_connection, SQL(astring))
  } else {
    dtelm <- dbGetQuery(elmer_connection, SQL(astring))
  }
  dbDisconnect(elmer_connection)
  dtelm
}

# Create a simplified crosstab from one variable, calculate counts, totals, shares, and MOE for categorical data
create_table_one_var_simp= function(var1, table_temp, table_type) {
  #table_temp = recategorize_var_upd(var2,table_temp)
  #print(table_temp)
  if (table_type == "household" | table_type == "person" ) {
    weight_2017 = "hh_wt_revised"
    weight_2019 = "hh_wt_2019"
    weight_comb = "hh_wt_combined"
  } else if (table_type == "trip") {
    weight_2017 = "trip_weight_revised"
    weight_2019 = "trip_wt_2019"
    weight_comb = "trip_wt_combined"  
  } 
  
  temp = table_temp %>% 
    dplyr::select(!!sym(var1), all_of(weight_2019)) %>% 
    filter(!.[[1]] %in% missing_codes, !is.na(.[[1]])) %>% 
    group_by(!!sym(var1)) %>% 
    summarise(SampleSize=n(),
              Weighted_2019 = sum(.data[[weight_2019]],na.rm = TRUE), 
              Weighted_2019 = round(Weighted_2019, 0)) %>% 
    mutate(WeightedPercent = Weighted_2019/sum(Weighted_2019)*100, 
           WeightedPercent = round(WeightedPercent, 2)) %>% 
    ungroup() %>%  
    mutate(MOE=1.65*(0.25/sum(SampleSize))^(1/2)*100, 
           MOE=round(MOE, 2)) %>% 
    arrange(var1)
  return(temp)
}

# Create a crosstab from two variables, calculate counts, totals, and shares for categorical data
cross_tab_categorical <- function(table, var1, var2, wt_field) {
  expanded <- table %>% 
    group_by(.data[[var1]],.data[[var2]]) %>%
    dplyr::summarize(Count= n(),Total=sum(.data[[wt_field]])) %>%
    group_by(.data[[var1]])%>%
    mutate(Percentage=Total/sum(Total)*100)
  
  
  expanded_pivot <-expanded%>%
    pivot_wider(names_from=.data[[var2]], values_from=c(Percentage,Total, Count))
  
  return (expanded_pivot)
  
} 

# Create margins of error for dataset
categorical_moe <- function(sample_size_group){
  sample_w_MOE<-sample_size_group %>%
    mutate(p_col=p_MOE) %>%
    mutate(MOE_calc1= (p_col*(1-p_col))/sample_size) %>%
    mutate(MOE_Percent=z*sqrt(MOE_calc1)*100)
  
  sample_w_MOE<- dplyr::select(sample_w_MOE, -c(p_col, MOE_calc1))
  
  return(sample_w_MOE)
}

# create table with bivariate analysis stats
bivariate_Pvalue <- function(outcome, explanatory){
  model_output <- polr(as.factor(outcome) ~ explanatory, Hess=T)
  ctable <- coef(summary(model_output))
  # calculate and store p values
  p <- pnorm(abs(ctable[,"t value"]), lower.tail = F)*2
  p_round <- round(pnorm(abs(ctable[,"t value"]), lower.tail = F)*2,4)
  #odds ratio
  oddsratio <- round(exp(coef(model_output)),4)
  #combine elements
  ctable <- cbind(round(ctable,4), "p-value"=p, "simp p."=p_round, "odds ratio"=oddsratio)
  return(ctable)
}

# create stargazer table
stargazer_table <- function(outcome, explanatory, table_title){
  stargazer::stargazer(
    polr(as.factor(outcome) ~ explanatory, Hess=T), type = "html",
                      title =  table_title,
                      notes.append =  FALSE, 
                      notes =  c("<sup>&sstarf;</sup>p<0.1; <sup>&sstarf;&sstarf;</sup>p<0.05; <sup>&sstarf;&sstarf;&sstarf;</sup>p<0.01"))

}

# Confidence Interval for mean = sample mean + Z-value for confidence level(sample st. dev/sqrt(number of elements in sample))

Mean.SD_CI <- function(x, cat_var, num_var, weight_var){
  cat_var <- enquo(cat_var)
  num_var <- enquo(num_var)
  weight_var <- enquo(weight_var)

  Mean.SD_CI_table <- x %>%
    group_by(!!cat_var) %>%
    summarize(n=n(),
              Group_weight = round(sum(!!weight_var),0),
              Weighted_percent = (round(Group_weight/sum(x$hh_wt_2019),4)*100),
              Weighted_avg = round(weighted.mean(!!num_var, !!weight_var),2),
              sd = round(sd(!!num_var),3),
              se = round(sd/sqrt(n),3),
              CI_90 = round(z*(sd/sqrt(n)),3),
              CI_95 = round(z_95*(sd/sqrt(n)),3))
    
  Mean.SD_CI_temp<- dplyr::select(Mean.SD_CI_table, -c(n))
  
  # return(Mean.SD_CI_temp)
  return(Mean.SD_CI_table)
}


# rounding function applied to categorical columns 
my.render.cat <- function(x) {
    c("", sapply(stats.default(x), function(y) with(y,
        sprintf("%d (%.0f%%)", FREQ, PCT))))}

# Statistical assumptions for margins of error
p_MOE <- 0.5
z <- 1.645 #90% CI
z_95 <- 1.96 #95% CI
missing_codes <- c('Missing: Technical Error', 'Missing: Non-response', 
                   'Missing: Skip logic', 'Children or missing', 'Prefer not to answer',
                   'Missing')
```

```{r read in and set up data}
#### Read in Data ####
#where you are running your R code
wrkdir <- "C:/Users/mrichards/Documents/GitHub/travel-studies/2019/analysis"


#where you want to output tables
file_loc <- 'C:/Users/mrichards/Documents/GitHub/travel-studies/2019/analysis/commute_travel/outputs'

sql.trip.query <- paste("SELECT hhincome_detailed, race_category, person_dim_id, mode_simple,
                        survey_year, o_purpose, d_purpose, trip_wt_2019 
                        FROM HHSurvey.v_trips_2017_2019")

trips <- read.dt(sql.trip.query, 'sqlquery')
trips_2019 <- trips %>% filter(survey_year==2019)

sql.person.query <- paste("SELECT employment, hhincome_detailed, hhincome_broad, race_category, person_dim_id, gender, age, age_category, education, vehicle_count,commute_auto_time,commute_auto_Distance, mode_freq_1,  mode_freq_2,  mode_freq_3, mode_freq_4, mode_freq_5, workplace, survey_year, sample_county, work_county,telecommute_freq, hh_wt_2019 FROM HHSurvey.v_persons_2017_2019")

persons<-read.dt(sql.person.query, 'sqlquery')
persons_2019 <- persons  %>% filter(survey_year==2019)
```
\

## Commuting Trips
This analysis will focus on the commute trips of people who usually work outside the home. 

* **What is the share of trips made for "commuting"?**

Commuting will be defined as the trips where one of the trip ends is a regular workplace - based on the survey logic, the origin or destination purpose is 'Went to primary workplace'
```{r, include=FALSE}
## SURVEY
survey_trips <- nrow(trips_2019) # 72,855

# filter to commute trips made by people who work outside the home usually
commute_trips <- trips_2019 %>%
  filter(o_purpose=='Went to primary workplace' | d_purpose=='Went to primary workplace')

survey_commute_trips <- nrow(commute_trips) # 14,501

survey_commute_trips/survey_trips #0.199

## WEIGHTED
trips_total <- sum(trips_2019$trip_wt_2019) #17,813,956

trips_commute <- trips_2019 %>% 
  filter(o_purpose=='Went to primary workplace' | d_purpose=='Went to primary workplace') %>%
  summarize(commute_trips=sum(trip_wt_2019)) 

trips_commute <- commute_trips %>% 
  summarize(weighted_commute_trips=sum(trip_wt_2019))
# 4,189,078	

trips_commute/trips_total # 24%
```
Based on the total number of trips (72,855) documented as part of the 2019 survey results, 20% (14,501) were recorded as being commute trips. When trip weights are taken into consideration, **24%** of the trips in the region are for commuting. 
\
\

## Defining Commuters
This subset of the respondents are based on the following survey questions: 

* **"Primary type of employment?"** 
* **"Please answer for your primary jobs where you work the most hours per week"**
    + "Usual work location?"
\

For this analysis, commuters are those with: 

1. full-time or part-time paid employment or those who identified as self-employed with workplaces.  
2. workplaces that are not at home  
3. commute distances less than 200 miles
\

```{r}
workers <- persons_2019 %>% 
  filter(employment=='Employed full time (35+ hours/week, paid)'|
           employment=='Employed part time (fewer than 35 hours/week, paid)'|
           employment=='Self-employed')

# work_loc_type <- workers %>% 
#   group_by(workplace) %>%
#   summarize(Sample_size=n(),
#             Weighted_2019=sum(hh_wt_2019)) %>%
#   mutate(Weighted_Percent=(round(Weighted_2019/sum(workers$hh_wt_2019),4)*100))

# get workers who don't work at home all the time with commute distances less than 200 miles
# remove outliers
not_home_workers <- workers %>% 
  filter(workplace!='At home (telecommute or self-employed with home office)') %>%
  filter(commute_auto_Distance<200)

# table1(~employment|workplace, data=not_home_workers,
#        render.categorical=my.render.cat)

not_home_workers %>%
  group_by(employment) %>%
  summarize(n=n(),
            Weighted_2019=round(sum(hh_wt_2019))) %>%
  mutate(Weighted_Percent=(round(Weighted_2019/sum(not_home_workers$hh_wt_2019),4)*100))
not_home_workers %>%
  group_by(workplace) %>%
  summarize(n=n(),
            Weighted_2019=round(sum(hh_wt_2019))) %>%
  mutate(Weighted_Percent=(round(Weighted_2019/sum(not_home_workers$hh_wt_2019),4)*100))
```
  
Out of a total 5,711 people who responded to the 2019 travel survey, **2,704** are individuals employed outside of the home who commute with automobile commutes of less than 200 miles. The commute distance limit was intended to remove outliers.
\
\

## General commute statistics
### Methodology

The following statistics uses the fields `commute_auto_Distance` and `commute_auto_time` to summarize the distribution of distance to work and commute duration. It is important to note that only 2019 data has the distances. 
\
\
The main question we will be attempting to answer is: **How does the distance or commute time from home to your primary workplace vary by income and race?**
\
\
Commute travel times and distances were found using the google distance matrix api (https://developers.google.com/maps/documentation/distance-matrix/overview).  Survey respondents reported home locations and work locations for all household workers.  Google travel time and distance data were pulled from the api given these locations, assuming an 8 am travel time in April of 2019..
\
\

### **Commute distance**
```{r}
mean_dist<-weighted.mean(not_home_workers$commute_auto_Distance,
                         w=not_home_workers$hh_wt_2019,na.rm=TRUE)

# hist(not_home_workers$commute_auto_Distance, main="Histogram of Commute Distance (miles)")

ggplot(not_home_workers, aes(x=commute_auto_Distance, weight=hh_wt_2019)) + 
  geom_histogram(fill='darkblue', binwidth=2)+xlim(c(0, 50)) +
  geom_vline(xintercept=mean_dist) +
  labs(title="Distribution of 2019 Commute Distances (from the HTS)",
      x ="Commute Distance (miles)", y = "Number of Workers (Weighted)")
```
\
_The vertical line at x=**13.97 miles** represents the mean distance traveled by workers_
\
\

Share of workers in the region with auto commutes
```{r}
not_home_workers <- not_home_workers %>%
  mutate(dist_bins = case_when(commute_auto_Distance<3 ~ "<3 miles (very short)",
                               commute_auto_Distance>=3 &
                                 commute_auto_Distance<10 ~ "3-10 miles (short)",
                               commute_auto_Distance>=10 &
                                 commute_auto_Distance<20 ~ "10-20 miles (average)",
                               commute_auto_Distance>=20 &
                                 commute_auto_Distance<50 ~ "20-50 miles (long)",
                               commute_auto_Distance>=50 ~ "50+ miles (very long)"))

not_home_workers$dist_bins <- factor(not_home_workers$dist_bins,
                               levels=c("<3 miles (very short)","3-10 miles (short)","10-20 miles (average)",
                                        "20-50 miles (long)","50+ miles (very long)"))

not_home_workers %>%
  group_by(dist_bins) %>%
  summarise(HouseholdWeight=round(sum(hh_wt_2019)),
            WeightedPercent=(round(HouseholdWeight/sum(not_home_workers$hh_wt_2019),4)*100),
            WeightedAvg=round(weighted.mean(commute_auto_Distance,
                         w=hh_wt_2019,na.rm=TRUE),2),
            min=min(commute_auto_Distance), 
            max=max(commute_auto_Distance),
            SampleSize=n(),
            SamplePercent=round((SampleSize/nrow(not_home_workers))*100,2),)

# table1(~dist_bins, data=not_home_workers)
```
\
\

#### 1. Income
```{r}
# reorder detailed income categories
not_home_workers$hhincome_detailed <- factor(not_home_workers$hhincome_detailed,
                                             levels=c("Under $10,000",
                                                      "$10,000-$24,999",
                                                      "$25,000-$34,999",
                                                      "$35,000-$49,999",
                                                      "$50,000-$74,999",
                                                      "$75,000-$99,999",
                                                      "$100,000-$149,999",
                                                      "$150,000-$199,999",
                                                      "$200,000-$249,999",
                                                      "$250,000 or more",
                                                      "Prefer not to answer"))
```
From workers with incomes less than $75,000 annually, lower income workers have shorter commutes.  For example, workers with incomes under $10,000 have commutes of around 14 minutes, whereas workers with incomes from $50,000-$74,999 have commutes around 27 minutes. The average travel time to work is about the same for workers in households with incomes greater than $75,000 – it’s around 33 minutes.
\
\

The average commuting distance by income:
```{r distance income with sample CI}
dist_income_stats <- Mean.SD_CI(not_home_workers, hhincome_detailed, commute_auto_Distance, hh_wt_2019)
dist_income_stats
```

```{r}
avg_dist_by_inc<-not_home_workers %>%
  group_by(hhincome_detailed) %>% 
  summarize(n=n(),
            Percent=round((n/nrow(not_home_workers))*100,2),
            HouseholdWeight=sum(hh_wt_2019),
            avg_weighted_dist=round(weighted.mean(commute_auto_Distance,hh_wt_2019),2))

ggplot(avg_dist_by_inc, aes(x=hhincome_detailed, y=avg_weighted_dist)) + 
  geom_col() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Distance by Income",
      x ="Income Categories", y = "Average Weighted Distance (miles)") +
  geom_errorbar(aes(ymin=dist_income_stats$Weighted_avg-dist_income_stats$CI_95, 
                    ymax=dist_income_stats$Weighted_avg+dist_income_stats$CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\

#### 2. Race
```{r}
# reorder race categories
not_home_workers$race_category <- 
  factor(not_home_workers$race_category,
         levels=c("African American",
                  "Asian",
                  "Hispanic",
                  "White Only",
                  "Other",
                  "Missing"))
```
People who identify as white have the longest commutes (average ~15 miles). The margins of error for African American and Hispanic groups are relatively large due to the small sample sizes. 
\

The average commuting distance by race:
```{r dist race with sample CI}
dist_race_stats <- Mean.SD_CI(not_home_workers, race_category, commute_auto_Distance, hh_wt_2019)
dist_race_stats
```

```{r}
avg_dist_by_race<-not_home_workers %>%
  group_by(race_category) %>% 
  summarize(n=n(),
            Percent=round((n/nrow(not_home_workers))*100,2),
            HouseholdWeight=sum(hh_wt_2019),
            avg_weighted_dist=round(weighted.mean(commute_auto_Distance,hh_wt_2019),2))

ggplot(avg_dist_by_race, aes(x=race_category, y=avg_weighted_dist)) + 
  geom_col() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Distance by Race",
      x ="Race Categories", y = "Average Weighted Distance (miles)") +
  geom_errorbar(aes(ymin=dist_race_stats$Weighted_avg-dist_race_stats$CI_95, 
                    ymax=dist_race_stats$Weighted_avg+dist_race_stats$CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\

##### Aggregated Race
Because of small sample sizes, some of the respondents have been aggregated.
\
\

###### <span style="color:#486CAB">African American and Hispanic</span>
For this analysis, respondents who identified as African American and Hispanic were aggregated.
```{r}
# reorder race categories
not_home_workers$race_category <- as.character(not_home_workers$race_category)
not_home_workers <- not_home_workers %>%
  mutate(race_agg_1 = case_when(race_category=="African American" |
                                  race_category=="Hispanic" ~ "POC",
                                TRUE~.$race_category))

not_home_workers$race_agg_1 <- factor(not_home_workers$race_agg_1,
                                      levels=c("Asian",
                                               "POC",
                                               "White Only",
                                               "Other",
                                               "Missing"))
```
People who identify as African American and Hispanic have shorter commutes (average ~12 miles), while White Only individuals have longer commutes (average ~15 minutes).

Although the margin of error for the POC group is better than when African American- and Hispanic-identifying individuals were analyzed separately, it is still not distinct from respondents identifying as Asian.
\

The average commuting distance by race:
```{r dist race_1 with sample CI}
dist_race_stats <- Mean.SD_CI(not_home_workers, race_agg_1, commute_auto_Distance, hh_wt_2019)
dist_race_stats
```

```{r}
avg_dist_by_race<-not_home_workers %>%
  group_by(race_agg_1) %>% 
  summarize(n=n(),
            Percent=round((n/nrow(not_home_workers))*100,2),
            HouseholdWeight=sum(hh_wt_2019),
            avg_weighted_dist=round(weighted.mean(commute_auto_Distance,hh_wt_2019),2))

ggplot(avg_dist_by_race, aes(x=race_agg_1, y=avg_weighted_dist)) + 
  geom_col() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Distance by Race",
      x ="Race Categories", y = "Average Weighted Distance (miles)") +
  geom_errorbar(aes(ymin=dist_race_stats$Weighted_avg-dist_race_stats$CI_95, 
                    ymax=dist_race_stats$Weighted_avg+dist_race_stats$CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\

###### <span style="color:#486CAB">African American, Hispanic, and Other</span>
For this analysis, respondents who identified as African American, Hispanic, and Other were aggregated. 
```{r}
# reorder race categories
not_home_workers$race_category <- as.character(not_home_workers$race_category)
not_home_workers <- not_home_workers %>%
  mutate(race_agg_2 = case_when(race_category=="African American" |
                                  race_category=="Hispanic" |
                                  race_category=="Other" ~ "POC",
                                TRUE~.$race_category))

not_home_workers$race_agg_2 <- factor(not_home_workers$race_agg_2,
                                      levels=c("Asian",
                                               "POC",
                                               "White Only",
                                               "Other",
                                               "Missing"))
```
People who identify as African American, Hispanic, or Other have shorter commutes (average ~11 miles), while White Only individuals have longer commutes (average ~15 minutes).

The margin of error for the POC group is better than when African American-, Hispanic-, and Other-identifying individuals were analyzed separately. Unlike the previous aggregation, this group is now distinct from respondents identifying as Asian.
\

The average commuting distance by race:
```{r dist race_2 with sample CI}
dist_race_stats <- Mean.SD_CI(not_home_workers, race_agg_2, commute_auto_Distance, hh_wt_2019)
dist_race_stats
```

```{r}
avg_dist_by_race<-not_home_workers %>%
  group_by(race_agg_2) %>% 
  summarize(n=n(),
            Percent=round((n/nrow(not_home_workers))*100,2),
            HouseholdWeight=sum(hh_wt_2019),
            avg_weighted_dist=round(weighted.mean(commute_auto_Distance,hh_wt_2019),2))

ggplot(avg_dist_by_race, aes(x=race_agg_2, y=avg_weighted_dist)) + 
  geom_col() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Distance by Race",
      x ="Race Categories", y = "Average Weighted Distance (miles)") +
  geom_errorbar(aes(ymin=dist_race_stats$Weighted_avg-dist_race_stats$CI_95, 
                    ymax=dist_race_stats$Weighted_avg+dist_race_stats$CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\


#### 3. Gender
```{r, include=FALSE, eval=FALSE}
# reorder gender categories
not_home_workers$gender <- 
  factor(not_home_workers$gender,
         levels=c("Female",
                  "Male",
                  "Another",
                  "Prefer not to answer")
```
People who identify as another gender have the shortest commute distances (sample size?), followed by females. People who identify as male have the longest commutes (15 miles).
\

The average commuting distance by gender:
```{r dist gender with sample CI}
dist_gender_stats <- Mean.SD_CI(not_home_workers, gender, commute_auto_Distance, hh_wt_2019)
dist_gender_stats
```
\

Because of the small sample sizes for respondents identifying as Another or choosing not to answer, these were removed from the graph:  
```{r}
avg_dist_by_gender<-not_home_workers %>%
  select(gender, commute_auto_Distance, hh_wt_2019) %>%
  group_by(gender)

temp <- avg_dist_by_gender %>%
  filter(gender=="Female" | gender=="Male") %>%
  summarize(n=n(),
            Percent=round((n/nrow(avg_dist_by_gender))*100,2),
              Group_weight = round(sum(hh_wt_2019),0),
              Weighted_percent = (round(Group_weight/sum(avg_dist_by_gender$hh_wt_2019),4)*100),
              Weighted_avg = round(weighted.mean(commute_auto_Distance, hh_wt_2019),2),
              sd = round(sd(commute_auto_Distance),3),
              se = round(sd/sqrt(n),3),
              CI_90 = round(z*(sd/sqrt(n)),3),
              CI_95 = round(z_95*(sd/sqrt(n)),3))

ggplot(temp, aes(x=gender, y=Weighted_avg)) + 
  geom_col() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Distance by Gender",
       x ="Gender Categories", 
       y = "Average Weighted Distance (miles)") +
  geom_errorbar(aes(ymin=Weighted_avg-CI_95, 
                    ymax=Weighted_avg+CI_95), 
                width=.2,
                position=position_dodge(.9))
```

```{r, include=FALSE, eval=FALSE}
avg_dist_by_gender<-not_home_workers %>%
  group_by(gender) %>% 
  summarize(n=n(),
            Percent=round((n/nrow(not_home_workers))*100,2),
            HouseholdWeight=sum(hh_wt_2019),
            avg_weighted_dist=round(weighted.mean(commute_auto_Distance,hh_wt_2019),2))

ggplot(avg_dist_by_gender, aes(x=gender, y=avg_weighted_dist)) + 
  geom_col() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Distance by Gender",
      x ="Gender Categories", y = "Average Weighted Distance (miles)") +
  geom_errorbar(aes(ymin=dist_gender_stats$Weighted_avg-dist_gender_stats$CI_95, 
                    ymax=dist_gender_stats$Weighted_avg+dist_gender_stats$CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\

### **Commute duration**
```{r}
mean_time<-weighted.mean(not_home_workers$commute_auto_time,
                         w=not_home_workers$hh_wt_2019,na.rm=TRUE)

# hist(not_home_workers$commute_auto_time, main="Histogram of Commute Duration (minutes)")

ggplot(not_home_workers, aes(x=commute_auto_time, weight=hh_wt_2019)) + geom_histogram(fill='darkblue', binwidth=5)+xlim(c(0, 200))+
  scale_x_continuous(breaks=seq(0,200,5))+
  geom_vline(xintercept=mean_time)+
  labs(title="Distribution of 2019 Commute Times (from the HTS)",
       x ="Commute Times(minutes)", y = "Number of Workers (Weighted)")
```
\
_The vertical line at x=**29.30 minutes** represents the mean commute duration for workers_
\
\

Share of workers with auto commutes
```{r}
not_home_workers <- not_home_workers %>%
  mutate(time_bins = case_when(commute_auto_time<10 ~ "<10 minutes",
                               commute_auto_time>=10 &
                                 commute_auto_time<20 ~ "10-20 minutes",
                               commute_auto_time>=20 &
                                 commute_auto_time<30 ~ "20-30 minutes",
                               commute_auto_time>=30 &
                                 commute_auto_time<45 ~ "30-45 minutes",
                               commute_auto_time>=45 &
                                 commute_auto_time<60 ~ "45-60 minutes",
                               commute_auto_time>=60 ~ "60+ minutes"))

not_home_workers %>%
  group_by(time_bins) %>%
  summarise(HouseholdWeight=round(sum(hh_wt_2019)),
            WeightedPercent=(round(HouseholdWeight/sum(not_home_workers$hh_wt_2019),4)*100),
            WeightedAvg=round(weighted.mean(commute_auto_time,
                         w=hh_wt_2019,na.rm=TRUE),2),
            min=min(commute_auto_time), 
            max=max(commute_auto_time),
            SampleSize=n(),
            SamplePercent=round((SampleSize/nrow(not_home_workers))*100,2),)

# table1(~time_bins, data=not_home_workers)
```
\
\

#### 1. Income
The distance to work increases with income up to 200K, then declines. The average distance to work for people in households with incomes less than $10,000 is 6 miles. The average distance to work for people in households with incomes $150,000-$200,000 is 18 miles (3 times as long).
\
\

The average commuting duration by income:
```{r time income with sample CI}
time_income_stats <- Mean.SD_CI(not_home_workers, hhincome_detailed, commute_auto_time, hh_wt_2019)
time_income_stats
```

```{r}
avg_time_by_inc<-not_home_workers %>% 
  group_by(hhincome_detailed) %>% 
  summarize(n=n(),
            Percent=round((n/nrow(not_home_workers))*100,2),
            HouseholdWeight=sum(hh_wt_2019),
            WeightedPercent=round((HouseholdWeight/sum(HouseholdWeight))*100,2),
            avg_weighted_time=round(weighted.mean(commute_auto_time,hh_wt_2019),2))

ggplot(avg_time_by_inc, aes(x=hhincome_detailed, y=avg_weighted_time)) + 
  geom_col() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Travel Time by Income",
      x ="Income Categories", y = "Weighted Average Commute Time (minutes)") +
  geom_errorbar(aes(ymin=time_income_stats$Weighted_avg-time_income_stats$CI_95, 
                    ymax=time_income_stats$Weighted_avg+time_income_stats$CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\

#### 2. Race
People who identify with by a race other than African American, Asian, Hispanic or White have the shortest commutes at about 18 minutes. African American, Asian, Hispanic, and White Only workers all have commutes around 30 minutes long.
\

The average commuting duration by race:
```{r time race with sample CI}
time_race_stats <- Mean.SD_CI(not_home_workers, race_category, commute_auto_time, hh_wt_2019)
time_race_stats
```

```{r}
avg_time_by_race<-not_home_workers %>%
  group_by(race_category) %>% 
  summarize(n=n(),
            Percent=round((n/nrow(not_home_workers))*100,2),
            HouseholdWeight=sum(hh_wt_2019),
            avg_weighted_time=round(weighted.mean(commute_auto_time,hh_wt_2019),2))

ggplot(avg_time_by_race, aes(x=race_category, y=avg_weighted_time)) + 
  geom_col() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Time by Race",
      x ="Race Categories", y = "Average Weighted Time (minutes)") +
  geom_errorbar(aes(ymin=time_race_stats$Weighted_avg-time_race_stats$CI_95, 
                    ymax=time_race_stats$Weighted_avg+time_race_stats$CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\


##### Aggregated Race
Because of small sample sizes, some of the respondents have been aggregated.
\
\

###### <span style="color:#486CAB">African American and Hispanic</span>
For this analysis, respondents who identified as African American and Hispanic were aggregated.

People who identify as African American and Hispanic have shorter commutes (average ~12 miles), while White Only individuals have longer commutes (average ~15 minutes).

Although the margin of error for the POC group is better than when African American- and Hispanic-identifying individuals were analyzed separately, it is still not distinct from respondents identifying as Asian.
\

The average commuting distance by race:
```{r time race_1 with sample CI}
time_race_stats <- Mean.SD_CI(not_home_workers, race_agg_1, commute_auto_time, hh_wt_2019)
time_race_stats
```

```{r}
avg_time_by_race<-not_home_workers %>%
  group_by(race_agg_1) %>% 
  summarize(n=n(),
            Percent=round((n/nrow(not_home_workers))*100,2),
            HouseholdWeight=sum(hh_wt_2019),
            avg_weighted_time=round(weighted.mean(commute_auto_time,hh_wt_2019),2))

ggplot(avg_time_by_race, aes(x=race_agg_1, y=avg_weighted_time)) + 
  geom_col() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Time by Race",
      x ="Race Categories", y = "Average Weighted Time (minutes)") +
  geom_errorbar(aes(ymin=time_race_stats$Weighted_avg-time_race_stats$CI_95, 
                    ymax=time_race_stats$Weighted_avg+time_race_stats$CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\

###### <span style="color:#486CAB">African American, Hispanic, and Other</span>
For this analysis, respondents who identified as African American, Hispanic, and Other were aggregated. 

People who identify as African American, Hispanic, or Other have shorter commutes (average ~11 miles), while White Only individuals have longer commutes (average ~15 minutes).

The margin of error for the POC group is better than when African American-, Hispanic-, and Other-identifying individuals were analyzed separately. Unlike the previous aggregation, this group is now distinct from respondents identifying as Asian.
\

The average commuting distance by race:
```{r time race_2 with sample CI}
time_race_stats <- Mean.SD_CI(not_home_workers, race_agg_2, commute_auto_time, hh_wt_2019)
time_race_stats
```

```{r}
avg_time_by_race<-not_home_workers %>%
  group_by(race_agg_2) %>% 
  summarize(n=n(),
            Percent=round((n/nrow(not_home_workers))*100,2),
            HouseholdWeight=sum(hh_wt_2019),
            avg_weighted_time=round(weighted.mean(commute_auto_time,hh_wt_2019),2))

ggplot(avg_time_by_race, aes(x=race_agg_2, y=avg_weighted_time)) + 
  geom_col() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Time by Race",
      x ="Race Categories", y = "Average Weighted Time (minutes)") +
  geom_errorbar(aes(ymin=time_race_stats$Weighted_avg-time_race_stats$CI_95, 
                    ymax=time_race_stats$Weighted_avg+time_race_stats$CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\


#### 3. Gender
```{r, include=FALSE, eval=FALSE}
# reorder gender categories
not_home_workers$gender <- 
  factor(not_home_workers$gender,
         levels=c("Female",
                  "Male",
                  "Another",
                  "Prefer not to answer")
```
People who identify as another gender have the shortest commute distances (sample size?), followed by females. People who identify as male have the longest commutes (31 minutes).
\

The average commuting distance by gender:
```{r time gender with sample CI}
time_gender_stats <- Mean.SD_CI(not_home_workers, gender, commute_auto_time, hh_wt_2019)
time_gender_stats
```
\

Because of the small sample sizes for respondents identifying as Another or choosing not to answer, these were removed from the graph:  
```{r}
avg_time_by_gender<-not_home_workers %>%
  select(gender, commute_auto_time, hh_wt_2019) %>%
  group_by(gender)

temp <- avg_time_by_gender %>%
  filter(gender=="Female" | gender=="Male") %>%
  summarize(n=n(),
            Percent=round((n/nrow(avg_time_by_gender))*100,2),
              Group_weight = round(sum(hh_wt_2019),0),
              Weighted_percent = (round(Group_weight/sum(avg_time_by_gender$hh_wt_2019),4)*100),
              Weighted_avg = round(weighted.mean(commute_auto_time, hh_wt_2019),2),
              sd = round(sd(commute_auto_time),3),
              se = round(sd/sqrt(n),3),
              CI_90 = round(z*(sd/sqrt(n)),3),
              CI_95 = round(z_95*(sd/sqrt(n)),3))

ggplot(temp, aes(x=gender, y=Weighted_avg)) + 
  geom_col() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Time by Gender",
       x ="Gender Categories", 
       y = "Average Weighted Time (minutes)") +
  geom_errorbar(aes(ymin=Weighted_avg-CI_95, 
                    ymax=Weighted_avg+CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\

## Multiple Variables

### **Commute Distance: Race and Gender**

#### All Race Categories
```{r}
# commute distance: gender and race
not_home_workers$race_category <- factor(not_home_workers$race_category,
                                         levels = c("African American",
                                                    "Asian",
                                                    "Hispanic",
                                                    "White Only",
                                                    "Other",
                                                    "Missing"))

avg_dist_by_gender_race<-not_home_workers %>%
  select(gender, race_category, race_agg_1, race_agg_2, commute_auto_Distance, hh_wt_2019) %>%
  group_by(gender, race_category, race_agg_1, race_agg_2)
```

```{r}
temp <- avg_dist_by_gender_race %>%
  filter(gender=="Female" | gender=="Male") %>%
  group_by(gender, race_category) %>%
  summarize(n=n(),
            Percent=round((n/nrow(avg_dist_by_gender_race))*100,2),
              Group_weight = round(sum(hh_wt_2019),0),
              Weighted_percent = (round(Group_weight/sum(avg_dist_by_gender_race$hh_wt_2019),4)*100),
              Weighted_avg = round(weighted.mean(commute_auto_Distance, hh_wt_2019),2),
              sd = round(sd(commute_auto_Distance),3),
              se = round(sd/sqrt(n),3),
              CI_90 = round(z*(sd/sqrt(n)),3),
              CI_95 = round(z_95*(sd/sqrt(n)),3))

ggplot(temp, aes(x=race_category, y=Weighted_avg, fill=gender)) + 
  geom_bar(position="dodge", stat="identity") +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Distance by Race",
       x ="Race Categories", 
       y = "Average Weighted Distance (miles)",
       fill = "Gender") +
  geom_errorbar(aes(ymin=Weighted_avg-CI_95, 
                    ymax=Weighted_avg+CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\

#### Aggregated Race Categories
Respondents who identified as African American and Hispanic were aggregated.
```{r}
temp <- avg_dist_by_gender_race %>%
  filter(gender=="Female" | gender=="Male") %>%
  group_by(gender, race_agg_1) %>%
  summarize(n=n(),
            Percent=round((n/nrow(avg_dist_by_gender_race))*100,2),
              Group_weight = round(sum(hh_wt_2019),0),
              Weighted_percent = (round(Group_weight/sum(avg_dist_by_gender_race$hh_wt_2019),4)*100),
              Weighted_avg = round(weighted.mean(commute_auto_Distance, hh_wt_2019),2),
              sd = round(sd(commute_auto_Distance),3),
              se = round(sd/sqrt(n),3),
              CI_90 = round(z*(sd/sqrt(n)),3),
              CI_95 = round(z_95*(sd/sqrt(n)),3))

ggplot(temp, aes(x=race_agg_1, y=Weighted_avg, fill=gender)) + 
  geom_bar(position="dodge", stat="identity") +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Distance by Race",
       x ="Race Categories", 
       y = "Average Weighted Distance (miles)",
       fill = "Gender") +
  geom_errorbar(aes(ymin=Weighted_avg-CI_95, 
                    ymax=Weighted_avg+CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\

#### Further Aggregated Race Categories
Respondents who identified as African American, Hispanic, or Other were aggregated.
```{r}
temp <- avg_dist_by_gender_race %>%
  filter(gender=="Female" | gender=="Male") %>%
  group_by(gender, race_agg_2) %>%
  summarize(n=n(),
            Percent=round((n/nrow(avg_dist_by_gender_race))*100,2),
              Group_weight = round(sum(hh_wt_2019),0),
              Weighted_percent = (round(Group_weight/sum(avg_dist_by_gender_race$hh_wt_2019),4)*100),
              Weighted_avg = round(weighted.mean(commute_auto_Distance, hh_wt_2019),2),
              sd = round(sd(commute_auto_Distance),3),
              se = round(sd/sqrt(n),3),
              CI_90 = round(z*(sd/sqrt(n)),3),
              CI_95 = round(z_95*(sd/sqrt(n)),3))

ggplot(temp, aes(x=race_agg_2, y=Weighted_avg, fill=gender)) + 
  geom_bar(position="dodge", stat="identity") +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Distance by Race",
       x ="Race Categories", 
       y = "Average Weighted Distance (miles)",
       fill = "Gender") +
  geom_errorbar(aes(ymin=Weighted_avg-CI_95, 
                    ymax=Weighted_avg+CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\


### **Commute Distance: Income and Race**

#### All Race Categories
For this analysis, broader income categories were used for easier visualization. The respondents who chose "Prefer not to answer" when asked for income were removed from this analysis.
```{r}
# commute distance: income and race
# unique(persons_2019$hhincome_broad)
not_home_workers$hhincome_broad <- factor(not_home_workers$hhincome_broad,
                                          levels=c("Under $25,000",
                                                   "$25,000-$49,999",
                                                   "$50,000-$74,999",
                                                   "$75,000-$99,999",
                                                   "$100,000 or more"))

avg_dist_by_income_race<-not_home_workers %>%
  select(hhincome_broad, race_category, race_agg_1, race_agg_2, commute_auto_Distance, hh_wt_2019) %>%
  group_by(hhincome_broad, race_category, race_agg_1, race_agg_2)
```

```{r}
temp <- avg_dist_by_income_race %>%
  filter(hhincome_broad!="Prefer not to answer") %>%
  group_by(hhincome_broad, race_category) %>%
  summarize(n=n(),
            Percent=round((n/nrow(avg_dist_by_income_race))*100,2),
              Group_weight = round(sum(hh_wt_2019),0),
              Weighted_percent = (round(Group_weight/sum(avg_dist_by_income_race$hh_wt_2019),4)*100),
              Weighted_avg = round(weighted.mean(commute_auto_Distance, hh_wt_2019),2),
              sd = round(sd(commute_auto_Distance),3),
              se = round(sd/sqrt(n),3),
              CI_90 = round(z*(sd/sqrt(n)),3),
              CI_95 = round(z_95*(sd/sqrt(n)),3))

ggplot(temp, aes(x=hhincome_broad, y=Weighted_avg, fill=race_category)) + 
  geom_bar(position="dodge", stat="identity") +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Distance by Income",
       x ="Income Categories", 
       y = "Average Weighted Distance (miles)",
       fill = "Race") +
  geom_errorbar(aes(ymin=Weighted_avg-CI_95, 
                    ymax=Weighted_avg+CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\

#### Aggregated Race Categories
Respondents who identified as African American and Hispanic were aggregated.
```{r}
temp <- avg_dist_by_income_race %>%
  filter(hhincome_broad!="Prefer not to answer") %>%
  group_by(hhincome_broad, race_agg_1) %>%
  summarize(n=n(),
            Percent=round((n/nrow(avg_dist_by_income_race))*100,2),
              Group_weight = round(sum(hh_wt_2019),0),
              Weighted_percent = (round(Group_weight/sum(avg_dist_by_income_race$hh_wt_2019),4)*100),
              Weighted_avg = round(weighted.mean(commute_auto_Distance, hh_wt_2019),2),
              sd = round(sd(commute_auto_Distance),3),
              se = round(sd/sqrt(n),3),
              CI_90 = round(z*(sd/sqrt(n)),3),
              CI_95 = round(z_95*(sd/sqrt(n)),3))

ggplot(temp, aes(x=hhincome_broad, y=Weighted_avg, fill=race_agg_1)) + 
  geom_bar(position="dodge", stat="identity") +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Distance by Income",
       x ="Income Categories", 
       y = "Average Weighted Distance (miles)",
       fill = "Aggregated Race") +
  geom_errorbar(aes(ymin=Weighted_avg-CI_95, 
                    ymax=Weighted_avg+CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\

#### Further Aggregated Race Categories
Respondents who identified as African American, Hispanic, or Other were aggregated.
```{r}
temp <- avg_dist_by_income_race %>%
  filter(hhincome_broad!="Prefer not to answer") %>%
  group_by(hhincome_broad, race_agg_2) %>%
  summarize(n=n(),
            Percent=round((n/nrow(avg_dist_by_income_race))*100,2),
              Group_weight = round(sum(hh_wt_2019),0),
              Weighted_percent = (round(Group_weight/sum(avg_dist_by_income_race$hh_wt_2019),4)*100),
              Weighted_avg = round(weighted.mean(commute_auto_Distance, hh_wt_2019),2),
              sd = round(sd(commute_auto_Distance),3),
              se = round(sd/sqrt(n),3),
              CI_90 = round(z*(sd/sqrt(n)),3),
              CI_95 = round(z_95*(sd/sqrt(n)),3))

ggplot(temp, aes(x=hhincome_broad, y=Weighted_avg, fill=race_agg_2)) + 
  geom_bar(position="dodge", stat="identity") +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Distance by Income",
       x ="Income Categories", 
       y = "Average Weighted Distance (miles)",
       fill = "Aggregated Race") +
  geom_errorbar(aes(ymin=Weighted_avg-CI_95, 
                    ymax=Weighted_avg+CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\


### **Commute Distance: Income and Gender**
```{r}
# commute distance: gender and race
avg_dist_by_gender_income<-not_home_workers %>%
  select(gender, hhincome_broad, commute_auto_Distance, hh_wt_2019) %>%
  group_by(gender, hhincome_broad)

temp <- avg_dist_by_gender_income %>%
  filter(gender=="Female" | gender=="Male") %>%
  filter(hhincome_broad!="Prefer not to answer") %>%
  group_by(gender, hhincome_broad) %>%
  summarize(n=n(),
            Percent=round((n/nrow(avg_dist_by_gender_income))*100,2),
              Group_weight = round(sum(hh_wt_2019),0),
              Weighted_percent = (round(Group_weight/sum(avg_dist_by_gender_income$hh_wt_2019),4)*100),
              Weighted_avg = round(weighted.mean(commute_auto_Distance, hh_wt_2019),2),
              sd = round(sd(commute_auto_Distance),3),
              se = round(sd/sqrt(n),3),
              CI_90 = round(z*(sd/sqrt(n)),3),
              CI_95 = round(z_95*(sd/sqrt(n)),3))

ggplot(temp, aes(x=hhincome_broad, y=Weighted_avg, fill=gender)) + 
  geom_bar(position="dodge", stat="identity") +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  labs(title="Average Auto Commute Distance by Income",
       x ="Income Categories", 
       y = "Average Weighted Distance (miles)",
       fill = "Gender") +
  geom_errorbar(aes(ymin=Weighted_avg-CI_95, 
                    ymax=Weighted_avg+CI_95), 
                width=.2,
                position=position_dodge(.9))
```
\
\


## Trends in Travel Time
The following table uses ACS data
```{r}
# Data components
Counties <- c("King", "Kitsap", "Pierce", "Snohomish", "Region")
years <- c(1980,1990,2000,2010,2018)
c <- rep(Counties,5)
y <- rep(years, each=5)
y_1980 <- c(23.0,25.0,21.4,23.5,22.9)
y_1990 <- c(24.2,25.1,24.0,25.4,24.4)
y_2000 <- c(25.3,31.0,27.4,28.5,26.7)
y_2010 <- c(26.1,29.2,27.7,28.2,27.0)
y_2018 <- c(30.2,30.0,33.7,33.3,31.5)

# Create table
trend_table <- cbind(Counties, y_1980, y_1990, y_2000, y_2010, y_2018)

# Format table
trend_table %>% 
  kable(format = "html",align = c('l',rep(('c'),5))) %>%
  kable_styling("bordered",full_width = F, position = "left") %>% 
  column_spec(1, bold=TRUE) %>%
  row_spec(5, italic=TRUE)
```

Travel time to work has increased in our region since 1980 from about 23 minutes to 32 minutes in 2018.
\
\

```{r}
# Set up data frame
trend_data <- data.frame ( 
   County = c,
   time_m = c(y_1980, y_1990, y_2000, y_2010, y_2018),
   year = y)

# Reorder legend counties
trend_data$County <- factor(trend_data$County, levels=Counties)

# Plot
trend_data %>%
  ggplot(aes(x=year, y=time_m, group=County, color=County)) +
  geom_point() +
  geom_line() +
  labs(x = "Year",
       y = "Travel Time to Work (minutes)",
       group="Counties",
       title = "Average Commute Travel Time by Year") +
  scale_x_continuous(limits = c(1975, 2020),breaks=years) +
  scale_y_continuous(limits=c(0,35), minor_breaks = 5)
```



## Background: Sample
5,711 people responded to the 2019 survey and provided their employment status and workplace.
\
\

### Employment Status
The following table shows the distribution of respondents' employment status.
```{r}
persons_2019$employment <- factor(persons_2019$employment,
                                 levels=c("Employed full time (35+ hours/week, paid)",
                                          "Employed part time (fewer than 35 hours/week, paid)", 
                                          "Self-employed",
                                          "Unpaid volunteer or intern",
                                          "Homemaker",
                                          "Retired",
                                          "Not currently employed",
                                          "Missing: Skip logic"))

# employment_stats<-persons_2019 %>%
#   filter(employment!="Missing: Skip logic") %>%
#   group_by(employment) %>% 
#   summarize(sample_size=n(), 
#             HouseholdWeight = round(sum(hh_wt_2019),0)) %>%
#   mutate(Percent=(sample_size/sum(sample_size))*100)
# employment_stats

create_table_one_var_simp('employment', persons_2019, 'person')
```
\

### Workplace
The following table shows where workers go for work.
```{r}
persons_2019$workplace <- factor(persons_2019$workplace,
                                 levels=c("At home (telecommute or self-employed with home office)",
                                          "Drives for a living (e.g., bus driver, salesperson)", 
                                          "Usually the same location (outside home)",
                                          "Workplace regularly varies (different offices or jobsites)",
                                          "Missing: Skip logic"))
  
# workplace_stats<-persons_2019 %>%
#   filter(workplace!="Missing: Skip logic") %>%
#   group_by(workplace) %>% 
#   summarize(sample_size=n(), 
#             HouseholdWeight = round(sum(hh_wt_2019),0)) %>%
#   mutate(Percent=(sample_size/sum(sample_size))*100)
# workplace_stats

create_table_one_var_simp('workplace', persons_2019, 'person')

```