---
title: "Household"
date: today
---

```{r}
#| include: false

source("functions.R")
tables <- readRDS("analysis_table.RDS")
```

## Household Info

:::{.panel-tabset}

### Survey Platform

```{r}
CreateTableOne(data = tables[['hh']],
               vars = c("signup_platform","diary_platform")
               )
```

```{r}
get_vars_summary("hh", c("hhgroup"))
```

### Survey Completeness
```{r}
get_vars_summary("hh", c("hh_is_complete"))
```


```{r}
CreateTableOne(data = tables[['hh']],
               vars = c("num_trips","numdayscomplete","num_days_complete_weekday","num_complete_tue","num_complete_thu","num_days_complete_weekend"),
               factorVars = c("numdayscomplete","num_days_complete_weekday","num_complete_tue","num_complete_thu","num_days_complete_weekend")
               )
```
:::

## Household Size

::: {.panel-tabset}

### Total
```{r}
get_vars_summary("hh", c("hhsize"))
```

### Number of Participants
```{r}
CreateTableOne(data = tables[['hh']],
               vars = c("num_participants"),
               factorVars = c("num_participants")
               )
```

### Adults/ Children
```{r}
get_vars_summary("hh", c("numadults","numchildren"))
```

### Workers/ Students
```{r}
get_vars_summary("hh", c("numworkers","num_students"))
```
:::

## Home Location

### In Region
```{r}
get_vars_summary("hh", c("home_in_region"),order = FALSE)
```

### By County
```{r}
get_vars_summary("hh", c("home_county"))
```


## Household Income
```{r}
get_vars_summary("hh", c("hhincome_broad"))
```

::: {.callout-note collapse=true}

## detailed household income

```{r}
get_vars_summary("hh", c("hhincome_followup","hhincome_detailed"))
```

:::

## Auto Ownership
```{r}
get_vars_summary("hh", c("vehicle_count"))
```

## Housing
```{r}
get_vars_summary("hh", c("rent_own","res_type","res_dur"))
```

### Previous Residence

- all "Missing Response" in `prev_home_notwa_state` (because we don't ask those moved from outside Washington State)

:::{.panel-tabset}

#### Previous Residence Info
```{r}
get_vars_summary("hh", c("prev_home_wa","prev_res_type","prev_rent_own"))
```

#### Decision to Move

- number of responses for `prev_res_factors_specify`: `{r} length(unique(tables[["person"]]$prev_res_factors_specify))`
```{r}
get_vars_summary("hh", c("prev_res_factors_housing_cost","prev_res_factors_income_change","prev_res_factors_community_change",
                         "prev_res_factors_hh_size","prev_res_factors_more_space","prev_res_factors_less_space",
                         "prev_res_factors_employment","prev_res_factors_amenities","prev_res_factors_telework",
                         "prev_res_factors_school","prev_res_factors_crime","prev_res_factors_quality",
                         "prev_res_factors_forced","prev_res_factors_other","prev_res_factors_no_answer"))
```

:::

