---
title: "Special Needs Travel Analysis (2025)"
author: "Michael Jensen"
format:
  html:
    toc: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
hts_survey_year <- 2025

# Apply larger defaults before sourcing so generated plots inherit sizes.
ggplot2::theme_update(
  axis.text = ggplot2::element_text(size = 12),
  axis.title = ggplot2::element_text(size = 13),
  legend.text = ggplot2::element_text(size = 12),
  legend.title = ggplot2::element_text(size = 13)
)
ggplot2::update_geom_defaults("text", list(size = 4.5))

source("../../../../2023/summary/special_needs/hts_2023_SNTC_analysis.R")

```

## Overview

These data update analysis presented to the Coordinated Mobility & Accessibility Committee in 2024. The core Household Transportation Survey supports analysis of households with: 
1) disabilities
2) low-income
3) no vehicles
4) age 65+
5) people of color
Due to constraints on children as respondents, households with children under age 18 are not reflected; the survey also does not reflect/sample those with low English proficiency.

---

## Population Composition

### Income composition of the no-vehicle population

```{r noveh-income, fig.width=6, fig.height=4}
composition_plots$p_noveh_income
```

---

## Mode Share

```{r mode-tabs, results='asis', fig.width=6, fig.height=3.5}
mode_tab_labels <- c(
  p_dis_mode  = "By disability status",
  p_race_mode = "By race/ethnicity",
  p_age_mode  = "By age group",
  p_veh_mode  = "By vehicle availability",
  p_inc_mode  = "By household income"
)

cat("\n:::{.panel-tabset}\n\n")
purrr::iwalk(mode_tab_labels, \(label, nm) {
  cat("###", label, "\n\n")
  print(mode_plots[[nm]])
  cat("\n\n")
})
cat(":::")
```

---

## Trip Purpose

```{r purpose-tabs, results='asis', fig.width=6, fig.height=3.5}
purpose_tab_labels <- c(
  p_dis_purpose  = "By disability status",
  p_race_purpose = "By race/ethnicity",
  p_age_purpose  = "By age group",
  p_veh_purpose  = "By vehicle availability",
  p_inc_purpose  = "By household income"
)

cat("\n:::{.panel-tabset}\n\n")
purrr::iwalk(purpose_tab_labels, \(label, nm) {
  cat("###", label, "\n\n")
  print(purpose_plots[[nm]])
  cat("\n\n")
})
cat(":::")
```
