---
title: "Special Needs Travel Analysis (2025)"
author: "Michael Jensen"
format:
  html:
    toc: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
hts_survey_year <- 2025
source("../../../../2023/summary/special_needs/hts_2023_SNTC_analysis.R")
```

## Overview

These data update analysis presented to the Coordinated Mobility & Accessibility Committee in 2024. The core Household Transportation Survey supports analysis of households with: 
1) disabilities
2) low-income
3) no vehicles
4) age 65+
5) people of color
Due to constraints on children as respondents, households with children under age 18 are not reflected; the survey also does not reflect/sample those with low English proficiency.

---

## Population Composition

### Income composition of the no-vehicle population

```{r noveh-income, fig.width=6, fig.height=4}
composition_plots$p_noveh_income
```

---

## Mode Share

### By disability status

```{r mode-disability, fig.width=6, fig.height=3.5}
mode_plots$p_dis_mode
```

### By race/ethnicity

```{r mode-race, fig.width=6, fig.height=4}
mode_plots$p_race_mode
```

### By age group

```{r mode-age, fig.width=6, fig.height=3.5}
mode_plots$p_age_mode
```

### By vehicle availability

```{r mode-vehicle, fig.width=6, fig.height=3.5}
mode_plots$p_veh_mode
```

### By household income

```{r mode-income, fig.width=6, fig.height=4}
mode_plots$p_inc_mode
```

---

## Trip Purpose

### By disability status

```{r purpose-disability, fig.width=6, fig.height=3.5}
purpose_plots$p_dis_purpose
```

### By race/ethnicity

```{r purpose-race, fig.width=6, fig.height=4}
purpose_plots$p_race_purpose
```

### By age group

```{r purpose-age, fig.width=6, fig.height=3.5}
purpose_plots$p_age_purpose
```

### By vehicle availability

```{r purpose-vehicle, fig.width=6, fig.height=3.5}
purpose_plots$p_veh_purpose
```

### By household income

```{r purpose-income, fig.width=6, fig.height=4}
purpose_plots$p_inc_purpose
```
