---
title: Walk & Bike/Micromobility Summary
date: today
author: Grant Gibson

params:
  errorbar: FALSE
---

```{r setup, include=FALSE}
library(tidyverse)
library(psrcplot)
install_psrc_fonts()
```

```{r load_data, include=FALSE}
source("walk_bike_data.R")
source("../util.R")
```

```{r palettes, include=FALSE}
psrc_fill_dec <- list(
  Walk = psrc_colors$purples_dec[1:5],
  `Bike/Micromobility` = psrc_colors$oranges_dec[1:5],
  survey_year = psrc_colors$blues_dec[1:5],
  mode_class_5 = psrc_colors$greens_dec[1:5]
  )

psrc_fill_segment <- list(
  income_binary = c("#91268F", "#8CC63E"),
  income = psrc_colors$pgnobgy_5,
  gender = c("#F05A28", "#00A7A0", "#91268F"),
  race_binary = c("#91268F", "#00A7A0"),
  race_ethnicity = psrc_colors$pognbgy_10,
  age = psrc_colors$gnbopgy_5,
  home_county = psrc_colors$obgnpgy_10,
  in_rgc = c("#00A7A0", "#8CC63E"),
  res_type_group = c("#91268F", "#F05A28")
  )
```

```{r figure_output, include=FALSE}
# single category
single_cat_output <- function(tbl, mode){
  print(
    ggplot(filter(tbl, mode_class_5 == mode), 
           aes(x=factor(survey_year), y=prop)) +
      geom_col(aes(fill = factor(survey_year)), 
               width=0.7 # narrows columns to 70% for more white space between columns
               ) +
      # add data value above each column
      geom_text(aes(label=scales::percent(prop, accuracy = 1)),
                vjust = -0.3 # mode text vertically upward so it sits above columns
                ) +
      # add thin grey horizontal line at 0
      geom_hline(yintercept = 0, color = "#cbcbcb", linewidth = 0.3) +
      
      # Error Bar Toggle:
      # To switch error on and off using params (TRUE/FALSE)
      errorbar_switch(
        
        geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe), width = 0.1), 
        
        on=params$errorbar # this reads the errorbar parameter from yaml header to include error bars or not
        ) +
      
      scale_y_continuous(expand = expansion(mult = c(0, 0.1))) + # add a 10% buffer at the top to prevent geom_text from clipping
      scale_fill_manual(values = psrc_fill_dec[[mode]]) + # Palette Mapping: select color vector in psrc_fill_dec color list
      psrc_style() + # apply PSRC formatting from psrcplot
      theme(legend.position = "none", # Removes the legend when colors map 1:1 to x-axis labels
            panel.grid.major.y = element_blank(), # Removes horizontal gridlines
            axis.text.y = element_blank(), # Removes y-axis tick values
            axis.title = element_blank()) # Removes both X and Y titles (use axis.title.x = element_blank() to keep y-axis title)
  )
}

# multi category
multi_cat_output <- function(tbl, mode, group_var, type = "both"){
  if( type %in% c("year","both") ){
    
    # grouping by year
    df_plot <- ggplot(filter(tbl, mode_class_5 == mode),
                      aes(x=mode, y=prop, fill=!!sym(group_var))) +
      geom_col(position = position_dodge(), width=0.9) +
      geom_text(aes(label=scales::percent(prop, accuracy = 1)),
                    position = position_dodge(0.9),
                vjust = -0.3, 
                size = 10, size.unit="pt") +
      errorbar_switch(geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                                    position = position_dodge(0.9),
                                    width = 0.1), 
                      on=params$errorbar) +
      geom_hline(yintercept = 0, color = "#cbcbcb", linewidth = 0.3) +
      facet_wrap(~ survey_year, ncol=5)+
      scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, 0.1))) +
      scale_fill_manual(values = psrc_fill_segment[[group_var]])+
      psrc_style() +
      theme(panel.grid.major.y = element_blank(),
            legend.text = element_text(size = 11), # adjust text size on legend
            axis.title = element_blank(),
            axis.text.x = element_blank(),
            axis.text.y = element_blank())

    print(df_plot)

  }

  if( type %in% c("cat","both") ){
    
    # grouping by category
    df_plot <- ggplot(filter(tbl, mode_class_5 == mode),
                      aes(x=factor(survey_year), y=prop, fill=factor(survey_year))) +
      geom_col(width = 0.8) +
      geom_text(aes(label=scales::percent(prop, accuracy = 1)),
                vjust = -0.3) +
      errorbar_switch(geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                                    width = 0.1), 
                      on=params$errorbar) +
      geom_hline(yintercept = 0, color = "#cbcbcb", linewidth = 0.3) +
      facet_wrap(as.formula(paste("~", group_var)), ncol=5)+
      scale_fill_manual(values = psrc_fill_dec[[mode]])+
      scale_y_continuous(labels = scales::percent, 
                         expand = expansion(mult = c(0, 0.1))) +
      psrc_style() +
      theme(legend.position = "none", # remove legend
            strip.text = element_text(size = 10.5), # adjust text size on facet title (variable values)
            panel.grid.major.y = element_blank(),
            panel.spacing.x = unit(0.5, "lines"), # add spacing between facet grids
            axis.title = element_blank(),
            axis.text.x = element_text(size = 10.5), # adjust text size for axis text (years)
            axis.text.y = element_blank())

    print(df_plot)

  }
}
```

```{r tbl_output, include=FALSE}
tbl_output <- function(tbl, mode){
  return(filter(tbl, mode_class_5 == mode) %>% 
         mutate(prop = scales::percent(prop, accuracy = 0.1),
                prop_moe = scales::percent(prop_moe, accuracy = 0.01),
                est = scales::number(est, accuracy = 1, big.mark = ","),
                count = scales::number(count, accuracy = 1, big.mark = ",")) %>%
         select(-est_moe) %>%
         rename(`survey year` = survey_year,
                mode = mode_class_5,
                share = prop,
                share_moe = prop_moe,
                estimates = est,
                sample = count) %>%
         knitr::kable()
    )
}
```

## Trips in Region

```{r trip_totals}
#| label: fig-total-trips
#| fig-cap: Total Trips in Region
#| fig-height: 3.5
#| fig-width: 4.5

ggplot(trip_summary, 
       aes(x = factor(survey_year), y = est_rounded)) +
  geom_col(aes(fill = factor(survey_year)), 
               width = 0.7
           ) +
  geom_text(aes(label = round(est_rounded, 1)),
                vjust = -0.3
            ) +
  geom_hline(yintercept = 0, color = "#cbcbcb", linewidth = 0.3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) + 
  scale_fill_manual(values = psrc_fill_dec[["survey_year"]]) + 
  psrc_style() + 
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title = element_blank())
```

```{r all_modes}
#| label: fig-all-modes
#| fig-cap: Trips by Mode
#| fig-height: 3.5
#| fig-width: 10

ggplot(filter(mode_summary, !is.na(mode_class_5)),
       aes(x=factor(survey_year), y=prop, fill=factor(survey_year))) +
  geom_col(width = 0.8) +
  geom_text(aes(label=scales::percent(prop, accuracy = 1)),
                vjust = -0.3) +
  geom_hline(yintercept = 0, color = "#cbcbcb", linewidth = 0.3) +
  facet_wrap(~ mode_class_5, ncol=5)+
  scale_fill_manual(values = psrc_fill_dec[["mode_class_5"]])+
  scale_y_continuous(labels = scales::percent, 
                     expand = expansion(mult = c(0, 0.1))) +
  psrc_style() +
  theme(legend.position = "none", # remove legend
        strip.text = element_text(size = 10.5), # adjust text size on facet title (variable values)
        panel.grid.major.y = element_blank(),
        panel.spacing.x = unit(0.5, "lines"), # add spacing between facet grids
        axis.title = element_blank(),
        axis.text.x = element_text(size = 10.5), # adjust text size for axis text (years)
        axis.text.y = element_blank())
```

## Walking and Bike/Micromobility Trips

- Share of walk trips in region is down from 2021.
- Share of bike/micro trips in region is up from 2021.

### Walking
```{r walk_total}
#| label: fig-walk-total
#| fig-cap: Walk Trips
#| fig-height: 2.5
#| fig-width: 3

single_cat_output(mode_summary, "Walk")
```

```{r walk_total_tbl}
#| label: tbl-walk-total

tbl_output(mode_summary, "Walk")
```

### Bike/Micromobility
```{r bike_total}
#| label: fig-bike-total
#| fig-cap: Bike/Micromobility Trips
#| fig-height: 2.5
#| fig-width: 3

single_cat_output(mode_summary, "Bike/Micromobility")
```

```{r bike_total_tbl}
#| label: tbl-bike-total

tbl_output(mode_summary, "Bike/Micromobility")
```

### Trips by Household Income

- Walk trips in 2025 for $50k-$150k groups are down from 2021.
- Walk trips in 2025 for $150k+ group are similar to 2021.
- Bike/micro trips in 2025 for under $50k group are up significantly from all prior years; hard to say anything for other groups.

#### Walking
```{r walk_income_combined}
#| label: fig-walk-income
#| fig-cap: Walk Trips by Income
#| fig-height: 3
#| fig-width: 8

multi_cat_output(walk_bike_by_income_combined, mode = "Walk", group_var = "income")
```

```{r walk_income_tbl}
#| label: tbl-walk-income

tbl_output(walk_bike_by_income_combined, mode = "Walk")
```

#### Bike/Micromobility
```{r bike_income_binary}
#| label: fig-bike-income
#| fig-cap: Bike/Micromobility Trips by Income
#| fig-height: 3
#| fig-width: 5

multi_cat_output(walk_bike_by_income_binary, mode = "Bike/Micromobility", group_var = "income_binary")
```

```{r bike_income_tbl}
#| label: tbl-bike-income

tbl_output(walk_bike_by_income_binary, mode = "Bike/Micromobility")
```

#### Bike/Micromobility (exp)
```{r bike_income_exp}
#| fig-cap: Bike/Micromobility Trips by Income (exp.)
#| fig-height: 3
#| fig-width: 9
#| fig-align: center

multi_cat_output(walk_bike_by_income_combined, mode = "Bike/Micromobility", group_var = "income")
```

### Trips by Race/Ethnicity

- Only Some Other Race shows an increase in walk trips from prior years, but 2021-2023-2025 looks very unusual
- Increase in bike/micro by POC group hard to defend when disaggregated

#### Walking (exp)
```{r walk_race_exp}
#| fig-cap: Walk Trips by Race (exp.)
#| fig-height: 3
#| fig-width: 10.5

multi_cat_output(walk_bike_by_race, mode = "Walk", group_var = "race_ethnicity")
```

#### Bike/Micromobility (exp)
```{r bike_race_exp}
#| fig-cap: Bike/Micromobility Trips by Race (exp.)
#| fig-height: 3
#| fig-width: 10.5

multi_cat_output(walk_bike_by_race, mode = "Bike/Micromobility", group_var = "race_ethnicity")
```

#### Walking
```{r walk_race}
#| fig-cap: Walk Trips by Race
#| fig-height: 3
#| fig-width: 5

multi_cat_output(walk_bike_by_race_binary, mode = "Walk", group_var = "race_binary")
```

#### Bike/Micromobility
```{r bike_race}
#| fig-cap: Bike/Micromobility Trips by Race
#| fig-height: 3
#| fig-width: 5

multi_cat_output(walk_bike_by_race_binary, mode = "Bike/Micromobility", group_var = "race_binary")
```

### Trips by Age

- Walk trips up for Under 18 group but down in others
- Large increase in bike/micro for 18-34 group after 2021

#### Walking
```{r walk_age}
#| label: fig-walk-age
#| fig-cap: Walk Trips by Age
#| fig-height: 3
#| fig-width: 8

multi_cat_output(walk_bike_by_age, mode = "Walk", group_var = "age")
```

```{r walk_age_tbl}
#| label: tbl-walk-age

tbl_output(walk_bike_by_age, mode = "Walk")
```

#### Bike/Micromobility
```{r bike_age}
#| label: fig-bike-age
#| fig-cap: Bike/Micromobility Trips by Age
#| fig-height: 3
#| fig-width: 8

multi_cat_output(walk_bike_by_age, mode = "Bike/Micromobility", group_var = "age")
```

```{r bike_age_tbl}
#| label: tbl-bike-age

tbl_output(walk_bike_by_age, mode = "Bike/Micromobility")
```

### Trips by Gender

- Non-binary/Other group appears to walk & bike more than other groups, but small sizes make it hard to say
- Walking trips in 2025 down for men; slight increase for women

#### Walking
```{r walk_gender}
#| label: fig-walk-gender
#| fig-cap: Walk Trips by Gender
#| fig-height: 3
#| fig-width: 6

multi_cat_output(walk_bike_by_gender, mode = "Walk", group_var = "gender")
```

```{r walk_rgc_tbl}
#| label: tbl-walk-gender

tbl_output(walk_bike_by_gender, mode = "Walk")
```

#### Bike/Micromobility
```{r bike_gender}
#| label: fig-bike-gender
#| fig-cap: Bike/Micromobility Trips by Gender
#| fig-height: 3
#| fig-width: 6

multi_cat_output(walk_bike_by_gender, mode = "Bike/Micromobility", group_var = "gender")
```

```{r bike_rgc_tbl}
#| label: tbl-bike-gender

tbl_output(walk_bike_by_gender, mode = "Bike/Micromobility")
```

### Trips based on Home Location

- Number of walk & bike trips by people living in RGCs is a fraction of trips made by people living outside them, but shares are much higher
- Walk trips for people in RGCs remain elevated after 2021
- Share of bike trips for people in RGCs shows significant growth after 2021

#### Walking
```{r walk_rgc}
#| label: fig-walk-rgc
#| fig-cap: Walk Trips by Home Location
#| fig-height: 3
#| fig-width: 5

multi_cat_output(walk_bike_by_rgc, mode = "Walk", group_var = "in_rgc")
```

```{r walk_rgc_tbl}
#| label: tbl-walk-rgc

tbl_output(walk_bike_by_rgc, mode = "Walk")
```

#### Bike/Micromobility
```{r bike_rgc}
#| label: fig-bike-rgc
#| fig-cap: Bike/Micromobility Trips by Home Location
#| fig-height: 3
#| fig-width: 5

multi_cat_output(walk_bike_by_rgc, mode = "Bike/Micromobility", group_var = "in_rgc")
```

```{r bike_rgc_tbl}
#| label: tbl-bike-rgc

tbl_output(walk_bike_by_rgc, mode = "Bike/Micromobility")
```

### Trips by County

- Walk share significantly decreased in Kitsap, Pierce in 2025

#### Walking
```{r walk_county}
#| label: fig-walk-county
#| fig-cap: Walk Trips by Home County
#| fig-height: 3
#| fig-width: 8

multi_cat_output(walk_bike_by_county, mode = "Walk", group_var = "home_county")
```

```{r walk_county_tbl}
#| label: tbl-walk-county

tbl_output(walk_bike_by_county, mode = "Walk")
```

#### Bike/Micromobility
```{r bike_county}
#| label: fig-bike-county
#| fig-cap: Bike/Micromobility Trips by Home County
#| fig-height: 3
#| fig-width: 8

multi_cat_output(walk_bike_by_county, mode = "Bike/Micromobility", group_var = "home_county")
```

```{r bike_county_tbl}
#| label: tbl-bike-county

tbl_output(walk_bike_by_county, mode = "Bike/Micromobility")
```
