---
title: "Transportation Operator Committee HHTS 2021 notes"
output:
  html_document:
    df_print: paged
---

https://www.psrc.org/committee/transportation-operators-committee

July 27 meeting

We are gathering materials information that could be useful to the transportation operators committee

```{r }
library(psrc.travelsurvey)
library(dplyr)
library(stringr)
library(ggplot2)
library(psrcplot)
library(forcats)
library(odbc)
library(DBI)
library(tidyr)
```

```{r Elmer variable metadata}
# connect to Elmer
db.connect <- function(adatabase) {
  elmer_connection <- dbConnect(odbc(),
                                driver = "SQL Server",
                                server = "AWS-PROD-SQL\\SOCKEYE",
                                database = adatabase,
                                trusted_connection = "yes"
  )
}

# read table
read.dt <- function(adatabase, atable) {
  elmer_connection <- db.connect(adatabase)
  dtelm <- dbReadTable(elmer_connection, SQL(atable))
  dbDisconnect(elmer_connection)
  return(dtelm)
}

# read-in variable metadata table for levels
vars_meta <- read.dt('Elmer', 'HHSurvey.variable_metadata')
```

Ashley Asmus presentation by Met Council on their 2022 transit on-board survey provides a lot of inspiration

https://metrocouncil.org/Council-Meetings/Committees/Transportation-Committee/2022/March-28,-2022/info-1_TransitOnBoardSurvey_2022-03-28.aspx

* Estimate of total transit trips regionally in 2019 vs 2021
 - by access mode
 - by geography?
 - by income group
 - by trip purpose
 - by time of day
 - by race category
 - by age category
 - by lifecycle
 - by gender
 - by worker status
 
 * Overall mode share shift in 2019 vs 2021
 - by access mode
 - by geography?
 - by income group
 - by trip purpose
 - by time of day
 - by race category
 - by age category
 - by gender
 - trip_length
  * commute_mode_pre_covid vs commute_mode
   - by income, race, age, gender
 * commute_freq_pre_covid vs commute_freq
    - by income, race, age, gender

* telecommute_freq_pre_covid vs telecommute_freq
  - by income, race, age, gender
 
 Pull in data that might be useful
 
 Origin County and Destination County are missing
 Home County appears like it may not be filled in everywhere for 2021?
```{r}
mode_vars<-c('mode_1', 'mode_simple')
other_vars<-c('final_home_rgcnum', 'hhsize', 'vehicle_count',  "hhincome_broad", 'rent_own', 'res_dur', 'student', 'education',  'hhincome_detailed', "age", "age_category", 'race_category', 'race_eth_broad', 'gender', 'employment',  'lifecycle', 'mode_acc', 'dest_purpose_cat', 'origin_purpose_cat', 'final_home_is_rgc', 'race_eth_poc_update', 'race_eth_apoc')
trip_path_dist<-'trip_path_distance'
all_vars<-c(mode_vars, other_vars, trip_path_dist)
```



 
 
```{r}
transit_data_17_19<- get_hhts("2017_2019", "t", vars=all_vars)%>% mutate(year=ifelse(survey=='2017_2019', '2017/2019', '2021'))
```


```{r}
transit_data_21<- get_hhts("2021", "t", vars=all_vars)%>% mutate(year=ifelse(survey=='2017_2019', '2017/2019', '2021'))

```
 
 
 Start simple - What was the total estimated transit trips in 2017/2019 vs 2021?
 
 These are trips made by adults only
 
```{r, include=TRUE}
trips_by_mode_17_19<-hhts_count(transit_data_17_19, group_vars='mode_simple')%>%
  filter(mode_simple!='Total')
trips_by_mode_21<-hhts_count(transit_data_21, group_vars='mode_simple')%>%filter(mode_simple!='Total')

trips_by_mode_17_19_21<-merge(trips_by_mode_17_19, trips_by_mode_21, by='mode_simple', suffixes=c('17_19', '21'))
trips_by_mode<-rbind(trips_by_mode_17_19, trips_by_mode_21)%>% mutate(year=ifelse(survey=='2017_2019', '2017/2019', '2021'))
                                                                               

```
```{r}
write.csv(trips_by_mode, 'trips_by_mode.csv')
```

```{r, include=TRUE}
p<-static_bar_chart(t=trips_by_mode, y='mode_simple', x='share', fill='year', moe='share_moe', est='percent')
  
  
p
```
From 2017/2019 to 2021, the transit share went down, drive shares went down, and walking shares went up.

```{r, include=TRUE}
trips_by_mode_no_drive<-trips_by_mode%>%filter(!mode_simple %in% c('Drive', 'Bike', 'Other'))
p<-static_bar_chart(t=trips_by_mode_no_drive, y='mode_simple', x='share', fill='year', moe='share_moe')
  

print(p)
```



Group variables into meaningful categories and redo analysis
```{r}
transit_data_17_19<-transit_data_17_19%>%mutate(NoVehicles=ifelse(vehicle_count=='0 (no vehicles)', 'No Vehicles', "Has Vehicles"))%>%
  mutate(hhsize_simple=case_when(hhsize== '4 people' ~'4 or more people',                                                                     hhsize== '5 people' ~'4 or more people',
                                  hhsize== '6 people' ~'4 or more people',
                                  hhsize== '7 people' ~'4 or more people',
                                  hhsize== '8 people' ~'4 or more people',
                                  hhsize== '12 people' ~'4 or more people',
                                  TRUE ~ hhsize))%>%
  mutate(hhincome_100= case_when(hhincome_broad=='$100,000-$199,000' ~ '$100,000 or more',
                                 hhincome_broad=='$200,000 or more' ~ '$100,000 or more',
                                 TRUE~hhincome_broad))%>%
  mutate(edu_simple= case_when(education=='Bachelor degree' ~ 'Bachelors or higher', 
                               education=='Graduate/Post-graduate degree' ~ 'Bachelors or higher',
                               TRUE ~ 'Less than Bachelors degree'))%>%
  mutate(age_grp= case_when(age=='75-84 years' ~ '75 years or older', 
                            age == '85 or years older' ~ '75 years or older',
                            TRUE ~ age))%>%
  mutate(gender_grp= case_when(gender == 'Prefer not to answer' ~ 'Non-binary, another, prefer not to answer',
                            gender=='Not listed here / prefer not to answer' ~ 'Non-binary, another, prefer not to answer',
                            gender=='Non-Binary'~ 'Non-binary, another, prefer not to answer',
                            gender=='Another'~ 'Non-binary, another, prefer not to answer',
                            TRUE ~ gender))%>%mutate(work_purpose=ifelse(dest_purpose_cat=='Work', 'Work', 'Not Work'))%>%

    mutate(race_short= str_extract(race_eth_broad,  "^[^ ]+"))%>%
   mutate(simple_purpose=ifelse(dest_purpose_cat=='Home', origin_purpose_cat, dest_purpose_cat))%>%
  mutate(simple_purpose=case_when(simple_purpose=='Work'~ 'Work/School',
                                  simple_purpose=='School'~ 'Work/School',
                                  simple_purpose=='Work-related'~ 'Work/School',
                                  simple_purpose=='Shop'~ 'Shop and Errands',
                                  simple_purpose=='Escort'~ 'Shop and Errands',
                                  simple_purpose=='Errand/Other'~ 'Shop and Errands',
                                  simple_purpose=='Change mode'~ 'Shop and Errands',
                                  simple_purpose=='Social/Recreation' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Meal' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Home' ~'Shop and Errands',
                                  is.na(simple_purpose) ~ 'Shop and Errands',
                                  TRUE ~ simple_purpose))%>%mutate(rgc=as.factor(final_home_is_rgc))%>%
  mutate(non_motorized_mode=ifelse((mode_simple=='Walk'|mode_simple=='Bike'),'Walk/Bike', 'Not Walk/Bike'))


transit_data_21<-transit_data_21%>%mutate(NoVehicles=ifelse(vehicle_count=='0 (no vehicles)', 'No Vehicles', "Has Vehicles"))%>%
  mutate(hhsize_simple=case_when(hhsize== '4 people' ~'4 or more people',                                                                    hhsize== '5 people' ~'4 or more people',
                                 hhsize== '6 people' ~'4 or more people',
                                 hhsize== '7 people' ~'4 or more people',
                                 hhsize== '8 people' ~'4 or more people',
                                 hhsize== '12 people' ~'4 or more people',
                                 TRUE ~ hhsize)) %>%
  mutate(hhincome_100= case_when(hhincome_broad=='$100,000-$199,000' ~ '$100,000 or more',
                                 hhincome_broad=='$200,000 or more' ~ '$100,000 or more',
                                 TRUE~hhincome_broad))%>%
  mutate(edu_simple= case_when(education=='Bachelor degree' ~ 'Bachelors or higher', 
                               education=='Graduate/Post-graduate degree' ~ 'Bachelors or higher',
                               TRUE ~ 'Less than Bachelors degree'))%>%
  mutate(age_grp= case_when(age=='75-84 years' ~ '75 years or older', 
                            age == '85 or years older' ~ '75 years or older',
                            TRUE ~ age))%>%
  mutate(gender_grp= case_when(gender == 'Prefer not to answer' ~ 'Non-binary, another, prefer not to answer',
                            gender=='Not listed here / prefer not to answer' ~ 'Non-binary, another, prefer not to answer',
                            gender=='Non-Binary'~ 'Non-binary, another, prefer not to answer', 
                            gender=='Another'~ 'Non-binary, another, prefer not to answer',
                            TRUE ~ gender))%>%
  mutate(work_purpose=ifelse(dest_purpose_cat=='Work', 'Work', 'Not Work'))%>%
  mutate(simple_purpose=ifelse(dest_purpose_cat=='Home', origin_purpose_cat, dest_purpose_cat))%>%
  mutate(simple_purpose=case_when(simple_purpose=='Work'~ 'Work/School',
                                  simple_purpose=='School'~ 'Work/School',
                                  simple_purpose=='Work-related'~ 'Work/School',
                                  simple_purpose=='Shop'~ 'Shop and Errands',
                                  simple_purpose=='Escort'~ 'Shop and Errands',
                                  simple_purpose=='Errand/Other'~ 'Shop and Errands',
                                  simple_purpose=='Change mode'~ 'Shop and Errands',
                                  simple_purpose=='Social/Recreation' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Meal' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Meal' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Home' ~'Shop and Errands',
                                  is.na(simple_purpose) ~ 'Shop and Errands',
                                  TRUE ~ simple_purpose))%>%mutate(rgc=as.factor(final_home_is_rgc))%>%
  mutate(non_motorized_mode=ifelse((mode_simple=='Walk'|mode_simple=='Bike'),'Walk/Bike', 'Not Walk/Bike'))

```


```{r}



simple_groupings<-c('race_eth_broad'='Race/Ethnicity','hhincome_100_f'='Household Income','edu_simple'= 'Education Level', 'hhsize_simple'= 'Household Size','NoVehicles'= 'Household Vehicles', 'simple_purpose'="Trip Purpose", 'rgc'='Regional Growth Center', 'gender_grp'='Gender')


```

```{r}
# bunch of ordering and recoding

transit_data_17_19$hhincome_100_f=factor(transit_data_17_19$hhincome_100, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))

transit_data_21$hhincome_100_f=factor(transit_data_21$hhincome_100, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))


```


```{r, include=TRUE}
trips_by_mode_17_19<-hhts_count(transit_data_17_19, group_vars=c('race_eth_poc_update', 'gender','mode_simple'))%>%
  filter(mode_simple!='Total')%>%filter(gender=='Male' | gender=='Female')%>%filter(race_eth_poc_update!='Total')
trips_by_mode_21<-hhts_count(transit_data_21, group_vars=c( 'race_eth_poc_update','gender','mode_simple'))%>%filter(mode_simple!='Total')%>%filter(gender=='Male' | gender=='Female')%>%filter(race_eth_poc_update!='Total')



                                                                               

```

```{r}
static_facet_column_chart(t=trips_by_mode_17_19, x= 'gender', y='share', fill= 'race_eth_poc_update', facet='mode_simple',  moe='share_moe')
```



```{r}
trips_by_mode<-rbind(trips_by_mode_17_19, trips_by_mode_21)%>% mutate(year=ifelse(survey=='2017_2019', '2017/2019', '2021'))
```




 
```{r, include=TRUE}
trips_by_mode_17_19<-hhts_count(transit_data_17_19, group_vars=c('race_eth_poc_update', 'gender','mode_simple'))%>%
  filter(mode_simple!='Total')%>%filter(gender=='Male' | gender=='Female')%>%filter(race_eth_poc_update!='Total')%>%filter(mode_simple=='Walk')%>%filter('race_eth_poc_update' !='No response')
trips_by_mode_21<-hhts_count(transit_data_21, group_vars=c( 'race_eth_poc_update','gender','mode_simple'))%>%filter(mode_simple!='Total')%>%filter(gender=='Male' | gender=='Female')%>%filter(race_eth_poc_update!='Total')%>%filter(mode_simple=='Walk')%>%filter('race_eth_poc_update' !='No response')



                                                                               

```


```{r}
trips_by_mode<-rbind(trips_by_mode_17_19, trips_by_mode_21)%>% mutate(year=ifelse(survey=='2017_2019', '2017/2019', '2021'))
```



```{r}
static_column_chart(t=trips_by_mode_17_19, x= 'gender', y='share', fill= 'race_eth_poc_update', moe='share_moe')
```
```{r}
static_column_chart(t=trips_by_mode_21, x= 'gender', y='share', fill= 'race_eth_poc_update', moe='share_moe')
```


```{r}
static_facet_column_chart(t=trips_by_mode, x= 'gender', y='share', fill= 'race_eth_poc_update', facet='year')+coord_cartesian(ylim=c(0, .15))
```

 
```{r, include=TRUE}
trips_by_mode_17_19<-hhts_count(transit_data_17_19, group_vars=c('race_eth_poc_update', 'gender','mode_simple'))%>%
  filter(mode_simple!='Total')%>%filter(gender=='Male' | gender=='Female')%>%filter(race_eth_poc_update!='Total')%>%filter(mode_simple=='Transit')%>%filter(race_eth_poc_update !='No response')
trips_by_mode_21<-hhts_count(transit_data_21, group_vars=c( 'race_eth_poc_update','gender','mode_simple'))%>%filter(mode_simple!='Total')%>%filter(gender=='Male' | gender=='Female')%>%filter(race_eth_poc_update!='Total')%>%filter(mode_simple=='Transit')%>%filter(race_eth_poc_update !='No response')



                                                                               

```


```{r}
trips_by_mode<-rbind(trips_by_mode_17_19, trips_by_mode_21)%>% mutate(year=ifelse(survey=='2017_2019', '2017/2019', '2021'))
```



```{r}
static_facet_column_chart(t=trips_by_mode, x= 'gender', y='share', fill= 'race_eth_poc_update', facet='year')+coord_cartesian(ylim=c(0, .08))
```
```{r}
static_column_chart(t=trips_by_mode_21, x= 'gender', y='share', fill= 'race_eth_poc_update', moe='share_moe')
```

 
```{r, include=TRUE}
trips_by_mode_17_19<-hhts_count(transit_data_17_19, group_vars=c('race_eth_poc_update', 'gender','mode_simple'))%>%
  filter(mode_simple!='Total')%>%filter(gender=='Male' | gender=='Female')%>%filter(race_eth_poc_update!='Total')%>%filter(mode_simple=='Walk')
trips_by_mode_21<-hhts_count(transit_data_21, group_vars=c( 'race_eth_poc_update','gender','mode_simple'))%>%filter(mode_simple!='Total')%>%filter(gender=='Male' | gender=='Female')%>%filter(race_eth_poc_update!='Total')%>%filter(mode_simple=='Walk')



                                                                               

```


```{r}
trips_by_mode<-rbind(trips_by_mode_17_19, trips_by_mode_21)%>% mutate(year=ifelse(survey=='2017_2019', '2017/2019', '2021'))
```



```{r}
static_column_chart(t=trips_by_mode_17_19, x= 'gender', y='share', fill= 'race_eth_poc_update', moe='share_moe')
```
```{r}
static_column_chart(t=trips_by_mode_21, x= 'gender', y='share', fill= 'race_eth_poc_update', moe='share_moe')
```


```{r}
static_facet_column_chart(t=trips_by_mode, x= 'gender', y='share', fill= 'race_eth_poc_update', facet='year')+coord_cartesian(ylim=c(0, .15))
```
