---
title: "Transportation Operator Committee HHTS 2021 notes"
output:
  html_document:
    df_print: paged
---

https://www.psrc.org/committee/transportation-operators-committee

July 27 meeting

We are gathering materials information that could be useful to the transportation operators committee

```{r }
library(psrc.travelsurvey)
library(dplyr)
library(stringr)
library(ggplot2)
library(psrcplot)
library(forcats)
library(odbc)
library(DBI)
library(tidyr)
```

```{r Elmer variable metadata}
# connect to Elmer
db.connect <- function(adatabase) {
  elmer_connection <- dbConnect(odbc(),
                                driver = "SQL Server",
                                server = "AWS-PROD-SQL\\SOCKEYE",
                                database = adatabase,
                                trusted_connection = "yes"
  )
}

# read table
read.dt <- function(adatabase, atable) {
  elmer_connection <- db.connect(adatabase)
  dtelm <- dbReadTable(elmer_connection, SQL(atable))
  dbDisconnect(elmer_connection)
  return(dtelm)
}

# read-in variable metadata table for levels
vars_meta <- read.dt('Elmer', 'HHSurvey.variable_metadata')
```

Ashley Asmus presentation by Met Council on their 2022 transit on-board survey provides a lot of inspiration

https://metrocouncil.org/Council-Meetings/Committees/Transportation-Committee/2022/March-28,-2022/info-1_TransitOnBoardSurvey_2022-03-28.aspx

* Estimate of total transit trips regionally in 2019 vs 2021
 - by access mode
 - by geography?
 - by income group
 - by trip purpose
 - by time of day
 - by race category
 - by age category
 - by lifecycle
 - by gender
 - by worker status
 
 * Overall mode share shift in 2019 vs 2021
 - by access mode
 - by geography?
 - by income group
 - by trip purpose
 - by time of day
 - by race category
 - by age category
 - by gender
 - trip_length
  * commute_mode_pre_covid vs commute_mode
   - by income, race, age, gender
 * commute_freq_pre_covid vs commute_freq
    - by income, race, age, gender

* telecommute_freq_pre_covid vs telecommute_freq
  - by income, race, age, gender
 
 Pull in data that might be useful
 
 Origin County and Destination County are missing
 Home County appears like it may not be filled in everywhere for 2021?
```{r}
mode_vars<-c('mode_1', 'mode_simple')
other_vars<-c('final_home_rgcnum', 'hhsize', 'vehicle_count',  "hhincome_broad", 'rent_own', 'res_dur', 'student', 'education',  'hhincome_detailed', "age", "age_category", 'race_category', 'race_eth_broad', 'gender', 'employment',  'lifecycle', 'mode_acc', 'dest_purpose_cat', 'origin_purpose_cat', 'final_home_is_rgc')
trip_path_dist<-'trip_path_distance'
all_vars<-c(mode_vars, other_vars, trip_path_dist)
```



 
 
```{r}
transit_data_17_19<- get_hhts("2017_2019", "t", vars=all_vars)%>% mutate(year=ifelse(survey=='2017_2019', '2017/2019', '2021'))
```


```{r}
transit_data_21<- get_hhts("2021", "t", vars=all_vars)%>% mutate(year=ifelse(survey=='2017_2019', '2017/2019', '2021'))

```
 
 
 Start simple - What was the total estimated transit trips in 2017/2019 vs 2021?
 
 These are trips made by adults only
 
```{r, include=TRUE}
trips_by_mode_17_19<-hhts_count(transit_data_17_19, group_vars='mode_simple')%>%
  filter(mode_simple!='Total')
trips_by_mode_21<-hhts_count(transit_data_21, group_vars='mode_simple')%>%filter(mode_simple!='Total')

trips_by_mode_17_19_21<-merge(trips_by_mode_17_19, trips_by_mode_21, by='mode_simple', suffixes=c('17_19', '21'))
trips_by_mode<-rbind(trips_by_mode_17_19, trips_by_mode_21)%>% mutate(year=ifelse(survey=='2017_2019', '2017/2019', '2021'))
                                                                               

```
```{r}
write.csv(trips_by_mode, 'trips_by_mode.csv')
```

```{r, include=TRUE}
p<-create_bar_chart(t=trips_by_mode, w.x='mode_simple', w.y='share', f='year', w.moe='share_moe', est.type='percent', w.color = 'psrc_light')
  
  


print(p)
psrc_print_palette()
```
From 2017/2019 to 2021, the transit share went down, drive shares went down, and walking shares went up.

```{r, include=TRUE}
trips_by_mode_no_drive<-trips_by_mode%>%filter(!mode_simple %in% c('Drive', 'Bike', 'Other'))
p<-create_bar_chart(t=trips_by_mode_no_drive, w.x='mode_simple', w.y='share', f='year', w.moe='share_moe', est.type='percent', w.color = 'psrc_light')
  

print(p)
```



Group variables into meaningful categories and redo analysis
```{r}
transit_data_17_19<-transit_data_17_19%>%mutate(NoVehicles=ifelse(vehicle_count=='0 (no vehicles)', 'No Vehicles', "Has Vehicles"))%>%
  mutate(hhsize_simple=case_when(hhsize== '4 people' ~'4 or more people',                                                                     hhsize== '5 people' ~'4 or more people',
                                  hhsize== '6 people' ~'4 or more people',
                                  hhsize== '7 people' ~'4 or more people',
                                  hhsize== '8 people' ~'4 or more people',
                                  hhsize== '12 people' ~'4 or more people',
                                  TRUE ~ hhsize))%>%
  mutate(hhincome_100= case_when(hhincome_broad=='$100,000-$199,000' ~ '$100,000 or more',
                                 hhincome_broad=='$200,000 or more' ~ '$100,000 or more',
                                 TRUE~hhincome_broad))%>%
  mutate(edu_simple= case_when(education=='Bachelor degree' ~ 'Bachelors or higher', 
                               education=='Graduate/Post-graduate degree' ~ 'Bachelors or higher',
                               TRUE ~ 'Less than Bachelors degree'))%>%
  mutate(age_grp= case_when(age=='75-84 years' ~ '75 years or older', 
                            age == '85 or years older' ~ '75 years or older',
                            TRUE ~ age))%>%
  mutate(gender_grp= case_when(gender == 'Prefer not to answer' ~ 'Non-binary, another, prefer not to answer',
                            gender=='Not listed here / prefer not to answer' ~ 'Non-binary, another, prefer not to answer',
                            gender=='Non-Binary'~ 'Non-binary, another, prefer not to answer',
                            gender=='Another'~ 'Non-binary, another, prefer not to answer',
                            TRUE ~ gender))%>%mutate(work_purpose=ifelse(dest_purpose_cat=='Work', 'Work', 'Not Work'))%>%

    mutate(race_short= str_extract(race_eth_broad,  "^[^ ]+"))%>%
   mutate(simple_purpose=ifelse(dest_purpose_cat=='Home', origin_purpose_cat, dest_purpose_cat))%>%
  mutate(simple_purpose=case_when(simple_purpose=='Work'~ 'Work/School',
                                  simple_purpose=='School'~ 'Work/School',
                                  simple_purpose=='Work-related'~ 'Work/School',
                                  simple_purpose=='Shop'~ 'Shop and Errands',
                                  simple_purpose=='Escort'~ 'Shop and Errands',
                                  simple_purpose=='Errand/Other'~ 'Shop and Errands',
                                  simple_purpose=='Change mode'~ 'Shop and Errands',
                                  simple_purpose=='Social/Recreation' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Meal' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Home' ~'Shop and Errands',
                                  is.na(simple_purpose) ~ 'Shop and Errands',
                                  TRUE ~ simple_purpose))%>%mutate(rgc=as.factor(final_home_is_rgc))%>%
  mutate(non_motorized_mode=ifelse((mode_simple=='Walk'|mode_simple=='Bike'),'Walk/Bike', 'Not Walk/Bike'))


transit_data_21<-transit_data_21%>%mutate(NoVehicles=ifelse(vehicle_count=='0 (no vehicles)', 'No Vehicles', "Has Vehicles"))%>%
  mutate(hhsize_simple=case_when(hhsize== '4 people' ~'4 or more people',                                                                    hhsize== '5 people' ~'4 or more people',
                                 hhsize== '6 people' ~'4 or more people',
                                 hhsize== '7 people' ~'4 or more people',
                                 hhsize== '8 people' ~'4 or more people',
                                 hhsize== '12 people' ~'4 or more people',
                                 TRUE ~ hhsize)) %>%
  mutate(hhincome_100= case_when(hhincome_broad=='$100,000-$199,000' ~ '$100,000 or more',
                                 hhincome_broad=='$200,000 or more' ~ '$100,000 or more',
                                 TRUE~hhincome_broad))%>%
  mutate(edu_simple= case_when(education=='Bachelor degree' ~ 'Bachelors or higher', 
                               education=='Graduate/Post-graduate degree' ~ 'Bachelors or higher',
                               TRUE ~ 'Less than Bachelors degree'))%>%
  mutate(age_grp= case_when(age=='75-84 years' ~ '75 years or older', 
                            age == '85 or years older' ~ '75 years or older',
                            TRUE ~ age))%>%
  mutate(gender_grp= case_when(gender == 'Prefer not to answer' ~ 'Non-binary, another, prefer not to answer',
                            gender=='Not listed here / prefer not to answer' ~ 'Non-binary, another, prefer not to answer',
                            gender=='Non-Binary'~ 'Non-binary, another, prefer not to answer', 
                            gender=='Another'~ 'Non-binary, another, prefer not to answer',
                            TRUE ~ gender))%>%
  mutate(work_purpose=ifelse(dest_purpose_cat=='Work', 'Work', 'Not Work'))%>%
  mutate(simple_purpose=ifelse(dest_purpose_cat=='Home', origin_purpose_cat, dest_purpose_cat))%>%
  mutate(simple_purpose=case_when(simple_purpose=='Work'~ 'Work/School',
                                  simple_purpose=='School'~ 'Work/School',
                                  simple_purpose=='Work-related'~ 'Work/School',
                                  simple_purpose=='Shop'~ 'Shop and Errands',
                                  simple_purpose=='Escort'~ 'Shop and Errands',
                                  simple_purpose=='Errand/Other'~ 'Shop and Errands',
                                  simple_purpose=='Change mode'~ 'Shop and Errands',
                                  simple_purpose=='Social/Recreation' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Meal' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Meal' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Home' ~'Shop and Errands',
                                  is.na(simple_purpose) ~ 'Shop and Errands',
                                  TRUE ~ simple_purpose))%>%mutate(rgc=as.factor(final_home_is_rgc))%>%
  mutate(non_motorized_mode=ifelse((mode_simple=='Walk'|mode_simple=='Bike'),'Walk/Bike', 'Not Walk/Bike'))

```


```{r}



simple_groupings<-c('race_eth_broad'='Race/Ethnicity','hhincome_100_f'='Household Income','edu_simple'= 'Education Level', 'hhsize_simple'= 'Household Size','NoVehicles'= 'Household Vehicles', 'simple_purpose'="Trip Purpose", 'rgc'='Regional Growth Center', 'gender_grp'='Gender')


```

```{r}
# bunch of ordering and recoding

transit_data_17_19$hhincome_100_f=factor(transit_data_17_19$hhincome_100, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))

transit_data_21$hhincome_100_f=factor(transit_data_21$hhincome_100, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))


```

```{r Updated ggplot2 style and order}
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  # cat_name <- eval(names(g))
  cat_name <- names(g)
  print(cat_name)
  transit_summs_2017_2019 <- hhts_count(transit_data_17_19, 
                                        group_vars=c(names(g),'mode_simple'), 
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter(cat_name!='Total') %>%
    filter(mode_simple=='Transit')
  

  
  transit_summs_2021 <- hhts_count(transit_data_21, 
                                   group_vars=c(names(g),'mode_simple'), 
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter(cat_name!='Total') %>%
    filter(mode_simple=='Transit')
  
  if(cat_name=='race_eth_broad'){

     transit_summs_2017_2019 <- transit_summs_2017_2019 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
     
        transit_summs_2021 <- transit_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  transit_summs_2017_2019$race_eth_broad<-factor(transit_summs_2017_2019$race_eth_broad, levels=
                                                  c('White', 'Other', 'Hispanic', 'Black', 'Asian'))
  
   transit_summs_2021$race_eth_broad<-factor(transit_summs_2021$race_eth_broad, levels=
                                                  c('White', 'Other', 'Hispanic', 'Black', 'Asian'))
 

  
  }
  
   if(cat_name=='hhincome_100_f'){
     
     transit_summs_2017_2019$hhincome_100_f=factor( transit_summs_2017_2019$hhincome_100_f, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))

 transit_summs_2021$hhincome_100_f=factor(transit_summs_2021$hhincome_100_f, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))
     
     
   }

  transit_summs_long<- rbind(transit_summs_2017_2019, transit_summs_2021) %>% 
    mutate(survey = str_replace_all(survey, "_", "/"))

  
   # factorize variable column. Adjust levels in function definition
  #transit_summs_long <- factor_variable_column(names(g), transit_summs_long)
  
  transit_summs_2021_only<-  transit_summs_2021%>%filter(mode_simple=='Transit')
  
 # p0<-create_bar_chart(t=transit_summs_2021_only , w.x=cat_name, w.y='share',  f=cat_name,w.moe='share_moe', est.type='percent', w.color='psrc_light')+ t#heme(axis.text.x = element_text(size=10,color="#4C4C4C"))+coord_flip()
#  print(p0)
  
   p0<-create_bar_chart(t=transit_summs_2021 , w.x=cat_name, w.y='share', f='mode_simple', w.moe='share_moe', est.type='percent', w.color = 'psrc_light', w.interactive=TRUE)+
  xlab(as.character(g[cat_name])) + ylab("Transit mode share")+scale_fill_manual(values="#73CFCB")+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+ theme(axis.title.x = element_text(size=12,color="#4C4C4C"))+theme(axis.title.y = element_text(size=12,color="#4C4C4C"))+coord_flip()
  print(p0)
  
  p1<-create_bar_chart(t=transit_summs_long , w.x=cat_name, w.y='share', f='survey', w.moe='share_moe', est.type='percent', w.color = 'psrc_light', w.interactive=TRUE)+
  xlab(as.character(g[cat_name])) + ylab("Transit mode share")+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+ theme(axis.title.x = element_text(size=12,color="#4C4C4C"))+theme(axis.title.y = element_text(size=12,color="#4C4C4C"))+coord_flip()
  print(p1)
  

 # p<-create_facet_bar_chart(t=transit_summs_long , w.x=cat_name, w.y='share', f='mode_simple', g='survey', w.moe='share_moe', est.type='percent', w.color = 'psrc_light')+ theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+coord_flip()

}
```




```{r Updated ggplot2 style and order}
  all_trips_summs_2017_2019 <- hhts_count(transit_data_17_19, 
                                        group_vars=c('simple_purpose', 'mode_simple'), 
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter(mode_simple!='Total')%>%filter(simple_purpose!='Total')%>%filter(mode_simple!='Drive')%>%filter(mode_simple!='Other')%>%filter(mode_simple!='Bike')

  

  
  all_trips_summs_2021 <- hhts_count(transit_data_21, 
                                   group_vars=c('simple_purpose', 'mode_simple'), 
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter(mode_simple!='Total')%>%filter(simple_purpose!='Total')%>%filter(mode_simple!='Drive')%>%filter(mode_simple!='Drive')%>%filter(mode_simple!='Other')%>%filter(mode_simple!='Bike')



  all_trips_summs_long<- rbind(all_trips_summs_2017_2019, all_trips_summs_2021) %>% 
    mutate(survey = str_replace_all(survey, "_", "/"))


  
for (g in unique(all_trips_summs_long$simple_purpose)){  
  
  trips_purp<-all_trips_summs_long%>%filter(simple_purpose==g)
p<-create_bar_chart(t=trips_purp , w.x='mode_simple', w.y='share', f='survey', w.moe='share_moe', est.type='percent', w.color = 'psrc_light')+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+
  ylab(as.character(paste(g, ' Mode Share'))) + xlab("Mode")+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+ theme(axis.title.x = element_text(size=12,color="#4C4C4C"))+theme(axis.title.y = element_text(size=12,color="#4C4C4C"))
  print(p)
  }
```

```{r}

all_trips_summs_2017_2019 <- hhts_median(transit_data_17_19,
                                         stat_var = 'trip_path_distance',
                                        group_vars=c('mode_simple'), 
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%filter(mode_simple!='Other')%>%filter(mode_simple!='Bike')%>%drop_na(mode_simple)

all_trips_summs_2021 <- hhts_median(transit_data_21,
                                         stat_var = 'trip_path_distance',
                                        group_vars=c('mode_simple'), 
                                        spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%filter(mode_simple!='Other')%>%filter(mode_simple!='Bike')%>%drop_na(mode_simple)

  all_trips_summs_long<- rbind(all_trips_summs_2017_2019, all_trips_summs_2021) %>% 
    mutate(survey = str_replace_all(survey, "_", "/"))

trips_purp<-all_trips_summs_long
p<-create_bar_chart(t=all_trips_summs_long , w.x='mode_simple', w.y='trip_path_distance_median', f='survey', w.moe='trip_path_distance_median_moe', est.type='number', w.color = 'psrc_light')+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+
  ylab(as.character('Trip Path Distance')) + xlab("Mode")+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+ theme(axis.title.x = element_text(size=12,color="#4C4C4C"))+theme(axis.title.y = element_text(size=12,color="#4C4C4C"))
  print(p)

```


```{r Updated ggplot2 style and order}
  all_trips_summs_2017_2019 <- hhts_count(transit_data_17_19, 
                                        group_vars=c('simple_purpose'), 
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter(simple_purpose!='Total')

  

  
  all_trips_summs_2021 <- hhts_count(transit_data_21, 
                                   group_vars=c('simple_purpose'), 
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter(simple_purpose!='Total')



  all_trips_summs_long<- rbind(all_trips_summs_2017_2019, all_trips_summs_2021) %>% 
    mutate(survey = str_replace_all(survey, "_", "/"))


  
  
  trips_purp<-all_trips_summs_long
p<-create_bar_chart(t=trips_purp , w.x='simple_purpose', w.y='share', f='survey', w.moe='share_moe', est.type='percent')+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+
  ylab('Share of All Trips') + xlab('Purpose')+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+ theme(axis.title.x = element_text(size=12,color="#4C4C4C"))+theme(axis.title.y = element_text(size=12,color="#4C4C4C"))+coord_flip()
  print(p)
  
```



same analysis but for walk/bike

```{r Updated ggplot2 style and order}
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  # cat_name <- eval(names(g))
  cat_name <- names(g)
  print(cat_name)
  transit_summs_2017_2019 <- hhts_count(transit_data_17_19, 
                                        group_vars=c(names(g),'non_motorized_mode'), 
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter(cat_name!='Total') %>%
    filter(non_motorized_mode=='Walk/Bike')
  

  
  transit_summs_2021 <- hhts_count(transit_data_21, 
                                   group_vars=c(names(g),'non_motorized_mode'), 
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter(cat_name!='Total') %>%
    filter(non_motorized_mode=='Walk/Bike')
  
  if(cat_name=='race_eth_broad'){

     transit_summs_2017_2019 <- transit_summs_2017_2019 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
     
        transit_summs_2021 <- transit_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  transit_summs_2017_2019$race_eth_broad<-factor(transit_summs_2017_2019$race_eth_broad, levels=
                                                  c('White', 'Other', 'Hispanic', 'Black', 'Asian'))
  
   transit_summs_2021$race_eth_broad<-factor(transit_summs_2021$race_eth_broad, levels=
                                                  c('White', 'Other', 'Hispanic', 'Black', 'Asian'))
 

  
  }
  
   if(cat_name=='hhincome_100_f'){
     
     transit_summs_2017_2019$hhincome_100_f=factor( transit_summs_2017_2019$hhincome_100_f, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))

 transit_summs_2021$hhincome_100_f=factor(transit_summs_2021$hhincome_100_f, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))
     
     
   }

  transit_summs_long<- rbind(transit_summs_2017_2019, transit_summs_2021) %>% 
    mutate(survey = str_replace_all(survey, "_", "/"))

  
   # factorize variable column. Adjust levels in function definition
  #transit_summs_long <- factor_variable_column(names(g), transit_summs_long)
  
  transit_summs_2021_only<-  transit_summs_2021%>%filter(non_motorized_mode=='Walk/Bike')
  
 # p0<-create_bar_chart(t=transit_summs_2021_only , w.x=cat_name, w.y='share',  f=cat_name,w.moe='share_moe', est.type='percent', w.color='psrc_light')+ t#heme(axis.text.x = element_text(size=10,color="#4C4C4C"))+coord_flip()
#  print(p0)
  
   p0<-create_bar_chart(t=transit_summs_2021 , w.x=cat_name, w.y='share', f='non_motorized_mode', w.moe='share_moe', est.type='percent', w.color = 'psrc_light', w.interactive=TRUE)+
  xlab(as.character(g[cat_name])) + ylab('Walk/Bike mode share')+scale_fill_manual(values="#73CFCB")+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+ theme(axis.title.x = element_text(size=12,color="#4C4C4C"))+theme(axis.title.y = element_text(size=12,color="#4C4C4C"))+coord_flip()
  print(p0)
  
  p1<-create_bar_chart(t=transit_summs_long , w.x=cat_name, w.y='share', f='survey', w.moe='share_moe', est.type='percent', w.color = 'psrc_light', w.interactive=TRUE)+
  xlab(as.character(g[cat_name])) + ylab("Walk/Bike mode share")+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+ theme(axis.title.x = element_text(size=12,color="#4C4C4C"))+theme(axis.title.y = element_text(size=12,color="#4C4C4C"))+coord_flip()
  print(p1)
  

 # p<-create_facet_bar_chart(t=transit_summs_long , w.x=cat_name, w.y='share', f='non_motorized_mode', g='survey', w.moe='share_moe', est.type='percent', w.color = 'psrc_light')+ theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+coord_flip()

}
```




```{r Updated ggplot2 style and order}
  all_trips_summs_2017_2019 <- hhts_count(transit_data_17_19, 
                                        group_vars=c('simple_purpose', 'non_motorized_mode'), 
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter(non_motorized_mode!='Total')%>%filter(simple_purpose!='Total')%>%filter(non_motorized_mode!='Drive')%>%filter(non_motorized_mode!='Other')%>%filter(non_motorized_mode!='Bike')

  

  
  all_trips_summs_2021 <- hhts_count(transit_data_21, 
                                   group_vars=c('simple_purpose', 'non_motorized_mode'), 
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter(non_motorized_mode!='Total')%>%filter(simple_purpose!='Total')%>%filter(non_motorized_mode!='Drive')%>%filter(non_motorized_mode!='Drive')%>%filter(non_motorized_mode!='Other')%>%filter(non_motorized_mode!='Bike')



  all_trips_summs_long<- rbind(all_trips_summs_2017_2019, all_trips_summs_2021) %>% 
    mutate(survey = str_replace_all(survey, "_", "/"))


  
for (g in unique(all_trips_summs_long$simple_purpose)){  
  
  trips_purp<-all_trips_summs_long%>%filter(simple_purpose==g)
p<-create_bar_chart(t=trips_purp , w.x='non_motorized_mode', w.y='share', f='survey', w.moe='share_moe', est.type='percent', w.color = 'psrc_light')+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+
  ylab(as.character(paste(g, ' Mode Share'))) + xlab("Mode")+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+ theme(axis.title.x = element_text(size=12,color="#4C4C4C"))+theme(axis.title.y = element_text(size=12,color="#4C4C4C"))
  print(p)
  }
```

```{r}

all_trips_summs_2017_2019 <- hhts_median(transit_data_17_19,
                                         stat_var = 'trip_path_distance',
                                        group_vars=c('non_motorized_mode'), 
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%filter(non_motorized_mode!='Other')%>%filter(non_motorized_mode!='Bike')%>%drop_na(non_motorized_mode)

all_trips_summs_2021 <- hhts_median(transit_data_21,
                                         stat_var = 'trip_path_distance',
                                        group_vars=c('non_motorized_mode'), 
                                        spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%filter(non_motorized_mode!='Other')%>%filter(non_motorized_mode!='Bike')%>%drop_na(non_motorized_mode)

  all_trips_summs_long<- rbind(all_trips_summs_2017_2019, all_trips_summs_2021) %>% 
    mutate(survey = str_replace_all(survey, "_", "/"))

trips_purp<-all_trips_summs_long
p<-create_bar_chart(t=all_trips_summs_long , w.x='non_motorized_mode', w.y='trip_path_distance_median', f='survey', w.moe='trip_path_distance_median_moe', est.type='number', w.color = 'psrc_light')+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+
  ylab(as.character('Trip Path Distance')) + xlab("Mode")+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+ theme(axis.title.x = element_text(size=12,color="#4C4C4C"))+theme(axis.title.y = element_text(size=12,color="#4C4C4C"))
  print(p)

```


```{r Updated ggplot2 style and order}
  all_trips_summs_2017_2019 <- hhts_count(transit_data_17_19, 
                                        group_vars=c('simple_purpose'), 
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter(simple_purpose!='Total')

  

  
  all_trips_summs_2021 <- hhts_count(transit_data_21, 
                                   group_vars=c('simple_purpose'), 
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter(simple_purpose!='Total')



  all_trips_summs_long<- rbind(all_trips_summs_2017_2019, all_trips_summs_2021) %>% 
    mutate(survey = str_replace_all(survey, "_", "/"))


  
  
  trips_purp<-all_trips_summs_long
p<-create_bar_chart(t=trips_purp , w.x='simple_purpose', w.y='share', f='survey', w.moe='share_moe', est.type='percent')+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+
  ylab('Share of All Trips') + xlab('Purpose')+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+ theme(axis.title.x = element_text(size=12,color="#4C4C4C"))+theme(axis.title.y = element_text(size=12,color="#4C4C4C"))+coord_flip()
  print(p)
  
```
walk/bike frequency
```{r}
mode_vars<-c('mode_freq_3', 'mode_freq_2')
other_vars<-c('final_home_rgcnum', 'hhsize', 'vehicle_count',  "hhincome_broad", 'rent_own', 'res_dur', 'student', 'education',  'hhincome_detailed', "age", "age_category", 'race_category', 'race_eth_broad', 'gender', 'employment',  'lifecycle', 'final_home_is_rgc', 'seattle_home', 'final_cnty')
all_vars<-c(mode_vars, other_vars)
```



 
 
```{r}
walk_bike_freq_data_17_19<- get_hhts("2017_2019", "p", vars=all_vars)%>% mutate(year=ifelse(survey=='2017_2019', '2017/2019', '2021'))
```


```{r}
walk_bike_freq_data_21<- get_hhts("2021", "p", vars=all_vars)%>% mutate(year=ifelse(survey=='2017_2019', '2017/2019', '2021'))

```


```{r Updated ggplot2 style and order}
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  # cat_name <- eval(names(g))
  cat_name <- names(g)
  print(cat_name)
  # does this include kids, i suppose we'll see with the summary itself
  wbfreq_summs_2017_2019 <- hhts_count(walk_bike_freq_data_17_19, 
                                        group_vars=c(names(g),'mode_freq_2'), 
                                        spec_wgt='hh_weight_2017_2019_v2021')%>%
    filter(cat_name!='Total')

  
  wbfreq_summs_2021 <- hhts_count(walk_bike_freq_data_21, 
                                   group_vars=c(names(g),'mode_freq_2'), 
                                   spec_wgt='person_weight_2021_ABS_Panel_adult')%>%
    filter(cat_name!='Total') %>%

  
  if(cat_name=='race_eth_broad'){
print('here)
     wbfreq_summs_2017_2019 <- wbfreq_summs_2017_2019  %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
     
        wbfreq_summs_2021 <- wbfreq_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  wbfreq_summs_2017_2019$race_eth_broad<-factor(wbfreq_summs_2017_2019$race_eth_broad, levels=
                                                  c('White', 'Other', 'Hispanic', 'Black', 'Asian'))
  
   wbfreq_summs_2021$race_eth_broad<-factor(wbfreq_summs_2021$race_eth_broad, levels=
                                                  c('White', 'Other', 'Hispanic', 'Black', 'Asian'))
 

  
  }
  
if(cat_name=='hhincome_100_f'){
     
     wbfreq_summs_2017_2019$hhincome_100_f=factor( wbfreq_summs_2017_2019$hhincome_100_f, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))

 wbfreq_summs_2021$hhincome_100_f=factor(wbfreq_summs_2021$hhincome_100_f, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))
     
     
   }

  wbfreq_summs_long<- rbind(wbfreq_summs_2017_2019, wbfreq_summs_2021) %>% 
    mutate(survey = str_replace_all(survey, "_", "/"))

  
   # factorize variable column. Adjust levels in function definition
  #transit_summs_long <- factor_variable_column(names(g), transit_summs_long)
  

  
   p0<-create_bar_chart(t=wbfreq_summs_2021 , w.x=cat_name, w.y='share', f='non_motorized_mode', w.moe='share_moe', est.type='percent', w.color = 'psrc_light', w.interactive=TRUE)+
  xlab(as.character(g[cat_name])) + ylab('Walk/Bike mode share')+scale_fill_manual(values="#73CFCB")+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+ theme(axis.title.x = element_text(size=12,color="#4C4C4C"))+theme(axis.title.y = element_text(size=12,color="#4C4C4C"))+coord_flip()
  print(p0)
  
  p1<-create_bar_chart(t=wbfreq_summs_long , w.x=cat_name, w.y='share', f='survey', w.moe='share_moe', est.type='percent', w.color = 'psrc_light', w.interactive=TRUE)+
  xlab(as.character(g[cat_name])) + ylab("Walk/Bike mode share")+theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+ theme(axis.title.x = element_text(size=12,color="#4C4C4C"))+theme(axis.title.y = element_text(size=12,color="#4C4C4C"))+coord_flip()
  print(p1)
  

 # p<-create_facet_bar_chart(t=transit_summs_long , w.x=cat_name, w.y='share', f='non_motorized_mode', g='survey', w.moe='share_moe', est.type='percent', w.color = 'psrc_light')+ theme(axis.text.x = element_text(size=10,color="#4C4C4C"))+coord_flip()

}
```
