---
title: "commute_subsidy"
author: "suzanne"
date: "2024-08-23"
output: html_document
---

```{r}
workers<-person%>%filter(!workplace %in% c('Missing: Skip logic', "Missing: Skip Logic"))%>%drop_na(workplace)
hh_23<-hh%>%filter(survey_year=='2023')
person_23<-workers%>%filter(survey_year=='2023')
worker_list<-list(hh=hh, person=workers)

```

```{r}
variable_list<-variable_list%>%mutate(is_checkbox=0, shared_name=variable)
workers_respond<-workers%>%filter(commute_subsidy_1 %in% c('Selected', 'Not selected'))%>%filter(!is.na(commute_subsidy_1))
hh<-hh%>%mutate(hhincome_detailed=factor(hhincome_detailed, levels= c( 'Under $10,000' ,'$10,000-$24,999','$25,000-$34,999','$35,000-$49,999','$50,000-$74,999','$75,000-$99,999', '$100,000-$149,999','$150,000-$199,999', '$200,000-$249,999', '$250,000 or more' )))
worker_list<-list(hh=hh, person=workers_respond)

test<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'commute_subsidy_1',
                               summarize_by = 'hhincome_broad',
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight')
```


```{r}
transit_pass<-test$summary$wtd%>%filter(commute_subsidy_1=='Selected')%>%filter(hhincome_broad!='Prefer not to answer')
transit_pass
```
```{r}
static<-static_bar_chart(transit_pass, x='prop', fill='hhincome_broad', y='hhincome_broad', dec=0, color='pgnobgy_10') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
static
```

```{r}
static<-static_bar_chart(parking_pass, x='prop', fill='hhincome_broad', y='hhincome_broad', dec=0, color='pgnobgy_10') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
static
```

``````{r}
variable_list<-variable_list%>%mutate(is_checkbox=0, shared_name=variable)
workers_respond<-workers%>%filter(commute_subsidy_3 %in% c('Selected', 'Not selected'))%>%filter(!is.na(commute_subsidy_1))
hh<-hh%>%mutate(hhincome_detailed=factor(hhincome_detailed, levels= c( 'Under $10,000' ,'$10,000-$24,999','$25,000-$34,999','$35,000-$49,999','$50,000-$74,999','$75,000-$99,999', '$100,000-$149,999','$150,000-$199,999', '$200,000-$249,999', '$250,000 or more' )))
worker_list<-list(hh=hh, person=workers_respond)

test<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'commute_subsidy_3',
                               summarize_by = 'hhincome_broad',
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight')
parking<-test$summary$wtd%>%filter(commute_subsidy_3=='Selected')
parking

```
```{r}
parking_pass<-parking%>%filter(hhincome_broad!='Prefer not to answer')
parking_pass
```

```{r}
static<-static_bar_chart(parking_pass, x='prop', fill='hhincome_broad', y='hhincome_broad', dec=0, color='pgnobgy_10') + theme(axis.text.x=element_text(size=16), axis.text.y=element_text(size=16),legend.text = element_text(size=16), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
static
```
```{r}
test<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'commute_subsidy_1',
                               summarize_by = 'race_category_2023_c',
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight')
```
```{r}
test<- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'commute_subsidy_1',
                               summarize_by = 'commute_subsidy_3',
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight', 'person_weight'),
                               wtname= 'person_weight')
```
```{r}
test$summary$wtd
```

