---
title: "KC-Metro-Examples"
author: "suzanne"
date: "2023-04-11"
output: html_document
---
We use a few packages to make it smoother to work with complex travel survey data and summarize it.
Here's instructions on install the two packages:
https://psrc.github.io/psrc.travelsurvey/
https://psrc.github.io/psrcplot/

Put the sqlite database called hhsurvey.db in the directory where this code resides.

```{r}
# read in libraries
library(psrc.travelsurvey)
library(psrcplot)
library(ggplot2)
library(RSQLite)
library(config)
library(dplyr)

```

Set your configuration to use sqlite connection properties. This way the psrc.travelsurvey package knows to access the data from the sqlite database, instead of an internal SQL Server database.
```{r}
#Sys.setenv(R_CONFIG_ACTIVE = 'shinyapps')
#Sys.setenv(R_CONFIG_ACTIVE='default')

```


Telecommuting rates for people working in King County
2017/2019


```{r}

work_transitpass_person_df_17_19<-get_hhts(survey = '2017_2019',
                    level = "p",
                    vars = c("age_category",
                             "work_county",
                             "workplace",
                             "gender",
                             "hhincome_broad",
                             "industry", 
                             "final_cnty",
                              "benefits_3"))%>%filter(work_county=='King County')



```



Telecommuting rates for people working in King County
2021
```{r}

work_transitpass_person_df_21<-get_hhts(survey = '2021',
                    level = "p",
                    vars = c("age_category",
                             "work_county",
                             "workplace",
                             "gender",
                             "hhincome_broad",
                             "industry",
                             "benefits_3",
                     'final_cnty'))%>%filter(work_county=='KING')



```



```{r}
work_transitpass_sum_17_19<-hhts_count(df=work_transitpass_person_df_17_19, group_vars=c('benefits_3'), incl_na=FALSE)%>%filter(benefits_3!='Total')
work_transitpass_sum_21<-hhts_count(df=work_transitpass_person_df_21, group_vars=c('benefits_3'), incl_na=FALSE)%>%filter(benefits_3!='Total')
```                                  
                                  
                                  
```{r}
work_transitpass_trend<- rbind(work_transitpass_sum_17_19, work_transitpass_sum_21)
```

```{r}

work_transitpass_trend_chart<-static_column_chart(t=work_transitpass_trend, x='benefits_3', y='share', fill='survey', moe='share_moe')
work_transitpass_trend_chart
```

