---
title: "EV Charging Analysis"
format:
  html:
    code-overflow: wrap  
editor: visual
toc: true
toc-location: right
---

```{r, include=FALSE}
source("ev_analysis2024.R")
```

## EV Charging Location

This shows ev charging location based on a hierarchial approach. If a hh had access to charging at work, they are in this category If a hh had access to public/other, they are in this category The rest of the hh only charged at home

```{r}
#| echo: false
#| warning: false


kbl(ev_charge) %>% 
   kable_styling(bootstrap_options = c("striped", "hover"), html_font = "Poppins") %>%
  column_spec(c(3), background = psrc_colors$purples_inc[1]) %>%
  column_spec(c(4), background = psrc_colors$greens_inc[1]) %>% 
  column_spec(c(5), background = psrc_colors$purples_inc[1]) %>% 
  column_spec(c(6), background = psrc_colors$greens_inc[1])

static_column_chart(ev_charge, x='top_ev_loc', y='prop', fill='top_ev_loc', moe="prop_moe") +
  scale_y_continuous(labels=scales::percent) +
  labs(title = "EV Typical Charging Locations", caption = "data: Household Travel Survey 2023", x="EV Typical Charge", y="Share of Households") +
  theme(axis.text=element_text(size=20), axis.title=element_text(size=20), plot.title=element_text(size=20), plot.caption = element_text(size=20))

```

This method I found the ratio of hh who charged their ev at a location compared to the rest of the hh who did not charge at that location

This is saying 96% of hh have access to at home charging

14% of hh have access to charging at work

28% of hh have access to charging at a public or other location

```{r}
#| echo: false
#| warning: false


kbl(ev_charge1) %>% 
   kable_styling(bootstrap_options = c("striped", "hover"), html_font = "Poppins") %>%
  column_spec(c(3), background = psrc_colors$purples_inc[1]) %>%
  column_spec(c(4), background = psrc_colors$greens_inc[1]) %>% 
  column_spec(c(5), background = psrc_colors$purples_inc[1]) %>% 
  column_spec(c(6), background = psrc_colors$greens_inc[1])

static_column_chart(ev_charge1, x='ev_loc', y='prop', fill='ev_loc', moe="prop_moe") +
  scale_y_continuous(labels=scales::percent) +
  labs(title = "EV Typical Charging Locations", caption = "data: Household Travel Survey 2023", x="Charging Location", y="Share of Households") +
  theme(axis.text=element_text(size=20), axis.title=element_text(size=20), plot.title=element_text(size=20), plot.caption = element_text(size=20))

```

## Home vs. Not Home

```{r}
#| warning: false
#| echo: false
static_column_chart(home_comparison, x='home_comp', y='prop', fill='home_comp', moe='moe') +
  scale_y_continuous(labels=scales::percent) +
    labs(title = "Households with Home Access to EV Charging", caption = "data: Household Travel Survey 2023", x="EV Charging Location", y="Percentage of Households") +
    theme(axis.text=element_text(size=20), axis.title=element_text(size=20), plot.title=element_text(size=20), plot.caption = element_text(size=20))

# Just at Home
static_column_chart(at_home, x='home_comp', y='prop', fill='home_comp', moe='moe') +
  scale_y_continuous(labels=scales::percent) +
    labs(title = "Households with Home Access to EV Charging", caption = "data: Household Travel Survey 2023", x="EV Charging Location", y="Percentage of Households") +
    theme(axis.text=element_text(size=20), axis.title=element_text(size=20), plot.title=element_text(size=20), plot.caption = element_text(size=20))
```

## Vehicle Count

This summary shows the counts for the number of people who have a car and those that do not have a car

```{r}
#| echo: false
#| warning: false


veh_count <- veh_count_summary$wtd %>% mutate(moe = prop_se * 1.645)
veh_count$prop <- round(veh_count$prop, 4)
veh_count$est <- round(veh_count$est, 2)

kbl(veh_count) %>% 
   kable_styling(bootstrap_options = c("striped", "hover"), html_font = "Poppins") %>%
  column_spec(c(2), background = psrc_colors$purples_inc[1]) %>%
  column_spec(c(3), background = psrc_colors$greens_inc[1]) %>% 
  column_spec(c(4), background = psrc_colors$purples_inc[1]) %>% 
  column_spec(c(5), background = psrc_colors$greens_inc[1]) %>% 
  column_spec(c(6), background = psrc_colors$purples_inc[1]) %>% 
  column_spec(c(7), background = psrc_colors$greens_inc[1])

static_column_chart(veh_count, x='veh_count', y='prop', fill='veh_count', moe="moe") +
    scale_y_continuous(labels=scales::percent) +
    labs(title = "Number of Vehicles per a Household", caption = "data: Household Travel Survey 2023", x="Number of Vehicles", y="Share of Households") +
    theme(axis.text=element_text(size=20), axis.title=element_text(size=20), plot.title=element_text(size=20), plot.caption = element_text(size=20))
```

## Fuel Type

This summary displays the proportion and counts for each fuel type

```{r}
#| echo: false
#| warning: false

fuel_cat <- fuel_summary_df1
fuel_cat$prop <- round(fuel_cat$prop, 4)
fuel_cat$est <- round(fuel_cat$est, 2)
kbl(fuel_cat) %>% 
   kable_styling(bootstrap_options = c("striped", "hover"), html_font = "Poppins") %>%
  column_spec(c(2), background = psrc_colors$purples_inc[1]) %>%
  column_spec(c(3), background = psrc_colors$greens_inc[1]) %>% 
  column_spec(c(4), background = psrc_colors$purples_inc[1]) %>% 
  column_spec(c(5), background = psrc_colors$greens_inc[1]) %>% 
  column_spec(c(6), background = psrc_colors$purples_inc[1]) %>% 
  column_spec(c(7), background = psrc_colors$greens_inc[1])


# chart by fuel type
static_column_chart(fuel_summary_df1, x='fuel_type', y='prop', fill='fuel_type', moe="prop_moe") +
    scale_y_continuous(labels=scales::percent) +
    labs(title = "Share of vehicle fuel types by Household", caption = "data: Household Travel Survey 2023", x="Vehicle Fuel Type", y="Share of Households") +
    theme(axis.text=element_text(size=20), axis.title=element_text(size=20), plot.title=element_text(size=20), plot.caption = element_text(size=20))


```

Fuel Type By Household

```{r}
#| warning: false
#| echo: false

# This summary table displays for each household, the number of vehicles it has for each fuel type and the household's total vehicles
# step 1: creates a new column and adds a '1' if it satisfies this condition
# step 2: groups by household
# step 3: for each group of hh rows, it sums the column values for each vehicle type

veh_type_count <- vehicle %>% 
  mutate(
  total_ev = case_when(fuel == 'Electric (EV)' ~ 1, TRUE ~ 0),
  total_gas = case_when(fuel == 'Gas' ~ 1, TRUE ~ 0),
  total_diesel = case_when(fuel == 'Diesel' ~ 1, TRUE ~ 0),
  total_other = case_when(fuel == 'Other (e.g., natural gas, bio-diesel, Flex fuel (FFV))' ~ 1, TRUE ~ 0),
  total_hev = case_when(fuel == 'Hybrid (HEV)' ~ 1, TRUE ~ 0),
  total_phev = case_when(fuel == 'Plug-in hybrid (PHEV)' ~ 1, TRUE ~ 0)) %>% 
  group_by(hh_id) %>%
  summarise(across(c(total_ev, total_gas, total_diesel, total_other, total_hev, total_phev), sum)) %>% 
  mutate(total_vehicles = rowSums(.[2:7])) %>% 
  ungroup() %>% 
  full_join(hh, by="hh_id")

vtc_short <- head(veh_type_count)

kbl(vtc_short) %>% 
   kable_styling(bootstrap_options = c("striped", "hover"), html_font = "Poppins")
# used to verify the total number of vehicles for each hh is the same as in the vehicles table  
test <- vehicle %>% group_by(hh_id) %>% summarise(count = n())

```

Verifying if multiple people within a hh answered the 'EV typical charge' location question

```{r}
#| echo: false
#| warning: false
hh_response_count <- person %>% 
  group_by(hh_id) %>% 
  summarise(num_responses = n(),
            num_people = n_distinct(person_id))

sum(hh_response_count$num_people)
# length(unique(person_new$hh_id))

hhrc_short <- head(hh_response_count)

kbl(hhrc_short) %>% 
   kable_styling(bootstrap_options = c("striped", "hover"), html_font = "Poppins")
```

## Further Exploration Questions...

1.  hh income of ev households?

2.  do ev owners live in rgcs?

3.  what is their res_type?
