---
title: "Transportation Operator Committee HHTS 2021 notes"
output:
  html_document:
    df_print: paged
---

https://www.psrc.org/committee/transportation-operators-committee

July 27 meeting

We are gathering materials information that could be useful to the transportation operators committee


```{r }
library(psrc.travelsurvey)
library(dplyr)
library(stringr)
library(ggplot2)
library(psrcplot)
library(forcats)
library(odbc)
library(DBI)
library(tidyr)
library(tidycensus)
library(psrcslides)
library(officer)
library(DBI)
install_psrc_fonts()
```


```{r}
mode_vars<-c('mode_1', 'mode_simple')
other_vars<-c('sample_county', 'vehicle_count',  "hhincome_broad",  'hhincome_detailed', "age", "age_category", 'final_home_is_rgc', 'race_category', 'race_eth_broad', 'dest_purpose_cat', 'origin_purpose_cat', 'trip_path_distance', 'person_id')
all_vars<-c(mode_vars, other_vars)
person_vars=c('vehicle_count',  "hhincome_broad",  'hhincome_detailed', "age", "age_category", 'final_home_is_rgc', 'race_category', 'race_eth_broad', 'person_id')
```

 
```{r}
transit_data_17_19<- get_hhts("2017_2019", "t", vars=all_vars)
```


```{r}
transit_data_21<- get_hhts("2021", "t", vars=all_vars)
```

```{r}
person_data_17_19<- get_hhts("2017_2019", "p", vars=person_vars)
```


```{r}
person_data_21<- get_hhts("2021", "p", vars=person_vars)
```


```{r}
 elmer_connect<-DBI::dbConnect(odbc::odbc(),
                 driver = "ODBC Driver 17 for SQL Server",
                 server = "AWS-PROD-SQL\\Sockeye",
                 database = "Elmer",
                 trusted_connection = "yes",
                 port = 1433)
  


```

```{r}
disab_sql<-'with hhlds_disability as 
(
    select v.household_id, survey_year, count(*) as disability_vehicle_count
    from HHSurvey.v_vehicles v
    where v.disability = \'Yes\'
    group by v.household_id, v.survey_year
)
select hhlds_disability.household_id, p.person_id,isnull(hhlds_disability.disability_vehicle_count, 0) as disability_vehicle_count
from HHSurvey.v_persons as p
left join hhlds_disability on p.household_id = hhlds_disability.household_id and p.survey_year = hhlds_disability.survey_year'

```

```{r}
people_disab_vehs <- DBI::dbGetQuery(elmer_connect, DBI::SQL(disab_sql))

```

```{r}
person_data_17_19<-merge(person_data_17_19, people_disab_vehs, by='person_id')
person_data_21<-merge(person_data_21, people_disab_vehs, by = 'person_id')
```

```{r}
transit_data_17_19<-merge(transit_data_17_19, people_disab_vehs, by='person_id')
transit_data_21<-merge(transit_data_21, people_disab_vehs, by = 'person_id')
```

```{r}

psrc_pres = read_pptx(system.file('extdata', 'psrc-trends-template.pptx', package='psrcslides'))

```
 
 
 Start simple - What was the total estimated transit trips in 2017/2019 vs 2021?
 
 These are trips made by adults only
 
```{r, include=TRUE}
trips_by_mode_17_19<-hhts_count(transit_data_17_19, group_vars='mode_simple')%>%
  filter(mode_simple!='Total')
trips_by_mode_21<-hhts_count(transit_data_21, group_vars='mode_simple')%>%filter(mode_simple!='Total')

trips_by_mode_17_19_21<-merge(trips_by_mode_17_19, trips_by_mode_21, by='mode_simple', suffixes=c('17_19', '21'))
trips_by_mode<-rbind(trips_by_mode_17_19, trips_by_mode_21)

```
```{r}
write.csv(trips_by_mode, 'trips_by_mode.csv')
```

```{r, include=TRUE}
p<-ggplot(trips_by_mode, aes(x=mode_simple, y=share, fill=survey))+geom_bar(stat="identity", position=position_dodge())+xlab('Trip Mode')+ggtitle('Trip Mode Share', 'by Survey Year')+scale_fill_discrete_psrc('psrc_light')+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe), width=.2,
                 position=position_dodge(.9)) +scale_y_continuous(labels=scales::percent)

print(p)
```
From 2017/2019 to 2021, the transit share went down, drive shares went down, and walking shares went up.

```{r, include=TRUE}
trips_by_mode_no_drive<-trips_by_mode%>%filter(!mode_simple %in% c('Drive', 'Bike', 'Other'))
p<-ggplot(trips_by_mode_no_drive, aes(x=mode_simple, y=share, fill=survey))+geom_bar(stat="identity", position=position_dodge())+xlab('Trip Mode')+ggtitle('Trip Mode Share', 'by Survey Year')+scale_fill_discrete_psrc('psrc_light')+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe), width=.2,
                 position=position_dodge(.9))  +scale_y_continuous(labels=scales::percent)

print(p)
```





 


Mode Share by different characteristics

findings: people with zero cars kept using transit more
people in smaller households?
lower income


first doing detailed groupings, to figure out how to group the data


Group variables into meaningful categories and redo analysis
```{r}
transit_data_17_19<-transit_data_17_19%>%mutate(NoVehicles=ifelse(vehicle_count=='0 (no vehicles)', 'No Vehicles', "Has Vehicles")) %>%mutate(hhincome_100= case_when(hhincome_broad=='$100,000-$199,000' ~ '$100,000 or more',
                                 hhincome_broad=='$200,000 or more' ~ '$100,000 or more',
                                 TRUE~hhincome_broad))%>%
  mutate(age_grp= case_when(age=='75-84 years' ~ '75 years or older', 
                             age == '85 or years older' ~ '75 years or older',
                            TRUE ~ age))%>%mutate(simple_purpose=ifelse(dest_purpose_cat=='Home', origin_purpose_cat, dest_purpose_cat))%>%
  mutate(simple_purpose=case_when(simple_purpose=='Work'~ 'Work/School',
                                  simple_purpose=='School'~ 'Work/School',
                                  simple_purpose=='Work-related'~ 'Work/School',
                                  simple_purpose=='Shop'~ 'Shop',
                                  simple_purpose=='Escort'~ 'Errands',
                                  simple_purpose=='Errand/Other'~ 'Errands',
                                  simple_purpose=='Change mode'~ 'Errands',
                                  simple_purpose=='Social/Recreation' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Meal' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Meal' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Home' ~'Errands',
                                  is.na(simple_purpose) ~ 'Errands',
                                  TRUE ~ simple_purpose))%>%
  mutate(has_disab_plates=ifelse(disability_vehicle_count>0, 'Has Disability Vehicle Plates', 'No Disability Vehicle Plates'))
 

transit_data_21<-transit_data_21%>%mutate(NoVehicles=ifelse(vehicle_count=='0 (no vehicles)', 'No Vehicles', "Has Vehicles")) %>%
  mutate(hhincome_100= case_when(hhincome_broad=='$100,000-$199,000' ~ '$100,000 or more',
                                 hhincome_broad=='$200,000 or more' ~ '$100,000 or more',
                                 TRUE~hhincome_broad))%>%
  mutate(age_grp= case_when(age=='75-84 years' ~ '75 years or older', 
                            age == '85 or years older' ~ '75 years or older',
                            TRUE ~ age))%>% mutate(simple_purpose=ifelse(dest_purpose_cat=='Home', origin_purpose_cat, dest_purpose_cat))%>%
  mutate(simple_purpose=case_when(simple_purpose=='Work'~ 'Work/School',
                                  simple_purpose=='School'~ 'Work/School',
                                  simple_purpose=='Work-related'~ 'Work/School',
                                  simple_purpose=='Shop'~ 'Shop',
                                  simple_purpose=='Escort'~ 'Errands',
                                  simple_purpose=='Errand/Other'~ 'Errands',
                                  simple_purpose=='Change mode'~ 'Errands',
                                  simple_purpose=='Social/Recreation' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Meal' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Meal' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Home' ~'Errands',
                                  is.na(simple_purpose) ~ 'Errands',
                                  TRUE ~ simple_purpose))%>%
  mutate(has_disab_plates=ifelse(disability_vehicle_count>0, 'Has Disability Vehicle Plates', 'No Disability Vehicle Plates'))
 

transit_data_17_19$hhincome_100_f=factor(transit_data_17_19$hhincome_100,levels=c("Under $25,000","$25,000-$49,999",   "$50,000-$74,999","$75,000-$99,999","$100,000 or more", "Prefer not to answer"))

transit_data_21$hhincome_100_f=factor(transit_data_21$hhincome_100,levels=c("Under $25,000","$25,000-$49,999",   "$50,000-$74,999","$75,000-$99,999","$100,000 or more", "Prefer not to answer"))


transit_data_17_19$NoVehicles=factor(transit_data_17_19$NoVehicles, levels= c('No Vehicles', 'Has Vehicles'))

transit_data_21$NoVehicles=factor(transit_data_21$NoVehicles, levels=c('No Vehicles', 'Has Vehicles'))


simple_groupings<-c('race_eth_broad'='Race/Ethnicity','hhincome_100_f'='Household Income','NoVehicles'= 'Household Vehicles', 'age_grp'= 'Age Group', 'sample_county' ='Home County', 'age_category'='Age Category', 'final_home_is_rgc'='Home in Regional Growth Center', has_disab_plates='Has Disability Vehicle Plates' )
```
Trip Mode Shares by Demographic
```{r }
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  # cat_name <- eval(names(g))
  cat_name <- names(g)
  print(cat_name)
  transit_summs_2017_2019 <- hhts_count(transit_data_17_19,
                                        group_vars=c(names(g),'mode_simple'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter(cat_name!='Total') %>%filter(mode_simple!='Total')%>%drop_na(cat_name)%>%mutate(cat_name=str_wrap(cat_name, width=10))



  transit_summs_2021 <- hhts_count(transit_data_21,
                                   group_vars=c(names(g),'mode_simple'),
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter(cat_name!='Total') %>%filter(mode_simple!='Total')%>%drop_na(cat_name)%>%mutate(cat_name=str_wrap(cat_name, width=10))

  if(cat_name=='race_eth_broad'){

     transit_summs_2017_2019 <- transit_summs_2017_2019 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')

        transit_summs_2021 <- transit_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  #transit_summs_2017_2019$race_eth_broad<-factor(transit_summs_2017_2019$race_eth_broad, levels=
                                                  #c('White', 'Other', 'Hispanic', 'Black', 'Asian'))

   #transit_summs_2021$race_eth_broad<-factor(transit_summs_2021$race_eth_broad, levels=
                                                  #c('White', 'Other', 'Hispanic', 'Black', 'Asian'))



  }

 #  if(cat_name=='hhincome_100_f'){

  #   transit_summs_2017_2019$hhincome_100_f=factor( transit_summs_2017_2019$hhincome_100_f, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))

 #transit_summs_2021$hhincome_100_f=factor(transit_summs_2021$hhincome_100_f, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))


#   }

  transit_summs_long<- rbind(transit_summs_2017_2019, transit_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))


   # factorize variable column. Adjust levels in function definition
  #transit_summs_long <- factor_variable_column(names(g), transit_summs_long)

  #transit_summs_2021_only<-  transit_summs_2021%>%filter(non_motorized_mode=='Walk/Bike')

 # p0<-create_bar_chart(t=transit_summs_2021_only , x=cat_name, y='share',  f=cat_name,moe='share_moe', est='percent', #color='psrc_light')+ t#heme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()


   p0<-create_column_chart(t=transit_summs_2021 , x='mode_simple', y='share', f=cat_name, moe='share_moe', est='percent', color = 'psrc_light', interactive=FALSE)  +theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)


  p1<-create_facet_bar_chart(t=transit_summs_long , f='mode_simple',  w.x=cat_name, w.y='share', g='survey', w.moe='share_moe', est.type='percent', w.color = 'psrc_light') +theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)

}
 # p<-create_facet_bar_chart(t=transit_summs_long , x=cat_name, y='share', f='non_motorized_mode', g='survey', moe='share_moe', est='percent', color = 'psrc_light')+ theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()



```

# Walk Trip Mode Shares by Demographic
```{r }
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  # cat_name <- eval(names(g))
  cat_name <- names(g)
  print(cat_name)
  transit_summs_2017_2019 <- hhts_count(transit_data_17_19,
                                        group_vars=c(names(g),'mode_simple'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter(cat_name!='Total') %>%
    filter(mode_simple=='Walk')%>%drop_na(cat_name)%>%mutate(cat_name=str_wrap(cat_name, width=10))



  transit_summs_2021 <- hhts_count(transit_data_21,
                                   group_vars=c(names(g),'mode_simple'),
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter(cat_name!='Total') %>%
    filter(mode_simple=='Walk')%>%drop_na(cat_name)%>%mutate(cat_name=str_wrap(cat_name, width=10))

  if(cat_name=='race_eth_broad'){

     transit_summs_2017_2019 <- transit_summs_2017_2019 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')

        transit_summs_2021 <- transit_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  #transit_summs_2017_2019$race_eth_broad<-factor(transit_summs_2017_2019$race_eth_broad, levels=
                                                  #c('White', 'Other', 'Hispanic', 'Black', 'Asian'))

   #transit_summs_2021$race_eth_broad<-factor(transit_summs_2021$race_eth_broad, levels=
                                                  #c('White', 'Other', 'Hispanic', 'Black', 'Asian'))



  }

 #  if(cat_name=='hhincome_100_f'){

  #   transit_summs_2017_2019$hhincome_100_f=factor( transit_summs_2017_2019$hhincome_100_f, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))

 #transit_summs_2021$hhincome_100_f=factor(transit_summs_2021$hhincome_100_f, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))


#   }

  transit_summs_long<- rbind(transit_summs_2017_2019, transit_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))


   # factorize variable column. Adjust levels in function definition
  #transit_summs_long <- factor_variable_column(names(g), transit_summs_long)

  #transit_summs_2021_only<-  transit_summs_2021%>%filter(non_motorized_mode=='Walk/Bike')

 # p0<-create_bar_chart(t=transit_summs_2021_only , x=cat_name, y='share',  f=cat_name,moe='share_moe', est='percent', #color='psrc_light')+ t#heme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()


   p0<-create_column_chart(t=transit_summs_2021 , x=cat_name, y='share', f='mode_simple', moe='share_moe', est='percent', color = 'psrc_light', interactive=FALSE) + 
  xlab(as.character(g[cat_name])) + ylab('Walk mode share')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  


  p1<-create_column_chart(t=transit_summs_long , x=cat_name, y='share', f='survey', moe='share_moe', est='percent', color = 'psrc_light', interactive=FALSE)+ 
  xlab(as.character(g[cat_name])) + ylab("Walk mode share")+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)

}
 # p<-create_facet_bar_chart(t=transit_summs_long , x=cat_name, y='share', f='non_motorized_mode', g='survey', moe='share_moe', est='percent', color = 'psrc_light')+ theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()



#print(psrc_pres, target = "C:/Users/schildress/OneDrive - Puget Sound Regional Council/Documents/hhts_sntc_sept22_new2.pptx")
```
```{r }
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  # cat_name <- eval(names(g))
  cat_name <- names(g)
  print(cat_name)
  transit_summs_2017_2019 <- hhts_count(transit_data_17_19,
                                        group_vars=c(names(g),'mode_simple'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter(cat_name!='Total') %>%
    filter(mode_simple=='Transit')%>%drop_na(cat_name)

  transit_summs_2021 <- hhts_count(transit_data_21,
                                   group_vars=c(names(g),'mode_simple'),
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter(cat_name!='Total') %>%
    filter(mode_simple=='Transit')%>%drop_na(cat_name)
  if(cat_name=='race_eth_broad'){

     transit_summs_2017_2019 <- transit_summs_2017_2019 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')

        transit_summs_2021 <- transit_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  transit_summs_2017_2019$race_eth_broad<-factor(transit_summs_2017_2019$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))

   transit_summs_2021$race_eth_broad<-factor(transit_summs_2021$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))



  }



  transit_summs_long<- rbind(transit_summs_2017_2019, transit_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))


   # factorize variable column. Adjust levels in function definition
  #transit_summs_long <- factor_variable_column(names(g), transit_summs_long)

  #transit_summs_2021_only<-  transit_summs_2021%>%filter(non_motorized_mode=='Walk/Bike')

 # p0<-create_bar_chart(t=transit_summs_2021_only , x=cat_name, y='share',  f=cat_name,moe='share_moe', est='percent', #color='psrc_light')+ t#heme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()


   p0<-create_column_chart(t=transit_summs_2021 , x=cat_name, y='share', f='mode_simple', moe='share_moe', est='percent', color = 'psrc_light', interactive=FALSE) + 
  xlab(as.character(g[cat_name])) + ylab('Transit mode share')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  


  p1<-create_column_chart(t=transit_summs_long , x=cat_name, y='share', f='survey', moe='share_moe', est='percent', color = 'psrc_light', interactive=FALSE)+ 
  xlab(as.character(g[cat_name])) + ylab("Transit mode share")+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)

}
```



```{r}
  all_trips_summs_2017_2019 <- hhts_count(transit_data_17_19,
                                        group_vars=c('simple_purpose'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter(simple_purpose!='Total')




  all_trips_summs_2021 <- hhts_count(transit_data_21,
                                   group_vars=c('simple_purpose'),
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter(simple_purpose!='Total')



  all_trips_summs_long<- rbind(all_trips_summs_2017_2019, all_trips_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))



p<-create_column_chart(t=all_trips_summs_long , x='simple_purpose', y='share', f='survey', moe='share_moe', est='percent', color='psrc_dark')+
  ylab('Share of All Trips') + xlab('Purpose')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+ theme(axis.title.x = element_text(size=50,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p)
  
  



  


```

```{r }
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  # cat_name <- eval(names(g))
  cat_name <- names(g)
  print(cat_name)
  transit_summs_2017_2019 <- hhts_count(transit_data_17_19,
                                        group_vars=c(names(g),'simple_purpose'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
    filter(cat_name!='Total') %>%drop_na(c(cat_name, 'simple_purpose'))%>%filter(simple_purpose!='Total')

  transit_summs_2021 <- hhts_count(transit_data_21,
                                   group_vars=c(names(g),'simple_purpose'),
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
    filter(cat_name!='Total') %>%drop_na(cat_name) %>%drop_na(c(cat_name, 'simple_purpose'))%>%filter(simple_purpose!='Total')
  
  if(cat_name=='race_eth_broad'){

     transit_summs_2017_2019 <- transit_summs_2017_2019 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')

        transit_summs_2021 <- transit_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  transit_summs_2017_2019$race_eth_broad<-factor(transit_summs_2017_2019$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))

   transit_summs_2021$race_eth_broad<-factor(transit_summs_2021$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))



  }



  transit_summs_long<- rbind(transit_summs_2017_2019, transit_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))


   # factorize variable column. Adjust levels in function definition
  #transit_summs_long <- factor_variable_column(names(g), transit_summs_long)

  #transit_summs_2021_only<-  transit_summs_2021%>%filter(non_motorized_mode=='Walk/Bike')

 # p0<-create_bar_chart(t=transit_summs_2021_only , x=cat_name, y='share',  f=cat_name,moe='share_moe', est='percent', #color='psrc_dark')+ t#heme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()


   p0<-create_column_chart(t=transit_summs_2021 , x='simple_purpose', y='share', f=cat_name, moe='share_moe', est='percent', color = 'psrc_dark', interactive=FALSE) + 
  xlab(as.character(g[cat_name])) + ylab('Trip Purpose')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  


  p1<-create_facet_bar_chart(t=transit_summs_long , w.x=cat_name, w.y='share', f='survey',g='simple_purpose', w.moe='share_moe', est.type='percent', w.color = 'psrc_dark')+ 
  xlab(as.character(g[cat_name])) + ylab("Transit mode share")+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)

}
```


```{r }
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  # cat_name <- eval(names(g))
  cat_name <- names(g)
  print(cat_name)
  transit_summs_2017_2019 <- hhts_median(transit_data_17_19,
                                         stat_var = 'trip_path_distance',
                                        group_vars=c(cat_name),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%
filter(cat_name!='Other')%>%drop_na(cat_name)

  

  transit_summs_2021<- hhts_median(transit_data_21,
                                         stat_var = 'trip_path_distance',
                                        group_vars=c(cat_name),
                                        spec_wgt='trip_weight_2021_ABS_Panel_adult')%>%
filter(cat_name!='Other')%>%drop_na(cat_name)
  
  if(cat_name=='race_eth_broad'){

     transit_summs_2017_2019 <- transit_summs_2017_2019 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')

        transit_summs_2021 <- transit_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  transit_summs_2017_2019$race_eth_broad<-factor(transit_summs_2017_2019$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))

   transit_summs_2021$race_eth_broad<-factor(transit_summs_2021$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))



  }



  transit_summs_long<- rbind(transit_summs_2017_2019, transit_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))


   # factorize variable column. Adjust levels in function definition
  #transit_summs_long <- factor_variable_column(names(g), transit_summs_long)

  #transit_summs_2021_only<-  transit_summs_2021%>%filter(non_motorized_mode=='Walk/Bike')

 # p0<-create_bar_chart(t=transit_summs_2021_only , x=cat_name, y='share',  f=cat_name,moe='share_moe', est='percent', #color='psrc_dark')+ t#heme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()


   p0<-create_column_chart(t=transit_summs_2021 , x=cat_name, y='trip_path_distance_median', f='survey', moe='trip_path_distance_median_moe', est='number', color = 'psrc_pairs', interactive=FALSE) + 
  xlab(as.character(g[cat_name])) + ylab('Median Distance')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  


  p1<-create_column_chart(t=transit_summs_long , x=cat_name, y='trip_path_distance_median', f='survey', moe='trip_path_distance_median_moe', est='number', color = 'psrc_pairs', interactive=FALSE)+ 
  xlab(as.character(g[cat_name])) + ylab("Median Distance")+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)

}
```

```{r }
print(psrc_pres, target = "C:/Users/schildress/OneDrive - Puget Sound Regional Council/Documents/hhts_sntc_sept22_new6.pptx")
```

```{r}
person_data_17_19<-person_data_17_19%>%mutate(NoVehicles=ifelse(vehicle_count=='0 (no vehicles)', 'No Vehicles', "Has Vehicles")) %>%mutate(hhincome_100= case_when(hhincome_broad=='$100,000-$199,000' ~ '$100,000 or more',
                                 hhincome_broad=='$200,000 or more' ~ '$100,000 or more',
                                 TRUE~hhincome_broad))%>%
  mutate(age_grp= case_when(age=='75-84 years' ~ '75 years or older', 
                             age == '85 or years older' ~ '75 years or older',
                            TRUE ~ age))%>%
  mutate(has_disab_plates=ifelse(disability_vehicle_count>0, 'Has Disability Vehicle Plates', 'No Disability Vehicle Plates'))
 

person_data_21<-person_data_21%>%mutate(NoVehicles=ifelse(vehicle_count=='0 (no vehicles)', 'No Vehicles', "Has Vehicles")) %>%
  mutate(hhincome_100= case_when(hhincome_broad=='$100,000-$199,000' ~ '$100,000 or more',
                                 hhincome_broad=='$200,000 or more' ~ '$100,000 or more',
                                 TRUE~hhincome_broad))%>%
  mutate(age_grp= case_when(age=='75-84 years' ~ '75 years or older', 
                             age == '85 or years older' ~ '75 years or older',
                            TRUE ~ age))%>%
  mutate(has_disab_plates=ifelse(disability_vehicle_count>0, 'Has Disability Vehicle Plates', 'No Disability Vehicle Plates'))
 

person_data_17_19$hhincome_100_f=factor(person_data_17_19$hhincome_100,levels=c("Under $25,000","$25,000-$49,999",   "$50,000-$74,999","$75,000-$99,999","$100,000 or more", "Prefer not to answer"))

person_data_21$hhincome_100_f=factor(person_data_21$hhincome_100,levels=c("Under $25,000","$25,000-$49,999",   "$50,000-$74,999","$75,000-$99,999","$100,000 or more", "Prefer not to answer"))


person_data_17_19$NoVehicles=factor(person_data_17_19$NoVehicles, levels= c('No Vehicles', 'Has Vehicles'))

person_data_21$NoVehicles=factor(person_data_21$NoVehicles, levels=c('No Vehicles', 'Has Vehicles'))



simple_groupings<-c('race_eth_broad'='Race/Ethnicity','hhincome_100_f'='Household Income','NoVehicles'= 'Household Vehicles', 'age_grp'= 'Age Group', 'sample_county' ='Home County', 'age_category'='Age Category', 'final_home_is_rgc'='Home in Regional Growth Center', has_disab_plates='Has Disability Vehicle Plates' )
```


Trips per Person
```{r }
for(i in seq(1, length(simple_groupings))){
  g <- simple_groupings[i]
  # cat_name <- eval(names(g))
  cat_name <- names(g)
  print(cat_name)
  transit_summs_2017_2019 <- hhts_count(transit_data_17_19,
                                        group_vars=c(names(g),'simple_purpose'),
                                        spec_wgt='trip_weight_2017_2019_v2021_adult')%>%drop_na(c(cat_name, 'simple_purpose'))%>%filter(cat_name!='Total')

  transit_summs_2021 <- hhts_count(transit_data_21,
                                   group_vars=c(names(g),'simple_purpose'),
                                   spec_wgt='trip_weight_2021_ABS_Panel_adult') %>%drop_na(cat_name) %>%drop_na(c(cat_name, 'simple_purpose'))%>%filter(cat_name!='Total')
  
  
    person_summs_2017_2019 <- hhts_count(person_data_17_19,
                                        group_vars=c(names(g)),
                                        spec_wgt='hh_weight_2017_2019_v2021_adult')%>%
    filter(cat_name!='Total') %>%drop_na(cat_name)

  person_summs_2021 <- hhts_count(person_data_21,
                                   group_vars=c(names(g)),
                                   spec_wgt='person_weight_2021_ABS_Panel_adult')%>%
    filter(cat_name!='Total') %>%drop_na(cat_name)

  
  if(cat_name=='race_eth_broad'){

     transit_summs_2017_2019 <- transit_summs_2017_2019 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')

        transit_summs_2021 <- transit_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  transit_summs_2017_2019$race_eth_broad<-factor(transit_summs_2017_2019$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))

   transit_summs_2021$race_eth_broad<-factor(transit_summs_2021$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))

        person_summs_2017_2019 <- person_summs_2017_2019 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')

        person_summs_2021 <- person_summs_2021 %>%
  mutate(race_eth_broad= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  filter(race_eth_broad!='No')
  person_summs_2017_2019$race_eth_broad<-factor(person_summs_2017_2019$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))

   person_summs_2021$race_eth_broad<-factor(person_summs_2021$race_eth_broad, levels=
                                                  c('Asian', 'Black', 'Hispanic', 'Other', 'White'))



  }

 trips_persons_2017_2019<- merge(transit_summs_2017_2019, person_summs_2017_2019, by=cat_name, suffixes=c('_trips','_people' ))
 trips_persons_2021<- merge(transit_summs_2021, person_summs_2021, by=cat_name, suffixes=c('_trips','_people' ))
 
 trips_persons_2017_2019<- trips_persons_2017_2019 %>%mutate(trips_per_person=count_trips/count_people,moe_trips_person=moe_ratio(count_trips, count_people, count_moe_trips, count_moe_people))
 
 
  trips_persons_2021<- trips_persons_2021 %>%mutate(trips_per_person=count_trips/count_people,moe_trips_person=moe_ratio(count_trips, count_people, count_moe_trips, count_moe_people))
  
 trips_persons_long<- rbind(trips_persons_2017_2019, trips_persons_2021) %>%
    mutate(survey_trips = str_replace_all(survey_trips, "_", "/"))

trips_persons_long_all_trips<-trips_persons_long%>%filter(simple_purpose=='Total')

 trips_persons_2021_purpose<-trips_persons_2021%>%filter(simple_purpose!='Total')
   # factorize variable column. Adjust levels in function definition
  #transit_summs_long <- factor_variable_column(names(g), transit_summs_long)

  #transit_summs_2021_only<-  transit_summs_2021%>%filter(non_motorized_mode=='Walk/Bike')

 # p0<-create_bar_chart(t=transit_summs_2021_only , x=cat_name, y='share',  f=cat_name,moe='share_moe', est='percent', #color='psrc_dark')+ t#heme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()
 trips_persons_long_shop_trips<-trips_persons_long%>%filter(simple_purpose=='Shop')

   p0<-create_column_chart(t=trips_persons_2021_purpose , x=cat_name, y='trips_per_person', f='simple_purpose', moe='moe_trips_person', est='number', color = 'psrc_pairs', interactive=FALSE) + 
  xlab(as.character(g[cat_name])) + ylab('Trips per Person')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  
  
 p1<-create_facet_bar_chart(t=trips_persons_long, w.x=cat_name, w.y='trips_per_person', f='survey_trips',g='simple_purpose', w.moe='moe_trips_person', est.type='number', w.color = 'psrc_dark')




   p2<-create_column_chart(t=trips_persons_long_all_trips , x=cat_name, y='trips_per_person', f='survey_trips', moe='moe_trips_person', est='number', color = 'psrc_pairs', interactive=FALSE) + 
  xlab(as.character(g[cat_name])) + ylab('Trips per Person')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p2)

  
   p3<-create_column_chart(t=trips_persons_long_shop_trips , x=cat_name, y='trips_per_person', f='survey_trips', moe='moe_trips_person', est='number', color = 'psrc_pairs', interactive=FALSE) + 
  xlab(as.character(g[cat_name])) + ylab('Trips per Person')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p2)
  

  
  

  
}

print(psrc_pres, target = "C:/Users/schildress/OneDrive - Puget Sound Regional Council/Documents/hhts_sntc_sept22_test456.pptx")
```

```{r}
transit_summs_long<- rbind(transit_summs_2017_2019, transit_summs_2021) %>%
    mutate(survey = str_replace_all(survey, "_", "/"))


   # factorize variable column. Adjust levels in function definition
  #transit_summs_long <- factor_variable_column(names(g), transit_summs_long)

  #transit_summs_2021_only<-  transit_summs_2021%>%filter(non_motorized_mode=='Walk/Bike')

 # p0<-create_bar_chart(t=transit_summs_2021_only , x=cat_name, y='share',  f=cat_name,moe='share_moe', est='percent', #color='psrc_dark')+ t#heme(axis.text.x = element_text(size=35,color="#4C4C4C"))+coord_flip()


   p0<-create_column_chart(t=transit_summs_2021 , x='simple_purpose', y='share', f=cat_name, moe='share_moe', est='percent', color = 'psrc_dark', interactive=FALSE) + 
  xlab(as.character(g[cat_name])) + ylab('Trip Purpose')+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p0)
  


  p1<-create_facet_bar_chart(t=transit_summs_long , w.x=cat_name, w.y='share', f='survey',g='simple_purpose', w.moe='share_moe', est.type='percent', w.color = 'psrc_dark')+ 
  xlab(as.character(g[cat_name])) + ylab("Transit mode share")+theme(axis.text.x = element_text(size=35,color="#4C4C4C"))+theme(axis.title.y = element_text(size=50,color="#4C4C4C"))+theme(axis.text.y.left =   element_text(size=35,color="#4C4C4C"))+theme(legend.text=element_text(size=35))+
  scale_x_discrete(labels=function(x) str_wrap(x, width=10))
  print(p1)

```

