---
title: "HHTS: transit ridership (2017, 2019, 2021) Analysis"
author: "Mary Richards"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 6
    toc_float: true
---

```{r rmarkdown setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE) # formatting
```

```{r library setup, include=FALSE}
devtools::install_github("psrc/psrc.travelsurvey")

library(tidyverse)
# library(psrccensus)
library(psrc.travelsurvey)
# library(rlang) #required for psrccensus
# library(emmeans) #required for rlang
library(magrittr)
library(knitr)
library(kableExtra)
library(ggplot2)

# library(survey)
library(summarytools) #freq
library(table1)  #nice descriptive summary table
library(scales) #number formatting
library(ggpubr) #graphing - ggarrange fx
library(forcats) #for factor releveling
```

```{r surveys}
# global variables
survey_a <- list(survey = '2017', label = '2017')
survey_b <- list(survey = '2019', label = '2019')
survey_ab <- list(survey = '2017_2019', label = '2017/2019')
survey_c <- list(survey = '2021', label = '2021')
```

```{r functions}
# rounding function applied to categorical columns 
my.render.cat <- function(x) {
    c("", sapply(stats.default(x), function(y) with(y,
        sprintf("%d (%.0f%%)", FREQ, PCT))))}

# hhts_count and prep for graphing
count_prep_fx <- function(data, grp1, grp2) {
  
  hhts_count(data,
             group_vars = c(grp1, grp2)) %>%
    mutate(period = data$label) %>%
    filter(!!sym(grp2)!="Total") %>% 
    filter(!!sym(grp1)!="NA")
}

# proportion graphs
share_plot <- function(dt, grp_var, num_var, legend_name, tbl_name){
  
  fill_group <- dt[[grp_var]]
  x_axis_grp <- dt[[num_var]]
  
  share_plot <- ggplot(dt,
                       aes(x=x_axis_grp,
                           y=share,
                           fill=fill_group)) +
  geom_bar(stat="identity",
           position="dodge2") +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        axis.title.x = element_blank()) +
  labs(y = "Share",
       fill = legend_name) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                size=.5, width=.2,
                position=position_dodge(0.9))
  
  annotate_figure(share_plot, top = text_grob(tbl_name, 
               color = "blue", face = "bold", size = 14))
}

# side-by-side count and proportion graphs
count_and_share_plot <- function(dt, grp_var, num_var, legend_name, tbl_name){
  
  fill_group <- dt[[grp_var]]
  x_axis_grp <- dt[[num_var]]
  
  count_plot <- ggplot(dt,
                       aes(x=x_axis_grp,
                           y=count,
                           fill=fill_group)) +
  geom_bar(stat="identity",
           position="dodge2") +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        axis.title.x = element_blank()) +
  labs(y = "Number",
       fill = legend_name) +
  scale_y_continuous(labels = scales::comma) +
  geom_errorbar(aes(ymin=count-count_moe, ymax=count+count_moe),
                size=.5,width=.2,
                position=position_dodge(.9))
  
  
  share_plot <- ggplot(dt,
                       aes(x=x_axis_grp,
                           y=share,
                           fill=fill_group)) +
  geom_bar(stat="identity",
           position="dodge2") +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        axis.title.x = element_blank()) +
  labs(y = "Share",
       fill = legend_name) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                size=.5, width=.2,
                position=position_dodge(0.9))
  
  autos <- ggarrange(count_plot, share_plot, 
                     # ncol = 2, nrow = 1,
                     common.legend = TRUE,
                     legend = "right")
  annotate_figure(autos, top = text_grob(tbl_name, 
               color = "blue", face = "bold", size = 14))
}
```

# Introduction
The work presented in this document contains analysis of transit ridership in relation to auto ownership and geography. The analysis includes 2021 data and utilizes PSRC's R package, [psrc.travelsurvey](https://psrc.github.io/psrc.travelsurvey/index.html).

* **Has commuting by transit changed with COVID-19?**
  + Who uses transit? Have there been shifts over time?
* **How does vehicle ownership influence transit commuting use over time?**

```{r variables}
all_vars_p <- c("commute_mode",
                "workplace",
                # "personid",
                "race_category",
                "race_eth_broad",
                # "hhid",
                "license",
                "hhsize",
                "numadults",
                "numworkers",
                "vehicle_count",
                "hhincome_broad",
                "hhincome_detailed",
                "lifecycle",
                "final_home_rgcnum",
                # "final_home_is_rgc",
                "education")
```

*Filtering out children and those who do not commute* 
```{r 2017 survey}
psurvey_17 <- get_hhts(survey = survey_a$survey,
                       level="p",
                       vars=all_vars_p)
  
## POC/Non-POC
psurvey_17_simp_2grp <- psurvey_17 %>%
  filter(race_eth_broad!="Child -- no race specified") %>% 
  drop_na(commute_mode) %>% 
  mutate(race_eth_broad = factor(case_when(race_eth_broad=="Asian only, non-Hispanic/Latinx" |
                                             race_eth_broad=="Black or African American only, non-Hispanic/Latinx" |
                                             race_eth_broad=="Other race, including multi-race non-Hispanic" |
                                             race_eth_broad=="Hispanic or Latinx"~ "Persons of Color",
                                           TRUE ~ "White only, non-Hispanic/Latinx"))) %>% 
  mutate(commute_mode = factor(case_when(commute_mode=="Bus (public transit)" |
                                           commute_mode=="Commuter rail (Sounder, Amtrak)" |
                                           commute_mode=="Ferry or water taxi" |
                                           commute_mode=="Paratransit" |
                                           commute_mode=="Streetcar" |
                                           commute_mode=="Urban rail (Link light rail, monorail)" ~ "2-Public Transit",
                                         commute_mode=="Bicycle or e-bike" |
                                           commute_mode=="Motorcycle/moped/scooter" |
                                           commute_mode=="Walk, jog, or wheelchair" ~ "1-Active/micro-transit",
                                         commute_mode=="Drive alone" ~ "3-Drive alone",
                                         TRUE ~ "4-Other modes"))) %>% 
  mutate(final_home_rgcnum = factor(case_when(final_home_rgcnum=="Not RGC" ~ "Not RGC",
                                              TRUE ~ "RGC")))


## Asian POC/Non-Asian POC/White
psurvey_17_simp_3grp <- psurvey_17 %>%
  filter(race_eth_broad!="Child -- no race specified") %>% 
  drop_na(commute_mode) %>% 
  mutate(race_eth_broad = factor(case_when(race_eth_broad=="Asian only, non-Hispanic/Latinx" ~ "Asian POC",
                                           race_eth_broad=="Black or African American only, non-Hispanic/Latinx" |
                                             race_eth_broad=="Other race, including multi-race non-Hispanic" |
                                             race_eth_broad=="Hispanic or Latinx"~ "Non-Asian POC",
                                           TRUE ~ "White only, non-Hispanic/Latinx"))) %>% 
  mutate(commute_mode = factor(case_when(commute_mode=="Bus (public transit)" |
                                           commute_mode=="Commuter rail (Sounder, Amtrak)" |
                                           commute_mode=="Ferry or water taxi" |
                                           commute_mode=="Paratransit" |
                                           commute_mode=="Streetcar" |
                                           commute_mode=="Urban rail (Link light rail, monorail)" ~ "2-Public Transit",
                                         commute_mode=="Bicycle or e-bike" |
                                           commute_mode=="Motorcycle/moped/scooter" |
                                           commute_mode=="Walk, jog, or wheelchair" ~ "1-Active/micro-transit",
                                         commute_mode=="Drive alone" ~ "3-Drive alone",
                                         TRUE ~ "4-Other modes"))) %>% 
  mutate(final_home_rgcnum = factor(case_when(final_home_rgcnum=="Not RGC" ~ "Not RGC",
                                              TRUE ~ "RGC")))

# freq(psurvey_17_simp$race_eth_broad)
# freq(psurvey_17$commute_mode)
# freq(psurvey_17_simp$commute_mode)

# freq(psurvey_17$final_home_rgcnum)
# freq(psurvey_17_simp$final_home_rgcnum)
```

```{r 2019 survey}
psurvey_19 <- get_hhts(survey = survey_b$survey,
                       level="p",
                       vars=all_vars_p)
  

## POC/Non-POC
psurvey_19_simp_2grp <- psurvey_19 %>%
  filter(race_eth_broad!="Child -- no race specified") %>% 
  drop_na(commute_mode) %>% 
  mutate(race_eth_broad = factor(case_when(race_eth_broad=="Asian only, non-Hispanic/Latinx" |
                                             race_eth_broad=="Black or African American only, non-Hispanic/Latinx" |
                                             race_eth_broad=="Other race, including multi-race non-Hispanic" |
                                             race_eth_broad=="Hispanic or Latinx"~ "Persons of Color",
                                           TRUE ~ "White only, non-Hispanic/Latinx"))) %>% 
  mutate(commute_mode = factor(case_when(commute_mode=="Bus (public transit)" |
                                           commute_mode=="Commuter rail (Sounder, Amtrak)" |
                                           commute_mode=="Ferry or water taxi" |
                                           commute_mode=="Paratransit" |
                                           commute_mode=="Streetcar" |
                                           commute_mode=="Urban rail (Link light rail, monorail)" ~ "2-Public Transit",
                                         commute_mode=="Bicycle or e-bike" |
                                           commute_mode=="Motorcycle/moped/scooter" |
                                           commute_mode=="Walk, jog, or wheelchair" ~ "1-Active/micro-transit",
                                         commute_mode=="Drive alone" ~ "3-Drive alone",
                                         TRUE ~ "4-Other modes"))) %>% 
  mutate(final_home_rgcnum = factor(case_when(final_home_rgcnum=="Not RGC" ~ "Not RGC",
                                              TRUE ~ "RGC")))


## Asian POC/Non-Asian POC/White
psurvey_19_simp_3grp <- psurvey_19 %>%
  filter(race_eth_broad!="Child -- no race specified") %>% 
  drop_na(commute_mode) %>% 
  mutate(race_eth_broad = factor(case_when(race_eth_broad=="Asian only, non-Hispanic/Latinx" ~ "Asian POC",
                                           race_eth_broad=="Black or African American only, non-Hispanic/Latinx" |
                                             race_eth_broad=="Other race, including multi-race non-Hispanic" |
                                             race_eth_broad=="Hispanic or Latinx"~ "Non-Asian POC",
                                           TRUE ~ "White only, non-Hispanic/Latinx"))) %>% 
  mutate(commute_mode = factor(case_when(commute_mode=="Bus (public transit)" |
                                           commute_mode=="Commuter rail (Sounder, Amtrak)" |
                                           commute_mode=="Ferry or water taxi" |
                                           commute_mode=="Paratransit" |
                                           commute_mode=="Streetcar" |
                                           commute_mode=="Urban rail (Link light rail, monorail)" ~ "2-Public Transit",
                                         commute_mode=="Bicycle or e-bike" |
                                           commute_mode=="Motorcycle/moped/scooter" |
                                           commute_mode=="Walk, jog, or wheelchair" ~ "1-Active/micro-transit",
                                         commute_mode=="Drive alone" ~ "3-Drive alone",
                                         TRUE ~ "4-Other modes"))) %>% 
  mutate(final_home_rgcnum = factor(case_when(final_home_rgcnum=="Not RGC" ~ "Not RGC",
                                              TRUE ~ "RGC")))
```

```{r 2017-19 survey}
psurvey_1719 <- get_hhts(survey = survey_ab$survey,
                       level="p",
                       vars=all_vars_p)
  

## POC/Non-POC
psurvey_1719_simp_2grp <- psurvey_1719 %>%
  filter(race_eth_broad!="Child -- no race specified") %>% 
  drop_na(commute_mode) %>% 
  mutate(race_eth_broad = factor(case_when(race_eth_broad=="Asian only, non-Hispanic/Latinx" |
                                             race_eth_broad=="Black or African American only, non-Hispanic/Latinx" |
                                             race_eth_broad=="Other race, including multi-race non-Hispanic" |
                                             race_eth_broad=="Hispanic or Latinx"~ "Persons of Color",
                                           TRUE ~ "White only, non-Hispanic/Latinx"))) %>% 
  mutate(commute_mode = factor(case_when(commute_mode=="Bus (public transit)" |
                                           commute_mode=="Commuter rail (Sounder, Amtrak)" |
                                           commute_mode=="Ferry or water taxi" |
                                           commute_mode=="Paratransit" |
                                           commute_mode=="Streetcar" |
                                           commute_mode=="Urban rail (Link light rail, monorail)" ~ "2-Public Transit",
                                         commute_mode=="Bicycle or e-bike" |
                                           commute_mode=="Motorcycle/moped/scooter" |
                                           commute_mode=="Walk, jog, or wheelchair" ~ "1-Active/micro-transit",
                                         commute_mode=="Drive alone" ~ "3-Drive alone",
                                         TRUE ~ "4-Other modes"))) %>% 
  mutate(final_home_rgcnum = factor(case_when(final_home_rgcnum=="Not RGC" ~ "Not RGC",
                                              TRUE ~ "RGC")))


## Asian POC/Non-Asian POC/White
psurvey_1719_simp_3grp <- psurvey_1719 %>%
  filter(race_eth_broad!="Child -- no race specified") %>% 
  drop_na(commute_mode) %>% 
  mutate(race_eth_broad = factor(case_when(race_eth_broad=="Asian only, non-Hispanic/Latinx" ~ "Asian POC",
                                           race_eth_broad=="Black or African American only, non-Hispanic/Latinx" |
                                             race_eth_broad=="Other race, including multi-race non-Hispanic" |
                                             race_eth_broad=="Hispanic or Latinx"~ "Non-Asian POC",
                                           TRUE ~ "White only, non-Hispanic/Latinx"))) %>% 
  mutate(commute_mode = factor(case_when(commute_mode=="Bus (public transit)" |
                                           commute_mode=="Commuter rail (Sounder, Amtrak)" |
                                           commute_mode=="Ferry or water taxi" |
                                           commute_mode=="Paratransit" |
                                           commute_mode=="Streetcar" |
                                           commute_mode=="Urban rail (Link light rail, monorail)" ~ "2-Public Transit",
                                         commute_mode=="Bicycle or e-bike" |
                                           commute_mode=="Motorcycle/moped/scooter" |
                                           commute_mode=="Walk, jog, or wheelchair" ~ "1-Active/micro-transit",
                                         commute_mode=="Drive alone" ~ "3-Drive alone",
                                         TRUE ~ "4-Other modes"))) %>% 
  mutate(final_home_rgcnum = factor(case_when(final_home_rgcnum=="Not RGC" ~ "Not RGC",
                                              TRUE ~ "RGC")))
```

```{r 2021 survey}
psurvey_21 <- get_hhts(survey = survey_c$survey,
                       level="p",
                       vars=all_vars_p)
  

## POC/Non-POC
psurvey_21_simp_2grp <- psurvey_21 %>%
  filter(race_eth_broad!="Child -- no race specified") %>% 
  drop_na(commute_mode) %>% 
  mutate(race_eth_broad = factor(case_when(race_eth_broad=="Asian only, non-Hispanic/Latinx" |
                                             race_eth_broad=="Black or African American only, non-Hispanic/Latinx" |
                                             race_eth_broad=="Other race, including multi-race non-Hispanic" |
                                             race_eth_broad=="Hispanic or Latinx"~ "Persons of Color",
                                           TRUE ~ "White only, non-Hispanic/Latinx"))) %>% 
  mutate(commute_mode = factor(case_when(commute_mode=="Bus (public transit)" |
                                           commute_mode=="Commuter rail (Sounder, Amtrak)" |
                                           commute_mode=="Ferry or water taxi" |
                                           commute_mode=="Paratransit" |
                                           commute_mode=="Streetcar" |
                                           commute_mode=="Urban rail (Link light rail, monorail)" ~ "2-Public Transit",
                                         commute_mode=="Bicycle or e-bike" |
                                           commute_mode=="Motorcycle/moped/scooter" |
                                           commute_mode=="Walk, jog, or wheelchair" ~ "1-Active/micro-transit",
                                         commute_mode=="Drive alone" ~ "3-Drive alone",
                                         TRUE ~ "4-Other modes"))) %>% 
  mutate(final_home_rgcnum = factor(case_when(final_home_rgcnum=="Not RGC" ~ "Not RGC",
                                              TRUE ~ "RGC")))


## Asian POC/Non-Asian POC/White
psurvey_21_simp_3grp <- psurvey_21 %>%
  filter(race_eth_broad!="Child -- no race specified") %>% 
  drop_na(commute_mode) %>% 
  mutate(race_eth_broad = factor(case_when(race_eth_broad=="Asian only, non-Hispanic/Latinx" ~ "Asian POC",
                                           race_eth_broad=="Black or African American only, non-Hispanic/Latinx" |
                                             race_eth_broad=="Other race, including multi-race non-Hispanic" |
                                             race_eth_broad=="Hispanic or Latinx"~ "Non-Asian POC",
                                           TRUE ~ "White only, non-Hispanic/Latinx"))) %>% 
  mutate(commute_mode = factor(case_when(commute_mode=="Bus (public transit)" |
                                           commute_mode=="Commuter rail (Sounder, Amtrak)" |
                                           commute_mode=="Ferry or water taxi" |
                                           commute_mode=="Paratransit" |
                                           commute_mode=="Streetcar" |
                                           commute_mode=="Urban rail (Link light rail, monorail)" ~ "2-Public Transit",
                                         commute_mode=="Bicycle or e-bike" |
                                           commute_mode=="Motorcycle/moped/scooter" |
                                           commute_mode=="Walk, jog, or wheelchair" ~ "1-Active/micro-transit",
                                         commute_mode=="Drive alone" ~ "3-Drive alone",
                                         TRUE ~ "4-Other modes"))) %>% 
  mutate(final_home_rgcnum = factor(case_when(final_home_rgcnum=="Not RGC" ~ "Not RGC",
                                              TRUE ~ "RGC")))
```

# Descriptive Stats: 2017
## Transit
Using 'commute_mode' defined as: 'Usual way of commuting to current location/office'
\
    *survey logic*: if workplace = fixed/varied/telework some days

Workplace options: 

* **Only one work location outside of home**
* Work at home ONLY (telework, self-employed)
* **Telework some days and travel to a work location some days**
* **Work location regularly varies (different offices/jobsites)**
* Drive/travel for work (driver, sales)
\
\

```{r}
# set year
df <- psurvey_17
df_simp_2 <- psurvey_17_simp_2grp
df_simp_3 <- psurvey_17_simp_3grp
srvy_yr <- 2017

freq(df$commute_mode)
# freq(psurvey_17$workplace)
```
The NAs represent those who only work at home and those who drive for their jobs.
\
\

When the commute modes are simplified:
```{r}
# table(psurvey_17_simp$commute_mode, psurvey_17_simp$commute_mode)
# freq(psurvey_17_simp$commute_mode)
freq(df_simp_2$commute_mode)
```
<span style="color:#00716c">*do we count hired services/taxis as SOV?* it's currently categorized as 'other modes'</span> 
\
\

### Race - POC/Non-POC
```{r}
race_com_cnt_2017 <-count_prep_fx(df_simp_2, 'race_eth_broad', 'commute_mode')

# count_and_share_plot(race_com_cnt_2017, 'race_eth_broad', 'commute_mode', 'Race', 'Commute Mode and Race: 2017')
share_plot(race_com_cnt_2017, 'race_eth_broad', 'commute_mode', 'Race', 'Commute Mode and Race: 2017')
```
\
\

#### Focusing on Active/Public Transit
To look at numbers more closely
```{r}
active_public_cnt_2017 <- race_com_cnt_2017 %>% 
  filter(commute_mode=="1-Active/micro-transit" |
           commute_mode=="2-Public Transit")

share_plot(active_public_cnt_2017, 'race_eth_broad', 'commute_mode', 'Race', 'Commute Mode and Race: 2017')
```
\
\

### Race - Asian POC/Non-Asian POC/White
```{r}
race_com_cnt_2017 <-count_prep_fx(df_simp_3, 'race_eth_broad', 'commute_mode')

# count_and_share_plot(race_com_cnt_2017, 'race_eth_broad', 'commute_mode', 'Race', 'Commute Mode and Race: 2017')
share_plot(race_com_cnt_2017, 'race_eth_broad', 'commute_mode', 'Race', 'Commute Mode and Race: 2017')
```
\
\

### Income
```{r}
race_com_cnt_2017 <- count_prep_fx(df_simp_2, 'hhincome_broad', 'commute_mode')
  
race_com_cnt_2017$hhincome_broad <- factor(race_com_cnt_2017$hhincome_broad,
                                                 levels=c("Under $25,000",
                                                          "$25,000-$49,999",
                                                          "$50,000-$74,999",
                                                          "$75,000-$99,999",
                                                          "$100,000 or more",
                                                          "Prefer not to answer"))

share_plot(race_com_cnt_2017, 'hhincome_broad', 'commute_mode', 'Income', 'Commute Mode and Income: 2017')

```
\
\

#### Focusing on Active/Public Transit
To look at numbers more closely
```{r}
active_public_cnt_2017 <- race_com_cnt_2017 %>% 
  filter(commute_mode=="1-Active/micro-transit" |
           commute_mode=="2-Public Transit")

share_plot(active_public_cnt_2017, 'hhincome_broad', 'commute_mode', 'Income', 'Commute Mode and Income: 2017')
```
\
\

### Education
```{r}
race_com_cnt_2017 <- count_prep_fx(df_simp_2, 'education', 'commute_mode')

race_com_cnt_2017$education <- factor(race_com_cnt_2017$education,
                                                 levels=c("Less than high school",
                                                          "High school graduate",
                                                          "Vocational/technical training",
                                                          "Some college",
                                                          "Associates degree",
                                                          "Bachelor degree",
                                                          "Graduate/post-graduate degree"))

share_plot(race_com_cnt_2017, 'education', 'commute_mode', 'Education', 'Commute Mode and Education: 2017')
```
\
\

#### Focusing on Active/Public Transit
To look at numbers more closely
```{r}
active_public_cnt_2017 <- race_com_cnt_2017 %>% 
  filter(commute_mode=="1-Active/micro-transit" |
           commute_mode=="2-Public Transit")

share_plot(active_public_cnt_2017, 'education', 'commute_mode', 'Education', 'Commute Mode and Education: 2017')
```
\
\

### Regional Growth Center - home location
```{r}
race_com_cnt_2017 <- count_prep_fx(df_simp_2, 'final_home_rgcnum', 'commute_mode')

race_com_cnt_2017$final_home_rgcnum <- factor(race_com_cnt_2017$final_home_rgcnum,
                                                 levels=c("RGC",
                                                          "Not RGC"))

share_plot(race_com_cnt_2017, 'final_home_rgcnum', 'commute_mode', 'Regional Growth Center', 'Commute Mode and RGC: 2017')
```
\
\

#### Focusing on Active/Public Transit
To look at numbers more closely
```{r}
active_public_cnt_2017 <- race_com_cnt_2017 %>% 
  filter(commute_mode=="1-Active/micro-transit" |
           commute_mode=="2-Public Transit")

share_plot(active_public_cnt_2017, 'final_home_rgcnum', 'commute_mode', 'Regional Growth Center', 'Commute Mode and RGC: 2017')
```
\
\

### No-vehicle households
survey logic: vehicle_count=="0 (no vehicles)"
```{r}
race_com_2017_noveh_2grp <- psurvey_17_simp_2grp %>% 
  filter(vehicle_count=="0 (no vehicles)")

race_com_2017_noveh_cnt_2 <- count_prep_fx(race_com_2017_noveh_2grp,'race_eth_broad', 'commute_mode')

share_plot(race_com_2017_noveh_cnt_2, 'race_eth_broad', 'commute_mode', 'Race', 'Commute Mode and Race: 2017, No household vehicles')
```
\
```{r}
race_com_2017_noveh_3grp <- psurvey_17_simp_3grp %>% 
  filter(vehicle_count=="0 (no vehicles)")

race_com_2017_noveh_cnt_3 <- count_prep_fx(race_com_2017_noveh_3grp,'race_eth_broad', 'commute_mode')

share_plot(race_com_2017_noveh_cnt_3, 'race_eth_broad', 'commute_mode', 'Race', 'Commute Mode and Race: 2017, No household vehicles')
```
\

#### Demographics of no-vehicle households
```{r}
race_com_2017_noveh_cnt <- count_prep_fx(race_com_2017_noveh_2grp,'race_eth_broad', 'hhincome_broad')

race_com_2017_noveh_cnt$hhincome_broad <- factor(race_com_2017_noveh_cnt$hhincome_broad,
                                                 levels=c("Under $25,000",
                                                          "$25,000-$49,999",
                                                          "$50,000-$74,999",
                                                          "$75,000-$99,999",
                                                          "$100,000 or more",
                                                          "Prefer not to answer"))

share_plot(race_com_2017_noveh_cnt, 'race_eth_broad', 'hhincome_broad', 'Race', 'Household income and Race: 2017, No household vehicles')
```
\

```{r}
race_com_2017_noveh_cnt <- count_prep_fx(race_com_2017_noveh_2grp,'race_eth_broad', 'education')

race_com_2017_noveh_cnt$education <- factor(race_com_2017_noveh_cnt$education,
                                                 levels=c("Less than high school",
                                                          "High school graduate",
                                                          "Vocational/technical training",
                                                          "Some college",
                                                          "Associates degree",
                                                          "Bachelor degree",
                                                          "Graduate/post-graduate degree"))

share_plot(race_com_2017_noveh_cnt, 'race_eth_broad', 'education', 'Race', 'Education attainment and Race: 2017, No household vehicles')
```
\
\

# Descriptive Stats: 2017, 2019, 2017-2019, 2021
## Race (2 categories) and transit mode
```{r}
# summarize data ----
all_transit_17 <- count_prep_fx(df_simp_2,'race_eth_broad', 'commute_mode')
all_transit_19 <- count_prep_fx(psurvey_19_simp_2grp,'race_eth_broad', 'commute_mode')
all_transit_1719 <- count_prep_fx(psurvey_1719_simp_2grp,'race_eth_broad', 'commute_mode')
all_transit_21 <- count_prep_fx(psurvey_21_simp_2grp,'race_eth_broad', 'commute_mode')

# merge dfs ----
all_commute <- bind_rows(all_transit_17, 
                         all_transit_19, 
                         all_transit_1719, 
                         all_transit_21) %>% 
  mutate(period = as.factor(survey))

# freq(all_commute$period)
```

```{r}
all_commute_no_other <- all_commute %>% 
  filter(commute_mode!="4-Other modes")

ggplot(all_commute_no_other,
       aes(x=period,
           y=share,
           fill=commute_mode)) +
  geom_bar(stat="identity",
           position="dodge2") +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        axis.title.x = element_blank()) +
  labs(y = "Share",
       fill = "Commute modes") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                size=.5, width=.2,
                position=position_dodge(0.9))+
  facet_grid(cols=vars(race_eth_broad))
```
\
\

## Race (3 categories) and transit mode
```{r}
# summarize data ----
all_transit_17 <- count_prep_fx(df_simp_3,'race_eth_broad', 'commute_mode')
all_transit_19 <- count_prep_fx(psurvey_19_simp_3grp,'race_eth_broad', 'commute_mode')
all_transit_1719 <- count_prep_fx(psurvey_1719_simp_3grp,'race_eth_broad', 'commute_mode')
all_transit_21 <- count_prep_fx(psurvey_21_simp_3grp,'race_eth_broad', 'commute_mode')

# merge dfs ----
all_commute <- bind_rows(all_transit_17,
                         all_transit_19,
                         all_transit_1719,
                         all_transit_21) %>%
  mutate(period = as.factor(survey))

# freq(all_commute$period)
```

```{r}
all_commute_no_other <- all_commute %>% 
  filter(commute_mode!="4-Other modes")

ggplot(all_commute_no_other,
       aes(x=period,
           y=share,
           fill=commute_mode)) +
  geom_bar(stat="identity",
           position="dodge2") +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        axis.title.x = element_blank()) +
  labs(y = "Share",
       fill = "Commute modes") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                size=.5, width=.2,
                position=position_dodge(0.9))+
  facet_grid(cols=vars(race_eth_broad))
```
\
\

# Vehicle Ownership
## Different ways to quantify vehicle ownership
1. **hhsize/vehicle_count** = Household size
2. **numadults/vehicle_count** = Number of adults (18+)
3. **numworkers/vehicle_count** = Number of workers (number of household members with 1+ jobs)
```{r}
race_com_2017_numveh_2 <- psurvey_17_simp_2grp %>% 
  filter(vehicle_count!="0 (no vehicles)") %>% 
  mutate(veh_num=case_when(vehicle_count=="10 or more vehicles" ~10,
                           TRUE~as.numeric(vehicle_count))) %>% 
  mutate(hhsize_num=as.numeric(substr(hhsize, 1, 1))) %>% 
  mutate(hhsz_veh=hhsize_num/veh_num) %>% 
  mutate(hhadults_veh=numadults/veh_num) %>% 
  mutate(numwork_veh=numworkers/veh_num) %>% 
  mutate(hhsz_veh_grp=case_when(hhsz_veh<1~"3-More vehicles than people",
                                hhsz_veh==1~"2-One vehicle per person",
                                hhsz_veh>1~"1-Fewer vehicles than people",
                                TRUE ~ as.character(hhsz_veh))) %>% 
  mutate(hhadults_veh_grp=case_when(hhadults_veh<1~"3-More vehicles than adults",
                                    hhadults_veh==1~"2-One vehicle per adults",
                                    hhadults_veh>1~"1-Fewer vehicles than adults",
                                    TRUE ~ as.character(hhadults_veh))) %>%  
  mutate(numwork_veh_grp=case_when(numwork_veh<1~"3-More vehicles than workers",
                                   numwork_veh==1~"2-One vehicle per worker",
                                   numwork_veh>1~"1-Fewer vehicles than worker",
                                   TRUE ~ as.character(numwork_veh)))

race_com_2017_numveh_3 <- psurvey_17_simp_3grp %>% 
  filter(vehicle_count!="0 (no vehicles)") %>% 
  mutate(veh_num=case_when(vehicle_count=="10 or more vehicles" ~10,
                           TRUE~as.numeric(vehicle_count))) %>% 
  mutate(hhsize_num=as.numeric(substr(hhsize, 1, 1))) %>% 
  mutate(hhsz_veh=hhsize_num/veh_num) %>% 
  mutate(hhadults_veh=numadults/veh_num) %>% 
  mutate(numwork_veh=numworkers/veh_num) %>% 
  mutate(hhsz_veh_grp=case_when(hhsz_veh<1~"3-More vehicles than people",
                                hhsz_veh==1~"2-One vehicle per person",
                                hhsz_veh>1~"1-Fewer vehicles than people",
                                TRUE ~ as.character(hhsz_veh))) %>% 
  mutate(hhadults_veh_grp=case_when(hhadults_veh<1~"3-More vehicles than adults",
                                    hhadults_veh==1~"2-One vehicle per adults",
                                    hhadults_veh>1~"1-Fewer vehicles than adults",
                                    TRUE ~ as.character(hhadults_veh))) %>%  
  mutate(numwork_veh_grp=case_when(numwork_veh<1~"3-More vehicles than workers",
                                   numwork_veh==1~"2-One vehicle per worker",
                                   numwork_veh>1~"1-Fewer vehicles than worker",
                                   TRUE ~ as.character(numwork_veh)))
# freq(race_com_2017_numveh$hhsz_veh)
```
\

Because we're interested in commute modes, we will be focusing on the number of workers within a household
\
```{r}
race_com_2017_noveh_2grp$numworkers_chr <- as.character(race_com_2017_noveh_2grp$numworkers)

race_com_2017_noveh_cnt_2 <- count_prep_fx(race_com_2017_noveh_2grp,'race_eth_broad', 'numworkers_chr')

share_plot(race_com_2017_noveh_cnt_2, 'race_eth_broad', 'numworkers_chr', 'Race', 'Number of workers and Race (2 categories): 2017, No household vehicles')
```

```{r}
race_com_2017_noveh_3grp$numworkers_chr <- as.character(race_com_2017_noveh_3grp$numworkers)

race_com_2017_noveh_cnt_3 <- count_prep_fx(race_com_2017_noveh_3grp,'race_eth_broad', 'numworkers_chr')

share_plot(race_com_2017_noveh_cnt_3, 'race_eth_broad', 'numworkers_chr', 'Race', 'Number of workers and Race (3 categories): 2017, No household vehicles')
```
\
\

The larger the number, the more people per vehicle in each household. Numbers less than 1 represent households with more vehicles than people. People in households without a vehicle are excluded from this analysis.
\
\

#### 3. Number of workers
```{r}
hist(race_com_2017_numveh_2$numwork_veh, 
     main="Number of household workers and vehicle ownership",
     xlab="Workers : Vehicles",
     breaks=23,
     ylim=c(0,3000))

race_com_2017_noveh_cnt <- count_prep_fx(race_com_2017_numveh_2,'race_eth_broad', 'numwork_veh_grp')

share_plot(race_com_2017_noveh_cnt, 'race_eth_broad', 'numwork_veh_grp', 'Race', 'Household vehicles and Race: 2017, Household workers/veh')
```
\
\

## Regional Growth Centers (RGCs)
```{r}

```
\
\

# Analysis
## Transit and Household Auto Ownership
```{r, include=FALSE, eval=FALSE}
# retrieve data
# hh_autos_all <- get_hhts(dyear=c(2017,2019,2021),
#                          level="h",
#                          vars=c("hh_race_category",
#                                 "hhsize",
#                                 "vehicle_count",
#                                 "survey_year"))



# Household-level data
hh_autos <- c(2017, 2019, 2021) %>%
  lapply(FUN=function(x) get_hhts(x, "h", 
                                  c("survey_year",
                                    "vehicle_count",
                                    "hh_race_category",
                                    "hhsize",
                                    "numworkers", #to be able to calculate ratio
                                    "hhincome_broad",
                                    "lifecycle",
                                    "numworkers", #to be able to calculate ratio
                                    # "final_home_rgcnum",
                                    "res_type")))

hh_autos_simp <- lapply(c(2017, 2019, 2021), FUN=function(x) get_hhts(x, "h", 
                                  c("hh_race_category",
                                    "hhsize",
                                    "vehicle_count",
                                    "survey_year",
                                    "numworkers", #to be able to calculate ratio
                                    "hhincome_broad",
                                    "hhincome_detailed",
                                    "lifecycle",
                                    "final_home_rgcnum")) %>% 
                          mutate(vehicle_count_simp=case_when(vehicle_count== "0 (no vehicles)" ~ 0,
                                      vehicle_count == "1" ~ 1,
                                      vehicle_count == "2" ~ 2,
                                      is.na(vehicle_count) ~ NA_real_,
                                      TRUE ~ 3)))

# Person-level data
#education, license 

```

```{r example from Michael 4/15, include=FALSE, eval=FALSE}
pums_1yr <- list()
pums_1yr <- lapply(2015:2019, FUN=function(x) get_psrc_pums(1, x, "p", c("PRACE", "SCHL", "AGEP")) %>%
               mutate(data_year=as.factor(x)) %>%
               mutate(POCYN=case_when(
                   PRACE=="White alone" ~ "Non-POC", 
                   !is.na(PRACE) ~ "POC")) %>%
               mutate(edu_simp=case_when(
                   grepl("(Bach|Mast|Prof|Doct)", SCHL) ~ "Bachelor's degree or higher",
                                           !is.na(SCHL) ~ "Less than a Bachelor's degree")))
poc_edu_pums <- function(dt){
                  temp <- filter(dt, AGEP > 24)
                  counties <- c("King","Kitsap","Pierce","Snohomish")
                  rs <- list()
                  rs <- lapply(counties, FUN=function(x){
                     ctytmp <- filter(temp, COUNTY==x)
                     rs1 <- psrc_pums_count(ctytmp, 
                                            group_var=c("edu_simp","POCYN"), 
                                            incl_counties=FALSE) %>%
                            mutate(County=unique(pull(ctytmp,COUNTY)))}) %>% rbindlist() %>%
                            mutate(data_year=unique(pull(temp, data_year))) %>% 
                            relocate(c("data_year", "County"))
return(rs)
}
poc_edu <- lapply(pums_1yr, poc_edu_pums) %>% rbindlist()
```

#### By Race
```{r, include=FALSE, eval=FALSE}
# stat function
# race_auto <- hhts_count(df=hh_autos_all,
#                         group_vars=c("hh_race_category", 
#                                      "vehicle_count"))

cnt_fx <- function(x){
  hhts_count(x, group_vars=c("hh_race_category","vehicle_count"))
}

race_auto <- lapply(hh_autos, cnt_fx) %>% data.table::rbindlist()

# re-order number of vehicles in numerical order
race_auto$vehicle_count <- factor(race_auto$vehicle_count,
                                  levels=c("0 (no vehicles)", "1", 
                                           "2", "3", "4", "5",
                                           "6", "7", "8", "9", 
                                           "10 or more vehicles"))
# freq(race_auto$vehicle_count)
```

##### Survey Numbers
```{r, include=FALSE, eval=FALSE}
race_auto_total_nototal <- race_auto %>% 
  filter(vehicle_count!="Total")

count_and_share_plot(race_auto_total_nototal,
                     "hh_race_category",
                     "vehicle_count",
                     "Household race category",
                     "Number of Household Automobiles by Race")
```

```{r workspace, include=FALSE, eval=FALSE, echo=FALSE}
# race_auto_plot_count <- ggplot(race_auto_total_nototal, 
#                                aes(x=vehicle_count,
#                                    y=count,
#                                    fill=hh_race_category)) +
#   geom_bar(stat="identity",
#            position="dodge2") +
#   theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
#         axis.title.x = element_blank()) +
#   labs(y = "Number of Households",
#        # x = "Number of vehicles",
#        # title = "Count in Region",
#        fill = "Household race category") +
#   scale_y_continuous(labels = scales::comma) +
#   geom_errorbar(aes(ymin=count-count_moe, ymax=count+count_moe),
#                 size=.5,width=.2,
#                 position=position_dodge(.9))
# 
# # print(race_auto_plot_count)
# 
# race_auto_plot_share <- ggplot(race_auto_total_nototal,
#                                aes(x=vehicle_count,
#                                    y=share,
#                                    fill=hh_race_category)) +
#   geom_bar(stat="identity",
#            position="dodge2") +
#   theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
#         axis.title.x = element_blank()) +
#   labs(y = "Share of Households",
#        # x = "Number of vehicles",
#        # title = "Proportion in Region",
#        fill = "Household race category") +
#   scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
#   geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
#                 size=.5, width=.2,
#                 position=position_dodge(0.9))
# 
# # print(race_auto_plot_nototal)
# 
# 
# autos <- ggarrange(race_auto_plot_count, race_auto_plot_share, ncol = 2, nrow = 1,
#                    common.legend = TRUE,
#                    legend = "bottom")
# annotate_figure(autos, top = text_grob("Number of Household Automobiles by Race", 
#                color = "blue", face = "bold", size = 14))
```
\
\

##### Simplified Vehicle Numbers
Because the survey provides the option for people to respond with the number of household vehicles up to 10, these categories have been simplified to 4 categories: 1, 2, 3+. 
```{r, include=FALSE, eval=FALSE}
# hh_autos_simp <- race_auto %>%
#   filter(vehicle_count!="Missing") %>% 
#   mutate(vehicle_count_simp=case_when(grepl("0", vehicle_count) ~ "0",
#                                       grepl("1", vehicle_count) ~ "1",
#                                       grepl("2", vehicle_count) ~ "2",
#                                       grepl("(3|4|5|6|7|8|9|10)", vehicle_count) ~ "3+",
#                                       TRUE~ vehicle_count))

hh_autos_simp <- race_auto %>%
  # filter(vehicle_count!="Missing") %>% 
  mutate(vehicle_count_simp=case_when(vehicle_count== "0 (no vehicles)" ~ 0,
                                      vehicle_count == "1" ~ 1,
                                      vehicle_count == "2" ~ 2,
                                      is.na(vehicle_count) ~ NA_real_,
                                      TRUE ~ 3))

table1(~vehicle_count |hh_race_category,
       data=race_auto,
       render.categorical=my.render.cat)

```

```{r, include=FALSE, eval=FALSE}
# unable to group by survey_year because of issue with variable, so requires additional steps - after retrieving the data, need to filter out by the year, then can perform the summary statistic, then rbind the tables together
hh_autos_simp_17 <- hh_autos_simp %>% 
  filter(survey_year==as.factor(2017))
hh_autos_simp_19 <- hh_autos_simp %>% 
  filter(survey_year==as.factor(2019))
hh_autos_simp_21 <- hh_autos_simp %>% 
  filter(survey_year==as.factor(2021))

auto_num_17 <- hhts_count(hh_autos_simp_17,
                          group_vars = c("vehicle_count_simp",
                                         "hh_race_category")) %>% 
  mutate(Year=2017)
auto_num_19 <- hhts_count(hh_autos_simp_19,
                          group_vars = c("vehicle_count_simp",
                                         "hh_race_category")) %>% 
  mutate(Year=2019)
auto_num_21 <- hhts_count(hh_autos_simp_21,
                          group_vars = c("vehicle_count_simp",
                                         "hh_race_category")) %>% 
  mutate(Year=2021)


auto_num_17_21 <- rbind(auto_num_17,
                        auto_num_19,
                        auto_num_21)


prop_tbl_by_auto <- auto_num_17_21 %>% 
  filter(vehicle_count_simp != "Total") %>% 
  filter(hh_race_category == "Total")

prop_tbl_by_auto_race <- auto_num_17_21 %>% 
  filter(vehicle_count_simp != "Total") %>% 
  filter(hh_race_category != "Total")


g1 <- ggplot(data=prop_tbl_by_auto, 
             aes(x=vehicle_count_simp,
                 y=share,
                 fill=Year))+
  geom_bar(stat="identity",
             position = "dodge2") +
    labs(x = "Number of Vehicles",
         y = "Percent of Households",
         fill = "Year",
         title = paste0("Auto Ownership by Household: Year")) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                       breaks = seq(0, 1, by = .2))

print(g1)

g2 <- ggplot(data=prop_tbl_by_auto_race, 
             aes(x=vehicle_count_simp,
                 y=share,
                 fill=hh_race_category))+
  geom_bar(stat="identity",
             position = "dodge2") +
    labs(x = "Number of Vehicles",
         y = "Percent of Households",
         fill = "Race/ethnicity",
         title = paste0("Auto Ownership by Household: Race/ethnicity")) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                       breaks = seq(0, 1, by = .2))

print(g2)



ggarrange(g1, g2, ncol = 2, nrow = 1,
          common.legend = TRUE,
          legend = "bottom")
```

<a href="#top">Back to top of the page</a>