---
title: "second-example"
author: "suzanne"
date: "2024-03-27"
output: html_document
---



```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
source('survey-23-preprocess.R')
```



```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad,survey_year, hh_weight
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category,
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, industry 
                from HHSurvey.v_persons_labels")
                
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,
person_id, mode_1, dest_purpose, survey_year, trip_weight 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)
setDT(day)
setDT(trip)
```
#race_category needs to be added to the codebook



## Read in Codebook
```{r pressure, echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/codebook/PSRC_Combined_Codebook_2023_packagable.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```

## Set IDs as characters
TO DO: functionalize convert all ids to characters, or store as characters upfront
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
day[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
day[, person_id := as.character(person_id)]
trip[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
trip[, day_id := as.character(day_id)]

trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
day <- day%>%mutate(survey_year=as.character(survey_year))
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```

Variables that need grouping:
telework_time: use group_2, group_2_value: telework_time_broad
gender: gender_grp group_1
telecommute_freq: telecommute_freq_simple: group_1
dest_purpose: dest_purpose_simple, group_1
mode-> mode_simple


```{r}

variable_list<-rbind(
    variable_list,
    data.table(
      variable = c("mode_simple", 'dest_purpose_simple', 'telework_time_broad', 'gender_grp', 'telecommute_freq_simple', 'hhincome_broad'),
      is_checkbox = c(0,0,0,0,0),
      hh = c(0,0,0,0,0),
      person = c(0,0,0,1, 1),
      day = c(0,0,1,0, 0),
      trip = c(1,1, 0, 0,0),
      vehicle = c(0,0,0,0,0),
      location = c(0,0,0,0,0),
      description = c("mode aggregation", "destination aggregation", "telework time aggregation", "gender group", 'telecommute frequency group'),
      logic = c('mode aggregation', 'destination aggregation',"telework time aggregation", "gender group", 'telecommute frequency group'),
      data_type = c("integer/categorical", "integer/categorical", "integer/categorical", "integer/categorical", "integer/categorical"),
      shared_name = c("mode_simple", 'dest_purpose_simple', 'telework_time_broad', 'gender_grp', 'telecommute_freq_simple')
    )
  )


```


# Add associated values
1. persons: telecommute_freq_simple, gender_grp
2. day: telework_time_broad
2. trips: mode_simple, dest_purpose_simple

# persons day table, gender_grp
1. find the labels for the group
2. add the labeled data to the underlying table
3. add the values code to the values list

```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='gender_grp')
values_labels<-add_values_code(group_name='gender_grp')
person<-grp_to_tbl(tbl=person, ungrouped_name='gender', grouped_name='gender_grp')
```

```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='mode_simple')
values_labels<-add_values_code(group_name='mode_simple', group_labels)
trip<-grp_to_tbl(tbl=trip, ungrouped_name='mode_1', grouped_name='mode_simple')
```


# Summaries from person table:
# workplace by survey year
# 2023
# workplace by gender
# workplace by hhincome


```{r}
workers<-person%>%filter(!workplace %in% c('Missing: Skip logic', "Missing: Skip Logic"))%>%drop_na(workplace)
worker_list<-list(person=workers)

```

```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```


```{r}

workplace <- summarize_weighted(hts_data= worker_list,
                               summarize_var = 'workplace',
                               summarize_by = 'survey_year',
                               id_cols=c('person_id'),
                               wt_cols=c('person_weight'),
                               wtname= 'person_weight'
                               )

```
## Visualize
```{r}

workplace_summary<-workplace$summary$wtd
static_bar_chart(workplace_summary, y='workplace', x='prop', fill='survey_year')

```
# to do: generalize, functionalize, just making it work for now;
```{r}
# make a generalized dataframe to store the outputs
summary_out<-workplace_summary%>%
        mutate(var_1_value=survey_year, var_1_name= 'survey_year',
         var_2_value= workplace, var_2_name= 'workplace')%>%
         select(-c(survey_year, workplace))

```

2023 only by gender
```{r}
workers_23<-workers%>%filter(survey_year =='2023')
hh_23<- hh%>%filter(survey_year== '2023')

worker_list_23<- list(hh=hh_23, person= workers_23)
```

```{r}
workplace_gender <- summarize_weighted(hts_data= worker_list_23,
                               summarize_var = 'workplace',
                               summarize_by = 'gender_grp',
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight','person_weight'),
                               wtname="person_weight"
                               )

```
```{r}
summary_out_gender<-workplace_gender_summary%>%
        mutate(var_1_value=gender_grp, var_1_name= 'gender_grp',
         var_2_value= workplace, var_2_name= 'workplace')%>%
         select(-c(gender_grp, workplace))

summary_out<-rbind(summary_out_gender, summary_out)
```


## Visualize
```{r}

workplace_gender_summary<-workplace_gender$summary$wtd
static_bar_chart(workplace_gender_summary, y='workplace', x='prop', fill='gender_grp')

```
```{r}
workplace_income <- summarize_weighted(hts_data= worker_list_23,
                               summarize_var = 'workplace',
                               summarize_by = 'hhincome_broad',
                               id_cols=c('hh_id', 'person_id'),
                               wt_cols=c('hh_weight','person_weight'),
                               wtname="person_weight"
                               )

```
## Visualize
```{r}

workplace_income_summary<-workplace_income$summary$wtd%>%
  mutate(hhincome_broad=factor(hhincome_broad,
levels=c('Under $25,000','$25,000-$49,999','$50,000-$74,999',
'$75,000-$99,999','$100,000 or more','$100,000-$199,000','$200,000 or more','Prefer not to answer',
'Missing: Skip Logic')))

static_bar_chart(workplace_income_summary, y='workplace', x='prop', fill='hhincome_broad', color='pgnobgy_10')

```

```{r}
summary_out_income<-workplace_income_summary%>%
        mutate(var_1_value=hhincome_broad, var_1_name= 'hhincome_broad',
         var_2_value= workplace, var_2_name= 'workplace')%>%
         select(-c(hhincome_broad, workplace))

summary_out<-rbind(summary_out_gender, summary_out)
```
