---
title: "how-to-use-the-2023-travelSurveyTools-package"
author: "suzanne"
date: "2024-04-24"
output: html_document
---
PSRC and RSG have built a new package called travelSurveyTools to work with travel survey data.
It is recommended to read through this documentation before you start using the package.
https://rsginc.github.io/travelSurveyTools/




## Load in Packages
YOU MIGHT need to download some special packages on github: travelSurveyTools, psrcelmer, and psrcplot.
IF you need to download them, here's how:
library(devtools)
devtools::install_github('rsgInc/travelSurveyTools')
devtools::install_github('psrc/psrcelmer')
devtools::install_github('psrc/psrcplot')

```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
source('../survey-23-preprocess.R')
```

survey-23-preprocess.R is some functions Suzanne wrote to make it easier to work with the package. They are checked in here: https://github.com/psrc/travel-studies/tree/master/2023/summary
Free free to add to the functions or improve them.

The travelSurveyTools package has very specific requirements for the codebook in terms of the names of columns and items included.

To summarize ANY variable and value you  must be in the internally created data table codebook represented. In other words, you don't need the changes written to the excel codebook. You just need the in memory codebook to have the variable and value represented.

The reason is the codebook carries important metadata to be used by the package to know how to handle the variables values.

PSRC has written some functions that make it easier to group data via using the grouping columns in the codebook.
You can feel free to add to the grouping columns in the Excel codebook to make some of the grouping exist more permanently for easier reuse. (But you don't have to, you can also group in code)

Before you get started, go over to: J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx

note the structure of the codebook. You will need to update the in-memory codebook if you want to change any variables or values.

## Read in Codebook
```{r , echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```

variable_list and value_labels will be used by the travelSurveyTools package whenever you summarize.

## Read in Data from Elmer

The 2017, 2019, 2021, and 2023 travel survey data is in Elmer. You will need to select exactly which columns you want to summarize. Note that the variables and values you pull in must match the codebook. (if for some reason in error, there is inconsistency we have to fix it).

The package uses DT and data.table for most operation, so you will usually want inputs to be in data.table format.

```{r }
hh<- get_query(sql= "select  household_id as hh_id, hhincome_broad,survey_year, hh_weight
                from HHSurvey.v_households_labels")


person<- get_query(sql= "select household_id as hh_id,race_category,
person_id, workplace, telecommute_freq, survey_year, person_weight, gender, age, industry,office_available, commute_freq, employment
                from HHSurvey.v_persons_labels")
                
day<- get_query(sql= "select person_id, day_id, household_id as hh_id, telework_time, day_weight , survey_year from HHSurvey.v_days_labels")

trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,
person_id, mode_1, dest_purpose, travelers_total, survey_year, distance_miles,trip_weight 
                from HHSurvey.v_trips_labels")
setDT(hh)
setDT(person)
setDT(day)
setDT(trip)
```



## Set IDs as characters
We should fix this by writing a function. But for now set all ids to characters for easy joining.
```{r}

hh[, hh_id:=as.character(hh_id)]
person[, hh_id:=as.character(hh_id)]
day[, hh_id:=as.character(hh_id)]
trip[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
day[, person_id := as.character(person_id)]
trip[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
trip[, day_id := as.character(day_id)]

trip[, trip_id := as.character(trip_id)]
hh <- hh%>%mutate(survey_year=as.character(survey_year))
person <- person%>%mutate(survey_year=as.character(survey_year))
day <- day%>%mutate(survey_year=as.character(survey_year))
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```


Before you get started on any summary, please review the survey instrument for the question wording and options.

Carefully pay attention to which survey participants were asked which question. Often only a subset of the participants answered a question.

Furthermore, if you are going to summarize any data across time, you will need to RECONCILE any changes.

The survey instrument is here:
J:/Projects/Surveys/HHTravel/Survey2023/Documents/Survey_Instrument/Puget_Sound_HTS_Questionnaire_2023_Final.docx


Open the codebook and review the variable and it's values.
J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx

# Careful with multi year summaries.
If you are doing a multi year summary, carefully look back at the questionnaire to note question changes.
Here are the questionnaires for 2019:file:///J:/Projects/Surveys/HHTravel/Survey2019/Planning/Survey%20Materials/PSRC_2019_Online_Questionnaire.html#employment_and_commute_questions
and 2021:J:/Projects/Surveys/HHTravel/Survey2021/Questionnaire/PSRC_2021_Questionnaire_032521.docx

#1. Summarize Trip Modes


Note that in the values table, we have groupings. 
These groupings can help you handle inconsistencies in the questions over time.

Also we generally have way too much detail in the responses to have statistically significant results, so most of the time you want to summarize grouped data.
You can add the groupings to the variables table by using the function get_grouped_labels
```{r}
variable_list<-add_variable(variable_list, 'mode_simple','trip')
group_labels<-get_grouped_labels(group_id='group_1', group_name='mode_simple')
value_labels<-add_values_code(group_name='mode_simple')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='mode_1', grouped_name='mode_simple')
```


```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```


```{r}
hts_data = list(hh = hh,
                person = person,
                day = day,
                trip = trip)



ids = c('hh_id', 'person_id', 'day_id', 'trip_id')

wts = c('hh_weight', 'person_weight', 'day_weight', 'trip_weight')
```

Now mode_simple is on the variables table, the values table, and the trip table. And we can run a summary.



##summarize data
```{r}

mode_summary = summarize_weighted(hts_data= hts_data,
                               summarize_var = 'mode_simple',
                               summarize_by = 'survey_year',
                               id_cols= ids,
                               wt_cols=wts,
                               wtname='trip_weight'
                               )
```


```{r}

common_modes<-mode_summary$summary$wtd%>%
  mutate(mode_simple= fct_reorder(mode_simple,-prop))%>%drop_na(mode_simple)

static_column_chart(common_modes, x='mode_simple', y='prop', fill='survey_year',ylabel= 'Share', xlabel='Trip Mode')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

```

What if we want to summarize a newly created variable that requires multiple fields, for example sov mode.


We still need to add the variable and values to the codebook.

```{r}
variable_list<-add_variable(variable_list, 'mode_w_sov','trip')
```

#to do: make this a function
```{r}

value<- c(0,1,2,3,4,5)
label=c('SOV', 'HOV', 'Walk', 'Transit', 'Bike', 'Other')
val_labels<-data.frame(value, label)
val_labels<-val_labels%>%mutate(variable='mode_w_sov')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  

```

# add mode_w_sov to the trip table
```{r}
trip<- trip%>%mutate(travelers_total_fix= ifelse(travelers_total!='1 traveler', 'More than 1', '1 traveler'))
trip<-trip%>%mutate(mode_simple= replace_na(mode_simple, 'Drive'))%>%
mutate(mode_w_sov=case_when(
  mode_simple=="Drive"& travelers_total=='1 traveler' ~ 'SOV',
  is.na(travelers_total) ~ 'SOV',
  mode_simple=="Drive"& travelers_total!='1 traveler'~  'HOV',
  .default= mode_simple
))
```


```{r}
hts_data = list(
                trip = trip)
ids = c('trip_id')
wts = c('trip_weight')
```

```{r}

mode_summary = summarize_weighted(hts_data= hts_data,
                               summarize_var = 'mode_w_sov',
                               summarize_by = 'survey_year',
                               id_cols=ids,
                               wt_cols=wts,
                               wtname='trip_weight'
                               )
```


```{r}

common_modes<-mode_summary$summary$wtd%>%
  mutate(mode_w_sov= fct_reorder(mode_w_sov,-prop))%>%
  drop_na(mode_w_sov)

static_column_chart(common_modes, x='mode_w_sov', y='prop', fill='survey_year',ylabel= 'Share', xlabel='Trip Mode with SOV')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

```
