---
title: "first-example-mode-summary"
author: "suzanne"
date: "2024-03-20"
output:
  pdf_document: default
  html_document: default
---

## Load in Packages
IF YOU need to download some packages on github: travelSurveyTools, psrcelmer, and psrcplot.
For this use the syntax:
library(devtools)
devtools::install_github('rsgInc/travelSurveyTools')
devtools::install_github('psrc/psrcelmer')
devtools::install_github('psrc/psrcplot')

```{r setup, include=FALSE}
library(data.table)
library(stringr)
library(travelSurveyTools)
library(psrcelmer)
library(dplyr)
library(psrcplot)
library(tidyverse)
source('../survey-23-preprocess.R')
```



## Read in data items from Elmer

The travelSurveyTools package expects everything to be a data.table, so run setDT all over the place.
```{r }
trip<- get_query(sql= "select trip_id, household_id as hh_id, day_id,
person_id, mode_1, dest_purpose,origin_purpose, survey_year, travelers_total, origin_rgcname, dest_rgcname, trip_weight 
                from HHSurvey.v_trips_labels")

setDT(trip)
```

## Read in Codebook
```{r pressure, echo=FALSE}
cb_path = str_glue("J:/Projects/Surveys/HHTravel/Survey2023/Data/data_published/PSRC_Codebook_2023_v1.xlsx")

variable_list = readxl::read_xlsx(cb_path, sheet = 'variable_list')
value_labels = readxl::read_xlsx(cb_path, sheet = 'value_labels')

setDT(variable_list)
setDT(value_labels)
```

## Set IDs as characters
I guess for joining?, functionalize convert all ids to charcets
```{r}
trip[, trip_id := as.character(trip_id)]
trip[, hh_id := as.character(hh_id)]
trip[, person_id := as.character(person_id)]
trip[, day_id := as.character(day_id)]
```


```{r}
trip <- trip%>%mutate(survey_year=as.character(survey_year))
```



```{r}
center.lyr <- st_read_elmergeo('URBAN_CENTERS')
```

```{r}

variable_list<-rbind(
    variable_list,
    data.table(
      variable = c("mode_simple", "mode_w_sov"),
      is_checkbox = c(0,0),
      hh = c(0,0),
      person = c(0,0),
      day = c(0,0),
      trip = c(1,1),
      vehicle = c(0,0),
      location = c(0,0),
      description = c("mode aggregation","SOV mode aggregation"),
      logic = c("mode aggregation","SOV mode aggregation"),
      data_type = c("integer/categorical","integer/categorical"),
      shared_name = c("mode_simple", "mode_w_sov")
    )
  )

```

trip to/from center by category: rgc_cat_od
work /not work: work, non_work

totals and shares

Filling in rgc_cat_od
```{r}
variable_list<-add_variable(variable_list, 'rgc_cat_od','trip')
```

```{r}
variable_list<-add_variable(variable_list, 'work_or_not','trip')
```

```{r}
group_labels<-get_grouped_labels(group_id='group_1', group_name='mode_simple')
value_labels<-add_values_code(group_name='mode_simple')
trip<-grp_to_tbl(tbl=trip, ungrouped_name='mode_1', grouped_name='mode_simple')
```

# add mode_w_sov to the trip table
```{r}
trip<- trip%>%mutate(travelers_total_fix= ifelse(travelers_total!='1 traveler', 'More than 1', '1 traveler'))
trip<-trip%>%mutate(mode_simple= replace_na(mode_simple, 'Drive'))%>%
mutate(mode_w_sov=case_when(
  mode_simple=="Drive"& travelers_total=='1 traveler' ~ 'SOV',
  is.na(travelers_total) ~ 'SOV',
  mode_simple=="Drive"& travelers_total!='1 traveler'~  'HOV',
  .default= mode_simple
))
```

```{r}

value<- c(0,1)
label=c( 'NotRGC','RGC')
val_labels<-data.frame(value, label)
val_labels<-val_labels%>%mutate(variable='rgc_cat_od')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  
```


```{r}

value<- c(0,1)
label=c('NotWork','Work')
val_labels<-data.frame(value, label)
val_labels<-val_labels%>%mutate(variable='work_or_not')%>%
    mutate(group_1_title = NA, group_1_value = NA,
           group_2_title = NA, group_2_value= NA,
           group_3_title = NA, group_3_value = NA)

all_value_labels<-value_labels%>%select(variable, value, label, group_1_title, group_1_value,
                                      group_2_title, group_2_value, group_3_title, group_3_value)
new_value_labels<-rbind(all_value_labels, val_labels)
new_value_labels[, val_order := seq_len(nrow(new_value_labels))]
value_labels<-new_value_labels  
```

```{r}
trip<-left_join(trip, center.lyr, by=join_by(origin_rgcname==name), suffix= c("", '.origin'))
```

```{r}
trip<-left_join(trip, center.lyr, by=join_by(dest_rgcname==name), suffix=c('', '.destination'))
```

```{r}
trip<-trip%>%mutate(rgc_cat_od= 
                      case_when(is.na(category)& is.na(category.destination) ~ 'Not RGC',
                                category=='Metro' | category.destination=='Metro' ~ 'RGC',
                                category=='Urban' | category.destination=='Urban' ~ 'RGC',
                                TRUE ~'Missing')
)
```

```{r}
work_purposes<-c('Went to work-related place (e.g., meeting, second job, delivery)',
                                                'Went to primary workplace',
                                                'Went to work-related activity (e.g., meeting, delivery, worksite)',
                                                'Went to other work-related activity')
trip<-trip%>%mutate(work_or_not=if_else(dest_purpose %in% work_purposes | origin_purpose %in% work_purposes, 'Work', 'NotWork'))

```



The package expects the data to be in a list of data.tables.
```{r}

hts_data = list(#hh = hh,
                #person = person,
                #day = day,
                trip = trip)
```


```{r}
#some how a duplicate snuck into the variable list not sure how
variable_list<-variable_list%>%distinct(variable, .keep_all=TRUE)
```

##summarize data
```{r}
trip_summary = summarize_weighted(hts_data= hts_data,
                               summarize_var = 'mode_w_sov',
                               summarize_by = c('survey_year', 'rgc_cat_od'),
                               id_cols='trip_id',
                               wt_cols='trip_weight',
                               wtname='trip_weight'
                               )

```


```{r}
trips_summary_wtd<-trip_summary$summary$wtd%>%filter(survey_year %in% c('2019', '2023'))%>%
  mutate(mode_w_sov= fct_reorder(mode_w_sov,-prop))%>%filter(mode_w_sov %in% c('Transit'))
static_column_chart(trips_summary_wtd, x='rgc_cat_od', y='prop', fill='survey_year', color='pognbgy_5')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
```


how many trips overall to centers vs not
```{r}
trip_summary = summarize_weighted(hts_data= hts_data,
                               summarize_var = 'work_or_not',
                               summarize_by = c('survey_year', 'rgc_cat_od'),
                               id_cols='trip_id',
                               wt_cols='trip_weight',
                               wtname='trip_weight'
                               )

```

## Visualize
```{r}

trips_summary_wtd<-trip_summary$summary$wtd%>%mutate(prop_moe=prop_se*1.645)

write.csv(trips_summary_wtd, 'trips_centers_trend.csv')


```
```{r}
trips_summary_wtd<-trips_summary_wtd%>%filter(work_or_not=='Work')%>%filter(survey_year!='2017')%>%mutate(est=round(est,-4))
static_column_chart(trips_summary_wtd, x= 'rgc_cat_od', y='est', fill='survey_year')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))
```

```{r}

mode_summary = summarize_weighted(hts_data= hts_data,
                               summarize_var = 'mode_w_sov',
                               summarize_by = c('survey_year', 'work_or_not', 'rgc_cat_od'),
                               id_cols='trip_id',
                               wt_cols='trip_weight',
                               wtname='trip_weight'
                               )



```


## Visualize
```{r}

common_modes<-mode_summary$summary$wtd%>%
  mutate(mode_w_sov= fct_reorder(mode_w_sov,-prop))%>%mutate(prop_moe=prop_se*1.645)

write.csv(common_modes, 'mode_centers_trend.csv')

work_modes<-common_modes%>%filter(work_or_not=='Work')%>%filter(survey_year!=2017)
not_work_modes<-common_modes%>%filter(work_or_not=='NotWork')%>%filter(survey_year!=2017)

static_facet_column_chart(work_modes, x='mode_w_sov', y='prop', fill='survey_year',ylabel= 'Share', xlabel='Trip Mode', facet='rgc_cat_od',scale='fixed')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

static_facet_column_chart(not_work_modes, x='mode_w_sov', y='prop', fill='survey_year',ylabel= 'Share', xlabel='Trip Mode', facet='rgc_cat_od', scale='fixed')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

```

```{r}


static_facet_column_chart(work_modes, x='mode_w_sov', y='est', fill='survey_year',ylabel= 'Share', xlabel='Trip Mode', facet='rgc_cat_od',scale='fixed')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

static_facet_column_chart(not_work_modes, x='mode_w_sov', y='est', fill='survey_year',ylabel= 'Share', xlabel='Trip Mode', facet='rgc_cat_od', scale='fixed')+ theme(axis.text.x=element_text(size=14), axis.text.y=element_text(size=14),legend.text = element_text(size=14), axis.title.y=element_text(size=20), axis.title.x=element_text(size=20))

```
