---
title: "Centers monitoring: 2023 data update"
date: today
format: 
  html:
    mainfont: Poppins
    theme: [cosmo, custom.scss]
    toc: true
    df-print: kable
    fig-asp: .7
    fig-width: 4
    warning: false
    echo: false
---

```{r}
source("get_rgc_update_data.R")
install_psrc_fonts()
```

## Travel Behaviors by home location

### mode share

```{r}
#| fig-asp: 0.6
#| fig-width: 9


df_plot <- psrc_hts_stat(df_hts_data_17_19, "trip", c("home_in_UGA_mu", "mode_class_simple"), incl_na = FALSE) %>%
  add_row(psrc_hts_stat(df_hts_data, "trip", c("home_in_UGA_mu", "mode_class_simple"), incl_na = FALSE))

dodge <- position_dodge(width=0.9)
ggplot(df_plot, aes(x=mode_class_simple, y=prop, fill=survey_year)) +
  geom_bar(position="dodge", stat="identity") +
  scale_y_continuous(labels = scales::label_percent()) +
  geom_errorbar(aes(ymin = prop-prop_moe, ymax = prop+prop_moe, group=survey_year), position = dodge, width = 0.15) +
  facet_wrap(~home_in_UGA_mu) +
  psrc_style() +
  scale_fill_manual(values = psrc_colors$purples_inc[2:6])
```

```{r}
df_plot %>%
  mutate(prop = scales::percent(prop, 0.01)) %>%
  pivot_wider(id_cols = c("home_in_UGA_mu", "mode_class_simple"), names_from="survey_year", values_from = "prop")
```


:::{.panel-tabset}

```{r}
plot_mode_share <- function(df, survey_year){
  ggplot(df, aes(x=fct_rev(home_in_UGA_mu), y=prop, fill=mode_class_simple)) +
    geom_bar(position=position_stack(reverse = TRUE), stat="identity") +
    scale_y_continuous(labels = scales::label_percent()) +
    psrc_style() +
    scale_fill_manual(values = psrc_colors$pognbgy_5) +
    coord_flip()+
    theme(
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_line(color="#cbcbcb")
    ) +
    labs(x="",y="")
}
```

#### 2017

```{r}
df_plot2 <- df_plot %>% filter(survey_year=="2017")
plot_mode_share(df_plot2,"2017")
```

#### 2017/2019

```{r}
df_plot2 <- df_plot %>% filter(survey_year=="2017/2019")
plot_mode_share(df_plot2,"2017/2019")
```

#### 2019

```{r}
df_plot2 <- df_plot %>% filter(survey_year=="2019")
plot_mode_share(df_plot2,"2019")
```

#### 2021

```{r}
df_plot2 <- df_plot %>% filter(survey_year=="2021")
plot_mode_share(df_plot2,2021)
```

#### 2023
```{r}
df_plot2 <- df_plot %>% filter(survey_year=="2023")
plot_mode_share(df_plot2,2023)
```

:::

### change in mode share

```{r}
df_plot <- psrc_hts_stat(df_hts_data_17_19, "trip", c("home_in_UGA_center", "mode_class_simple"), incl_na = FALSE) %>%
  add_row(psrc_hts_stat(df_hts_data, "trip", c("home_in_UGA_center", "mode_class_simple"), incl_na = FALSE))

mode_share_change <- df_plot %>%
  filter(survey_year%in% c("2021","2023")) %>%
  left_join(df_plot %>% filter(survey_year=="2017/2019") %>%
              rename(prop_2017_2019 = prop) %>%
              select(home_in_UGA_center,mode_class_simple,prop_2017_2019), by=c("home_in_UGA_center","mode_class_simple")) %>%
  mutate(change = (prop-prop_2017_2019)/prop_2017_2019)

ggplot(mode_share_change, aes(x=fct_rev(home_in_UGA_center), y=change, fill=mode_class_simple)) +
    geom_col(position = "dodge")+
  facet_wrap(~survey_year) +
    scale_y_continuous(labels=scales::percent) +
    scale_fill_discrete_psrc ("pognbgy_10")+
    psrc_style() +
    coord_flip()+
    theme(
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_line(color="#cbcbcb")
    )
```

```{r}
mode_share_change %>%
  mutate(change = scales::percent(change, 0.01)) %>%
  pivot_wider(id_cols = c("home_in_UGA_center", "mode_class_simple"), names_from="survey_year", values_from = "change")
```


### transit frequency

- only includes adults 18+

```{r}
#| fig-asp: 0.6
#| fig-width: 9

df_plot <- psrc_hts_stat(df_hts_data, "person", c("home_in_UGA_mu", "transit_freq_simple"), incl_na = FALSE) 
df_plot2 <- psrc_hts_stat(df_hts_data_17_19, "person", c("home_in_UGA_mu", "transit_freq_simple"), incl_na = FALSE) %>% add_row(df_plot)

ggplot(df_plot2, aes(x=transit_freq_simple, y=prop, fill=survey_year)) +
  geom_col(position = "dodge")+  
  geom_errorbar(aes(ymin=prop-prop_moe, 
                    ymax=prop+prop_moe),
                width=0.2, position = position_dodge(0.9))+
  facet_wrap(~home_in_UGA_mu) +
  scale_y_continuous(labels=scales::percent) +
  scale_fill_manual(values = psrc_colors$purples_inc[1:5]) +
  psrc_style() + 
  theme(plot.title = element_blank())
```

```{r}
df_plot2 %>%
  mutate(prop = scales::percent(prop, 0.01)) %>%
  pivot_wider(id_cols = c("home_in_UGA_mu", "transit_freq_simple"), names_from="survey_year", values_from = "prop")
```

### walk frequency

- only includes adults 18+

```{r}
#| fig-asp: 0.6
#| fig-width: 9

df_plot <- psrc_hts_stat(df_hts_data, "person", c("home_in_UGA_mu", "walk_freq_simple"), incl_na = FALSE)
df_plot2 <- psrc_hts_stat(df_hts_data_17_19, "person", c("home_in_UGA_mu", "walk_freq_simple"), incl_na = FALSE) %>% add_row(df_plot)

ggplot(df_plot2, aes(x=walk_freq_simple, y=prop, fill=survey_year)) +
  geom_col(position = "dodge")+
  geom_errorbar(aes(ymin=prop-prop_moe,
                    ymax=prop+prop_moe),
                width=0.2, position = position_dodge(0.9))+
  facet_wrap(~home_in_UGA_mu) +
  scale_y_continuous(labels=scales::percent) +
  scale_fill_manual(values = psrc_colors$purples_inc[1:5]) +
  psrc_style() +
  theme(plot.title = element_blank())
```

```{r}
df_plot2 %>%
  mutate(prop = scales::percent(prop, 0.01)) %>%
  pivot_wider(id_cols = c("home_in_UGA_mu", "walk_freq_simple"), names_from="survey_year", values_from = "prop")
```

### VMT

:::{.panel-tabset}

#### Own calculation: Suzanne's code

- assumptions:
  1. 5+ travelers is replaced with 7
  2. removed trips with distance longer than 200 miles

```{r}

# vmt by centers
total_vmt_mu <- df_hts_data$trip %>%
  add_row(df_hts_data_17_19$trip) %>%
  left_join(df_hts_data$hh %>% select(hh_id,home_in_UGA_mu), by="hh_id") %>%
  group_by(survey_year, home_in_UGA_mu) %>%
  summarize(total_vmt = sum(weighted_vmt), 
            total_vmt_7 = sum(weighted_vmt_7)) %>%
  ungroup()

people_mu <- df_hts_data$person %>%
  add_row(df_hts_data_17_19$person) %>%
  left_join(df_hts_data$hh %>% select(hh_id,home_in_UGA_mu), by="hh_id") %>%
  group_by(survey_year, home_in_UGA_mu) %>%
  summarise(people = sum(person_weight))

df_vmt_mu <- total_vmt_mu %>%
  left_join(people_mu, by=c("survey_year", "home_in_UGA_mu"))  %>%
  mutate(vmt_per_person = total_vmt/people,
         vmt_per_person_7 = total_vmt_7/people)

# region vmt
total_vmt <- df_hts_data$trip %>%
  add_row(df_hts_data_17_19$trip) %>%
  group_by(survey_year) %>%
  summarize(total_vmt = sum(weighted_vmt), 
            total_vmt_7 = sum(weighted_vmt_7)) %>%
  ungroup()

people <- df_hts_data$person %>%
  add_row(df_hts_data_17_19$person) %>%
  group_by(survey_year) %>%
  summarise(people = sum(person_weight))

df_vmt <- df_vmt_mu %>%
  add_row(total_vmt %>%
  left_join(people, by="survey_year") %>%
  mutate(vmt_per_person = total_vmt/people,
         vmt_per_person_7 = total_vmt_7/people) %>%
  mutate(home_in_UGA_mu = "Region", .after="survey_year")) %>%
  mutate(vmt_per_person = ifelse(survey_year=="2017/2019",vmt_per_person/2,vmt_per_person),
         vmt_per_person_7 = ifelse(survey_year=="2017/2019",vmt_per_person_7/2,vmt_per_person_7))


df_vmt %>%
  mutate(vmt_per_person_7 = scales::number(vmt_per_person_7, 0.01)) %>%
  pivot_wider(id_cols = "home_in_UGA_mu", names_from="survey_year", values_from = "vmt_per_person_7")
```

```{r}

# vmt by centers
total_vmt_mu <- df_hts_data$trip %>%
  add_row(df_hts_data_17_19$trip) %>%
  left_join(df_hts_data$hh %>% select(hh_id,home_regional_geog), by="hh_id") %>%
  group_by(survey_year, home_regional_geog) %>%
  summarize(total_vmt = sum(weighted_vmt), 
            total_vmt_7 = sum(weighted_vmt_7)) %>%
  ungroup()

people_mu <- df_hts_data$person %>%
  add_row(df_hts_data_17_19$person) %>%
  left_join(df_hts_data$hh %>% select(hh_id,home_regional_geog), by="hh_id") %>%
  group_by(survey_year, home_regional_geog) %>%
  summarise(people = sum(person_weight))

df_vmt_mu <- total_vmt_mu %>%
  left_join(people_mu, by=c("survey_year", "home_regional_geog"))  %>%
  mutate(vmt_per_person = total_vmt/people,
         vmt_per_person_7 = total_vmt_7/people)

# region vmt
total_vmt <- df_hts_data$trip %>%
  add_row(df_hts_data_17_19$trip) %>%
  group_by(survey_year) %>%
  summarize(total_vmt = sum(weighted_vmt), 
            total_vmt_7 = sum(weighted_vmt_7)) %>%
  ungroup()

people <- df_hts_data$person %>%
  add_row(df_hts_data_17_19$person) %>%
  group_by(survey_year) %>%
  summarise(people = sum(person_weight))

df_vmt <- df_vmt_mu %>%
  add_row(total_vmt %>%
  left_join(people, by="survey_year") %>%
  mutate(vmt_per_person = total_vmt/people,
         vmt_per_person_7 = total_vmt_7/people) %>%
  mutate(home_regional_geog = "Region", .after="survey_year")) %>%
  mutate(vmt_per_person = ifelse(survey_year=="2017/2019",vmt_per_person/2,vmt_per_person),
         vmt_per_person_7 = ifelse(survey_year=="2017/2019",vmt_per_person_7/2,vmt_per_person_7)) %>%
  mutate(home_regional_geog = factor(home_regional_geog, levels=c("Metro","Core","HCT","CitiesTowns","UU","Rural","Region")))


df_vmt %>%
  arrange(home_regional_geog) %>%
  mutate(vmt_per_person_7 = scales::number(vmt_per_person_7, 0.01)) %>%
  pivot_wider(id_cols = "home_regional_geog", names_from="survey_year", values_from = "vmt_per_person_7")
```

- sample count
```{r}
df_hts_data$hh %>%
  add_row(df_hts_data_17_19$hh) %>%
  mutate(home_regional_geog = factor(home_regional_geog, levels=c("Metro","Core","HCT","CitiesTowns","UU","Rural","Region"))) %>%
  group_by(survey_year,home_regional_geog) %>%
  summarise(count=n()) %>%
  arrange(home_regional_geog) %>%
  pivot_wider(id_cols = "home_regional_geog", names_from="survey_year", values_from = "count")
```


#### package

```{r}
# TODO: don't know why 2017/2019 not showing: look into if needed
df_plot <- psrc_hts_vmtrate(df_hts_data, "home_in_UGA_mu") %>%
  add_row(psrc_hts_vmtrate(df_hts_data) %>%
            mutate(home_in_UGA_mu = "Region", .after="survey_year"))
df_plot_region <- psrc_hts_vmtrate(df_hts_data) %>%
  mutate(home_in_UGA_mu = "Region", .after="survey_year")


df_plot %>%
  mutate(mean = scales::number(mean, 0.01)) %>%
  pivot_wider(id_cols = "survey_year", names_from="home_in_UGA_mu", values_from = "mean")
```

:::


### average trip distance & travel time

:::{.panel-tabset}

#### trip distance
```{r}
#| layout-ncol: 2

rs1 <- psrc_hts_stat(df_hts_data, 
                     analysis_unit="trip", 
                     group_vars="home_category", 
                     stat_var="distance_miles")
rs3 <- psrc_hts_stat(df_hts_data_17_19, 
                     analysis_unit="trip", 
                     group_vars="home_category", 
                     stat_var="distance_miles") %>% add_row(rs1)

ggplot(rs3, aes(x=home_category, y=mean, fill=survey_year)) +
  geom_col(position = "dodge")+  
  geom_errorbar(aes(ymin=mean-mean_moe, 
                    ymax=mean+mean_moe),
                width=0.2, position = position_dodge(0.9))+
  scale_fill_manual(values = psrc_colors$purples_inc[1:5]) +
  # facet_wrap(~home_category, scales = "free_y") +
  psrc_style() + 
  theme(plot.title = element_blank())

rs1 <- psrc_hts_stat(df_hts_data, 
                     analysis_unit="trip", 
                     group_vars="home_in_UGA_mu", 
                     stat_var="distance_miles")
rs3 <- psrc_hts_stat(df_hts_data_17_19, 
                     analysis_unit="trip", 
                     group_vars="home_in_UGA_mu", 
                     stat_var="distance_miles") %>% add_row(rs1)

ggplot(rs3, aes(x=home_in_UGA_mu, y=mean, fill=survey_year)) +
  geom_col(position = "dodge")+  
  geom_errorbar(aes(ymin=mean-mean_moe, 
                    ymax=mean+mean_moe),
                width=0.2, position = position_dodge(0.9))+
  scale_fill_manual(values = psrc_colors$purples_inc[1:5]) +
  # facet_wrap(~home_category, scales = "free_y") +
  psrc_style() + 
  theme(plot.title = element_blank())
```

#### travel time

```{r}
#| layout-ncol: 2

rs2 <- psrc_hts_stat(df_hts_data, 
                     analysis_unit="trip", 
                     group_vars="home_category", 
                     stat_var="duration_minutes")
rs3 <- psrc_hts_stat(df_hts_data_17_19, 
                     analysis_unit="trip", 
                     group_vars="home_category", 
                     stat_var="duration_minutes") %>% add_row(rs2)

ggplot(rs3, aes(x=home_category, y=mean, fill=survey_year)) +
  geom_col(position = "dodge")+  
  geom_errorbar(aes(ymin=mean-mean_moe, 
                    ymax=mean+mean_moe),
                width=0.2, position = position_dodge(0.9))+
  scale_fill_manual(values = psrc_colors$purples_inc[1:5]) +
  psrc_style() + 
  theme(plot.title = element_blank())

rs2 <- psrc_hts_stat(df_hts_data, 
                     analysis_unit="trip", 
                     group_vars="home_in_UGA_mu", 
                     stat_var="duration_minutes")
rs3 <- psrc_hts_stat(df_hts_data_17_19, 
                     analysis_unit="trip", 
                     group_vars="home_in_UGA_mu", 
                     stat_var="duration_minutes") %>% add_row(rs2)

ggplot(rs3, aes(x=home_in_UGA_mu, y=mean, fill=survey_year)) +
  geom_col(position = "dodge")+  
  geom_errorbar(aes(ymin=mean-mean_moe, 
                    ymax=mean+mean_moe),
                width=0.2, position = position_dodge(0.9))+
  scale_fill_manual(values = psrc_colors$purples_inc[1:5]) +
  psrc_style() + 
  theme(plot.title = element_blank())
```

:::

## Travel Behavior by destination location

### mode share

:::{.panel-tabset}

#### mode share with walk/bike/micromobility

```{r}
#| fig-asp: 0.6
#| fig-width: 9

df_plot <- psrc_hts_stat(df_hts_data_17_19, "trip", c("dest_in_UGA_mu", "mode_class_simple"), incl_na = FALSE) %>%
  add_row(psrc_hts_stat(df_hts_data, "trip", c("dest_in_UGA_mu", "mode_class_simple"), incl_na = FALSE))

dodge <- position_dodge(width=0.9)
ggplot(df_plot, aes(x=mode_class_simple, y=prop, fill=dest_in_UGA_mu)) +
  geom_bar(position="dodge", stat="identity") +
  geom_text(aes(label=scales::percent(prop,accuracy = 1), y=prop, group=dest_in_UGA_mu),vjust=-1.2, position= dodge)+
  geom_errorbar(aes(ymin = prop-prop_moe, ymax = prop+prop_moe, group=dest_in_UGA_mu), position = dodge, width = 0.15) +
  scale_y_continuous(labels = scales::label_percent()) +
  facet_wrap(~survey_year) +
  psrc_style() +
  scale_fill_manual(values = psrc_colors$pognbgy_5)
```

```{r}
#| fig-asp: 0.6
#| fig-width: 9
ggplot(df_plot, aes(x=mode_class_simple, y=prop, fill=survey_year)) +
  geom_bar(position="dodge", stat="identity") +
  geom_text(aes(label=scales::percent(prop,accuracy = 1), y=prop, group=survey_year),vjust=-0.6, position= dodge)+
  scale_y_continuous(labels = scales::label_percent()) +
  geom_errorbar(aes(ymin = prop-prop_moe, ymax = prop+prop_moe, group=survey_year), position = dodge, width = 0.15) +
  facet_wrap(~dest_in_UGA_mu) +
  psrc_style() +
  scale_fill_manual(values = psrc_colors$purples_inc)
```

```{r}
df_plot %>%
  mutate(prop = scales::percent(prop, 0.01)) %>%
  pivot_wider(id_cols = c("dest_in_UGA_mu", "mode_class_simple"), names_from="survey_year", values_from = "prop")
```


#### mode_class_5

```{r}
#| fig-asp: 0.6
#| fig-width: 9

df_plot <- psrc_hts_stat(df_hts_data_17_19, "trip", c("dest_in_UGA_mu", "mode_class_5"), incl_na = FALSE) %>%
  add_row(psrc_hts_stat(df_hts_data, "trip", c("dest_in_UGA_mu", "mode_class_5"), incl_na = FALSE))

dodge <- position_dodge(width=0.9)
ggplot(df_plot, aes(x=mode_class_5, y=prop, fill=dest_in_UGA_mu)) +
  geom_bar(position="dodge", stat="identity") +
  geom_text(aes(label=scales::percent(prop,accuracy = 1), y=prop, group=dest_in_UGA_mu),vjust=-1.2, position= dodge)+
  geom_errorbar(aes(ymin = prop-prop_moe, ymax = prop+prop_moe, group=dest_in_UGA_mu), position = dodge, width = 0.15) +
  scale_y_continuous(labels = scales::label_percent()) +
  facet_wrap(~survey_year) +
  psrc_style() +
  scale_fill_manual(values = psrc_colors$pognbgy_5)
```

```{r}
#| fig-asp: 0.6
#| fig-width: 9
ggplot(df_plot, aes(x=mode_class_5, y=prop, fill=survey_year)) +
  geom_bar(position="dodge", stat="identity") +
  geom_text(aes(label=scales::percent(prop,accuracy = 1), y=prop, group=survey_year),vjust=-0.6, position= dodge)+
  scale_y_continuous(labels = scales::label_percent()) +
  geom_errorbar(aes(ymin = prop-prop_moe, ymax = prop+prop_moe, group=survey_year), position = dodge, width = 0.15) +
  facet_wrap(~dest_in_UGA_mu) +
  psrc_style() +
  scale_fill_manual(values = psrc_colors$purples_inc)
```

```{r}
df_plot %>%
  mutate(prop = scales::percent(prop, 0.01)) %>%
  pivot_wider(id_cols = c("dest_in_UGA_mu", "mode_class_5"), names_from="survey_year", values_from = "prop")
```

:::

### number of trips per acre

- work trip includes `"Work","Work-related","School","School-related"`

:::{.panel-tabset}

#### Metro/Urban centers

```{r}
df_plot <- psrc_hts_stat(df_hts_data_17_19, "trip", c("dest_in_UGA_mu","dest_purpose_work"), incl_na = FALSE) %>%
  add_row(psrc_hts_stat(df_hts_data, "trip", c("dest_in_UGA_mu","dest_purpose_work"), incl_na = FALSE)) %>%
  left_join(df_metro_urban_area, by=c("dest_in_UGA_mu"="category")) %>%
  mutate(`trip/acre` = est/acres,
         dest_in_UGA_mu = factor(dest_in_UGA_mu,levels=c("Metro","Urban","Rest of UGA","not UGA")))

dodge <- position_dodge(width=0.9)
ggplot(df_plot, aes(x=dest_purpose_work, y=`trip/acre`, fill=dest_in_UGA_mu)) +
  geom_bar(position="dodge", stat="identity") +
  geom_text(aes(label=scales::number(`trip/acre`,accuracy = 0.1), y=`trip/acre`, group=dest_in_UGA_mu),vjust=-0.8, position= dodge)+
  facet_wrap(~survey_year) +
  psrc_style() +
  scale_fill_manual(values = psrc_colors$pognbgy_5)
```

```{r}
df_plot %>%
  mutate(prop = scales::percent(prop, 0.01)) %>%
  pivot_wider(id_cols = c("dest_in_UGA_mu", "dest_purpose_work"), names_from="survey_year", values_from = `trip/acre`)
```

#### RGC

```{r}
df_plot <- psrc_hts_stat(df_hts_data_17_19, "trip", c("dest_in_UGA_center","dest_purpose_work"), incl_na = FALSE) %>%
  add_row(psrc_hts_stat(df_hts_data, "trip", c("dest_in_UGA_center","dest_purpose_work"), incl_na = FALSE)) %>%
  left_join(df_center_area, by=c("dest_in_UGA_center"="rgcname")) %>%
  mutate(`trip/acre` = est/acres,
         dest_in_UGA_center = factor(dest_in_UGA_center,levels=c("RGC","Rest of UGA","not UGA")))

dodge <- position_dodge(width=0.9)
ggplot(df_plot %>% filter(dest_in_UGA_center!="not UGA"), aes(x=dest_purpose_work, y=`trip/acre`, fill=dest_in_UGA_center)) +
  geom_bar(position="dodge", stat="identity") +
  geom_text(aes(label=scales::number(`trip/acre`,accuracy = 1), y=`trip/acre`, group=dest_in_UGA_center),vjust=-0.8, position= dodge)+
  facet_wrap(~survey_year) +
  psrc_style() +
  scale_fill_manual(values = psrc_colors$pognbgy_5[c(1,2,5)])
```

```{r}
df_plot %>%
  mutate(prop = scales::percent(prop, 0.01)) %>%
  pivot_wider(id_cols = c("dest_in_UGA_center", "dest_purpose_work"), names_from="survey_year", values_from = `trip/acre`)
```


:::


